### 2012-09-24 problem_formulation.tex
starting paper

\documentclass[]{aiaa-tc} % insert '[draft]' option to show overfull boxes

 \title{An application of Graph Theory to MDAO problem formulation}

\author{
  David Pate, %
     \thanks{Georgia Tech}
  Justin Gray,%
     \thanks{Aerospace Engineer, MDAO Branch, Mail Stop 5-11, AIAA Member}
  Dr. Brian German 
     \thanks{Georgia Tech...}
 }

%\usepackage{setspace}
%\doublespace

\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{caption} 
\usepackage{amsmath}
\usepackage{lscape}
\usepackage{hyperref}
\usepackage{appendix}
\usepackage[section]{placeins}

\captionsetup[figure]{margin=5pt,font=small,labelfont=bf,textfont=bf,justification=justified,}
%\captionsetup[wrapfigure]{margin=5pt,font=small,labelfont=bf,justification=justified,singlelinecheck=off}
\captionsetup[table]{margin=5pt,font=small,labelfont=bf,textfont=bf,justification=justified,position=top}

\bibliographystyle{aiaa}

\usepackage{lettrine}
\usepackage{verbatim}

%\usepackage{hyperref} %allows for the creation of actual text links
\begin{document}

\maketitle

\begin{abstract}
   blah blah blah
\end{abstract}

\section*{Nomenclature}

\begin{tabular}{l l} 
    DOE      & Design of Experiments \\ 
    IDF      & Individual Design Feasible  \\
    MDF      & Multiple Design Feasible \\
    MDAO     & Multidisciplinary Design Analysis and Optimization \\
\end{tabular}

\section{Introduction}

    - Large Scale engineered systems 
    - Complex problems that are not obvious to set up
    - need for a formal and consistent way to represent the problem
      - contains *all* the information needed to formulate a solution method
    - Previous attempts to provide problem formulation

      \begin{itemize}
        \item Structural Matrix, 
        \item DSM
        \item Functional Dependence Table
        \item XDSM
        \item Flow Chart
        \item Adjacency Matrix
      \end{itemize}

   As the size and complexity of engineering systems grows the time and expense for setting up 
   analysis models grows with them. Multidisciplinary Design Analysis and Optimziation
   frameworks such as OpenMDAO and ModelCenter have enabled a new level of analysis tool integration 
   and paved the way for models of increasing integration and complexity.

\section{Graph Based Problem Formulation}

\section{Example Problem}

\section{Applications}

\section{Conclusions}

\end{document}

### 2012-09-24 problem_formulation.tex
adding in david's background notes from previous weeks

@@
      \thanks{Georgia Tech}
+  Dr. Brian German 
+     \thanks{Georgia Tech...}
   Justin Gray,%
      \thanks{Aerospace Engineer, MDAO Branch, Mail Stop 5-11, AIAA Member}   
-  Dr. Brian German 
-     \thanks{Georgia Tech...}
  }
@@
 \begin{tabular}{l l} 
-    DOE      & Design of Experiments \\ 
-    IDF      & Individual Design Feasible  \\
-    MDF      & Multiple Design Feasible \\
+    AAO      & All-At- \\
     MDAO     & Multidisciplinary Design Analysis and Optimization \\
+    FPF      & Fundamental Problem Formulation \\
 \end{tabular}
@@
     As the size and complexity of engineering systems grows the time and expense for setting up 
-   analysis models grows with them. Multidisciplinary Design Analysis and Optimziation
-   frameworks such as OpenMDAO and ModelCenter have enabled a new level of analysis tool integration 
-   and paved the way for models of increasing integration and complexity.
+    analysis models grows with them. Multidisciplinary Design Analysis and Optimization (MDAO)
+    frameworks such as OpenMDAO\cite{Gray2012} and ModelCenter have enabled a new level of analysis tool integration 
+    and paved the way for models of with more analysis tools and increasing numbers of multidisciplinary couplings. 
+    Such large models present a distinct challenge to properly set up in a manner that allows for an efficient solution. 
+    In fact as the size of a problem grows very large just determining what the proper solution strategy can be a daunting 
+    task.
+
+    What is needed, in order to address the complexity problem, is a method of specifying an engineering problem 
+    in a complete and totally general manner. A complete specification needs to include the following information: 
+    \begin{itemize}
+       \item Discipline Analyses 
+       \item Discipline State Variables and Residuals
+       \item Local \& Global Design Variables
+       \item Local \& Global Constraints
+       \item Objective or Objectives
+       \item Coupling Constraints
+       \item Local and Global Parameters
+    \end{itemize}
+    A general specification is one that states nothing specific about a problem solution path. It does not include any details about 
+    how an objective is to be minimized, how coupling constraints will be satisfied, or how residuals will managed. We 
+    define a complete and general problem specification to be the Fundamental Problem Formulation (FPF). By definition, the FPF for 
+    any given problem will be constant regardless of which MDAO framework or optimization architecture is used to solve the problem.
+
+    In this work we proposed a graph based syntax for the specification of the FPF. This graph based syntax provides several key
+    features that make it useful for working with large scale MDAO problems. It provides a rigid structure that can be easily manipulated 
+    in with a wide range of well establish graph-theory algorithms. By taking advantage of that structure it is possible to test 
+    any given problem specification to determine if it is in fact complete and general. It can be used to describe more specific problem 
+    setups, and given one can be used to derive the FPF from it. Lastly, a graph based specification for problem formulation lends itself
+    well to interacting with MDAO frameworks which makes it particularly powerful for integrating problem formulation analysis techniques 
+    into those frameworks.
+
+\section{Problem Formulation Syntax}
+    The most common problem formulation for an MDAO problem could be given as follows:
+
+    \begin{align}
+        min &\ \ f\left(X\right) \notag
+        \\ w.r.t. &\ \  X \notag
+        \\ s.t. &\ \ G\left(X\right) \leq 0
+        \\      &\ \ C\left(X\right) \leq 0
+    \end{align}
+
+    Where $f(X)$ is the objective function, $X$ is the vector of design variables, $G(X)$ are the optimization constraints, 
+    and $C(X)$ are the coupling constraints. This does represent a general problem formulation, since it says nothing about 
+    how the problem should be solved. Tedford and Martins used this format to specify a set of test problems and derived specific formulations 
+    for solving them with a number of optimization architectures\cite{Tedford2009}. Lamb and Martins also used this basic mathematical 
+    syntax to describe the All-At-Once (AAO) architecture \cite{Lambe2011}. Although the AAO specification does specify a solution path 
+    including a single optimizer to solve the entire problem, in every other way it fills the criterion for a general problem specification.
+
+    The challenge with using this traditional mathematical syntax is that it is not easily manipulated or analyzed algorithmically. 
+    A number of matrix based methods have been used successfully to manipulate problem formulations in a programmatic manner. Steward's 
+    Design Structure Matrix (DSM) is a form of an adjacency matrix which captures the relationship between design elements and 
+    design variables where off diagonal elments of the matrix indicate coupling\cite{Steward1981}. 
+    Although it does not have to provide a specific solution path, the DSM has been successfully 
+    applied to allow selection of an optimal ordering for analysis tools in a complex system. Rogers et. al developed DeMAID to manipulate a
+    DSM to minimizing the computational costs of solving highly coupled systems\cite{Rogers1996}. 
+    Since these matrix based methods operate on some form of an adjacency matrix, they can all be represented in an equivalent graph.
 

### 2012-09-24 notes/example3.py

@@ -1,0 +1,211 @@
+import networkx as nx
+import matplotlib.pyplot as plt
+import set_ops as so
+import numpy
+import pylab
+
+pos=dict()
+
+def cab(DG,s,m,n,w,p,pos):
+    #create an anlysis block
+    si=s+'i'
+    so = s+'o'
+    DG.add_node(si)
+    DG.add_node(so)
+    pos.update({si:numpy.array([p[0]-0.1,p[1]]),
+                so:numpy.array([p[0]+0.1,p[1]])})
+    DG.add_weighted_edges_from([(si,so,w)])
+    DG[si][so]['type'] = 'analysis'
+    x = numpy.arange(1,m+1)
+    y = numpy.arange(1,n+1)
+    for j in x:
+        DG.add_node(si+str(j));
+        if m>1:
+            pos.update({si+str(j):numpy.array([p[0]-0.27,p[1]+0.2-0.4/(m-1)*(j-1)])})
+        else:
+            pos.update({si+str(j):numpy.array([p[0]-0.27,p[1]])})
+        DG.add_weighted_edges_from([(si+str(j),si,0)])
+        DG[si+str(j)][si]['type'] = 'in2analysis'
+    for j in y:
+        DG.add_node(so+str(j));
+        if n>1:
+            pos.update({so+str(j):numpy.array([p[0]+0.25,p[1]+0.2-0.4/(n-1)*(j-1)])})
+        else:
+            pos.update({so+str(j):numpy.array([p[0]+0.25,p[1]])})
+        DG.add_weighted_edges_from([(so,so+str(j),0)])
+        DG[so][so+str(j)]['type'] = 'analysis2out'
+    return;
+
+def dict2tree(Gx,DG1,DG2):
+    DG2.add_nodes_from(DG1.nodes())
+    nds=Gx.keys()
+    vls=Gx.values()
+    j=0
+    for i in nds:
+        if len(vls[j])>1:
+            nd=vls[j][-2]
+            DG2.add_weighted_edges_from([(nd,i,DG1[nd][i]['weight'])])
+        j+=1
+    return DG2;
+
+# Build the sample graph. Treat x as the root of all inputs without loss of generality.
+DG=nx.DiGraph()
+DG.add_nodes_from(['x','y1','y2'])
+pos.update({'x':numpy.array([1,2]),
+            'y1':numpy.array([3,2.5]),
[…]

### 2012-09-24 problem_formulation.tex
throwing around a few ideas

@@
 
-    - Large Scale engineered systems 
-    - Complex problems that are not obvious to set up
-    - need for a formal and consistent way to represent the problem
-      - contains *all* the information needed to formulate a solution method
-    - Previous attempts to provide problem formulation
-
-      \begin{itemize}
-        \item Structural Matrix, 
-        \item DSM
-        \item Functional Dependence Table
-        \item XDSM
-        \item Flow Chart
-        \item Adjacency Matrix
-      \end{itemize}
-
     As the size and complexity of engineering systems grows the time and expense for setting up 
@@
     in with a wide range of well establish graph-theory algorithms. By taking advantage of that structure it is possible to test 
-    any given problem specification to determine if it is in fact complete and general. It can be used to describe more specific problem 
+    any given problem specification to determine if it is complete and general. It can be used to describe more specific problem 
     setups, and given one can be used to derive the FPF from it. Lastly, a graph based specification for problem formulation lends itself
@@
         \\ s.t. &\ \ G\left(X\right) \leq 0
-        \\      &\ \ C\left(X\right) \leq 0
+        \\      &\ \ C\left(X\right) = 0
     \end{align}
@@
     Design Structure Matrix (DSM) is a form of an adjacency matrix which captures the relationship between design elements and 
-    design variables where off diagonal elments of the matrix indicate coupling\cite{Steward1981}. 
+    design variables where off diagonal elements of the matrix indicate coupling\cite{Steward1981}. 
     Although it does not have to provide a specific solution path, the DSM has been successfully 
@@
     DSM to minimizing the computational costs of solving highly coupled systems\cite{Rogers1996}. 
-    Since these matrix based methods operate on some form of an adjacency matrix, they can all be represented in an equivalent graph.
+    Since the DSM is essentially adjacency matrix, they can be represented in an equivalent graph. One major challenge with the DSM and similar 
+    approaches is that they lack the capability to fully describe the problem formulation. For example, the DSM deals with variables and analysis tools, 
+    but ignored objectives and constraints. An alternate matrix based representation, called a Functional Dependence Table (FTD), represents the relationship
+    between variables, objectives, and constraints but ignores the discipline analyses\cite{Michelena1997}. Both DSM and FTD have been used 
+    successfully to implement problem decomposition techniques, but neither by itself is sufficient to provide a syntax to describe a FPF.
 
-\section{Graph Based Problem Formulation}
+\section{Graph Based Problem Formulation Syntax}
+    \subsection{Formulation Graph Syntax}
+    Rather than start with an adjacency, we chose to work directly with a directed cyclic graph to develop a syntax for the FPF. 
+    \begin{itemize}
+        \item Specify analyses
+        \item Connections between analyses 
+            \begin{itemize}
+                \item local variables
+                \item global variables? Use "fake" node that broadcasts out to the rest of the graph? 
+            \end{itemize}
+        \item Cycles indicate coupling
+        \item Cycles for design variables->objectives/constraints
+        \item Objectives/constraints are just outputs? Special nodes? 
+        \item Residuals are just outputs? Special nodes? 
+        \item Parameters are just input nodes that are not design variables (use identifies these)
+        \item FPF no solvers/optimizers anywhere in it
+    \end{itemize}
+
+    \subsection{Solution Graph Syntax}
+    What is the difference between a problem formulation and a problem solution method? Convert from a cyclic graph, to an acyclic graph
+    \begin{itemize}
+        \item cycles indicate convergence loops or design varaible loops
+        \item Problem can't be solved until all loops are *removed* by adding in solvers/optimizers
+        \item *special* nodes for solvers and optimizers that *break* loops (from an algorithmic point of view)
+        \item FPF represents the minimal amount of information necessary to define a problem
+        \item Any solution path grows the graph complexity by adding edges and nodes (or possibly have an empty solution graph, which you build up 
+        as you remove edges from problem formulation graph?)
+    \end{itemize}
 

### 2012-09-28 problem_formulation.tex
added a bibtex file
making some progress

@@
     and paved the way for models of with more analysis tools and increasing numbers of multidisciplinary couplings. 
-    Such large models present a distinct challenge to properly set up in a manner that allows for an efficient solution. 
-    In fact as the size of a problem grows very large just determining what the proper solution strategy can be a daunting 
+    Such complex models present a distinct challenge proper implementation of any given solution strategy . 
+    In fact as the size of a problem grows very large even determining what the proper solution strategy can be a daunting 
     task.
 
-    What is needed, in order to address the complexity problem, is a method of specifying an engineering problem 
-    in a complete and totally general manner. A complete specification needs to include the following information: 
+    What is needed, in order to address the complexity problem, is a syntax for specifying an engineering problem 
+    in a way that allows for both analysis of potential solution strategies and automated implementation of those strategies 
+    inside an MDAO framework. In order to serve that purpose, the problem formulation needs to include the following information: 
     \begin{itemize}
@@
     \end{itemize}
-    A general specification is one that states nothing specific about a problem solution path. It does not include any details about 
-    how an objective is to be minimized, how coupling constraints will be satisfied, or how residuals will managed. We 
-    define a complete and general problem specification to be the Fundamental Problem Formulation (FPF). By definition, the FPF for 
-    any given problem will be constant regardless of which MDAO framework or optimization architecture is used to solve the problem.
+    Notably absent from the preceding list are any kind of solvers, optimizers, or other iterative solution finding tools. 
+    At it most basic, a problem formulation includes only information about what is being sought after in a given problem, or what the
+    goals of a given problem are. It need not contain any information about solution paths or strategies to reach those goals. 
+    A general specification is one that states nothing specific about a problem solution path.
 
-    In this work we proposed a graph based syntax for the specification of the FPF. This graph based syntax provides several key
+    We define a complete and general problem specification striped down to it's most essential parts to be the 
+    Fundamental Problem Formulation (FPF). By definition, the FPF for any given problem will be constant regardless 
+    of which MDAO framework, optimization architecture, optimizer, or solver is used to solve the problem.
+
+    In this work we proposed a graph based syntax for the specification of problem formulation. This graph based syntax provides several key
     features that make it useful for working with large scale MDAO problems. It provides a rigid structure that can be easily manipulated 
-    in with a wide range of well establish graph-theory algorithms. By taking advantage of that structure it is possible to test 
-    any given problem specification to determine if it is complete and general. It can be used to describe more specific problem 
-    setups, and given one can be used to derive the FPF from it. Lastly, a graph based specification for problem formulation lends itself
-    well to interacting with MDAO frameworks which makes it particularly powerful for integrating problem formulation analysis techniques 
-    into those frameworks.
+    with a wide range of well establish graph-theory algorithms for the purposes of problem decomposition. The graph syntax also 
+    provides a means for algorithmically testing a given problem formulation to check if it is the FPF, and if not, to reduce it to the FPF. 
+    Lastly, a graph based specification for problem formulation lends itself well to interacting with MDAO frameworks which dramatically 
+    increases it's utility for real world applications.
 
 \section{Problem Formulation Syntax}
-    The most common problem formulation for an MDAO problem could be given as follows:
+    The most common problem formulation syntax for an MDAO problem could be given as follows:
 
@@
     and $C(X)$ are the coupling constraints. This does represent a general problem formulation, since it says nothing about 
-    how the problem should be solved. Tedford and Martins used this format to specify a set of test problems and derived specific formulations 
-    for solving them with a number of optimization architectures\cite{Tedford2009}. Lamb and Martins also used this basic mathematical 
-    syntax to describe the All-At-Once (AAO) architecture \cite{Lambe2011}. Although the AAO specification does specify a solution path 
-    including a single optimizer to solve the entire problem, in every other way it fills the criterion for a general problem specification.
+    how the problem should be solved. Tedford and Martins used this format to specify the FPF for a set of test problems and 
+    also to describe specific formulations for solving them with a number of optimization architectures\cite{Tedford2009}. Their
+    work demonstrates clearly how multiple specific problem formulations can all relate back to a common FPF.
 
-    The challenge with using this traditional mathematical syntax is that it is not easily manipulated or analyzed algorithmically. 
-    A number of matrix based methods have been used successfully to manipulate problem formulations in a programmatic manner. Steward's 
-    Design Structure Matrix (DSM) is a form of an adjacency matrix which captures the relationship between design elements and 
-    design variables where off diagonal elements of the matrix indicate coupling\cite{Steward1981}. 
-    Although it does not have to provide a specific solution path, the DSM has been successfully 
-    applied to allow selection of an optimal ordering for analysis tools in a complex system. Rogers et. al developed DeMAID to manipulate a
+    The challenge with using this traditional mathematical syntax is that it is not easily manipulated or analyzed by computer programs. 
+    A number of matrix based methods have been used successfully to translate the mathematical syntax into a more useful computational form. 
+    Steward's Design Structure Matrix (DSM) is a form of an adjacency matrix which captures the relationship between analysis tools where off 
+    diagonal elements of the matrix indicate coupling\cite{Steward1981}.
+
+    [[insert cartoon figure here of DSM]]
+
+    The DSM has been widely adopted within the engineering and optimization fields, and even outside of it. The DSM
+    syntax has been successfully applied to allow selection of beneficial orderings for analysis 
+    tools in a complex system. Rogers et. al developed DeMAID to manipulate a
     DSM to minimizing the computational costs of solving highly coupled systems\cite{Rogers1996}.
-    Since the DSM is essentially adjacency matrix, they can be represented in an equivalent graph. One major challenge with the DSM and similar 
-    approaches is that they lack the capability to fully describe the problem formulation. For example, the DSM deals with variables and analysis tools, 
-    but ignored objectives and constraints. An alternate matrix based representation, called a Functional Dependence Table (FTD), represents the relationship
-    between variables, objectives, and constraints but ignores the discipline analyses\cite{Michelena1997}. Both DSM and FTD have been used 
-    successfully to implement problem decomposition techniques, but neither by itself is sufficient to provide a syntax to describe a FPF.
+
+    [Need to cite more recent research on automatic paritioning, Peter's work was just published...]
+    [Include examples of DSM being used in the business world that Brian Talked about?]
+
+    Since a DSM essentially describes adjacency matrix, it can be represented in an equivalent directed graph where nodes represent analysis tools and 
+    edges represent information exchange between those tools. An alternate matrix based syntax, called a 
+    Functional Dependence Table (FDT), was proposed by Michelena and Papalambros. 
+    FDT represents the relationship between functions, including objectives and constraints, and their values\cite{Michelena1997}. Similar to DSM
+    FDT also describes an adjacency matrix of a graph. Unlike the DSM graph, however, a FDT graph is an undirected 
+    graph where nodes can represent analysis tools, objectives, or constraints. Edges between nodes represent a dependence on the same 
+    variable value. By searching the FDT graph for clusters of totally connected nodes Wagner and Papalambros were identify groups of 
+    analysis tools that were all dependent on the same input variables and used that to make partitioning decisions \cite{Wagner1993}.
+
+    [[insert cartoon figure of a FDT graph and matrix]]
 
@@
 
+\bibliography{library}
 \end{document}

### 2012-09-28 problem_formulation.tex
added in some figures

@@
     Steward's Design Structure Matrix (DSM) is a form of an adjacency matrix which captures the relationship between analysis tools where off 
-    diagonal elements of the matrix indicate coupling\cite{Steward1981}.
+    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Figure \ref{fig:dsm_simple} shows a simple notional example of a DSM. 
+    The DSM has been successfully applied in a number of ways to algorithmically modify problem formulations to 
+    improve computational efficiency. Rogers et. al developed DeMAID to manipulate a
+    DSM  minimizing the computational costs of solving highly coupled systems\cite{Rogers1996}. Lu and Martins demonstrated 
+    the use of a weighted DSM to make partitioning and coordinating choices that dramatically reduced the 
+    cost of solving a specific kind of optimization problem as the scale of the problem grew. 
+    They applied weights to both the nodes (diagonal terms) and the edges (off diagonal terms) of the 
+    DSM to indicated computational cost and coupling strength respectively.
 
-    [[insert cartoon figure here of DSM]]
+    \begin{figure}[!hbp]
+        \begin{center}
+        \includegraphics[width=.75\textwidth]{images/dsm_simple}
+        \caption{Graph and matrix based DSM specifications for a notional analysis \label{fig:dsm_simple}}
+        \end{center}
+    \end{figure}
 
-    The DSM has been widely adopted within the engineering and optimization fields, and even outside of it. The DSM
-    syntax has been successfully applied to allow selection of beneficial orderings for analysis 
-    tools in a complex system. Rogers et. al developed DeMAID to manipulate a
-    DSM to minimizing the computational costs of solving highly coupled systems\cite{Rogers1996}.
-
-    [Need to cite more recent research on automatic paritioning, Peter's work was just published...]
-    [Include examples of DSM being used in the business world that Brian Talked about?]
-
-    Since a DSM essentially describes adjacency matrix, it can be represented in an equivalent directed graph where nodes represent analysis tools and 
+    Since a DSM describes a square adjacency matrix, it can be represented in an equivalent directed graph where nodes represent analysis tools and 
     edges represent information exchange between those tools. An alternate matrix based syntax, called a 
@@
 
-    [[insert cartoon figure of a FDT graph and matrix]]
+    It is possible to combine the syntax of an FDT and a DSM by making a subtle extension to a traditional DSM specification. Normally, a DSM is given 
+    with analysis tools as nodes and variable dependencies given as edges. Lamb and Martins included the variables, objectives, and constraint functions
+    as nodes in an Extented DSM (XDSM)\cite{Lambe2012} in order to capture a more complete problem formulation for MDAO problems. With XDSM 
+    it is possible to also encompass the relationships defined in an FDT. Figure \ref{fig:dsm_full} 
+    illustrates the same notional analysis as in Figure \ref{fig:dsm_simple} represented as an XDSM. 
+    The dashed boxes indicate groups of nodes that could be collapsed back down to retrieve the simpler 
+    DSM from above. Beyond specifying relationships between data in a problem formulation, XDSM also includes syntax to describe the
+    procedural order of any given solution path. Figure \ref{fig:xdsm_full} shows the full XDSM specification for solving the notional 
+    analysis with a traditional Multidisciplinary Design Feasible (MDF) architecture.
+
+    \begin{figure}[!hbp]
+        \begin{center}
+        \includegraphics[width=.75\textwidth]{images/dsm_full}
+        \caption{Extended DSM graph for a notional analysis \label{fig:dsm_full}}
+        \end{center}
+    \end{figure}
 

### 2012-10-01 problem_formulation.tex
DJP: added to the graph syntax section

@@
-\documentclass[]{aiaa-tc} % insert '[draft]' option to show overfull boxes
-
- \title{An application of Graph Theory to MDAO problem formulation}
-
-\author{
-  David Pate, %
-     \thanks{Georgia Tech}
-  Dr. Brian German 
-     \thanks{Georgia Tech...}
-  Justin Gray,%
-     \thanks{Aerospace Engineer, MDAO Branch, Mail Stop 5-11, AIAA Member}   
- }
-
-%\usepackage{setspace}
-%\doublespace
-
-\usepackage{graphicx}
-\usepackage{wrapfig}
-\usepackage{caption} 
-\usepackage{amsmath}
-\usepackage{lscape}
-\usepackage{hyperref}
-\usepackage{appendix}
-\usepackage[section]{placeins}
-
-\captionsetup[figure]{margin=5pt,font=small,labelfont=bf,textfont=bf,justification=justified,}
-%\captionsetup[wrapfigure]{margin=5pt,font=small,labelfont=bf,justification=justified,singlelinecheck=off}
-\captionsetup[table]{margin=5pt,font=small,labelfont=bf,textfont=bf,justification=justified,position=top}
-
-\bibliographystyle{aiaa}
-
-\usepackage{lettrine}
-\usepackage{verbatim}
-
-%\usepackage{hyperref} %allows for the creation of actual text links
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-   blah blah blah
-\end{abstract}
-
-\section*{Nomenclature}
-
-\begin{tabular}{l l} 
-    AAO      & All-At- \\
-    MDAO     & Multidisciplinary Design Analysis and Optimization \\
-    FPF      & Fundamental Problem Formulation \\
-\end{tabular}
-
-\section{Introduction}
-
-    As the size and complexity of engineering systems grows the time and expense for setting up 
-    analysis models grows with them. Multidisciplinary Design Analysis and Optimization (MDAO)
-    frameworks such as OpenMDAO\cite{Gray2012} and ModelCenter have enabled a new level of analysis tool integration 
-    and paved the way for models of with more analysis tools and increasing numbers of multidisciplinary couplings. 
-    Such complex models present a distinct challenge proper implementation of any given solution strategy . 
-    In fact as the size of a problem grows very large even determining what the proper solution strategy can be a daunting 
-    task.
-
-    What is needed, in order to address the complexity problem, is a syntax for specifying an engineering problem 
-    in a way that allows for both analysis of potential solution strategies and automated implementation of those strategies 
-    inside an MDAO framework. In order to serve that purpose, the problem formulation needs to include the following information: 
-    \begin{itemize}
-       \item Discipline Analyses 
-       \item Discipline State Variables and Residuals
-       \item Local \& Global Design Variables
-       \item Local \& Global Constraints
-       \item Objective or Objectives
-       \item Coupling Constraints
-       \item Local and Global Parameters
-    \end{itemize}
-    Notably absent from the preceding list are any kind of solvers, optimizers, or other iterative solution finding tools. 
-    At it most basic, a problem formulation includes only information about what is being sought after in a given problem, or what the
-    goals of a given problem are. It need not contain any information about solution paths or strategies to reach those goals. 
-    A general specification is one that states nothing specific about a problem solution path.
-
-    We define a complete and general problem specification striped down to it's most essential parts to be the 
-    Fundamental Problem Formulation (FPF). By definition, the FPF for any given problem will be constant regardless 
-    of which MDAO framework, optimization architecture, optimizer, or solver is used to solve the problem.
-
-    In this work we proposed a graph based syntax for the specification of problem formulation. This graph based syntax provides several key
-    features that make it useful for working with large scale MDAO problems. It provides a rigid structure that can be easily manipulated 
-    with a wide range of well establish graph-theory algorithms for the purposes of problem decomposition. The graph syntax also 
-    provides a means for algorithmically testing a given problem formulation to check if it is the FPF, and if not, to reduce it to the FPF. 
-    Lastly, a graph based specification for problem formulation lends itself well to interacting with MDAO frameworks which dramatically 
-    increases it's utility for real world applications.
-
-\section{Problem Formulation Syntax}
-    The most common problem formulation syntax for an MDAO problem could be given as follows:
-
-    \begin{align}
-        min &\ \ f\left(X\right) \notag
-        \\ w.r.t. &\ \  X \notag
-        \\ s.t. &\ \ G\left(X\right) \leq 0
-        \\      &\ \ C\left(X\right) = 0
-    \end{align}
-
-    Where $f(X)$ is the objective function, $X$ is the vector of design variables, $G(X)$ are the optimization constraints, 
-    and $C(X)$ are the coupling constraints. This does represent a general problem formulation, since it says nothing about 
-    how the problem should be solved. Tedford and Martins used this format to specify the FPF for a set of test problems and 
-    also to describe specific formulations for solving them with a number of optimization architectures\cite{Tedford2009}. Their
-    work demonstrates clearly how multiple specific problem formulations can all relate back to a common FPF.
-
-    The challenge with using this traditional mathematical syntax is that it is not easily manipulated or analyzed by computer programs. 
-    A number of matrix based methods have been used successfully to translate the mathematical syntax into a more useful computational form. 
-    Steward's Design Structure Matrix (DSM) is a form of an adjacency matrix which captures the relationship between analysis tools where off 
-    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Figure \ref{fig:dsm_simple} shows a simple notional example of a DSM. 
-    The DSM has been successfully applied in a number of ways to algorithmically modify problem formulations to 
-    improve computational efficiency. Rogers et. al developed DeMAID to manipulate a
-    DSM  minimizing the computational costs of solving highly coupled systems\cite{Rogers1996}. Lu and Martins demonstrated 
-    the use of a weighted DSM to make partitioning and coordinating choices that dramatically reduced the 
-    cost of solving a specific kind of optimization problem as the scale of the problem grew. 
-    They applied weights to both the nodes (diagonal terms) and the edges (off diagonal terms) of the 
-    DSM to indicated computational cost and coupling strength respectively.
-
-    \begin{figure}[!hbp]
-        \begin{center}
-        \includegraphics[width=.75\textwidth]{images/dsm_simple}
-        \caption{Graph and matrix based DSM specifications for a notional analysis \label{fig:dsm_simple}}
-        \end{center}
-    \end{figure}
-
     Since a DSM describes a square adjacency matrix, it can be represented in an equivalent directed graph where nodes represent analysis tools and
@@
     Rather than start with an adjacency, we chose to work directly with a directed cyclic graph to develop a syntax for the FPF. 
+    The following information must be provided to start:
     \begin{itemize}
-        \item Specify analyses
-        \item Connections between analyses 
+        \item Analysis blocks: an analysis block represents any calculation, and each comes with
             \begin{itemize}
-                \item local variables
-                \item global variables? Use "fake" node that broadcasts out to the rest of the graph? 
+                \item local inputs
+                \item local outputs 
+                \item execution properties: the properties associated with running the code, such as run time.
             \end{itemize}
-        \item Cycles indicate coupling
-        \item Cycles for design variables->objectives/constraints
-        \item Objectives/constraints are just outputs? Special nodes? 
-        \item Residuals are just outputs? Special nodes? 
-        \item Parameters are just input nodes that are not design variables (use identifies these)
-        \item FPF no solvers/optimizers anywhere in it
+        \item Global parameters: these may serve as fixed inputs to the local inputs of analysis blocks
+        \item Global outputs: these may represent
+            \begin{itemize}
+                \item objectives
+                \item constraints
+                \item residuals: (not sure)
             \end{itemize}
+    \end{itemize}
+
+    The graph representation of a data flow is cast to utilize the extensive library of algorithms in graph theory to analyze a directed weighted graph. 
+    Edge weights are used to represent the metrics associated with a data flow
+    \begin{itemize}
+        \item Run time: this metric is a property of an analysis block
+        \item Fidelity: this metric is a property of an individual local output
+        \item Expected Convergence
+    \end{itemize}
+
+    The key assumption is that identical variables are recognized as such. This serves as the basis for creating a data flow by connecting compatible input and output nodes with a directed edge. 
+    In order to represent the fact that execution of an analysis code does not depend on the number of outputs being used, it is represented as depicted in the following figure (not made yet).
+
+    This information immediately leads to the maximal connectivity graph, which is formed by placing a directed edge from each local output or global parameter to each matching local input or global output.
+    Whenever multiple edges are connected to a single input a conflict occurs because only one may be used. Resolving these conflicts is one key challenge in creating a data flow.
+%    \begin{itemize}
+%        \item Specify analyses
+%        \item Connections between analyses 
+%            \begin{itemize}
+%                \item local variables
+%                \item global variables? Use "fake" node that broadcasts out to the rest of the graph? 
+%            \end{itemize}
+%        \item Cycles indicate coupling
+%        \item Cycles for design variables->objectives/constraints
+%        \item Objectives/constraints are just outputs? Special nodes? 
+%        \item Residuals are just outputs? Special nodes? 
+%        \item Parameters are just input nodes that are not design variables (use identifies these)
+%        \item FPF no solvers/optimizers anywhere in it
+%    \end{itemize}
 

### 2012-10-01 problem_formulation.tex
starting for the day

--- 2012-09-28 problem_formulation.tex
@@
     Such complex models present a distinct challenge proper implementation of any given solution strategy . 
-    In fact as the size of a problem grows very large even determining what the proper solution strategy can be a daunting 
+    In fact as the size of a problem grows very large even selecting an appropriate solution strategy can be a daunting 
     task.
 
-    What is needed, in order to address the complexity problem, is a syntax for specifying an engineering problem 
+    To help address in order to address the complexity problem we propose the use of graph based problem formulation 
     in a way that allows for both analysis of potential solution strategies and automated implementation of those strategies 
-    inside an MDAO framework. In order to serve that purpose, the problem formulation needs to include the following information: 
+    inside an MDAO framework. In order to serve that purpose, the problem  needs to include the following information: 
     \begin{itemize}
@@
        \item Discipline State Variables and Residuals
-       \item Local \& Global Design Variables
-       \item Local \& Global Constraints
+       \item Design Variables
+       \item Constraints
        \item Objective or Objectives
        \item Coupling Constraints
-       \item Local and Global Parameters
+       \item Parameters
     \end{itemize}
@@
 
-\section{Problem Formulation Syntax}
-    The most common problem formulation syntax for an MDAO problem could be given as follows:
+\section{Fundamental Problem Formulation}
+    Problem formulation is most commonly given in a mathematical syntax. For a simple, notional problem the problem formulation
+    can be given as follows:
 
     \begin{align}
-        min &\ \ f\left(X\right) \notag
-        \\ w.r.t. &\ \  X \notag
-        \\ s.t. &\ \ G\left(X\right) \leq 0
-        \\      &\ \ C\left(X\right) = 0
+        given & \ \ A.f\left(A.x,A.y\right) \notag
+        \\ & \ \  B.f\left(B.x,B.z\right) \notag
+        \\ & \ \  C.f1\left(A.f,B.f,C.s\right) \notag
+        \\ & \ \  C.f2\left(A.f,B.f,C.s\right) \notag
+        \\min &\ \ F\left(C.f1\right) \notag
+        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
+        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
+        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
+        \\      &\ \ E\left(A.y,C.f1\right) = 0 \notag
+        \\      &\ \ R\left(C.s,C.f2\right) = 0 
+        \label{eqn:simple}
     \end{align}
 
-    Where $f(X)$ is the objective function, $X$ is the vector of design variables, $G(X)$ are the optimization constraints, 
-    and $C(X)$ are the coupling constraints. This does represent a general problem formulation, since it says nothing about 
-    how the problem should be solved. Tedford and Martins used this format to specify the FPF for a set of test problems and 
-    also to describe specific formulations for solving them with a number of optimization architectures\cite{Tedford2009}. Their
-    work demonstrates clearly how multiple specific problem formulations can all relate back to a common FPF.
+    Where $A,B,C$ represent analysis tools, $F,G1,G2,E,R$ are objective and constraint functions, and each analysis tool 
+    has a set of input and output variables. While this does represent a complete problem formulation, some assumptions have
+    inherently been made about the solution strategy. A coupling constraint, $E$, and $C$ is directly dependent
+    $A.f$ and $B.f$ as it's inputs. This implies that first you must run $A$ and $B$, then $C$ and
+    iterate on the design variables in $A$ and $B$ until you satisfy $E$. However, a different solution could be equally valid and still represent
+    the same fundamental problem:
 
-    The challenge with using this traditional mathematical syntax is that it is not easily manipulated or analyzed by computer programs. 
+    \begin{align}
+        given & \ \ A.f\left(A.x,C.f1\right) \notag
+        \\ & \ \  B.f\left(B.x,B.z\right) \notag
+        \\ & \ \  C.f1\left(C.y,B.f,C.s\right) \notag
+        \\ & \ \  C.f2\left(C.q,B.f,C.s\right) \notag
+        \\min &\ \ F\left(C.f1\right) \notag
+        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
+        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
+        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
+        \\      &\ \ E\left(A.f,C.q\right) = 0 \notag
+        \\      &\ \ R\left(C.s,C.f2\right) = 0 
+        \label{eqn:simple_2}
+    \end{align}
+
+    Equation \ref{eqn:simple_2} differs only slightly from Eqn. \ref{eqn:simple}. $A$ is now directly dependent on the output of $C$, 
+    and the constraint $E$ has changed as well. Now the problem must be solved by running $B$ and $C$, then $A$ and iterating 
+    on the design variables in $B$ and $C$ until you satisfy $E$. Since the formulations in Eqn. \ref{eqn:simple} 
+    and \ref{eqn:simple_2} both describe the same problem there must be a more fundamental description of the problem that is common
+    between them. We present the FPF as follows:
+
+    \begin{align}
+        given & \ \ A.f\left(A.x,A.y\right) \notag
+        \\ & \ \  B.f\left(B.x,B.z\right) \notag
+        \\ & \ \  C.f1\left(C.y,C.r,C.s\right) \notag
+        \\ & \ \  C.f2\left(C.q,C.r,C.s\right) \notag
+        \\min &\ \ F\left(C.f1\right) \notag
+        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
+        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
+        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
+        \\      &\ \ E\left(A.f,C.\right) = 0 \notag
+        \\      &\ \ R\left(C.s,C.f2\right) = 0 
+        \label{eqn:simple_2}
+    \end{align}
+
+    \section{Problem Formulation Syntax}
+
+    The same mathematical syntax can be used to describe more complex problem formulations
+    such as ones that would decompose the FPF to a set of sub-problems. Tedford and Martins used this format to specify the 
+    FPF for a set of test problems and also to describe specific formulations for solving them with a 
+    number of optimization architectures\cite{Tedford2009}. Their work demonstrates clearly how multiple specific 
+    problem formulations can all relate back to a common FPF.
+
+    In the above problem formulation, there is no explicit relationship definition of local and global features. Instead they are 
+    implicitly defined through variable naming conventions. For instance, $x$ is a global variable because it is used in $A.x$ and $B.x$. 
+    Simliarly, $A.y$ is a local variable because it only affects the of $A$. If such a naming convention is not in place, 
+    then a more explicit definition of the relation between variables must also be given in the form of additional equality constraints between 
+    input variables. Similarly, both the constraint, $C$ and the residual, $R$ are presented as equality constraints. They could be classified as 
+    a global and a local constraint on the problem respectively, with the only difference being the number of analysis tools that affect them.  
+    Hence, the distinction between local and global properties are not necessary to form a fundamental 
+    problem formulation. On the other hand, they are relevant when deriving a specific solution path for a problem, so any definition of problem 
+    formulation should include the capability to detect the weather a given property is local or global.
+
+    The challenge with using this traditional mathematical syntax is that it is not easily manipulated or analyzed. 
     A number of matrix based methods have been used successfully to translate the mathematical syntax into a more useful computational form. 
-    Steward's Design Structure Matrix (DSM) is a form of an adjacency matrix which captures the relationship between analysis tools where off 
-    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Figure \ref{fig:dsm_simple} shows a simple notional example of a DSM. 
+    Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools where off 
+    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Figure \ref{fig:dsm_simple} shows a DSM for the above notional example. 
     The DSM has been successfully applied in a number of ways to algorithmically modify problem formulations to 
     improve computational efficiency. Rogers et. al developed DeMAID to manipulate a
-    DSM  minimizing the computational costs of solving highly coupled systems\cite{Rogers1996}. Lu and Martins demonstrated 
-    the use of a weighted DSM to make partitioning and coordinating choices that dramatically reduced the 
-    cost of solving a specific kind of optimization problem as the scale of the problem grew. 
-    They applied weights to both the nodes (diagonal terms) and the edges (off diagonal terms) of the 
-    DSM to indicated computational cost and coupling strength respectively.
+    DSM  minimizing the computational costs of solving highly coupled systems\cite{Rogers1996}. Notice however, that the DSM in 
+    Figure \ref{fig:dsm_simple} contains less information than the full problem formulation. Objective and constraint information 
+    is missing and it's not possible to distinguish individual variables.
 
@@
     variable value. By searching the FDT graph for clusters of totally connected nodes Wagner and Papalambros were identify groups of 
-    analysis tools that were all dependent on the same input variables and used that to make partitioning decisions \cite{Wagner1993}.
+    analysis tools that were all dependent on the same input variables and used that to make partitioning decisions \cite{Wagner1993}. FDT retains 
+    a greater portion of the information in the problem formulation, but it is still not complete. It ignores information about feedback coupling 
+    between codes.
 
-    It is possible to combine the syntax of an FDT and a DSM by making a subtle extension to a traditional DSM specification. Normally, a DSM is given 
+    It is possible to combine the syntax of an FDT and a DSM by making a small extension to a traditional DSM specification. Normally, a DSM is given 
     with analysis tools as nodes and variable dependencies given as edges. Lamb and Martins included the variables, objectives, and constraint functions
-    as nodes in an Extented DSM (XDSM)\cite{Lambe2012} in order to capture a more complete problem formulation for MDAO problems. With XDSM 
-    it is possible to also encompass the relationships defined in an FDT. Figure \ref{fig:dsm_full} 
-    illustrates the same notional analysis as in Figure \ref{fig:dsm_simple} represented as an XDSM. 
-    The dashed boxes indicate groups of nodes that could be collapsed back down to retrieve the simpler 
-    DSM from above. Beyond specifying relationships between data in a problem formulation, XDSM also includes syntax to describe the
-    procedural order of any given solution path. Figure \ref{fig:xdsm_full} shows the full XDSM specification for solving the notional 
-    analysis with a traditional Multidisciplinary Design Feasible (MDF) architecture.
+    as nodes in an Extended DSM (XDSM)\cite{Lambe2012} in order to capture a more complete problem formulation for MDAO problems. With XDSM 
+    it is possible to specify a complete problem formulation. Lu and Martins developed a process to perform ordering and partitioning on the fundamental 
+    problem formulation specified by XDSM in order to achieve significant reductions in computational costs for large scale optimization problems. 
+    Figure \ref{fig:dsm_full} illustrates the same notional analysis as in Figure \ref{fig:dsm_simple} represented as an XDSM.
 
@@
         \begin{center}
-        \includegraphics[width=.75\textwidth]{images/dsm_full}
-        \caption{Extended DSM graph for a notional analysis \label{fig:dsm_full}}
+        \includegraphics[width=.75\textwidth]{images/dsm_simple}
+        \caption{[TODO: Generate XDSM]Extended DSM graph for a notional analysis \label{fig:dsm_full}}
         \end{center}
@@
 
-\section{Graph Based Problem Formulation Syntax}
+\section{Graph Based Problem Formulation Using XDSM}
     \subsection{Formulation Graph Syntax}

### 2012-10-01 problem_formulation.tex
stuff

@@
     Where $A,B,C$ represent analysis tools, $F,G1,G2,E,R$ are objective and constraint functions, and each analysis tool 
-    has a set of input and output variables. While this does represent a complete problem formulation, some assumptions have
-    inherently been made about the solution strategy. A coupling constraint, $E$, and $C$ is directly dependent
+    has a set of input and output variables. While this does represent a complete problem formulation, 
+    it's not the fundamental one since some assumptions have inherently been made about the solution 
+    strategy. $C$ is directly dependent
     $A.f$ and $B.f$ as it's inputs. This implies that first you must run $A$ and $B$, then $C$ and

### 2012-10-01 problem_formulation.tex

@@
 
-\section{Graph Based Problem Formulation Using XDSM}
+\section{Graph Based Problem Formulation Syntax}
     \subsection{Formulation Graph Syntax}
     Rather than start with an adjacency, we chose to work directly with a directed cyclic graph to develop a syntax for the FPF. 
+    The following information must be provided to start:
     \begin{itemize}
-        \item Specify analyses
-        \item Connections between analyses 
+        \item Analysis blocks: an analysis block represents any calculation, and each comes with
             \begin{itemize}
-                \item local variables
-                \item global variables? Use "fake" node that broadcasts out to the rest of the graph? 
+                \item local inputs
+                \item local outputs 
+                \item execution properties: the properties associated with running the code, such as run time.
             \end{itemize}
-        \item Cycles indicate coupling
-        \item Cycles for design variables->objectives/constraints
-        \item Objectives/constraints are just outputs? Special nodes? 
-        \item Residuals are just outputs? Special nodes? 
-        \item Parameters are just input nodes that are not design variables (use identifies these)
-        \item FPF no solvers/optimizers anywhere in it
+        \item Global parameters: these may serve as fixed inputs to the local inputs of analysis blocks
+        \item Global outputs: these may represent
+            \begin{itemize}
+                \item objectives
+                \item constraints
+                \item residuals: (not sure)
             \end{itemize}
+    \end{itemize}
+
+    The graph representation of a data flow is cast to utilize the extensive library of algorithms in graph theory to analyze a directed weighted graph. 
+    Edge weights are used to represent the metrics associated with a data flow
+    \begin{itemize}
+        \item Run time: this metric is a property of an analysis block
+        \item Fidelity: this metric is a property of an individual local output
+        \item Expected Convergence
+    \end{itemize}
+
+    The key assumption is that identical variables are recognized as such. This serves as the basis for creating a data flow by connecting compatible input and output nodes with a directed edge. 
+    In order to represent the fact that execution of an analysis code does not depend on the number of outputs being used, it is represented as depicted in the following figure (not made yet).
+
+    This information immediately leads to the maximal connectivity graph, which is formed by placing a directed edge from each local output or global parameter to each matching local input or global output.
+    Whenever multiple edges are connected to a single input a conflict occurs because only one may be used. Resolving these conflicts is one key challenge in creating a data flow.
+%    \begin{itemize}
+%        \item Specify analyses
+%        \item Connections between analyses 
+%            \begin{itemize}
+%                \item local variables
+%                \item global variables? Use "fake" node that broadcasts out to the rest of the graph? 
+%            \end{itemize}
+%        \item Cycles indicate coupling
+%        \item Cycles for design variables->objectives/constraints
+%        \item Objectives/constraints are just outputs? Special nodes? 
+%        \item Residuals are just outputs? Special nodes? 
+%        \item Parameters are just input nodes that are not design variables (use identifies these)
+%        \item FPF no solvers/optimizers anywhere in it
+%    \end{itemize}
 

### 2012-10-02 graph_syntax.tex
breaking things up

--- 2012-10-01 problem_formulation.tex
@@
-\documentclass[]{aiaa-tc} % insert '[draft]' option to show overfull boxes
-
- \title{An application of Graph Theory to MDAO problem formulation}
-
-\author{
-  David Pate, %
-     \thanks{Georgia Tech}
-  Dr. Brian German 
-     \thanks{Georgia Tech...}
-  Justin Gray,%
-     \thanks{Aerospace Engineer, MDAO Branch, Mail Stop 5-11, AIAA Member}   
- }
-
-%\usepackage{setspace}
-%\doublespace
-
-\usepackage{graphicx}
-\usepackage{wrapfig}
-\usepackage{caption} 
-\usepackage{amsmath}
-\usepackage{lscape}
-\usepackage{hyperref}
-\usepackage{appendix}
-\usepackage[section]{placeins}
-
-\captionsetup[figure]{margin=5pt,font=small,labelfont=bf,textfont=bf,justification=justified,}
-%\captionsetup[wrapfigure]{margin=5pt,font=small,labelfont=bf,justification=justified,singlelinecheck=off}
-\captionsetup[table]{margin=5pt,font=small,labelfont=bf,textfont=bf,justification=justified,position=top}
-
-\bibliographystyle{aiaa}
-
-\usepackage{lettrine}
-\usepackage{verbatim}
-
-%\usepackage{hyperref} %allows for the creation of actual text links
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-   blah blah blah
-\end{abstract}
-
-\section*{Nomenclature}
-
-\begin{tabular}{l l} 
-    AAO      & All-At- \\
-    MDAO     & Multidisciplinary Design Analysis and Optimization \\
-    FPF      & Fundamental Problem Formulation \\
-\end{tabular}
-
-\section{Introduction}
-
-    As the size and complexity of engineering systems grows the time and expense for setting up 
-    analysis models grows with them. Multidisciplinary Design Analysis and Optimization (MDAO)
-    frameworks such as OpenMDAO\cite{Gray2012} and ModelCenter have enabled a new level of analysis tool integration 
-    and paved the way for models of with more analysis tools and increasing numbers of multidisciplinary couplings. 
-    Such complex models present a distinct challenge proper implementation of any given solution strategy . 
-    In fact as the size of a problem grows very large even selecting an appropriate solution strategy can be a daunting 
-    task.
-
-    To help address in order to address the complexity problem we propose the use of graph based problem formulation 
-    in a way that allows for both analysis of potential solution strategies and automated implementation of those strategies 
-    inside an MDAO framework. In order to serve that purpose, the problem  needs to include the following information: 
-    \begin{itemize}
-       \item Discipline Analyses 
-       \item Discipline State Variables and Residuals
-       \item Design Variables
-       \item Constraints
-       \item Objective or Objectives
-       \item Coupling Constraints
-       \item Parameters
-    \end{itemize}
-    Notably absent from the preceding list are any kind of solvers, optimizers, or other iterative solution finding tools. 
-    At it most basic, a problem formulation includes only information about what is being sought after in a given problem, or what the
-    goals of a given problem are. It need not contain any information about solution paths or strategies to reach those goals. 
-    A general specification is one that states nothing specific about a problem solution path.
-
-    We define a complete and general problem specification striped down to it's most essential parts to be the 
-    Fundamental Problem Formulation (FPF). By definition, the FPF for any given problem will be constant regardless 
-    of which MDAO framework, optimization architecture, optimizer, or solver is used to solve the problem.
-
-    In this work we proposed a graph based syntax for the specification of problem formulation. This graph based syntax provides several key
-    features that make it useful for working with large scale MDAO problems. It provides a rigid structure that can be easily manipulated 
-    with a wide range of well establish graph-theory algorithms for the purposes of problem decomposition. The graph syntax also 
-    provides a means for algorithmically testing a given problem formulation to check if it is the FPF, and if not, to reduce it to the FPF. 
-    Lastly, a graph based specification for problem formulation lends itself well to interacting with MDAO frameworks which dramatically 
-    increases it's utility for real world applications.
-
-\section{Fundamental Problem Formulation}
-    Problem formulation is most commonly given in a mathematical syntax. For a simple, notional problem the problem formulation
-    can be given as follows:
-
-    \begin{align}
-        given & \ \ A.f\left(A.x,A.y\right) \notag
-        \\ & \ \  B.f\left(B.x,B.z\right) \notag
-        \\ & \ \  C.f1\left(A.f,B.f,C.s\right) \notag
-        \\ & \ \  C.f2\left(A.f,B.f,C.s\right) \notag
-        \\min &\ \ F\left(C.f1\right) \notag
-        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
-        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
-        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
-        \\      &\ \ E\left(A.y,C.f1\right) = 0 \notag
-        \\      &\ \ R\left(C.s,C.f2\right) = 0 
-        \label{eqn:simple}
-    \end{align}
-
-    Where $A,B,C$ represent analysis tools, $F,G1,G2,E,R$ are objective and constraint functions, and each analysis tool 
-    has a set of input and output variables. While this does represent a complete problem formulation, 
-    it's not the fundamental one since some assumptions have inherently been made about the solution 
-    strategy. $C$ is directly dependent
-    $A.f$ and $B.f$ as it's inputs. This implies that first you must run $A$ and $B$, then $C$ and
-    iterate on the design variables in $A$ and $B$ until you satisfy $E$. However, a different solution could be equally valid and still represent
-    the same fundamental problem:
-
-    \begin{align}
-        given & \ \ A.f\left(A.x,C.f1\right) \notag
-        \\ & \ \  B.f\left(B.x,B.z\right) \notag
-        \\ & \ \  C.f1\left(C.y,B.f,C.s\right) \notag
-        \\ & \ \  C.f2\left(C.q,B.f,C.s\right) \notag
-        \\min &\ \ F\left(C.f1\right) \notag
-        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
-        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
-        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
-        \\      &\ \ E\left(A.f,C.q\right) = 0 \notag
-        \\      &\ \ R\left(C.s,C.f2\right) = 0 
-        \label{eqn:simple_2}
-    \end{align}
-
-    Equation \ref{eqn:simple_2} differs only slightly from Eqn. \ref{eqn:simple}. $A$ is now directly dependent on the output of $C$, 
-    and the constraint $E$ has changed as well. Now the problem must be solved by running $B$ and $C$, then $A$ and iterating 
-    on the design variables in $B$ and $C$ until you satisfy $E$. Since the formulations in Eqn. \ref{eqn:simple} 
-    and \ref{eqn:simple_2} both describe the same problem there must be a more fundamental description of the problem that is common
-    between them. We present the FPF as follows:
-
-    \begin{align}
-        given & \ \ A.f\left(A.x,A.y\right) \notag
-        \\ & \ \  B.f\left(B.x,B.z\right) \notag
-        \\ & \ \  C.f1\left(C.y,C.r,C.s\right) \notag
-        \\ & \ \  C.f2\left(C.q,C.r,C.s\right) \notag
-        \\min &\ \ F\left(C.f1\right) \notag
-        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
-        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
-        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
-        \\      &\ \ E\left(A.f,C.\right) = 0 \notag
-        \\      &\ \ R\left(C.s,C.f2\right) = 0 
-        \label{eqn:simple_2}
-    \end{align}
-
-    \section{Problem Formulation Syntax}
-
-    The same mathematical syntax can be used to describe more complex problem formulations
-    such as ones that would decompose the FPF to a set of sub-problems. Tedford and Martins used this format to specify the 
-    FPF for a set of test problems and also to describe specific formulations for solving them with a 
-    number of optimization architectures\cite{Tedford2009}. Their work demonstrates clearly how multiple specific 
-    problem formulations can all relate back to a common FPF.
-
-    In the above problem formulation, there is no explicit relationship definition of local and global features. Instead they are 
-    implicitly defined through variable naming conventions. For instance, $x$ is a global variable because it is used in $A.x$ and $B.x$. 
-    Simliarly, $A.y$ is a local variable because it only affects the of $A$. If such a naming convention is not in place, 
-    then a more explicit definition of the relation between variables must also be given in the form of additional equality constraints between 
-    input variables. Similarly, both the constraint, $C$ and the residual, $R$ are presented as equality constraints. They could be classified as 
-    a global and a local constraint on the problem respectively, with the only difference being the number of analysis tools that affect them.  
-    Hence, the distinction between local and global properties are not necessary to form a fundamental 
-    problem formulation. On the other hand, they are relevant when deriving a specific solution path for a problem, so any definition of problem 
-    formulation should include the capability to detect the weather a given property is local or global.
-
-    The challenge with using this traditional mathematical syntax is that it is not easily manipulated or analyzed. 
-    A number of matrix based methods have been used successfully to translate the mathematical syntax into a more useful computational form. 
-    Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools where off 
-    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Figure \ref{fig:dsm_simple} shows a DSM for the above notional example. 
-    The DSM has been successfully applied in a number of ways to algorithmically modify problem formulations to 
-    improve computational efficiency. Rogers et. al developed DeMAID to manipulate a
-    DSM  minimizing the computational costs of solving highly coupled systems\cite{Rogers1996}. Notice however, that the DSM in 
-    Figure \ref{fig:dsm_simple} contains less information than the full problem formulation. Objective and constraint information 
-    is missing and it's not possible to distinguish individual variables.
-
-    \begin{figure}[!hbp]
-        \begin{center}
-        \includegraphics[width=.75\textwidth]{images/dsm_simple}
-        \caption{Graph and matrix based DSM specifications for a notional analysis \label{fig:dsm_simple}}
-        \end{center}
-    \end{figure}
-
-    Since a DSM describes a square adjacency matrix, it can be represented in an equivalent directed graph where nodes represent analysis tools and 
-    edges represent information exchange between those tools. An alternate matrix based syntax, called a 
-    Functional Dependence Table (FDT), was proposed by Michelena and Papalambros. 
-    FDT represents the relationship between functions, including objectives and constraints, and their values\cite{Michelena1997}. Similar to DSM
-    FDT also describes an adjacency matrix of a graph. Unlike the DSM graph, however, a FDT graph is an undirected 
-    graph where nodes can represent analysis tools, objectives, or constraints. Edges between nodes represent a dependence on the same 
-    variable value. By searching the FDT graph for clusters of totally connected nodes Wagner and Papalambros were identify groups of 
-    analysis tools that were all dependent on the same input variables and used that to make partitioning decisions \cite{Wagner1993}. FDT retains 
-    a greater portion of the information in the problem formulation, but it is still not complete. It ignores information about feedback coupling 
-    between codes.
-
-    It is possible to combine the syntax of an FDT and a DSM by making a small extension to a traditional DSM specification. Normally, a DSM is given 
-    with analysis tools as nodes and variable dependencies given as edges. Lamb and Martins included the variables, objectives, and constraint functions
-    as nodes in an Extended DSM (XDSM)\cite{Lambe2012} in order to capture a more complete problem formulation for MDAO problems. With XDSM 
-    it is possible to specify a complete problem formulation. Lu and Martins developed a process to perform ordering and partitioning on the fundamental 
-    problem formulation specified by XDSM in order to achieve significant reductions in computational costs for large scale optimization problems. 
-    Figure \ref{fig:dsm_full} illustrates the same notional analysis as in Figure \ref{fig:dsm_simple} represented as an XDSM.
-
-    \begin{figure}[!hbp]
-        \begin{center}
-        \includegraphics[width=.75\textwidth]{images/dsm_simple}
-        \caption{[TODO: Generate XDSM]Extended DSM graph for a notional analysis \label{fig:dsm_full}}
-        \end{center}
-    \end{figure}
-
 \section{Graph Based Problem Formulation Syntax}
@@
     \end{itemize}
-
-\section{Example Problem}
-
-\section{Applications}
-
-\section{Conclusions}
-
-\bibliography{library}
-\end{document}

### 2012-10-02 problem_formulation.tex
breaking things up

--- 2012-10-01 problem_formulation.tex
@@
 
-\section{Graph Based Problem Formulation Using XDSM}
-    \subsection{Formulation Graph Syntax}
-    Rather than start with an adjacency, we chose to work directly with a directed cyclic graph to develop a syntax for the FPF. 
-    \begin{itemize}
-        \item Specify analyses
-        \item Connections between analyses 
-            \begin{itemize}
-                \item local variables
-                \item global variables? Use "fake" node that broadcasts out to the rest of the graph? 
-            \end{itemize}
-        \item Cycles indicate coupling
-        \item Cycles for design variables->objectives/constraints
-        \item Objectives/constraints are just outputs? Special nodes? 
-        \item Residuals are just outputs? Special nodes? 
-        \item Parameters are just input nodes that are not design variables (use identifies these)
-        \item FPF no solvers/optimizers anywhere in it
-    \end{itemize}
-
-    \subsection{Solution Graph Syntax}
-    What is the difference between a problem formulation and a problem solution method? Convert from a cyclic graph, to an acyclic graph
-    \begin{itemize}
-        \item cycles indicate convergence loops or design varaible loops
-        \item Problem can't be solved until all loops are *removed* by adding in solvers/optimizers
-        \item *special* nodes for solvers and optimizers that *break* loops (from an algorithmic point of view)
-        \item FPF represents the minimal amount of information necessary to define a problem
-        \item Any solution path grows the graph complexity by adding edges and nodes (or possibly have an empty solution graph, which you build up 
-        as you remove edges from problem formulation graph?)
-    \end{itemize}
+\input{graph_syntax}
 

### 2012-10-02 intro.tex
more refactoring

--- 2012-10-02 problem_formulation.tex
@@
-\documentclass[]{aiaa-tc} % insert '[draft]' option to show overfull boxes
-
- \title{An application of Graph Theory to MDAO problem formulation}
-
-\author{
-  David Pate, %
-     \thanks{Georgia Tech}
-  Dr. Brian German 
-     \thanks{Georgia Tech...}
-  Justin Gray,%
-     \thanks{Aerospace Engineer, MDAO Branch, Mail Stop 5-11, AIAA Member}   
- }
-
-%\usepackage{setspace}
-%\doublespace
-
-\usepackage{graphicx}
-\usepackage{wrapfig}
-\usepackage{caption} 
-\usepackage{amsmath}
-\usepackage{lscape}
-\usepackage{hyperref}
-\usepackage{appendix}
-\usepackage[section]{placeins}
-
-\captionsetup[figure]{margin=5pt,font=small,labelfont=bf,textfont=bf,justification=justified,}
-%\captionsetup[wrapfigure]{margin=5pt,font=small,labelfont=bf,justification=justified,singlelinecheck=off}
-\captionsetup[table]{margin=5pt,font=small,labelfont=bf,textfont=bf,justification=justified,position=top}
-
-\bibliographystyle{aiaa}
-
-\usepackage{lettrine}
-\usepackage{verbatim}
-
-%\usepackage{hyperref} %allows for the creation of actual text links
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-   blah blah blah
-\end{abstract}
-
-\section*{Nomenclature}
-
-\begin{tabular}{l l} 
-    AAO      & All-At- \\
-    MDAO     & Multidisciplinary Design Analysis and Optimization \\
-    FPF      & Fundamental Problem Formulation \\
-\end{tabular}
-
 \section{Introduction}
@@
 \section{Fundamental Problem Formulation}
-    Problem formulation is most commonly given in a mathematical syntax. For a simple, notional problem the problem formulation
-    can be given as follows:
+    Problem formulation is most commonly given in a mathematical syntax. For a simple, notional problem with two analyses
+    the problem formulation is as follows
 
@@
     \end{figure}
-
-\input{graph_syntax}
-
-\section{Example Problem}
-
-\section{Applications}
-
-\section{Conclusions}
-
-\bibliography{library}
-\end{document}

### 2012-10-02 problem_formulation.tex
more refactoring

@@
 
-\section{Introduction}
-
-    As the size and complexity of engineering systems grows the time and expense for setting up 
-    analysis models grows with them. Multidisciplinary Design Analysis and Optimization (MDAO)
-    frameworks such as OpenMDAO\cite{Gray2012} and ModelCenter have enabled a new level of analysis tool integration 
-    and paved the way for models of with more analysis tools and increasing numbers of multidisciplinary couplings. 
-    Such complex models present a distinct challenge proper implementation of any given solution strategy . 
-    In fact as the size of a problem grows very large even selecting an appropriate solution strategy can be a daunting 
-    task.
-
-    To help address in order to address the complexity problem we propose the use of graph based problem formulation 
-    in a way that allows for both analysis of potential solution strategies and automated implementation of those strategies 
-    inside an MDAO framework. In order to serve that purpose, the problem  needs to include the following information: 
-    \begin{itemize}
-       \item Discipline Analyses 
-       \item Discipline State Variables and Residuals
-       \item Design Variables
-       \item Constraints
-       \item Objective or Objectives
-       \item Coupling Constraints
-       \item Parameters
-    \end{itemize}
-    Notably absent from the preceding list are any kind of solvers, optimizers, or other iterative solution finding tools. 
-    At it most basic, a problem formulation includes only information about what is being sought after in a given problem, or what the
-    goals of a given problem are. It need not contain any information about solution paths or strategies to reach those goals. 
-    A general specification is one that states nothing specific about a problem solution path.
-
-    We define a complete and general problem specification striped down to it's most essential parts to be the 
-    Fundamental Problem Formulation (FPF). By definition, the FPF for any given problem will be constant regardless 
-    of which MDAO framework, optimization architecture, optimizer, or solver is used to solve the problem.
-
-    In this work we proposed a graph based syntax for the specification of problem formulation. This graph based syntax provides several key
-    features that make it useful for working with large scale MDAO problems. It provides a rigid structure that can be easily manipulated 
-    with a wide range of well establish graph-theory algorithms for the purposes of problem decomposition. The graph syntax also 
-    provides a means for algorithmically testing a given problem formulation to check if it is the FPF, and if not, to reduce it to the FPF. 
-    Lastly, a graph based specification for problem formulation lends itself well to interacting with MDAO frameworks which dramatically 
-    increases it's utility for real world applications.
-
-\section{Fundamental Problem Formulation}
-    Problem formulation is most commonly given in a mathematical syntax. For a simple, notional problem the problem formulation
-    can be given as follows:
-
-    \begin{align}
-        given & \ \ A.f\left(A.x,A.y\right) \notag
-        \\ & \ \  B.f\left(B.x,B.z\right) \notag
-        \\ & \ \  C.f1\left(A.f,B.f,C.s\right) \notag
-        \\ & \ \  C.f2\left(A.f,B.f,C.s\right) \notag
-        \\min &\ \ F\left(C.f1\right) \notag
-        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
-        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
-        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
-        \\      &\ \ E\left(A.y,C.f1\right) = 0 \notag
-        \\      &\ \ R\left(C.s,C.f2\right) = 0 
-        \label{eqn:simple}
-    \end{align}
-
-    Where $A,B,C$ represent analysis tools, $F,G1,G2,E,R$ are objective and constraint functions, and each analysis tool 
-    has a set of input and output variables. While this does represent a complete problem formulation, 
-    it's not the fundamental one since some assumptions have inherently been made about the solution 
-    strategy. $C$ is directly dependent
-    $A.f$ and $B.f$ as it's inputs. This implies that first you must run $A$ and $B$, then $C$ and
-    iterate on the design variables in $A$ and $B$ until you satisfy $E$. However, a different solution could be equally valid and still represent
-    the same fundamental problem:
-
-    \begin{align}
-        given & \ \ A.f\left(A.x,C.f1\right) \notag
-        \\ & \ \  B.f\left(B.x,B.z\right) \notag
-        \\ & \ \  C.f1\left(C.y,B.f,C.s\right) \notag
-        \\ & \ \  C.f2\left(C.q,B.f,C.s\right) \notag
-        \\min &\ \ F\left(C.f1\right) \notag
-        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
-        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
-        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
-        \\      &\ \ E\left(A.f,C.q\right) = 0 \notag
-        \\      &\ \ R\left(C.s,C.f2\right) = 0 
-        \label{eqn:simple_2}
-    \end{align}
-
-    Equation \ref{eqn:simple_2} differs only slightly from Eqn. \ref{eqn:simple}. $A$ is now directly dependent on the output of $C$, 
-    and the constraint $E$ has changed as well. Now the problem must be solved by running $B$ and $C$, then $A$ and iterating 
-    on the design variables in $B$ and $C$ until you satisfy $E$. Since the formulations in Eqn. \ref{eqn:simple} 
-    and \ref{eqn:simple_2} both describe the same problem there must be a more fundamental description of the problem that is common
-    between them. We present the FPF as follows:
-
-    \begin{align}
-        given & \ \ A.f\left(A.x,A.y\right) \notag
-        \\ & \ \  B.f\left(B.x,B.z\right) \notag
-        \\ & \ \  C.f1\left(C.y,C.r,C.s\right) \notag
-        \\ & \ \  C.f2\left(C.q,C.r,C.s\right) \notag
-        \\min &\ \ F\left(C.f1\right) \notag
-        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
-        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
-        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
-        \\      &\ \ E\left(A.f,C.\right) = 0 \notag
-        \\      &\ \ R\left(C.s,C.f2\right) = 0 
-        \label{eqn:simple_2}
-    \end{align}
-
-    \section{Problem Formulation Syntax}
-
-    The same mathematical syntax can be used to describe more complex problem formulations
-    such as ones that would decompose the FPF to a set of sub-problems. Tedford and Martins used this format to specify the 
-    FPF for a set of test problems and also to describe specific formulations for solving them with a 
-    number of optimization architectures\cite{Tedford2009}. Their work demonstrates clearly how multiple specific 
-    problem formulations can all relate back to a common FPF.
-
-    In the above problem formulation, there is no explicit relationship definition of local and global features. Instead they are 
-    implicitly defined through variable naming conventions. For instance, $x$ is a global variable because it is used in $A.x$ and $B.x$. 
-    Simliarly, $A.y$ is a local variable because it only affects the of $A$. If such a naming convention is not in place, 
-    then a more explicit definition of the relation between variables must also be given in the form of additional equality constraints between 
-    input variables. Similarly, both the constraint, $C$ and the residual, $R$ are presented as equality constraints. They could be classified as 
-    a global and a local constraint on the problem respectively, with the only difference being the number of analysis tools that affect them.  
-    Hence, the distinction between local and global properties are not necessary to form a fundamental 
-    problem formulation. On the other hand, they are relevant when deriving a specific solution path for a problem, so any definition of problem 
-    formulation should include the capability to detect the weather a given property is local or global.
-
-    The challenge with using this traditional mathematical syntax is that it is not easily manipulated or analyzed. 
-    A number of matrix based methods have been used successfully to translate the mathematical syntax into a more useful computational form. 
-    Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools where off 
-    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Figure \ref{fig:dsm_simple} shows a DSM for the above notional example. 
-    The DSM has been successfully applied in a number of ways to algorithmically modify problem formulations to 
-    improve computational efficiency. Rogers et. al developed DeMAID to manipulate a
-    DSM  minimizing the computational costs of solving highly coupled systems\cite{Rogers1996}. Notice however, that the DSM in 
-    Figure \ref{fig:dsm_simple} contains less information than the full problem formulation. Objective and constraint information 
-    is missing and it's not possible to distinguish individual variables.
-
-    \begin{figure}[!hbp]
-        \begin{center}
-        \includegraphics[width=.75\textwidth]{images/dsm_simple}
-        \caption{Graph and matrix based DSM specifications for a notional analysis \label{fig:dsm_simple}}
-        \end{center}
-    \end{figure}
-
-    Since a DSM describes a square adjacency matrix, it can be represented in an equivalent directed graph where nodes represent analysis tools and 
-    edges represent information exchange between those tools. An alternate matrix based syntax, called a 
-    Functional Dependence Table (FDT), was proposed by Michelena and Papalambros. 
-    FDT represents the relationship between functions, including objectives and constraints, and their values\cite{Michelena1997}. Similar to DSM
-    FDT also describes an adjacency matrix of a graph. Unlike the DSM graph, however, a FDT graph is an undirected 
-    graph where nodes can represent analysis tools, objectives, or constraints. Edges between nodes represent a dependence on the same 
-    variable value. By searching the FDT graph for clusters of totally connected nodes Wagner and Papalambros were identify groups of 
-    analysis tools that were all dependent on the same input variables and used that to make partitioning decisions \cite{Wagner1993}. FDT retains 
-    a greater portion of the information in the problem formulation, but it is still not complete. It ignores information about feedback coupling 
-    between codes.
-
-    It is possible to combine the syntax of an FDT and a DSM by making a small extension to a traditional DSM specification. Normally, a DSM is given 
-    with analysis tools as nodes and variable dependencies given as edges. Lamb and Martins included the variables, objectives, and constraint functions
-    as nodes in an Extended DSM (XDSM)\cite{Lambe2012} in order to capture a more complete problem formulation for MDAO problems. With XDSM 
-    it is possible to specify a complete problem formulation. Lu and Martins developed a process to perform ordering and partitioning on the fundamental 
-    problem formulation specified by XDSM in order to achieve significant reductions in computational costs for large scale optimization problems. 
-    Figure \ref{fig:dsm_full} illustrates the same notional analysis as in Figure \ref{fig:dsm_simple} represented as an XDSM.
-
-    \begin{figure}[!hbp]
-        \begin{center}
-        \includegraphics[width=.75\textwidth]{images/dsm_simple}
-        \caption{[TODO: Generate XDSM]Extended DSM graph for a notional analysis \label{fig:dsm_full}}
-        \end{center}
-    \end{figure}
-
+\input{intro}
 \input{graph_syntax}

### 2012-10-02 intro.tex
updated the simple problem to use better icons

@@
 \section{Fundamental Problem Formulation}
-    Problem formulation is most commonly given in a mathematical syntax. For a simple, notional problem with two analyses
-    the problem formulation is as follows
+    Problem formulation is most commonly given in a mathematical syntax. For a simple, notional problem 
+    the problem formulation is as follows:
 
     \begin{align}
-        given & \ \ A.f\left(A.x,A.y\right) \notag
-        \\ & \ \  B.f\left(B.x,B.z\right) \notag
-        \\ & \ \  C.f1\left(A.f,B.f,C.s\right) \notag
-        \\ & \ \  C.f2\left(A.f,B.f,C.s\right) \notag
-        \\min &\ \ F\left(C.f1\right) \notag
-        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
-        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
-        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
-        \\      &\ \ E\left(A.y,C.f1\right) = 0 \notag
-        \\      &\ \ R\left(C.s,C.f2\right) = 0 
+        given & \ \ A: {x,y} \rightarrow {z}
+              & \ \ B: {x,z} \rightarrow {y^t}
+        min. &\ \ f(m,t) \notag
+        w.r.t. & \ \ x,y,z \notag
+        s.t. & \ \ G(y^t,y) = 0
         \label{eqn:simple}
@@
 
-    Where $A,B,C$ represent analysis tools, $F,G1,G2,E,R$ are objective and constraint functions, and each analysis tool 
-    has a set of input and output variables. While this does represent a complete problem formulation, 
+    Where $A$ and $B$ represent analysis tools, $F,G$ are the objective and constraint functions respectively. 
+    While this does represent a complete problem formulation, 
     it's not the fundamental one since some assumptions have inherently been made about the solution 
-    strategy. $C$ is directly dependent
-    $A.f$ and $B.f$ as it's inputs. This implies that first you must run $A$ and $B$, then $C$ and
-    iterate on the design variables in $A$ and $B$ until you satisfy $E$. However, a different solution could be equally valid and still represent
+    strategy. Analysis $A$ outputs $z$ which is an input to $B$. Hence, $A$ should be run before $B$ with 
+    the $y$ being iterated on to convergence. However, a different solution could be equally valid and still represent
     the same fundamental problem:
@@
     \begin{align}
-        given & \ \ A.f\left(A.x,C.f1\right) \notag
-        \\ & \ \  B.f\left(B.x,B.z\right) \notag
-        \\ & \ \  C.f1\left(C.y,B.f,C.s\right) \notag
-        \\ & \ \  C.f2\left(C.q,B.f,C.s\right) \notag
-        \\min &\ \ F\left(C.f1\right) \notag
-        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
-        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
-        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
-        \\      &\ \ E\left(A.f,C.q\right) = 0 \notag
-        \\      &\ \ R\left(C.s,C.f2\right) = 0 
-        \label{eqn:simple_2}
+        given & \ \ B: {x,z} \rightarrow {y}
+              & \ \ A: {x,y} \rightarrow {z^t}
+        min. &\ \ f(m,t) \notag
+        w.r.t. & \ \ x,y,z \notag
+        s.t. & \ \ G(z^t,z) = 0
+        \label{eqn:simple2}
     \end{align}
 
-    Equation \ref{eqn:simple_2} differs only slightly from Eqn. \ref{eqn:simple}. $A$ is now directly dependent on the output of $C$, 
-    and the constraint $E$ has changed as well. Now the problem must be solved by running $B$ and $C$, then $A$ and iterating 
-    on the design variables in $B$ and $C$ until you satisfy $E$. Since the formulations in Eqn. \ref{eqn:simple} 
-    and \ref{eqn:simple_2} both describe the same problem there must be a more fundamental description of the problem that is common
-    between them. We present the FPF as follows:
+    Equation \ref{eqn:simple2} differs only slightly from Eqn. \ref{eqn:simple}. $A$ is now dependent on the output of $B$, 
+    and $z$ will be iterated on to meet $G$. Now the problem can be solved by running $B$ first and then $A$.
+    Since the formulations in Eqn. \ref{eqn:simple} and \ref{eqn:simple_2} both describe the same problem 
+    there must be a more fundamental description of the problem that is common between them. We present the FPF as follows:
 
     \begin{align}
-        given & \ \ A.f\left(A.x,A.y\right) \notag
-        \\ & \ \  B.f\left(B.x,B.z\right) \notag
-        \\ & \ \  C.f1\left(C.y,C.r,C.s\right) \notag
-        \\ & \ \  C.f2\left(C.q,C.r,C.s\right) \notag
-        \\min &\ \ F\left(C.f1\right) \notag
-        \\ w.r.t. &\ \  A.x,A.y,B.x,B.z \notag
-        \\ s.t. &\ \ G1\left(A.x,B.z\right) \leq 0 \notag
-        \\      &\ \ G2\left(A.y\right) \leq 0 \notag
-        \\      &\ \ E\left(A.f,C.\right) = 0 \notag
-        \\      &\ \ R\left(C.s,C.f2\right) = 0 
-        \label{eqn:simple_2}
+        given & \ \ A: {x,y} \rightarrow {z^t}
+              & \ \ B: {x,z} \rightarrow {y^t}
+        min. &\ \ f(m,t) \notag
+        w.r.t. & \ \ x,y,z \notag
+        s.t. & \ \ G1(z^t,z) = 0 \notag
+             & \ \ G2(y^t,y) = 0
+        \label{eqn:simple_fpf}
     \end{align}
+
+    The FPF in Eqn. \ref{eqn:simple_fpf} differs from both Eqn. \ref{eqn:simple} and Eqn. \ref{eqn:simple2} because it has 
+    two constraints which both must be met. The presence of both of these constraints fully decouples the problem so that 
+    either $A$ or $B$ could be run first or both could be run simultaneously. By removing either constraint, and replacing 
+    it with a direct dependence between the two analyses you could regain the earlier two problem formulations.
 

### 2012-10-02 intro.tex
fixed eqn's
updating the simple DSM

@@
 
-\section{Fundamental Problem Formulation}
+\subsection{Fundamental Problem Formulation}
     Problem formulation is most commonly given in a mathematical syntax. For a simple, notional problem 
@@
         given & \ \ A: {x,y} \rightarrow {z}
-              & \ \ B: {x,z} \rightarrow {y^t}
-        min. &\ \ f(m,t) \notag
-        w.r.t. & \ \ x,y,z \notag
-        s.t. & \ \ G(y^t,y) = 0
+        \\      & \ \ B: {x,z} \rightarrow {y^t}
+        \\min. &\ \ f(m,t) \notag
+        \\w.r.t. & \ \ x,y,z \notag
+        \\s.t. & \ \ G(y^t,y) = 0
         \label{eqn:simple}
@@
         given & \ \ B: {x,z} \rightarrow {y}
-              & \ \ A: {x,y} \rightarrow {z^t}
-        min. &\ \ f(m,t) \notag
-        w.r.t. & \ \ x,y,z \notag
-        s.t. & \ \ G(z^t,z) = 0
+        \\      & \ \ A: {x,y} \rightarrow {z^t}
+        \\min. &\ \ f(m,t) \notag
+        \\w.r.t. & \ \ x,y,z \notag
+        \\s.t. & \ \ G(z^t,z) = 0
         \label{eqn:simple2}
@@
         given & \ \ A: {x,y} \rightarrow {z^t}
-              & \ \ B: {x,z} \rightarrow {y^t}
-        min. &\ \ f(m,t) \notag
-        w.r.t. & \ \ x,y,z \notag
-        s.t. & \ \ G1(z^t,z) = 0 \notag
-             & \ \ G2(y^t,y) = 0
+        \\      & \ \ B: {x,z} \rightarrow {y^t}
+        \\min. &\ \ f(m,t) \notag
+        \\w.r.t. & \ \ x,y,z \notag
+        \\s.t. & \ \ G1(z^t,z) = 0 \notag
+        \\     & \ \ G2(y^t,y) = 0
         \label{eqn:simple_fpf}
@@
 
-    \section{Problem Formulation Syntax}
+    \subsection{Problem Formulation Syntax}
 
-    The same mathematical syntax can be used to describe more complex problem formulations
-    such as ones that would decompose the FPF to a set of sub-problems. Tedford and Martins used this format to specify the 
+    As shown above, the mathematical language for specifying problem formulations is very general and can be used both for 
+    fundamental and specific problem formulations. Tedford and Martins used the above mathematical syntax to specify the 
     FPF for a set of test problems and also to describe specific formulations for solving them with a 
@@
 
+    \begin{comment}
     In the above problem formulation, there is no explicit relationship definition of local and global features. Instead they are 
@@
     formulation should include the capability to detect the weather a given property is local or global. 
+    \end{comment}
 
@@
     improve computational efficiency. Rogers et. al developed DeMAID to manipulate a
-    DSM  minimizing the computational costs of solving highly coupled systems\cite{Rogers1996}. Notice however, that the DSM in 
-    Figure \ref{fig:dsm_simple} contains less information than the full problem formulation. Objective and constraint information 
+    DSM  to find an ordering for analysis tools that reduces the cost of solving highly coupled systems\cite{Rogers1996}. This re-ordering yields
+    multiple specific problem formulations which all solve the same FPF. In other words, manipulation of the DSM does not fundamentally alter
+    the problem formulation, which makes DSM an excellent foundation for specifying the FPF itself. Notice however, that the DSM in 
+    Figure \ref{fig:dsm_simple} contains less information than the Eqn. \ref{eqn:simple_fpf}. Objective and constraint information 
     is missing and it's not possible to distinguish individual variables.
@@
         \includegraphics[width=.75\textwidth]{images/dsm_simple}
-        \caption{Graph and matrix based DSM specifications for a notional analysis \label{fig:dsm_simple}}
+        \caption{Graph and matrix based DSM specifications for Eqn. \ref{eqn:simple_FPF} \label{fig:dsm_simple}}
         \end{center}

### 2012-10-02 XDSM/diagram_styles.tex
starting xdsm for simple

% Define all the styles used to produce XDSMs for MDO

% Component types
\tikzstyle{Optimization} = [rounded rectangle,draw,fill=blue!20,inner sep=6pt,minimum height=1cm,text badly centered]
\tikzstyle{LP_Optimization} = [rectangle,draw,fill=blue!20,inner sep=6pt,minimum height=1cm,text badly centered]
\tikzstyle{Analysis} = [rectangle,draw,fill=green!20,inner sep=6pt,minimum height=1cm,text badly centered]
\tikzstyle{Function} = [rectangle,draw,fill=purple!20,inner sep=6pt,minimum height=1cm,text badly centered]
\tikzstyle{MDA} = [rounded rectangle,draw,fill=orange!20,inner sep=6pt,minimum height=1cm,text badly centered]
\tikzstyle{Metamodel} = [rectangle,draw,fill=yellow!20,inner sep=6pt,minimum height=1cm,text badly centered]
\tikzstyle{DOE} = [rounded rectangle,draw,fill=yellow!20,inner sep=6pt,minimum height=1cm,text badly centered]
%\tikzstyle{OptFunction} = [rectangle,draw,fill=red!20,inner sep=6pt,minimum height=1cm,text badly centered]

%% A simple command to give the repeated structure look for components and data
\tikzstyle{stack} = [double copy shadow]

%% A simple command to fade components and data, e.g. demonstrating a sequence of steps in an animation
\tikzstyle{faded} = [draw=black!50,fill=white,text opacity=0.5]

%% Simple fading commands for the lines
\tikzstyle{fadeddata} = [color=black!20]
\tikzstyle{fadedprocess} = [color=black!50]

% Component types for repeated structures (i.e. for parallel structures)
\tikzstyle{Optimization_i} = [double copy shadow, Optimization]
\tikzstyle{LP_Optimization_i} = [double copy shadow, LP_Optimization]
\tikzstyle{Analysis_i} = [double copy shadow, Analysis]
\tikzstyle{Function_i} = [double copy shadow, Function]
\tikzstyle{MDA_i} = [double copy shadow, MDA]
\tikzstyle{Metamodel_i} = [double copy shadow, Metamodel]
\tikzstyle{DOE_i} = [double copy shadow, DOE]

% Faded component types for, e.g. demonstrations of each step. We use these style definitions to "gray out" large parts of the diagram.
\tikzstyle{Optimization_fade} = [Optimization,fill=blue!10,draw=black!30,text opacity=0.3]
\tikzstyle{Analysis_fade} = [Analysis,fill=green!10,draw=black!30,text opacity=0.3]
\tikzstyle{Function_fade} = [Function,fill=purple!10,draw=black!30,text opacity=0.3]
\tikzstyle{MDA_fade} = [MDA,fill=orange!10,draw=black!30,text opacity=0.3]
\tikzstyle{Metamodel_fade} = [Metamodel,fill=yellow!10,draw=black!30,text opacity=0.3]
\tikzstyle{DOE_fade} = [DOE,fill=yellow!10,draw=black!30,text opacity=0.3]

\tikzstyle{Optimization_i_fade} = [Optimization_i,fill=blue!10,draw=black!30,text opacity=0.3]
\tikzstyle{Analysis_i_fade} = [Analysis_i,fill=green!10,draw=black!30,text opacity=0.3]
\tikzstyle{Function_i_fade} = [Function_i,fill=purple!10,draw=black!30,text opacity=0.3]
\tikzstyle{MDA_i_fade} = [MDA_i,fill=orange!10,draw=black!30,text opacity=0.3]
\tikzstyle{Metamodel_i_fade} = [Metamodel_i,fill=yellow!10,draw=black!30,text opacity=0.3]
\tikzstyle{DOE_i_fade} = [DOE_i,fill=yellow!10,draw=black!30,text opacity=0.3]

% Data types
\tikzstyle{DataInter} = [trapezium,trapezium left angle=75,trapezium right angle=105,draw,fill=black!10]
\tikzstyle{DataIO} = [trapezium,trapezium left angle=75,trapezium right angle=105,draw,fill=white]

% Data types for repeated structures
\tikzstyle{DataInter_i} = [double copy shadow, DataInter]
\tikzstyle{DataIO_i} = [double copy shadow, DataIO]

% Faded data types
\tikzstyle{DataInter_fade} = [DataInter,draw=black!30,fill=white,text opacity=0.3]
\tikzstyle{DataIO_fade} = [DataIO_i,draw=black!30,fill=white,text opacity=0.3]

\tikzstyle{DataInter_i_fade} = [DataInter_i,draw=black!30,fill=white,text opacity=0.3]
\tikzstyle{DataIO_i_fade} = [DataIO_i,draw=black!30,fill=white,text opacity=0.3]

% Edges
\tikzstyle{DataLine} = [color=black!40,line width=5pt]
\tikzstyle{ProcessHV} = [-,line width=1pt,to path={-| (\tikztotarget)}]
\tikzstyle{ProcessTip} = [-,line width=1pt]

% Faded edges
\tikzstyle{DataLine_fade} = [DataLine,color=black!10]
\tikzstyle{ProcessHV_fade} = [ProcessHV,color=black!30]
\tikzstyle{ProcessTip_fade} = [ProcessTip,color=black!30]

% Matrix options
\tikzstyle{MatrixSetup} = [row sep=3mm, column sep=2mm]

% Declare a background layer for showing node connections
\pgfdeclarelayer{data}
\pgfdeclarelayer{process}
\pgfsetlayers{data,process,main}

% A new command to split the component text over multiple lines
\newcommand{\MultilineComponent}[3]
{
	\begin{minipage}{#1}
	\begin{center}
		#2 \linebreak #3
	\end{center}
	\end{minipage}
}

% A new command to split the component text over multiple columns
\newcommand{\MultiColumnComponent}[5]
{
	\begin{minipage}{#1}
	\begin{center}
	#2 \linebreak #3
	\end{center}
	\begin{minipage}{0.49\textwidth}
	\begin{center}
	#4
	\end{center}
	\end{minipage}
	\begin{minipage}{0.49\textwidth}
	\begin{center}
		#5
	\end{center}
	\end{minipage}
	\end{minipage}
}

### 2012-10-02 intro.tex
starting xdsm for simple

@@
         \\w.r.t. & \ \ x,y,z \notag
-        \\s.t. & \ \ G(y^t,y) = 0
+        \\s.t. & \ \ g(y^t,y) = 0
         \label{eqn:simple}
@@
 
-    Where $A$ and $B$ represent analysis tools, $F,G$ are the objective and constraint functions respectively. 
+    Where $A$ and $B$ represent analysis tools, $f,g$ are the objective and constraint functions respectively. 
     While this does represent a complete problem formulation, 
@@
         \\w.r.t. & \ \ x,y,z \notag
-        \\s.t. & \ \ G(z^t,z) = 0
+        \\s.t. & \ \ g(z^t,z) = 0
         \label{eqn:simple2}
@@
     Equation \ref{eqn:simple2} differs only slightly from Eqn. \ref{eqn:simple}. $A$ is now dependent on the output of $B$, 
-    and $z$ will be iterated on to meet $G$. Now the problem can be solved by running $B$ first and then $A$.
-    Since the formulations in Eqn. \ref{eqn:simple} and \ref{eqn:simple_2} both describe the same problem 
+    and $z$ will be iterated on to meet $g$. Now the problem can be solved by running $B$ first and then $A$.
+    Since the formulations in Eqn. \ref{eqn:simple} and \ref{eqn:simple2} both describe the same problem 
     there must be a more fundamental description of the problem that is common between them. We present the FPF as follows:
@@
         \\w.r.t. & \ \ x,y,z \notag
-        \\s.t. & \ \ G1(z^t,z) = 0 \notag
-        \\     & \ \ G2(y^t,y) = 0
+        \\s.t. & \ \ g1(z^t,z) = 0 \notag
+        \\     & \ \ g2(y^t,y) = 0
         \label{eqn:simple_fpf}
@@
     Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools where off 
-    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Figure \ref{fig:dsm_simple} shows a DSM for the above notional example. 
-    The DSM has been successfully applied in a number of ways to algorithmically modify problem formulations to 
-    improve computational efficiency. Rogers et. al developed DeMAID to manipulate a
-    DSM  to find an ordering for analysis tools that reduces the cost of solving highly coupled systems\cite{Rogers1996}. This re-ordering yields
-    multiple specific problem formulations which all solve the same FPF. In other words, manipulation of the DSM does not fundamentally alter
+    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Figure \ref{fig:dsm_simple} shows a DSM for the above notional example 
+    with and without the coupling cycle identified. In order to derive the formulations in Eqn. \ref{eqn:simple} and Eq. \ref{eqn:simple2} 
+    you would remove one of the output target variables ($y^t$ or $z^t$) and replace it with a direct connection which would select either $A$
+    or $B$ to be executed first. For more complex problems, choosing the proper order to run analysis tools is a non-trival task. 
+    Rogers et. al developed DeMAID to manipulate a DSM to find an ordering for analysis tools that 
+    reduces the cost of solving highly coupled systems\cite{Rogers1996}. This re-ordering yields multiple specific problem 
+    formulations which all solve the same FPF. In other words, manipulation of the DSM does not fundamentally alter
     the problem formulation, which makes DSM an excellent foundation for specifying the FPF itself. Notice however, that the DSM in 
-    Figure \ref{fig:dsm_simple} contains less information than the Eqn. \ref{eqn:simple_fpf}. Objective and constraint information 
-    is missing and it's not possible to distinguish individual variables.
+    Figure \ref{fig:dsm_simple} contains less information than the Eqn. \ref{eqn:simple_fpf}. While it does fully capture 
+    the coupling information, objective and constraint information is missing.
 
@@
         \begin{center}
-        \includegraphics[width=.75\textwidth]{images/dsm_simple}
-        \caption{Graph and matrix based DSM specifications for Eqn. \ref{eqn:simple_FPF} \label{fig:dsm_simple}}
+        \includegraphics[height=.25\textheight]{images/dsm_simple}
+        \caption{DSM specifications for Eqn. \ref{eqn:simple_fpf}. Coupling cycle is identified by dashed lines. \label{fig:dsm_simple}}
         \end{center}
@@
     graph where nodes can represent analysis tools, objectives, or constraints. Edges between nodes represent a dependence on the same 
-    variable value. By searching the FDT graph for clusters of totally connected nodes Wagner and Papalambros were identify groups of 
+    variable value. By searching the FDT graph for clusters of totally connected nodes Wagner and Papalambros identified groups of 
     analysis tools that were all dependent on the same input variables and used that to make partitioning decisions \cite{Wagner1993}. FDT retains 
-    a greater portion of the information in the problem formulation, but it is still not complete. It ignores information about feedback coupling 
-    between codes.
+    a greater portion of the information in the problem formulation, but it is still not complete. It ignores information about coupling 
+    between analyses.
 
@@
     with analysis tools as nodes and variable dependencies given as edges. Lamb and Martins included the variables, objectives, and constraint functions
-    as nodes in an Extended DSM (XDSM)\cite{Lambe2012} in order to capture a more complete problem formulation for MDAO problems. With XDSM 
-    it is possible to specify a complete problem formulation. Lu and Martins developed a process to perform ordering and partitioning on the fundamental 
-    problem formulation specified by XDSM in order to achieve significant reductions in computational costs for large scale optimization problems. 
-    Figure \ref{fig:dsm_full} illustrates the same notional analysis as in Figure \ref{fig:dsm_simple} represented as an XDSM.
+    as nodes in an Extended DSM (XDSM)\cite{Lambe2012} in order to capture a more complete problem formulation for MDAO problems. Lu made use of 
+    XDSM to perform ordering and partitioning on the fundamental problem formulation specified by XDSM in order to achieve 
+    significant reductions in computational costs for large scale optimization problems. Figure \ref{fig:dsm_full} illustrates 
+    the same notional analysis as in Figure \ref{fig:dsm_simple} represented as an XDSM.
 
@@
         \includegraphics[width=.75\textwidth]{images/dsm_simple}
-        \caption{[TODO: Generate XDSM]Extended DSM graph for a notional analysis \label{fig:dsm_full}}
+        \caption{[TODO: Generate XDSM]Extended DSM graph for Eqn. \ref{eqn:simple_fpf} \label{fig:dsm_full}}
         \end{center}

### 2012-10-02 intro.tex
fixing simple problem formulation
fixing some eqn numbers
tweaking the wording

@@
     Such complex models present a distinct challenge proper implementation of any given solution strategy .
-    In fact as the size of a problem grows very large even selecting an appropriate solution strategy can be a daunting 
-    task.
 
-    To help address in order to address the complexity problem we propose the use of graph based problem formulation 
-    in a way that allows for both analysis of potential solution strategies and automated implementation of those strategies 
-    inside an MDAO framework. In order to serve that purpose, the problem  needs to include the following information: 
+    We propose the use of graph based problem formulation that generally describes a
+    given problem, supports algorithmic manipulation to find beneficial solutions strategies, and enables automated 
+    implementation of those strategies inside an MDAO framework. In order to serve that purpose, the graph includes the following information: 
     \begin{itemize}
@@
     goals of a given problem are. It need not contain any information about solution paths or strategies to reach those goals.
-    A general specification is one that states nothing specific about a problem solution path.
 
-    We define a complete and general problem specification striped down to it's most essential parts to be the 
+    We define a problem specification striped down to it's most essential parts to be the 
     Fundamental Problem Formulation (FPF). By definition, the FPF for any given problem will be constant regardless 
-    of which MDAO framework, optimization architecture, optimizer, or solver is used to solve the problem.
+    of which MDAO framework, optimization architecture, optimization algorithm, or iterative solver is used to solve the problem.
 
-    In this work we proposed a graph based syntax for the specification of problem formulation. This graph based syntax provides several key
-    features that make it useful for working with large scale MDAO problems. It provides a rigid structure that can be easily manipulated 
-    with a wide range of well establish graph-theory algorithms for the purposes of problem decomposition. The graph syntax also 
-    provides a means for algorithmically testing a given problem formulation to check if it is the FPF, and if not, to reduce it to the FPF. 
-    Lastly, a graph based specification for problem formulation lends itself well to interacting with MDAO frameworks which dramatically 
-    increases it's utility for real world applications.
+    The graph based syntax used in this work is based on the Design Structure Matrix (DSM). 
+    This provides several key features that make it useful for working with large scale MDAO problems. 
+    It provides a rigid structure that can be easily manipulated with a wide range of well establish graph-theory algorithms 
+    for the purposes of problem decomposition. It also provides a means for testing a given 
+    problem formulation to check if it is the FPF, and if not to reduce it to the FPF. Lastly, it lends itself 
+    well to interacting with MDAO frameworks.
 
-\subsection{Fundamental Problem Formulation}
-    Problem formulation is most commonly given in a mathematical syntax. For a simple, notional problem 
-    the problem formulation is as follows:
+\section{Specific vs Fundamental Problem Formulation }
+    For a simple, notional problem  is given as follows:
 
     \begin{align}
-        given & \ \ A: {x,y} \rightarrow {z}
-        \\      & \ \ B: {x,z} \rightarrow {y^t}
-        \\min. &\ \ f(m,t) \notag
+        given & \ \ A: {x,y} \rightarrow {m,z} \notag
+        \\      & \ \ B: {x,z} \rightarrow {y^t} \notag
+        \\min. &\ \ f(m) \notag
         \\w.r.t. & \ \ x,y,z \notag
@@
     Where $A$ and $B$ represent analysis tools, $f,g$ are the objective and constraint functions respectively. 
-    While this does represent a complete problem formulation, 
-    it's not the fundamental one since some assumptions have inherently been made about the solution 
-    strategy. Analysis $A$ outputs $z$ which is an input to $B$. Hence, $A$ should be run before $B$ with 
-    the $y$ being iterated on to convergence. However, a different solution could be equally valid and still represent
+    Equation \ref{eqn:simple} makes an inherent assumption about the solution strategy for the problem. 
+    Analysis $A$ outputs $z$ which is an input to $B$. Hence, $A$ should be run before $B$ with 
+    the $y$ being iterated on to convergence with $y^t$. However, a different solution could be equally valid and still represent
     the same fundamental problem:
@@
     \begin{align}
-        given & \ \ B: {x,z} \rightarrow {y}
-        \\      & \ \ A: {x,y} \rightarrow {z^t}
-        \\min. &\ \ f(m,t) \notag
+        given & \ \ B: {x,z} \rightarrow {y} \notag
+        \\      & \ \ A: {x,y} \rightarrow {m,z^t} \notag
+        \\min. &\ \ f(m) \notag
         \\w.r.t. & \ \ x,y,z \notag
@@
     Equation \ref{eqn:simple2} differs only slightly from Eqn. \ref{eqn:simple}. $A$ is now dependent on the output of $B$, 
-    and $z$ will be iterated on to meet $g$. Now the problem can be solved by running $B$ first and then $A$.
-    Since the formulations in Eqn. \ref{eqn:simple} and \ref{eqn:simple2} both describe the same problem 
-    there must be a more fundamental description of the problem that is common between them. We present the FPF as follows:
+    and $z$ will be iterated on to satisfy $g$. Now the problem can be solved by running $B$ first and then $A$.
+    Since the formulations in Eqn. \ref{eqn:simple} and \ref{eqn:simple2} both describe the same problem and neither can be the FPF. 
+    There must be a more fundamental description of the problem that is common between them. We present the FPF as follows:
 
     \begin{align}
-        given & \ \ A: {x,y} \rightarrow {z^t}
-        \\      & \ \ B: {x,z} \rightarrow {y^t}
-        \\min. &\ \ f(m,t) \notag
+        given & \ \ A: {x,y} \rightarrow {m,z^t} \notag
+        \\      & \ \ B: {x,z} \rightarrow {y^t} \notag
+        \\min. &\ \ f(m) \notag
         \\w.r.t. & \ \ x,y,z \notag
@@
 
-    \subsection{Problem Formulation Syntax}
+\section{Using DSM for Problem Formulation}
 
@@
 
-    \begin{comment}
-    In the above problem formulation, there is no explicit relationship definition of local and global features. Instead they are 
-    implicitly defined through variable naming conventions. For instance, $x$ is a global variable because it is used in $A.x$ and $B.x$. 
-    Simliarly, $A.y$ is a local variable because it only affects the of $A$. If such a naming convention is not in place, 
-    then a more explicit definition of the relation between variables must also be given in the form of additional equality constraints between 
-    input variables. Similarly, both the constraint, $C$ and the residual, $R$ are presented as equality constraints. They could be classified as 
-    a global and a local constraint on the problem respectively, with the only difference being the number of analysis tools that affect them.  
-    Hence, the distinction between local and global properties are not necessary to form a fundamental 
-    problem formulation. On the other hand, they are relevant when deriving a specific solution path for a problem, so any definition of problem 
-    formulation should include the capability to detect the weather a given property is local or global. 
-    \end{comment}
-
     The challenge with using this traditional mathematical syntax is that it is not easily manipulated or analyzed. 
@@
     Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools where off 
-    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Figure \ref{fig:dsm_simple} shows a DSM for the above notional example 
-    with and without the coupling cycle identified. In order to derive the formulations in Eqn. \ref{eqn:simple} and Eq. \ref{eqn:simple2} 
-    you would remove one of the output target variables ($y^t$ or $z^t$) and replace it with a direct connection which would select either $A$
-    or $B$ to be executed first. For more complex problems, choosing the proper order to run analysis tools is a non-trival task. 
+    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Since a DSM describes a square adjacency matrix, 
+    it can be represented in an equivalent directed graph where nodes represent analysis tools and 
+    edges represent information dependence between those tools. The ordering of elements in a DSM can be used to indicate 
+    execution order.  For more complex problems, choosing the proper order to run analysis tools is a non-trival task. 
     Rogers et. al developed DeMAID to manipulate a DSM to find an ordering for analysis tools that 
@@
     formulations which all solve the same FPF. In other words, manipulation of the DSM does not fundamentally alter
-    the problem formulation, which makes DSM an excellent foundation for specifying the FPF itself. Notice however, that the DSM in 
-    Figure \ref{fig:dsm_simple} contains less information than the Eqn. \ref{eqn:simple_fpf}. While it does fully capture 
-    the coupling information, objective and constraint information is missing.
+    the problem formulation, which makes DSM an excellent foundation for specifying the FPF itself. Traditional DSM only captures information 
+    about data dependency between analyses. Objective and constraint information is missing from the description of the problem.
 
-    \begin{figure}[!hbp]
-        \begin{center}
-        \includegraphics[height=.25\textheight]{images/dsm_simple}
-        \caption{DSM specifications for Eqn. \ref{eqn:simple_fpf}. Coupling cycle is identified by dashed lines. \label{fig:dsm_simple}}
-        \end{center}
-    \end{figure}
+    An alternate matrix based syntax, called a Functional Dependency Table (FDT), was proposed by Michelena and Papalambros. 
+    FDT represents the relationship between functions, including objectives and constraints, and specific variables that affect 
+    them\cite{Michelena1997}. Similar to DSM, FDT also describes an adjacency matrix of a graph. Unlike the DSM graph, 
+    however, the graph is undirected and nodes can represent analysis tools, objectives, 
+    or constraints. Edges between nodes represent a dependence on the same 
+    variable. Wagner and Papalambros made use of the FDT to solve a graph partitioning problem that yielded 
+    more efficient optimization problem decompositions\cite{Wagner1993}. While FDT succeeds at capturing the 
+    information about objectives and constraints, it neglects the data dependency captured by DSM. 
+    [TODO: There is more recent research to cite here, like Allison's Work.]
 
-    Since a DSM describes a square adjacency matrix, it can be represented in an equivalent directed graph where nodes represent analysis tools and 
-    edges represent information exchange between those tools. An alternate matrix based syntax, called a 
-    Functional Dependence Table (FDT), was proposed by Michelena and Papalambros. 
-    FDT represents the relationship between functions, including objectives and constraints, and their values\cite{Michelena1997}. Similar to DSM
-    FDT also describes an adjacency matrix of a graph. Unlike the DSM graph, however, a FDT graph is an undirected 
-    graph where nodes can represent analysis tools, objectives, or constraints. Edges between nodes represent a dependence on the same 
-    variable value. By searching the FDT graph for clusters of totally connected nodes Wagner and Papalambros identified groups of 
-    analysis tools that were all dependent on the same input variables and used that to make partitioning decisions \cite{Wagner1993}. FDT retains 
-    a greater portion of the information in the problem formulation, but it is still not complete. It ignores information about coupling 
-    between analyses.
+    Lamb and Martins included the variables, objectives, and constraint functions as nodes in an Extended 
+    DSM (XDSM) in order to capture a more complete problem formulation for MDAO problems\cite{Lambe2012}. By adding in those 
+    elements, they partially combined a traditional DSM with an FDT. 
+    This allows XDSM to represent data dependency between multiple analysis tools as well as between analysis tools and
+    objective/constraint functions. With the additional information included in an XDSM Lu and Martins applied both ordering and partitioning 
+    algorithms on an MDAO test problem named the Scalable Problem \cite{Lu2012}.
 
-    It is possible to combine the syntax of an FDT and a DSM by making a small extension to a traditional DSM specification. Normally, a DSM is given 
-    with analysis tools as nodes and variable dependencies given as edges. Lamb and Martins included the variables, objectives, and constraint functions
-    as nodes in an Extended DSM (XDSM)\cite{Lambe2012} in order to capture a more complete problem formulation for MDAO problems. Lu made use of 
-    XDSM to perform ordering and partitioning on the fundamental problem formulation specified by XDSM in order to achieve 
-    significant reductions in computational costs for large scale optimization problems. Figure \ref{fig:dsm_full} illustrates 
-    the same notional analysis as in Figure \ref{fig:dsm_simple} represented as an XDSM.
-
-    \begin{figure}[!hbp]
-        \begin{center}
-        \includegraphics[width=.75\textwidth]{images/dsm_simple}
-        \caption{[TODO: Generate XDSM]Extended DSM graph for Eqn. \ref{eqn:simple_fpf} \label{fig:dsm_full}}
-        \end{center}
-    \end{figure}
+    Although XDSM captures part of the FDT formulation, it still leaves out the variables themselves. As a result, all variable information is aggregated 
+    so that for the problem from Eqn. \ref{eqn:simple_fpf} you can say that $A$ depends on $B$ or vice versa, but you can't identify which 
+    individual variables are interacting in the dependency cycle. Without the detailed variable information, you can't construct the compatibility 
+    constraints necessary to implement the problem. Hence, to complete the DSM for use with a fundamental problem formulation you need 
+    to add in nodes for each variable in the analysis. In the following sections we describe the necessary syntax for specifying a 
+    problem formulation with DSM syntax and show how to make use of the resulting graph to find the FPF from any given graph.

### 2012-10-03 intro.tex
tweaking the intro some more

@@
     information about objectives and constraints, it neglects the data dependency captured by DSM. 
-    [TODO: There is more recent research to cite here, like Allison's Work.]
+    [TODO: There is more recent research to cite here. e.g. Allison's Work.]
+
+    [TODO: insert FDT of simple problem here]
 
     Lamb and Martins included the variables, objectives, and constraint functions as nodes in an Extended 
-    DSM (XDSM) in order to capture a more complete problem formulation for MDAO problems\cite{Lambe2012}. By adding in those 
+    DSM (XDSM) in order to capture a more complete description of solution strategies for MDAO problems\cite{Lambe2012}. By adding in those 
     elements, they partially combined a traditional DSM with an FDT. 
@@
 
-    Although XDSM captures part of the FDT formulation, it still leaves out the variables themselves. As a result, all variable information is aggregated 
+    [TODO: insert XDSM for simple problem solved with MDF here]
+
+    Although XDSM captures part functional aspects of FDT, it  leaves out the variables themselves. As a result, all variable information is aggregated 
     so that for the problem from Eqn. \ref{eqn:simple_fpf} you can say that $A$ depends on $B$ or vice versa, but you can't identify which 
     individual variables are interacting in the dependency cycle. Without the detailed variable information, you can't construct the compatibility 
-    constraints necessary to implement the problem. Hence, to complete the DSM for use with a fundamental problem formulation you need 
-    to add in nodes for each variable in the analysis. In the following sections we describe the necessary syntax for specifying a 
-    problem formulation with DSM syntax and show how to make use of the resulting graph to find the FPF from any given graph.
+    constraints necessary to implement the problem. Additionally, XDSM requires the use of solver and optimizer blocks to represent 
+    the relationship between design variables and objectives/constraints. By introducing solver or optimizer blocks XDSM automatically provides
+    some kind of solution strategy. Hence XDSM is too specific for use with a fundamental problem formulation.
+
+    By taking XDSM, removing the requirement for solver/optimizer nodes and adding in nodes for each variable we are able to fully describe 
+    the problem formulation for a general MDAO problem. In the following sections we describe the necessary syntax for specifying a 
+    problem formulation withing a DSM and show how to make use of the resulting graph to find the FPF from any given graph. For completeness we 
+    also present algorithms for converting between the general DSM and the more specific formulations discussed above.

### 2012-10-04 intro.tex
adding XDSM

@@
     more efficient optimization problem decompositions\cite{Wagner1993}. While FDT succeeds at capturing the 
-    information about objectives and constraints, it neglects the data dependency captured by DSM. 
+    information about objectives and constraints, it neglects the data dependency captured by DSM. For instance, 
+    although we know from the FPF in Eqn. \ref{eqn:simple_fpf} that the objective, $f$, is dependent on the 
+    output of analysis $A$, you could not determine that from the FDT in Fig. \ref{fig:FDT_simple} alone.
+
     [TODO: There is more recent research to cite here. e.g. Allison's Work.]
 
-    [TODO: insert FDT of simple problem here]
+    \begin{figure}
+        \begin{center}
+        \begin{tabular}{|c|c|c|c|c|c|c|}
+            \hline
+                 & $x$ & $y$ & $y^t$ & $z$ & $z^t$ & m \\ \hline
+            $A$  & 1  & 1    &       &     &       &   \\ \hline
+            $B$  & 1  &      &       & 1   &       &   \\ \hline
+            $f$  &    &      &       &     &       & 1 \\ \hline
+            $g1$ &    &      &       & 1   & 1     &   \\ \hline
+            $g2$ &    & 1    & 1     &     &       &   \\
+            \hline
+        \end{tabular}
+        \caption{Functional Dependency Table (FDT) for Eqn. \ref{eqn:simple_fpf} \label{fig:FDT_simple}}
+        \end{center}
+    \end{figure}
 

### 2012-10-10 intro.tex
Editing changes

@@
 
-    As the size and complexity of engineering systems grows the time and expense for setting up 
-    analysis models grows with them. Multidisciplinary Design Analysis and Optimization (MDAO)
+    As the size and complexity of engineering systems grow, the time and expense for setting up 
+    analysis models grow with them. Multidisciplinary Design Analysis and Optimization (MDAO)
     frameworks such as OpenMDAO\cite{Gray2012} and ModelCenter have enabled a new level of analysis tool integration 
-    and paved the way for models of with more analysis tools and increasing numbers of multidisciplinary couplings. 
-    Such complex models present a distinct challenge proper implementation of any given solution strategy .
+    and paved the way for models with more analysis tools and increasing numbers of multidisciplinary couplings. 
+    Such complex models present a distinct challenge for proper implementation of any given solution strategy.
 
-    We propose the use of graph based problem formulation that generally describes a
+    We propose the use of graph-based problem formulation that generally describes a
     given problem, supports algorithmic manipulation to find beneficial solutions strategies, and enables automated 
@@
     Notably absent from the preceding list are any kind of solvers, optimizers, or other iterative solution finding tools. 
-    At it most basic, a problem formulation includes only information about what is being sought after in a given problem, or what the
+    At its most basic, a problem formulation includes only information about what is being sought after in a given problem or what the
     goals of a given problem are. It need not contain any information about solution paths or strategies to reach those goals.
 
-    We define a problem specification striped down to it's most essential parts to be the 
+    We define a problem specification striped down to its most essential parts to be the 
     Fundamental Problem Formulation (FPF). By definition, the FPF for any given problem will be constant regardless 
@@
 
-    The graph based syntax used in this work is based on the Design Structure Matrix (DSM). 
-    This provides several key features that make it useful for working with large scale MDAO problems. 
-    It provides a rigid structure that can be easily manipulated with a wide range of well establish graph-theory algorithms 
+    The graph-based syntax used in this work is based on the Design Structure Matrix (DSM). 
+    This provides several key features that make it useful for working with large-scale MDAO problems. 
+    It provides a rigid structure that can be easily manipulated with a wide range of well-established graph-theory algorithms 
     for the purposes of problem decomposition. It also provides a means for testing a given 
-    problem formulation to check if it is the FPF, and if not to reduce it to the FPF. Lastly, it lends itself 
+    problem formulation to check if it is the FPF, and if not, to reduce it to the FPF. Lastly, it lends itself 
     well to interacting with MDAO frameworks.
@@
 \section{Specific vs Fundamental Problem Formulation }
-    For a simple, notional problem  is given as follows:
+    A simple, notional problem is given as follows:
 
@@
 
-    Where $A$ and $B$ represent analysis tools, $f,g$ are the objective and constraint functions respectively. 
+    Where $A$ and $B$ represent analysis tools, and $f$ and $g$ are the objective and constraint functions respectively. 
     Equation \ref{eqn:simple} makes an inherent assumption about the solution strategy for the problem. 
-    Analysis $A$ outputs $z$ which is an input to $B$. Hence, $A$ should be run before $B$ with 
+    Analysis $A$ outputs $z$, which is an input to $B$. Hence, $A$ should be run before $B$ with 
     the $y$ being iterated on to convergence with $y^t$. However, a different solution could be equally valid and still represent
@@
 
-    Equation \ref{eqn:simple2} differs only slightly from Eqn. \ref{eqn:simple}. $A$ is now dependent on the output of $B$, 
+    Equation \ref{eqn:simple2} differs only slightly from Eq. \ref{eqn:simple}. $A$ is now dependent on the output of $B$, 
     and $z$ will be iterated on to satisfy $g$. Now the problem can be solved by running $B$ first and then $A$.
-    Since the formulations in Eqn. \ref{eqn:simple} and \ref{eqn:simple2} both describe the same problem and neither can be the FPF. 
-    There must be a more fundamental description of the problem that is common between them. We present the FPF as follows:
+    Since the formulations in Eqs. \ref{eqn:simple} and \ref{eqn:simple2} both describe the same problem and neither can be the
+    FPF, there must be a more fundamental description of the problem that is common between them. We present the FPF as follows:
 
@@
 
-    The FPF in Eqn. \ref{eqn:simple_fpf} differs from both Eqn. \ref{eqn:simple} and Eqn. \ref{eqn:simple2} because it has 
+    The FPF in Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} and Eq. \ref{eqn:simple2} because it has 
     two constraints which both must be met. The presence of both of these constraints fully decouples the problem so that 
-    either $A$ or $B$ could be run first or both could be run simultaneously. By removing either constraint, and replacing 
+    either $A$ or $B$ could be run first or both could be run simultaneously. By removing either constraint and replacing 
     it with a direct dependence between the two analyses you could regain the earlier two problem formulations.
@@
     The challenge with using this traditional mathematical syntax is that it is not easily manipulated or analyzed. 
-    A number of matrix based methods have been used successfully to translate the mathematical syntax into a more useful computational form. 
+    A number of matrix-based methods have been used successfully to translate the mathematical syntax into a more useful computational form. 
     Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools where off 
@@
 
-    An alternate matrix based syntax, called a Functional Dependency Table (FDT), was proposed by Michelena and Papalambros. 
+    An alternative matrix-based syntax, called a Functional Dependency Table (FDT), was proposed by Michelena and Papalambros. 
     FDT represents the relationship between functions, including objectives and constraints, and specific variables that affect 
@@
     information about objectives and constraints, it neglects the data dependency captured by DSM. For instance, 
-    although we know from the FPF in Eqn. \ref{eqn:simple_fpf} that the objective, $f$, is dependent on the 
+    although we know from the FPF in Eq. \ref{eqn:simple_fpf} that the objective, $f$, is dependent on the 
     output of analysis $A$, you could not determine that from the FDT in Fig. \ref{fig:FDT_simple} alone.
@@
         \end{tabular}
-        \caption{Functional Dependency Table (FDT) for Eqn. \ref{eqn:simple_fpf} \label{fig:FDT_simple}}
+        \caption{Functional Dependency Table (FDT) for Eq. \ref{eqn:simple_fpf} \label{fig:FDT_simple}}
         \end{center}
@@
     This allows XDSM to represent data dependency between multiple analysis tools as well as between analysis tools and
-    objective/constraint functions. With the additional information included in an XDSM Lu and Martins applied both ordering and partitioning 
+    objective/constraint functions. With the additional information included in an XDSM, Lu and Martins applied both ordering and partitioning 
     algorithms on an MDAO test problem named the Scalable Problem \cite{Lu2012}.
@@
     Although XDSM captures part functional aspects of FDT, it leaves out the variables themselves. As a result, all variable information is aggregated 
-    so that for the problem from Eqn. \ref{eqn:simple_fpf} you can say that $A$ depends on $B$ or vice versa, but you can't identify which 
+    so that for the problem from Eq. \ref{eqn:simple_fpf} you can say that $A$ depends on $B$ or vice versa, but you can't identify which 
     individual variables are interacting in the dependency cycle. Without the detailed variable information, you can't construct the compatibility 
     constraints necessary to implement the problem. Additionally, XDSM requires the use of solver and optimizer blocks to represent 
-    the relationship between design variables and objectives/constraints. By introducing solver or optimizer blocks XDSM automatically provides
-    some kind of solution strategy. Hence XDSM is too specific for use with a fundamental problem formulation.
+    the relationship between design variables and objectives/constraints. By introducing solver or optimizer blocks, XDSM automatically provides
+    some kind of solution strategy. Hence XDSM is too specific to use with a fundamental problem formulation.
 
-    By taking XDSM, removing the requirement for solver/optimizer nodes and adding in nodes for each variable we are able to fully describe 
+    By taking XDSM, removing the requirement for solver/optimizer nodes, and adding nodes for each variable, we are able to fully describe 
     the problem formulation for a general MDAO problem. In the following sections we describe the necessary syntax for specifying a 
-    problem formulation withing a DSM and show how to make use of the resulting graph to find the FPF from any given graph. For completeness we 
+    problem formulation within a DSM and show how to use the resulting graph to find the FPF from any given graph. For completeness we 
     also present algorithms for converting between the general DSM and the more specific formulations discussed above.

### 2012-10-12 intro.tex
stuff

--- 2012-10-04 intro.tex
@@
     DSM (XDSM) in order to capture a more complete description of solution strategies for MDAO problems\cite{Lambe2012}. By adding in those 
-    elements, they partially combined a traditional DSM with an FDT. 
-    This allows XDSM to represent data dependency between multiple analysis tools as well as between analysis tools and
-    objective/constraint functions. With the additional information included in an XDSM Lu and Martins applied both ordering and partitioning 
+    elements, they partially combined a traditional DSM with an FDT. This allows XDSM to represent data 
+    dependency between multiple analysis tools as well as between analysis tools and objective/constraint functions. 
+    With the additional information included in an XDSM Lu and Martins applied both ordering and partitioning 
     algorithms on an MDAO test problem named the Scalable Problem \cite{Lu2012}.
 
-    [TODO: insert XDSM for simple problem solved with MDF here]
+    \begin{figure}
+        \begin{center}
+        \includegraphics[height=.25\textheight]{XDSM/simple}
+        \caption{XDSM for Eqn. \ref{eqn:simple}, with a gauss-siedel iteration and MDF solution architecture. \label{fig:XDSM_simple}}
+        \end{center}
+    \end{figure}
 
@@
     the relationship between design variables and objectives/constraints. By introducing solver or optimizer blocks XDSM automatically provides
-    some kind of solution strategy. Hence XDSM is too specific for use with a fundamental problem formulation.
+    some kind of solution strategy. The XDSM for Eqn. \ref{eqn:simple} is given in Figure \ref{fig:XDSM_simple}. 
+    This diagram is shown with an assumed gauss-siedel iteration scheme and a MDF solution architecture. 
+    Hence XDSM is too specific for use with a fundamental problem formulation.
 
-    By taking XDSM, removing the requirement for solver/optimizer nodes and adding in nodes for each variable we are able to fully describe 
+    By taking XDSM, removing the solver/optimizer nodes and adding in nodes for each variable we are able to fully describe 
     the problem formulation for a general MDAO problem. In the following sections we describe the necessary syntax for specifying a 

### 2012-10-12 graph_syntax.tex

@@
-\section{Graph Based Problem Formulation Syntax}
+\section{Graph-based Problem Formulation Syntax}
     \subsection{Formulation Graph Syntax}
@@
                 \item local outputs 
-                \item execution properties: the properties associated with running the code, such as run time.
+                \item execution properties: these are the properties associated with running the code, such as run time
             \end{itemize}
@@
     The graph representation of a data flow is cast to utilize the extensive library of algorithms in graph theory to analyze a directed weighted graph. 
-    Edge weights are used to represent the metrics associated with a data flow
+    Edge weights are used to represent the metrics associated with a data flow:
     \begin{itemize}
@@
     The key assumption is that identical variables are recognized as such. This serves as the basis for creating a data flow by connecting compatible input and output nodes with a directed edge. 
-    In order to represent the fact that execution of an analysis code does not depend on the number of outputs being used, it is represented as depicted in the following figure (not made yet).
+    To represent the fact that execution of an analysis code does not depend on the number of outputs being used, we have created the following figure (not made yet).
 
     This information immediately leads to the maximal connectivity graph, which is formed by placing a directed edge from each local output or global parameter to each matching local input or global output.
-    Whenever multiple edges are connected to a single input a conflict occurs because only one may be used. Resolving these conflicts is one key challenge in creating a data flow.
+    Whenever multiple edges are connected to a single input, a conflict occurs because only one may be used. Resolving these conflicts is one key challenge in creating a data flow.
 %    \begin{itemize}
@@
     \begin{itemize}
-        \item cycles indicate convergence loops or design varaible loops
-        \item Problem can't be solved until all loops are *removed* by adding in solvers/optimizers
-        \item *special* nodes for solvers and optimizers that *break* loops (from an algorithmic point of view)
+        \item Cycles indicate convergence loops or design variable loops
+        \item Problem can't be solved until all loops are *removed* by adding solvers/optimizers
+        \item *Special* nodes for solvers and optimizers that *break* loops (from an algorithmic point of view)
         \item FPF represents the minimal amount of information necessary to define a problem

### 2012-10-12 intro.tex

--- 2012-10-10 intro.tex
@@
 
-    [TODO: insert XDSM for simple problem solved with MDF here]
+    \begin{figure}
+        \begin{center}
+        \includegraphics[height=.25\textheight]{XDSM/simple}
+        \caption{XDSM for Eqn. \ref{eqn:simple}, with a gauss-siedel iteration and MDF solution architecture. \label{fig:XDSM_simple}}
+        \end{center}
+    \end{figure}
 
-    Although XDSM captures part functional aspects of FDT, it leaves out the variables themselves. As a result, all variable information is aggregated 
+    Although XDSM captures part of the functional aspects of FDT, it leaves out the variables themselves. 
+    As a result, all variable information is aggregated 
     so that for the problem from Eq. \ref{eqn:simple_fpf} you can say that $A$ depends on $B$ or vice versa, but you can't identify which 
@@
     constraints necessary to implement the problem. Additionally, XDSM requires the use of solver and optimizer blocks to represent 
-    the relationship between design variables and objectives/constraints. By introducing solver or optimizer blocks, XDSM automatically provides
-    some kind of solution strategy. Hence XDSM is too specific to use with a fundamental problem formulation.
+    the relationship between design variables and objectives/constraints. By introducing solver or optimizer blocks XDSM automatically provides
+    some kind of solution strategy. The XDSM for Eqn. \ref{eqn:simple} is given in Figure \ref{fig:XDSM_simple}. 
+    This diagram is shown with an assumed gauss-siedel iteration scheme and a MDF solution architecture. 
+    Hence XDSM is too specific for use with a fundamental problem formulation.
 

### 2012-10-12 intro.tex
added in a new reference, fixed two others

@@
         frameworks such as OpenMDAO\cite{Gray2012} and ModelCenter have enabled a new level of analysis tool integration 
-    and paved the way for models with more analysis tools and increasing numbers of multidisciplinary couplings. 
+        and paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. 
         Such complex models present a distinct challenge for proper implementation of any given solution strategy.
@@
         given problem, supports algorithmic manipulation to find beneficial solutions strategies, and enables automated 
-    implementation of those strategies inside an MDAO framework. In order to serve that purpose, the graph includes the following information: 
+        implementation of those strategies inside an MDAO framework. In order to serve that purpose, the graph must include the following information: 
         \begin{itemize}
@@
         either $A$ or $B$ could be run first or both could be run simultaneously. By removing either constraint and replacing 
-    it with a direct dependence between the two analyses you could regain the earlier two problem formulations.
+        it with a direct dependence between the two analyses you could regain the earlier two problem formulations. Alexandrov
+        and Lewis demonstrated the value of a more modular approach to problem formulation because it enables one to transition between different 
+        MDAO solution strategies depending on the specifics of the problem\cite{Alex2000}.
 
@@
 
-    An alternative matrix-based syntax, called a Functional Dependency Table (FDT), was proposed by Michelena and Papalambros. 
+        An alternative matrix-based syntax, called a Functional Dependency Table (FDT), was proposed by Wagner and Papalambros. 
         FDT represents the relationship between functions, including objectives and constraints, and specific variables that affect 
-    them\cite{Michelena1997}. Similar to DSM, FDT also describes an adjacency matrix of a graph. Unlike the DSM graph, 
+        them\cite{Wagner1993}. Similar to DSM, FDT also describes an adjacency matrix of a graph. Unlike the DSM graph, 
         however, the graph is undirected and nodes can represent analysis tools, objectives, 
         or constraints. Edges between nodes represent a dependence on the same 
-    variable. Wagner and Papalambros made use of the FDT to solve a graph partitioning problem that yielded 
-    more efficient optimization problem decompositions\cite{Wagner1993}. While FDT succeeds at capturing the 
+        variable. Michelena and Papalambros made use of the FDT to solve a graph partitioning problem that yielded 
+        more efficient optimization problem decompositions\cite{Michelena1997}. While FDT succeeds at capturing the 
         information about objectives and constraints, it neglects the data dependency captured by DSM. For instance, 
         although we know from the FPF in Eq. \ref{eqn:simple_fpf} that the objective, $f$, is dependent on the 
-    output of analysis $A$, you could not determine that from the FDT in Fig. \ref{fig:FDT_simple} alone.
-
-    [TODO: There is more recent research to cite here. e.g. Allison's Work.]
+        output of analysis $A$, you could not determine that from the FDT in Fig. \ref{fig:FDT_simple} alone. This make FDT 
+        applicable for partitioning decisions, but not for ordering problems.
 
@@
             \includegraphics[height=.25\textheight]{XDSM/simple}
-        \caption{XDSM for Eqn. \ref{eqn:simple}, with a gauss-siedel iteration and MDF solution architecture. \label{fig:XDSM_simple}}
+            \caption{XDSM for Eq. \ref{eqn:simple}, with a gauss-siedel iteration and MDF solution architecture. \label{fig:XDSM_simple}}
             \end{center}
@@
         the relationship between design variables and objectives/constraints. By introducing solver or optimizer blocks XDSM automatically provides
-    some kind of solution strategy. The XDSM for Eqn. \ref{eqn:simple} is given in Figure \ref{fig:XDSM_simple}. 
+        some kind of solution strategy. The XDSM for Eq. \ref{eqn:simple} is given in Figure \ref{fig:XDSM_simple}. 
         This diagram is shown with an assumed gauss-siedel iteration scheme and a MDF solution architecture. 

### 2012-10-15 graph_syntax.tex
DJP adding to my section

@@
 \section{Graph-based Problem Formulation Syntax}
+	For a large scale problem with many analyses, obtaining the FPF is likely to be a challenge. In general, at set of analysis codes will be given and the desired outputs will be specified as
+	    \begin{align}
+            given & \ \ A: \lbrace x_i \lvert i \in A_\textrm{input} \rbrace \rightarrow \lbrace x_i \lvert i \in A_\textrm{output} \rbrace \notag
+            \\    & \ \ B: \lbrace x_i \lvert i \in B_\textrm{input} \rbrace \rightarrow \lbrace x_i \lvert i \in B_\textrm{output} \rbrace \notag
+			\\    & \ \ \quad \quad \quad \quad \quad \quad  \quad \quad \vdots
+            \\min. &\ \ f(x_i), i \in \mathcal{O} \notag
+            \\w.r.t. & \ \ x_i, i \in \mathcal{I}, \notag
+            \label{eqn:preFPF}
+        \end{align}
+	where $\mathcal{O}$ is the index set of the global outputs and $\mathcal{I}$ is the index set of the global inputs.
+For this to be an FPF, there must be no \emph{conflicts} or \emph{holes}.
+A conflict arises when different analyses produce the same output; two analysis codes $A$ and $B$ are in conflict if
+		\begin{equation}
+			A_\textrm{output} \cap B_\textrm{output} \neq \emptyset.
+		\end{equation}
+	A hole arises when a local input to an analysis block is neither a global input nor a local output of any analysis block. If we let $\mathcal{H}$ denote the index set of local inputs which are holes
+		\begin{equation}
+			i \notin \bigcup (\mathcal{I},A_\textrm{output},B_\textrm{output},\ldots)  \implies  i \in \mathcal{H}.
+		\end{equation}
+	If the given set of analyses contains any holes, an FPF cannot be obtained.
+On the contrary, if there are conflicts multiple FPFs may be obtained.
+This is because every conflict represents a choice of which analysis to use and each choice could (potentially) yeild a different but valid FPF.
+The following section develops an application of graphy theory to obtain multiple FPFs from a set of analyses and desired outputs.
+
     \subsection{Formulation Graph Syntax}

### 2012-12-03 max_connect_graph.py
Update to python code

@@ -82,6 +82,7 @@ def create_sample_graph():
     a.append('x');  b.append('Ci1');
     a.append('Co1');b.append('Gi1');
     a.append('Co1');b.append('Hi1');
+    a.append('Co1');b.append('Ai2');
 
     cab(DG,'D',3,1,1,(2,1),pos)
     a.append('x');  b.append('Di1');
@@ -128,7 +129,6 @@ def create_sample_graph():
         ('Fo2','Gi2',w),
         ('Ko2','y1',w),
         ('Ho3','Di3',w),
-        ('Co1','Ai2',w),
         ('Fo3','Di2',w),
         ('Io2','Ki2',w),
         ('Ho2','y2',w),
@@ -146,8 +146,8 @@ def remove_node(DG,i):
     return DG;
 
 def resolve_conflict(edgs,DG):
-    print len(edgs)
-    print edgs
+    #print len(edgs)
+    print "Conflicted edges:",edgs
     # We must determine which of the edges to choose from are fixed
     fxd = [];
     nfxd = 0;
@@ -155,33 +155,35 @@ def resolve_conflict(edgs,DG):
         check = DG.edge[edg[0]][edg[1]]['fixed']
         fxd.append(check);
         if check==True:
-            nfxd=+1;
+            nfxd=nfxd+1;
+        #print 'nfxd',nfxd
     if nfxd>1:
-        print 'More than one edge at node',edg[1],'is fixed'
+        print '  More than one edge at node',edg[1],'is fixed'
         quit()
     elif nfxd==1:
         chosen_edge = edgs[fxd.index(True)] # choose the edge that is not fixed
-        print 'Edge',edg,'is fixed'
+        print '  Edge',edg,'is fixed'
     else:
         # ---- A decision only needs to be made when none of the edges are fixed ----
-        choice = 'rank';
+        choice = 'user';
         if choice=='rank':
             rnk = [];
             for edg in edgs:
                 rnk.append(DG.edge[edg[0]][edg[1]]['rank'])
             hghst = max(rnk);
             chosen_edge = edgs[rnk.index(hghst)];
-            print 'Edge',chosen_edge,'was chosen'
         elif choice=='random':
-            print 'No edges are fixed'
             chosen_edge = edgs[0]; # "randomly" choose the first element
         elif choice=='user':
-            print 'user'
+            print "Please enter the index of the connection to resolve the conflict: "
+            inp_int = int(raw_input())
+            chosen_edge = edgs[inp_int];
         elif choice=='metric':
[…]

### 2012-12-03 max_connect_graph.py
added if __name__=='__main__'

@@ -273,36 +273,39 @@ def find_cycles(DG):
             CG.add_edge(i[j],i[j+1])
     return [CG,cycles];
 
-out = create_sample_graph()
-maxDG = out[0]
-pos = out[1]
 
-FPF = obtain_fpf(maxDG)
+if __name__== "__main__": 
+    
+    out = create_sample_graph()
+    maxDG = out[0]
+    pos = out[1]
 
-SG = extract_simple_graph(maxDG)
+    FPF = obtain_fpf(maxDG)
 
-out = find_cycles(SG) 
-CGSG = out[0]
-cyclesSG = out[1]
-print cyclesSG
-print CGSG.edges()
+    SG = extract_simple_graph(maxDG)
 
-out = find_cycles(FPF) 
-CG = out[0]
-cycles = out[1]
-#print len(cycles)
-#print cycles
-#print FPF.edge['Co1']['Gi1']
+    out = find_cycles(SG) 
+    CGSG = out[0]
+    cyclesSG = out[1]
+    print cyclesSG
+    print CGSG.edges()
 
-pylab.figure(1)
-#nx.draw(maxDG,pos,width=3)
-#nx.draw(FPF,pos,width=1.5,edge_color='b')
-#nx.draw_networkx_edges(CG,pos,width=0.75,edge_color='y')
-nx.draw(SG,pos,width=2)
-#nx.draw_networkx_edges(FPF,pos,width=1.25,edge_color='b')
-nx.draw_networkx_edges(CGSG,pos,width=0.75,edge_color='y')
+    out = find_cycles(FPF) 
+    CG = out[0]
+    cycles = out[1]
+    #print len(cycles)
+    #print cycles
+    #print FPF.edge['Co1']['Gi1']
 
-pylab.figure(2)
-nx.draw(maxDG,pos,width = 0.5)
+    pylab.figure(1)
+    #nx.draw(maxDG,pos,width=3)
+    #nx.draw(FPF,pos,width=1.5,edge_color='b')
+    #nx.draw_networkx_edges(CG,pos,width=0.75,edge_color='y')
+    nx.draw(SG,pos,width=2)
+    #nx.draw_networkx_edges(FPF,pos,width=1.25,edge_color='b')
+    nx.draw_networkx_edges(CGSG,pos,width=0.75,edge_color='y')
 
-pylab.show()
+    pylab.figure(2)
+    nx.draw(maxDG,pos,width = 0.5)
+
+    pylab.show()

### 2012-12-03 max_connect_graph.py
added some simple error raising

@@ -158,8 +158,7 @@ def resolve_conflict(edgs,DG):
             nfxd=nfxd+1;
         #print 'nfxd',nfxd
     if nfxd>1:
-        print '  More than one edge at node',edg[1],'is fixed'
-        quit()
+        raise  RuntimeError('More than one edge at node %s is fixed'%str(edg[1])
     elif nfxd==1:
         chosen_edge = edgs[fxd.index(True)] # choose the edge that is not fixed
         print '  Edge',edg,'is fixed'
@@ -181,8 +180,8 @@ def resolve_conflict(edgs,DG):
         elif choice=='metric':
             print 'metric'
         else:
-            print 'Invalid choice for conflict resolution'
-            quit()
+            raise RuntimeError('Invalid choice for conflict resolution')
+            
     print '  Edge',chosen_edge,'was chosen'
     edgs.remove(chosen_edge)
     DG.remove_edges_from(edgs) # Remove all the edges except for the one that was selected
@@ -275,7 +274,7 @@ def find_cycles(DG):
 
 
 if __name__== "__main__": 
-    
+
     out = create_sample_graph()
     maxDG = out[0]
     pos = out[1]

### 2013-02-12 graph_syntax.tex
DJP: a couple changes
writing update

--- 2012-10-12 graph_syntax.tex
@@
 \section{Graph-based Problem Formulation Syntax}
-    \subsection{Formulation Graph Syntax}
-    Rather than start with an adjacency, we chose to work directly with a directed cyclic graph to develop a syntax for the FPF. 
-    The following information must be provided to start:
+We begin by discussing the general syntax of graph theory and then expand upon this syntax to represent the data flow of an MDAO problem.
+A graph is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, which means that the elements of $E$ are 2--element subsets of $V$ \cite{Diestel2010}. 
+For a directed graph we introduce two maps $S:E \to V$ and $F:E \to V$ to denote the starting and final nodes of each edge, respectively. 
+\begin{figure}[htb!]
+	\begin{center}
+	\includegraphics[width=1.5in]{images/example_directed_graph}
+	\end{center}
+	\vspace{-20pt}
+\caption{Example directed graph.}
+\label{f:example directed graph}
+\end{figure}
+As an example, for the directed graph shown in Fig.~\ref{f:example directed graph} we have
+\begin{IEEEeqnarray*}{rCl}
+V & = & \{1,2,3,4,5\}, \\
+E & = & \big\{\{1,2\},\{2,3\},\{1,3\},\{3,4\}\big\}, \\
+S & = & \Big\{ \big(\{1,2\},2\big),\big(\{2,3\},3\big),\big(\{1,3\},1\big),\big(\{3,4\},3\big) \Big\}, \\
+F & = & \Big\{ \big(\{1,2\},1\big),\big(\{2,3\},2\big),\big(\{1,3\},3\big),\big(\{3,4\},4\big) \Big\}.
+\end{IEEEeqnarray*}
+
+We now assign additional properties to the nodes and edges so that they may represent the data flow of an MDAO problem. 
+The first property assigned to nodes and edges is the \emph{type}. The possible node types are:
 \begin{itemize}
-        \item Analysis blocks: an analysis block represents any calculation, and each comes with
+\item global input
+\item global output
+\item local input
+\item local output
+\item analysis start
+\item analysis finish
+\end{itemize}
+and the possible edge types are
 \begin{itemize}
-                \item local inputs
-                \item local outputs 
-                \item execution properties: these are the properties associated with running the code, such as run time
+\item connection
+\item in
+\item analysis
+\item out
 \end{itemize}
-        \item Global parameters: these may serve as fixed inputs to the local inputs of analysis blocks
-        \item Global outputs: these may represent
-            \begin{itemize}
-                \item objectives
-                \item constraints
-                \item residuals: (not sure)
-            \end{itemize}
-    \end{itemize}
+Then, considering sets $T_\txt{node}$ and $T_\txt{edge}$ to be sets containing the possible node types and edge types, respectively, we use mappings $t_\txt{node}:V \to T_\txt{node}$ and $t_\txt{edge}:E \to T_\txt{edge}$ to assign a type to each node and edge.
+Nodes and edges are named according to their type (e.g.
+connection edge).
+Additional properties will be discussed is subsequent sections.
 
-    The graph representation of a data flow is cast to utilize the extensive library of algorithms in graph theory to analyze a directed weighted graph. 
-    Edge weights are used to represent the metrics associated with a data flow:
-    \begin{itemize}
-        \item Run time: this metric is a property of an analysis block
-        \item Fidelity: this metric is a property of an individual local output
-        \item Expected Convergence
-    \end{itemize}
+Next, we introduce the graph $A$ which represents a single analysis block broken down into its individual inputs and outputs.
+An example graph is shown in Fig.~\ref{f:analysis block} to demonstrate the basic structure.
+This structure is derived from the need to represent each input and output individually while representing the actual analysis with a single node or edge.
+The `local input' nodes each represent a single input of the analysis block.
+Each of these nodes is connected to a single `analysis start' node via an `in' edge.
+Next the analysis start node is directed to the `analysis finish' node via the `analysis edge', which represents the actual analysis.
+Finally, the analysis finish node is directed to `local output' nodes via individual `out' edges.
+\begin{figure}[htb!]
+	\begin{center}
+	\includegraphics[width=4in]{images/analysis_block}
+	\end{center}
+	\vspace{-20pt}
+\caption{Example directed graph. The each node type and edge type is labeled, with nodes being label in italics.}
+\label{f:analysis block}
+\end{figure}
+`Connection' edges connect local output nodes from one analysis block to either the local input nodes of other analysis blocks or global output node.
+In this way, the analysis block graph is the fundamental building block of the MDAO problem data flow.
 
-    The key assumption is that identical variables are recognized as such. This serves as the basis for creating a data flow by connecting compatible input and output nodes with a directed edge. 
-    To represent the fact that execution of an analysis code does not depend on the number of outputs being used, we have created the following figure (not made yet).
+\subsection{Maximal Connectivity Graph}
+Now we define the node property \emph{variable name}, which represents the unique identity of the variables passed between analysis blocks as represented by nodes of type local input, local output, global input, and global output.
+Let $N$ be the set containing each unique variable name (no multiplicity), and let $n:V \to N$ be a mapping that assigns a variable name to a node.
 
-    This information immediately leads to the maximal connectivity graph, which is formed by placing a directed edge from each local output or global parameter to each matching local input or global output.
-    Whenever multiple edges are connected to a single input, a conflict occurs because only one may be used. Resolving these conflicts is one key challenge in creating a data flow.
+Next, Consider an MDAO problem defined by a set of codes, global inputs, and global outputs, as represented by analysis blocks $A_i=(V_{A_i},E_{A_i}),i=1,\ldots,m$, the set of nodes $\mathcal I$, and the set of nodes $\mathcal O$, respectively.
+Then we may construct the \emph{maximal connectivity graph} $M=(V_M,E_M)$ as
+\begin{IEEEeqnarray*}{rCl}
+V_M & = & \mathcal I \cup \mathcal O \cup \left( \bigcup_{i = 1}^m V_{A_i} \right), \\
+E_M & = & C \cup \left( \bigcup_{i=1}^m E_{A_i} \right),
+\end{IEEEeqnarray*}
+where $C$ is a set containing connection edges defined for $v_1,v_2 \in V_M$ as:
+\begin{IEEEeqnarray*}{rl}
+c = & \{v_1,v_2\} \in C \txt{ if:} \\
+& t_\txt{node}(v_1) = \txt{`local output' or `global input'}, \\
+& t_\txt{node}(v_2) = \txt{`local input' or `global output'},\txt{ and} \\
+& n(v_1) = n(v_2),
+\end{IEEEeqnarray*}
+in which case $S(c) = v_1$ and $F(c) = v_2$.
+The maximal connectivity graph contains all possible connections that may exist between the given analysis blocks, and these connections are made by connecting variables with the same name, which requires a consistent naming convention between analysis blocks.
+
+\subsection{Fundamental Problem Formulation Graph}
+We may now define the graph representing the fundamental problem formulation $P=(V_P,E_P)$ as a subgraph of $M$ such that there are no \emph{holes} or \emph{collisions}. 
+Before defining these terms, let us first define $\txt{deg}^-(v)$ as the \emph{indegree} $\txt{deg}^+(v)$ as the \emph{outdegree} of vertex $v$, where indegree and outdegree refer to the number of edges directed into vertex $v$ and the number of edges directed out of vertex $v$, respectively.
+Let $H$ be the set of all nodes which are holes in $M$ and define $H$ as
+\begin{equation}
+v \in H \txt{ if } t_\txt{node}(v) = \txt{`local input' or `global output'} \txt{ and } \txt{deg}^-(v) = 0.
+\end{equation}
+Next, let $L$ be the set of all nodes which are collisions in $M$ and define $L$ as
+\begin{equation}
+v \in L \txt{ if } t_\txt{node}(v) = \txt{`local input' or `global output'} \txt{ and } \txt{deg}^-(v) > 1.
+\end{equation}
+This means that exactly one edge must be directed into each local input node and each global output node.
+We also require $P$ to contain all of the global output nodes, meaning $P\subset M$ is a fundamental problem formulation if
+\begin{equation}
+\mathcal O \subset V_P \ \txt{ and } \ v \notin H, v \notin L \ \forall v \in V_P.
+\end{equation}
+
+\subsection{Obtaining the Fundamental Problem Formulation Graph}
+In general, the fundamental problem formulation graph $P$ may not exist, or there may be multiple different solutions. Then the first problem is to determine if there exists a graph $P$.
+In practice, it is likely that this definition will not produce a unique solution. 
+A collision represents a choice.
+
+% \begin{figure}[htb!]
+	% \begin{center}
+	% \includegraphics[width=2in]{images/hour_glass}
+	% \end{center}
+	% \vspace{-20pt}
+% \caption{(()).}
+% \label{f:our glass}
+% \end{figure}
+% \subsection{old}
+	% For a large scale problem with many analyses, obtaining the FPF is likely to be a challenge. In general, a set of analysis codes will be given and the desired outputs will be specified as
+	    % \begin{align}
+            % given & \ \ A: \lbrace x_i \lvert i \in A_\textrm{input} \rbrace \rightarrow \lbrace x_i \lvert i \in A_\textrm{output} \rbrace \notag
+            % \\    & \ \ B: \lbrace x_i \lvert i \in B_\textrm{input} \rbrace \rightarrow \lbrace x_i \lvert i \in B_\textrm{output} \rbrace \notag
+			% \\    & \ \ \quad \quad \quad \quad \quad \quad  \quad \quad \vdots
+            % \\min. &\ \ f(x_i), i \in \mathcal{O} \notag
+            % \\w.r.t. & \ \ x_i, i \in \mathcal{I}, \notag
+            % \label{eqn:preFPF}
+        % \end{align}
+	% where $\mathcal{O}$ is the index set of the global outputs and $\mathcal{I}$ is the index set of the global inputs.
+For this to be an FPF, there must be no \emph{conflicts} or \emph{holes}.
+A conflict arises when different analyses produce the same output; two analysis codes $A$ and $B$ are in conflict if
+		% \begin{equation}
+			% A_\textrm{output} \cap B_\textrm{output} \neq \emptyset.
+		% \end{equation}
+	% A hole arises when a local input to an analysis block is neither a global input nor a local output of any analysis block.
+If we let $\mathcal{H}$ denote the index set of local inputs which are holes
+		% \begin{equation}
+			% i \in \bigcup \{A_\textrm{input},B_\textrm{input},\ldots\} \textrm{ and } i \notin \bigcup \{\mathcal{I},A_\textrm{output},B_\textrm{output},\ldots\}  \implies  i \in \mathcal{H}.
+		% \end{equation}
+	% If the given set of analyses contains any holes, an FPF cannot be obtained.
+On the contrary, if there are conflicts multiple FPFs may be obtained.
+This is because every conflict represents a choice of which analysis to use and each choice could (potentially) yeild a different but valid FPF.
+The following section develops an application of graphy theory to obtain multiple FPFs from a set of analyses and desired outputs.
+
+    % \subsection{Formulation Graph Syntax}
+    % Rather than start with an adjacency, we chose to work directly with a directed cyclic graph to develop a syntax for the FPF. 
+    % The following information must be provided to start:
+    % \begin{itemize}
+        % \item Analysis blocks: an analysis block represents any calculation, and each comes with
+            % \begin{itemize}
+                % \item local inputs
+                % \item local outputs 
+                % \item execution properties: these are the properties associated with running the code, such as run time
+            % \end{itemize}
+        % \item Global parameters: these may serve as fixed inputs to the local inputs of analysis blocks
+        % \item Global outputs: these may represent
+            % \begin{itemize}
+                % \item objectives
+                % \item constraints
+                % \item residuals: (not sure)
+            % \end{itemize}
+    % \end{itemize}
+
+    % The graph representation of a data flow is cast to utilize the extensive library of algorithms in graph theory to analyze a directed weighted graph. 
+    % Edge weights are used to represent the metrics associated with a data flow:
+    % \begin{itemize}
+        % \item Run time: this metric is a property of an analysis block
+        % \item Fidelity: this metric is a property of an individual local output
+        % \item Expected Convergence
+    % \end{itemize}
+
+    % The key assumption is that identical variables are recognized as such. This serves as the basis for creating a data flow by connecting compatible input and output nodes with a directed edge. 
+    % To represent the fact that execution of an analysis code does not depend on the number of outputs being used, we have created the following figure (not made yet).
+
+    % This information immediately leads to the maximal connectivity graph, which is formed by placing a directed edge from each local output or global parameter to each matching local input or global output.
+    % Whenever multiple edges are connected to a single input, a conflict occurs because only one may be used. Resolving these conflicts is one key challenge in creating a data flow.
 %    \begin{itemize}

### 2013-02-21 graph_syntax.tex
update

@@
 A graph is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, which means that the elements of $E$ are 2--element subsets of $V$ \cite{Diestel2010}. 
-For a directed graph we introduce two maps $S:E \to V$ and $F:E \to V$ to denote the starting and final nodes of each edge, respectively. 
+For a directed graph we construct $E$ as a set of ordered pairs instead of a set of sets.
+Each ordered pair represents an edge starting at the node indicated by the first entry and directed to the node indicated by the second entry.
+Therefore, $E$ is a mapping from $V$ to $V$.
 \begin{figure}[htb!]
@@
 V & = & \{1,2,3,4,5\}, \\
-E & = & \big\{\{1,2\},\{2,3\},\{1,3\},\{3,4\}\big\}, \\
-S & = & \Big\{ \big(\{1,2\},2\big),\big(\{2,3\},3\big),\big(\{1,3\},1\big),\big(\{3,4\},3\big) \Big\}, \\
-F & = & \Big\{ \big(\{1,2\},1\big),\big(\{2,3\},2\big),\big(\{1,3\},3\big),\big(\{3,4\},4\big) \Big\}.
+E & = & \big\{(1,2),(3,2),(1,3),(3,4)\big\}.
 \end{IEEEeqnarray*}
@@
 \item out
+\item ((consider eliminating `in' and `out' and using connection))
 \end{itemize}
@@
 \begin{IEEEeqnarray*}{rl}
-c = & \{v_1,v_2\} \in C \txt{ if:} \\
+c = & (v_1,v_2) \in C \txt{ if:} \\
 & t_\txt{node}(v_1) = \txt{`local output' or `global input'}, \\
 & t_\txt{node}(v_2) = \txt{`local input' or `global output'},\txt{ and} \\
-& n(v_1) = n(v_2),
+& n(v_1) = n(v_2).
 \end{IEEEeqnarray*}
-in which case $S(c) = v_1$ and $F(c) = v_2$.
 The maximal connectivity graph contains all possible connections that may exist between the given analysis blocks, and these connections are made by connecting variables with the same name, which requires a consistent naming convention between analysis blocks.
@@
 \subsection{Fundamental Problem Formulation Graph}
-We may now define the graph representing the fundamental problem formulation $P=(V_P,E_P)$ as a subgraph of $M$ such that there are no \emph{holes} or \emph{collisions}. 
-Before defining these terms, let us first define $\txt{deg}^-(v)$ as the \emph{indegree} $\txt{deg}^+(v)$ as the \emph{outdegree} of vertex $v$, where indegree and outdegree refer to the number of edges directed into vertex $v$ and the number of edges directed out of vertex $v$, respectively.
-Let $H$ be the set of all nodes which are holes in $M$ and define $H$ as
+Before defining the the fundamental problem formulation graph, let us first define $\txt{deg}^-(v)$ as the \emph{indegree} $\txt{deg}^+(v)$ as the \emph{outdegree} of vertex $v$, where indegree and outdegree refer to the number of edges directed into vertex $v$ and the number of edges directed out of vertex $v$, respectively.
+Then, define a mapping $d:V|_\txt{local input} \to \mathbb{N}$ to specifiy the maximum number of connections that are allowed into a local input node.
+If $d(v)=1$ for each local input node, then exactly one edge must be directed into each local input node and each global output node. 
+If $d(v)>1$ for a local input node, this means that multiple inputs are accepted, such as the case for a multifidelity formulation.
+
+Next, let $H$ be the set of all nodes which are \emph{holes} in $M$ and define $H$ as
 \begin{equation}
-v \in H \txt{ if } t_\txt{node}(v) = \txt{`local input' or `global output'} \txt{ and } \txt{deg}^-(v) = 0.
+H=\big\{v \in M \big| t_\txt{node}(v) = \txt{`local input' or `global output'} \txt{ and } \txt{deg}^-(v) = 0\big\}.
 \end{equation}
-Next, let $L$ be the set of all nodes which are collisions in $M$ and define $L$ as
+Also, let $L$ be the set of all nodes which are \emph{collisions} in $M$ and define $L$ as
 \begin{equation}
-v \in L \txt{ if } t_\txt{node}(v) = \txt{`local input' or `global output'} \txt{ and } \txt{deg}^-(v) > 1.
+L = \big\{v \in M \big| t_\txt{node}(v) = \txt{`local input' or `global output'} \txt{ and } \txt{deg}^-(v) > d(v) \big\}.
 \end{equation}
-This means that exactly one edge must be directed into each local input node and each global output node.
-We also require $P$ to contain all of the global output nodes, meaning $P\subset M$ is a fundamental problem formulation if
+We may now define the graph representing the fundamental problem formulation $F=(V_F,E_F)$ as a subgraph of $M$ such that there are no holes or collisions.
+
+We also require $F$ to contain all of the global output nodes, meaning $F\subset M$ is a fundamental problem formulation if
 \begin{equation}
-\mathcal O \subset V_P \ \txt{ and } \ v \notin H, v \notin L \ \forall v \in V_P.
+\mathcal O \subset V_F \ \txt{ and } \ v \notin H, v \notin L \ \forall v \in V_F.
 \end{equation}
-
+((this is missing any edges!!!))
 \subsection{Obtaining the Fundamental Problem Formulation Graph}

### 2013-02-25 graph_syntax.tex
writing update

@@
 connection edge).
-Additional properties will be discussed is subsequent sections.
+Additional properties will be discussed in subsequent sections.
 
@@
 V_M & = & \mathcal I \cup \mathcal O \cup \left( \bigcup_{i = 1}^m V_{A_i} \right), \\
-E_M & = & C \cup \left( \bigcup_{i=1}^m E_{A_i} \right),
+E_M & = & C_M \cup \left( \bigcup_{i=1}^m E_{A_i} \right),
 \end{IEEEeqnarray*}
@@
 \begin{IEEEeqnarray*}{rl}
-c = & (v_1,v_2) \in C \txt{ if:} \\
+c = & (v_1,v_2) \in C_M \txt{ if:} \\
 & t_\txt{node}(v_1) = \txt{`local output' or `global input'}, \\
@@
 Before defining the the fundamental problem formulation graph, let us first define $\txt{deg}^-(v)$ as the \emph{indegree} $\txt{deg}^+(v)$ as the \emph{outdegree} of vertex $v$, where indegree and outdegree refer to the number of edges directed into vertex $v$ and the number of edges directed out of vertex $v$, respectively.
-Then, define a mapping $d:V|_\txt{local input} \to \mathbb{N}$ to specifiy the maximum number of connections that are allowed into a local input node.
-If $d(v)=1$ for each local input node, then exactly one edge must be directed into each local input node and each global output node. 
+Then, define a mapping $d:V \to \mathbb{N}$ to specifiy the maximum number of connections that are allowed into a local input node.
 If $d(v)>1$ for a local input node, this means that multiple inputs are accepted, such as the case for a multifidelity formulation.
 
-Next, let $H$ be the set of all nodes which are \emph{holes} in $M$ and define $H$ as
-\begin{equation}
-H=\big\{v \in M \big| t_\txt{node}(v) = \txt{`local input' or `global output'} \txt{ and } \txt{deg}^-(v) = 0\big\}.
-\end{equation}
-Also, let $L$ be the set of all nodes which are \emph{collisions} in $M$ and define $L$ as
-\begin{equation}
-L = \big\{v \in M \big| t_\txt{node}(v) = \txt{`local input' or `global output'} \txt{ and } \txt{deg}^-(v) > d(v) \big\}.
-\end{equation}
-We may now define the graph representing the fundamental problem formulation $F=(V_F,E_F)$ as a subgraph of $M$ such that there are no holes or collisions.
+% Next, let $H$ be the set of all nodes which are \emph{holes} in $M$ and define $H$ as
+% \begin{equation}
+% H=\big\{v \in M \big| t_\txt{node}(v) = \txt{`local input' or `global output'} \txt{ and } \txt{deg}^-(v) = 0\big\}.
+% \end{equation}
+% Also, let $L$ be the set of all nodes which are \emph{collisions} in $M$ and define $L$ as
+% \begin{equation}
+% L = \big\{v \in M \big| t_\txt{node}(v) = \txt{`local input' or `global output'} \txt{ and } \txt{deg}^-(v) > d(v) \big\}.
+% \end{equation}
+We may now define the \emph{fundamental problem formulation graph}, $F=(V_F,E_F)$, as a directed graph meeting the following conditions
+\begin{enumerate}
+\item[(1)] $\displaystyle{V_F = \mathcal I_F \cup \mathcal O \cup \left( \bigcup_{i \in \mathcal A} V_{A_i} \right),\ \mathcal I_F \subset \mathcal I}$
+\item[(2)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in \mathcal A} E_{A_i} \right)}$
+\item[(3)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`local input' or `global output,'}\quad 0 < \txt{deg}^-(v) \leq d(v)}$
+\end{enumerate}
+The set $\mathcal A$ is an index set containing the indices of the analysis blocks in $F$.
+The first requirement for $F$ is that $V_F$ be composed of all of the global outputs (which are given as required), the nodes from each analysis block in $F$, and any global input nodes that are needed, $\mathcal I_F$.
+The second requirement suggests that the edges in $F$ comprise the edges for each analysis block and the connections between them, the global inputs, and the global outputs.
+The final requirement is specific to the number of connection edges directed into a local input node or a global outputs node.
+If there are no connections directed inward, the node is called a \emph{hole}, and if more connections are directed in than are allowed, the node is called a \emph{collision}.
 
-We also require $F$ to contain all of the global output nodes, meaning $F\subset M$ is a fundamental problem formulation if
-\begin{equation}
-\mathcal O \subset V_F \ \txt{ and } \ v \notin H, v \notin L \ \forall v \in V_F.
-\end{equation}
-((this is missing any edges!!!))
 \subsection{Obtaining the Fundamental Problem Formulation Graph}
-In general, the fundamental problem formulation graph $P$ may not exist, or there may be multiple different solutions. Then the first problem is to determine if there exists a graph $P$.
-In practice, it is likely that this definition will not produce a unique solution. 
+In general, there may be multiple different graphs that satisfy the FPF conditions, though there may be none at all.
+Here, we describe a process for obtaining an FPF by starting with the MCG and removing nodes and edges until the FPF conditions are met.
+Then the problem is reduced to deciding which nodes and edges to remove, and this is centered on the local input nodes, which are treated as decision switches.
+The first step is to detect holes and remove the corresponding analysis blocks, which could not be used in a
+
 A collision represents a choice.
+Delete edges and analysis blocks so that there are no nodes or collisions.
 

### 2013-03-04 graph_syntax.tex
update for consolidation of node and edge types

@@
 The first property assigned to nodes and edges is the \emph{type}. The possible node types are:
-\begin{itemize}
-\item global input
-\item global output
-\item local input
-\item local output
-\item analysis start
-\item analysis finish
-\end{itemize}
-and the possible edge types are
-\begin{itemize}
-\item connection
-\item in
-\item analysis
-\item out
-\item ((consider eliminating `in' and `out' and using connection))
-\end{itemize}
+\begin{description}
+\item[input] The `input' node has specific rules for the number of edges that may be directed into it.
+\item[simple] The `simple' node has no additional properties and is used as a building block.
+\end{description}
+
+The possible edge types are:
+\begin{description}
+\item[connection] The `connection' edge represents the passing of data between analysis blocks.
+\item[data] The `data' edge represents the passing of data within an analysis block.
+\item[analysis] The `analysis' edge contains additional properties to represent the execution of an analysis block.
+\end{description}
 Then, considering sets $T_\txt{node}$ and $T_\txt{edge}$ to be sets containing the possible node types and edge types, respectively, we use mappings $t_\txt{node}:V \to T_\txt{node}$ and $t_\txt{edge}:E \to T_\txt{edge}$ to assign a type to each node and edge.
-Nodes and edges are named according to their type (e.g.
-connection edge).
+Nodes and edges are herein referred to by their type (e.g.
+`connection' edge).
 Additional properties will be discussed in subsequent sections.
@@
 This structure is derived from the need to represent each input and output individually while representing the actual analysis with a single node or edge.
-The `local input' nodes each represent a single input of the analysis block.
-Each of these nodes is connected to a single `analysis start' node via an `in' edge.
-Next the analysis start node is directed to the `analysis finish' node via the `analysis edge', which represents the actual analysis.
-Finally, the analysis finish node is directed to `local output' nodes via individual `out' edges.
+The `input' nodes each represent a single input of the analysis block.
+Each of these nodes is connected to a single `simple' node via a `data' edge, which represents the gathering of the inputs to begin the analysis.
+Next the `simple' node is directed to another `simple' node via the `analysis edge', which represents the actual analysis.
+Finally, this `simple' node is directed to several `simple' nodes via individual `data' edges, and this represents individual local outputs which may then be connected to other analysis blocks.
 \begin{figure}[htb!]
@@
 \end{figure}
-`Connection' edges connect local output nodes from one analysis block to either the local input nodes of other analysis blocks or global output node.
-In this way, the analysis block graph is the fundamental building block of the MDAO problem data flow.
+`Connection' edges connect the local outputs from one analysis block to input nodes. In this way, the analysis block graph is the fundamental building block of the MDAO problem data flow.
 
 \subsection{Maximal Connectivity Graph}
-Now we define the node property \emph{variable name}, which represents the unique identity of the variables passed between analysis blocks as represented by nodes of type local input, local output, global input, and global output.
-Let $N$ be the set containing each unique variable name (no multiplicity), and let $n:V \to N$ be a mapping that assigns a variable name to a node.
-
-Next, Consider an MDAO problem defined by a set of codes, global inputs, and global outputs, as represented by analysis blocks $A_i=(V_{A_i},E_{A_i}),i=1,\ldots,m$, the set of nodes $\mathcal I$, and the set of nodes $\mathcal O$, respectively.
+Next, Consider an MDAO problem defined by a set of codes, global inputs, global outputs.
+The codes are represented by analysis blocks $A_i=(V_{A_i},E_{A_i}), \ i=1,\ldots,m$, the global inputs are represented a set of `simple' nodes $I$, and the global outputs are represented by a set of `input' nodes $O$.
+We assume that $O$, $I$, and $A_i$ are given, and that any potential connection edge is given in the form of the set $C_M$.
 Then we may construct the \emph{maximal connectivity graph} $M=(V_M,E_M)$ as
 \begin{IEEEeqnarray*}{rCl}
-V_M & = & \mathcal I \cup \mathcal O \cup \left( \bigcup_{i = 1}^m V_{A_i} \right), \\
-E_M & = & C_M \cup \left( \bigcup_{i=1}^m E_{A_i} \right),
+V_M & = & I \cup O \cup \left( \bigcup_{i = 1}^m V_{A_i} \right), \\
+E_M & = & C_M \cup \left( \bigcup_{i=1}^m E_{A_i} \right).
 \end{IEEEeqnarray*}
-where $C$ is a set containing connection edges defined for $v_1,v_2 \in V_M$ as:
-\begin{IEEEeqnarray*}{rl}
-c = & (v_1,v_2) \in C_M \txt{ if:} \\
-& t_\txt{node}(v_1) = \txt{`local output' or `global input'}, \\
-& t_\txt{node}(v_2) = \txt{`local input' or `global output'},\txt{ and} \\
-& n(v_1) = n(v_2).
-\end{IEEEeqnarray*}
-The maximal connectivity graph contains all possible connections that may exist between the given analysis blocks, and these connections are made by connecting variables with the same name, which requires a consistent naming convention between analysis blocks.
+The maximal connectivity graph contains all possible connections that may exist between the given analysis blocks.
 
@@
 Before defining the the fundamental problem formulation graph, let us first define $\txt{deg}^-(v)$ as the \emph{indegree} $\txt{deg}^+(v)$ as the \emph{outdegree} of vertex $v$, where indegree and outdegree refer to the number of edges directed into vertex $v$ and the number of edges directed out of vertex $v$, respectively.
-Then, define a mapping $d:V \to \mathbb{N}$ to specifiy the maximum number of connections that are allowed into a local input node.
-If $d(v)>1$ for a local input node, this means that multiple inputs are accepted, such as the case for a multifidelity formulation.
+Then, define a mapping $d:V \to \mathbb{N}$ to specifiy the maximum number of connections that are allowed into an input node.
+If $d(v)>1$ for an input node, this means that multiple inputs are accepted, such as the case for a multifidelity formulation.
 
@@
 \begin{enumerate}
-\item[(1)] $\displaystyle{V_F = \mathcal I_F \cup \mathcal O \cup \left( \bigcup_{i \in \mathcal A} V_{A_i} \right),\ \mathcal I_F \subset \mathcal I}$
+\item[(1)] $\displaystyle{V_F = I_F \cup O \cup \left( \bigcup_{i \in \mathcal A} V_{A_i} \right),\ I_F \subset I}$
 \item[(2)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in \mathcal A} E_{A_i} \right)}$
-\item[(3)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`local input' or `global output,'}\quad 0 < \txt{deg}^-(v) \leq d(v)}$
+\item[(3)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`input,'}\quad 0 < \txt{deg}^-(v) \leq d(v)}$
 \end{enumerate}
 The set $\mathcal A$ is an index set containing the indices of the analysis blocks in $F$.
-The first requirement for $F$ is that $V_F$ be composed of all of the global outputs (which are given as required), the nodes from each analysis block in $F$, and any global input nodes that are needed, $\mathcal I_F$.
+The first requirement for $F$ is that $V_F$ be composed of all of the global outputs (which are given as required), the nodes from each analysis block in $F$, and any global input nodes that are needed, $I_F$.
 The second requirement suggests that the edges in $F$ comprise the edges for each analysis block and the connections between them, the global inputs, and the global outputs.
@@
 If there are no connections directed inward, the node is called a \emph{hole}, and if more connections are directed in than are allowed, the node is called a \emph{collision}.
+The final requirement is therefore a requirement on $C_F$.
 
@@
 Here, we describe a process for obtaining an FPF by starting with the MCG and removing nodes and edges until the FPF conditions are met.
-Then the problem is reduced to deciding which nodes and edges to remove, and this is centered on the local input nodes, which are treated as decision switches.
-The first step is to detect holes and remove the corresponding analysis blocks, which could not be used in a
+Then the problem is reduced to deciding which nodes and edges to remove, and this is centered on the input nodes, which are treated as decision switches.
+
+Then, mathematically, the FPF starts with $\mathcal A_1 = \{1,\ldots,m\}$, $I_{F,1} = I$, and $C_{F,1} = C_M$.
+
+\begin{enumerate}
+\item The first step is to detect holes and remove the corresponding analysis blocks, which could not be used because the required inputs are not supplied.
+\begin{equation}
+\forall i \in \mathcal A_1,\txt{ and }\forall v \in V_{A_i},\txt{ if deg}^-(v)=0 \txt{ then } i \in \mathcal B
+\end{equation}
+\begin{equation}
+\mathcal A_2 = \mathcal A_1 \setminus \mathcal B
+\end{equation}
+\end{enumerate}
 

### 2013-03-04 graph_syntax.tex
next update

@@
 Each ordered pair represents an edge starting at the node indicated by the first entry and directed to the node indicated by the second entry.
-Therefore, $E$ is a mapping from $V$ to $V$.
+While $E$ is effectively a mapping from $V$ to $V$, we reference this mapping as $e$.
 \begin{figure}[htb!]
@@
 \item The first step is to detect holes and remove the corresponding analysis blocks, which could not be used because the required inputs are not supplied.
+Any analysis blocks with holes may be stored in an index set $\mathcal H$ as
 \begin{equation}
-\forall i \in \mathcal A_1,\txt{ and }\forall v \in V_{A_i},\txt{ if deg}^-(v)=0 \txt{ then } i \in \mathcal B
+\forall i \in \mathcal A_1,\txt{ and }\forall v \in V_{A_i},\txt{ if deg}^-(v)=0 \txt{ then } i \in \mathcal H.
+\end{equation}
+Then an updated index set $\mathcal A_2$ is created using set difference notation as
+\begin{equation}
+\mathcal A_2 = \mathcal A_1 \setminus \mathcal H
+\end{equation}The connection edges directed from the removed analysis blocks are removed as
+\begin{equation}
+H = \{c \in C_{F,1} |c = (e^{-1}(v),v), \txt{ where } v \in V_{A_i} \txt{ for some } i \in \mathcal H\},
 \end{equation}
 \begin{equation}
-\mathcal A_2 = \mathcal A_1 \setminus \mathcal B
+C_{F,2} = C_{F,1} \setminus H
 \end{equation}
+Finally, the FPF is updated as
+\begin{equation}
+V_{F,2} = I_{F,2} \cup O \cup \left( \bigcup_{i \in \mathcal A_2} V_{A_i} \right),
+\end{equation}
+\begin{equation}
+E_{F,2} = C_{F,2} \cup \left( \bigcup_{i \in \mathcal A_2} E_{A_i} \right)
+\end{equation}
+This process may require multiple iterations because removing an analysis block may create holes upstream.
+It is assumed that the process as been repeated sufficiently such that $\mathcal A_2$ does not have any holes.
+
+\item The next step is to resolve conflicts. The set $C$ of input nodes containing conflicts is
+\begin{equation}
+C = \{v \in V_{F,2} | t_\txt{node}(v) = \txt{`input'} \txt{ and } \txt{deg}^-(v) > d(v)\}
+\end{equation}
+%\begin{equation}
+%for v \in c let 
+%\end{equation}
 \end{enumerate}
 
-A collision represents a choice.
-Delete edges and analysis blocks so that there are no nodes or collisions.
+\subsection{Classification of the FPF}
+
+%A collision represents a choice.
+%Delete edges and analysis blocks so that there are no nodes or collisions.
 

### 2013-03-05 graph_syntax.tex
update with graph theory

@@
 We begin by discussing the general syntax of graph theory and then expand upon this syntax to represent the data flow of an MDAO problem.
-A graph is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, which means that the elements of $E$ are 2--element subsets of $V$ \cite{Diestel2010}. 
-For a directed graph we construct $E$ as a set of ordered pairs instead of a set of sets.
+The present notation is adapted from Diestel \cite{Diestel2010}.
+A \emph{graph} is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, which means that the elements of $E$ are 2--element subsets of $V$.
+The set $V$ contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
+For a \emph{directed graph*} we construct $E$ as a set of ordered pairs instead of a set of sets.
 Each ordered pair represents an edge starting at the node indicated by the first entry and directed to the node indicated by the second entry.
-While $E$ is effectively a mapping from $V$ to $V$, we reference this mapping as $e$.
 \begin{figure}[htb!]
@@
 \end{IEEEeqnarray*}
+The edges directed out of $v \in V$ are given by $E(v)$ and the edges directed into $v$ are given by $E^{-1}(v)$.
+The \emph{indegree} is the number of edges directed in and is denoted as $\txt{deg}^-(v)$, and the \emph{outdegree} is the number of edges directed out and it is denoted as $\txt{deg}^+(v)$.
+Edge $e$ = $(x,y)$ may be referred to simply as $xy$.
+A \emph{path} is a nonempty graph $P = (V,E)$ with $V = \{x_0,x_1,\ldots,x_k\}$ and $E = \{x_0x_1,x_1x_2,\ldots,x_{k-1}x_k\}$, where $x_i$ are all distinct.
+A graph $G$ is \emph{connected} if any two of its vertices are linked by a path in $G$.
+A \emph{tree} is a connected and acyclic graph.
 
-We now assign additional properties to the nodes and edges so that they may represent the data flow of an MDAO problem. 
+The MDAO problem formulation concepts we wish to represent are:
+\begin{description}
+\item[global input] A global input is a variable that is taken as given and is used by multiple analyses.
+\item[design variable]
+\item[function call] A function call requires a set of inputs to be fulfilled and has certain properties, such as run time, which are independent of the number of outputs actually being used.
+\item[passing of distinct variables] Instead of connecting analyses, the individual variables as are connected
+\item[objective \& constraint] These are variables that mush be produced by the problem formulation for it to be valid
+\item[collision] A collision occurs when an input has multiple sources
+\item[hole] A hole occurs when an input has no sources
+\end{description}
+
+In order to represent the listed problem formulation concepts with a graph, we assign properties and rules for these nodes and edges.
 The first property assigned to nodes and edges is the \emph{type}. The possible node types are:
 \begin{description}
-\item[input] The `input' node has specific rules for the number of edges that may be directed into it.
-\item[simple] The `simple' node has no additional properties and is used as a building block.
+\item[variable node] The variable node represents passing of data.
+\item[model node] The model node represents the handling of data.
 \end{description}
@@
 \begin{description}
-\item[connection] The `connection' edge represents the passing of data between analysis blocks.
-\item[data] The `data' edge represents the passing of data within an analysis block.
-\item[analysis] The `analysis' edge contains additional properties to represent the execution of an analysis block.
+\item[fixed edge] A fixed edge may not be removed.
+\item[free edge] A free edge may be removed.
 \end{description}
+
+For these types of nodes and edges we impose the following rules:
+\begin{itemize}
+\item A model node can only have one edge directed to or from another model node.
+\item A model node can only have fixed edges directed in or out.
+\item A model node must have at least one edge directed in and at least one edge directed out.
+\item An objective or constraint is indicated by a variable node with no outgoing edges and with the incoming edges being directed from only other variable nodes.
+\item A global input is a variable node with no edges directed in and with at least two edges directed out to different variable nodes.
+\end{itemize}
+Additionally, the problem formulation concepts represented by the edges are described in Table \ref{t:edge representation}
+\begin{table}[h!]
+ \begin{center}
+  \caption{Problem formulation concept represented by an edge}
+  \label{t:edge representation}
+  \begin{tabular}{ccc} \hline 
+from node & to node & edge representation \\ \hline
+variable & variable & connection/passing of a variable\\
+variable & model & local input \\
+model & variable & local output \\
+model & model & function call \\ \hline
+  \end{tabular}
+ \end{center}
+ \vspace{-15pt}
+\end{table}
+
+Any tree containing only fixed edges and at least one model node is an analysis block.
+
 Then, considering sets $T_\txt{node}$ and $T_\txt{edge}$ to be sets containing the possible node types and edge types, respectively, we use mappings $t_\txt{node}:V \to T_\txt{node}$ and $t_\txt{edge}:E \to T_\txt{edge}$ to assign a type to each node and edge.
@@
 `connection' edge).
+
 Additional properties will be discussed in subsequent sections.
@@
 This structure is derived from the need to represent each input and output individually while representing the actual analysis with a single node or edge.
-The `input' nodes each represent a single input of the analysis block.
-Each of these nodes is connected to a single `simple' node via a `data' edge, which represents the gathering of the inputs to begin the analysis.
-Next the `simple' node is directed to another `simple' node via the `analysis edge', which represents the actual analysis.
-Finally, this `simple' node is directed to several `simple' nodes via individual `data' edges, and this represents individual local outputs which may then be connected to other analysis blocks.
+The variable nodes each represent a single input of the analysis block.
+Each of these nodes is connected to a single model node via a `data' edge, which represents the gathering of the inputs to begin the analysis.
+Next the model node is directed to another model node via the `analysis edge', which represents the actual analysis.
+Finally, this model node is directed to several model nodes via individual `data' edges, and this represents individual local outputs which may then be connected to other analysis blocks.
 \begin{figure}[htb!]
@@
 Next, Consider an MDAO problem defined by a set of codes, global inputs, global outputs.
-The codes are represented by analysis blocks $A_i=(V_{A_i},E_{A_i}), \ i=1,\ldots,m$, the global inputs are represented a set of `simple' nodes $I$, and the global outputs are represented by a set of `input' nodes $O$.
+The codes are represented by analysis blocks $A_i=(V_{A_i},E_{A_i}), \ i=1,\ldots,m$, the global inputs are represented a set of `simple' nodes $I$, and the global outputs are represented by a set of variable nodes $O$.
 We assume that $O$, $I$, and $A_i$ are given, and that any potential connection edge is given in the form of the set $C_M$.

### 2013-03-11 graph_syntax.tex

@@
 \section{Graph-based Problem Formulation Syntax}
+\subsection{Graph Theory Basics}
 We begin by discussing the general syntax of graph theory and then expand upon this syntax to represent the data flow of an MDAO problem.
@@
 The \emph{indegree} is the number of edges directed in and is denoted as $\txt{deg}^-(v)$, and the \emph{outdegree} is the number of edges directed out and it is denoted as $\txt{deg}^+(v)$.
-Edge $e$ = $(x,y)$ may be referred to simply as $xy$.
+Edge $e$ = $(x,y)$ may be referred to simply as $xy$ and we say $y = E(x)$.
+Also, $E(E(v))$ is denoted as $E^2(v)$, and likewise for additonal levels.
+
  A \emph{path} is a nonempty graph $P = (V,E)$ with $V = \{x_0,x_1,\ldots,x_k\}$ and $E = \{x_0x_1,x_1x_2,\ldots,x_{k-1}x_k\}$, where $x_i$ are all distinct.
@@
 
+\subsection{Problem Formulation Concepts to Represent}
 The MDAO problem formulation concepts we wish to represent are:
@@
 \item[passing of distinct variables] Instead of connecting analyses, the individual variables as are connected
-\item[objective \& constraint] These are variables that mush be produced by the problem formulation for it to be valid
+\item[objective \& constraint] These are variables that must be produced by the problem formulation for it to be valid
 \item[collision] A collision occurs when an input has multiple sources
 \item[hole] A hole occurs when an input has no sources
+\item[coupling] Couple is the mutual dependence between a set of analysis
+\item[multi--fidelity] It may be desirable for the same variable to be calculated by separate analyses and to use both results.
 \end{description}
 
-In order to represent the listed problem formulation concepts with a graph, we assign properties and rules for these nodes and edges.
+\subsection{Graph Theory Representation of a Problem Formulation}
+In order to represent the listed problem formulation concepts with a graph, we assign properties and rules to nodes and edges.
 The first property assigned to nodes and edges is the \emph{type}. The possible node types are:
@@
 \end{description}
-
+To keep track of which nodes and edges are which type, let $T_\txt{node}$ and $T_\txt{edge}$ be sets containing the possible node types and edge types, respectively, and then define mappings $t_\txt{node}:V \to T_\txt{node}$ and $t_\txt{edge}:E \to T_\txt{edge}$ to assign a type to each node and edge.
 For these types of nodes and edges we impose the following rules:
@@
 \end{itemize}
-Additionally, the problem formulation concepts represented by the edges are described in Table \ref{t:edge representation}
+Additional problem formulation concepts represented by the edges are described in Table \ref{t:edge representation}
 \begin{table}[h!]
@@
 
-Any tree containing only fixed edges and at least one model node is an analysis block.
-
-Then, considering sets $T_\txt{node}$ and $T_\txt{edge}$ to be sets containing the possible node types and edge types, respectively, we use mappings $t_\txt{node}:V \to T_\txt{node}$ and $t_\txt{edge}:E \to T_\txt{edge}$ to assign a type to each node and edge.
-Nodes and edges are herein referred to by their type (e.g.
-`connection' edge).
-
-Additional properties will be discussed in subsequent sections.
-
-Next, we introduce the graph $A$ which represents a single analysis block broken down into its individual inputs and outputs.
-An example graph is shown in Fig.~\ref{f:analysis block} to demonstrate the basic structure.
+Finally, any tree containing only fixed edges and at least one model node is an \emph{analysis block}, which is used to represent a contributing analysis.
+An example analysis block is shown in Fig.~\ref{f:analysis block} to demonstrate the basic structure.
 This structure is derived from the need to represent each input and output individually while representing the actual analysis with a single node or edge.
-The variable nodes each represent a single input of the analysis block.
-Each of these nodes is connected to a single model node via a `data' edge, which represents the gathering of the inputs to begin the analysis.
-Next the model node is directed to another model node via the `analysis edge', which represents the actual analysis.
-Finally, this model node is directed to several model nodes via individual `data' edges, and this represents individual local outputs which may then be connected to other analysis blocks.
+The variable nodes each represent a single input or output of the analysis block.
+Each of these nodes is connected to a single model node via a fixed edge, which represents the gathering of the inputs to begin the analysis or the disseminating of outputs after the analysis.
+Finally, free edges are used to connect the variable nodes of the analysis block to other variables nodes outside of the analysis block.
+In this way, the analysis block graph is the fundamental building block of the MDAO problem data flow.
 \begin{figure}[htb!]
@@
 \end{figure}
-`Connection' edges connect the local outputs from one analysis block to input nodes. In this way, the analysis block graph is the fundamental building block of the MDAO problem data flow.
+%`Connection' edges connect the local outputs from one analysis block to input nodes.
 
@@
 Next, Consider an MDAO problem defined by a set of codes, global inputs, global outputs.
-The codes are represented by analysis blocks $A_i=(V_{A_i},E_{A_i}), \ i=1,\ldots,m$, the global inputs are represented a set of `simple' nodes $I$, and the global outputs are represented by a set of variable nodes $O$.
-We assume that $O$, $I$, and $A_i$ are given, and that any potential connection edge is given in the form of the set $C_M$.
+The codes are represented by analysis blocks $A_i=(V_{A_i},E_{A_i}), \ i=1,\ldots,m$, the global inputs are represented a set of variable nodes $I$, and the global outputs are represented by a set of variable nodes $O$.
+We assume that $O$, $I$, and $A_i$ are given, and that any potential connection between variables is given in the form of the free edges in the set $C_M$.
 Then we may construct the \emph{maximal connectivity graph} $M=(V_M,E_M)$ as
@@
 \subsection{Fundamental Problem Formulation Graph}
-Before defining the the fundamental problem formulation graph, let us first define $\txt{deg}^-(v)$ as the \emph{indegree} $\txt{deg}^+(v)$ as the \emph{outdegree} of vertex $v$, where indegree and outdegree refer to the number of edges directed into vertex $v$ and the number of edges directed out of vertex $v$, respectively.
-Then, define a mapping $d:V \to \mathbb{N}$ to specifiy the maximum number of connections that are allowed into an input node.
-If $d(v)>1$ for an input node, this means that multiple inputs are accepted, such as the case for a multifidelity formulation.
+Before defining the the fundamental problem formulation graph, let us first define the \emph{upper degree limit} 
+\begin{equation}
+\txt{deg}_u^-(v):V \to \mathbb{N}
+\end{equation} 
+and the \emph{lower degree limit}
+\begin{equation}
+\txt{deg}_l^-(v):V \to \mathbb{N}.
+\end{equation}
+These user-specified limits set the number of edges that may be directed into a node for a problem formulation to be valid.
+For example, a variable node $v$ will have $\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$, unless it is a multi--fidelity variable, in which case the user may specify some upper limit higher than one.
 
-% Next, let $H$ be the set of all nodes which are \emph{holes} in $M$ and define $H$ as
-% \begin{equation}
-% H=\big\{v \in M \big| t_\txt{node}(v) = \txt{`local input' or `global output'} \txt{ and } \txt{deg}^-(v) = 0\big\}.
-% \end{equation}
-% Also, let $L$ be the set of all nodes which are \emph{collisions} in $M$ and define $L$ as
-% \begin{equation}
-% L = \big\{v \in M \big| t_\txt{node}(v) = \txt{`local input' or `global output'} \txt{ and } \txt{deg}^-(v) > d(v) \big\}.
-% \end{equation}
 We may now define the \emph{fundamental problem formulation graph}, $F=(V_F,E_F)$, as a directed graph meeting the following conditions
@@
 \item[(2)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in \mathcal A} E_{A_i} \right)}$
-\item[(3)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`input,'}\quad 0 < \txt{deg}^-(v) \leq d(v)}$
+\item[(3)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'}\quad \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
 \end{enumerate}
@@
 The first requirement for $F$ is that $V_F$ be composed of all of the global outputs (which are given as required), the nodes from each analysis block in $F$, and any global input nodes that are needed, $I_F$.
-The second requirement suggests that the edges in $F$ comprise the edges for each analysis block and the connections between them, the global inputs, and the global outputs.
-The final requirement is specific to the number of connection edges directed into a local input node or a global outputs node.
-If there are no connections directed inward, the node is called a \emph{hole}, and if more connections are directed in than are allowed, the node is called a \emph{collision}.
+The second requirement suggests that the edges in $F$ comprise the edges for each analysis block and the edges between them, the global inputs, and the global outputs.
+The final requirement is specific to the number of edges directed into a variable.
+If there are no edges directed inward, the node is called a \emph{hole}, and if more connections are directed in than are allowed, the node is called a \emph{collision}; a hole is allowed if $\txt{deg}_l^-(v)=0$.
 The final requirement is therefore a requirement on $C_F$.
@@
 In general, there may be multiple different graphs that satisfy the FPF conditions, though there may be none at all.
-Here, we describe a process for obtaining an FPF by starting with the MCG and removing nodes and edges until the FPF conditions are met.
-Then the problem is reduced to deciding which nodes and edges to remove, and this is centered on the input nodes, which are treated as decision switches.
+Here, we describe a process for obtaining an FPF by starting with the MCG and disconnecting free edges until the FPF conditions are met.
+Then the problem is reduced to deciding which free edges to remove.
 
-Then, mathematically, the FPF starts with $\mathcal A_1 = \{1,\ldots,m\}$, $I_{F,1} = I$, and $C_{F,1} = C_M$.
+%Then, mathematically, the FPF starts with $\mathcal A_1 = \{1,\ldots,m\}$, $I_{F,1} = I$, and $C_{F,1} = C_M$.
+Then, mathematically, the FPF starts with $C_{F,0} = C_M$.
+First address the nodes where $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ then the ones where $\txt{deg}^-(v) > \txt{deg}_u^-(v)$
 
 \begin{enumerate}
-\item The first step is to detect holes and remove the corresponding analysis blocks, which could not be used because the required inputs are not supplied.
+\item The first step is to detect holes and disconnect the free edges following them.
+These free edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs.
+The set of variable nodes which are holes is created as
+\begin{equation}
+H = \{v \in V | \txt{deg}^-(v) < \txt{deg}_l^-(v) \}
+\end{equation}
+Then the updated set of edges is
+\begin{equation}
+C_{F,1} = C_{F,0} \setminus \{\ e \in E, \ e=xy | x=E^3(v) \txt{ for } v \in H\}
+\end{equation}
+%deg-v is now recaculated for the new C. 
+This step is demonstrated by Fig.~\ref{f:holes}.
+\end{enumerate}
+\begin{figure}[htb!]
+	\begin{center}
+	\includegraphics[width=3in]{images/analysis_block_hole}
+	\end{center}
+	\vspace{-20pt}
+\caption{Example variable node indicating a hole.}
+\label{f:holes}
+\end{figure}
+
+%\begin{enumerate}
+%\item The first step is to detect holes and remove the corresponding analysis blocks, which could not be used because the required inputs are not supplied.
 Any analysis blocks with holes may be stored in an index set $\mathcal H$ as
-\begin{equation}
-\forall i \in \mathcal A_1,\txt{ and }\forall v \in V_{A_i},\txt{ if deg}^-(v)=0 \txt{ then } i \in \mathcal H.
-\end{equation}
-Then an updated index set $\mathcal A_2$ is created using set difference notation as
-\begin{equation}
-\mathcal A_2 = \mathcal A_1 \setminus \mathcal H
-\end{equation}The connection edges directed from the removed analysis blocks are removed as
-\begin{equation}
-H = \{c \in C_{F,1} |c = (e^{-1}(v),v), \txt{ where } v \in V_{A_i} \txt{ for some } i \in \mathcal H\},
-\end{equation}
-\begin{equation}
-C_{F,2} = C_{F,1} \setminus H
-\end{equation}
-Finally, the FPF is updated as
-\begin{equation}
-V_{F,2} = I_{F,2} \cup O \cup \left( \bigcup_{i \in \mathcal A_2} V_{A_i} \right),
-\end{equation}
-\begin{equation}
-E_{F,2} = C_{F,2} \cup \left( \bigcup_{i \in \mathcal A_2} E_{A_i} \right)
-\end{equation}
-This process may require multiple iterations because removing an analysis block may create holes upstream.
+%\begin{equation}
+%\forall i \in \mathcal A_1,\txt{ and }\forall v \in V_{A_i},\txt{ if deg}^-(v)=0 \txt{ then } i \in \mathcal H.
+%\end{equation}
+%Then an updated index set $\mathcal A_2$ is created using set difference notation as
+%\begin{equation}
+%\mathcal A_2 = \mathcal A_1 \setminus \mathcal H
+%\end{equation}The connection edges directed from the removed analysis blocks are removed as
+%\begin{equation}
+%H = \{c \in C_{F,1} |c = (e^{-1}(v),v), \txt{ where } v \in V_{A_i} \txt{ for some } i \in \mathcal H\},
+%\end{equation}
+%\begin{equation}
+%C_{F,2} = C_{F,1} \setminus H
+%\end{equation}
+%Finally, the FPF is updated as
+%\begin{equation}
+%V_{F,2} = I_{F,2} \cup O \cup \left( \bigcup_{i \in \mathcal A_2} V_{A_i} \right),
+%\end{equation}
+%\begin{equation}
+%E_{F,2} = C_{F,2} \cup \left( \bigcup_{i \in \mathcal A_2} E_{A_i} \right)
+%\end{equation}
+%This process may require multiple iterations because removing an analysis block may create holes upstream.
 It is assumed that the process as been repeated sufficiently such that $\mathcal A_2$ does not have any holes.
-
-\item The next step is to resolve conflicts. The set $C$ of input nodes containing conflicts is
-\begin{equation}
-C = \{v \in V_{F,2} | t_\txt{node}(v) = \txt{`input'} \txt{ and } \txt{deg}^-(v) > d(v)\}
-\end{equation}
+%
+%\item The next step is to resolve conflicts. The set $C$ of input nodes containing conflicts is
 %\begin{equation}
-%for v \in c let 
+%C = \{v \in V_{F,2} | t_\txt{node}(v) = \txt{`input'} \txt{ and } \txt{deg}^-(v) > d(v)\}
 %\end{equation}
-\end{enumerate}
+%%\begin{equation}
+%%for v \in c let 
+%%\end{equation}
+%\end{enumerate}
 

### 2013-03-14 graph_syntax.tex
most of suggested major changes

@@
-\section{Graph-based Problem Formulation Syntax}
+\section{Graph--based Problem Formulation Syntax}
+A problem formulation will be represented by nodes and edges which will be assigned certain properties and rules.
+There are many MDAO problem formulation concepts that we wish to represent.
+An analysis code should be represented by a distinct structure to distinguish between the input, output, and function evaluation, processes.
+The variables passed between analysis codes should be represented separately to provide insight and control over the dataflow.
+For example, it is important to distinguish between the case where two analysis codes are supplying the same variable to a third code, or the case where the two analysis codes are providing different variables to the third code.
+Local inputs, local outputs, and design variables(??) should be distinguished from global inputs, which are input that are used by multiple analysis codes.
+Objectives and constaints should be represented as separate from local variables because they must be provided in order for the problem formulation to be valid.
+The coupling between sets of analysis codes is important to capture, and multi--fidelity analysis should also be allowed.
+
 \subsection{Graph Theory Basics}
@@
 Each ordered pair represents an edge starting at the node indicated by the first entry and directed to the node indicated by the second entry.
+Edge $e$ = $(x,y)$ may be referred to simply as $xy$ and we say $y = E(x)$.
 \begin{figure}[htb!]
@@
 \end{IEEEeqnarray*}
-The edges directed out of $v \in V$ are given by $E(v)$ and the edges directed into $v$ are given by $E^{-1}(v)$.
-The \emph{indegree} is the number of edges directed in and is denoted as $\txt{deg}^-(v)$, and the \emph{outdegree} is the number of edges directed out and it is denoted as $\txt{deg}^+(v)$.
-Edge $e$ = $(x,y)$ may be referred to simply as $xy$ and we say $y = E(x)$.
-Also, $E(E(v))$ is denoted as $E^2(v)$, and likewise for additonal levels.
 
- A \emph{path} is a nonempty graph $P = (V,E)$ with $V = \{x_0,x_1,\ldots,x_k\}$ and $E = \{x_0x_1,x_1x_2,\ldots,x_{k-1}x_k\}$, where $x_i$ are all distinct.
-A graph $G$ is \emph{connected} if any two of its vertices are linked by a path in $G$.
-A \emph{tree} is a connected and acyclic graph.
-
-\subsection{Problem Formulation Concepts to Represent}
-The MDAO problem formulation concepts we wish to represent are:
-\begin{description}
-\item[global input] A global input is a variable that is taken as given and is used by multiple analyses.
-\item[design variable]
-\item[function call] A function call requires a set of inputs to be fulfilled and has certain properties, such as run time, which are independent of the number of outputs actually being used.
-\item[passing of distinct variables] Instead of connecting analyses, the individual variables as are connected
-\item[objective \& constraint] These are variables that must be produced by the problem formulation for it to be valid
-\item[collision] A collision occurs when an input has multiple sources
-\item[hole] A hole occurs when an input has no sources
-\item[coupling] Couple is the mutual dependence between a set of analysis
-\item[multi--fidelity] It may be desirable for the same variable to be calculated by separate analyses and to use both results.
-\end{description}
+% \subsection{Problem Formulation Concepts to Represent}
+% The MDAO problem formulation concepts we wish to represent are:
+% \begin{description}
+% \item[global input] A global input is a variable that is taken as given and is used by multiple analyses.
+% \item[design variable]
+% \item[function call] A function call requires a set of inputs to be fulfilled and has certain properties, such as run time, which are independent of the number of outputs actually being used.
+% \item[passing of distinct variables] Instead of connecting analyses, the individual variables as are connected
+% \item[objective \& constraint] These are variables that must be produced by the problem formulation for it to be valid
+% \item[collision] A collision occurs when an input has multiple sources
+% \item[hole] A hole occurs when an input has no sources
+% \item[coupling] Couple is the mutual dependence between a set of analysis
+% \item[multi--fidelity] It may be desirable for the same variable to be calculated by separate analyses and to use both results.
+% \end{description}
 
@@
 In order to represent the listed problem formulation concepts with a graph, we assign properties and rules to nodes and edges.
-The first property assigned to nodes and edges is the \emph{type}. The possible node types are:
+The first property assigned to nodes and edges is the \emph{type}. \\
+The possible node types are:
 \begin{description}
@@
 \end{description}
-
 The possible edge types are:
@@
 \end{description}
-To keep track of which nodes and edges are which type, let $T_\txt{node}$ and $T_\txt{edge}$ be sets containing the possible node types and edge types, respectively, and then define mappings $t_\txt{node}:V \to T_\txt{node}$ and $t_\txt{edge}:E \to T_\txt{edge}$ to assign a type to each node and edge.
-For these types of nodes and edges we impose the following rules:
+Model nodes must be used with the following rules:
 \begin{itemize}
@@
 \item A model node must have at least one edge directed in and at least one edge directed out.
-\item An objective or constraint is indicated by a variable node with no outgoing edges and with the incoming edges being directed from only other variable nodes.
-\item A global input is a variable node with no edges directed in and with at least two edges directed out to different variable nodes.
 \end{itemize}
-Additional problem formulation concepts represented by the edges are described in Table \ref{t:edge representation}
+The MDO problem formulation concepts represented by nodes are given in Table \ref{t:node representation}, and the concepts represented by nodes are given in Table \ref{t:edge representation}.
+% \begin{itemize}
+% \item An objective or constraint is indicated by a variable node with no outgoing edges and with the incoming edges being directed from only other variable nodes.
+% \item A global input is a variable node with no edges directed in and with at least two edges directed out to different variable nodes.
+% \item A design variable is represented by a variable node with no edges directed in and on fixed edge directed out.
+% \end{itemize}
+\begin{table}[h!]
+ \begin{center}
+  \caption{Problem formulation concept represented by a node}
+  \label{t:node representation}
+  \begin{tabular}{ccc} \hline 
+edges directed in & edges directed out & node representation \\ \hline
+only free edges & none & objective or constraint\\
+none & at least two free edges & global input \\
+none & one fixed edge & design variable \\ \hline
+  \end{tabular}
+ \end{center}
+ \vspace{-15pt}
+\end{table}
 \begin{table}[h!]
@@
 
-Finally, any tree containing only fixed edges and at least one model node is an \emph{analysis block}, which is used to represent a contributing analysis.
-An example analysis block is shown in Fig.~\ref{f:analysis block} to demonstrate the basic structure.
+Finally, an analysis code is represented by an \emph{analysis block}, which is a collection of model nodes, variable nodes, and fixed edges with a specific structure.
 This structure is derived from the need to represent each input and output individually while representing the actual analysis with a single node or edge.
@@
 Finally, free edges are used to connect the variable nodes of the analysis block to other variables nodes outside of the analysis block.
-In this way, the analysis block graph is the fundamental building block of the MDAO problem data flow.
+In this way, the analysis block graph is a fundamental building block of the MDAO problem data flow.
 \begin{figure}[htb!]
@@
 	\end{center}
-	\vspace{-20pt}
-\caption{Example directed graph. The each node type and edge type is labeled, with nodes being label in italics.}
+	\vspace{-10pt}
+\caption{Example analysis block. The each node type and edge type is labeled in italics and annotated parenthetically.}
 \label{f:analysis block}
@@
 
-\subsection{Maximal Connectivity Graph}
-Next, Consider an MDAO problem defined by a set of codes, global inputs, global outputs.
-The codes are represented by analysis blocks $A_i=(V_{A_i},E_{A_i}), \ i=1,\ldots,m$, the global inputs are represented a set of variable nodes $I$, and the global outputs are represented by a set of variable nodes $O$.
-We assume that $O$, $I$, and $A_i$ are given, and that any potential connection between variables is given in the form of the free edges in the set $C_M$.
-Then we may construct the \emph{maximal connectivity graph} $M=(V_M,E_M)$ as
-\begin{IEEEeqnarray*}{rCl}
-V_M & = & I \cup O \cup \left( \bigcup_{i = 1}^m V_{A_i} \right), \\
-E_M & = & C_M \cup \left( \bigcup_{i=1}^m E_{A_i} \right).
-\end{IEEEeqnarray*}
-The maximal connectivity graph contains all possible connections that may exist between the given analysis blocks.
+\section{Graph--based Representation of the Fundamental Problem Formulation}
+Next we provide the meaningful graphs that can be created from the suggested syntax.
+The first graph is the \emph{maximal connectivity graph} which represents the full potential of all the analysis codes being considered, i.e.
+each analysis code is represented by an analysis block, and each potential connections between variables is represented by a free edge.
+The second graph that may be represented is \emph{fundamental problem formulation graph}, which is the graph with the fewest number of edges and nodes needed to provide a valid problem formulation, as disscussed subsequently.
+Finally, a \emph{solution problem formulation} may be represented by including additional edges and node times to represent the optimization architecture, though this is beyond the scope of this paper.
 
-\subsection{Fundamental Problem Formulation Graph}
-Before defining the the fundamental problem formulation graph, let us first define the \emph{upper degree limit} 
+The relationship between these three graphs is depicted in Figs.~\ref{f:tree} and \ref{f:hourglass}.
+The tree diagram demonstrates the fact that it is generally possible to obtain multiple FPFs from a single maximal connectivity graph.
+This  may correspond to different down--selections of analysis codes, different connections between them, or both.
+Then, for each FPF, different SPFs may be obtained by implementing different optimization architectures.
+The hourglass shape indicates that the FPF graph is the smallest graph that may represent a specific problem formulation.
+The FPF is obtained from the MCG by removing nodes and edges, and the SPF is obtained from the FPF by adding nodes and edges.
+\begin{figure}[htb!]
+	\centering
+	\subfigure[number of possible graphs]{
+	\includegraphics[width=2.0in]{images/tree}
+	\label{f:tree}
+	}
+	\subfigure[graph size]{
+	\includegraphics[width=2.0in]{images/hourglass}
+	\label{f:hourglass}
+	}
+\caption{The relationship between the MCG, FPF, and SPF.}
+\end{figure}
+
+In this section we introduce the remaining necessary notion and give graph--theory based definitions of the MCG and FPF, and then process to demonstrate how the FPF is obtained from the MCG.
+
+\subsection{Additional Notation}
+For node $v \in V$ the edges directed out are given by $E(v)$ and the edges directed into $v$ are given by $E^{-1}(v)$; $E(E(v))$ is denoted as $E^2(v)$, and likewise for additional levels. 
+The \emph{indegree} of a node is the number of edges directed in and is denoted as $\txt{deg}^-(v)$, and the \emph{outdegree} is the number of edges directed out and it is denoted as $\txt{deg}^+(v)$.
+We also define the \emph{upper indegree limit} 
 \begin{equation}
@@
 \end{equation} 
-and the \emph{lower degree limit}
+and the \emph{lower indegree limit}
 \begin{equation}
@@
 
-We may now define the \emph{fundamental problem formulation graph}, $F=(V_F,E_F)$, as a directed graph meeting the following conditions
+To keep track of which nodes and edges are which type, let $T_\txt{node}$ and $T_\txt{edge}$ be sets containing the possible node types and edge types, respectively, and then define mappings $t_\txt{node}:V \to T_\txt{node}$ and $t_\txt{edge}:E \to T_\txt{edge}$ to assign a type to each node and edge.
+Then, for example, for a variable node $v$ we have $t_\txt{node}(v) = \txt{`variable'}$.
+
+ % A \emph{path} is a nonempty graph $P = (V,E)$ with $V = \{x_0,x_1,\ldots,x_k\}$ and $E = \{x_0x_1,x_1x_2,\ldots,x_{k-1}x_k\}$, where $x_i$ are all distinct.
+A graph $G$ is \emph{connected} if any two of its vertices are linked by a path in $G$.
+A \emph{tree} is a connected and acyclic graph.
+%%% this needs to be changed to fit the definition using a directed graph
+
+\subsection{Maximal Connectivity Graph}
+To construct the maximal connectivity graph, we assume that a set of codes, global inputs, and objectives and constraints (collectively called global outputs).
+The codes are represented by analysis blocks $A_i=(V_{A_i},E_{A_i}), \ i=1,\ldots,m$, the global inputs are represented a set of variable nodes $I$, and the global outputs are represented by a set of variable nodes $O$.
+We assume that $O$, $I$, and $A_i$ are given, and that any potential connection between variables is given in the form of the free edges in the set $C_M$.
+Then we may construct the maximal connectivity graph $M=(V_M,E_M)$ as
+\begin{IEEEeqnarray*}{rCl}
+V_M & = & I \cup O \cup \left( \bigcup_{i = 1}^m V_{A_i} \right), \\
+E_M & = & C_M \cup \left( \bigcup_{i=1}^m E_{A_i} \right),
+\end{IEEEeqnarray*}
+The MCG $M$ is uniquely determined by the given set of analysis blocks, the required outputs, and the given global inputs.
+In the cases where the set of global inputs $I$ is not known a priori, the process of obtaining the FPF will reveal the required inputs, as discussed subsequently.
+
+\subsection{Fundamental Problem Formulation Graph}
+We now define the fundamental problem formulation graph, $F=(V_F,E_F)$, as a directed graph meeting the following conditions
 \begin{enumerate}
@@
 \end{enumerate}
-The set $\mathcal A$ is an index set containing the indices of the analysis blocks in $F$.
-The first requirement for $F$ is that $V_F$ be composed of all of the global outputs (which are given as required), the nodes from each analysis block in $F$, and any global input nodes that are needed, $I_F$.
+The set $\mathcal A$ is an index set containing the indices of the analysis blocks in $F$; for the case where all of the analysis blocks are used, we would have $\mathcal A = \{1,2,\ldots,m\}$.
+The first requirement for $F$ is that $V_F$ be composed of all of the global outputs (the required objectives and constraints), the nodes from each analysis block in $F$, and any global input nodes that are needed, $I_F$.
 The second requirement suggests that the edges in $F$ comprise the edges for each analysis block and the edges between them, the global inputs, and the global outputs.
@@
 This step is demonstrated by Fig.~\ref{f:holes}.
+\item (())
 \end{enumerate}
@@
 
-\subsection{Classification of the FPF}
+%\subsection{Classification of the FPF}
 

### 2013-03-15 graph_syntax.tex
updated introduction and started editing a couple of lower sections

@@
-\section{Graph--based Problem Formulation Syntax}
-A problem formulation will be represented by nodes and edges which will be assigned certain properties and rules.
-There are many MDAO problem formulation concepts that we wish to represent.
-An analysis code should be represented by a distinct structure to distinguish between the input, output, and function evaluation, processes.
-The variables passed between analysis codes should be represented separately to provide insight and control over the dataflow.
-For example, it is important to distinguish between the case where two analysis codes are supplying the same variable to a third code, or the case where the two analysis codes are providing different variables to the third code.
-Local inputs, local outputs, and design variables(??) should be distinguished from global inputs, which are input that are used by multiple analysis codes.
-Objectives and constaints should be represented as separate from local variables because they must be provided in order for the problem formulation to be valid.
-The coupling between sets of analysis codes is important to capture, and multi--fidelity analysis should also be allowed.
+\section{New Graph--based Syntax}
+The ultimate goal of the new graph--based syntax presented here is to be able to 
+fully describe the fundamental problem formulation for a complex design task, 
+while still being able to accomodate the more specific case when a solution 
+strategy is applied to the problem formulation. In order to achieve that goal 
+the graph needs to accomodate a number of constructs from MDAO problem:
+
+\begin{itemize}
+    \item Analysis tools 
+    \item Local and global variables 
+    \item Coupling between analyses
+    \item Design variables
+    \item Objective or objectives 
+    \item Local and global Constraints
+    \item Iterative loops from solvers and optimizers
+\end{itemize}
+
+The following section provides an overview of graph theory basics, then describes 
+in detail the different node and edge types, as well as rules governing allowable 
+egdegs between nodes, used as the foundation of the new graph syntax. These basic graph 
+primatives are them mapped to the MDAO constructs stated above, and used to 
+build three distinct graph forms:
+
+\begin{itemize}
+    \item Maximal Connectivity
+    \item Fundamental Problem Formulation 
+    \item Specific Solution
+\end{itemize}
+
+Each of these forms serves a different purpose in the design process, and we 
+present some transformations that can be applied to transition 
+between any two of the three graph forms.
 
 \subsection{Graph Theory Basics}
-We begin by discussing the general syntax of graph theory and then expand upon this syntax to represent the data flow of an MDAO problem.
-The present notation is adapted from Diestel \cite{Diestel2010}.
-A \emph{graph} is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, which means that the elements of $E$ are 2--element subsets of $V$.
-The set $V$ contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
-For a \emph{directed graph*} we construct $E$ as a set of ordered pairs instead of a set of sets.
-Each ordered pair represents an edge starting at the node indicated by the first entry and directed to the node indicated by the second entry.
-Edge $e$ = $(x,y)$ may be referred to simply as $xy$ and we say $y = E(x)$.
+The notation used in this work is adapted from Diestel \cite{Diestel2010}. 
+A \emph{graph} is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, 
+which means that the elements of $E$ are 2--element subsets of $V$. The set $V$ 
+contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
+For a \emph{directed graph*} we construct $E$ as a set of ordered pairs instead 
+of a set of sets. Each ordered pair represents an edge starting at the node 
+indicated by the first entry and directed to the node indicated by the second 
+entry. Edge $e$ = $(x,y)$ may be referred to simply as $xy$ and we say $y = E(x)$.
+
 \begin{figure}[htb!]
@@
 
-\subsection{Graph Theory Representation of a Problem Formulation}
-In order to represent the listed problem formulation concepts with a graph, we assign properties and rules to nodes and edges.
-The first property assigned to nodes and edges is the \emph{type}. \\
-The possible node types are:
+\subsection{MDAO Graph Primatives}
+There are three node types:  
 \begin{description}
-\item[variable node] The variable node represents passing of data.
-\item[model node] The model node represents the handling of data.
+\item[variable node] represents passing of data.
+\item[model node] represents the handling of data.
+\item[driver node] represents constrol structures capable of managing iteration
 \end{description}
-The possible edge types are:
+There are two edge types: 
 \begin{description}
@@
 \end{description}
-Model nodes must be used with the following rules:
-\begin{itemize}
+
+A number of restrictions are places on the model node with respect to the number and 
+type of edges that can be connected to it: 
+\begin{enumerate}
 \item A model node can only have one edge directed to or from another model node.
@@
 \item A model node must have at least one edge directed in and at least one edge directed out.
-\end{itemize}
-The MDO problem formulation concepts represented by nodes are given in Table \ref{t:node representation}, and the concepts represented by nodes are given in Table \ref{t:edge representation}.
+\end{enumerate}
+
+The MDO problem formulation concepts represented by nodes are given in 
+Table \ref{t:node representation}, and the concepts represented by edges are 
+given in Table \ref{t:edge representation}.
 % \begin{itemize}

### 2013-03-15 intro.tex
updated introduction and started editing a couple of lower sections

@@
     frameworks such as OpenMDAO\cite{Gray2012} and ModelCenter have enabled a new level of analysis tool integration 
-        and paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. 
-        Such complex models present a distinct challenge for proper implementation of any given solution strategy.
+    and paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. That 
+    new capability has created a new challenge since configuring a model with the larger nubmer of analyses, associated inputs and outputs
+    can be very difficult. Even for models with 10's of analyses, there could be hundreds of variables that need to be managed. 
+    It is not hard to imagine that the task of combining all the analyses into a consistent system model capable of solving a relevant 
+    engineering design problem could easily become much more costly than createing the discipline analyses themselves. Thus, for a large system
+    as the couplings between the disciplines begin to dominate the design space, the couplings between the analyses begin to dominate 
+    the job of setting up the model.
 
-        We propose the use of graph-based problem formulation that generally describes a
-        given problem, supports algorithmic manipulation to find beneficial solutions strategies, and enables automated 
-        implementation of those strategies inside an MDAO framework. In order to serve that purpose, the graph must include the following information: 
-        \begin{itemize}
-           \item Discipline Analyses 
-           \item Discipline State Variables and Residuals
-           \item Design Variables
-           \item Constraints
-           \item Objective or Objectives
-           \item Coupling Constraints
-           \item Parameters
-        \end{itemize}
-        Notably absent from the preceding list are any kind of solvers, optimizers, or other iterative solution finding tools. 
-        At its most basic, a problem formulation includes only information about what is being sought after in a given problem or what the
+    When building a model, coupling represents the reciprical flow of information between two analyses. 
+    We propose the use of graph-based syntax to describe the flow of information from inputs, 
+    through analyses, to outputs. If you allow for any free inputs to become design
+    variables and some set of the outputs to become design objectives and constraints, then the graph 
+    represents the problem formulation for a given design effort. 
+    The graph-based syntax presented has application in all phases of the design process. 
+    It can be used early on, to help construct a valid problem formulation from the set of available analysis tools. 
+    It also provides a consistent structure to describe problem formulation that provides a foundation to apply
+    algorithmic methods for selecting and then implementing effective solution strategies inside an MDAO framework.
+
+    We've already identifed that a problem formulation graph should include design variables, analysis tools and their 
+    associated inputs and outputs, and objectives and constraints. Notably absent from that list are any kind of 
+    solvers, optimizers, or other iterative solution finding tools. At its most fundamental state, a problem formulation 
+    includes only information about what is being sought after in a design or what the
     goals of a given problem are. It need not contain any information about solution paths or strategies to reach those goals. 
-
-        We define a problem specification striped down to its most essential parts to be the 
-        Fundamental Problem Formulation (FPF). By definition, the FPF for any given problem will be constant regardless 
-        of which MDAO framework, optimization architecture, optimization algorithm, or iterative solver is used to solve the problem.
-
-        The graph-based syntax used in this work is based on the Design Structure Matrix (DSM). 
-        This provides several key features that make it useful for working with large-scale MDAO problems. 
-        It provides a rigid structure that can be easily manipulated with a wide range of well-established graph-theory algorithms 
-        for the purposes of problem decomposition. It also provides a means for testing a given 
-        problem formulation to check if it is the FPF, and if not, to reduce it to the FPF. Lastly, it lends itself 
-        well to interacting with MDAO frameworks.
+    While solution information can be represented with the graph syntax presented here, a unique aspect of this syntax is that 
+    such information always remains seperable from the underlying problem formulation. In other words, a given a graph representing 
+    a specific solution strategy for a specific problem formulation, it is possible to remove the solution information from the 
+    graph and reclaim the more basic problem formulation.
 
 \section{Specific vs Fundamental Problem Formulation }
-        A simple, notional problem is given as follows:
+
+    The Fundamental Problem Formulation (FPF) for any given problem will be constant regardless 
+    of which MDAO framework, optimization algorithm, iterative solver, or solution strategy
+    is used to solve the problem. For example, consider the following notional problem:
 
@@
 
-        Where $A$ and $B$ represent analysis tools, and $f$ and $g$ are the objective and constraint functions respectively. 
+    $A$ and $B$ represent analysis tools, and $f$ and $g$ are the objective and constraint functions respectively. 
     Equation \ref{eqn:simple} makes an inherent assumption about the solution strategy for the problem. 
     Analysis $A$ outputs $z$, which is an input to $B$. Hence, $A$ should be run before $B$ with 
-        the $y$ being iterated on to convergence with $y^t$. However, a different solution could be equally valid and still represent
-        the same fundamental problem:
+    with $y$ being iterated on to convergence with $y^t$. However, a slightly different formulation is 
+    equally valid and still represents the exact same problem:
 
@@
     Equation \ref{eqn:simple2} differs only slightly from Eq. \ref{eqn:simple}. $A$ is now dependent on the output of $B$, 
-        and $z$ will be iterated on to satisfy $g$. Now the problem can be solved by running $B$ first and then $A$.
+    and $z$ will be iterated on to convergence with $z^t$. Now the problem can be solved by running $B$ first and then $A$.
     Since the formulations in Eqs. \ref{eqn:simple} and \ref{eqn:simple2} both describe the same problem and neither can be the
-        FPF, there must be a more fundamental description of the problem that is common between them. We present the FPF as follows:
+    FPF. They are both specific versions of the same more fundamental description of 
+    the problem that is common between them. We present the FPF as follows:
 
@@
 
-    \section{Using DSM for Problem Formulation}
+\section{Existing Graph--Based Problem Formulation Syntax}
 
@@
     number of optimization architectures\cite{Tedford2009}. Their work demonstrates clearly how multiple specific 
-        problem formulations can all relate back to a common FPF.
+    problem formulations can all relate back to a common FPF. The challenge with using this 
+    traditional mathematical syntax is that it is not easily manipulated or analyzed. 
+    A number of graph-based methods have been used successfully to translate the 
+    mathematical syntax into a more useful computational form.
 
-        The challenge with using this traditional mathematical syntax is that it is not easily manipulated or analyzed. 
-        A number of matrix-based methods have been used successfully to translate the mathematical syntax into a more useful computational form. 
     Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools where off 
@@
     Rogers et. al developed DeMAID to manipulate a DSM to find an ordering for analysis tools that 
-        reduces the cost of solving highly coupled systems\cite{Rogers1996}. This re-ordering yields multiple specific problem 
+    reduces the cost of solving highly coupled systems\cite{Rogers1996}. This re-ordering is done through 
+    row operations on the DSM matrix and yields multiple specific problem 
     formulations which all solve the same FPF. In other words, manipulation of the DSM does not fundamentally alter
-        the problem formulation, which makes DSM an excellent foundation for specifying the FPF itself. Traditional DSM only captures information 
-        about data dependency between analyses. Objective and constraint information is missing from the description of the problem.
+    the problem formulation, which makes DSM an excellent foundation for specifying the FPF itself.
 
+    Despite it's attractive properties, a DSM by itself is insufficient to describe complete problem formulations. 
+    Traditional DSM only captures information about data dependency between analyses. 
+    Objective and constraint information is missing from the description of the problem. 
     An alternative matrix-based syntax, called a Functional Dependency Table (FDT), was proposed by Wagner and Papalambros. 
@@
     more efficient optimization problem decompositions\cite{Michelena1997}. While FDT succeeds at capturing the 
-        information about objectives and constraints, it neglects the data dependency captured by DSM. For instance, 
-        although we know from the FPF in Eq. \ref{eqn:simple_fpf} that the objective, $f$, is dependent on the 
-        output of analysis $A$, you could not determine that from the FDT in Fig. \ref{fig:FDT_simple} alone. This make FDT 
-        applicable for partitioning decisions, but not for ordering problems.
+    information about objectives and constraints, it can not capture the coupled data dependency that DSM captures. For instance, 
+    we know from the FPF in Eq. \ref{eqn:simple_fpf} that the objective, $f$, is dependent on the 
+    output, $m$, of analysis $A$. You could not determine that from the FDT in Fig. \ref{fig:FDT_simple} alone. This missing 
+    information means that, while FDT is very useful for partitioning problems, it is also not sufficient to contain a 
+    complete problem formulation.
 
@@
 
-        Lamb and Martins included the variables, objectives, and constraint functions as nodes in an Extended 
-        DSM (XDSM) in order to capture a more complete description of solution strategies for MDAO problems\cite{Lambe2012}. By adding in those 
-        elements, they partially combined a traditional DSM with an FDT. 
+    Lamb and Martins included objectives and constraint functions as nodes in an Extended 
+    DSM (XDSM) in order to capture a more complete description of solution strategies for MDAO problems\cite{Lambe2012}. 
+    XDSM retains the square adjacency matrix form from DSM, but by adding in the 
+    new elements they partially combined a traditional DSM with an FDT. 
     This allows XDSM to represent data dependency between multiple analysis tools as well as between analysis tools and
-        objective/constraint functions. With the additional information included in an XDSM, Lu and Martins applied both ordering and partitioning 
+    objective/constraint functions. With the additional information included in an 
+    XDSM, Lu and Martins applied both ordering and partitioning 
     algorithms on an MDAO test problem named the Scalable Problem \cite{Lu2012}.
@@
     As a result, all variable information is aggregated 
-        so that for the problem from Eq. \ref{eqn:simple_fpf} you can say that $A$ depends on $B$ or vice versa, but you can't identify which 
-        individual variables are interacting in the dependency cycle. Without the detailed variable information, you can't construct the compatibility 
-        constraints necessary to implement the problem. Additionally, XDSM requires the use of solver and optimizer blocks to represent 
-        the relationship between design variables and objectives/constraints. By introducing solver or optimizer blocks XDSM automatically provides
-        some kind of solution strategy. The XDSM for Eq. \ref{eqn:simple} is given in Figure \ref{fig:XDSM_simple}. 
-        This diagram is shown with an assumed gauss-siedel iteration scheme and a MDF solution architecture. 
+    so that for the problem from Eq. \ref{eqn:simple_fpf} you can say that $A$ 
+    depends on $B$ or vice versa, but you can't identify which 
+    individual variables are interacting in the dependency cycle. 
+    Without the detailed variable information, you can't construct the compatibility 
+    constraints necessary to implement the problem. Additionally, XDSM 
+    requires the use of solver and optimizer blocks to represent 
+    the relationship between design variables and objectives/constraints. 
+    By introducing solver or optimizer blocks XDSM automatically provides
+    some kind of solution strategy. The XDSM for Eq. \ref{eqn:simple} is 
+    given in Figure \ref{fig:XDSM_simple}. This diagram is shown with an 
+    assumed gauss-siedel iteration scheme and a MDF solution architecture. 
     Hence XDSM is too specific for use with a fundamental problem formulation.
 
-        By taking XDSM, removing the requirement for solver/optimizer nodes, and adding nodes for each variable, we are able to fully describe 
-        the problem formulation for a general MDAO problem. In the following sections we describe the necessary syntax for specifying a 
-        problem formulation within a DSM and show how to use the resulting graph to find the FPF from any given graph. For completeness we 
-        also present algorithms for converting between the general DSM and the more specific formulations discussed above.
+    Of the three methods, XDSM comes the closest to fully describing a problem formulation. 
+    Thought it does not contain specific variable information, this could be easily added back in
+    simply by putting the variables back into the graph and drawing the apprpriat connections
+    between them and the relevant analysis nodes. However, it's dependence on 
+    solution strategy specifics to represent couplings can not be removed without 
+    stripping it of the useful information contained in an FDT. In the following sections
+    we describe a new graph syntax that retains the features of XDSM, without the dependence on 
+    solution strategy (although it does allow for that information to be present). Thus 
+    the new syntax allows for the compelete specification of an FPF.

### 2013-03-15 graph_syntax.tex
merged

@@
 
+Let $I$ be a nonempty set such that for each $i \in I$ there is a corresponding set $A_i$.
+The family $\{A_i | i \in I\}$ is an indexed family of sets with index $i$ and indexing set $I$, and we can say $\mathcal A = \{A_i | i \in I\}$\cite{smith2006}.
+Then the union over this family of sets is
+\begin{equation}
+\bigcup_{i \in I} A_i = \bigcup_{A \in \mathcal A} A = \{x | x \in A \txt{ for } A \in \mathcal A\}.
+\end{equation}
+Lastly, the cardinality of a set $B$ is the number of elements in $B$ and is denoted as $|B|$.
+
 % \subsection{Problem Formulation Concepts to Represent}
@@
 
-A number of restrictions are places on the model node with respect to the number and 
+A number of restrictions are placed on the model node with respect to the number and 
 type of edges that can be connected to it: 
@@
 To construct the maximal connectivity graph, we assume that a set of codes, global inputs, and objectives and constraints (collectively called global outputs).
-The codes are represented by analysis blocks $A_i=(V_{A_i},E_{A_i}), \ i=1,\ldots,m$, the global inputs are represented a set of variable nodes $I$, and the global outputs are represented by a set of variable nodes $O$.
-We assume that $O$, $I$, and $A_i$ are given, and that any potential connection between variables is given in the form of the free edges in the set $C_M$.
+The codes are represented by analysis blocks $A_i=(V_{A_i},E_{A_i}), \ i\in \{1,2,\ldots,m\}=I$, the global inputs are represented a set of variable nodes $V_\txt{in}$, and the global outputs are represented by a set of variable nodes $V_\txt{out}$.
+We assume that $V_\txt{in}$, $V_\txt{out}$, and $A_i$ are given, and that any potential connection between variables is given in the form of the free edges in the set $C_M$.
 Then we may construct the maximal connectivity graph $M=(V_M,E_M)$ as
 \begin{IEEEeqnarray*}{rCl}
-V_M & = & I \cup O \cup \left( \bigcup_{i = 1}^m V_{A_i} \right), \\
-E_M & = & C_M \cup \left( \bigcup_{i=1}^m E_{A_i} \right),
+V_M & = & V_\txt{in} \cup V_\txt{out} \cup \left( \bigcup_{i \in I} V_{A_i} \right), \\
+E_M & = & C_M \cup \left( \bigcup_{i \in I} E_{A_i} \right),
 \end{IEEEeqnarray*}
@@
 \begin{enumerate}
-\item[(1)] $\displaystyle{V_F = I_F \cup O \cup \left( \bigcup_{i \in \mathcal A} V_{A_i} \right),\ I_F \subset I}$
-\item[(2)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in \mathcal A} E_{A_i} \right)}$
-\item[(3)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'}\quad \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
+\item[(1)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} | \txt{deg}^+(v) > 0\} }$
+\item[(2)] $\displaystyle{I_F = \{i \in I | \exists v \in V_{A_i} \txt{ with } \txt{deg}^+(v)>0 \txt{ and } t_\txt{node}(u)=\txt{`model'} \txt{ for } (u,v) \in E_{A_i}\} }$
+\item[(2)] $\displaystyle{V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right)}$
+\item[(3)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right)}$
+\item[(4)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
 \end{enumerate}
-The set $\mathcal A$ is an index set containing the indices of the analysis blocks in $F$; for the case where all of the analysis blocks are used, we would have $\mathcal A = \{1,2,\ldots,m\}$.
-The first requirement for $F$ is that $V_F$ be composed of all of the global outputs (the required objectives and constraints), the nodes from each analysis block in $F$, and any global input nodes that are needed, $I_F$.
+The set $I_F$ is an index set containing the indices of the analysis blocks in $F$; for the case where all of the analysis blocks are used, we would have $I_F = I = \{1,2,\ldots,m\}$. 
+The set $V_{\txt{in},F}$ is the set of global inputs used by the analysis blocks in $F$.
+Because it is possible for some global inputs to be unused, we only require $V_{\txt{in},F} \subset V_\txt{in}$.
+
+The first requirement creates the set $V_{\txt{in},F}$ which is the set of only the global inputs that are used.
+The second requirement creates set $I_F$ which is an index set corresponding to only that analysis blocks
+The first requirement for $F$ is that $V_F$ be composed of all of the global outputs (the required objectives and constraints), the nodes from each analysis block in $F$, and any global input nodes that are needed.
 The second requirement suggests that the edges in $F$ comprise the edges for each analysis block and the edges between them, the global inputs, and the global outputs. 
@@
 \begin{equation}
-H = \{v \in V | \txt{deg}^-(v) < \txt{deg}_l^-(v) \}
+H = \{v \in V | t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
 \end{equation}
-Then the updated set of edges is
+which is the set of variable nodes with few incoming edges than are allowed by the lower indegree limit.
+Then the updated set of edges is created by removing the first set of free edges following the hole:
 \begin{equation}
-C_{F,1} = C_{F,0} \setminus \{\ e \in E, \ e=xy | x=E^3(v) \txt{ for } v \in H\}
+C_{F,1} = C_{F,0} \setminus \{(x,y) \in E | x=E^3(v) \txt{ for } v \in H\},
 \end{equation}
 %deg-v is now recaculated for the new C. 
-This step is demonstrated by Fig.~\ref{f:holes}.
-\item (())
+and this step is demonstrated by Fig.~\ref{f:holes}.
+\item The second step is to detect collisions and to then disconnect enough free edges so that the collisions are resolved. The set of variable nodes which containing collisions is created as
+\begin{equation}
+S_\txt{nodes} = \{v \in V | t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
+\end{equation}
+For each collision node we can construct a set contained the edges directed in. The set containing all of these sets is constructed as
+\begin{equation}
+S_\txt{edges} = \big \{ \{(x,y) \in E\} | y \in S_\txt{nodes} \big \}
+\end{equation}
+
 \end{enumerate}

### 2013-03-15 graph_syntax.tex
update

@@
 \begin{enumerate}
-\item[(1)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} | \txt{deg}^+(v) > 0\} }$
-\item[(2)] $\displaystyle{I_F = \{i \in I | \exists v \in V_{A_i} \txt{ with } \txt{deg}^+(v)>0 \txt{ and } t_\txt{node}(u)=\txt{`model'} \txt{ for } (u,v) \in E_{A_i}\} }$
-\item[(2)] $\displaystyle{V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right)}$
-\item[(3)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right)}$
-\item[(4)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
+\item[(1)] $\displaystyle{I_F \subset I}$ such that the following conditions hold
+\item[(2)] $\displaystyle{ \forall i \in I, \ \exists v \in V_{A_i} \txt{ with } t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v)>0 }$
+%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_{A_i} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
+%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_\txt{out} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
+\item[(3)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} | \txt{deg}^+(v) > 0\} }$
+\item[(4)] $\displaystyle{V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right)}$
+\item[(5)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right), \ C_F \subset C_M}$
+\item[(6)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
 \end{enumerate}
-The set $I_F$ is an index set containing the indices of the analysis blocks in $F$; for the case where all of the analysis blocks are used, we would have $I_F = I = \{1,2,\ldots,m\}$. 
-The set $V_{\txt{in},F}$ is the set of global inputs used by the analysis blocks in $F$.
-Because it is possible for some global inputs to be unused, we only require $V_{\txt{in},F} \subset V_\txt{in}$.
+The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6).
+Requirement (2) stipulates that the analysis blocks in $F$ must each have at least one local output that is beging used.
+Requirement (3) stipulates that only the global inputs that are being used should be included.
+Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper indegree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.
 
-The first requirement creates the set $V_{\txt{in},F}$ which is the set of only the global inputs that are used.
-The second requirement creates set $I_F$ which is an index set corresponding to only that analysis blocks
-The first requirement for $F$ is that $V_F$ be composed of all of the global outputs (the required objectives and constraints), the nodes from each analysis block in $F$, and any global input nodes that are needed.
-The second requirement suggests that the edges in $F$ comprise the edges for each analysis block and the edges between them, the global inputs, and the global outputs. 
-The final requirement is specific to the number of edges directed into a variable. 
-If there are no edges directed inward, the node is called a \emph{hole}, and if more connections are directed in than are allowed, the node is called a \emph{collision}; a hole is allowed if $\txt{deg}_l^-(v)=0$.
-The final requirement is therefore a requirement on $C_F$.
+The set $C_F$ is a set containing free edges representing connections.
+Since $C_M$ is the set of all potential connections, we must have $C_F \subset C_M$.
+While no other conditions are explicitly stated for $C_F$, requirement (6) is actually a requirement on both $I$ and $C_F$.
 
@@
 %deg-v is now recaculated for the new C. 
-and this step is demonstrated by Fig.~\ref{f:holes}.
-\item The second step is to detect collisions and to then disconnect enough free edges so that the collisions are resolved. The set of variable nodes which containing collisions is created as
-\begin{equation}
-S_\txt{nodes} = \{v \in V | t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
-\end{equation}
-For each collision node we can construct a set contained the edges directed in. The set containing all of these sets is constructed as
-\begin{equation}
-S_\txt{edges} = \big \{ \{(x,y) \in E\} | y \in S_\txt{nodes} \big \}
-\end{equation}
-
-\end{enumerate}
+and this step is demonstrated by Fig.~\ref{f:holes}. Because removing these edges can create new holes, this step must be repeated until no more holes are found.
 \begin{figure}[htb!]
@@
 \end{figure}
+\item The second step is to detect collisions and to then disconnect enough free edges so that the collisions are resolved. The set of variable nodes which contain collisions is created as
+\begin{equation}
+S_\txt{nodes} = \{v \in V | t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
+\end{equation}
+For each collision node we can construct a set contained the edges directed in. The set containing all of these sets is constructed as
+\begin{equation}
+S_\txt{edges} = \big \{ \{(x,y) \in E\} \big | y \in S_\txt{nodes} \big \}
+\end{equation}
+Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$.
+We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$.
+Then we may construct sets of edges as
+\begin{equation}
+B_j = \big \{e_k \in S_{\txt{edges},j} \ | \ k \in \{1,2,\ldots,K\} \txt{ with } K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
+\end{equation}
+which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper indegree limit of $v_j$.
+The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include.
+Then let
+\begin{equation}
+C_{F,2} = \{ e \in C_{F,1} \ | \ e \in B_j \txt{ for some } j \in J\}.
+\end{equation}
+\item The third and final step is to remove any unused nodes and edges. The set of unused variable nodes representing local outputs is
+\begin{equation}
+V_\txt{empty} = \{v \in V_M \ | \ t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) = 0\},
+\end{equation}
+and so then the indexing set for only the analysis blocks with at least one used local output may be constructed as
+\begin{equation}
+I_3 =  \{i \in I \ | \ \nexists v \in V_{A_i} \txt{ such that }  v \in V_\txt{empty} \}.
+\end{equation}
+Finally, the set of nodes and edges describing a fundamental problem formulation is given by
+\begin{equation}
+V_{F,3} = V_M \setminus V_\txt{empty}
+\end{equation}
+\begin{equation}
+C_{F,3} = \{ (x,y) \in C_{F,2} \ | \ x \in V_{A_i} \txt{ or } y \in V_{A_i} \txt{ for some } i \in I_3 \}
+\end{equation}
+\begin{equation}
+F_3 = (V_{F,3},C_{F,3})
+\end{equation}
 
-%\begin{enumerate}
-%\item The first step is to detect holes and remove the corresponding analysis blocks, which could not be used because the required inputs are not supplied.
-Any analysis blocks with holes may be stored in an index set $\mathcal H$ as
-%\begin{equation}
-%\forall i \in \mathcal A_1,\txt{ and }\forall v \in V_{A_i},\txt{ if deg}^-(v)=0 \txt{ then } i \in \mathcal H.
-%\end{equation}
-%Then an updated index set $\mathcal A_2$ is created using set difference notation as
-%\begin{equation}
-%\mathcal A_2 = \mathcal A_1 \setminus \mathcal H
-%\end{equation}The connection edges directed from the removed analysis blocks are removed as
-%\begin{equation}
-%H = \{c \in C_{F,1} |c = (e^{-1}(v),v), \txt{ where } v \in V_{A_i} \txt{ for some } i \in \mathcal H\},
-%\end{equation}
-%\begin{equation}
-%C_{F,2} = C_{F,1} \setminus H
-%\end{equation}
-%Finally, the FPF is updated as
-%\begin{equation}
-%V_{F,2} = I_{F,2} \cup O \cup \left( \bigcup_{i \in \mathcal A_2} V_{A_i} \right),
-%\end{equation}
-%\begin{equation}
-%E_{F,2} = C_{F,2} \cup \left( \bigcup_{i \in \mathcal A_2} E_{A_i} \right)
-%\end{equation}
-%This process may require multiple iterations because removing an analysis block may create holes upstream.
-It is assumed that the process as been repeated sufficiently such that $\mathcal A_2$ does not have any holes.
-%
-%\item The next step is to resolve conflicts. The set $C$ of input nodes containing conflicts is
-%\begin{equation}
-%C = \{v \in V_{F,2} | t_\txt{node}(v) = \txt{`input'} \txt{ and } \txt{deg}^-(v) > d(v)\}
-%\end{equation}
-%%\begin{equation}
-%%for v \in c let 
-%%\end{equation}
-%\end{enumerate}
+%Let $F_2 = (V_M,C_{F,2})$ be an undirected graph, which is to say to disregard the directions associated with $C_{F,2}$.
 
-%\subsection{Classification of the FPF}
+%Fixed edges are fixed only with respect to the analysis block in which they reside, it is okay to remove them along with the entire analysis block.
 
-%A collision represents a choice.
-%Delete edges and analysis blocks so that there are no nodes or collisions.
-
-% \begin{figure}[htb!]
-	% \begin{center}
-	% \includegraphics[width=2in]{images/hour_glass}
-	% \end{center}
-	% \vspace{-20pt}
-% \caption{(()).}
-% \label{f:our glass}
-% \end{figure}
-% \subsection{old}
-	% For a large scale problem with many analyses, obtaining the FPF is likely to be a challenge. In general, a set of analysis codes will be given and the desired outputs will be specified as
-	    % \begin{align}
-            % given & \ \ A: \lbrace x_i \lvert i \in A_\textrm{input} \rbrace \rightarrow \lbrace x_i \lvert i \in A_\textrm{output} \rbrace \notag
-            % \\    & \ \ B: \lbrace x_i \lvert i \in B_\textrm{input} \rbrace \rightarrow \lbrace x_i \lvert i \in B_\textrm{output} \rbrace \notag
-			% \\    & \ \ \quad \quad \quad \quad \quad \quad  \quad \quad \vdots
-            % \\min. &\ \ f(x_i), i \in \mathcal{O} \notag
-            % \\w.r.t. & \ \ x_i, i \in \mathcal{I}, \notag
-            % \label{eqn:preFPF}
-        % \end{align}
-	% where $\mathcal{O}$ is the index set of the global outputs and $\mathcal{I}$ is the index set of the global inputs.
-For this to be an FPF, there must be no \emph{conflicts} or \emph{holes}.
-A conflict arises when different analyses produce the same output; two analysis codes $A$ and $B$ are in conflict if
-		% \begin{equation}
-			% A_\textrm{output} \cap B_\textrm{output} \neq \emptyset.
-		% \end{equation}
-	% A hole arises when a local input to an analysis block is neither a global input nor a local output of any analysis block.
-If we let $\mathcal{H}$ denote the index set of local inputs which are holes
-		% \begin{equation}
-			% i \in \bigcup \{A_\textrm{input},B_\textrm{input},\ldots\} \textrm{ and } i \notin \bigcup \{\mathcal{I},A_\textrm{output},B_\textrm{output},\ldots\}  \implies  i \in \mathcal{H}.
-		% \end{equation}
-	% If the given set of analyses contains any holes, an FPF cannot be obtained.
-On the contrary, if there are conflicts multiple FPFs may be obtained.
-This is because every conflict represents a choice of which analysis to use and each choice could (potentially) yeild a different but valid FPF.
-The following section develops an application of graphy theory to obtain multiple FPFs from a set of analyses and desired outputs.
-
-    % \subsection{Formulation Graph Syntax}
-    % Rather than start with an adjacency, we chose to work directly with a directed cyclic graph to develop a syntax for the FPF. 
-    % The following information must be provided to start:
-    % \begin{itemize}
-        % \item Analysis blocks: an analysis block represents any calculation, and each comes with
-            % \begin{itemize}
-                % \item local inputs
-                % \item local outputs 
-                % \item execution properties: these are the properties associated with running the code, such as run time
-            % \end{itemize}
-        % \item Global parameters: these may serve as fixed inputs to the local inputs of analysis blocks
-        % \item Global outputs: these may represent
-            % \begin{itemize}
-                % \item objectives
-                % \item constraints
-                % \item residuals: (not sure)
-            % \end{itemize}
-    % \end{itemize}
-
-    % The graph representation of a data flow is cast to utilize the extensive library of algorithms in graph theory to analyze a directed weighted graph. 
-    % Edge weights are used to represent the metrics associated with a data flow:
-    % \begin{itemize}
-        % \item Run time: this metric is a property of an analysis block
-        % \item Fidelity: this metric is a property of an individual local output
-        % \item Expected Convergence
-    % \end{itemize}
-
-    % The key assumption is that identical variables are recognized as such. This serves as the basis for creating a data flow by connecting compatible input and output nodes with a directed edge. 
-    % To represent the fact that execution of an analysis code does not depend on the number of outputs being used, we have created the following figure (not made yet).
-
-    % This information immediately leads to the maximal connectivity graph, which is formed by placing a directed edge from each local output or global parameter to each matching local input or global output.
-    % Whenever multiple edges are connected to a single input, a conflict occurs because only one may be used. Resolving these conflicts is one key challenge in creating a data flow.
-%    \begin{itemize}
-%        \item Specify analyses
-%        \item Connections between analyses 
-%            \begin{itemize}
-%                \item local variables
-%                \item global variables? Use "fake" node that broadcasts out to the rest of the graph? 
-%            \end{itemize}
-%        \item Cycles indicate coupling
-%        \item Cycles for design variables->objectives/constraints
-%        \item Objectives/constraints are just outputs? Special nodes? 
-%        \item Residuals are just outputs? Special nodes? 
-%        \item Parameters are just input nodes that are not design variables (use identifies these)
-%        \item FPF no solvers/optimizers anywhere in it
-%    \end{itemize}
+\end{enumerate}
 

### 2013-03-15 graph_syntax.tex
more updates

--- 2013-03-15 graph_syntax.tex
@@
 \begin{description}
-\item[fixed edge] A fixed edge may not be removed.
-\item[free edge] A free edge may be removed.
+\item[fixed edge] A fixed edge may never be removed from the graph.
+\item[free edge] A free edge may be removed removed from the graph.
 \end{description}
@@
 \end{enumerate}
+
+Analysis tools take in a number input variables, then perform some work calculate 
+the values for their respective outputs. In an MDAO graph, this process is 
+represented by an group of nodes and edges called an \emph{analysis block}, 
+shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
+node represents a single input or output and is connected 
+to a single model node via a fixed edge. Note that in Fig. \ref{f:analysis block}, 
+the analysis block contains two model nodes, with a single edge connecting them. 
+This part of the graph represents the necessary calculations to map given inputs 
+to proper output values.
+
+\begin{figure}[htb!]
+    \begin{center}
+    \includegraphics[width=4in]{images/analysis_block}
+    \end{center}
+    \vspace{-10pt}
+\caption{Example analysis block. The each node type and edge type is labeled in italics and annotated parenthetically.}
+\label{f:analysis block}
+\end{figure}
+
+Finally, free edges are used to connect the variable nodes of the analysis block 
+to other variables nodes outside of the analysis block. In this way, the 
+analysis block graph is a fundamental building block of the MDAO problem data flow.
 
@@
 \end{table}
-
-Finally, an analysis code is represented by an \emph{analysis block}, which is a collection of model nodes, variable nodes, and fixed edges with a specific structure.
-This structure is derived from the need to represent each input and output individually while representing the actual analysis with a single node or edge.
-The variable nodes each represent a single input or output of the analysis block.
-Each of these nodes is connected to a single model node via a fixed edge, which represents the gathering of the inputs to begin the analysis or the disseminating of outputs after the analysis.
-Finally, free edges are used to connect the variable nodes of the analysis block to other variables nodes outside of the analysis block.
-In this way, the analysis block graph is a fundamental building block of the MDAO problem data flow.
-\begin{figure}[htb!]
-	\begin{center}
-	\includegraphics[width=4in]{images/analysis_block}
-	\end{center}
-	\vspace{-10pt}
-\caption{Example analysis block. The each node type and edge type is labeled in italics and annotated parenthetically.}
-\label{f:analysis block}
-\end{figure}
 %`Connection' edges connect the local outputs from one analysis block to input nodes.

### 2013-03-15 graph_syntax.tex
reorganized the graph section a bit.

@@
     \item Local and global variables 
+    \item Information passing between analyses
     \item Coupling between analyses
@@
 
-The following section provides an overview of graph theory basics, then describes 
-in detail the different node and edge types, as well as rules governing allowable 
-egdegs between nodes, used as the foundation of the new graph syntax. These basic graph 
-primatives are them mapped to the MDAO constructs stated above, and used to 
-build three distinct graph forms:
+Beyond those basic constructs, there are also multiple three phases for a design process that 
+all need to be representable with the new syntax. Firstly there is the initial problem definition
+phase where the specific analysis tools and design goals are identified. At the end of this phase, 
+a single formal problem formulation is selected specifying design variables, constraints, objectives, 
+analysis tools, etc. Lastly some specifc procedure for solving the problem is selected, for example 
+picking an MDAO optimization architecture. Using the proposed graph syntax, each of these phases 
+can be represented with a related graph.
 
@@
     \item Fundamental Problem Formulation 
-    \item Specific Solution
+    \item Specific Solution Formulation
 \end{itemize}
 
-Each of these forms serves a different purpose in the design process, and we 
-present some transformations that can be applied to transition 
-between any two of the three graph forms.
+We present some transformations that can be applied to transition any graph
+between any two adjacent design phases. The \emph{maximal connectivity graph} 
+represents the full potential of all the analysis codes being considered, i.e. 
+a large set of analysis codes being considered for a given problem are included in
+a graph, with all possible connections between them also present. The second graph 
+is \emph{fundamental problem formulation graph}, which is the smallest possible graph 
+that still fully defines a given problem formulation. Finally, a \emph{specific problem formulation} 
+may be represented by including additional edges and node types to represent the 
+solution strategy being employed to solve the problem.
+
+The relationship between these three graphs is depicted in Figs.~\ref{f:tree} and \ref{f:hourglass}. 
+The tree diagram demonstrates the fact that it is generally possible to obtain 
+multiple FPFs from a single maximal connectivity graph. This  may correspond to 
+different down--selections of analysis codes, different connections between them, 
+or both. Each down--selection shrinks the number of possible FPFs that could be reached 
+untill only one is remaining. Then, a single FPF, different SPFs may be obtained by implementing 
+different solution strategies. The hourglass shape in Fig. \ref{f:hourglass} illustrates how 
+the MCG gets reduced to a single FPF, then multiple possible SPFs exist to solve the problem.
+In other words the FPF is obtained from the MCG by removing nodes and edges, 
+and the SPF is obtained from the FPF by adding nodes and edges.
+
+\begin{figure}[htb!]
+    \centering
+    \subfigure[number of possible graphs]{
+    \includegraphics[width=2.0in]{images/tree}
+    \label{f:tree}
+    }
+    \subfigure[graph size]{
+    \includegraphics[width=2.0in]{images/hourglass}
+    \label{f:hourglass}
+    }
+\caption{The relationship between the MCG, FPF, and SPF.}
+\end{figure}
 
@@
 \item A model node must have at least one edge directed in and at least one edge directed out.
+\item If a variable node with an outgoing edge to a model node may not have any other outgoing edges.
 \end{enumerate}
 
-Analysis tools take in a number input variables, then perform some work calculate 
+Analysis tools take in a number input variables, then perform some work to calculate 
 the values for their respective outputs. In an MDAO graph, this process is 
@@
 This part of the graph represents the necessary calculations to map given inputs 
-to proper output values.
+to proper output values. Although analysis blocks are comprised of multiple nodes and edges, since all 
+the edges within them are fixed, they become fixed structures within a graph.
 
@@
 
-Finally, free edges are used to connect the variable nodes of the analysis block 
-to other variables nodes outside of the analysis block. In this way, the 
-analysis block graph is a fundamental building block of the MDAO problem data flow.
+Via the analysis block structure, we can also distinguish between input and ouptut
+variable nodes. Inputs are any variable nodes that have an outgoing edge into a model 
+node. Conversely, outputs are any variable nodes that have an incoming edge from a model node. 
+Note that by this definition, inputs and outputs can only exist when variable nodes poses an 
+edge joining them to a model node. In Fig. \ref{f:analysis block} these are labeled out as 
+local inputs and local outputs. The necessity to be connected to a specific model node forces
+the variable nodes to be local in nature. When information from the output of one analysis 
+block is passed, or connected, to the input of another analysis block, a new free edge is added 
+connecting the two variable nodes involved in the exchange. The edge is free because, unlike the edges 
+within an analysis block, it could be added or removed depending on the demands of the specific problem.
 
@@
 \section{Graph--based Representation of the Fundamental Problem Formulation}
-Next we provide the meaningful graphs that can be created from the suggested syntax.
-The first graph is the \emph{maximal connectivity graph} which represents the full potential of all the analysis codes being considered, i.e.
-each analysis code is represented by an analysis block, and each potential connections between variables is represented by a free edge.
-The second graph that may be represented is \emph{fundamental problem formulation graph}, which is the graph with the fewest number of edges and nodes needed to provide a valid problem formulation, as disscussed subsequently.
-Finally, a \emph{solution problem formulation} may be represented by including additional edges and node times to represent the optimization architecture, though this is beyond the scope of this paper.
-
-The relationship between these three graphs is depicted in Figs.~\ref{f:tree} and \ref{f:hourglass}.
-The tree diagram demonstrates the fact that it is generally possible to obtain multiple FPFs from a single maximal connectivity graph.
-This  may correspond to different down--selections of analysis codes, different connections between them, or both.
-Then, for each FPF, different SPFs may be obtained by implementing different optimization architectures.
-The hourglass shape indicates that the FPF graph is the smallest graph that may represent a specific problem formulation.
-The FPF is obtained from the MCG by removing nodes and edges, and the SPF is obtained from the FPF by adding nodes and edges.
-\begin{figure}[htb!]
-	\centering
-	\subfigure[number of possible graphs]{
-	\includegraphics[width=2.0in]{images/tree}
-	\label{f:tree}
-	}
-	\subfigure[graph size]{
-	\includegraphics[width=2.0in]{images/hourglass}
-	\label{f:hourglass}
-	}
-\caption{The relationship between the MCG, FPF, and SPF.}
-\end{figure}
 

### 2013-03-16 graph_syntax.tex
a few more re-org changes.

@@
 strategy is applied to the problem formulation. In order to achieve that goal 
-the graph needs to accomodate a number of constructs from MDAO problem:
+the graph needs to accomodate a number of constructs from MDAO problems:
 
 \begin{itemize}
-    \item Analysis tools 
-    \item Local and global variables 
-    \item Information passing between analyses
+    \item Analysis tools and connections between them
     \item Coupling between analyses
-    \item Design variables
-    \item Objective or objectives 
-    \item Local and global Constraints
-    \item Iterative loops from solvers and optimizers
+    \item Design Variable
+    \item Objectives and constraints
+    \item Local and global properties
 \end{itemize}
 
-Beyond those basic constructs, there are also multiple three phases for a design process that 
+Beyond those basic constructs, there are also three phases of a design process that 
 all need to be representable with the new syntax. Firstly there is the initial problem definition
@@
 
-We present some transformations that can be applied to transition any graph
-between any two adjacent design phases. The \emph{maximal connectivity graph} 
-represents the full potential of all the analysis codes being considered, i.e. 
-a large set of analysis codes being considered for a given problem are included in
-a graph, with all possible connections between them also present. The second graph 
+The \emph{maximal connectivity graph} represents the first phase with all the 
+analysis codes being considered with all possible connections between them also present. The second graph 
 is \emph{fundamental problem formulation graph}, which is the smallest possible graph 
@@
 \end{figure}
+
+In the following sections we present some graph theory basics, then demonstrate how to build 
+the relevant MDAO constructs from the given grah primatives. Next we show how to construct 
+the MCG, FPF, and SPF graphs from those constructs and demonstrate methods to transition between 
+them.
 
@@
 \end{description}
-There are two edge types: 
+
+Only the variable node and the model node are allowed to be used in an MCG or FPF. 
+All three nodes are required for any SPF. IN addition to the three node types, there 
+are two edge types:
+
 \begin{description}
@@
 \end{enumerate}
+
+\subsubsection{Analysis Blocks and Connections}
 
@@
 within an analysis block, it could be added or removed depending on the demands of the specific problem.
+
+\subsubsection{Design Variables, Objectives, and Constraints}
+
+\subsubsection{Local vs Global}
+
+\subsubsection{}
 

### 2013-03-16 graph_syntax.tex
a little more re-org, again...

@@
-\section{New Graph--based Syntax}
+\section{Requirements for New Graph Syntax}
 The ultimate goal of the new graph--based syntax presented here is to be able to 
@@
     \item Analysis tools and connections between them
+    \item Design variables, objectives, and constraints
+    \item Local and global properties
     \item Coupling between analyses
-    \item Design Variable
-    \item Objectives and constraints
-    \item Local and global properties
+    \item Multi-fidelity analyses
 \end{itemize}
@@
 
-\subsection{MDAO Graph Primatives}
+\section{MDAO Graph Primatives}
 There are three node types:  
@@
 
-\subsubsection{Analysis Blocks and Connections}
+\subsection{Analysis Blocks and Connections}
 
@@
 
-\subsubsection{Design Variables, Objectives, and Constraints}
-
-\subsubsection{Local vs Global}
-
-\subsubsection{}
+\subsection{Design Variables, Objectives, and Constraints}
 
@@
 
-\section{Graph--based Representation of the Fundamental Problem Formulation}
+\subsection{Local vs Global Properties}
 
-In this section we introduce the remaining necessary notion and give graph--theory based definitions of the MCG and FPF, and then process to demonstrate how the FPF is obtained from the MCG.
+\subsection{Coupling Between Analyses}
 
-\subsection{Additional Notation}
+\subsection{Multi-fidelity Analyses}
+
 For node $v \in V$ the edges directed out are given by $E(v)$ and the edges directed into $v$ are given by $E^{-1}(v)$; $E(E(v))$ is denoted as $E^2(v)$, and likewise for additional levels. 
@@
 %%% this needs to be changed to fit the definition using a directed graph
+
+\section{Building MCG, FPF, and SPF Graphs}
 

### 2013-03-18 graph_syntax.tex
updated the Desgin Variable, Objectives, Constraints section with some new notes. Two new node types

@@
 \section{MDAO Graph Primatives}
-There are three node types:  
+There are four node types:  
 \begin{description}
-\item[variable node] represents passing of data.
-\item[model node] represents the handling of data.
-\item[driver node] represents constrol structures capable of managing iteration
+\item[variable] represents scalar or array data
+\item[model] responsible for mapping inputs to outputs
+\item[expression] simple algebreic combinations of variables (such as for constraint of objective functions)
+\item[driver] constrol structures capable of managing iteration
 \end{description}
 
-Only the variable node and the model node are allowed to be used in an MCG or FPF. 
-All three nodes are required for any SPF. IN addition to the three node types, there 
-are two edge types:
+Only the variable, model nodes are allowed to be used in an MCG. In a FPF expression 
+nodes are also required. All four nodes are required for any SPF. In addition to 
+the three node types, there are two edge types:
 
@@
 edge joining them to a model node. In Fig. \ref{f:analysis block} these are labeled out as 
-local inputs and local outputs. The necessity to be connected to a specific model node forces
-the variable nodes to be local in nature. When information from the output of one analysis 
+local inputs and local outputs. The variables are local because they are explicitely tied
+to the analysis block and have no relevance without it.
+
+MDAO problems require that information be passed between sets of analyses. When information from the output of one analysis 
 block is passed, or connected, to the input of another analysis block, a new free edge is added 
@@
 
+Design variables, objectives, and constraints all share the property, that they are 
+require external decisions to be made about a given system to identify them. Consider 
+a notional problem given by the graph in Fig. [insert ref]a. There are three input variable 
+nodes which, could potentially become design variables: w, x, and y. Assume that the 
+output y is connected to the input y, yeilding Fig. [insert ref]b. Now only w and x could 
+be design variables. We define a design variable to be any input variable node that does 
+not have an incoming connection.
+
+[[INSERT GRAPH HERE]]
+
+Objective and constraints are also constructs that need to be identified by engineers 
+as part of the design process. In the case of objective functions single output values 
+could be selected, but commonly multiple values are assembeled together via simple 
+athrimetic opperations to form the final objective function. For example, the problem from 
+Fig. [insert ref] might have an objective function of $z^2$ or $2y + z$. Constraints are 
+always given in the form of either an inequality of an equality, e.g. 
+$\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer explictly identifies variables 
+or groups of variables and assigns an simple, explicit relationship between them.
+
+For the proposed syntax objectives and constraints are represented through the use of 
+expression nodes. These nodes contain the actual mathematical expression used to define the 
+objective or constraint value. You can distinguish between an objective or constraint node 
+via the presence of a simple expression (objective) vs an inequality/equality (constraint). 
+Fig. [insert new ref] shows the notional problem graph augmented with objective and 
+constraint values.
+
+[[INSERT ANOTHER GRAPH HERE]]
+
+\begin{comment}
 The MDO problem formulation concepts represented by nodes are given in 
@@
 %`Connection' edges connect the local outputs from one analysis block to input nodes.
+\end{comment}
+\subsection{Local vs Global Properties}
 
-\subsection{Local vs Global Properties}
+\begin{itemize}
+  \item what can be local or global? Variables, objectives, constraints, 
+  \item What does it really mean to be local vs global? 
+  \item What does the graph structure of local vs global look like? 
+  \item local is a frame of reference... a variable could be global when looking at two analyses, but local to the two when looking at the three
+\end{itemize}
 

### 2013-03-18 graph_syntax.tex
tweak
update and merger

@@
+\newcommand{\st}{ \ | \ }
+
 \section{Requirements for New Graph Syntax}
@@
 Let $I$ be a nonempty set such that for each $i \in I$ there is a corresponding set $A_i$.
-The family $\{A_i | i \in I\}$ is an indexed family of sets with index $i$ and indexing set $I$, and we can say $\mathcal A = \{A_i | i \in I\}$\cite{smith2006}.
+The family $\{A_i \st i \in I\}$ is an indexed family of sets with index $i$ and indexing set $I$, and we can say $\mathcal A = \{A_i \st i \in I\}$\cite{smith2006}.
 Then the union over this family of sets is
 \begin{equation}
-\bigcup_{i \in I} A_i = \bigcup_{A \in \mathcal A} A = \{x | x \in A \txt{ for } A \in \mathcal A\}.
+\bigcup_{i \in I} A_i = \bigcup_{A \in \mathcal A} A = \{x \st x \in A \txt{ for } A \in \mathcal A\}.
 \end{equation}
@@
 
-Only the variable, model nodes are allowed to be used in an MCG. In a FPF expression 
-nodes are also required. All four nodes are required for any SPF. In addition to 
-the three node types, there are two edge types:
+Only the variable, model, and expression nodes are allowed to be used in an MCG or FPF. 
+All four nodes are required for any SPF. In addition to the three node types, there are two edge types:
 
@@
 %\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_\txt{out} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
-\item[(3)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} | \txt{deg}^+(v) > 0\} }$
+\item[(3)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} \st \txt{deg}^+(v) > 0\} }$ ((maybe this should be 1 instead of 0 to match the previous definition of global input))
 \item[(4)] $\displaystyle{V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right)}$
@@
 \begin{equation}
-H = \{v \in V | t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
+H = \{v \in V \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
 \end{equation}
@@
 \begin{equation}
-C_{F,1} = C_{F,0} \setminus \{(x,y) \in E | x=E^3(v) \txt{ for } v \in H\},
+C_{F,1} = C_{F,0} \setminus \{(x,y) \in E \st x=E^3(v) \txt{ for } v \in H\},
 \end{equation}
 %deg-v is now recaculated for the new C. 
-and this step is demonstrated by Fig.~\ref{f:holes}. Because removing these edges can create new holes, this step must be repeated until no more holes are found.
+and this step is demonstrated by Fig.~\ref{f:hole}. Because removing these edges can create new holes, this step must be repeated until no more holes are found.
 \begin{figure}[htb!]
@@
 \caption{Example variable node indicating a hole.}
-\label{f:holes}
+\label{f:hole}
 \end{figure}
+
 \item The second step is to detect collisions and to then disconnect enough free edges so that the collisions are resolved. The set of variable nodes which contain collisions is created as
 \begin{equation}
-S_\txt{nodes} = \{v \in V | t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
+S_\txt{nodes} = \{v \in V \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
 \end{equation}
@@
 \begin{equation}
-S_\txt{edges} = \big \{ \{(x,y) \in E\} \big | y \in S_\txt{nodes} \big \}
+S_\txt{edges} = \big \{ \{(x,y) \in E\} \st y \in S_\txt{nodes} \big \}
 \end{equation}
 Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$.
+An examle collision is shown in Fig.~\ref{f:collision} to indicate the definition of $S_{\txt{edges},j}$.
 We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$.
@@
 \begin{equation}
-B_j = \big \{e_k \in S_{\txt{edges},j} \ | \ k \in \{1,2,\ldots,K\} \txt{ with } K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
+B_j = \big \{e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \txt{ with } K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
 \end{equation}
@@
 \begin{equation}
-C_{F,2} = \{ e \in C_{F,1} \ | \ e \in B_j \txt{ for some } j \in J\}.
+C_{F,2} = \{ e \in C_{F,1} \st e \in B_j \txt{ for some } j \in J\}.
 \end{equation}
+\begin{figure}[htb!]
+	\begin{center}
+	\includegraphics[width=3in]{images/analysis_block_collision}
+	\end{center}
+	\vspace{-20pt}
+\caption{Example variable node indicating a collision.}
+\label{f:collision}
+\end{figure}
+
 \item The third and final step is to remove any unused nodes and edges. The set of unused variable nodes representing local outputs is
 \begin{equation}
-V_\txt{empty} = \{v \in V_M \ | \ t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) = 0\},
+V_\txt{empty} = \{v \in V_M \st t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) = 0\},
 \end{equation}
@@
 \begin{equation}
-I_3 =  \{i \in I \ | \ \nexists v \in V_{A_i} \txt{ such that }  v \in V_\txt{empty} \}.
+I_F =  \{i \in I \ | \ \nexists v \in V_{A_i} \txt{ such that }  v \in V_\txt{empty} \}.
 \end{equation}
+The set of used global inputs is constructed as
+\begin{equation}
+V_{\txt{in},F} = \{v \in V_\txt{in} \st |\{(x,y) \in E(v) \st y \in V_{A_i} \txt{ for } i \in I_F  \}| > 1\},
+\end{equation}
+which requires that at least two edges directed out of the nodes in $V_{\txt{in},F}$ must be two the analysis blocks indexed by $I_F$.
+
 Finally, the set of nodes and edges describing a fundamental problem formulation is given by
 \begin{equation}
-V_{F,3} = V_M \setminus V_\txt{empty}
+V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right),
 \end{equation}
 \begin{equation}
-C_{F,3} = \{ (x,y) \in C_{F,2} \ | \ x \in V_{A_i} \txt{ or } y \in V_{A_i} \txt{ for some } i \in I_3 \}
+C_{F} = \{ (x,y) \in C_{F,2} \st x,y \in V_F\},
 \end{equation}
 \begin{equation}
-F_3 = (V_{F,3},C_{F,3})
+F = (V_F,C_F).
 \end{equation}

### 2013-03-19 graph_syntax.tex
added a new edge type

@@
 \begin{description}
-\item[variable] represents scalar or array data
-\item[model] responsible for mapping inputs to outputs
-\item[expression] simple algebreic combinations of variables (such as for constraint of objective functions)
-\item[driver] constrol structures capable of managing iteration
+\item[variable: ] represents scalar or array data
+\item[model:] responsible for mapping inputs to outputs
+\item[expression:] simple algebreic combinations of variables (such as for constraint of objective functions)
+\item[driver:] constrol structures capable of managing iteration
 \end{description}
 
-Only the variable, model, and expression nodes are allowed to be used in an MCG or FPF. 
-All four nodes are required for any SPF. In addition to the three node types, there are two edge types:
+In addition to the four node types, there are three edge types:
 
 \begin{description}
-\item[fixed edge] A fixed edge may never be removed from the graph.
-\item[free edge] A free edge may be removed removed from the graph.
+\item[fixed edge:] Fundamental, non-changeable property of a problem, such as 
+  connections between related variable nodes. These 
+edges can never be removed from a graph. 
+\item[explicit edge:] Explicit exchange of information between two nodes. No node 
+  can have more than a single incomming explicit edge. These edges are free and
+  can be added or removed from the graph. 
+\item [implicit edge:] Implicit passing of information from a driver node to a 
+  variable node. A single variable node can have many incomming implicit edges. These edges are 
+  free and can be added or removed from the graph. 
 \end{description}
 
-A number of restrictions are placed on the model node with respect to the number and 
-type of edges that can be connected to it: 
+The driver node and the implicit edge are only allowed to be pressent in SPF graphs. All 
+other node and edge types can be present in in any of the three graphs, subject to a 
+few restrictions: 
 \begin{enumerate}
@@
 \item A model node can only have fixed edges directed in or out.
-\item A model node must have at least one edge directed in and at least one edge directed out.
-\item If a variable node with an outgoing edge to a model node may not have any other outgoing edges.
+\item A model node must have at least one edge directed in and at least one edge 
+  directed out.
+\item If a variable node has an outgoing edge to a model node then it may not have 
+  any other outgoing edges.
 \end{enumerate}
@@
 
-Design variables, objectives, and constraints all share the property, that they are 
-require external decisions to be made about a given system to identify them. Consider 
-a notional problem given by the graph in Fig. [insert ref]a. There are three input variable 
-nodes which, could potentially become design variables: w, x, and y. Assume that the 
-output y is connected to the input y, yeilding Fig. [insert ref]b. Now only w and x could 
-be design variables. We define a design variable to be any input variable node that does 
-not have an incoming connection.
+Design variables, objectives, and constraints all require external decisions 
+to be made about a given system to identify them. Consider 
+a notional problem given by the graph in Fig. \ref{f:design vars} There are three input variables 
+which, could potentially become design variables: $w$, $x$, and $y$. Because the output, $x$, of $A$
+is connected to the input, $x$, of $B$ x can't actually be a design variable. The inputs $w$ and $y$
+have no incomming edges and could each be selected as design variables.
 
-[[INSERT GRAPH HERE]]
+\begin{figure}[htb!]
+  \begin{center}
+    \includegraphics[width=.6\textwidth]{images/design_vars_graph}
+  \end{center}
+  \caption{Simple notional problem with two potential design variables \label{f:design vars}}
+\end{figure}
+
+More formally, all inputs that have no incomming fixed or explicit edges are defined
+as design variables.
 

### 2013-03-19 graph_syntax.tex
more updates to the designvars,constraints,objectives section

@@
 More formally, all inputs that have no incomming fixed or explicit edges are defined
-as design variables.
+as design variables. When working with a MCG or FPF, only fixed and explicit edges are allowed, 
+in which case a design variable is simply a variable node with no incomming edges. For a SPF 
+any design variable node would have at least one incomming implicit edge from a driver node.
 
@@
 athrimetic opperations to form the final objective function. For example, the problem from 
-Fig. [insert ref] might have an objective function of $z^2$ or $2y + z$. Constraints are 
+Fig. \ref{f:design vars} might have an objective function of $z^2$ or $2y + z$. Constraints are 
 always given in the form of either an inequality of an equality, e.g. 
@@
 expression nodes. These nodes contain the actual mathematical expression used to define the 
-objective or constraint value. You can distinguish between an objective or constraint node 
+objective or constraint value. Objective or constraint nodes are distinguished
 via the presence of a simple expression (objective) vs an inequality/equality (constraint). 
-Fig. [insert new ref] shows the notional problem graph augmented with objective and 
+Fig. [insert ref] shows the notional problem graph augmented with objective and 
 constraint values.

### 2013-03-19 graph_syntax.tex
typo
changed all instances of 'free' to explicit

@@
 MDAO problems require that information be passed between sets of analyses. When information from the output of one analysis 
-block is passed, or connected, to the input of another analysis block, a new free edge is added 
-connecting the two variable nodes involved in the exchange. The edge is free because, unlike the edges 
-within an analysis block, it could be added or removed depending on the demands of the specific problem.
+block is passed, or connected, to the input of another analysis block, a new explicit edge is added 
+connecting the two variable nodes involved in the exchange. Explicit edges are free because, unlike the edges 
+within an analysis block, they could be added or removed depending on the demands of the specific problem. 
+In other words, explicit edges are created by engineers linking the output of one tool to the input 
+of another one.
 
@@
 $\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer explictly identifies variables 
-or groups of variables and assigns an simple, explicit relationship between them.
+or groups of variables and assigns a simple, explicit relationship between them.
 
-For the proposed syntax objectives and constraints are represented through the use of 
+For the proposed syntax, objectives and constraints are represented through the use of 
 expression nodes. These nodes contain the actual mathematical expression used to define the 
@@
 edges directed in & edges directed out & node representation \\ \hline
-only free edges & none & objective or constraint\\
-none & at least two free edges & global input \\
+only explicit edges & none & objective or constraint\\
+none & at least two explicit edges & global input \\
 none & one fixed edge & design variable \\ \hline
@@
 The codes are represented by analysis blocks $A_i=(V_{A_i},E_{A_i}), \ i\in \{1,2,\ldots,m\}=I$, the global inputs are represented a set of variable nodes $V_\txt{in}$, and the global outputs are represented by a set of variable nodes $V_\txt{out}$.
-We assume that $V_\txt{in}$, $V_\txt{out}$, and $A_i$ are given, and that any potential connection between variables is given in the form of the free edges in the set $C_M$.
+We assume that $V_\txt{in}$, $V_\txt{out}$, and $A_i$ are given, and that any potential connection between variables is given in the form of the explicit edges in the set $C_M$.
 Then we may construct the maximal connectivity graph $M=(V_M,E_M)$ as
@@
 
-The set $C_F$ is a set containing free edges representing connections.
+The set $C_F$ is a set containing explicit edges representing connections.
 Since $C_M$ is the set of all potential connections, we must have $C_F \subset C_M$.
@@
 In general, there may be multiple different graphs that satisfy the FPF conditions, though there may be none at all.
-Here, we describe a process for obtaining an FPF by starting with the MCG and disconnecting free edges until the FPF conditions are met.
-Then the problem is reduced to deciding which free edges to remove.
+Here, we describe a process for obtaining an FPF by starting with the MCG and disconnecting explicit edges until the FPF conditions are met.
+Then the problem is reduced to deciding which explicit edges to remove.
 
@@
 \begin{enumerate}
-\item The first step is to detect holes and disconnect the free edges following them.
-These free edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs.
+\item The first step is to detect holes and disconnect the explicit edges following them.
+These explicit edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs.
 The set of variable nodes which are holes is created as
@@
 which is the set of variable nodes with few incoming edges than are allowed by the lower indegree limit.
-Then the updated set of edges is created by removing the first set of free edges following the hole:
+Then the updated set of edges is created by removing the first set of explicit edges following the hole:
 \begin{equation}
@@
 
-\item The second step is to detect collisions and to then disconnect enough free edges so that the collisions are resolved. The set of variable nodes which contain collisions is created as
+\item The second step is to detect collisions and to then disconnect enough explicit edges so that the collisions are resolved. The set of variable nodes which contain collisions is created as
 \begin{equation}

### 2013-03-19 graph_syntax.tex
a bit more re-org, moved the indegree stuff into the basic graph syntax

@@
 \begin{itemize}
-    \item Maximal Connectivity
-    \item Fundamental Problem Formulation 
-    \item Specific Solution Formulation
+    \item Maximal Connectivity Graph (MCG)
+    \item Fundamental Problem  (FPG)
+    \item Problem Solution Graph (SPG)
 \end{itemize}
 
-The \emph{maximal connectivity graph} represents the first phase with all the 
-analysis codes being considered with all possible connections between them also present. The second graph 
+The \emph{maximal connectivity graph} represents the first phase with all 
+analysis codes being considered and all possible connections between them also present. The second graph 
 is \emph{fundamental problem formulation graph}, which is the smallest possible graph 
@@
 The tree diagram demonstrates the fact that it is generally possible to obtain 
-multiple FPFs from a single maximal connectivity graph. This  may correspond to 
+multiple FPGs from a single maximal connectivity graph. This  may correspond to 
 different down--selections of analysis codes, different connections between them, 
-or both. Each down--selection shrinks the number of possible FPFs that could be reached 
-untill only one is remaining. Then, a single FPF, different SPFs may be obtained by implementing 
+or both. Each down--selection shrinks the number of possible FPGs that could be reached 
+untill only one is remaining. Then, from a single FPG, different PSGs may be obtained by implementing 
 different solution strategies. The hourglass shape in Fig. \ref{f:hourglass} illustrates how 
-the MCG gets reduced to a single FPF, then multiple possible SPFs exist to solve the problem.
-In other words the FPF is obtained from the MCG by removing nodes and edges, 
-and the SPF is obtained from the FPF by adding nodes and edges.
+the MCG gets reduced to a single FPG, then multiple possible PSGs exist to solve the problem.
+In other words the FPG is obtained from the MCG by removing nodes and edges, 
+and the PSG is obtained from the FPG by adding nodes and edges.
 
@@
     }
-\caption{The relationship between the MCG, FPF, and SPF.}
+\caption{The relationship between the MCG, FPG, and PSG.}
 \end{figure}
 
-In the following sections we present some graph theory basics, then demonstrate how to build 
-the relevant MDAO constructs from the given grah primatives. Next we show how to construct 
-the MCG, FPF, and SPF graphs from those constructs and demonstrate methods to transition between 
-them.
+%the relevant MDAO constructs from the given grah primatives. Next we show how to construct 
+%the MCG, FPG, and SPF graphs from those constructs and demonstrate methods to transition between 
+%them.
 
-\subsection{Graph Theory Basics}
+\section{Syntax Definition}
+In the rest of this section we present the necessary graph theory fundamentals to 
+construct any of the different graphs discussed above. 
 The notation used in this work is adapted from Diestel \cite{Diestel2010}. 
@@
 Let $I$ be a nonempty set such that for each $i \in I$ there is a corresponding set $A_i$. 
-The family $\{A_i \st i \in I\}$ is an indexed family of sets with index $i$ and indexing set $I$, and we can say $\mathcal A = \{A_i \st i \in I\}$\cite{smith2006}.
+The family $\{A_i \st i \in I\}$ is an indexed family of sets with index $i$ and 
+indexing set $I$, and we can say $\mathcal A = \{A_i \st i \in I\}$\cite{smith2006}. 
 Then the union over this family of sets is
+
 \begin{equation}
@@
 
-% \subsection{Problem Formulation Concepts to Represent}
-% The MDAO problem formulation concepts we wish to represent are:
-% \begin{description}
-% \item[global input] A global input is a variable that is taken as given and is used by multiple analyses.
-% \item[design variable]
-% \item[function call] A function call requires a set of inputs to be fulfilled and has certain properties, such as run time, which are independent of the number of outputs actually being used.
-% \item[passing of distinct variables] Instead of connecting analyses, the individual variables as are connected
-% \item[objective \& constraint] These are variables that must be produced by the problem formulation for it to be valid
-% \item[collision] A collision occurs when an input has multiple sources
-% \item[hole] A hole occurs when an input has no sources
-% \item[coupling] Couple is the mutual dependence between a set of analysis
-% \item[multi--fidelity] It may be desirable for the same variable to be calculated by separate analyses and to use both results.
-% \end{description}
-
-\section{MDAO Graph Primatives}
 There are four node types:  
@@
 \item[model:] responsible for mapping inputs to outputs
-\item[expression:] simple algebreic combinations of variables (such as for constraint of objective functions)
+\item[expression:] simple algebreic combinations of variables (such as for 
+  constraint of objective functions)
 \item[driver:] constrol structures capable of managing iteration
@@
 
-The driver node and the implicit edge are only allowed to be pressent in SPF graphs. All 
-other node and edge types can be present in in any of the three graphs, subject to a 
+[[Insert legend for node and edge types]]
+
+The driver node and the implicit edge are only allowed to be pressent in PSGs. All 
+other node and edge types can be present in any of the three graphs, subject to a 
 few restrictions: 
@@
   any other outgoing edges.
+\item If a variable node has any incomming explicit edges, it may not have any 
+  incomming implicit edges. 
 \end{enumerate}
+
+For node $v \in V$ the edges directed out are given by $E(v)$ and the edges 
+directed into $v$ are given by $E^{-1}(v)$; $E(E(v))$ is denoted as $E^2(v)$, 
+and likewise for additional levels. The \emph{indegree} of a node is the number 
+of explicit edges directed in and is denoted as $\txt{deg}^-(v)$, and the \emph{outdegree} 
+is the number of explicit edges directed out and it is denoted as $\txt{deg}^+(v)$.
+We also define the \emph{upper indegree limit}
+
+\begin{equation}
+\txt{deg}_u^-(v):V \to \mathbb{N}
+\end{equation} 
+and the \emph{lower indegree limit}
+\begin{equation}
+\txt{deg}_l^-(v):V \to \mathbb{N}.
+\end{equation}
+
+These user-specified limits govern the number of explicit edges that may be directed into a 
+node for a valid graph. Consider a variable node $v$ with
+$\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, that node must have one, and only one
+incomming explicit edge or the graph is not valid. Two possiblities exists for voilating 
+the limit conditions: 
+\begin{description}
+  \item[hole: ] $\txt{deg}_u^-(v) < \txt{deg}_l^-(v)$. The number of incomming edges 
+  is less than $\txt{deg}_l^-(v)$. 
+  \item[collision: ]$\txt{deg}_u^-(v) > \txt{deg}_u^-(v)$. The number of incomming edges 
+  is greater than $ \txt{deg}_u^-(v)$.
+\end{description}
+
+\section{Graph Representation of MDAO Features}
 
@@
 More formally, all inputs that have no incomming fixed or explicit edges are defined
-as design variables. When working with a MCG or FPF, only fixed and explicit edges are allowed, 
-in which case a design variable is simply a variable node with no incomming edges. For a SPF 
+as design variables. When working with a MCG or FPG, only fixed and explicit edges are allowed, 
+in which case a design variable is simply a variable node with no incomming edges. For a valid PSG 
 any design variable node would have at least one incomming implicit edge from a driver node.
@@
 
-For node $v \in V$ the edges directed out are given by $E(v)$ and the edges directed into $v$ are given by $E^{-1}(v)$; $E(E(v))$ is denoted as $E^2(v)$, and likewise for additional levels. 
-The \emph{indegree} of a node is the number of edges directed in and is denoted as $\txt{deg}^-(v)$, and the \emph{outdegree} is the number of edges directed out and it is denoted as $\txt{deg}^+(v)$.
-We also define the \emph{upper indegree limit} 
-\begin{equation}
-\txt{deg}_u^-(v):V \to \mathbb{N}
-\end{equation} 
-and the \emph{lower indegree limit}
-\begin{equation}
-\txt{deg}_l^-(v):V \to \mathbb{N}.
-\end{equation}
-These user-specified limits set the number of edges that may be directed into a node for a problem formulation to be valid.
-For example, a variable node $v$ will have $\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$, unless it is a multi--fidelity variable, in which case the user may specify some upper limit higher than one.
-
-To keep track of which nodes and edges are which type, let $T_\txt{node}$ and $T_\txt{edge}$ be sets containing the possible node types and edge types, respectively, and then define mappings $t_\txt{node}:V \to T_\txt{node}$ and $t_\txt{edge}:E \to T_\txt{edge}$ to assign a type to each node and edge.
-Then, for example, for a variable node $v$ we have $t_\txt{node}(v) = \txt{`variable'}$.
-
- % A \emph{path} is a nonempty graph $P = (V,E)$ with $V = \{x_0,x_1,\ldots,x_k\}$ and $E = \{x_0x_1,x_1x_2,\ldots,x_{k-1}x_k\}$, where $x_i$ are all distinct.
-A graph $G$ is \emph{connected} if any two of its vertices are linked by a path in $G$.
-A \emph{tree} is a connected and acyclic graph.
-%%% this needs to be changed to fit the definition using a directed graph
-
-\section{Building MCG, FPF, and SPF Graphs}
+\section{Building MCG, FPG, and PSG}
 
@@
 The MCG $M$ is uniquely determined by the given set of analysis blocks, the required outputs, and the given global inputs.
-In the cases where the set of global inputs $I$ is not known a priori, the process of obtaining the FPF will reveal the required inputs, as discussed subsequently.
+In the cases where the set of global inputs $I$ is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.
 
@@
 \subsection{Obtaining the Fundamental Problem Formulation Graph}
-In general, there may be multiple different graphs that satisfy the FPF conditions, though there may be none at all.
-Here, we describe a process for obtaining an FPF by starting with the MCG and disconnecting explicit edges until the FPF conditions are met.
+In general, there may be multiple different graphs that satisfy the FPG conditions, though there may be none at all.
+Here, we describe a process for obtaining an FPG by starting with the MCG and disconnecting explicit edges until the FPG conditions are met.
 Then the problem is reduced to deciding which explicit edges to remove.
 
-%Then, mathematically, the FPF starts with $\mathcal A_1 = \{1,\ldots,m\}$, $I_{F,1} = I$, and $C_{F,1} = C_M$.
-Then, mathematically, the FPF starts with $C_{F,0} = C_M$.
+%Then, mathematically, the FPG starts with $\mathcal A_1 = \{1,\ldots,m\}$, $I_{F,1} = I$, and $C_{F,1} = C_M$.
+Then, mathematically, the FPG starts with $C_{F,0} = C_M$.
 First address the nodes where $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ then the ones where $\txt{deg}^-(v) > \txt{deg}_u^-(v)$
@@
         \item *Special* nodes for solvers and optimizers that *break* loops (from an algorithmic point of view)
-        \item FPF represents the minimal amount of information necessary to define a problem
+        \item FPG represents the minimal amount of information necessary to define a problem
         \item Any solution path grows the graph complexity by adding edges and nodes (or possibly have an empty solution graph, which you build up

### 2013-03-21 graph_syntax.tex
wrote design vars and local vs global sections

@@
 The ultimate goal of the new graph--based syntax presented here is to be able to 
-fully describe the fundamental problem formulation for a complex design task, 
+fully describe the general structure of an MDAO problem indpendently of any solution information, 
 while still being able to accomodate the more specific case when a solution 
-strategy is applied to the problem formulation. In order to achieve that goal 
-the graph needs to accomodate a number of constructs from MDAO problems:
+strategy is applied. In order to achieve that goal 
+the graph needs to accomodate a number of constructs of MDAO problems:
 
@@
 untill only one is remaining. Then, from a single FPG, different PSGs may be obtained by implementing 
-different solution strategies. The hourglass shape in Fig. \ref{f:hourglass} illustrates how 
+different solution strategies. If you consider the size of a graph to be the sum of all its
+edges and nodes then the hourglass shape in Fig. \ref{f:hourglass} illustrates how
 the MCG gets reduced to a single FPG, then multiple possible PSGs exist to solve the problem.
@@
 indicated by the first entry and directed to the node indicated by the second 
-entry. Edge $e$ = $(x,y)$ may be referred to simply as $xy$ and we say $y = E(x)$.
+entry. Edge $e$ = $(x,y)$ may be referred to simply as $xy$. For node $v \in V$ 
+the edges directed out are given by $E(v)$ and the edges directed into $v$ are given 
+by $E^{-1}(v)$; $E(E(v))$ is denoted as $E^2(v)$, and likewise for additional levels.
 
@@
 
+%does this belong in the analysis block section???? 
 Let $I$ be a nonempty set such that for each $i \in I$ there is a corresponding set $A_i$. 
@@
 \end{equation}
+
 Lastly, the cardinality of a set $B$ is the number of elements in $B$ and is denoted as $|B|$.
+
+\subsection{Node and Edge Types}
 
@@
   constraint of objective functions)
-\item[driver:] constrol structures capable of managing iteration
+\item[driver:] control logic capable of managing iteration
 \end{description}
@@
 
+\subsection{Rules for Nodes and Edges}
+
 The driver node and the implicit edge are only allowed to be pressent in PSGs. All 
@@
 
-For node $v \in V$ the edges directed out are given by $E(v)$ and the edges 
-directed into $v$ are given by $E^{-1}(v)$; $E(E(v))$ is denoted as $E^2(v)$, 
-and likewise for additional levels. The \emph{indegree} of a node is the number 
-of explicit edges directed in and is denoted as $\txt{deg}^-(v)$, and the \emph{outdegree} 
-is the number of explicit edges directed out and it is denoted as $\txt{deg}^+(v)$.
-We also define the \emph{upper indegree limit}
+\subsection{In-degree and Out-degree}
+
+The \emph{in-degree} of a node is the number of fixed or explicit edges directed in and 
+is denoted as $\txt{deg}^-(v)$, and the \emph{out-degree} 
+is the number of fixed or explicit edges directed out and it is denoted as $\txt{deg}^+(v)$.
+We also define the \emph{upper in-degree limit}
 
@@
 \end{equation} 
-and the \emph{lower indegree limit}
+and the \emph{lower in-degree limit}
 \begin{equation}
@@
 \begin{description}
-  \item[hole: ] $\txt{deg}_u^-(v) < \txt{deg}_l^-(v)$. The number of incomming edges 
-  is less than $\txt{deg}_l^-(v)$. 
-  \item[collision: ]$\txt{deg}_u^-(v) > \txt{deg}_u^-(v)$. The number of incomming edges 
-  is greater than $ \txt{deg}_u^-(v)$.
+  \item[hole: ] The number of incomming edges is less than $\txt{deg}_l^-(v)$. 
+    \begin{equation} \txt{deg}_u^-(v) < \txt{deg}_l^-(v) \label{e:hole} \end{equation}
+
+  \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
+    \begin{equation} \txt{deg}_u^-(v) > \txt{deg}_u^-(v) \end{equation}
 \end{description}
 
-\section{Graph Representation of MDAO Features}
+\section{Graph Representation of MDAO Constructs}
 
@@
 the analysis block contains two model nodes, with a single edge connecting them. 
-This part of the graph represents the necessary calculations to map given inputs 
-to proper output values. Although analysis blocks are comprised of multiple nodes and edges, since all 
-the edges within them are fixed, they become fixed structures within a graph.
+This edge represents the necessary calculations to map given inputs 
+to proper output values. If constructing a weighted graph, the weight calculation edge 
+provides the opportunity to computational cost. However, it is also allowed to skip 
+this edge and have both incomming and outgoing edges to variables through a single model 
+node for a given analysis block. Although analysis blocks are comprised of multiple nodes and edges, since all 
+the edges within them are fixed, they become fixed sub-graphs within a graph.
 
@@
 In other words, explicit edges are created by engineers linking the output of one tool to the input 
-of another one.
+of another one. Note that it is allowed for a single output to have outgoing explicit edges to multiple 
+other inputs. In Fig [insert ref], an output from analysis block A is connected 
+to the inputs for both blocks B and C.
+
+[[insert graph fig here]]
 
@@
 
-More formally, all inputs that have no incomming fixed or explicit edges are defined
-as design variables. When working with a MCG or FPG, only fixed and explicit edges are allowed, 
+More formally, any input variable that can be defined as a hole by Eqn. \ref{e:hole}
+would be considered a design variable in the graph. When working with a MCG or FPG, only fixed and explicit edges are allowed, 
 in which case a design variable is simply a variable node with no incomming edges. For a valid PSG 
-any design variable node would have at least one incomming implicit edge from a driver node.
+any design variable node would have at least one (but possible multiple) incomming 
+implicit edge from a driver node.
 
@@
 \end{comment}
-\subsection{Local vs Global Properties}
+\subsection{Local vs Global Nodes}
 
-\begin{itemize}
-  \item what can be local or global? Variables, objectives, constraints, 
-  \item What does it really mean to be local vs global? 
-  \item What does the graph structure of local vs global look like? 
-  \item local is a frame of reference... a variable could be global when looking at two analyses, but local to the two when looking at the three
-\end{itemize}
+A simple definition of a global node is any node that involves information 
+from more than one analysis block. Likewise, a local node is any node 
+that involves information from only a single analysis block. As stated before, variable 
+nodes with fixed edges directed to model nodes (as part of an analysis block) are inherently local 
+and similarly model nodes are also local. Expression and driver nodes
+are not inherently local or global though. For these, their locality is determined 
+by the incomming and outgoing edges they posses. Consider the graph in Fig [insert ref], 
+where the expression node $g$ represents some kind of constraint for the optimziation problem. 
+Since $g$ only had edges comming from analysis block $B$, it would be considered a local constraint 
+on block $B$.
+
+[[insert graph here]]
+
+Although we've already established that variable nodes in an analysis block are local to that 
+block, global design variables are a part of many MDAO problems. We've previously defined 
+a design variable as any
+
+While this simple definition is valid, it glosses over a more subtle aspect of 
+some MDAO problems. Consider the graph in Fig. [insert ref], with three analysis 
+blocks A, B, and C. The expression node, $g$, is a constraint that requires information 
+from $A$ and $B$, so it would be considered global by the previous definition. Since $g$
+has incomming edges from block $C$, it would not be global with respect to $C$. 
+You could say that $g$ is local to the group $A,B$. In this sense, locality 
+becomes a relative term. A given node can be local to a specific sub-graph of a large
+problem. This subtle structure can become significant when determining how to partition 
+an MDAO problem for use with certain MDAO architectures. Martins and Marriage 
+demonstrate how Collaborative Optimization outperforms MultiDisciplinary Feasible
+when they separated a problem into two sub-problems with variables, objectives, and 
+constraints local to each one\cite{Marriage2008a}.
 
@@
 Requirement (3) stipulates that only the global inputs that are being used should be included.
-Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper indegree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.
+Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper in-degree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.
 
@@
 \end{equation}
-which is the set of variable nodes with few incoming edges than are allowed by the lower indegree limit.
+which is the set of variable nodes with few incoming edges than are allowed by the lower in-degree limit.
 Then the updated set of edges is created by removing the first set of explicit edges following the hole:
@@
 \end{equation}
-which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper indegree limit of $v_j$.
+which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper in-degree limit of $v_j$.
 The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include.

### 2013-03-21 intro.tex
a few more re-org changes.
wrote design vars and local vs global sections

@@
 
-\section{Existing Graph--Based Problem Formulation Syntax}
+\section{Existing Graph--Based Syntax}
 
@@
     Of the three methods, XDSM comes the closest to fully describing a problem formulation. 
-    Thought it does not contain specific variable information, this could be easily added back in
-    simply by putting the variables back into the graph and drawing the apprpriat connections
+    Thought it does not contain specific variable information on the diaganol of the DSM, this could be easily added back in
+    simply by putting the variables back into the graph and drawing the appropriate connections
     between them and the relevant analysis nodes. However, it's dependence on 
     solution strategy specifics to represent couplings can not be removed without 
-    stripping it of the useful information contained in an FDT. In the following sections
-    we describe a new graph syntax that retains the features of XDSM, without the dependence on 
-    solution strategy (although it does allow for that information to be present). Thus 
-    the new syntax allows for the compelete specification of an FPF.
+    stripping it of the useful information contained in an FDT. We propose
+    a new graph syntax that retains the features of XDSM, without the dependence on 
+    solution strategy (although it does allow for that information to be present). This 
+    new syntax allows for the compelete specification of an MDAO problem.

### 2013-03-22 graph_syntax.tex
added a graph for constraints and objectives

@@
   \end{center}
-  \caption{Simple notional problem with two potential design variables \label{f:design vars}}
+  \caption{Notional graph with two potential design variables \label{f:design vars}}
 \end{figure}
@@
 via the presence of a simple expression (objective) vs an inequality/equality (constraint). 
-Fig. [insert ref] shows the notional problem graph augmented with objective and 
+Fig. \ref{f:obj-cons} shows the notional problem graph augmented with objective and 
 constraint values.
 
-[[INSERT ANOTHER GRAPH HERE]]
+\begin{figure}[htb!]
+  \begin{center}
+    \includegraphics[width=.6\textwidth]{images/obj_const_graph}
+  \end{center}
+  \caption{Notional graph with and objective and constraint nodes \label{f:obj-cons}}
+\end{figure}
 
@@
 \end{comment}
+
 \subsection{Local vs Global Nodes}

### 2013-03-28 graph_syntax.tex
Edits to the wording and math to improve readability

@@
 %does this belong in the analysis block section???? 
-Let $I$ be a nonempty set such that for each $i \in I$ there is a corresponding set $A_i$. 
-The family $\{A_i \st i \in I\}$ is an indexed family of sets with index $i$ and 
-indexing set $I$, and we can say $\mathcal A = \{A_i \st i \in I\}$\cite{smith2006}. 
-Then the union over this family of sets is
-
+Let $I$ be a nonempty set of counting numbers such that for each $i \in I$ there is a corresponding set $A_i$. 
+The set of sets $\mathcal A = \{A_i \st i \in I\}$ is called an indexed family of sets with index $i$ and 
+indexing set $I$\cite{smith2006}. 
+The union over this family of sets can be described in a few different ways:
 \begin{equation}
-\bigcup_{i \in I} A_i = \bigcup_{A \in \mathcal A} A = \{x \st x \in A \txt{ for } A \in \mathcal A\}.
+\bigcup_{i \in I} A_i = \bigcup_{A \in \mathcal A} A = \{x \st x \in A \txt{ for some } A \in \mathcal A\}.
 \end{equation}
@@
 \end{equation}
-
 These user-specified limits govern the number of explicit edges that may be directed into a 
 node for a valid graph. Consider a variable node $v$ with
-$\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, that node must have one, and only one
-incomming explicit edge or the graph is not valid. Two possiblities exists for voilating 
+$\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, that node must have exactly one
+incomming explicit edge or the graph is invalid. Two possiblities exist for voilating 
 the limit conditions: 
@@
   \item[hole: ] The number of incomming edges is less than $\txt{deg}_l^-(v)$. 
-    \begin{equation} \txt{deg}_u^-(v) < \txt{deg}_l^-(v) \label{e:hole} \end{equation}
+    \begin{equation} \txt{deg}^-(v) < \txt{deg}_l^-(v) \label{e:hole} \end{equation}
 
   \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
-    \begin{equation} \txt{deg}_u^-(v) > \txt{deg}_u^-(v) \end{equation}
+    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \end{equation}
 \end{description}
@@
 \subsection{Maximal Connectivity Graph}
-To construct the maximal connectivity graph, we assume that a set of codes, global inputs, and objectives and constraints (collectively called global outputs).
-The codes are represented by analysis blocks $A_i=(V_{A_i},E_{A_i}), \ i\in \{1,2,\ldots,m\}=I$, the global inputs are represented a set of variable nodes $V_\txt{in}$, and the global outputs are represented by a set of variable nodes $V_\txt{out}$.
-We assume that $V_\txt{in}$, $V_\txt{out}$, and $A_i$ are given, and that any potential connection between variables is given in the form of the explicit edges in the set $C_M$.
+To construct the maximal connectivity graph, we assume that a set of codes, global inputs, and objectives and constraints (collectively called global outputs) are provided.
+The codes are represented by analysis block graphs ((ABGs?)) $A_i=(V_{A_i},E_{A_i}), \ i\in I$, where $m$ is the number of codes and $I=\{1,2,\ldots,m\}$.
+The global inputs are represented as a set of variable nodes $V_\txt{in}$, and the global outputs are represented by a set of variable nodes $V_\txt{out}$.
+We assume that $V_\txt{in}$, $V_\txt{out}$, and each $A_i$ are given, and that any potential connection between variables is given in the form of explicit edges in the set $C_M$. 
 Then we may construct the maximal connectivity graph $M=(V_M,E_M)$ as
@@
 The MCG $M$ is uniquely determined by the given set of analysis blocks, the required outputs, and the given global inputs.
-In the cases where the set of global inputs $I$ is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.
+In the cases where the set of global inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.
 
 \subsection{Fundamental Problem Formulation Graph}
-We now define the fundamental problem formulation graph, $F=(V_F,E_F)$, as a directed graph meeting the following conditions
+We now define the fundamental problem formulation graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions, which are explained subsequently,
 \begin{enumerate}
-\item[(1)] $\displaystyle{I_F \subset I}$ such that the following conditions hold
-\item[(2)] $\displaystyle{ \forall i \in I, \ \exists v \in V_{A_i} \txt{ with } t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v)>0 }$
+\item[(1)] $\displaystyle{I_F \subset I}$ such that the remaining conditions hold
+\item[(2)] $\displaystyle{ \forall i \in I_F, \ \exists v \in V_{A_i} \txt{ with } t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v)>0 }$
 %\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_{A_i} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
@@
 The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). 
-Requirement (2) stipulates that the analysis blocks in $F$ must each have at least one local output that is beging used.
+Requirement (2) stipulates that each analysis blocks in $F$ must have at least one local output that is being used. 
 Requirement (3) stipulates that only the global inputs that are being used should be included. 
+Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
 Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper in-degree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.
@@
 Since $C_M$ is the set of all potential connections, we must have $C_F \subset C_M$.
-While no other conditions are explicitly stated for $C_F$, requirement (6) is actually a requirement on both $I$ and $C_F$.
+While no other conditions are explicitly stated for $C_F$, requirement (6) is actually a requirement on both $I_F$ and $C_F$.
 
@@
 %Then, mathematically, the FPG starts with $\mathcal A_1 = \{1,\ldots,m\}$, $I_{F,1} = I$, and $C_{F,1} = C_M$.
-Then, mathematically, the FPG starts with $C_{F,0} = C_M$.
-First address the nodes where $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ then the ones where $\txt{deg}^-(v) > \txt{deg}_u^-(v)$
-
+Then, mathematically, the FPG starts with $C_{F,0} = C_M$, $E_{F,0} = E_M$, $V_{F,0} = V_M$.
+The nodes where $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ are considered first.
 \begin{enumerate}
@@
 \begin{equation}
-H = \{v \in V \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
+H = \{v \in V_M \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
 \end{equation}
-which is the set of variable nodes with few incoming edges than are allowed by the lower in-degree limit.
+which is the set of variable nodes with fewer incoming edges than are allowed by the lower in-degree limit.
 Then the updated set of edges is created by removing the first set of explicit edges following the hole:
 \begin{equation}
-C_{F,1} = C_{F,0} \setminus \{(x,y) \in E \st x=E^3(v) \txt{ for } v \in H\},
+C_{F,1} = C_{F,0} \setminus \{(x,y) \in E_M \st x=E_M^3(v) \txt{ for } v \in H\},
 \end{equation}
@@
 \begin{equation}
-S_\txt{nodes} = \{v \in V \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
+S_\txt{nodes} = \{v \in V_M \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
 \end{equation}
@@
 \begin{equation}
-S_\txt{edges} = \big \{ \{(x,y) \in E\} \st y \in S_\txt{nodes} \big \}
+S_\txt{edges} = \big \{ \{(x,y) \in E_M\} \st y \in S_\txt{nodes} \big \}
 \end{equation}
 Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$. 
-An examle collision is shown in Fig.~\ref{f:collision} to indicate the definition of $S_{\txt{edges},j}$.
-We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$.
+An example collision is shown in Fig.~\ref{f:collision} to indicate the definition of $S_{\txt{edges},j}$. 
+We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$ (which are sets). 
 Then we may construct sets of edges as
@@
 \end{equation}
-which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper in-degree limit of $v_j$.
+which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper in--degree limit of $v_j$.
 The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include.
@@
 \begin{equation}
-C_{F,2} = \{ e \in C_{F,1} \st e \in B_j \txt{ for some } j \in J\}.
+C_{F,2} = \{ e \in C_{F,1} \st e \in B_j \txt{ for some } j \in J\},
+\end{equation}
+\begin{equation}
+E_{F,2} = E_M \setminus (C_M \setminus C_{F,2})
+\end{equation}
+and
+\begin{equation}
+F_2 = (V_M,E_{F,2})
 \end{equation}
@@
 
-\item The third and final step is to remove any unused nodes and edges. The set of unused variable nodes representing local outputs is
+\item The third and final step is to remove any unused nodes and edges. The indexing set for only the analysis blocks with at least one used local output may be constructed as
 \begin{equation}
-V_\txt{empty} = \{v \in V_M \st t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) = 0\},
+I_F =  \{i \in I \ | \ \exists v \in V_{A_i} \txt{ such that }  t_\txt{node}(E_{F,2}^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) > 0 \},
 \end{equation}
-and so then the indexing set for only the analysis blocks with at least one used local output may be constructed as
-\begin{equation}
-I_F =  \{i \in I \ | \ \nexists v \in V_{A_i} \txt{ such that }  v \in V_\txt{empty} \}.
-\end{equation}
+where the degree is calculated with respect to $F_2$. This set excludes both the analysis blocks that were unused in the original MCG and those that became unused in step 1 or step 2.
+
 The set of used global inputs is constructed as
 \begin{equation}
-V_{\txt{in},F} = \{v \in V_\txt{in} \st |\{(x,y) \in E(v) \st y \in V_{A_i} \txt{ for } i \in I_F  \}| > 1\},
+V_{\txt{in},F} = \Big\{v \in V_\txt{in} \ \Big| \ |\{(x,y) \in E_{F,2}(v) \st y \in V_{A_i} \txt{ for } i \in I_F  \}| > 1 \Big\},
 \end{equation}
-which requires that at least two edges directed out of the nodes in $V_{\txt{in},F}$ must be two the analysis blocks indexed by $I_F$.
+which requires that at least two edges directed out of the nodes in $V_{\txt{in},F}$ must be to the analysis blocks indexed by $I_F$.
 
@@
 \begin{equation}
-F = (V_F,C_F).
+E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right),
+\end{equation}
+\begin{equation}
+F = (V_F,E_F).
 \end{equation}

### 2013-04-01 graph_syntax.tex
staring work on a monday morning

--- 2013-03-22 graph_syntax.tex
@@
 
-\subsection{Obtaining the Fundamental Problem Formulation Graph}
+\subsection{Obtaining the Fundamental Problem Graph}
 In general, there may be multiple different graphs that satisfy the FPG conditions, though there may be none at all.

### 2013-04-01 example_problem.tex

\section{Example Problem}
This section presents an example problem to demonstrate the process of obtaining the FPG and the advantages of doing so. The task is the conceptual sizing of a single--aisle subsonic transport using an MDAO framework with objectives for performance and total weight for a required mission. The full set of analysis codes available for use is given in Table \ref{t:analysis codes}.
\begin{table}[htbp]
  \centering
  \caption{Analysis codes for subsonic transport sizing}
    \begin{tabular}{cc}
    \toprule
    analysis code & description \\
    \midrule
    VSP   & parametric geometry \\
    PDCYL & wing and fuselage weight estimation \\
    NPSS  & engine sizing and performance \\
    VORLAX & aerodynamics using the vortex lattice method \\
    PMARC & aerodynamics using a low-order panel method \\
    WATE  & engine weight estimation \\
    FLOPSa & mission performance and \\
      & engine sizing and weight estimation \\
    FLOPSb & mission performance only \\
    \bottomrule
    \end{tabular}%
  \label{t:analysis codes}%
\end{table}%
Each of these analysis codes contribute individual disciplinary analysis but they are not mutually exclusive. For example, VORLAX and PMARC are both aerodynamics codes that predict inviscid drag but with different levels of fidelity. The Flight Optimization System (FLOPS) is included twice to represent two different uses. FLOPSa denotes FLOPS being executed for both mission performance and engine analysis, while FLOPSb indicates that FLOPS is executed for only mission performance analysis.

In this example, the MCG is formed using a consistant naming convention and then connecting every variables with the same name, though this is not the only way to form an MCG. Table \ref{t:ins and outs} presents the full list of variables in the leftmost column and then indicates whether the variable is an input or an output of each analysis code. Some of the variables, such as geometry and performance, represent many different variables which have been bundled together due to there similarity. This bundling is not necessary and does not limit the generality of this example, it is done to simplify the presentation.
\begin{table}[htb!]
  \centering
  \caption{Analysis code input and output description}
    \begin{tabular}{ccccccccc}
    \toprule
    variable & \multicolumn{8}{c}{analysis code} \\
    \midrule
          & VSP   & PDCYL & NPSS  & VORLAX & PMARC & WATE  & FLOPSa & FLOPSb \\
    geometry & in    & in    &       & in    & in    & in    & in    & in \\
    number of engines &       &       & in    &       &       &       & in    & in \\
    mission &       &       &       &       &       &       & in    & in \\
    fuselage weight &       & out   &       &       &       &       & in    & in \\
    wing weight &       & out   &       &       &       &       & in    & in \\
    engine weight &       &       &       &       &       & out   & out   & in \\
    wetted area & out   &       &       &       &       &       & in    & in \\
    inviscid drag &       &       &       & out   & out   &       &       & in \\
    drag  &       &       & in    &       &       &       & out   & out \\
    engine performance &       &       & out   &       &       & in    & out   & in \\
    performance &       &       &       &       &       &       & out   & out \\
    total weight &       & in    &       &       &       &       & out   & out \\
    \bottomrule
    \end{tabular}
  \label{t:ins and outs}
\end{table}

The MCG $M$ is formed in four steps:
\begin{enumerate}
\item An analysis block is created for each analysis code using the information in each column of Table \ref{t:ins and outs}. Each analysis block is formed by creating variable nodes for each input and directing them into a single model node. This model node is then directed into another model node, which is then directed into a variable node for each output. A sample analysis block is shown for analysis code FLOPSb in Fig.~\ref{f:FLOPSb analysis block}.
\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=.6\textwidth]{images/FLOPSb_analysis_block}
  \end{center}
  \caption{Sample analysis block for analysis code FLOPSb using Table \ref{t:ins and outs}}
\label{f:FLOPSb analysis block}
\end{figure}

\item Variable nodes are created to represent the objectives for performance and total weight.

\item Variable nodes are created to represent the geometry variable and mission variable global inputs.

\item Explicit edges are created from each variable node to every other variable node representing variables with the same name.
\end{enumerate}
The resulting MCG is shown in Fig.~\ref{f:design vars}. Emerging cycles are shown in yellow and indicate potential cycles that may exist when an FPG is formed.
\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=.6\textwidth]{images/MCG}
  \end{center}
  \caption{Maximal connectivity graph for the subsonic transport example problem}
\end{figure}

\subsection{Obtaining an FPG}
As presented in Sec.(()), the first step to obtaining an FPG is to detect holes and, to start, we set the lower indegree limit for every node to be one. In this case, it turns out that the variable nodes representing the number of engines is a hole for NPSS, FLOPSa, and FLOPSb. This implies than an FPG cannot be obtained from the given MCG. Therefore, we may now create another variable node to represent the number of engines as a global input and also create explicit edges to supply this new input. Now this step is repeated with the new MCG and no holes are found.

While Sec.(()) indicated the process for obtaining an FPG, it did not specify how to decide which edges to keep when resolving a collision, that freedom is left to the specific implementation. The benefit of this approach is that standard graph theory algorithms can be used to make these decisions. The first example of this is to obtain an FPG with the fewest possible number of cycles, which are found using the implementation of Johnson's algorithm \cite{Johnson1975} in the Python package NetworkX. The resulting FPG is shown in Fig.~\ref{f:FPG fewest cycles} and the only cycle is shown in yellow. A similar alternative would be to minimize the number of analysis blocks involved in cycles, counting multiplicity. While resulting graph has only one cycle, it may be more desirable to include the higher fidelity codes, such as NPSS.
\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=.6\textwidth]{images/FPG_fewest_cycles}
  \end{center}
  \caption{FPG with the fewest number of cycles.}
\label{f:FPG fewest cycles}
\end{figure}

One method to ensure that the edges directed from certain analysis blocks are the ones chosen to resolve the conflicts is to use a ranking system. For this case, consider the rankings for each analysis code given in Table \ref{t:rankings}.
\begin{table}[htbp]
  \centering
  \caption{Example ranking of importance}
    \begin{tabular}{cc}
    \toprule
    analysis code & rank \\
    \midrule
    VSP   & 5 \\
    PDCYL & 5 \\
    NPSS  & 4 \\
    PMARC & 4 \\
    FLOPSb & 4 \\
    VORLAX & 3 \\
    WATE  & 3 \\
    FLOPSa & 2 \\
    \bottomrule
    \end{tabular}%
  \label{t:rankings}%
\end{table}% 
Step two in Sec.(()) (the resolution of collisions) is now conducted by selecting the edges directed from the analysis blocks with the highest rank for each collision. The resulting FPG is shown in Fig.~\ref{f:FPG highest rank} and this graph has four cycles, which are enumerated in Table \ref{t:cycles}.
\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=.6\textwidth]{images/FPG_highest_rank}
  \end{center}
  \caption{Simple notional problem with two potential design variables \label{f:design vars}}
\label{f:FPG highest rank}
\end{figure}
\begin{table}[htbp]
  \centering
  \caption{Cycles for the FPG obtained using ranking}
    \begin{tabular}{ccccccc}
    \toprule
    output/input & analysis code & output/input & analysis code & output/input & analysis code & output/input \\
    \midrule
    engine weight & FLOPSb & drag  & NPSS  & engine performance & WATE  & engine weight \\
    engine performance & FLOPSb & drag  & NPSS  & engine performance &       &  \\
    fuselage weight & FLOPSb & total weight & PDCYL & fuselage weight &       &  \\
    wing weight & FLOBSb & total weight & PDCYL & wing weight &       &  \\
    \bottomrule
    \end{tabular}%
  \label{t:cycles}
\end{table}%

Finally, now consider the case where it is desired to have the inviscid drag input into FLOPSb be a multi--fidelity input. This is done simply by setting the upper indegree limit for that node to two and repreating the (automated) process. In this case, both VORLAX and PMARC are retained, resulting in an FPG that omits only FLOPSa.

%\begin{figure}[htb!]
%  \begin{center}
%    \includegraphics[width=.6\textwidth]{images/FPG_highest_rank_simplified}
%  \end{center}
%  \caption{Simple notional problem with two potential design variables \label{f:design vars}}
%\end{figure}

%\begin{table}[h!]
% \begin{center}
%  \caption{Example problem analysis codes}
%  \label{t:analysis codes}
%  \begin{tabular}{ccc} \hline 
%Analysis code & input variables & output variables \\ \hline
%VSP & geometry & wetted area \\
%PDCYL & geometry & fuselage weight \\
%	& total weight & wing weight \\
%NPSS & number of engines & engine performance \\
%	& drag &	\\
%VORLAX & geometry & inviscid drag \\
%PMARC & geometry & inviscid drag \\
%WATE & geometry & engine weight \\
%	& engine performance &	\\ 
%FLOPSa & number of engines & engine weight \\
%	& mission & performance \\
%	& geometry & total weight \\
%	& wetted area & engine performance \\
%	& fuselage weight & drag \\
%FLOPSb & number of engines & drag \\
%	& mission & performance \\
%	& geometry & total weight \\
%	& wetted area & 	 \\
%	& engine weight & 	\\
%	& wing weight & 	\\
%	& inviscid drag & 	\\
%	& engine performance & 	\\ \hline
%  \end{tabular}
% \end{center}
% \vspace{-15pt}
%\end{table}

### 2013-04-02 graph_syntax.tex
updates to the intro with brian's stuff

@@
   any other outgoing edges.
-\item If a variable node has any incomming explicit edges, it may not have any 
-  incomming implicit edges. 
 \end{enumerate}
@@
 to be made about a given system to identify them. Consider 
-a notional problem given by the graph in Fig. \ref{f:design vars} There are three input variables 
+a notional problem given by the graph in Fig. \ref{f:designvars} There are three input variables 
 which, could potentially become design variables: $w$, $x$, and $y$. Because the output, $x$, of $A$
@@
   \end{center}
-  \caption{Notional graph with two potential design variables \label{f:design vars}}
+  \caption{Notional graph with two potential design variables \label{f:designvars}}
 \end{figure}
@@
 athrimetic opperations to form the final objective function. For example, the problem from 
-Fig. \ref{f:design vars} might have an objective function of $z^2$ or $2y + z$. Constraints are 
+Fig. \ref{f:designvars} might have an objective function of $z^2$ or $2y + z$. Constraints are 
 always given in the form of either an inequality of an equality, e.g. 

### 2013-04-02 intro.tex
updates to the intro with brian's stuff

@@
 \section{Introduction}
+
+    The number of analysis tools required in multidisciplinary design optimization (MDO) studies is growing
+    in parallel with the increasing scope of typical MDO problems. An example of this growth in scope can
+    be observed in the historical evolution of the disciplines involved in MDO problems in aircraft design.
+    Multidisciplinary optimization emerged as a separate field from structural optimization through the
+    need to introduce formal techniques for managing the coupling of aerodynamic loads and structural
+    deformations, through the linking of aerodynamic vortex lattice or panel methods with structural finite
+    element models [[REF]]. Subsequently, flight performance and life cycle economics tools were integrated
+    into MDO analysis workflows for conceptual and preliminary design studies [[REF]]. Currently, MDO
+    problems for aircraft design also often include tools for aircraft noise and emissions. In sum, it is now
+    commonplace for 5-10 analysis tools to be employed for typical aircraft design optimization studies [[REF]].
+
+    The number of analysis tools is expected to grow in the future as the scope of MDO problems continues
+    to evolve, and as computing is increasingly commoditized. This expansion of scope will be driven, in
+    part, by consideration of additional disciplines. Current trends on the horizon for aircraft MDO studies
+    include incorporation of manufacturing analyses [[REF]], subsystem performance [[REF]], and fleet-level
+    aggregate models of emissions, noise, or economics [[REF]].
 
@@
     and paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. That 
-    new capability has created a new challenge since configuring a model with the larger nubmer of analyses, associated inputs and outputs
-    can be very difficult. Even for models with 10's of analyses, there could be hundreds of variables that need to be managed. 
-    It is not hard to imagine that the task of combining all the analyses into a consistent system model capable of solving a relevant 
-    engineering design problem could easily become much more costly than createing the discipline analyses themselves. Thus, for a large system
-    as the couplings between the disciplines begin to dominate the design space, the couplings between the analyses begin to dominate 
-    the job of setting up the model.
+    new capability has created a new challenge. Even for models with 10's of analyses there could be hundreds or thousands
+    of variables that need to be linked together. Commonly different analyses will provide competing values for the 
+    same physical quantity, and these conflicts need to be resolved somehow. To make matters worse, the competeing values 
+    may not even be in the same format. These occurrences are particularly acute when analysis tools have differing fidelities. For
+    example, an abstracted aerodynamic analysis such as an empirical drag buildup model may return only
+    integrated drag, whereas a CFD tool may return pressure and shear stress distributions across the entire
+    surface grid. If an analysis downstream of the aerodynamics tool needs only integrated drag as an input,
+    then the designer has a free choice of which of the two possible aerodynamics analysis tools to select to
+    provide the drag estimate (presuming that drag can be computed from surface distributions by a simple
+    integration algorithm). On the other hand, if a downstream analysis needs a pressure distribution in
+    order to compute pitching moment, for example, then any feasible MDO problem formulation must
+    include CFD or similar analysis in the data flow, regardless of whether the empirical drag buildup model
+    is also included.
 
-    When building a model, coupling represents the reciprical flow of information between two analyses. 
-    We propose the use of graph-based syntax to describe the flow of information from inputs, 
-    through analyses, to outputs. If you allow for any free inputs to become design
-    variables and some set of the outputs to become design objectives and constraints, then the graph 
-    represents the problem formulation for a given design effort. 
-    The graph-based syntax presented has application in all phases of the design process. 
-    It can be used early on, to help construct a valid problem formulation from the set of available analysis tools. 
-    It also provides a consistent structure to describe problem formulation that provides a foundation to apply
-    algorithmic methods for selecting and then implementing effective solution strategies inside an MDAO framework.
+    With all the added complexity from larger models it is not hard to imagine that the task of combining all the analyses into a 
+    consistent system model capable of solving a relevant engineering design 
+    problem could easily become much more costly than createing the discipline 
+    analyses themselves. Just as for a large system the couplings between the 
+    disciplines begin to dominate the design space, the couplings between the 
+    analyses begin to dominate the job of setting up the model. It is this problem of 
+    determining sets of analysis tools and their interconnectivities to form realistic
+    multidisciplinary problems that is the subject of this paper. We are motivated by the following notional
+    but real problem of organizing an MDO study for a complex system:
 
-    We've already identifed that a problem formulation graph should include design variables, analysis tools and their 
-    associated inputs and outputs, and objectives and constraints. Notably absent from that list are any kind of 
-    solvers, optimizers, or other iterative solution finding tools. At its most fundamental state, a problem formulation 
-    includes only information about what is being sought after in a design or what the
-    goals of a given problem are. It need not contain any information about solution paths or strategies to reach those goals. 
-    While solution information can be represented with the graph syntax presented here, a unique aspect of this syntax is that 
-    such information always remains seperable from the underlying problem formulation. In other words, a given a graph representing 
-    a specific solution strategy for a specific problem formulation, it is possible to remove the solution information from the 
-    graph and reclaim the more basic problem formulation.
+    A new system is being designed for which there is little or no historical precedent. The system
+    is complex, as measured by the number of disciplines and/or components involved in the
+    analysis. A general optimization problem statement has been formulated based on system-
+    level objectives and constraints; however, it is unclear which engineering analysis tools should
+    be interconnected in order to solve the optimization problem. A team of disciplinary and/
+    or component design engineers has been formed in which each engineer has expertise in a
+    particular analysis tool or component model. The engineers meet to discuss the approach to
+    interconnecting their tools to achieve the required system-level MDO model.
+
+    Our goal is to develop formalism for expressing analysis interconnectivity and for determining feasible
+    data flows to assist an engineering team conducting this task. Because the problem deals with
+    interconnectivity, we base our approach on the representations and techniques of graph theory. The 
+    proposed graph syntax is closely related to the REMS system proposed by 
+    Alexandrov and Lewis\cite{alexandrov2004}. The approach taken here should be viewed as an extention 
+    of the concepts proposed in REMS with the goal of allowing the graphs to more closely interact with 
+    existing design processes and MDAO frameworks. The approach begins by 
+    constructing the \emph{maximal connectivity graph (MCG)} describing all possible
+    interconnections between the analysis tools proposed by the engineers. Graph operations are then
+    conducted reduce the MCG down to a \emph{fundamental problem graphs (FPG)} that describes the minimum set of analysis
+    tools needed to solve the specified system-level design problem. The FPG does not predispose
+    any particular solution procedure for the problem; any relevant MDO solution architecture, e.g. MDF,
+    IDF, CO, BLISS, could be selected \emph{post facto} to implement the system-level optimization by
+    constructing a \emph{problem solution graph (PSG)} and the corresponding problem solution approach.
+    The concept of the FPG and the identification of feasible FPGs are the main contributions of the paper.
+
+    The paper is organized as follows. First, we describe the differences between a fundamental problem
+    formulation, which is based only on the system-level optimization problem statement that the
+    engineers desire to solve and on the available analysis tools, and a specific problem formulation, which
+    additionally presumes a specific solution to the problem. Next, we survey the literature in applying
+    graph theoretic and formal language approaches to multidisciplinary design problem formulation. 
+    We then discuss our graph syntax and representation of MDO problems and describe the procedures for 
+    determining the MCG and FPG. Finally, we present an example problem based on an MDO analysis of a 
+    commercial aircraft and discuss additional applications enabled by our approach.
 
@@
     two constraints which both must be met. The presence of both of these constraints fully decouples the problem so that 
-    either $A$ or $B$ could be run first or both could be run simultaneously. By removing either constraint and replacing 
-    it with a direct dependence between the two analyses you could regain the earlier two problem formulations. Alexandrov
-    and Lewis demonstrated the value of a more modular approach to problem formulation because it enables one to transition between different 
-    MDAO solution strategies depending on the specifics of the problem\cite{Alex2000}.
+    either $A$ or $B$ could be run first or both could be run simultaneously. Alexandrov and Lewis investigated problem 
+    formulation in this fully decoupled way as part of their work on 
+    Reconfigurability in MDO Problem Synthesis (REMS) \cite{alexandrov2004}.
+    In this work they show how removing either constraint and replacing 
+    it with a direct dependence between the two analyses, a process they term ``collapsing'', 
+    you could regain the earlier two problem formulations.
 
-\section{Existing Graph--Based Syntax}
+\section{Existing Graph-Based Syntax}
 
@@
     number of optimization architectures\cite{Tedford2009}. Their work demonstrates clearly how multiple specific 
-    problem formulations can all relate back to a common FPF. The challenge with using this 
+    problem formulations can all relate back to a common FPG. The challenge with using this 
     traditional mathematical syntax is that it is not easily manipulated or analyzed. 
@@
 
-    Lamb and Martins included objectives and constraint functions as nodes in an Extended 
+    Alexandrov and Lewis introduced a graph based syntax called REMS which 
+    incorporates objectives and constraints into a graph, effectively combining 
+    FDT and DSM\cite{alexandrov2004}. REMS retains the square adjacency 
+    matrix form from DSM, but by adding in the new elements it partially 
+    combined a traditional DSM with an FDT. This allows REMS to represent data 
+    dependency between multiple analysis tools as well as between analysis tools and
+    objective/constraint functions. One significant conclusion from REMS is 
+    the need for variables that are local to discipline analyses. This allows 
+    all information exchange to be stated as equality constraints, which become 
+    edges in the graph. Additionaly, REMS address the need to be able to 
+    transition between multiple solution strategies while keeping a constant 
+    graph representation of the fundamental problem formulation. REMS only provides
+    syntax for variables and functions, but does not provide for inclusion 
+    of solvers or optimizers in the graph.
+
+    Lamb and Martins also included objectives and constraint functions as nodes in an Extended 
     DSM (XDSM) in order to capture a more complete description of solution strategies for MDAO problems\cite{Lambe2012}. 
-    XDSM retains the square adjacency matrix form from DSM, but by adding in the 
-    new elements they partially combined a traditional DSM with an FDT. 
-    This allows XDSM to represent data dependency between multiple analysis tools as well as between analysis tools and
-    objective/constraint functions. With the additional information included in an 
-    XDSM, Lu and Martins applied both ordering and partitioning 
-    algorithms on an MDAO test problem named the Scalable Problem \cite{Lu2012}.
+    Unlike REMS, XDSM also includes nodes for solvers and optimziers to enable complete definition of 
+    MDAO architectures. Martins uses XDSM to describe 13 different optimziation architectures in 
+    a survey paper that provides a novel classification of the different 
+    techniques \cite{Lambe2011}. With the additional information included in an 
+    XDSM, Lu and Martins applied both ordering and partitioning algorithms on an 
+    MDAO test problem named the Scalable Problem \cite{Lu2012}.
 
@@
 
-    Of the three methods, XDSM comes the closest to fully describing a problem formulation. 
-    Thought it does not contain specific variable information on the diaganol of the DSM, this could be easily added back in
-    simply by putting the variables back into the graph and drawing the appropriate connections
-    between them and the relevant analysis nodes. However, it's dependence on 
-    solution strategy specifics to represent couplings can not be removed without 
-    stripping it of the useful information contained in an FDT. We propose
-    a new graph syntax that retains the features of XDSM, without the dependence on 
-    solution strategy (although it does allow for that information to be present). This 
-    new syntax allows for the compelete specification of an MDAO problem.
+    Of all the methods, REMS comes the closest to fully describing a problem formulation,
+    but lacks the ability to represent optimziers and solvers in more specific problem 
+    formulations. While XDSM does include optimizers and constraints, it relies on 
+    their presence and can't be used for a more fundamental problem statement. 
+    We propose a new graph syntax that combines the key features of REMS and XDSM
+    and provides new language to enable more complete description of MDAO problems.

### 2013-04-04 graph_syntax.tex
technical edits and theoretical comments

@@
 the graph needs to accomodate a number of constructs of MDAO problems: 
-
 \begin{itemize}
@@
 picking an MDAO optimization architecture. Using the proposed graph syntax, each of these phases 
-can be represented with a related graph.
-
+can be represented with the following three graphs:
 \begin{itemize}
@@
 analysis codes being considered and all possible connections between them also present. The second graph 
-is \emph{fundamental problem formulation graph}, which is the smallest possible graph 
+is the \emph{fundamental problem formulation graph}, which is the smallest possible graph 
 that still fully defines a given problem formulation. Finally, a \emph{specific problem formulation} 
@@
 or both. Each down--selection shrinks the number of possible FPGs that could be reached 
-untill only one is remaining. Then, from a single FPG, different PSGs may be obtained by implementing 
-different solution strategies. If you consider the size of a graph to be the sum of all its
+until only one is remaining. Then, from a single FPG, different PSGs may be obtained by implementing 
+different solution strategies. If you consider the size of a graph to be the sum of all of its
 edges and nodes then the hourglass shape in Fig. \ref{f:hourglass} illustrates how
@@
 contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
-For a \emph{directed graph*} we construct $E$ as a set of ordered pairs instead 
+For a \emph{directed graph*} (or \emph{digraph}) we construct $E$ as a set of ordered pairs instead 
 of a set of sets. Each ordered pair represents an edge starting at the node 
@@
 by $E^{-1}(v)$; $E(E(v))$ is denoted as $E^2(v)$, and likewise for additional levels. 
-
+If $E$ is not a one--to--one mapping, $E(v)$ may be the empty set, a single node, or a set of nodes.
 \begin{figure}[htb!]
@@
 \subsection{Node and Edge Types}
-
+We now define different types of nodes and edges with specific properties to compose graphs used to represent an MDAO problem. 
 There are four node types:  
@@
 \item[model:] responsible for mapping inputs to outputs
-\item[expression:] simple algebreic combinations of variables (such as for 
-  constraint of objective functions)
+\item[expression:] simple algebraic combinations of variables (such as for 
+  constraint or objective functions)
 \item[driver:] control logic capable of managing iteration
@@
 In addition to the four node types, there are three edge types: 
-
 \begin{description}
@@
 \item[explicit edge:] Explicit exchange of information between two nodes. No node 
-  can have more than a single incomming explicit edge. These edges are free and
+  can have more than a single incomming explicit edge ((?)). These edges are free and
   can be added or removed from the graph. 
@@
 \subsection{Rules for Nodes and Edges}
-
-The driver node and the implicit edge are only allowed to be pressent in PSGs. All 
-other node and edge types can be present in any of the three graphs, subject to a 
-few restrictions: 
+There are specific rules for the usage of these nodes and edges.
+The driver node and the implicit edge are only allowed to be present in PSGs. All 
+other node and edge types can be present in any of the three graphs, subject to these restrictions: 
 \begin{enumerate}
@@
 \subsection{In-degree and Out-degree}
-
 The \emph{in-degree} of a node is the number of fixed or explicit edges directed in and 
 is denoted as $\txt{deg}^-(v)$, and the \emph{out-degree} 
-is the number of fixed or explicit edges directed out and it is denoted as $\txt{deg}^+(v)$.
-We also define the \emph{upper in-degree limit}
-
+is the number of fixed or explicit edges directed out and is denoted as $\txt{deg}^+(v)$.
+We also define the \emph{upper in-degree limit} as
 \begin{equation}
@@
 \end{equation} 
-and the \emph{lower in-degree limit}
+and the \emph{lower in-degree limit} as
 \begin{equation}
@@
 \end{equation}
-These user-specified limits govern the number of explicit edges that may be directed into a 
+These user-specified limits govern the number of explicit edges that may be directed into a variable
 node for a valid graph. Consider a variable node $v$ with
-$\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, that node must have exactly one
+$\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, $v$ must have exactly one
 incomming explicit edge or the graph is invalid. Two possiblities exist for voilating 
-the limit conditions: 
+these limit conditions: 
 \begin{description}
-  \item[hole: ] The number of incomming edges is less than $\txt{deg}_l^-(v)$. 
+  \item[hole: ] The number of incomming edges is less than the lower indegree limit:
     \begin{equation} \txt{deg}^-(v) < \txt{deg}_l^-(v) \label{e:hole} \end{equation} 
-
-  \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
+  \item[collision: ] The number of incomming edges is greater than the upper indegree limit:
     \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \end{equation}
 \end{description} 
+These violations will be managed in order to seek a working problem formulation.
 
 \section{Graph Representation of MDAO Constructs}
-
+%As discussed in Sec.((intro)), the digraph must be able to describe the individual aspects of an MDAO problem.
 \subsection{Analysis Blocks and Connections}
-
-Analysis tools take in a number input variables, then perform some work to calculate 
+Analysis tools take in a number input variables and then perform some work to calculate 
 the values for their respective outputs. In an MDAO graph, this process is 
-represented by an group of nodes and edges called an \emph{analysis block}, 
+represented by a group of nodes and edges called an \emph{analysis block}, 
 shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
@@
 Via the analysis block structure, we can also distinguish between input and ouptut
-variable nodes. Inputs are any variable nodes that have an outgoing edge into a model 
+variable nodes, as shown in Fig.~\ref{f:analysis block}. Inputs are any variable nodes that have an outgoing edge into a model 
 node. Conversely, outputs are any variable nodes that have an incoming edge from a model node. 
 Note that by this definition, inputs and outputs can only exist when variable nodes poses an 
-edge joining them to a model node. In Fig. \ref{f:analysis block} these are labeled out as 
-local inputs and local outputs. The variables are local because they are explicitely tied
+edge joining them to a model node. These variables are local because they are explicitely tied
 to the analysis block and have no relevance without it.
@@
 \subsection{Design Variables, Objectives, and Constraints}
-
 Design variables, objectives, and constraints all require external decisions 
@@
 a notional problem given by the graph in Fig. \ref{f:designvars} There are three input variables 
-which, could potentially become design variables: $w$, $x$, and $y$. Because the output, $x$, of $A$
-is connected to the input, $x$, of $B$ x can't actually be a design variable. The inputs $w$ and $y$
+which, could potentially become design variables: $w$, $x$, and $y$. Because the variable $x$ is connected from $A$ to $B$,
+ $x$ can't actually be a design variable. On the other hand, the inputs $w$ and $y$
 have no incomming edges and could each be selected as design variables.
-
 \begin{figure}[htb!]
@@
 More formally, any input variable that can be defined as a hole by Eqn. \ref{e:hole}
-would be considered a design variable in the graph. When working with a MCG or FPG, only fixed and explicit edges are allowed, 
+would be considered a design variable in the graph. When working with an MCG or FPG, only fixed and explicit edges are allowed, 
 in which case a design variable is simply a variable node with no incomming edges. For a valid PSG 
-any design variable node would have at least one (but possible multiple) incomming 
+any design variable node would have at least one incomming 
 implicit edge from a driver node.
 
-Objective and constraints are also constructs that need to be identified by engineers 
+((DP: I think we are struggling to define the exact boundary between a hole and design variable and when one can become the other.
+This stems from the fact that they are identical with respect to the graph structure: they are both variable nodes representing local inputs without any edges directed in.
+Therefore, I suggest we let the difference be strictly determined by the user--set indegree limits: given a node in question, it is a hole if the lower indegree limit is greater than zero. 
+Then, a few things can happen.
+A) the hole can become irrelevant if the analysis block is not needed.
+B) the hole can cause the graph to be invalid and the user can change the lower indegree limit or change the maximal connectivity graph to include a new edge (and presumably other new nodes and edges as well).
+C) the user could have set the \emph{upper} indegree limit to be zero, meaning that it is supposed to be a design variable.
+Having written this, I feel that (if you agree) we should probably have it in the paper somewhere)).
+
+Objectives and constraints are also constructs that need to be identified by engineers 
 as part of the design process. In the case of objective functions single output values 
@@
 athrimetic opperations to form the final objective function. For example, the problem from 
-Fig. \ref{f:designvars} might have an objective function of $z^2$ or $2y + z$. Constraints are 
-always given in the form of either an inequality of an equality, e.g. 
+Fig.~\ref{f:designvars} may have an objective function of $z^2$ or $2y + z$. Constraints are 
+always given in the form of either an inequality or an equality, e.g., 
 $\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer explictly identifies variables 
 or groups of variables and assigns a simple, explicit relationship between them. 
+((DP:I think we can probably just say ``alegbraic expression'' and not have to show any, especially since the variables aren't given meaning))
 
@@
 constraint values.
+
+((DP: I think this definition of objective/constraint is disharmoneous with our other node types. I think it should also have local inputs represented by variables nodes))
 
@@
 \subsection{Local vs Global Nodes}
-
 A simple definition of a global node is any node that involves information 
@@
 constraints local to each one\cite{Marriage2008a}. 
+((DP: it is not clear to me how the difference between local and global should change in the process of obtaining the FPG from the MCG. For example,
+in the MCG a node representing a global input may have several edges directed out to multiple analysis blocks, but in the FPG all but one have been removed. 
+From my point of view, I was thinking that the difference between global and local wasn't a node's current state, but what it has the potential to be.
+Another way of saying this is: the definition of global vs local shouldn't depend on explicit edges because they are free to be removed, unless we also want a node to be able to change from being global or local.
+Lastly, a global input in an FPG that only has one edge directed out could be removed and the node it was directed in to could become a design variable.))
+
+((DP: along these lines, what if a design variable is a separate variable node outside of the analysis block.
+Then it doesn't need any indegree limits and being global or local can be directly tied to the number of edges directed out?))
 

### 2013-04-04 graph_syntax.tex
working...

--- 2013-04-02 graph_syntax.tex
@@
 that still fully defines a given problem formulation. Finally, a \emph{specific problem formulation} 
-may be represented by including additional edges and node types to represent the 
+may be represented by including additional edges and nodes to represent the 
 solution strategy being employed to solve the problem.
@@
 
-In addition to the four node types, there are three edge types:
+In addition to the four node types, there are two edge types:
 
 \begin{description}
-\item[fixed edge:] Fundamental, non-changeable property of a problem, such as 
-  connections between related variable nodes. These 
-edges can never be removed from a graph. 
-\item[explicit edge:] Explicit exchange of information between two nodes. No node 
-  can have more than a single incomming explicit edge. These edges are free and
-  can be added or removed from the graph. 
+\item[connection edge:] Exchange of information between two nodes. These edges 
+can either be fixed (can not be removed from graph) or free (can be removed). 
 \item [implicit edge:] Implicit passing of information from a driver node to a 
@@
 [[Insert legend for node and edge types]]
+
+In Alexandrov and Lewis's REMS syntax only two node types were present, variable 
+and function, and only one edge type\cite{alexandrov2004}. We've chosen to rename the ``function'' node 
+type to ``model'' to be more consistent with modern MDAO frameworks. This work 
+adds two more node types and one more edge type to the syntax to allow descriptive
+graphs for all three phases of the design process.
 
@@
 \item A model node can only have one edge directed to or from another model node.
-\item A model node can only have fixed edges directed in or out.
+\item A model node can only have fixed connection edges directed in or out.
 \item A model node must have at least one edge directed in and at least one edge 
@@
 \subsection{In-degree and Out-degree}
+\label{s:indegree-outdegree}
+The \emph{in-degree} of a node is the number of connection edges directed in and 
+is denoted as $\txt{deg}^-(v)$, and the \emph{out-degree} 
+is the number of connection edges directed out and it is denoted as $\txt{deg}^+(v)$.
+The degree of a given node is only a function of the connection edges 
+attached to it. The number of implicit edges is not relevant since any number 
+of drivers could be involved, at different parts of an solution process. For 
+example in a sequential optimization where a global optimizer is first applied
+and a gradient based one is applied second, design variables would have incomming 
+implicit edges from both optimizers but this would not affect the in-degree of those
+variable nodes.
 
-The \emph{in-degree} of a node is the number of fixed or explicit edges directed in and 
-is denoted as $\txt{deg}^-(v)$, and the \emph{out-degree} 
-is the number of fixed or explicit edges directed out and it is denoted as $\txt{deg}^+(v)$.
 We also define the \emph{upper in-degree limit}
@@
 \end{equation}
-These user-specified limits govern the number of explicit edges that may be directed into a 
-node for a valid graph. Consider a variable node $v$ with
+These user-specified limits govern the number of fixed and connection edges that 
+may be directed into a node for a valid graph. Consider a variable node $v$ with
 $\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, that node must have exactly one
-incomming explicit edge or the graph is invalid. Two possiblities exist for voilating 
+incomming edge or the graph is invalid. Two possiblities exist for voilating 
 the limit conditions: 
@@
   \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
-    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \end{equation}
+    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \label{e:collision} \end{equation}
 \end{description}
@@
 edge joining them to a model node. In Fig. \ref{f:analysis block} these are labeled out as 
-local inputs and local outputs. The variables are local because they are explicitely tied
+local inputs and local outputs. The variables are local because they are explicitly tied
 to the analysis block and have no relevance without it.
 
-MDAO problems require that information be passed between sets of analyses. When information from the output of one analysis 
-block is passed, or connected, to the input of another analysis block, a new explicit edge is added 
-connecting the two variable nodes involved in the exchange. Explicit edges are free because, unlike the edges 
-within an analysis block, they could be added or removed depending on the demands of the specific problem. 
-In other words, explicit edges are created by engineers linking the output of one tool to the input 
-of another one. Note that it is allowed for a single output to have outgoing explicit edges to multiple 
-other inputs. In Fig [insert ref], an output from analysis block A is connected 
-to the inputs for both blocks B and C.
-
-[[insert graph fig here]]
+MDAO problems require that information be passed between sets of analyses. When 
+information from the output of one analysis block is passed, or connected, to the 
+input of another analysis block, a new connection edge is added connecting the two 
+variable nodes involved in the exchange. These connection edges are free, unlike the edges 
+within an analysis block, they could be added or removed depending on the needs
+of a given design problem. In other words, free connection edges are created by 
+engineers linking the output of one tool to the input of another one. Note that 
+it is allowed for a single output to have outgoing connection edges to multiple 
+downstream inputs.
 
@@
 Design variables, objectives, and constraints all require external decisions 
-to be made about a given system to identify them. Consider 
-a notional problem given by the graph in Fig. \ref{f:designvars} There are three input variables 
-which, could potentially become design variables: $w$, $x$, and $y$. Because the output, $x$, of $A$
-is connected to the input, $x$, of $B$ x can't actually be a design variable. The inputs $w$ and $y$
-have no incomming edges and could each be selected as design variables.
+to be made about a given system to identify them. Consider a notional problem 
+given by the graph in Fig. \ref{f:designvars} There are three input variables 
+which, could potentially become design variables: $w$, $x$, and $y$. Because the 
+output, $x$, of $A$ is connected to the input, $x$, of $B$ x can't actually be a 
+design variable. The inputs $w$ and $y$ have no incomming edges and could each 
+be selected as design variables.
 
@@
 More formally, any input variable that can be defined as a hole by Eqn. \ref{e:hole}
-would be considered a design variable in the graph. When working with a MCG or FPG, only fixed and explicit edges are allowed, 
-in which case a design variable is simply a variable node with no incomming edges. For a valid PSG 
-any design variable node would have at least one (but possible multiple) incomming 
-implicit edge from a driver node.
+would be considered a design variable in the graph. When working with a MCG or 
+FPG, only connection edges are allowed, in which case a design variable 
+is simply a variable node with no incomming edges. For a valid PSG 
+any design variable node would have at least one, but possibly multiple, incomming 
+implicit edges from one or more driver nodes.
 
@@
 always given in the form of either an inequality of an equality, e.g. 
-$\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer explictly identifies variables 
-or groups of variables and assigns a simple, explicit relationship between them.
+$\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer identifies variables 
+or groups of variables and assigns a simple relationship between them.
 
@@
 
-\begin{comment}
-The MDO problem formulation concepts represented by nodes are given in 
-Table \ref{t:node representation}, and the concepts represented by edges are 
-given in Table \ref{t:edge representation}.
-% \begin{itemize}
-% \item An objective or constraint is indicated by a variable node with no outgoing edges and with the incoming edges being directed from only other variable nodes.
-% \item A global input is a variable node with no edges directed in and with at least two edges directed out to different variable nodes.
-% \item A design variable is represented by a variable node with no edges directed in and on fixed edge directed out.
-% \end{itemize}
-\begin{table}[h!]
- \begin{center}
-  \caption{Problem formulation concept represented by a node}
-  \label{t:node representation}
-  \begin{tabular}{ccc} \hline 
-edges directed in & edges directed out & node representation \\ \hline
-only explicit edges & none & objective or constraint\\
-none & at least two explicit edges & global input \\
-none & one fixed edge & design variable \\ \hline
-  \end{tabular}
- \end{center}
- \vspace{-15pt}
-\end{table}
-\begin{table}[h!]
- \begin{center}
-  \caption{Problem formulation concept represented by an edge}
-  \label{t:edge representation}
-  \begin{tabular}{ccc} \hline 
-from node & to node & edge representation \\ \hline
-variable & variable & connection/passing of a variable\\
-variable & model & local input \\
-model & variable & local output \\
-model & model & function call \\ \hline
-  \end{tabular}
- \end{center}
- \vspace{-15pt}
-\end{table}
-%`Connection' edges connect the local outputs from one analysis block to input nodes.
-\end{comment}
-
 \subsection{Local vs Global Nodes}
@@
   are not inherently local or global though. For these, their locality is determined 
-by the incomming and outgoing edges they posses. Consider the graph in Fig [insert ref], 
-where the expression node $g$ represents some kind of constraint for the optimziation problem. 
-Since $g$ only had edges comming from analysis block $B$, it would be considered a local constraint 
-on block $B$.
+  by the incomming and outgoing edges they possess. Consider the graph in Fig. \ref{f:local_global}, 
+  where the expression node $g_1$ represents some kind of constraint for the optimziation problem. 
+  Since $g_1$ only has edges comming from analysis block $B$, it would be considered a local constraint 
+  on block $B$. $g_2$ has incomming edges from both $A$ and $B$ analysis blocks, so it would be a global constraint.
 
+  \begin{figure}
+      \begin{center}
+      %\includegraphics[height=.25\textheight]{XDSM/simple}
       [[insert graph here]]
+      \caption{A simple graph with local and global nodes \label{f:local_global}}
+      \end{center}
+  \end{figure}
 
   Although we've already established that variable nodes in an analysis block are local to that 
-block, global design variables are a part of many MDAO problems. We've previously defined 
-a design variable as any
+  block, global variables are a part of many MDAO problems and must be representable. A global variable 
+  would be any variable node with multiple outgoing connection edges. 
+  For example, in Fig. \ref{f:local_global} the global node $w^*$ has outgoing edges to 
+  the $w$ variable nodes feeding both $A$ and $B$ analysis blocks.
 
   While this simple definition is valid, it glosses over a more subtle aspect of 
-some MDAO problems. Consider the graph in Fig. [insert ref], with three analysis 
-blocks A, B, and C. The expression node, $g$, is a constraint that requires information 
-from $A$ and $B$, so it would be considered global by the previous definition. Since $g$
-has incomming edges from block $C$, it would not be global with respect to $C$. 
-You could say that $g$ is local to the group $A,B$. In this sense, locality 
+  some MDAO problems. Fig. \ref{f:local_global}, has three analysis 
+  blocks A, B, and C. Above we identified node $g_2$ as global,  but
+  it has no incomming edges from block $C$. So you could say that $g_2$ is 
+  local to the the set of blocks $A,B$. In this sense, locality 
   becomes a relative term. A given node can be local to a specific sub-graph of a large
-problem. This subtle structure can become significant when determining how to partition 
-an MDAO problem for use with certain MDAO architectures. Martins and Marriage 
-demonstrate how Collaborative Optimization outperforms MultiDisciplinary Feasible
-when they separated a problem into two sub-problems with variables, objectives, and 
-constraints local to each one\cite{Marriage2008a}.
+  problem. This subtle structure can become significant when determining how to 
+  effectively partition an MDAO problem into sub-problems 
+  \cite{krishnamachari1997optimal,allison2009optimal,michelena1997hypergraph}.
 
@@
 
-\subsection{Multi-fidelity Analyses}
+  Coupling exists in a design problem when a set of two or more analyses each depend on the
+  others outputs. Consider the notional problem from section \ref{sec:spg_vs_fpg} where two 
+  analyses, $A$ and $B$ are coupled through their respective variables. The graph of this 
+  is shown in Fig. \ref{f:coupling}. In the graph $A.z$ has an outgoing connection 
+  edge to $B.z$, indicating that there is a connection between the two. 
+  Similarly a connection exists between $B.y$ and $A.y$. These two connections form 
+  a cycle which represent the coupling between them. This cycle means that blocks $A$ and $B$ need
+  to be iterated on untill they converge.
 
-\section{Building MCG, FPG, and PSG}
+  \begin{figure}
+      \begin{center}
+      %\includegraphics[height=.25\textheight]{}
+      [INSERT GRAPH HERE]
+      \caption{Graph of a notional problem with simple coupling \label{f:coupling}}
+      \end{center}
+  \end{figure}
+
+  It is important to note that the cycle in graph in Fig. \ref{f:coupling} does not contain 
+  any driver nodes. That means that no solver, optimizer, or other iterative construct is 
+  invovled in the coupling. This absence is intentional, as the coupling is an inherant part 
+  of the problem. This corresponds directly To Eqn. \ref{eqn:simple_fpf}, where 
+  each equality constraint is one edge in the graph. Transitionining to either 
+  Eqn. \ref{eqn:simple} or Eqn. \ref{eqn:simple2} means selecting one of the nodes as the starting point 
+  for the iteration, indicating the associated analysis block should be executed first.
+
+  Larger problems can contain more complex cycles in thier FPG, indicating more 
+  complex coupling between analyses. A cycle can contain more than 
+  just two analysis blocks. Multiple independent cycles can exist, indicating 
+  independent coupling relationships. Cycles can also overlap, where the same analysis 
+  blocks are involved in multiple different coupling cycles. All of these situations
+  arise naturally as the size of problems grows, and managing all of this coupling 
+  becomes difficult. From the perspective of the graph, the problem reduces down to 
+  a ordering challenge. For each cycle in the graph, you need to pick a analysis block
+  to start the iteration from. Rogers DEMAID tool uses a genetic algorithm to find an ordering 
+  that minimizes the overall coupling of the system by separating independent 
+  cycles in the graph \cite{rogers1996,rogers1996demaid}.
+
+\subsection{Multi-fidelity Problems}
+
+  A multi-fidelity problem can be characterized by two or more different analyses 
+  each calculating the same data. Mapping this type of multi-fidelity situation 
+  onto a graph will yeild two analysis blocks, each with thier own output 
+  variable nodes, having outgoing edges to a third analysis block that needs the 
+  data as input. Figure \ref{f:collision-example} shows a notional problem 
+  that exhibits this structure, where the key feature is that variable node $y$ in 
+  analysis block $C$ has two incomming edges.
+
+  \begin{figure}
+      \begin{center}
+      %\includegraphics[height=.25\textheight]{}
+      [INSERT GRAPH HERE]
+      \caption{Graph of a notional multi-fidelity problem \label{f:collision-example}}
+      \end{center}
+  \end{figure}
+
+  When an input variable node has more than one incomming edge, designers must make
+  some kind of decision about how to manage the flow of information in their problem. 
+  In Eqn. \ref{e:collision} we defined a collision as the situation where
+  there are more incomming connection edges than the ${deg}_u^-(y)$. So desginers can
+  indicate how to handle conflicting edges in a problem graph by setting the ${deg}_u^-(y)$
+  for a given node. In more concrete terms, setting this limit equates to defining 
+  the part of the graph in question as single or multi-fidelity.
+
+  When ${deg}_u^-(y)=1$, then any conflicting edges in the graph will cause a collision
+  and a choice between the colliding analysis blocks must be made. This results in a 
+  single fidelity design problem. On the other hand if ${deg}_u^-(y)=2$, then two 
+  different analysis blocks can co-exist with a given graph without causing a 
+  collision. The design problem is then a multi-fidelity problem.
+
+  Multi-fidelity problems are always charaterized by graphs with variable 
+  nodes that have an upper-degree-limit greater than one. These problems require 
+  special techniques for resolving the confliciting edges by introducing some mechanism
+  to manage when each of the different fidelity analyses are 
+  run\cite{march2012provably,alexandrov2001approximation,Huang_Allen_Notz_Miller_2006}.
+
+\section{Building MCG and FPG}
 
@@
 The global inputs are represented as a set of variable nodes $V_\txt{in}$, and the global outputs are represented by a set of variable nodes $V_\txt{out}$.
-We assume that $V_\txt{in}$, $V_\txt{out}$, and each $A_i$ are given, and that any potential connection between variables is given in the form of explicit edges in the set $C_M$. 
+  We assume that $V_\txt{in}$, $V_\txt{out}$, and each $A_i$ are given, and that any potential connection between variables is given in the form of connection edges in the set $C_M$. 
   Then we may construct the maximal connectivity graph $M=(V_M,E_M)$ as
@@
 
-The set $C_F$ is a set containing explicit edges representing connections.
+The set $C_F$ is a set containing connection edges representing connections.
 Since $C_M$ is the set of all potential connections, we must have $C_F \subset C_M$.
@@
 In general, there may be multiple different graphs that satisfy the FPG conditions, though there may be none at all.
-Here, we describe a process for obtaining an FPG by starting with the MCG and disconnecting explicit edges until the FPG conditions are met.
-Then the problem is reduced to deciding which explicit edges to remove.
+Here, we describe a process for obtaining an FPG by starting with the MCG and disconnecting connection edges until the FPG conditions are met.
+Then the problem is reduced to deciding which connection edges to remove.
 
@@
 \begin{enumerate}
-\item The first step is to detect holes and disconnect the explicit edges following them.
-These explicit edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs.
+\item The first step is to detect holes and disconnect the connection edges following them.
+These connection edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs.
 The set of variable nodes which are holes is created as
@@
 which is the set of variable nodes with fewer incoming edges than are allowed by the lower in-degree limit.
-Then the updated set of edges is created by removing the first set of explicit edges following the hole:
+Then the updated set of edges is created by removing the first set of connection edges following the hole:
 \begin{equation}
@@
 
-\item The second step is to detect collisions and to then disconnect enough explicit edges so that the collisions are resolved. The set of variable nodes which contain collisions is created as
+\item The second step is to detect collisions and to then disconnect enough connection edges so that the collisions are resolved. The set of variable nodes which contain collisions is created as
 \begin{equation}
@@
 \end{enumerate}
-
-    \subsection{Solution Graph Syntax}
-    What is the difference between a problem formulation and a problem solution method? Convert from a cyclic graph, to an acyclic graph
-    \begin{itemize}
-        \item Cycles indicate convergence loops or design variable loops
-        \item Problem can't be solved until all loops are *removed* by adding solvers/optimizers
-        \item *Special* nodes for solvers and optimizers that *break* loops (from an algorithmic point of view)
-        \item FPG represents the minimal amount of information necessary to define a problem
-        \item Any solution path grows the graph complexity by adding edges and nodes (or possibly have an empty solution graph, which you build up
-        as you remove edges from problem formulation graph?)
-    \end{itemize}

### 2013-04-04 intro.tex
working...

@@
     include incorporation of manufacturing analyses [[REF]], subsystem performance [[REF]], and fleet-level
-    aggregate models of emissions, noise, or economics [[REF]].
+    aggregate models of emissions, noise, or economics [[REF]]. A recent survey by Shan and Wang covered 
+    a range of different techniques proposed to solve High Dimentional, Expensive, Black-box (HEB) problems. 
+    The large amount of relevant research they identified demonstrates the growing interest large scale 
+    MDO problems that arise from multiple disciplines\cite{Shan2010}.
 
@@
     interconnectivity, we base our approach on the representations and techniques of graph theory. The 
-    proposed graph syntax is closely related to the REMS system proposed by 
-    Alexandrov and Lewis\cite{alexandrov2004}. The approach taken here should be viewed as an extention 
-    of the concepts proposed in REMS with the goal of allowing the graphs to more closely interact with 
-    existing design processes and MDAO frameworks. The approach begins by 
-    constructing the \emph{maximal connectivity graph (MCG)} describing all possible
+    proposed graph syntax is similar to the REMS system proposed by 
+    Alexandrov and Lewis\cite{alexandrov2004}. The approach taken here could be viewed as an extention 
+    of the concepts proposed in REMS with the goals of providing greater insight into design problem 
+    structure.  The approach begins by fconstructing the \emph{maximal connectivity graph (MCG)} describing all possible
     interconnections between the analysis tools proposed by the engineers. Graph operations are then
@@
     The concept of the FPG and the identification of feasible FPGs are the main contributions of the paper.
+    For completeness we include node and edge types for building a PSG, but the details of how a given 
+    problem can be decomposed are beyond the scope of this work.
 
-    The paper is organized as follows. First, we describe the differences between a fundamental problem
-    formulation, which is based only on the system-level optimization problem statement that the
-    engineers desire to solve and on the available analysis tools, and a specific problem formulation, which
-    additionally presumes a specific solution to the problem. Next, we survey the literature in applying
-    graph theoretic and formal language approaches to multidisciplinary design problem formulation. 
-    We then discuss our graph syntax and representation of MDO problems and describe the procedures for 
-    determining the MCG and FPG. Finally, we present an example problem based on an MDO analysis of a 
-    commercial aircraft and discuss additional applications enabled by our approach.
+    The paper is organized as follows. First, we describe the differences between 
+    a fundamental problem formulation based only on the design problem
+    that the engineers desire to solve and on the available analysis tools, 
+    and a specific problem formulation, which additionally 
+    presumes a specific solution to the problem. Next, we survey the literature in applying
+    graph theoretic and formal language approaches to multidisciplinary design 
+    problem formulation. We then discuss our graph syntax and representation of 
+    MDO problems and describe how different structures within a graph map to important
+    features of an MDAO problem. Next we present a procedure for determining possible 
+    FPG from a given MCG. Finally, we present the graph based formulation of commercial 
+    aircraft design problem and discuss additional applications enabled by our approach.
 
 \section{Specific vs Fundamental Problem Formulation }
+    \label{sec:spg_vs_fpg}
 
@@
 
-    Equation \ref{eqn:simple2} differs only slightly from Eq. \ref{eqn:simple}. $A$ is now dependent on the output of $B$, 
-    and $z$ will be iterated on to convergence with $z^t$. Now the problem can be solved by running $B$ first and then $A$.
-    Since the formulations in Eqs. \ref{eqn:simple} and \ref{eqn:simple2} both describe the same problem and neither can be the
-    FPF. They are both specific versions of the same more fundamental description of 
-    the problem that is common between them. We present the FPF as follows:
+    Equation \ref{eqn:simple2} differs only slightly from Eq. \ref{eqn:simple}. 
+    $A$ is now dependent on the output of $B$, and $z$ will be iterated on to 
+    convergence with $z^t$. Now the problem can be solved by running $B$ first 
+    and then $A$.Since the formulations in Eqs. \ref{eqn:simple} and 
+    \ref{eqn:simple2} both describe the same problem then neither can be the
+    FPF. They are both specific versions of a more fundamental description of 
+    the problem that is common between them. We present the fundamental 
+    formulation as follows:
 
@@
 
-    The FPF in Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} and Eq. \ref{eqn:simple2} because it has 
-    two constraints which both must be met. The presence of both of these constraints fully decouples the problem so that 
-    either $A$ or $B$ could be run first or both could be run simultaneously. Alexandrov and Lewis investigated problem 
-    formulation in this fully decoupled way as part of their work on 
-    Reconfigurability in MDO Problem Synthesis (REMS) \cite{alexandrov2004}.
-    In this work they show how removing either constraint and replacing 
-    it with a direct dependence between the two analyses, a process they term ``collapsing'', 
-    you could regain the earlier two problem formulations.
+    The formulation in Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} 
+    and Eq. \ref{eqn:simple2} because it has two constraints which both must be 
+    met. The presence of both of these constraints fully decouples the problem so that 
+    either $A$ or $B$ could be run first or both could be run simultaneously. 
+    Alexandrov and Lewis, in their work with REMS, assert that fully decoupling 
+    the analyses is key to achieving reconfigurability in the problem 
+    formulation\cite{alexandrov2004}.
 
@@
     Rogers et. al developed DeMAID to manipulate a DSM to find an ordering for analysis tools that 
-    reduces the cost of solving highly coupled systems\cite{Rogers1996}. This re-ordering is done through 
+    reduces the cost of solving highly coupled systems\cite{Rogers1996,rogers1996demaid}. This re-ordering is done through 
     row operations on the DSM matrix and yields multiple specific problem 

### 2013-04-09 graph_syntax.tex

@@
 \subsection{Node and Edge Types}
-
-There are four node types:  
+We now define different types of nodes and edges with specific properties to compose graphs used to represent an MDAO problem. 
+There are three node types:  
 \begin{description}
@@
 \item[model:] responsible for mapping inputs to outputs
-\item[expression:] simple algebreic combinations of variables (such as for 
-  constraint of objective functions)
 \item[driver:] control logic capable of managing iteration
@@
 
-In addition to the four node types, there are two edge types:
+In addition to the three node types, there are two edge types:
 
@@
 \end{equation}
-These user-specified limits govern the number of fixed and connection edges that 
-may be directed into a node for a valid graph. Consider a variable node $v$ with
-$\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, that node must have exactly one
-incomming edge or the graph is invalid. Two possiblities exist for voilating 
-the limit conditions: 
+These user-specified limits govern the number of connection edges that may be directed into a variable
+node for a valid graph. Consider a variable node $v$ with
+$\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, $v$ must have exactly one
+incomming explicit edge or the graph is invalid. Two possiblities exist for voilating 
+these limit conditions: 
 \begin{description}
-  \item[hole: ] The number of incomming edges is less than $\txt{deg}_l^-(v)$. 
+  \item[hole: ] The number of incomming edges is less than the lower indegree limit:
     \begin{equation} \txt{deg}^-(v) < \txt{deg}_l^-(v) \label{e:hole} \end{equation}
@@
   \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
-    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \label{e:collision} \end{equation}
+    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \end{equation}
 \end{description} 
+These violations will be managed in order to seek a working problem formulation.
 
 \section{Graph Representation of MDAO Constructs}
-
+%As discussed in Sec.((intro)), the digraph must be able to describe the individual aspects of an MDAO problem.
 \subsection{Analysis Blocks and Connections}
-
-Analysis tools take in a number input variables, then perform some work to calculate 
+Analysis tools take in a number input variables and then perform some work to calculate 
 the values for their respective outputs. In an MDAO graph, this process is 
-represented by an group of nodes and edges called an \emph{analysis block}, 
+represented by a group of nodes and edges called an \emph{analysis block}, 
 shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
@@
 Via the analysis block structure, we can also distinguish between input and ouptut
-variable nodes. Inputs are any variable nodes that have an outgoing edge into a model 
+variable nodes, as shown in Fig.~\ref{f:analysis block}. Inputs are any variable nodes that have an outgoing edge into a model 
 node. Conversely, outputs are any variable nodes that have an incoming edge from a model node. 
 Note that by this definition, inputs and outputs can only exist when variable nodes poses an 
-edge joining them to a model node. In Fig. \ref{f:analysis block} these are labeled out as 
-local inputs and local outputs. The variables are local because they are explicitly tied
+edge joining them to a model node. These are labeled out as 
+local inputs and local outputs. The variables are local because they are explicitely tied
 to the analysis block and have no relevance without it.
@@
 
-\subsection{Design Variables, Objectives, and Constraints}
-
-Design variables, objectives, and constraints all require external decisions 
-to be made about a given system to identify them. Consider a notional problem 
-given by the graph in Fig. \ref{f:designvars} There are three input variables 
-which, could potentially become design variables: $w$, $x$, and $y$. Because the 
-output, $x$, of $A$ is connected to the input, $x$, of $B$ x can't actually be a 
-design variable. The inputs $w$ and $y$ have no incomming edges and could each 
-be selected as design variables.
-
+\subsection{Design Variables}
+Design variables require external decisions 
+to be made about a given system to identify them. Consider 
+a notional problem given by the graph in Fig. \ref{f:designvars} There are three input variables 
+which, could potentially become design variables: $w$, $x$, and $y$. Because the variable $x$ is connected from $A$ to $B$,
+ $x$ can't actually be a design variable. On the other hand, the inputs $w$ and $y$
+have no incomming edges and could each be selected as design variables. 
 \begin{figure}[htb!]
@@
 
-More formally, any input variable that can be defined as a hole by Eqn. \ref{e:hole}
-would be considered a design variable in the graph. When working with a MCG or 
-FPG, only connection edges are allowed, in which case a design variable 
-is simply a variable node with no incomming edges. For a valid PSG 
+Given a MCG with the ${deg}_l^-(v)=1$ for all nodes, there will be a set of holes. 
+Each hole represents a node that could potentially become a design variable. 
+It is up to the designer to examine that variable and decide if it is appropriate to 
+use it as a design variable. If a variable should be a design variable, then 
+set it's ${deg}_l^-(v)=0$. However, some variable holes might not make good design 
+variables. For instance, if your aircraft mission analysis code requires a drag input 
+which ends up being a hole then it's likely that some new analysis tool is needed
+to predict drag before the design process can continue. In that case, you leave
+${deg}_l^-(v)=1$ and find some tool to fill the hole in graph.
+
+So by definition, a design variable is any variable node with ${deg}_l^-(v)=0$, 
+When working with a MCG or FPG, only connection edges are allowed, so a design 
+variable in either graph will be a variable with no edges. For a valid PSG 
 any design variable node would have at least one, but possibly multiple, incomming 
@@
 
-Objective and constraints are also constructs that need to be identified by engineers 
+\subsection{Objectives, and Constraints}
+Objectives and constraints are also constructs that need to be identified by engineers 
 as part of the design process. In the case of objective functions single output values 
 could be selected, but commonly multiple values are assembeled together via simple 
-athrimetic opperations to form the final objective function. For example, the problem from 
+algebraic expression to form a objective function. For example, the problem from 
 Fig. \ref{f:designvars} might have an objective function of $z^2$ or $2y + z$. Constraints are 
 always given in the form of either an inequality of an equality, e.g. 
-$\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer identifies variables 
-or groups of variables and assigns a simple relationship between them.
+$\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer compiles an
+algebraic expression that is a function of existing variable nodes in the graph.
 

### 2013-04-09 graph_syntax.tex
re-wrote objective and constraint section without expression nodes

@@
 $\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer compiles an
-algebraic expression that is a function of existing variable nodes in the graph.
+algebraic expression that can be evaluated to produce an output.
 
-For the proposed syntax, objectives and constraints are represented through the use of 
-expression nodes. These nodes contain the actual mathematical expression used to define the 
-objective or constraint value. Objective or constraint nodes are distinguished
-via the presence of a simple expression (objective) vs an inequality/equality (constraint). 
-Fig. \ref{f:obj-cons} shows the notional problem graph augmented with objective and 
-constraint values.
-
-((DP: I think this definition of objective/constraint is disharmoneous with our other node types. I think it should also have local inputs represented by variables nodes))
+These operations, though very simple, are effectively the same kind of task 
+performed by an analysis block. A constraint or objective function can therfore 
+be represented as another analysis block on the graph, with it's own input and 
+output variable nodes. Although fundamentally no different than an analysis block, 
+for clarity and convience it is useful to distinguish between regular analysis 
+blocks and those that arise from the addition of objectives or constraints to 
+the graph. So we define an expression block, as the collection of variable and model 
+nodes related to a given objective or constraint function. Fig. \ref{f:obj-cons}
+shows a notional problem graph augmented with expression blocks and appropriate 
+connection edges.
 

### 2013-04-09 example_problem.tex
example problem update

@@
     WATE  & engine weight estimation \\
-    FLOPSa & mission performance and \\
-      & engine sizing and weight estimation \\
+    FLOPSa & mission performance, \\
+      & engine sizing, and weight estimation \\
     FLOPSb & mission performance only \\
@@
 FLOPSa denotes FLOPS being executed for both mission performance and engine analysis, while FLOPSb indicates that FLOPS is executed for only mission performance analysis.
+This sort of representation is useful for ``supercodes'' that care capable of being used in different ways and with different sets of inputs and outputs.
 
-In this example, the MCG is formed using a consistant naming convention and then connecting every variables with the same name, though this is not the only way to form an MCG.
+In this example, the MCG is formed using a consistant variable naming convention and then connecting every variable with the same name, though this is not the only way to form an MCG. 
 Table \ref{t:ins and outs} presents the full list of variables in the leftmost column and then indicates whether the variable is an input or an output of each analysis code. 
@@
 \item An analysis block is created for each analysis code using the information in each column of Table \ref{t:ins and outs}.
-Each analysis block is formed by creating variable nodes for each input and directing them into a single model node.
-This model node is then directed into another model node, which is then directed into a variable node for each output.
+Each analysis block is formed by creating variable nodes for each input and directing them into a single model node, which is directed into another model node, which is then directed into variable nodes corresponding to each output.
 A sample analysis block is shown for analysis code FLOPSb in Fig.~\ref{f:FLOPSb analysis block}.
@@
 
-\item Variable nodes are created to represent the objectives for performance and total weight.
+\item Variable ((change)) nodes are created to represent the objectives for performance and total weight.
 
-\item Variable nodes are created to represent the geometry variable and mission variable global inputs.
+\item Variable nodes are created to represent the geometry variable and mission variable as global inputs.
 
 \item Explicit edges are created from each variable node to every other variable node representing variables with the same name.
+The direction is determined by whether the variable node has an edge directed into or out of a model node (local input or local output).
 \end{enumerate}
@@
 \subsection{Obtaining an FPG}
-As presented in Sec.(()), the first step to obtaining an FPG is to detect holes and, to start, we set the lower indegree limit for every node to be one.
+The process of obtaining an FPG from the given MCG, but it is expected to include the user because of required changes to the MCG or make decisions
+
+((I should show a figure of the MCG with edges and nodes colored as to when they were removed))
+((mention python))
+
+As presented in Sec.(()), the first step to obtaining an FPG is to detect holes.
+To start, we set the lower indegree limit for every node to be one to find any potential holes.
 In this case, it turns out that the variable nodes representing the number of engines is a hole for NPSS, FLOPSa, and FLOPSb.
-This implies than an FPG cannot be obtained from the given MCG.
+This implies than an FPG cannot be obtained from the given MCG, as shown in Fig.~\ref{f:MCG nengines}.
+In this figure, the green nodes indicate analysis blocks with a hole, and the cyan nodes indicate analysis blocks that became unused after the invalid analyses were removed ((connections were removed, color edges too)).
+\begin{figure}[htb!]
+  \begin{center}
+    \includegraphics[width=.6\textwidth]{images/MCG_nengines}
+  \end{center}
+  \caption{Simple notional problem with two potential design variables \label{f:design vars}}
+\label{f:MCG nengines}
+\end{figure}
+
 Therefore, we may now create another variable node to represent the number of engines as a global input and also create explicit edges to supply this new input.
@@
 The first example of this is to obtain an FPG with the fewest possible number of cycles, which are found using the implementation of Johnson's algorithm \cite{Johnson1975} in the Python package NetworkX.
-The resulting FPG is shown in Fig.~\ref{f:FPG fewest cycles} and the only cycle is shown in yellow.
+The FPG with the fewest cycles is shown in Fig.~\ref{f:FPG fewest cycles}, which indicates one cycle shown in yellow.
 A similar alternative would be to minimize the number of analysis blocks involved in cycles, counting multiplicity.
-While resulting graph has only one cycle, it may be more desirable to include the higher fidelity codes, such as NPSS.
 \begin{figure}[htb!]
@@
 
+Alternatively, it may be desirable to resolving collisions by including the higher fidelity codes, such as NPSS in this case.
 One method to ensure that the edges directed from certain analysis blocks are the ones chosen to resolve the conflicts is to use a ranking system.
@@
 Step two in Sec.(()) (the resolution of collisions) is now conducted by selecting the edges directed from the analysis blocks with the highest rank for each collision.
-The resulting FPG is shown in Fig.~\ref{f:FPG highest rank} and this graph has four cycles, which are enumerated in Table \ref{t:cycles}.
+The resulting FPG is shown in Fig.~\ref{f:FPG highest rank}, and this graph has four cycles, which are enumerated in Table \ref{t:cycles}.
 \begin{figure}[htb!]
@@
 
-Finally, now consider the case where it is desired to have the inviscid drag input into FLOPSb be a multi--fidelity input.
-This is done simply by setting the upper indegree limit for that node to two and repreating the (automated) process.
+Finally, now consider the case where it is desired to have the inviscid drag input into FLOPSb be a multi--fidelity input, meaning that it is sought to use multiple analysis codes to calculate the same variable.
+This is done simply by setting the upper indegree limit for this node as two and then repreating the (automated) process.
 In this case, both VORLAX and PMARC are retained, resulting in an FPG that omits only FLOPSa.

### 2013-04-09 graph_syntax.tex
another round of edits

@@
 The ultimate goal of the new graph--based syntax presented here is to be able to 
-fully describe the general structure of an MDAO problem indpendently of any solution information, 
+fully describe the general structure of an MDAO problem independently of any solution information, 
 while still being able to accomodate the more specific case when a solution 
@@
 a single formal problem formulation is selected specifying design variables, constraints, objectives, 
-analysis tools, etc. Lastly some specifc procedure for solving the problem is selected, for example 
+analysis tools, etc. Lastly some specific procedure for solving the problem is selected, for example 
 picking an MDAO optimization architecture. Using the proposed graph syntax, each of these phases 
@@
     \item Maximal Connectivity Graph (MCG)
-    \item Fundamental Problem  (FPG)
+    \item Fundamental Problem Graph (FPG)
     \item Problem Solution Graph (SPG)
@@
 analysis codes being considered and all possible connections between them also present. The second graph 
-is the \emph{fundamental problem formulation graph}, which is the smallest possible graph 
+is the \emph{fundamental problem graph}, which is the smallest possible graph 
 that still fully defines a given problem formulation. Finally, a \emph{specific problem formulation} 
@@
 \section{Syntax Definition}
-In the rest of this section we present the necessary graph theory fundamentals to 
-construct any of the different graphs discussed above. 
+In this section we present the necessary graph theory fundamentals to 
+construct the graphs discussed above. 
 The notation used in this work is adapted from Diestel \cite{Diestel2010}. 
@@
 and function, and only one edge type\cite{alexandrov2004}. We've chosen to rename the ``function'' node 
-type to ``model'' to be more consistent with modern MDAO frameworks. This work 
+type to ``model'' to be more consistent with modern MDAO frameworks. The present work 
 adds two more node types and one more edge type to the syntax to allow descriptive
@@
 \label{s:indegree-outdegree}
-The \emph{in-degree} of a node is the number of connection edges directed in and 
-is denoted as $\txt{deg}^-(v)$, and the \emph{out-degree} 
+The \emph{in--degree} of a node is the number of connection edges directed in and 
+is denoted as $\txt{deg}^-(v)$, and the \emph{out--degree} 
 is the number of connection edges directed out and it is denoted as $\txt{deg}^+(v)$.
@@
 and a gradient based one is applied second, design variables would have incomming 
-implicit edges from both optimizers but this would not affect the in-degree of those
+implicit edges from both optimizers but this would not affect the in--degree of those
 variable nodes. 
+((DP:I started to remove any reference to the edge type in these definitions but then saw that you needed it.
+The reason I did that was because in--degree and out--degree are basic graph theory terms.
+What if the reference to implicit edges is handled with the upper in--degree limit, i.e.
+after the FPG is obtained the upper indegree limit for the design variable nodes is relaxed (increased) so that it isn't a problem what is or isn't counted?))
 
-We also define the \emph{upper in-degree limit}
+((DP:the definitions were in/out degree were not in my book, I had to get them from wikipedia, where they are spelled indegree and outdegree))
+
+We also define the \emph{upper in--degree limit}
 
@@
 \end{equation} 
-and the \emph{lower in-degree limit} as
+and the \emph{lower in--degree limit} as
 \begin{equation}
@@
     \begin{equation} \txt{deg}^-(v) < \txt{deg}_l^-(v) \label{e:hole} \end{equation}
+	This represents a lack of information being supplied to the node.
 
@@
     \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \end{equation}
+	This represents an overload of information such that it is unclear which source to use.
 \end{description} 
-These violations will be managed in order to seek a working problem formulation.
+An algorithm for detecting and managing these violations if proposed in Sec.(()).
 
@@
 This edge represents the necessary calculations to map given inputs 
-to proper output values. If constructing a weighted graph, the weight calculation edge 
-provides the opportunity to computational cost. However, it is also allowed to skip 
+to proper output values. This edge can be given a weight to represent the computational cost of the analysis block. However, it is also allowed to skip 
 this edge and have both incomming and outgoing edges to variables through a single model 
@@
 
-Via the analysis block structure, we can also distinguish between input and ouptut
-variable nodes, as shown in Fig.~\ref{f:analysis block}. Inputs are any variable nodes that have an outgoing edge into a model 
+Variable nodes in an analysis block can be distinguished as either an input or an output by the manner in which they are connected.
+As shown in Fig.~\ref{f:analysis block}, inputs are any variable nodes that have an outgoing edge into a model
 node. Conversely, outputs are any variable nodes that have an incoming edge from a model node. 
 Note that by this definition, inputs and outputs can only exist when variable nodes poses an 
-edge joining them to a model node. These are labeled out as 
+edge joining them to a model node. 
+((DP:I'm not sure what the sentences before or after this comment mean))
+These are labeled out as 
 local inputs and local outputs. The variables are local because they are explicitely tied
@@
 input of another analysis block, a new connection edge is added connecting the two 
-variable nodes involved in the exchange. These connection edges are free, unlike the edges 
-within an analysis block, they could be added or removed depending on the needs
+variable nodes involved in the exchange. These connection edges are free---unlike the edges 
+within an analysis block---which means they could be added or removed depending on the needs
 of a given design problem. In other words, free connection edges are created by 
@@
 \subsection{Design Variables}
+((DP:I'm thinking we might also need the upper indegree limit to be zero as well, because setting only the lower indegree limit makes the node a design variable if it didn't have any incoming edges.
+It's seems possible that in the MCG a node has incoming edges but the user wants it to be a design variable.
+I can make the change if you agree.))
 Design variables require external decisions 
 to be made about a given system to identify them. Consider 
-a notional problem given by the graph in Fig. \ref{f:designvars} There are three input variables 
-which, could potentially become design variables: $w$, $x$, and $y$. Because the variable $x$ is connected from $A$ to $B$,
+a notional problem given by the graph in Fig.~\ref{f:designvars}. There are three input variables 
+which could potentially become design variables: $w$, $x$, and $y$. Because the variable $x$ is connected from $A$ to $B$,
  $x$ can't actually be a design variable. On the other hand, the inputs $w$ and $y$
@@
 
-Given a MCG with the ${deg}_l^-(v)=1$ for all nodes, there will be a set of holes. 
+Given a MCG with the $\txt{deg}_l^-(v)=1$ for all nodes, there will likely be a set of holes preventing a valid FPG from being obtained.
 Each hole represents a node that could potentially become a design variable. 
-It is up to the designer to examine that variable and decide if it is appropriate to 
-use it as a design variable. If a variable should be a design variable, then 
-set it's ${deg}_l^-(v)=0$. However, some variable holes might not make good design 
-variables. For instance, if your aircraft mission analysis code requires a drag input 
-which ends up being a hole then it's likely that some new analysis tool is needed
-to predict drag before the design process can continue. In that case, you leave
-${deg}_l^-(v)=1$ and find some tool to fill the hole in graph.
+It is up to the designer to examine each and decide if it is appropriate to 
+consider it a design variable, in which case the designer shall set the lower indegree limit to be zero.
+However, some variable holes may not actually be suitable as a design variable.
+For instance, if an aircraft mission analysis code requires a drag input
+which ends up being a hole then it's likely best to leave
+$\txt{deg}_l^-(v)=1$ and find a new analysis code to fill the hole in graph.
 
-So by definition, a design variable is any variable node with ${deg}_l^-(v)=0$, 
-When working with a MCG or FPG, only connection edges are allowed, so a design 
-variable in either graph will be a variable with no edges. For a valid PSG 
+A design variable is defined is any variable node with $\txt{deg}_l^-(v)=0$. ((DP: it seems like this should be have been at the beginning of this subsection.)) 
+When working with a MCG or FPG, connection edges are the only free edges, so a design 
+variable in either graph will be a variable node ((only variable nodes that are local inputs)) with no incoming edges. For a valid PSG 
 any design variable node would have at least one, but possibly multiple, incomming 
@@
 \subsection{Objectives, and Constraints}
-Objectives and constraints are also constructs that need to be identified by engineers 
+Objectives and constraints are also MDAO constructs that need to be identified by engineers 
 as part of the design process. In the case of objective functions single output values 
 could be selected, but commonly multiple values are assembeled together via simple 
-algebraic expression to form a objective function. For example, the problem from 
-Fig. \ref{f:designvars} might have an objective function of $z^2$ or $2y + z$. Constraints are 
+algebraic expression to form a objective function. For example, the problem depicted in 
+Fig.~\ref{f:designvars} could have an objective function of $z^2$ or $2y + z$. Constraints are 
 always given in the form of either an inequality of an equality, e.g. 
@@
 blocks and those that arise from the addition of objectives or constraints to 
-the graph. So we define an expression block, as the collection of variable and model 
-nodes related to a given objective or constraint function. Fig. \ref{f:obj-cons}
+the graph. Therefore, we define an \emph{expression block} as the collection of variable and model 
+nodes related to a given objective or constraint function. Fig.~\ref{f:obj-cons}
 shows a notional problem graph augmented with expression blocks and appropriate 
@@
 
-  A simple definition of a global node is any node that involves information 
+  A simple definition of a global node is any node that directly involves information 
   from more than one analysis block. Likewise, a local node is any node 
@@
   Coupling exists in a design problem when a set of two or more analyses each depend on the
-  others outputs. Consider the notional problem from section \ref{sec:spg_vs_fpg} where two 
+  others outputs. Consider the notional problem from Sec.~\ref{sec:spg_vs_fpg} where two 
   analyses, $A$ and $B$ are coupled through their respective variables. The graph of this 
@@
   a cycle which represent the coupling between them. This cycle means that blocks $A$ and $B$ need
-  to be iterated on untill they converge.
+  to be iterated on until they converge.
 
@@
 
-  It is important to note that the cycle in graph in Fig. \ref{f:coupling} does not contain 
+  It is important to note that the cycle in graph in Fig.~\ref{f:coupling} does not contain 
   any driver nodes. That means that no solver, optimizer, or other iterative construct is 
-  invovled in the coupling. This absence is intentional, as the coupling is an inherant part 
-  of the problem. This corresponds directly To Eqn. \ref{eqn:simple_fpf}, where 
-  each equality constraint is one edge in the graph. Transitionining to either 
-  Eqn. \ref{eqn:simple} or Eqn. \ref{eqn:simple2} means selecting one of the nodes as the starting point 
-  for the iteration, indicating the associated analysis block should be executed first.
+  invovled in the coupling. This absence is intentional, as the coupling is an inherent part 
+  of the problem. This corresponds directly to Eqn.~\ref{eqn:simple_fpf}, where 
+  each equality constraint is one edge in the graph. Transitioning to either 
+  Eqn.~\ref{eqn:simple} or Eqn.~\ref{eqn:simple2} means selecting one of the nodes as the starting point 
+  for the iteration and indicating which associated analysis block should be executed first.
 
-  Larger problems can contain more complex cycles in thier FPG, indicating more 
+  Larger problems can contain more complex cycles in their FPG, indicating more 
   complex coupling between analyses. A cycle can contain more than 
@@
   blocks are involved in multiple different coupling cycles. All of these situations
-  arise naturally as the size of problems grows, and managing all of this coupling 
-  becomes difficult. From the perspective of the graph, the problem reduces down to 
-  a ordering challenge. For each cycle in the graph, you need to pick a analysis block
-  to start the iteration from. Rogers DEMAID tool uses a genetic algorithm to find an ordering 
+  arise naturally as the size of problems grows, and managing this coupling may
+  become difficult. From the perspective of the graph, the problem reduces down to 
+  an ordering challenge. For each cycle in the graph, an analysis block needs to be selected 
+  to start the iteration from. Rogers' DEMAID tool uses a genetic algorithm to find an ordering 
   that minimizes the overall coupling of the system by separating independent 
   cycles in the graph \cite{rogers1996,rogers1996demaid}.
+In the present work, Sec.(()) will suggest that an FPG may be obtained with respect to the number of cycles or the length of these cycles, etc.
 
-\subsection{Multi-fidelity Problems}
+\subsection{Multi--fidelity Problems}
 
-  A multi-fidelity problem can be characterized by two or more different analyses 
-  each calculating the same data. Mapping this type of multi-fidelity situation 
-  onto a graph will yeild two analysis blocks, each with thier own output 
+  A multi--fidelity problem can be characterized by two or more different analyses 
+  each calculating the same data. Mapping this type of multi--fidelity situation 
+  onto a graph will yield two analysis blocks, each with their own output 
   variable nodes, having outgoing edges to a third analysis block that needs the 
@@
   that exhibits this structure, where the key feature is that variable node $y$ in 
-  analysis block $C$ has two incomming edges.
+  analysis block $C$ has two incoming edges.
 
@@
 
-  When an input variable node has more than one incomming edge, designers must make
+  When an input variable node has more than one incoming edge, designers must make
   some kind of decision about how to manage the flow of information in their problem. 
-  In Eqn. \ref{e:collision} we defined a collision as the situation where
-  there are more incomming connection edges than the ${deg}_u^-(y)$. So desginers can
-  indicate how to handle conflicting edges in a problem graph by setting the ${deg}_u^-(y)$
+  In Eqn.~\ref{e:collision} we defined a collision as the situation where
+  there are more incoming connection edges than the upper indegree limit. So desginers can
+  indicate how to handle conflicting edges in a problem graph by setting the upper indegree limit
   for a given node. In more concrete terms, setting this limit equates to defining 
-  the part of the graph in question as single or multi-fidelity.
+  the part of the graph in question as single or multi--fidelity.
 
-  When ${deg}_u^-(y)=1$, then any conflicting edges in the graph will cause a collision
+  When $\txt{deg}_u^-(y)=1$, then any conflicting edges in the graph will cause a collision
   and a choice between the colliding analysis blocks must be made. This results in a 
-  single fidelity design problem. On the other hand if ${deg}_u^-(y)=2$, then two 
+  single fidelity design problem. On the other hand if $\txt{deg}_u^-(y)=2$, then two 
   different analysis blocks can co-exist with a given graph without causing a 
@@
 
-\item The third and final step is to remove any unused nodes and edges. The indexing set for only the analysis blocks with at least one used local output may be constructed as
+\item The third and final step is to remove any unused nodes and edges. The indexing set for only the analysis blocks with at least one used local output may be constructed as ((pruning))
 \begin{equation}
@@
 
-%Let $F_2 = (V_M,C_{F,2})$ be an undirected graph, which is to say to disregard the directions associated with $C_{F,2}$.
-
-%Fixed edges are fixed only with respect to the analysis block in which they reside, it is okay to remove them along with the entire analysis block.
-
 \end{enumerate}

### 2013-04-10 graph_syntax.tex
cross referencing

@@
 \section{Requirements for New Graph Syntax}
+	\label{s:requirements}
 	The ultimate goal of the new graph--based syntax presented here is to be able to 
@@
 
-%the relevant MDAO constructs from the given grah primatives. Next we show how to construct 
-%the MCG, FPG, and SPF graphs from those constructs and demonstrate methods to transition between 
-%them.
-
 \section{Syntax Definition}
+	\label{s:syntax definition}
 	In this section we present the necessary graph theory fundamentals to 
@@
 \subsection{Node and Edge Types}
+	\label{ss:types}
 	We now define different types of nodes and edges with specific properties to compose graphs used to represent an MDAO problem. 
@@
 \subsection{Rules for Nodes and Edges}
+	\label{ss:rules}
 	There are specific rules for the usage of these nodes and edges.
@@
 	\end{description} 
-An algorithm for detecting and managing these violations if proposed in Sec.(()).
+	An algorithm for detecting and managing these violations if proposed in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}.
 
 \section{Graph Representation of MDAO Constructs}
-%As discussed in Sec.((intro)), the digraph must be able to describe the individual aspects of an MDAO problem.
+	\label{s:graph representation}
+	%As discussed in Sec.~\ref{s:intro}, the digraph must be able to describe the individual aspects of an MDAO problem.
 \subsection{Analysis Blocks and Connections}
+	\label{ss:analysis blocks and connections}
 	Analysis tools take in a number input variables and then perform some work to calculate 
@@
 	the edges within them are fixed, they become fixed sub-graphs within a graph.
-
 	\begin{figure}[htb!]
@@
 \subsection{Design Variables}
+	\label{ss:design variables}
 	((DP:I'm thinking we might also need the upper indegree limit to be zero as well, because setting only the lower indegree limit makes the node a design variable if it didn't have any incoming edges.
@@
 \subsection{Objectives, and Constraints}
+	\label{ss:objectives and constraints}
 	Objectives and constraints are also MDAO constructs that need to be identified by engineers 
@@
 	connection edges. 
-
 	\begin{figure}[htb!]
@@
 \subsection{Local vs Global Nodes}
-
+	\label{ss:local vs global}
 	A simple definition of a global node is any node that directly involves information 
@@
 	on block $B$. $g_2$ has incomming edges from both $A$ and $B$ analysis blocks, so it would be a global constraint. 
-
 	\begin{figure}
@@
 \subsection{Coupling Between Analyses}
-
+	\label{ss:coupling}
 	Coupling exists in a design problem when a set of two or more analyses each depend on the
-  others outputs. Consider the notional problem from Sec.~\ref{sec:spg_vs_fpg} where two 
+	others outputs. Consider the notional problem from Sec.~\ref{s:specific vs fundamental} where two 
 	analyses, $A$ and $B$ are coupled through their respective variables. The graph of this 
-  is shown in Fig. \ref{f:coupling}. In the graph $A.z$ has an outgoing connection 
+	is shown in Fig.~\ref{f:coupling}. In the graph $A.z$ has an outgoing connection 
 	edge to $B.z$, indicating that there is a connection between the two. 
@@
 	to be iterated on until they converge. 
-
 	\begin{figure}
@@
 	cycles in the graph \cite{rogers1996,rogers1996demaid}.
-In the present work, Sec.(()) will suggest that an FPG may be obtained with respect to the number of cycles or the length of these cycles, etc.
+In the present work, Sec.~\ref{s:example problem}.\ref{ss:obtaining an FPG} will suggest that an FPG may be obtained with respect to the number of cycles or the length of these cycles, etc.
 
 \subsection{Multi--fidelity Problems}
-
+	\label{ss:multi-fideliy problems}
 	A multi--fidelity problem can be characterized by two or more different analyses 
@@
 	analysis block $C$ has two incoming edges. 
-
 	\begin{figure}
@@
 \section{Building MCG and FPG}
-
+	\label{s:building graphs}
 \subsection{Maximal Connectivity Graph}
+	\label{ss:MCG}
 	To construct the maximal connectivity graph, we assume that a set of codes, global inputs, and objectives and constraints (collectively called global outputs) are provided.
@@
 
-\subsection{Fundamental Problem Formulation Graph}
-We now define the fundamental problem formulation graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions, which are explained subsequently,
+\subsection{Fundamental Problem Graph}
+	\label{ss:FPG}
+	We now define the fundamental problem graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions, which are explained subsequently,
 	\begin{enumerate}
@@
 \subsection{Obtaining the Fundamental Problem Graph}
+	\label{ss:obtaining FPG}
 	In general, there may be multiple different graphs that satisfy the FPG conditions, though there may be none at all.

### 2013-04-10 intro.tex
cross referencing

--- 2013-04-02 intro.tex
@@
 \section{Introduction}
-
+    \label{s:intro}
     The number of analysis tools required in multidisciplinary design optimization (MDO) studies is growing
@@
 \section{Specific vs Fundamental Problem Formulation }
-
+	\label{s:specific vs fundamental}
     The Fundamental Problem Formulation (FPF) for any given problem will be constant regardless 
@@
 \section{Existing Graph-Based Syntax}
-
+	\label{s:existing syntax}
     As shown above, the mathematical language for specifying problem formulations is very general and can be used both for 

### 2013-04-13 graph_syntax.tex
update sellar problem

--- 2013-04-09 graph_syntax.tex
@@
   \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
-    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \end{equation}
+    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \label{e:collision} \end{equation}
 \end{description} 
@@
 \section{Graph Representation of MDAO Constructs}
+In this section we will make use of the Sellar problem, described in 
+Eqn. \ref{eqn:simple_fpf}. The graph representation of the Sellar problem, 
+using the new stynax, is presented in Fig. \ref{f:sellar_graph_full}. Throughout
+the rest of this section, different sub-graphs or slight modifications of 
+the full graph will be used to demonstrate the important MDAO constructs within
+this graph syntax.
+
+\begin{figure}[htb!]
+    \begin{center}
+    \includegraphics[width=4in]{images/sellar_graph_full}
+    \end{center}
+    \vspace{-10pt}
+\caption{Graph of the Sellar problem formulation}
+\label{f:sellar_graph_full}
+\end{figure}
+
 %As discussed in Sec.((intro)), the digraph must be able to describe the individual aspects of an MDAO problem.
@@
 the values for their respective outputs. In an MDAO graph, this process is 
-represented by a group of nodes and edges called an \emph{analysis block}, 
-shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
+represented by a group of nodes and edges called an \emph{analysis block}. 
+The Sellar problem has two analyses, $D_1$ and $D_2$, so the graph will have 
+two analysis blocks in it. Fig. \ref{f:analysis block} shows the analysis block for $D_2$
+with some important annotations. Within an analysis block each variable 
 node represents a single input or output and is connected 
@@
 to proper output values. If constructing a weighted graph, the weight calculation edge 
-provides the opportunity to computational cost. However, it is also allowed to skip 
+provides the opportunity to represent computational cost. However, it is also fine to skip 
 this edge and have both incomming and outgoing edges to variables through a single model 
@@
     \vspace{-10pt}
-\caption{Example analysis block. The each node type and edge type is labeled in italics and annotated parenthetically.}
+\caption{Sellar problem $D_2$ analysis block The each node type and edge type is labeled in italics and annotated parenthetically.}
 \label{f:analysis block}
@@
 of a given design problem. In other words, free connection edges are created by 
-engineers linking the output of one tool to the input of another one. Note that 
+linking the output of one tool to the input of another one. Note that 
 it is allowed for a single output to have outgoing connection edges to multiple 
@@
 \subsection{Design Variables}
-Design variables require external decisions 
-to be made about a given system to identify them. Consider 
-a notional problem given by the graph in Fig. \ref{f:designvars} There are three input variables 
-which, could potentially become design variables: $w$, $x$, and $y$. Because the variable $x$ is connected from $A$ to $B$,
- $x$ can't actually be a design variable. On the other hand, the inputs $w$ and $y$
-have no incomming edges and could each be selected as design variables. 
+For the Sellar problem, there are three design variables, $x_1^*$, $z_1^*$, and $z_2^*$. In Fig. 
+\ref{f:sellar_graph_full}, these are the only three holes (nodes without out 
+incomming edges) in the graph. Based on this example it's tempting to simply 
+call all holes in the graph design variables. However, in the general case this is 
+not completely correct. For any given problem, the set of holes only identifies 
+all the potential design variables. However, some variable holes might not make good design 
+variables. For instance, if your aircraft mission analysis code requires a drag input 
+which ends up being a hole then it's likely that some new analysis tool is needed
+to predict drag before the design process can continue.
+
 \begin{figure}[htb!]
   \begin{center}
+  [NOTE: Re-work this one to match $D_2$ from Sellar]
     \includegraphics[width=.6\textwidth]{images/design_vars_graph}
@@
 
-Given a MCG with the ${deg}_l^-(v)=1$ for all nodes, there will be a set of holes. 
-Each hole represents a node that could potentially become a design variable. 
-It is up to the designer to examine that variable and decide if it is appropriate to 
-use it as a design variable. If a variable should be a design variable, then 
-set it's ${deg}_l^-(v)=0$. However, some variable holes might not make good design 
-variables. For instance, if your aircraft mission analysis code requires a drag input 
-which ends up being a hole then it's likely that some new analysis tool is needed
-to predict drag before the design process can continue. In that case, you leave
+More formally stated, given a MCG with the ${deg}_l^-(v)=1$ for all nodes, 
+there will be a set of holes. Each hole represents a node that could potentially 
+become a design variable. It is up to the designer to examine that variable and 
+decide if it is appropriate to use it as a design variable. If a variable should 
+be a design variable, then set it's ${deg}_l^-(v)=0$. Otherwise you leave 
 ${deg}_l^-(v)=1$ and find some tool to fill the hole in graph.
 
-So by definition, a design variable is any variable node with ${deg}_l^-(v)=0$, 
-When working with a MCG or FPG, only connection edges are allowed, so a design 
-variable in either graph will be a variable with no edges. For a valid PSG 
+So a design variable is defined as any variable node with ${deg}_l^-(v)=0$. 
+This, by definition, means that design variables are not holes in the graph, since 
+they have not violated their lower in-degree limit. Regardless, 
+when working with a MCG or FPG, only connection edges are allowed, so a design 
+variable in either graph will be a variable with no incomming edges edges. For a valid PSG 
 any design variable node would have at least one, but possibly multiple, incomming 
@@
 \subsection{Objectives, and Constraints}
-Objectives and constraints are also constructs that need to be identified by engineers 
-as part of the design process. In the case of objective functions single output values 
-could be selected, but commonly multiple values are assembeled together via simple 
-algebraic expression to form a objective function. For example, the problem from 
-Fig. \ref{f:designvars} might have an objective function of $z^2$ or $2y + z$. Constraints are 
-always given in the form of either an inequality of an equality, e.g. 
-$\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer compiles an
-algebraic expression that can be evaluated to produce an output.
+Like design variables, objectives and constraints are also constructs that need 
+to be identified when setting up a given design problem. In the case of 
+objective functions single output values could be selected, but commonly multiple 
+values are assembeled together via simple algebraic expressions to form a composite 
+objective function. (e.g. The objective function from Eqn. \ref{eqn:simple_fpf} is
+$x_1^2+z_2+y_1+e^{-y_2}$). Constraints are always given in the 
+form of either an inequality or an equality. (e.g. a constraint from Eqn. 
+\ref{eqn:simple_fpf} is $\frac{y_1}{3.16}-1\geq0$.) By convention, 
+constraints are usually given such that some expression will be less than or equal 
+to zero, so any positive value would violate the constraint. For both objective 
+and constraints, these simple expressions take some inputs and map them to an
+output value of significance to the overall design problem.
 
-These operations, though very simple, are effectively the same kind of task 
+These operations, though very simple, are effectively the same task 
 performed by an analysis block. A constraint or objective function can therfore 
-be represented as another analysis block on the graph, with it's own input and 
+be represented as another analysis block in the graph, with it's own input and 
 output variable nodes. Although fundamentally no different than an analysis block, 
@@
 nodes related to a given objective or constraint function. Fig. \ref{f:obj-cons}
-shows a notional problem graph augmented with expression blocks and appropriate 
-connection edges.
+shows the objective and constraint blocks from the Sellar problem in isolation 
+from the rest of the problem. Note how each block has it's own input variables, 
+which in the full graph would be linked via connection edges. Treating objectives 
+and constraints in this way introduces extra variables to the problem formulation,
+but provides an important benefit for describing local and global properties of 
+MDAO problems.
 
@@
   \begin{center}
+    [NOTE: Show just the G1, G2, and F blocks from Sellar Graph]
     \includegraphics[width=.6\textwidth]{images/obj_const_graph}
@@
 
-\subsection{Local vs Global Nodes}
+\subsection{Local vs Global Properties}
 
@@
   nodes with fixed edges directed to model nodes (as part of an analysis block) are inherently local 
-  and similarly model nodes are also local. Expression and driver nodes
-  are not inherently local or global though. For these, their locality is determined 
-  by the incomming and outgoing edges they possess. Consider the graph in Fig. \ref{f:local_global}, 
-  where the expression node $g_1$ represents some kind of constraint for the optimziation problem. 
-  Since $g_1$ only has edges comming from analysis block $B$, it would be considered a local constraint 
-  on block $B$. $g_2$ has incomming edges from both $A$ and $B$ analysis blocks, so it would be a global constraint.
+  and similarly model nodes are also local. Global variables are a part of many 
+  MDAO problems and must be representable in the graph. Follwoing our simple definition
+  any variable node that had multiple incomming or outgoing edges, would be global. 
+  In Fig. \ref{f:sellar_graph_full} the variable nodes for $x_1^*$, $z_1^*$, and $z_2^*$ 
+  have multiple outgoing edges and would be considered global.
+
+  Note that the variable $x_1$ from the Sellar problem is usually refered to as 
+  a local variable, which is contradicted by the given graph. This discrepancy 
+  arises from the explicit treatment of objectives and constraints as separate 
+  expression blocks. Normally in the Sellar problem $x_1$ is considered a local 
+  variable because it only directly affects analysis block $D_1$. By forcing
+  the expansion of $F$ into an expression block with its own inputs, then $F$ 
+  gets it's own copy of the $x_1$ variable. This necessitates the creation 
+  of the global $x_1^*$ node in the graph with two outgoing edges to link the 
+  two $x_1$ inputs nodes in the different blocks. Interestingly, using this 
+  definition, where $x_1^*$ is a global variable fits nicely within 
+  the structure of Braun's Colaborative Optimization (CO) archtiecture \cite{braun_thesis}. 
+  The rules of setting up a problem with CO require that each discipline be 
+  given an independent local copy of all global variables, while any local 
+  variables are retained uniquely within their respective disciplines. But 
+  when a local design variable shows up explicitely in the objective function 
+  or global constraint then a global variable is created, and the discipline again 
+  gets an independent local copy of it. Given our definition of global vs local
+  variable nodes, no such exception for local variables is necessary. When you 
+  include a given variable in an expression node, it becomes global by definition 
+  and should be treated as such in CO.
+
+  Analysis and expression blocks can also have a locality defined for the set of 
+  nodes and edges within the block. Their locality is determined by the 
+  incomming and outgoing edges from the block as a whole. Fig. \ref{f:sellar_graph_full} has expression blocks for each constraint, $G_1$ and $G_2$. Both constraints are 
+  local to their respective analysis blocks. But the expression block for $F$ 
+  has incomming edges from $D_1$ and $D_2$, so it would considered global.
 
@@
 
-  Although we've already established that variable nodes in an analysis block are local to that 
-  block, global variables are a part of many MDAO problems and must be representable. A global variable 
-  would be any variable node with multiple outgoing connection edges. 
-  For example, in Fig. \ref{f:local_global} the global node $w^*$ has outgoing edges to 
-  the $w$ variable nodes feeding both $A$ and $B$ analysis blocks.
-
-  While this simple definition is valid, it glosses over a more subtle aspect of 
-  some MDAO problems. Fig. \ref{f:local_global}, has three analysis 
-  blocks A, B, and C. Above we identified node $g_2$ as global,  but
-  it has no incomming edges from block $C$. So you could say that $g_2$ is 
-  local to the the set of blocks $A,B$. In this sense, locality 
-  becomes a relative term. A given node can be local to a specific sub-graph of a large
-  problem. This subtle structure can become significant when determining how to 
-  effectively partition an MDAO problem into sub-problems 
-  \cite{krishnamachari1997optimal,allison2009optimal,michelena1997hypergraph}.
+  Simply stating that any node or block with incomming or outgoing edges
+  from more than one other node or block is global works glosses over a more 
+  subtle aspect of some MDAO problems. The locality of any node in a graph is 
+  a relative property. For instance, a single node might have outgoing edges 
+  to two separate analysis blocks, $A$ and $B$, but none to a third, $C$. Then 
+  you could say that this node was global with respect to $A$ and $B$, but local 
+  to the group $A,B$ with respect to $C$. This situation produces a natural 
+  hierarchy in a graph, where you can form increasingly smaller groups as you 
+  segment problem into finer localities. When solving a problem using 
+  a heirarchical decomposition approach, there are various techniques for 
+  partitioning that are effective in different scenarios 
+  \cite{krishnamachari1997optimal,Perez2004,sobieszczanski1997,allison2009optimal,michelena1997hypergraph}.
 
@@
   is shown in Fig. \ref{f:coupling}. In the graph $A.z$ has an outgoing connection 
-  edge to $B.z$, indicating that there is a connection between the two. 
-  Similarly a connection exists between $B.y$ and $A.y$. These two connections form 
-  a cycle which represent the coupling between them. This cycle means that blocks $A$ and $B$ need
-  to be iterated on untill they converge.
+  edge to $B.z$, which refers to the coupling constraint $g1$ from Eqn. \ref{eqn:simple_fpf}. Similarly a connection exists between $B.y$ and $A.y$ refering to the coupling constraint
+  These two connections form a cycle which represent the coupling between them. This cycle means that blocks 
+  $A$ and $B$ need to be iterated on until the two equality constraints from Eqn. \ref{eqn:simple_fpf}.
 

### 2013-04-13 intro.tex
update sellar problem

--- 2013-04-04 intro.tex
@@
     of which MDAO framework, optimization algorithm, iterative solver, or solution strategy
-    is used to solve the problem. For example, consider the following notional problem:
+    is used to solve the problem. However, it is possible to formulate a problem
+    with an assumed execution ordering that implies a solution strategy. 
+    For example, consider the Sellar test problem \cite{AIAA:sellar}:
 
     \begin{align}
-        given & \ \ A: {x,y} \rightarrow {m,z} \notag
-        \\      & \ \ B: {x,z} \rightarrow {y^t} \notag
-        \\min. &\ \ f(m) \notag
-        \\w.r.t. & \ \ x,y,z \notag
-        \\s.t. & \ \ g(y^t,y) = 0
+        given & \ \ y_1 = D_1(x_1,y_2,z_1,z_2) \notag
+        \\      & \ \ y_2^t = D_2(y_1,z_1,z_2) \notag
+        \\min. &\ \ F(x_1,y_1,y_2,z_2) \notag
+        \\w.r.t. & \ \ x_1,y_1,y_2,z_1,z_2 \notag
+        \\s.t. & \ \ G(y_2,y_2^t) = 0
         \label{eqn:simple}
@@
 
-    $A$ and $B$ represent analysis tools, and $f$ and $g$ are the objective and constraint functions respectively. 
-    Equation \ref{eqn:simple} makes an inherent assumption about the solution strategy for the problem. 
-    Analysis $A$ outputs $z$, which is an input to $B$. Hence, $A$ should be run before $B$ with 
-    with $y$ being iterated on to convergence with $y^t$. However, a slightly different formulation is 
-    equally valid and still represents the exact same problem:
+    $D_1$ and $D_2$ represent analysis tools, and $F$ and $G$ are the objective 
+    and constraint functions respectively. Equation \ref{eqn:simple} makes an 
+    inherent assumption about the solution strategy for the problem. Analysis 
+    $D_1$ outputs $y_1$, which is an input to $D_2$. Hence, $D_1$ should be run 
+    before $D_2$ with with $y_2$ being iterated on to convergence with $y_2^t$. 
+    However, a slightly different formulation is equally valid and still 
+    represents the exact same problem:
 
     \begin{align}
-        given & \ \ B: {x,z} \rightarrow {y} \notag
-        \\      & \ \ A: {x,y} \rightarrow {m,z^t} \notag
-        \\min. &\ \ f(m) \notag
-        \\w.r.t. & \ \ x,y,z \notag
-        \\s.t. & \ \ g(z^t,z) = 0
-        \label{eqn:simple2}
+        given & \ \ y_2 = D_2(y_1,z_1,z_2) \notag
+        \\      & \ \ y_1^t = D_1(x_1,y_2,z_1,z_2) \notag
+        \\min. &\ \ F(x_1,y_1,y_2,z_2) \notag
+        \\w.r.t. & \ \ x_1,y_1,y_2,z_1,z_2 \notag
+        \\s.t. & \ \ G(y_1,y_1^t) = 0
+        \label{eqn:simple}
     \end{align}
@@
     Equation \ref{eqn:simple2} differs only slightly from Eq. \ref{eqn:simple}. 
-    $A$ is now dependent on the output of $B$, and $z$ will be iterated on to 
-    convergence with $z^t$. Now the problem can be solved by running $B$ first 
-    and then $A$.Since the formulations in Eqs. \ref{eqn:simple} and 
+    $D_1$ is now dependent on the output of $D_2$, and $y_1$ will be iterated on to 
+    convergence with $y_1^t$. Now the problem can be solved by running $D_2$ first 
+    and then $D_1$. Since the formulations in Eqs. \ref{eqn:simple} and 
     \ref{eqn:simple2} both describe the same problem then neither can be the
@@
     \begin{align}
-        given & \ \ A: {x,y} \rightarrow {m,z^t} \notag
-        \\      & \ \ B: {x,z} \rightarrow {y^t} \notag
-        \\min. &\ \ f(m) \notag
-        \\w.r.t. & \ \ x,y,z \notag
-        \\s.t. & \ \ g1(z^t,z) = 0 \notag
-        \\     & \ \ g2(y^t,y) = 0
+        given & \ \ D_1.y_1 = D_1(x_1,y_2,z_1,z_2) \notag
+        \\      & \ \ D_2.y_2 = D_2(y_1,z_1,z_2) \notag
+        \\min. &\ \ F(x_1,y_1,y_2,z_2) \notag
+        \\w.r.t. & \ \ x_1,y_1,y_2,z_1,z_2 \notag
+        \\s.t. & \ \ G_1(D_1.y_1,D_2.y_1) = 0 \notag
+        \\     & \ \ G_2(D_1.y_2,D_2.y_2) = 0
         \label{eqn:simple_fpf}
@@
     met. The presence of both of these constraints fully decouples the problem so that 
-    either $A$ or $B$ could be run first or both could be run simultaneously. 
+    either $D_1$ or $D_2$ could be run first or both could be run simultaneously. 
     Alexandrov and Lewis, in their work with REMS, assert that fully decoupling 
     the analyses is key to achieving reconfigurability in the problem 
-    formulation\cite{alexandrov2004}.
+    formulation\cite{alexandrov2004}. The key feature they identify to allow this decoupling 
+    is the creation of separate variables for each analysis so that a variable 
+    is identified by its analysis and it's name. For example, in Eqn. \ref{eqn:simple}
+    there is only one $y_1$ variable vs Eqn. \ref{eqn:simple_fpf} where 
+    there are two ($D_1.y_1$ vs $D_2.y_1$).
 
@@
     Rogers et. al developed DeMAID to manipulate a DSM to find an ordering for analysis tools that 
-    reduces the cost of solving highly coupled systems\cite{Rogers1996,rogers1996demaid}. This re-ordering is done through 
+    reduces the cost of solving highly coupled systems\cite{rogers1996,rogers1996demaid}. This re-ordering is done through 
     row operations on the DSM matrix and yields multiple specific problem 

### 2013-04-13 example_problem.tex
label fix

@@
 \section{Example Problem}
+	\label{s:example problem}
 	This section presents an example problem to demonstrate the process of obtaining the FPG and the advantages of doing so.
-The task is the conceptual sizing of a single--aisle subsonic transport using an MDAO framework with objectives for performance and total weight for a required mission.
+The algorithm from Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG} was implemented in the NetworkX package of the Python programming language.
+Any language with the ability to manipulate a graph should be applicable.
+NetworkX was chosen due to the built--in features, such as cycle detection.
+
+	The example task is the conceptual sizing of a single--aisle subsonic transport using an MDAO framework with objectives for performance and total weight for a required mission.
 The full set of analysis codes available for use is given in Table \ref{t:analysis codes}.
@@
 \subsection{Obtaining an FPG}
-The process of obtaining an FPG from the given MCG, but it is expected to include the user because of required changes to the MCG or make decisions
+	\label{ss:obtaining an FPG}
+	The process of obtaining an FPG from the given MCG is expected to be an iterative process in which the user applies the algorithm to determine whether or not an FPG can be obtained, and then modifies the indegree limits or the MCG to obtain a satisfactory FPG.
+For example, after determining the location of a hole, the user can change the indegree limit to make the node a design variable, or the user could change the MCG by adding a new analysis code such that the hole is resolved.
 
-((I should show a figure of the MCG with edges and nodes colored as to when they were removed))
-((mention python))
-
-As presented in Sec.(()), the first step to obtaining an FPG is to detect holes.
-To start, we set the lower indegree limit for every node to be one to find any potential holes.
+	As presented in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}, the first step to obtaining an FPG is to detect holes.
+To start, we set the lower indegree limit for every node to be unity to find any potential holes.
 In this case, it turns out that the variable nodes representing the number of engines is a hole for NPSS, FLOPSa, and FLOPSb.
@@
 	  \end{center}
-  \caption{Simple notional problem with two potential design variables \label{f:design vars}}
+	  \caption{The MCG with one global variable missing. The resulting holes are shown in green and unused analyses shown in cyan.}
 	\label{f:MCG nengines}
@@
 
-While Sec.(()) indicated the process for obtaining an FPG, it did not specify how to decide which edges to keep when resolving a collision, that freedom is left to the specific implementation.
+	While Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG} indicated the process for obtaining an FPG, it did not specify how to decide which edges to keep when resolving a collision, that freedom is left to the specific implementation.
 The benefit of this approach is that standard graph theory algorithms can be used to make these decisions.
@@
 	\end{table}% 
-Step two in Sec.(()) (the resolution of collisions) is now conducted by selecting the edges directed from the analysis blocks with the highest rank for each collision.
+	Step two in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG} (the resolution of collisions) is now conducted by selecting the edges directed from the analysis blocks with the highest rank for each collision.
 The resulting FPG is shown in Fig.~\ref{f:FPG highest rank}, and this graph has four cycles, which are enumerated in Table \ref{t:cycles}.
@@
 	  \end{center}
-  \caption{Simple notional problem with two potential design variables \label{f:design vars}}
+	  \caption{FPG obtained using the ranking system.}
 	\label{f:FPG highest rank}

### 2013-04-13 graph_syntax.tex

--- 2013-04-10 graph_syntax.tex
@@
     \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \end{equation}
-		This represents an overload of information such that it is unclear which source to use.
 \end{description} 
-	An algorithm for detecting and managing these violations if proposed in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}.
+These violations will be managed in order to seek a working problem formulation.
 
@@
 \label{s:graph representation}
-	%As discussed in Sec.~\ref{s:intro}, the digraph must be able to describe the individual aspects of an MDAO problem.
+In this section we will make use of the Sellar problem, described in 
+Eqn. \ref{eqn:simple_fpf}. The graph representation of the Sellar problem, 
+using the new stynax, is presented in Fig. \ref{f:sellar_graph_full}. Throughout
+the rest of this section, different sub-graphs or slight modifications of 
+the full graph will be used to demonstrate the important MDAO constructs within
+this graph syntax.
+
+\begin{figure}[htb!]
+    \begin{center}
+    \includegraphics[width=4in]{images/sellar_graph_full}
+    \end{center}
+    \vspace{-10pt}
+\caption{Graph of the Sellar problem formulation}
+\label{f:sellar_graph_full}
+\end{figure}
+
+%As discussed in Sec.((intro)), the digraph must be able to describe the individual aspects of an MDAO problem.
 \subsection{Analysis Blocks and Connections}
@@
 This edge represents the necessary calculations to map given inputs 
-	to proper output values. This edge can be given a weight to represent the computational cost of the analysis block. However, it is also allowed to skip 
+to proper output values. If constructing a weighted graph, the weight calculation edge 
+provides the opportunity to computational cost. However, it is also allowed to skip 
 this edge and have both incomming and outgoing edges to variables through a single model 
@@
 the edges within them are fixed, they become fixed sub-graphs within a graph.
+
 \begin{figure}[htb!]
@@
 input of another analysis block, a new connection edge is added connecting the two 
-	variable nodes involved in the exchange. These connection edges are free---unlike the edges 
-	within an analysis block---which means they could be added or removed depending on the needs
+variable nodes involved in the exchange. These connection edges are free, unlike the edges 
+within an analysis block, they could be added or removed depending on the needs
 of a given design problem. In other words, free connection edges are created by 
@@
 \subsection{Design Variables}
-	\label{ss:design variables}
 ((DP:I'm thinking we might also need the upper indegree limit to be zero as well, because setting only the lower indegree limit makes the node a design variable if it didn't have any incoming edges.
@@
 I can make the change if you agree.))
-	Design variables require external decisions 
-	to be made about a given system to identify them. Consider 
-	a notional problem given by the graph in Fig.~\ref{f:designvars}. There are three input variables 
-	which could potentially become design variables: $w$, $x$, and $y$. Because the variable $x$ is connected from $A$ to $B$,
-	 $x$ can't actually be a design variable. On the other hand, the inputs $w$ and $y$
-	have no incomming edges and could each be selected as design variables. 
+For the Sellar problem, there are three design variables, $x_1^*$, $z_1^*$, and $z_2^*$. In Fig. 
+\ref{f:sellar_graph_full}, these are the only three holes (nodes without out 
+incomming edges) in the graph. Based on this example it's tempting to simply 
+call all holes in the graph design variables. However, in the general case this is 
+not completely correct. For any given problem, the set of holes only identifies 
+all the potential design variables. However, some variable holes might not make good design 
+variables. For instance, if your aircraft mission analysis code requires a drag input 
+which ends up being a hole then it's likely that some new analysis tool is needed
+to predict drag before the design process can continue.
+
 \begin{figure}[htb!]
   \begin{center}
+  [NOTE: Re-work this one to match $D_2$ from Sellar]
     \includegraphics[width=.6\textwidth]{images/design_vars_graph}
@@
 
-	Given a MCG with the $\txt{deg}_l^-(v)=1$ for all nodes, there will likely be a set of holes preventing a valid FPG from being obtained.
-	Each hole represents a node that could potentially become a design variable. 
-	It is up to the designer to examine each and decide if it is appropriate to 
-	consider it a design variable, in which case the designer shall set the lower indegree limit to be zero.
-However, some variable holes may not actually be suitable as a design variable.
-For instance, if an aircraft mission analysis code requires a drag input
-	which ends up being a hole then it's likely best to leave
-	$\txt{deg}_l^-(v)=1$ and find a new analysis code to fill the hole in graph.
+More formally stated, given a MCG with the $\txt{deg}_l^-(v)=1$ for all nodes, 
+there will be a set of holes. Each hole represents a node that could potentially 
+become a design variable. It is up to the designer to examine that variable and 
+decide if it is appropriate to use it as a design variable. If a variable should 
+be a design variable, then set it's ${deg}_l^-(v)=0$. Otherwise you leave 
+${deg}_l^-(v)=1$ and find some tool to fill the hole in graph.
 
-	A design variable is defined is any variable node with $\txt{deg}_l^-(v)=0$. ((DP: it seems like this should be have been at the beginning of this subsection.)) 
-	When working with a MCG or FPG, connection edges are the only free edges, so a design 
-	variable in either graph will be a variable node ((only variable nodes that are local inputs)) with no incoming edges. For a valid PSG 
+So a design variable is defined as any variable node with ${deg}_l^-(v)=0$. 
+This, by definition, means that design variables are not holes in the graph, since 
+they have not violated their lower in-degree limit. Regardless, 
+when working with a MCG or FPG, only connection edges are allowed, so a design 
+variable in either graph will be a variable with no incomming edges edges. For a valid PSG 
 any design variable node would have at least one, but possibly multiple, incomming 
@@
 \label{ss:objectives and constraints}
-	Objectives and constraints are also MDAO constructs that need to be identified by engineers 
-	as part of the design process. In the case of objective functions single output values 
-	could be selected, but commonly multiple values are assembeled together via simple 
-	algebraic expression to form a objective function. For example, the problem depicted in 
-	Fig.~\ref{f:designvars} could have an objective function of $z^2$ or $2y + z$. Constraints are 
-	always given in the form of either an inequality of an equality, e.g. 
-	$\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer compiles an
-	algebraic expression that can be evaluated to produce an output.
+Like design variables, objectives and constraints are also constructs that need 
+to be identified when setting up a given design problem. In the case of 
+objective functions single output values could be selected, but commonly multiple 
+values are assembeled together via simple algebraic expressions to form a composite 
+objective function. (e.g. The objective function from Eqn. \ref{eqn:simple_fpf} is
+$x_1^2+z_2+y_1+e^{-y_2}$). Constraints are always given in the 
+form of either an inequality or an equality. (e.g. a constraint from Eqn. 
+\ref{eqn:simple_fpf} is $\frac{y_1}{3.16}-1\geq0$.) By convention, 
+constraints are usually given such that some expression will be less than or equal 
+to zero, so any positive value would violate the constraint. For both objective 
+and constraints, these simple expressions take some inputs and map them to an
+output value of significance to the overall design problem.
 
@@
 blocks and those that arise from the addition of objectives or constraints to 
-	the graph. Therefore, we define an \emph{expression block} as the collection of variable and model 
-	nodes related to a given objective or constraint function. Fig.~\ref{f:obj-cons}
+the graph. So we define an expression block, as the collection of variable and model 
+nodes related to a given objective or constraint function. Fig. \ref{f:obj-cons}
 shows a notional problem graph augmented with expression blocks and appropriate 
 connection edges.
+
 \begin{figure}[htb!]
@@
 \subsection{Local vs Global Nodes}
-	\label{ss:local vs global}
-	A simple definition of a global node is any node that directly involves information 
+
+  A simple definition of a global node is any node that involves information 
   from more than one analysis block. Likewise, a local node is any node 
@@
   on block $B$. $g_2$ has incomming edges from both $A$ and $B$ analysis blocks, so it would be a global constraint.
+
   \begin{figure}
@@
 \subsection{Coupling Between Analyses}
-	\label{ss:coupling}
+
   Coupling exists in a design problem when a set of two or more analyses each depend on the
-	others outputs. Consider the notional problem from Sec.~\ref{s:specific vs fundamental} where two 
+  others outputs. Consider the notional problem from section \ref{sec:spg_vs_fpg} where two 
   analyses, $A$ and $B$ are coupled through their respective variables. The graph of this 
-	is shown in Fig.~\ref{f:coupling}. In the graph $A.z$ has an outgoing connection 
+  is shown in Fig. \ref{f:coupling}. In the graph $A.z$ has an outgoing connection 
   edge to $B.z$, indicating that there is a connection between the two. 
@@
   a cycle which represent the coupling between them. This cycle means that blocks $A$ and $B$ need
-	to be iterated on until they converge. 
+  to be iterated on untill they converge.
+
   \begin{figure}

### 2013-04-13 intro.tex

--- 2013-04-10 intro.tex
@@
 
-    $A$ and $B$ represent analysis tools, and $f$ and $g$ are the objective and constraint functions respectively. 
-    Equation \ref{eqn:simple} makes an inherent assumption about the solution strategy for the problem. 
-    Analysis $A$ outputs $z$, which is an input to $B$. Hence, $A$ should be run before $B$ with 
-    with $y$ being iterated on to convergence with $y^t$. However, a slightly different formulation is 
-    equally valid and still represents the exact same problem:
+    $D_1$ and $D_2$ represent analysis tools, and $F$ and $G$ are the objective 
+    and constraint functions respectively. Equation \ref{eqn:simple} makes an 
+    inherent assumption about the solution strategy for the problem. Analysis 
+    $D_1$ outputs $y_1$, which is an input to $D_2$. Hence, $D_1$ should be run 
+    before $D_2$ with with $y_2$ being iterated on to convergence with $y_2^t$. 
+    However, a slightly different formulation is equally valid and still 
+    represents the exact same problem:
 
     \begin{align}
-        given & \ \ B: {x,z} \rightarrow {y} \notag
-        \\      & \ \ A: {x,y} \rightarrow {m,z^t} \notag
-        \\min. &\ \ f(m) \notag
-        \\w.r.t. & \ \ x,y,z \notag
-        \\s.t. & \ \ g(z^t,z) = 0
-        \label{eqn:simple2}
+        given & \ \ y_2 = D_2(y_1,z_1,z_2) \notag
+        \\      & \ \ y_1^t = D_1(x_1,y_2,z_1,z_2) \notag
+        \\min. &\ \ F(x_1,y_1,y_2,z_2) \notag
+        \\w.r.t. & \ \ x_1,y_1,y_2,z_1,z_2 \notag
+        \\s.t. & \ \ G(y_1,y_1^t) = 0
+        \label{eqn:simple}
     \end{align}
 
-    Equation \ref{eqn:simple2} differs only slightly from Eq. \ref{eqn:simple}. $A$ is now dependent on the output of $B$, 
-    and $z$ will be iterated on to convergence with $z^t$. Now the problem can be solved by running $B$ first and then $A$.
-    Since the formulations in Eqs. \ref{eqn:simple} and \ref{eqn:simple2} both describe the same problem and neither can be the
-    FPF. They are both specific versions of the same more fundamental description of 
-    the problem that is common between them. We present the FPF as follows:
+    Equation \ref{eqn:simple2} differs only slightly from Eq. \ref{eqn:simple}. 
+    $D_1$ is now dependent on the output of $D_2$, and $y_1$ will be iterated on to 
+    convergence with $y_1^t$. Now the problem can be solved by running $D_2$ first 
+    and then $D_1$. Since the formulations in Eqs. \ref{eqn:simple} and 
+    \ref{eqn:simple2} both describe the same problem then neither can be the
+    FPF. They are both specific versions of a more fundamental description of 
+    the problem that is common between them. We present the fundamental 
+    formulation as follows:
 
     \begin{align}
-        given & \ \ A: {x,y} \rightarrow {m,z^t} \notag
-        \\      & \ \ B: {x,z} \rightarrow {y^t} \notag
-        \\min. &\ \ f(m) \notag
-        \\w.r.t. & \ \ x,y,z \notag
-        \\s.t. & \ \ g1(z^t,z) = 0 \notag
-        \\     & \ \ g2(y^t,y) = 0
+        given & \ \ D_1.y_1 = D_1(x_1,y_2,z_1,z_2) \notag
+        \\      & \ \ D_2.y_2 = D_2(y_1,z_1,z_2) \notag
+        \\min. &\ \ F(x_1,y_1,y_2,z_2) \notag
+        \\w.r.t. & \ \ x_1,y_1,y_2,z_1,z_2 \notag
+        \\s.t. & \ \ G_1(D_1.y_1,D_2.y_1) = 0 \notag
+        \\     & \ \ G_2(D_1.y_2,D_2.y_2) = 0
         \label{eqn:simple_fpf}
@@
 
-    The FPF in Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} and Eq. \ref{eqn:simple2} because it has 
-    two constraints which both must be met. The presence of both of these constraints fully decouples the problem so that 
-    either $A$ or $B$ could be run first or both could be run simultaneously. Alexandrov and Lewis investigated problem 
-    formulation in this fully decoupled way as part of their work on 
-    Reconfigurability in MDO Problem Synthesis (REMS) \cite{alexandrov2004}.
-    In this work they show how removing either constraint and replacing 
-    it with a direct dependence between the two analyses, a process they term ``collapsing'', 
-    you could regain the earlier two problem formulations.
+    The formulation in Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} 
+    and Eq. \ref{eqn:simple2} because it has two constraints which both must be 
+    met. The presence of both of these constraints fully decouples the problem so that 
+    either $D_1$ or $D_2$ could be run first or both could be run simultaneously. 
+    Alexandrov and Lewis, in their work with REMS, assert that fully decoupling 
+    the analyses is key to achieving reconfigurability in the problem 
+    formulation\cite{alexandrov2004}. The key feature they identify to allow this decoupling 
+    is the creation of separate variables for each analysis so that a variable 
+    is identified by its analysis and it's name. For example, in Eqn. \ref{eqn:simple}
+    there is only one $y_1$ variable vs Eqn. \ref{eqn:simple_fpf} where 
+    there are two ($D_1.y_1$ vs $D_2.y_1$).
 

### 2013-04-13 graph_syntax.tex
cleaning up merge conflict stuff some more

@@
 \ref{f:sellar_graph_full}, these are the only three holes (nodes without out 
-incomming edges) in the graph. Based on this example it's tempting to simply 
-call all holes in the graph design variables. However, in the general case this is 
-not completely correct. For any given problem, the set of holes only identifies 
-all the potential design variables. However, some variable holes might not make good design 
-variables. For instance, if your aircraft mission analysis code requires a drag input 
-which ends up being a hole then it's likely that some new analysis tool is needed
-to predict drag before the design process can continue.
+incomming edges) in the graph.
 
@@
 More formally stated, given a MCG with the $\txt{deg}_l^-(v)=1$ for all nodes, 
-there will be a set of holes. Each hole represents a node that could potentially 
-become a design variable. It is up to the designer to examine that variable and 
-decide if it is appropriate to use it as a design variable. If a variable should 
-be a design variable, then set it's ${deg}_l^-(v)=0$. Otherwise you leave 
-${deg}_l^-(v)=1$ and find some tool to fill the hole in graph.
+there will likely be a set of holes preventing a valid FPG from being obtained. 
+Each hole represents a node that could potentially become a design variable. 
+It is up to the designer to examine each and decide if it is appropriate to 
+consider it a design variable, in which case the designer shall set the lower 
+indegree limit to be zero. However, some variable holes may not actually be 
+suitable as a design variable. For instance, if an aircraft mission analysis 
+code requires a drag input which ends up being a hole then it's likely best to 
+leave$\txt{deg}_l^-(v)=1$ and find a new analysis code to fill the hole in graph.
 
-So a design variable is defined as any variable node with ${deg}_l^-(v)=0$. 
+So a design variable is defined as any variable node with $\txt{deg}_l^-(v)=0$. 
 This, by definition, means that design variables are not holes in the graph, since 

### 2013-04-13 graph_syntax.tex
fixing stuff from change to sellar

@@
 
-\subsection{In-degree and Out-degree}
+\subsection{indegree and outdegree}
 	\label{s:indegree-outdegree}
-	The \emph{in--degree} of a node is the number of connection edges directed in and 
-	is denoted as $\txt{deg}^-(v)$, and the \emph{out--degree} 
+	The \emph{indegree} of a node is the number of connection edges directed in and 
+	is denoted as $\txt{deg}^-(v)$, and the \emph{outdegree} 
 	is the number of connection edges directed out and it is denoted as $\txt{deg}^+(v)$.
@@
 	and a gradient based one is applied second, design variables would have incomming 
-	implicit edges from both optimizers but this would not affect the in--degree of those
+	implicit edges from both optimizers but this would not affect the indegree of those
 	variable nodes. 
 	((DP:I started to remove any reference to the edge type in these definitions but then saw that you needed it.
-The reason I did that was because in--degree and out--degree are basic graph theory terms.
-What if the reference to implicit edges is handled with the upper in--degree limit, i.e.
+The reason I did that was because indegree and outdegree are basic graph theory terms.
+What if the reference to implicit edges is handled with the upper indegree limit, i.e.
 after the FPG is obtained the upper indegree limit for the design variable nodes is relaxed (increased) so that it isn't a problem what is or isn't counted?))
 
-	((DP:the definitions were in/out degree were not in my book, I had to get them from wikipedia, where they are spelled indegree and outdegree))
-
-	We also define the \emph{upper in--degree limit}
+	We also define the \emph{upper indegree limit}
 
@@
 	\end{equation} 
-	and the \emph{lower in--degree limit} as
+	and the \emph{lower indegree limit} as
 	\begin{equation}
@@
   \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
-    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \end{equation}
+    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \label{e:collision}\end{equation}
 \end{description} 
@@
 
-%As discussed in Sec.((intro)), the digraph must be able to describe the individual aspects of an MDAO problem.
 \subsection{Analysis Blocks and Connections}
@@
 node represents a single input or output and is connected 
-to a single model node via a fixed edge. Note that in Fig. \ref{f:analysis block}, 
+to a single model node via a fixed edge. Note that in Fig.~\ref{f:analysis block}, 
 the analysis block contains two model nodes, with a single edge connecting them. 
@@
 
-	Variable nodes in an analysis block can be distinguished as either an input or an output by the manner in which they are connected.
-As shown in Fig.~\ref{f:analysis block}, inputs are any variable nodes that have an outgoing edge into a model
+Variable nodes in an analysis block can be distinguished as either an input or 
+an output by the manner in which they are connected. As shown in Fig.~\ref{f:analysis block}, 
+inputs are any variable nodes that have an outgoing edge into a model 
 node. Conversely, outputs are any variable nodes that have an incoming edge from a model node.
-	Note that by this definition, inputs and outputs can only exist when variable nodes poses an 
-	edge joining them to a model node. 
-	((DP:I'm not sure what the sentences before or after this comment mean))
-	These are labeled out as 
-	local inputs and local outputs. The variables are local because they are explicitely tied
-	to the analysis block and have no relevance without it.
 
@@
 This, by definition, means that design variables are not holes in the graph, since 
-they have not violated their lower in-degree limit. Regardless, 
+they have not violated their lower indegree limit. Regardless, 
 when working with a MCG or FPG, only connection edges are allowed, so a design 
@@
   nodes and edges within the block. Their locality is determined by the 
-  incomming and outgoing edges from the block as a whole. Fig. \ref{f:sellar_graph_full} has expression blocks for each constraint, $G_1$ and $G_2$. Both constraints are 
-  local to their respective analysis blocks. But the expression block for $F$ 
-  has incomming edges from $D_1$ and $D_2$, so it would considered global.
+  incomming and outgoing edges from the block as a whole. Fig. 
+  \ref{f:sellar_graph_full} has expression blocks for each constraint, 
+  $G_1$ and $G_2$. Both constraints are local to their respective analysis 
+  blocks. But the expression block for $F$ has incomming edges from $D_1$ and 
+  $D_2$, so it would considered global.
 
@@
   to the group $A,B$ with respect to $C$. This situation produces a natural 
-  hierarchy in a graph, where you can form increasingly smaller groups as you 
-  segment problem into finer localities. When solving a problem using 
-  a heirarchical decomposition approach, there are various techniques for 
-  partitioning that are effective in different scenarios 
-  \cite{krishnamachari1997optimal,Perez2004,sobieszczanski1997,allison2009optimal,michelena1997hypergraph}.
+  hierarchy in a graph, where you could form increasingly smaller groups as you 
+  segment problem into finer localities.
+
+  When solving a problem using a heirarchical decomposition approach, 
+  there are various techniques for partitioning that are effective in different 
+  scenarios \cite{krishnamachari1997optimal,michelena1997hypergraph,
+  sobieszczanski1997,Perez2004,allison2009optimal}. Tosserams et al proposed a 
+  language based syntax for describing the partitioning of problems, named $\Psi$
+  \cite{tosserams2010specification}. $\Psi$ provides an opposing perspective to 
+  this work, where they compose hierarchies from the bottom up, into larger and 
+  larger groups. However, $\Psi$ provides a compiler that can produce a standard 
+  problem representation of the assembled problem, and a number of example converters
+  from that standard representation to application specific formats. It would be 
+  possible to provide such a converter to build a graph form the $\Psi$ format.
+
+  [[Could probably build this and include it in appendix]]
 
@@
   Coupling exists in a design problem when a set of two or more analyses each depend on the
-  others outputs. Consider the notional problem from section \ref{sec:spg_vs_fpg} where two 
-  analyses, $A$ and $B$ are coupled through their respective variables. The graph of this 
-  is shown in Fig. \ref{f:coupling}. In the graph $A.z$ has an outgoing connection 
-  edge to $B.z$, indicating that there is a connection between the two. 
-  Similarly a connection exists between $B.y$ and $A.y$. These two connections form 
-  a cycle which represent the coupling between them. This cycle means that blocks $A$ and $B$ need
-  to be iterated on untill they converge.
+  others outputs. In the Sellar problem from Eqn \ref{eqn:simple_fpf} the two 
+  coupling constraints, $G_1$ and $G_2$ provide a reciprical dependence between 
+  $D_1$ and $D_2$. In a graph, these constraints show up as connection edges between 
+  the outputs of each analysis block. The relevant connection edges are highlighted in 
+  Fig. \ref{f:coupling}, where they form a cycle. Cycles in a graph are the characteristic 
+  structure that represents coupling.
 
@@
 
-	It is important to note that the cycle in graph in Fig.~\ref{f:coupling} does not contain 
-	any driver nodes. That means that no solver, optimizer, or other iterative construct is 
-	invovled in the coupling. This absence is intentional, as the coupling is an inherent part 
-	of the problem. This corresponds directly to Eqn.~\ref{eqn:simple_fpf}, where 
-	each equality constraint is one edge in the graph. Transitioning to either 
-	Eqn.~\ref{eqn:simple} or Eqn.~\ref{eqn:simple2} means selecting one of the nodes as the starting point 
-	for the iteration and indicating which associated analysis block should be executed first.
+	It is important to note that the coupling cycles do not contain 
+	any driver nodes. Solver, optimizer, or other iterative process is 
+	invovled in the coupling definition--though one will be necessary for a valid 
+  PSG. Additionally, a coupling cycle has no inherant start or end. It would be
+  valid to pick any node in the cycle as a starting point, and proceed around the
+  loop untill you get back to your starting point. For the Sellar Problem, picking 
+  $D_1$ as the starting point would yeild a problem as given in 
+	Eqn.~\ref{eqn:simple}, while picking $D_2$ would yield the problem as givne in 
+  Eqn.~\ref{eqn:simple2}.
 
@@
 	arise naturally as the size of problems grows, and managing this coupling may
-	become difficult. From the perspective of the graph, the problem reduces down to 
-	an ordering challenge. For each cycle in the graph, an analysis block needs to be selected 
-	to start the iteration from. Rogers' DEMAID tool uses a genetic algorithm to find an ordering 
-	that minimizes the overall coupling of the system by separating independent 
-	cycles in the graph \cite{rogers1996,rogers1996demaid}.
-In the present work, Sec.~\ref{s:example problem}.\ref{ss:obtaining an FPG} will suggest that an FPG may be obtained with respect to the number of cycles or the length of these cycles, etc.
+	become difficult. In the present work, Sec.~\ref{s:example problem}.\ref{ss:obtaining an FPG} 
+  will demonstrate how building an FPG from an MCG provides an opportunity to 
+  identify and potentailly reduce the number of cycles in a graph.
+
+  If large amounts of coupling are present it can help convergence 
+  rates to look for an effective ordering for the execution of analyses.
+  Rogers' DEMAID tool uses a genetic algorithm to find an ordering that minimizes 
+  the overall coupling of the system by separating independent cycles in the 
+  graph \cite{rogers1996,rogers1996demaid}. Rodgers work focused on the matrix 
+  form of the DSM for ordering optimization. Lu and Martins more recent work 
+  worked with a weighted form of the DSM and used an iterative clustering approach 
+  to perform a similar task to DEMAID \cite{Lu2012}.
+
+  However, when using a graph representation, the ordering problem can be greatly 
+  simplified. For any given coupling cycle, a Gauss-Seidel iteration pattern 
+  is automatically given simply folowing the cycle around the graph. If only 
+  once cycle is present, then the task reduces down to selecting a starting point
+  in the cycle. When sub-cycles are present each sub-cycle ordering is also 
+  given automatically by following the graph, but now a there is freedom at which 
+  level to manage the iteration over the cycles. Ordering of many cycles and 
+  sub-cycles is still a challenge, but the total search space is reduced by 
+  always following the ordering from the graph from a selected starting node.
 
@@
 	onto a graph will yield two analysis blocks, each with their own output 
-	variable nodes, having outgoing edges to a third analysis block that needs the 
-	data as input. Figure \ref{f:collision-example} shows a notional problem 
-	that exhibits this structure, where the key feature is that variable node $y$ in 
-	analysis block $C$ has two incoming edges. 
+	variable nodes, having outgoing edges to a single variable in a third analysis 
+  block that needs the data as input. Figure \ref{f:collision_example} shows 
+  a modified version of the Sellar problem with a new analysis, $D_0$, representing a 
+  low fidelity version of $D_1$.
+
 	\begin{figure}
@@
 	  [INSERT GRAPH HERE]
-	  \caption{Graph of a notional multi-fidelity problem \label{f:collision-example}}
+	  \caption{Graph of a notional multi-fidelity problem \label{f:collision_example}}
 	  \end{center}
@@
 	Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
-	Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper in-degree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.
+	Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper indegree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.
 
@@
 	\end{equation}
-	which is the set of variable nodes with fewer incoming edges than are allowed by the lower in-degree limit.
+	which is the set of variable nodes with fewer incoming edges than are allowed by the lower indegree limit.
 	Then the updated set of edges is created by removing the first set of connection edges following the hole:
@@
 	\end{equation}
-	which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper in--degree limit of $v_j$.
+	which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper indegree limit of $v_j$.
 The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include.

### 2013-04-13 intro.tex
fixing stuff from change to sellar

@@
         \\s.t. & \ \ G(y_1,y_1^t) = 0
-        \label{eqn:simple}
+        \label{eqn:simple2}
     \end{align}

### 2013-04-17 graph_syntax.tex
merged changes

@@
 
+	A \emph{path} $P=(V,E)$ from $x_0$ to $x_k$ in graph $G$ is a subgraph of $G$ with $V = \{x_0,x_1,\ldots,x_k\}$ and $E = \{(x_0,x_1),(x_1,x_2),\ldots,(x_{k-1},x_k)\}$.
+	A \emph{reverse path} $P_R$ in graph $G$ is a path on $R$, where $R$ is the reverse graph of $G$ obtained by switching the orientation of every edge.
+
 	%does this belong in the analysis block section???? 
@@
 	\item[connection edge:] Exchange of information between two nodes. These edges 
-	can either be fixed (can not be removed from graph) or free (can be removed). 
+	can either be fixed (can not be removed from graph) or free (can be removed). ((this needs to be modified, as analysis blocks cannot be altered but they can be removed as a group))
 	\item [implicit edge:] Implicit passing of information from a driver node to a 
@@
 \end{description} 
-These violations will be managed in order to seek a working problem formulation.
+	An algorithm for detecting and managing these violations if proposed in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}.
 
@@
 node for a given analysis block. Although analysis blocks are comprised of multiple nodes and edges, since all 
-the edges within them are fixed, they become fixed sub-graphs within a graph.
-
+	the edges within them are fixed, they become fixed sub-graphs within a graph. In this way, they cannot be altered but they can be removed as a group.
 \begin{figure}[htb!]
@@
 variable nodes involved in the exchange. These connection edges are free, unlike the edges 
-within an analysis block, they could be added or removed depending on the needs
+within an analysis block, can be added or removed depending on the needs
 of a given design problem. In other words, free connection edges are created by 
@@
 blocks and those that arise from the addition of objectives or constraints to 
-the graph. So we define an expression block, as the collection of variable and model 
-nodes related to a given objective or constraint function. Fig. \ref{f:obj-cons}
+	the graph. Therefore, we define an \emph{expression block} as the collection of variable and model 
+	nodes related to a given objective or constraint function. Fig.~\ref{f:obj-cons}
 shows a notional problem graph augmented with expression blocks and appropriate 
@@
   $D_1$ as the starting point would yeild a problem as given in 
-	Eqn.~\ref{eqn:simple}, while picking $D_2$ would yield the problem as givne in 
+	Eqn.~\ref{eqn:simple}, while picking $D_2$ would yield the problem as given in 
   Eqn.~\ref{eqn:simple2}.
@@
 	\begin{enumerate}
-	\item[(1)] $\displaystyle{I_F \subset I}$ such that the remaining conditions hold
-	\item[(2)] $\displaystyle{ \forall i \in I_F, \ \exists v \in V_{A_i} \txt{ with } t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v)>0 }$
-	%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_{A_i} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
-	%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_\txt{out} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
-	\item[(3)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} \st \txt{deg}^+(v) > 0\} }$ ((maybe this should be 1 instead of 0 to match the previous definition of global input))
-	\item[(4)] $\displaystyle{V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right)}$
-	\item[(5)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right), \ C_F \subset C_M}$
-	\item[(6)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
+	\item[(1)] $F \subset M$ and $V_\txt{out} \subset V_F$
+	\item[(2)] $\displaystyle{\forall i \in I, \txt{ if } F \cap A_i \neq \emptyset \txt{ then } A_i \subset F}$
+	\item[(3)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
+	\item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_\txt{out}}$
 	\end{enumerate}
-	The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). 
-	Requirement (2) stipulates that each analysis blocks in $F$ must have at least one local output that is being used. 
-	Requirement (3) stipulates that only the global inputs that are being used should be included. 
-	Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
-	Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper indegree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.
+	%The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). 
+	%Requirement (2) stipulates that each analysis blocks in $F$ must have at least one local output that is being used. 
+	%Requirement (3) stipulates that only the global inputs that are being used should be included. 
+	%Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
+	Requirement (1) asserts that only the nodes and edges provided by the maximal connectivity graph can be used in the fundamental problem graph and that every global output must be included. 
+	Requirement (2) requires that each analysis block must be entirely present or entirely absent. 
+	Requirement (3) stipulates that the number of edges directed into each variable node must be within the lower and upper in-degree limits; 
+	if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is a \emph{collision}.
+	Lastly, requirement (4) ensures that only the nodes that are being used are included in the FPG by requiring that for every node a reverse path exists from at least one global output to that node. 
+	The reverse graph $R_F$ is obtained from $F$ by simply switching the orientation of every edge in $E_F$.
 
-	The set $C_F$ is a set containing connection edges representing connections.
-Since $C_M$ is the set of all potential connections, we must have $C_F \subset C_M$.
-While no other conditions are explicitly stated for $C_F$, requirement (6) is actually a requirement on both $I_F$ and $C_F$.
+	%The set $C_F$ is a set containing connection edges representing connections. Since $C_M$ is the set of all potential connections, we must have $C_F \subset C_M$. 
+	%While no other conditions are explicitly stated for $C_F$, requirement (6) is actually a requirement on both $I_F$ and $C_F$.
 
@@
 	\label{ss:obtaining FPG}
-	In general, there may be multiple different graphs that satisfy the FPG conditions, though there may be none at all.
+	In general, there may be multiple different graphs that satisfy the FPG conditions in Sec.~\ref{ss:FPG}, though there may be none at all. 
 	Here, we describe a process for obtaining an FPG by starting with the MCG and disconnecting connection edges until the FPG conditions are met. 
@@
 
-	%Then, mathematically, the FPG starts with $\mathcal A_1 = \{1,\ldots,m\}$, $I_{F,1} = I$, and $C_{F,1} = C_M$.
-	Then, mathematically, the FPG starts with $C_{F,0} = C_M$, $E_{F,0} = E_M$, $V_{F,0} = V_M$.
-	The nodes where $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ are considered first.
-	\begin{enumerate}
-	\item The first step is to detect holes and disconnect the connection edges following them.
+((describe pruning as a subroutine enforcing step 4 and then step 2))
+
+	\begin{description}
+	\item[Subroutine: Pruning] 
+		Given a digraph $G_1 = (V_1,E_1)$, let the subroutine denoted as $P_\txt{prune}$ operate on $G_1$ to produce a digraph $G_2 = (V_2,E_2)$ that satisfies requirements (2) 
+		and (4), and write $G_2 = P_\txt{prune} ( G_1 )$. 
+		This is accomplished by first creating the reverse graph of $G_1$, $R = (V_1,E_R)$. Then adding a new node $x$ to $V_R$ and adding edges directed from this node to each of the global outputs, i.e.,
+	\begin{equation}
+		\forall y \in V_\txt{out}, (x,y) \in E_R
+	\end{equation}
+		The set of nodes that may be reached from the global outputs is constructed as
+	\begin{equation}
+		U = \{ n \in V_1 \st \exists P \txt{ a path from $x$ to $n$}, \ P \subset R \}.
+		\end{equation}
+		The list of analysis blocks with at least one node that may be reach is constructed as
+		\begin{equation}
+		I_1 = \{ i \in I \st V_{A_i} \cap U \neq \emptyset \}
+		\end{equation}
+		The set of nodes to remove is
+		\begin{equation}
+		N = \{ v \in V_1 \st v \in V_{A_i} \txt{ for } i \in I_1, \txt{ or } v \in V_\txt{in}\setminus U \},
+		\end{equation}
+		which means that $N$ is composed of unused analysis block nodes and unused global input nodes. Then the new set of nodes is created as
+		\begin{equation}
+		V_2 = V_1 \setminus N,
+		\end{equation}
+		and edges involving the removed nodes are also deleted
+		\begin{equation}
+		E_2 = E_1 \setminus \{ (a,b) \st a \in N \txt{ or } b \in N  \}.
+		\end{equation}
+		The set of connection edges can be obtained as
+		\begin{equation}
+		C_2 = \{(x,y) \in V_{2} \st \sim(x \in V_{A_i} \txt{ and } y \in V_{A_i} \txt{ for some } i \in I_1)\},
+		\end{equation}
+		which means $C_2 \cap V_{A_i} = \emptyset  \ \forall i \in I_1$, and $E_2 = C_2 \cup \left( \bigcup_{i \in I_1} E_{A_i} \right)$.
+
+	\item[Step 1: Holes] 
+		The first step is to detect holes and disconnect the connection edges following them. 
 		These connection edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs.
+
+		To start, the an initial graph is created as a pruned copy of the MCG
+		\begin{equation}
+		F_0 = P_\txt{prune}(M),
+		\end{equation}
+		where $F_0 = (V_{F,0},E_{F,0})$, and $C_{F,0}$ is also obtained as described. 
 		The set of variable nodes which are holes is created as
 		\begin{equation}
-	H = \{v \in V_M \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
+		H = \{v \in V_{F,0} \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
 		\end{equation}
-	which is the set of variable nodes with fewer incoming edges than are allowed by the lower indegree limit.
-	Then the updated set of edges is created by removing the first set of connection edges following the hole:
+		which is the set of variable nodes with fewer incoming edges than are allowed by the lower in-degree limit.
+		Then the updated set of edges is created by removing the edges preceding or succeeding the analysis block:
 		\begin{equation}
-	C_{F,1} = C_{F,0} \setminus \{(x,y) \in E_M \st x=E_M^3(v) \txt{ for } v \in H\},
+		C_{F,1} = C_{F,0} \setminus \{(x,y) \in C_{F,0} \st x \in V_{A_i} \txt{ or } y \in V_{A_i}, \txt{ and } H\cap V_{A_i} \neq \emptyset \},
 	\end{equation}
@@
 	and this step is demonstrated by Fig.~\ref{f:hole}. Because removing these edges can create new holes, this step must be repeated until no more holes are found.
+		If this step every finds a global output to be a hole, meaning $V_\txt{out} \cap H \neq \emptyset$, then the algorithm stops because an FPG cannot be obtained. 
+		Otherwise, it is guaranteed that an FPG can be obtained.
 	\begin{figure}[htb!]
@@
 
-	\item The second step is to detect collisions and to then disconnect enough connection edges so that the collisions are resolved. The set of variable nodes which contain collisions is created as
+	\item[Step 2: Collisions] 
+		The second step is to detect collisions and to then disconnect enough connection edges so that the collisions are resolved. 
+		The set of variable nodes which contain collisions is created as
 	\begin{equation}
@@
 	\begin{equation}
-	S_\txt{edges} = \big \{ \{(x,y) \in E_M\} \st y \in S_\txt{nodes} \big \}
+		S_\txt{edges} = \big \{ \{(x,y) \in E_M\} \ \big| \ y \in S_\txt{nodes} \big \}
 	\end{equation}
@@
 	\end{equation}
-	which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper indegree limit of $v_j$.
-The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include.
-Then let
+		which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper in--degree limit of $v_j$. 
+		The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include. Then let
 	\begin{equation}
@@
 
-	\item The third and final step is to remove any unused nodes and edges. The indexing set for only the analysis blocks with at least one used local output may be constructed as ((pruning))
+	\item[Step 3: Finalize] 
+		The third and final step is to prune the graph one last time
 	\begin{equation}
-	I_F =  \{i \in I \ | \ \exists v \in V_{A_i} \txt{ such that }  t_\txt{node}(E_{F,2}^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) > 0 \},
+		F = P_\txt{prune}(F_2),
 	\end{equation}
-	where the degree is calculated with respect to $F_2$. This set excludes both the analysis blocks that were unused in the original MCG and those that became unused in step 1 or step 2.
+		and $F = (V_F,E_F)$.
+		The indices of the used analysis blocks can then be found as
+	\begin{equation}
+		I_F = \{ i \in I \st A_i \cap F \neq \emptyset \}
+	\end{equation}
+%		\begin{
+%		\begin{equation}
+%		I_F =  \{i \in I \ | \ \exists v \in V_{A_i} \txt{ such that }  t_\txt{node}(E_{F,2}^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) > 0 \},
+%		\end{equation}
+%		where the degree is calculated with respect to $F_2$. 
+%		This set excludes both the analysis blocks that were unused in the original MCG and those that became unused in step 1 or step 2.
+%
+%		The set of used global inputs is constructed as
+%		\begin{equation}
+%		V_{\txt{in},F} = \big\{v \in V_\txt{in} \ \big| \ |\{(x,y) \in E_{F,2}(v) \st y \in V_{A_i} \txt{ for } i \in I_F  \}| > 1 \big\},
+%		\end{equation}
+%		which requires that at least two edges directed out of the nodes in $V_{\txt{in},F}$ must be to the analysis blocks indexed by $I_F$.
+%
+%		Finally, the set of nodes and edges describing a fundamental problem graph is given by
+%		\begin{equation}
+%		V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right),
+%		\end{equation}
+%		\begin{equation}
+%		C_{F} = \{ (x,y) \in C_{F,2} \st x,y \in V_F\},
+%		\end{equation}
+%		\begin{equation}
+%		E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right),
+%		\end{equation}
+%		\begin{equation}
+%		F = (V_F,E_F).
+%		\end{equation}
 
-	The set of used global inputs is constructed as
-	\begin{equation}
-	V_{\txt{in},F} = \Big\{v \in V_\txt{in} \ \Big| \ |\{(x,y) \in E_{F,2}(v) \st y \in V_{A_i} \txt{ for } i \in I_F  \}| > 1 \Big\},
-	\end{equation}
-	which requires that at least two edges directed out of the nodes in $V_{\txt{in},F}$ must be to the analysis blocks indexed by $I_F$.
+	\end{description}
 
-	Finally, the set of nodes and edges describing a fundamental problem formulation is given by
-	\begin{equation}
-	V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right),
-	\end{equation}
-	\begin{equation}
-	C_{F} = \{ (x,y) \in C_{F,2} \st x,y \in V_F\},
-	\end{equation}
-	\begin{equation}
-	E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right),
-	\end{equation}
-	\begin{equation}
-	F = (V_F,E_F).
-	\end{equation}
-
-	\end{enumerate}
+This process will always provide an FPG if one exists. If an FPG does not exist, this process will reveal the limiting factors preventing a valid formulation.

### 2013-04-18 graph_syntax.tex
stuff

--- 2013-04-13 graph_syntax.tex
@@
 nodes related to a given objective or constraint function. Fig. \ref{f:obj-cons}
-shows a notional problem graph augmented with expression blocks and appropriate 
-connection edges.
+highlights the objective and constraint blocks from the Sellar problem.
 
@@
   definition, where $x_1^*$ is a global variable fits nicely within 
-  the structure of Braun's Colaborative Optimization (CO) archtiecture \cite{braun_thesis}. 
+  the structure of Braun's Colaborative Optimization (CO) archtiecture \cite{braun1996thesis}. 
   The rules of setting up a problem with CO require that each discipline be 
-  given an independent local copy of all global variables, while any local 
-  variables are retained uniquely within their respective disciplines. But 
-  when a local design variable shows up explicitely in the objective function 
-  or global constraint then a global variable is created, and the discipline again 
+  given an independent local copy of all global variables. All local 
+  variables are retained uniquely within their respective disciplines, except 
+  when a local design variable shows up explicitely in the objective 
+  or global constraint. In that case a global variable is created, and the discipline again 
   gets an independent local copy of it. Given our definition of global vs local
-  variable nodes, no such exception for local variables is necessary. When you 
-  include a given variable in an expression node, it becomes global by definition 
-  and should be treated as such in CO.
+  variable nodes, no such exception for local variables in global expressions is 
+  necessary. When you include a given variable in an expression block, it becomes 
+  global by definition and should be treated as such in CO.
 
@@
 
-  \begin{figure}
-      \begin{center}
-      %\includegraphics[height=.25\textheight]{XDSM/simple}
-      [[insert graph here]]
-      \caption{A simple graph with local and global nodes \label{f:local_global}}
-      \end{center}
-  \end{figure}
-
   Simply stating that any node or block with incomming or outgoing edges
@@
   to two separate analysis blocks, $A$ and $B$, but none to a third, $C$. Then 
-  you could say that this node was global with respect to $A$ and $B$, but local 
-  to the group $A,B$ with respect to $C$. This situation produces a natural 
+  you could say that this node was global with respect to $A$ and $B$ separately, but local 
+  to the group $A,B$. This situation produces a natural 
   hierarchy in a graph, where you could form increasingly smaller groups as you 
@@
   problem representation of the assembled problem, and a number of example converters
-  from that standard representation to application specific formats. It would be 
-  possible to provide such a converter to build a graph form the $\Psi$ format.
+  from that standard representation to application specific formats.
 

### 2013-04-18 graph_syntax.tex

--- 2013-04-17 graph_syntax.tex
@@
 blocks and those that arise from the addition of objectives or constraints to 
-	the graph. Therefore, we define an \emph{expression block} as the collection of variable and model 
-	nodes related to a given objective or constraint function. Fig.~\ref{f:obj-cons}
-shows a notional problem graph augmented with expression blocks and appropriate 
-connection edges.
+the graph. Therefore, we define an \emph{expression block}, as the collection of variable and model 
+nodes related to a given objective or constraint function. Fig. \ref{f:obj-cons}
+highlights the objective and constraint blocks from the Sellar problem.
 

### 2013-04-18 graph_syntax.tex
split graph_syntax into two files

--- 2013-04-18 graph_syntax.tex
@@
 blocks and those that arise from the addition of objectives or constraints to 
-the graph. So we define an expression block, as the collection of variable and model 
+the graph. Therefore, we define an \emph{expression block}, as the collection of variable and model 
 nodes related to a given objective or constraint function. Fig. \ref{f:obj-cons}
@@
 	run\cite{march2012provably,alexandrov2001approximation,Huang_Allen_Notz_Miller_2006}.
-
-\section{Building MCG and FPG}
-	\label{s:building graphs}
-\subsection{Maximal Connectivity Graph}
-	\label{ss:MCG}
-	To construct the maximal connectivity graph, we assume that a set of codes, global inputs, and objectives and constraints (collectively called global outputs) are provided.
-The codes are represented by analysis block graphs ((ABGs?)) $A_i=(V_{A_i},E_{A_i}), \ i\in I$, where $m$ is the number of codes and $I=\{1,2,\ldots,m\}$.
-The global inputs are represented as a set of variable nodes $V_\txt{in}$, and the global outputs are represented by a set of variable nodes $V_\txt{out}$.
-	We assume that $V_\txt{in}$, $V_\txt{out}$, and each $A_i$ are given, and that any potential connection between variables is given in the form of connection edges in the set $C_M$. 
-	Then we may construct the maximal connectivity graph $M=(V_M,E_M)$ as
-	\begin{IEEEeqnarray*}{rCl}
-	V_M & = & V_\txt{in} \cup V_\txt{out} \cup \left( \bigcup_{i \in I} V_{A_i} \right), \\
-	E_M & = & C_M \cup \left( \bigcup_{i \in I} E_{A_i} \right),
-	\end{IEEEeqnarray*}
-	The MCG $M$ is uniquely determined by the given set of analysis blocks, the required outputs, and the given global inputs.
-In the cases where the set of global inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.
-
-\subsection{Fundamental Problem Graph}
-	\label{ss:FPG}
-	We now define the fundamental problem graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions, which are explained subsequently,
-	\begin{enumerate}
-	\item[(1)] $\displaystyle{I_F \subset I}$ such that the remaining conditions hold
-	\item[(2)] $\displaystyle{ \forall i \in I_F, \ \exists v \in V_{A_i} \txt{ with } t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v)>0 }$
-	%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_{A_i} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
-	%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_\txt{out} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
-	\item[(3)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} \st \txt{deg}^+(v) > 0\} }$ ((maybe this should be 1 instead of 0 to match the previous definition of global input))
-	\item[(4)] $\displaystyle{V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right)}$
-	\item[(5)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right), \ C_F \subset C_M}$
-	\item[(6)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
-	\end{enumerate}
-	The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). 
-	Requirement (2) stipulates that each analysis blocks in $F$ must have at least one local output that is being used. 
-	Requirement (3) stipulates that only the global inputs that are being used should be included. 
-	Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
-	Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper indegree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.
-
-	The set $C_F$ is a set containing connection edges representing connections.
-Since $C_M$ is the set of all potential connections, we must have $C_F \subset C_M$.
-While no other conditions are explicitly stated for $C_F$, requirement (6) is actually a requirement on both $I_F$ and $C_F$.
-
-\subsection{Obtaining the Fundamental Problem Graph}
-	\label{ss:obtaining FPG}
-	In general, there may be multiple different graphs that satisfy the FPG conditions, though there may be none at all.
-Here, we describe a process for obtaining an FPG by starting with the MCG and disconnecting connection edges until the FPG conditions are met.
-Then the problem is reduced to deciding which connection edges to remove.
-
-	%Then, mathematically, the FPG starts with $\mathcal A_1 = \{1,\ldots,m\}$, $I_{F,1} = I$, and $C_{F,1} = C_M$.
-	Then, mathematically, the FPG starts with $C_{F,0} = C_M$, $E_{F,0} = E_M$, $V_{F,0} = V_M$.
-	The nodes where $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ are considered first.
-	\begin{enumerate}
-	\item The first step is to detect holes and disconnect the connection edges following them.
-These connection edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs.
-The set of variable nodes which are holes is created as
-	\begin{equation}
-	H = \{v \in V_M \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
-	\end{equation}
-	which is the set of variable nodes with fewer incoming edges than are allowed by the lower indegree limit.
-	Then the updated set of edges is created by removing the first set of connection edges following the hole:
-	\begin{equation}
-	C_{F,1} = C_{F,0} \setminus \{(x,y) \in E_M \st x=E_M^3(v) \txt{ for } v \in H\},
-	\end{equation}
-	%deg-v is now recaculated for the new C. 
-	and this step is demonstrated by Fig.~\ref{f:hole}. Because removing these edges can create new holes, this step must be repeated until no more holes are found.
-	\begin{figure}[htb!]
-		\begin{center}
-		\includegraphics[width=3in]{images/analysis_block_hole}
-		\end{center}
-		\vspace{-20pt}
-	\caption{Example variable node indicating a hole.}
-	\label{f:hole}
-	\end{figure}
-
-	\item The second step is to detect collisions and to then disconnect enough connection edges so that the collisions are resolved. The set of variable nodes which contain collisions is created as
-	\begin{equation}
-	S_\txt{nodes} = \{v \in V_M \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
-	\end{equation}
-	For each collision node we can construct a set contained the edges directed in. The set containing all of these sets is constructed as
-	\begin{equation}
-	S_\txt{edges} = \big \{ \{(x,y) \in E_M\} \st y \in S_\txt{nodes} \big \}
-	\end{equation}
-	Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$. 
-	An example collision is shown in Fig.~\ref{f:collision} to indicate the definition of $S_{\txt{edges},j}$. 
-	We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$ (which are sets). 
-	Then we may construct sets of edges as
-	\begin{equation}
-	B_j = \big \{e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \txt{ with } K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
-	\end{equation}
-	which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper indegree limit of $v_j$.
-The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include.
-Then let
-	\begin{equation}
-	C_{F,2} = \{ e \in C_{F,1} \st e \in B_j \txt{ for some } j \in J\},
-	\end{equation}
-	\begin{equation}
-	E_{F,2} = E_M \setminus (C_M \setminus C_{F,2})
-	\end{equation}
-	and
-	\begin{equation}
-	F_2 = (V_M,E_{F,2})
-	\end{equation}
-	\begin{figure}[htb!]
-		\begin{center}
-		\includegraphics[width=3in]{images/analysis_block_collision}
-		\end{center}
-		\vspace{-20pt}
-	\caption{Example variable node indicating a collision.}
-	\label{f:collision}
-	\end{figure}
-
-	\item The third and final step is to remove any unused nodes and edges. The indexing set for only the analysis blocks with at least one used local output may be constructed as ((pruning))
-	\begin{equation}
-	I_F =  \{i \in I \ | \ \exists v \in V_{A_i} \txt{ such that }  t_\txt{node}(E_{F,2}^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) > 0 \},
-	\end{equation}
-	where the degree is calculated with respect to $F_2$. This set excludes both the analysis blocks that were unused in the original MCG and those that became unused in step 1 or step 2.
-
-	The set of used global inputs is constructed as
-	\begin{equation}
-	V_{\txt{in},F} = \Big\{v \in V_\txt{in} \ \Big| \ |\{(x,y) \in E_{F,2}(v) \st y \in V_{A_i} \txt{ for } i \in I_F  \}| > 1 \Big\},
-	\end{equation}
-	which requires that at least two edges directed out of the nodes in $V_{\txt{in},F}$ must be to the analysis blocks indexed by $I_F$.
-
-	Finally, the set of nodes and edges describing a fundamental problem formulation is given by
-	\begin{equation}
-	V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right),
-	\end{equation}
-	\begin{equation}
-	C_{F} = \{ (x,y) \in C_{F,2} \st x,y \in V_F\},
-	\end{equation}
-	\begin{equation}
-	E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right),
-	\end{equation}
-	\begin{equation}
-	F = (V_F,E_F).
-	\end{equation}
-
-	\end{enumerate}

### 2013-04-18 mcg_fpg.tex
split graph_syntax into two files

\section{Building MCG and FPG}
    \label{s:building graphs}
\subsection{Maximal Connectivity Graph}
    \label{ss:MCG}
    To construct the maximal connectivity graph, we assume that a set of codes, global inputs, and objectives and constraints (collectively called global outputs) are provided. 
    The codes are represented by analysis block graphs ((ABGs?)) $A_i=(V_{A_i},E_{A_i}), \ i\in I$, where $m$ is the number of codes and $I=\{1,2,\ldots,m\}$. 
    The global inputs are represented as a set of variable nodes $V_\txt{in}$, and the global outputs are represented by a set of variable nodes $V_\txt{out}$. 
    We assume that $V_\txt{in}$, $V_\txt{out}$, and each $A_i$ are given, and that any potential connection between variables is given in the form of connection edges in the set $C_M$. 
    Then we may construct the maximal connectivity graph $M=(V_M,E_M)$ as
    \begin{IEEEeqnarray*}{rCl}
    V_M & = & V_\txt{in} \cup V_\txt{out} \cup \left( \bigcup_{i \in I} V_{A_i} \right), \\
    E_M & = & C_M \cup \left( \bigcup_{i \in I} E_{A_i} \right),
    \end{IEEEeqnarray*}
    The MCG $M$ is uniquely determined by the given set of analysis blocks, the required outputs, and the given global inputs. 
    In the cases where the set of global inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.

\subsection{Fundamental Problem Graph}
    \label{ss:FPG}
    We now define the fundamental problem graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions, which are explained subsequently,
    \begin{enumerate}
    \item[(1)] $F \subset M$ and $V_\txt{out} \subset V_F$
    \item[(2)] $\displaystyle{\forall i \in I, \txt{ if } F \cap A_i \neq \emptyset \txt{ then } A_i \subset F}$
    \item[(3)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_\txt{out}}$
    \end{enumerate}
    %The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). 
    %Requirement (2) stipulates that each analysis blocks in $F$ must have at least one local output that is being used. 
    %Requirement (3) stipulates that only the global inputs that are being used should be included. 
    %Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
    Requirement (1) asserts that only the nodes and edges provided by the maximal connectivity graph can be used in the fundamental problem graph and that every global output must be included. 
    Requirement (2) requires that each analysis block must be entirely present or entirely absent. 
    Requirement (3) stipulates that the number of edges directed into each variable node must be within the lower and upper in-degree limits; 
    if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is a \emph{collision}.
    Lastly, requirement (4) ensures that only the nodes that are being used are included in the FPG by requiring that for every node a reverse path exists from at least one global output to that node. 
    The reverse graph $R_F$ is obtained from $F$ by simply switching the orientation of every edge in $E_F$.

    %The set $C_F$ is a set containing connection edges representing connections. Since $C_M$ is the set of all potential connections, we must have $C_F \subset C_M$. 
    %While no other conditions are explicitly stated for $C_F$, requirement (6) is actually a requirement on both $I_F$ and $C_F$.

\subsection{Obtaining the Fundamental Problem Graph}
    \label{ss:obtaining FPG}
    In general, there may be multiple different graphs that satisfy the FPG conditions in Sec.~\ref{ss:FPG}, though there may be none at all. 
    Here, we describe a process for obtaining an FPG by starting with the MCG and disconnecting connection edges until the FPG conditions are met. 
    Then the problem is reduced to deciding which connection edges to remove.

((describe pruning as a subroutine enforcing step 4 and then step 2))

    \begin{description}
    \item[Subroutine: Pruning] 
        Given a digraph $G_1 = (V_1,E_1)$, let the subroutine denoted as $P_\txt{prune}$ operate on $G_1$ to produce a digraph $G_2 = (V_2,E_2)$ that satisfies requirements (2) 
        and (4), and write $G_2 = P_\txt{prune} ( G_1 )$. 
        This is accomplished by first creating the reverse graph of $G_1$, $R = (V_1,E_R)$. Then adding a new node $x$ to $V_R$ and adding edges directed from this node to each of the global outputs, i.e.,
    \begin{equation}
        \forall y \in V_\txt{out}, (x,y) \in E_R
    \end{equation}
        The set of nodes that may be reached from the global outputs is constructed as
    \begin{equation}
        U = \{ n \in V_1 \st \exists P \txt{ a path from $x$ to $n$}, \ P \subset R \}.
        \end{equation}
        The list of analysis blocks with at least one node that may be reach is constructed as
        \begin{equation}
        I_1 = \{ i \in I \st V_{A_i} \cap U \neq \emptyset \}
        \end{equation}
        The set of nodes to remove is
        \begin{equation}
        N = \{ v \in V_1 \st v \in V_{A_i} \txt{ for } i \in I_1, \txt{ or } v \in V_\txt{in}\setminus U \},
        \end{equation}
        which means that $N$ is composed of unused analysis block nodes and unused global input nodes. Then the new set of nodes is created as
        \begin{equation}
        V_2 = V_1 \setminus N,
        \end{equation}
        and edges involving the removed nodes are also deleted
        \begin{equation}
        E_2 = E_1 \setminus \{ (a,b) \st a \in N \txt{ or } b \in N  \}.
        \end{equation}
        The set of connection edges can be obtained as
        \begin{equation}
        C_2 = \{(x,y) \in V_{2} \st \sim(x \in V_{A_i} \txt{ and } y \in V_{A_i} \txt{ for some } i \in I_1)\},
        \end{equation}
        which means $C_2 \cap V_{A_i} = \emptyset  \ \forall i \in I_1$, and $E_2 = C_2 \cup \left( \bigcup_{i \in I_1} E_{A_i} \right)$.

    \item[Step 1: Holes] 
        The first step is to detect holes and disconnect the connection edges following them. 
        These connection edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs.

        To start, the an initial graph is created as a pruned copy of the MCG
        \begin{equation}
        F_0 = P_\txt{prune}(M),
        \end{equation}
        where $F_0 = (V_{F,0},E_{F,0})$, and $C_{F,0}$ is also obtained as described. 
        The set of variable nodes which are holes is created as
        \begin{equation}
        H = \{v \in V_{F,0} \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
        \end{equation}
        which is the set of variable nodes with fewer incoming edges than are allowed by the lower in-degree limit.
        Then the updated set of edges is created by removing the edges preceding or succeeding the analysis block:
        \begin{equation}
        C_{F,1} = C_{F,0} \setminus \{(x,y) \in C_{F,0} \st x \in V_{A_i} \txt{ or } y \in V_{A_i}, \txt{ and } H\cap V_{A_i} \neq \emptyset \},
    \end{equation}
    %deg-v is now recaculated for the new C. 
    and this step is demonstrated by Fig.~\ref{f:hole}. Because removing these edges can create new holes, this step must be repeated until no more holes are found.
        If this step every finds a global output to be a hole, meaning $V_\txt{out} \cap H \neq \emptyset$, then the algorithm stops because an FPG cannot be obtained. 
        Otherwise, it is guaranteed that an FPG can be obtained.
    \begin{figure}[htb!]
        \begin{center}
        \includegraphics[width=3in]{images/analysis_block_hole}
        \end{center}
        \vspace{-20pt}
    \caption{Example variable node indicating a hole.}
    \label{f:hole}
    \end{figure}

    \item[Step 2: Collisions] 
        The second step is to detect collisions and to then disconnect enough connection edges so that the collisions are resolved. 
        The set of variable nodes which contain collisions is created as
    \begin{equation}
    S_\txt{nodes} = \{v \in V_M \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
    \end{equation}
    For each collision node we can construct a set contained the edges directed in. The set containing all of these sets is constructed as
    \begin{equation}
        S_\txt{edges} = \big \{ \{(x,y) \in E_M\} \ \big| \ y \in S_\txt{nodes} \big \}
    \end{equation}
    Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$. 
    An example collision is shown in Fig.~\ref{f:collision} to indicate the definition of $S_{\txt{edges},j}$. 
    We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$ (which are sets). 
    Then we may construct sets of edges as
    \begin{equation}
    B_j = \big \{e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \txt{ with } K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
    \end{equation}
        which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper in--degree limit of $v_j$. 
        The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include. Then let
    \begin{equation}
    C_{F,2} = \{ e \in C_{F,1} \st e \in B_j \txt{ for some } j \in J\},
    \end{equation}
    \begin{equation}
    E_{F,2} = E_M \setminus (C_M \setminus C_{F,2})
    \end{equation}
    and
    \begin{equation}
    F_2 = (V_M,E_{F,2})
    \end{equation}
    \begin{figure}[htb!]
        \begin{center}
        \includegraphics[width=3in]{images/analysis_block_collision}
        \end{center}
        \vspace{-20pt}
    \caption{Example variable node indicating a collision.}
    \label{f:collision}
    \end{figure}

    \item[Step 3: Finalize] 
        The third and final step is to prune the graph one last time
    \begin{equation}
        F = P_\txt{prune}(F_2),
    \end{equation}
        and $F = (V_F,E_F)$.
        The indices of the used analysis blocks can then be found as
    \begin{equation}
        I_F = \{ i \in I \st A_i \cap F \neq \emptyset \}
    \end{equation}
%       \begin{
%       \begin{equation}
%       I_F =  \{i \in I \ | \ \exists v \in V_{A_i} \txt{ such that }  t_\txt{node}(E_{F,2}^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) > 0 \},
%       \end{equation}
%       where the degree is calculated with respect to $F_2$. 
%       This set excludes both the analysis blocks that were unused in the original MCG and those that became unused in step 1 or step 2.
%
%       The set of used global inputs is constructed as
%       \begin{equation}
%       V_{\txt{in},F} = \big\{v \in V_\txt{in} \ \big| \ |\{(x,y) \in E_{F,2}(v) \st y \in V_{A_i} \txt{ for } i \in I_F  \}| > 1 \big\},
%       \end{equation}
%       which requires that at least two edges directed out of the nodes in $V_{\txt{in},F}$ must be to the analysis blocks indexed by $I_F$.
%
%       Finally, the set of nodes and edges describing a fundamental problem graph is given by
%       \begin{equation}
%       V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right),
%       \end{equation}
%       \begin{equation}
%       C_{F} = \{ (x,y) \in C_{F,2} \st x,y \in V_F\},
%       \end{equation}
%       \begin{equation}
%       E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right),
%       \end{equation}
%       \begin{equation}
%       F = (V_F,E_F).
%       \end{equation}

    \end{description}

This process will always provide an FPG if one exists. If an FPG does not exist, this process will reveal the limiting factors preventing a valid formulation.

### 2013-04-18 graph_syntax.tex
working on formatting for SMO

@@
   \end{equation}
-	These user-specified limits govern the number of connection edges that may be directed into a variable
-	node for a valid graph. Consider a variable node $v$ with
-	$\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, $v$ must have exactly one
-	incomming explicit edge or the graph is invalid. Two possiblities exist for voilating 
-	these limit conditions: 
+  These user-specified limits govern the number of connection edges that may 
+  be directed into a variable node for a valid graph. Consider a variable 
+  node $v$ with $\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, $v$ 
+  must have exactly one incomming explicit edge or the graph is invalid. 
+  Two possiblities exist for voilating these limit conditions:
+
   \begin{description}
@@
       This represents a lack of information being supplied to the node.
-
     \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
@@
   \end{description}
+
   An algorithm for detecting and managing these violations if proposed in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}.
@@
 to proper output values. If constructing a weighted graph, the weight calculation edge 
-provides the opportunity to computational cost. However, it is also allowed to skip 
-this edge and have both incomming and outgoing edges to variables through a single model 
-node for a given analysis block. Although analysis blocks are comprised of multiple nodes and edges, since all 
-	the edges within them are fixed, they become fixed sub-graphs within a graph. In this way, they cannot be altered but they can be removed as a group.
+provides the opportunity to computational cost. However, it is also allowed to 
+skip this edge and have both incomming and outgoing edges to variables through a 
+single model node for a given analysis block. Although analysis blocks are 
+comprised of multiple nodes and edges, since all the edges within them are 
+fixed, they become fixed sub-graphs within a graph. In this way, they cannot be 
+altered but they can be removed as a group.
+
 \begin{figure}[htb!]
@@
   there are various techniques for partitioning that are effective in different 
-  scenarios \cite{krishnamachari1997optimal,michelena1997hypergraph,
-  sobieszczanski1997,Perez2004,allison2009optimal}. Tosserams et al proposed a 
-  language based syntax for describing the partitioning of problems, named $\Psi$
+  scenarios \cite{krishnamachari1997optimal,michelena1997hypergraph,sobieszczanski1997,Perez2004,allison2009optimal}. 
+  Tosserams et al proposed a language based syntax for describing the partitioning of problems, named $\Psi$
   \cite{tosserams2010specification}. $\Psi$ provides an opposing perspective to 

### 2013-04-18 problem_formulation.tex
writing update
fixing a few things to get it to compile
most of suggested major changes
updated introduction and started editing a couple of lower sections
updated the Desgin Variable, Objectives, Constraints section with some new notes. Two new node types
put in example problem using a new .tex file
split graph_syntax into two files
working on formatting for SMO

@@
-\documentclass[]{aiaa-tc} % insert '[draft]' option to show overfull boxes
+\documentclass[referee]{svjour3} % insert '[draft]' option to show overfull boxes
 
- \title{An application of Graph Theory to MDAO problem formulation}
+ \title{A Graph Theory Based Approach to MDAO Problem Formulation}
 
@@
   David Pate, %
-     \thanks{Georgia Tech}
+  Justin Gray,%
   Dr. Brian German 
-     \thanks{Georgia Tech...}
-  Justin Gray,%
-     \thanks{Aerospace Engineer, MDAO Branch, Mail Stop 5-11, AIAA Member}   
  }
+
+ \institute { Justin Gray \at NASA Glenn Research Center, Mail Stop 5-11, 21000 Brookpark Rd Clevland OH 44135 %
+  \and David Pate \at Gatech \and Dr. Brian German \at Gatech}
+
+\journalname{Structrual and Multidisciplinary Optimization}
 
@@
 \usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsfonts}
 \usepackage{lscape}
@@
 \usepackage[section]{placeins}
+\usepackage[retainorgcmds]{IEEEtrantools}
+\usepackage{subfigure}
+\usepackage{comment}
+\usepackage{booktabs}
 
@@
 \usepackage{verbatim}
+
+\newcommand{\txt}{\textrm}
 
@@
 \input{graph_syntax}
-
-\section{Example Problem}
+\input{mcg_fpg}
+\input{example_problem}
 

### 2013-04-19 graph_syntax.tex
updated per brian's suggestions

@@
   \begin{description}
-  \item[variable: ] represents scalar or array data
+    \item[variable:] represents scalar or array data
     \item[model:] responsible for mapping inputs to outputs
@@
   \item[connection edge:] Exchange of information between two nodes. These edges 
-  can either be fixed (can not be removed from graph) or free (can be removed). ((this needs to be modified, as analysis blocks cannot be altered but they can be removed as a group))
-  \item [implicit edge:] Implicit passing of information from a driver node to a 
-    variable node. A single variable node can have many incomming implicit edges. These edges are 
+  can either be fixed (can not be removed from graph) or free (can be removed). 
+  \item [driven edge:] Passing of information from a driver node to a 
+    variable node. A single variable node can have many incomming driven edges. These edges are 
     free and can be added or removed from the graph. 
@@
   type to ``model'' to be more consistent with modern MDAO frameworks. The present work 
-  adds two more node types and one more edge type to the syntax to allow descriptive
-  graphs for all three phases of the design process.
+  adds one more node types and one more edge type to the syntax to allow descriptive
+  graphs for all three phases of the design process. In addition, we introduce the 
+  concept of fixed vs. free connection edges to represent the
+  static relationship between variable and model nodes from a single analysis, 
+  vs the flexible relationship between variables from different analyses.
 
@@
   There are specific rules for the usage of these nodes and edges.
-  The driver node and the implicit edge are only allowed to be present in PSGs. All 
+  The driver node and the driven edge are only allowed to be present in PSGs. All 
   other node and edge types can be present in any of the three graphs, subject to these restrictions: 
@@
   The degree of a given node is only a function of the connection edges 
-  attached to it. The number of implicit edges is not relevant since any number 
+  attached to it. The number of driven edges is not relevant since any number 
   of drivers could be involved, at different parts of an solution process. For 
@@
   and a gradient based one is applied second, design variables would have incomming 
-  implicit edges from both optimizers but this would not affect the indegree of those
+  driven edges from both optimizers but this would not affect the indegree of those
   variable nodes.
-  ((DP:I started to remove any reference to the edge type in these definitions but then saw that you needed it.
-The reason I did that was because indegree and outdegree are basic graph theory terms.
-What if the reference to implicit edges is handled with the upper indegree limit, i.e.
-after the FPG is obtained the upper indegree limit for the design variable nodes is relaxed (increased) so that it isn't a problem what is or isn't counted?))
 
@@
 \subsection{Design Variables}
-((DP:I'm thinking we might also need the upper indegree limit to be zero as well, because setting only the lower indegree limit makes the node a design variable if it didn't have any incoming edges.
-It's seems possible that in the MCG a node has incoming edges but the user wants it to be a design variable.
-I can make the change if you agree.))
 For the Sellar problem, there are three design variables, $x_1^*$, $z_1^*$, and $z_2^*$. In Fig. 
@@
 any design variable node would have at least one, but possibly multiple, incomming 
-implicit edges from one or more driver nodes.
+driven edges from one or more driver nodes.
 
@@
   others outputs. In the Sellar problem from Eqn \ref{eqn:simple_fpf} the two 
-  coupling constraints, $G_1$ and $G_2$ provide a reciprical dependence between 
+  coupling constraints, $H_1$ and $H_2$ provide a reciprical dependence between 
   $D_1$ and $D_2$. In a graph, these constraints show up as connection edges between 
   the outputs of each analysis block. The relevant connection edges are highlighted in 
-  Fig. \ref{f:coupling}, where they form a cycle. Cycles in a graph are the characteristic 
+  Fig. \ref{f:coupling}, where they form a cycle. Cycles of connection edges are the characteristic 
   structure that represents coupling.

### 2013-04-19 intro.tex
updated per brian's suggestions

@@
     The Fundamental Problem Formulation (FPF) for any given problem will be constant regardless 
-    of which MDAO framework, optimization algorithm, iterative solver, or solution strategy
-    is used to solve the problem. However, it is possible to formulate a problem
-    with an assumed execution ordering that implies a solution strategy. 
+    of which MDAO framework, optimization algorithm, iterative solver, or execution sequence
+    is used to solve the problem. However, it is very easy to formulate a problem without 
+    any reference to a specific solution strategy, but still impose a a specific execution sequence.
     For example, consider the Sellar test problem \cite{AIAA:sellar}:
@@
     \begin{align}
-        given & \ \ y_1 = D_1(x_1,y_2,z_1,z_2) \notag
-        \\      & \ \ y_2^t = D_2(y_1,z_1,z_2) \notag
+        given & \ \ y_1 = D_1(x_1,\hat{y_2},z_1,z_2) \notag
+        \\      & \ \ y_2 = D_2(y_1,z_1,z_2) \notag
         \\min. &\ \ F(x_1,y_1,y_2,z_2) \notag
-        \\w.r.t. & \ \ x_1,y_1,y_2,z_1,z_2 \notag
-        \\s.t. & \ \ G(y_2,y_2^t) = 0
+        \\w.r.t. & \ \ x_1,y_1,\hat{y_2},z_1,z_2 \notag
+        \\s.t. & \ \ H(y_2,\hat{y_2}) = 0 \notag 
+        \\     & \ \ G_1(y_1) \geq 0 \notag
+        \\     & \ \ G_2(y_2) \geq 0
         \label{eqn:simple}
@@
 
-    $D_1$ and $D_2$ represent analysis tools, and $F$ and $G$ are the objective 
-    and constraint functions respectively. Equation \ref{eqn:simple} makes an 
-    inherent assumption about the solution strategy for the problem. Analysis 
-    $D_1$ outputs $y_1$, which is an input to $D_2$. Hence, $D_1$ should be run 
-    before $D_2$ with with $y_2$ being iterated on to convergence with $y_2^t$. 
-    However, a slightly different formulation is equally valid and still 
-    represents the exact same problem:
+    $D_1$ and $D_2$ represent analysis tools, $F$, $G_1$, and $G_2$ are the objective 
+    and constraint functions respectively, and $H$ is the coupling constraint. 
+    Equation \ref{eqn:simple} makes an inherent assumption about the execution 
+    order for the problem. Analysis $D_1$ outputs $y_1$, which is an input to $D_2$. 
+    Hence, $D_1$ should be run before $D_2$ with with $y_2$ being iterated on 
+    to convergence with $\hat{y_2}$. However, a slightly different formulation 
+    is equally valid and still represents the exact same problem:
 
     \begin{align}
-        given & \ \ y_2 = D_2(y_1,z_1,z_2) \notag
-        \\      & \ \ y_1^t = D_1(x_1,y_2,z_1,z_2) \notag
+        given & \ \ y_2 = D_2(\hat{y_1},z_1,z_2) \notag
+        \\      & \ \ y_1 = D_1(x_1,y_2,z_1,z_2) \notag
         \\min. &\ \ F(x_1,y_1,y_2,z_2) \notag
-        \\w.r.t. & \ \ x_1,y_1,y_2,z_1,z_2 \notag
-        \\s.t. & \ \ G(y_1,y_1^t) = 0
+        \\w.r.t. & \ \ x_1,\hat{y_1},y_2,z_1,z_2 \notag
+        \\s.t. & \ \ H(y_1,\hat{y_1}) = 0 \notag 
+        \\     & \ \ G_1(y_1) \geq 0 \notag
+        \\     & \ \ G_2(y_2) \geq 0
         \label{eqn:simple2}
@@
     Equation \ref{eqn:simple2} differs only slightly from Eq. \ref{eqn:simple}. 
+    Notice that the formulation from Eq. \ref{eqn:simple} includes 
+    $\hat{y_2}$, while Eq. \ref{eqn:simple2} includes $\hat{y_2}$. The result is that
     $D_1$ is now dependent on the output of $D_2$, and $y_1$ will be iterated on to 
-    convergence with $y_1^t$. Now the problem can be solved by running $D_2$ first 
+    convergence with $\hat{y_1}$. Now the problem has to be solved by running $D_2$ first 
     and then $D_1$. Since the formulations in Eqs. \ref{eqn:simple} and 
@@
     \begin{align}
-        given & \ \ D_1.y_1 = D_1(x_1,y_2,z_1,z_2) \notag
-        \\      & \ \ D_2.y_2 = D_2(y_1,z_1,z_2) \notag
+        given & \ \ y_1 = D_1(x_1,\hat{y_2},z_1,z_2) \notag
+        \\      & \ \ y_2 = D_2(\hat{y_1},z_1,z_2) \notag
         \\min. &\ \ F(x_1,y_1,y_2,z_2) \notag
-        \\w.r.t. & \ \ x_1,y_1,y_2,z_1,z_2 \notag
-        \\s.t. & \ \ G_1(D_1.y_1,D_2.y_1) = 0 \notag
-        \\     & \ \ G_2(D_1.y_2,D_2.y_2) = 0
+        \\w.r.t. & \ \ x_1,\hat{y_1},\hat{y_2},z_1,z_2 \notag
+        \\s.t. & \ \ H_1(y_1,\hat{y_1}) = 0 \notag
+        \\     & \ \ H_2(y_2,\hat{y_2}) = 0 \notag 
+        \\     & \ \ G_1(y_1) \geq 0 \notag
+        \\     & \ \ G_2(y_2) \geq 0
         \label{eqn:simple_fpf}
@@
     The formulation in Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} 
-    and Eq. \ref{eqn:simple2} because it has two constraints which both must be 
-    met. The presence of both of these constraints fully decouples the problem so that 
+    and Eq. \ref{eqn:simple2} because it has two coupling constraints which both must be 
+    met. The result is that Eq. \ref{eqn:simple_fpf} required the presence of both 
+    $\hat{y_1}$ and $\hat{y_2}$, which means that $D_1$ and $D_2$ each now operate with 
+    their own local copies of the coupling variables. This fully decouples the problem so that 
     either $D_1$ or $D_2$ could be run first or both could be run simultaneously. 
@@
     formulation\cite{alexandrov2004}. The key feature they identify to allow this decoupling 
-    is the creation of separate variables for each analysis so that a variable 
-    is identified by its analysis and it's name. For example, in Eqn. \ref{eqn:simple}
-    there is only one $y_1$ variable vs Eqn. \ref{eqn:simple_fpf} where 
-    there are two ($D_1.y_1$ vs $D_2.y_1$).
+    is the creation of independent variables for each analysis. Tosserams et. al. 
+    in their development of the $\psi$ language for heirarchical problem composition
+    make the same requirement that all analyses get their own local copies of 
+    variables in order to use them as the fundamental building blocks for 
+    problems\cite{tosserams2010specification}. Martins and Hwang derive the same requirement,
+    from a different perspective, in their work on computing analytic derivatives 
+    for MDAO problems\cite{Martins2012}. They determined that local copies of all variables 
+    allows for derivation of all relevant methods for calculating analytic derivatives 
+    from the basic chain rule.
 
@@
     row operations on the DSM matrix and yields multiple specific problem 
-    formulations which all solve the same FPF. In other words, manipulation of the DSM does not fundamentally alter
-    the problem formulation, which makes DSM an excellent foundation for specifying the FPF itself.
+    formulations which all solve the same FPF.
 
@@
     we know from the FPF in Eq. \ref{eqn:simple_fpf} that the objective, $f$, is dependent on the 
-    output, $m$, of analysis $A$. You could not determine that from the FDT in Fig. \ref{fig:FDT_simple} alone. This missing 
-    information means that, while FDT is very useful for partitioning problems, it is also not sufficient to contain a 
-    complete problem formulation.
+    output, $m$, of analysis $A$. The dependence of $m$ on $A$ is now captured 
+    in FDT in Fig. \ref{fig:FDT_simple} alone. This missing information means 
+    that, while FDT is very useful for partitioning problems, it is also not 
+    sufficient to contain a complete problem formulation.
 
@@
     dependency between multiple analysis tools as well as between analysis tools and
-    objective/constraint functions. One significant conclusion from REMS is 
-    the need for variables that are local to discipline analyses. This allows 
-    all information exchange to be stated as equality constraints, which become 
-    edges in the graph. Additionaly, REMS address the need to be able to 
+    objective/constraint functions. Additionaly, REMS address the need to be able to 
     transition between multiple solution strategies while keeping a constant 
@@
 
-    Lamb and Martins also included objectives and constraint functions as nodes in an Extended 
-    DSM (XDSM) in order to capture a more complete description of solution strategies for MDAO problems\cite{Lambe2012}. 
-    Unlike REMS, XDSM also includes nodes for solvers and optimziers to enable complete definition of 
-    MDAO architectures. Martins uses XDSM to describe 13 different optimziation architectures in 
-    a survey paper that provides a novel classification of the different 
-    techniques \cite{Lambe2011}. With the additional information included in an 
-    XDSM, Lu and Martins applied both ordering and partitioning algorithms on an 
-    MDAO test problem named the Scalable Problem \cite{Lu2012}.
+    Lamb and Martins also included objectives and constraint functions as nodes 
+    in an Extended DSM (XDSM) in order to capture a more complete description 
+    of solution strategies for MDAO problems\cite{Lambe2012}. Unlike REMS, 
+    XDSM also includes nodes for solvers and optimziers to enable complete 
+    definition of MDAO architectures. Martins uses XDSM to describe 13 different 
+    optimziation architectures in a survey paper that provides a novel 
+    classification of the different techniques \cite{Lambe2011}. With the 
+    additional information included in an XDSM, Lu and Martins applied both 
+    ordering and partitioning algorithms on an MDAO test problem named the 
+    Scalable Problem \cite{Lu2012}.
 
@@
 
-    Although XDSM captures part of the functional aspects of FDT, it leaves out the variables themselves.  
-    As a result, all variable information is aggregated 
-    so that for the problem from Eq. \ref{eqn:simple_fpf} you can say that $A$ 
-    depends on $B$ or vice versa, but you can't identify which 
-    individual variables are interacting in the dependency cycle. 
-    Without the detailed variable information, you can't construct the compatibility 
-    constraints necessary to implement the problem. Additionally, XDSM 
+    Although XDSM captures part of the functional aspects of FDT, it 
     requires the use of solver and optimizer blocks to represent 
@@
 
-    Of all the methods, REMS comes the closest to fully describing a problem formulation,
-    but lacks the ability to represent optimziers and solvers in more specific problem 
-    formulations. While XDSM does include optimizers and constraints, it relies on 
-    their presence and can't be used for a more fundamental problem statement. 
-    We propose a new graph syntax that combines the key features of REMS and XDSM
-    and provides new language to enable more complete description of MDAO problems.
+    Of all the methods, REMS comes the closest to fully describing a fundamental 
+    problem formulation, but lacks the ability to represent optimziers and 
+    solvers in more specific problem formulations. While XDSM does include 
+    optimizers and constraints, it relies on their presence and can't be used 
+    for a more fundamental problem statement. We propose a new graph syntax that 
+    combines the key features of REMS and XDSM while retaining the flexiblity to 
+    represent a fundamental problem formulation free of solution specific information.

### 2013-04-19 example_problem.tex
Dr. German's edits

@@
 	This section presents an example problem to demonstrate the process of obtaining the FPG and the advantages of doing so. 
-The algorithm from Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG} was implemented in the NetworkX package of the Python programming language.
-Any language with the ability to manipulate a graph should be applicable.
-NetworkX was chosen due to the built--in features, such as cycle detection.
+	For this example the algorithm from Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG} was implemented in the NetworkX package of the Python programming language. 
+	While any language with the ability to manipulate a graph should be applicable, NetworkX was chosen due to its built--in features, such as cycle detection and shortest path algorithms.
 
@@
 		\midrule
-		VSP   & parametric geometry \\
+		VSP   & parametric vehicle geometry \\
 		PDCYL & wing and fuselage weight estimation \\
@@
 	FLOPSa denotes FLOPS being executed for both mission performance and engine analysis, while FLOPSb indicates that FLOPS is executed for only mission performance analysis. 
-This sort of representation is useful for ``supercodes'' that care capable of being used in different ways and with different sets of inputs and outputs.
+	This representation is useful for ``supercodes'' that are capable of being used in different ways and with different sets of inputs and outputs.
 
@@
 	Some of the variables, such as geometry and performance, represent many different variables which have been bundled together due to there similarity. 
-	This bundling is not necessary and does not limit the generality of this example, it is done to simplify the presentation.
+	This bundling is not necessary and does not limit the generality of this example; it is done to simplify the presentation.
 	\begin{table}[htb!]
@@
 	  \begin{center}
-		\includegraphics[width=.6\textwidth]{images/FLOPSb_analysis_block}
+		\includegraphics[width=6in]{images/MCG_edit_holes}
 	  \end{center}
@@
 
-	\item Variable ((change)) nodes are created to represent the objectives for performance and total weight.
+	\item Variable nodes are created to represent the objectives for performance and total weight.
 
@@
 
-	\item Explicit edges are created from each variable node to every other variable node representing variables with the same name.
+	\item Connection edges are created from each variable node to every other variable node representing variables with the same name. 
 	The direction is determined by whether the variable node has an edge directed into or out of a model node (local input or local output).
 	\end{enumerate}
-	The resulting MCG is shown in Fig.~\ref{f:design vars}. Emerging cycles are shown in yellow and indicate potential cycles that may exist when an FPG is formed.
+	The resulting MCG is shown in Fig.~\ref{f:MCG}. 
+	Emerging cycles are shown in yellow and indicate potential cycles that may exist when an FPG is formed.
 	\begin{figure}[htb!]
 	  \begin{center}
-		\includegraphics[width=.6\textwidth]{images/MCG}
+		\includegraphics[width=6in]{images/MCG_edit}
 	  \end{center}
 	  \caption{Maximal connectivity graph for the subsonic transport example problem}
+	\label{f:MCG}
 	\end{figure}
@@
 	\label{ss:obtaining an FPG}
-	The process of obtaining an FPG from the given MCG is expected to be an iterative process in which the user applies the algorithm to determine whether or not an FPG can be obtained, and then modifies the indegree limits or the MCG to obtain a satisfactory FPG.
+	The process of obtaining an FPG from the given MCG is expected to be an iterative process in which the user applies the algorithm to determine whether or not an FPG can be obtained, 
+	and then modifies the indegree limits or the MCG to obtain a satisfactory FPG. 
 	For example, after determining the location of a hole, the user can change the indegree limit to make the node a design variable, or the user could change the MCG by adding a new analysis code such that the hole is resolved.
@@
 	As presented in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}, the first step to obtaining an FPG is to detect holes. 
-To start, we set the lower indegree limit for every node to be unity to find any potential holes.
+	To start, we set the lower indegree limit for every node to be unity. 
 	In this case, it turns out that the variable nodes representing the number of engines is a hole for NPSS, FLOPSa, and FLOPSb. 
-This implies than an FPG cannot be obtained from the given MCG, as shown in Fig.~\ref{f:MCG nengines}.
+	This implies that an FPG cannot be obtained from the given MCG, as shown in Fig.~\ref{f:MCG nengines}. 
 	In this figure, the green nodes indicate analysis blocks with a hole, and the cyan nodes indicate analysis blocks that became unused after the invalid analyses were removed ((connections were removed, color edges too)).
@@
 	\end{figure}
-
-	Therefore, we may now create another variable node to represent the number of engines as a global input and also create explicit edges to supply this new input.
-Now this step is repeated with the new MCG and no holes are found.
+	Therefore, we shall now create another variable node to represent the number of engines as a global input and also create connection edges to supply this new input.
 
@@
 	The FPG with the fewest cycles is shown in Fig.~\ref{f:FPG fewest cycles}, which indicates one cycle shown in yellow. 
-A similar alternative would be to minimize the number of analysis blocks involved in cycles, counting multiplicity.
+	A similar alternative would be to minimize the number of analysis blocks involved in cycles, counting multiplicity, or to minimize the length of the longest cycle.
+Since these graphs will always be relatively small (on the order of hundres of nodes) the
+	computational expense for examining all possible FPGs is negligble.
 	\begin{figure}[htb!]
@@
 
-	Alternatively, it may be desirable to resolving collisions by including the higher fidelity codes, such as NPSS in this case.
-	One method to ensure that the edges directed from certain analysis blocks are the ones chosen to resolve the conflicts is to use a ranking system.
+	Alternatively, it may be desirable to resolve collisions by choosing the higher fidelity codes, such as NPSS in this case.
+	One method to ensure that the edges directed from certain analysis blocks are those chosen to resolve conflicts is to use a ranking system. 
 	For this case, consider the rankings for each analysis code given in Table \ref{t:rankings}.
@@
 	In this case, both VORLAX and PMARC are retained, resulting in an FPG that omits only FLOPSa.
-
-%\begin{figure}[htb!]
-%  \begin{center}
-%    \includegraphics[width=.6\textwidth]{images/FPG_highest_rank_simplified}
-%  \end{center}
-%  \caption{Simple notional problem with two potential design variables \label{f:design vars}}
-%\end{figure}
+%\clearpage
+%\newtheorem{name}{Printed output}
+%\begin{name}
+%asdf
+%\end{name}
+%\begin{proof}[asdf]
+%asdf
+%\end{proof}
 

### 2013-04-19 graph_syntax.tex
one last edit
Dr. German's edits

@@
   \label{s:requirements}
-  The ultimate goal of the new graph--based syntax presented here is to be able to 
-  fully describe the general structure of an MDAO problem independently of any solution information, 
+  The goal of the new graph--based syntax presented here is to enable the general structure of an MDAO problem to be described independently of any solution information, 
   while still being able to accomodate the more specific case when a solution 
-  strategy is applied. In order to achieve that goal 
-  the graph needs to accomodate a number of constructs of MDAO problems: 
+  strategy is applied. In order to achieve this goal, 
+  the graph syntax needs to accomodate a number of MDAO problem constructs: 
   \begin{itemize}
-    \item Analysis tools and connections between them
+    \item Analysis tools and their interconnections
     \item Design variables, objectives, and constraints
-    \item Local and global properties
+    \item Local and global information
     \item Coupling between analyses
@@
 
-  Beyond those basic constructs, there are also three phases of a design process that 
-  all need to be representable with the new syntax. Firstly there is the initial problem definition
-  phase where the specific analysis tools and design goals are identified. At the end of this phase, 
-  a single formal problem formulation is selected specifying design variables, constraints, objectives, 
-  analysis tools, etc. Lastly some specific procedure for solving the problem is selected, for example 
-  picking an MDAO optimization architecture. Using the proposed graph syntax, each of these phases 
+The new syntax is intended to represent three phases of the design problem formulation process. 
+In the initial problem definition
+  phase, the specific analysis tools and design goals are identified. Next, 
+  a single formal problem formulation is identified that specifies design variables, constraints, objectives, 
+  analysis tools, and all other elements required to represent the overall MDAO problem. 
+Lastly, a specific procedure for solving the problem is selected, e.g. selecting an MDAO optimization architecture. Using the proposed graph syntax, these phases 
   can be represented with the following three graphs:
@@
     \item Fundamental Problem Graph (FPG)
-    \item Problem Solution Graph (SPG)
+    \item Problem Solution Graph (PSG)
   \end{itemize}
 
-  The \emph{maximal connectivity graph} represents the first phase with all 
+  The \emph{maximal connectivity graph} represents the first phase of the problem formulation with all 
   analysis codes being considered and all possible connections between them also present. The second graph 
   is the \emph{fundamental problem graph}, which is the smallest possible graph 
-  that still fully defines a given problem formulation. Finally, a \emph{specific problem formulation} 
+  that still fully defines a given problem formulation that could in principle be solved by an MDAO technique. Finally, a \emph{specific problem formulation} 
   may be represented by including additional edges and nodes to represent the 
@@
   edges and nodes then the hourglass shape in Fig. \ref{f:hourglass} illustrates how
-  the MCG gets reduced to a single FPG, then multiple possible PSGs exist to solve the problem.
-  In other words the FPG is obtained from the MCG by removing nodes and edges, 
+the FPG is obtained from the MCG by removing nodes and edges, 
   and the PSG is obtained from the FPG by adding nodes and edges.
@@
   \label{s:syntax definition}
-  In this section we present the necessary graph theory fundamentals to 
-  construct the graphs discussed above. 
+  In this section, we present the necessary graph theoretic fundamentals to 
+  construct the graphs discussed in Sec.~\ref{s:requirements}. 
   The notation used in this work is adapted from Diestel \cite{Diestel2010}. 
@@
   contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
-  For a \emph{directed graph*} (or \emph{digraph}) we construct $E$ as a set of ordered pairs instead 
+  For a \emph{directed graph*}((DJP: why the asterisk?)) (or \emph{digraph}) we construct $E$ as a set of ordered pairs instead 
   of a set of sets. Each ordered pair represents an edge starting at the node 
@@
   entry. Edge $e$ = $(x,y)$ may be referred to simply as $xy$. For node $v \in V$ 
-  the edges directed out are given by $E(v)$ and the edges directed into $v$ are given 
-  by $E^{-1}(v)$; $E(E(v))$ is denoted as $E^2(v)$, and likewise for additional levels. 
+  the edges directed out from $v$ are denoted by $E(v)$ and the edges directed into $v$ are given 
+  by $E^{-1}(v)$. 
   If $E$ is not a one--to--one mapping, $E(v)$ may be the empty set, a single node, or a set of nodes.
@@
   V & = & \{1,2,3,4,5\}, \\
-  E & = & \big\{(1,2),(3,2),(1,3),(3,4)\big\}.
+  E & = & \big\{(2,1),(3,2),(1,3),(3,4)\big\}.
   \end{IEEEeqnarray*}
@@
   %does this belong in the analysis block section???? 
-  Let $I$ be a nonempty set of counting numbers such that for each $i \in I$ there is a corresponding set $A_i$. 
+  Let $I$ be a nonempty set such that for each $i \in I$ there is a corresponding set $A_i$. 
   The set of sets $\mathcal A = \{A_i \st i \in I\}$ is called an indexed family of sets with index $i$ and 
@@
   \item[connection edge:] Exchange of information between two nodes. These edges 
-  can either be fixed (can not be removed from graph) or free (can be removed). 
+  can either be fixed (cannot be removed from graph) or free (can be removed). 
   \item [driven edge:] Passing of information from a driver node to a 
@@
 
-  In Alexandrov and Lewis's REMS syntax only two node types were present, variable 
-  and function, and only one edge type\cite{alexandrov2004}. We've chosen to rename the ``function'' node 
-  type to ``model'' to be more consistent with modern MDAO frameworks. The present work 
-  adds one more node types and one more edge type to the syntax to allow descriptive
-  graphs for all three phases of the design process. In addition, we introduce the 
+  In Alexandrov and Lewis's REMS syntax only two node types (variable and function) and one edge type are included \cite{alexandrov2004}. 
+ We have chosen to adopt the terminology ``model'' node instead of ``function'' node to be consistent with modern MDAO framework terminonlogy. The present work 
+  adds one additional node types and one additional edge type to the syntax to allow descriptive
+  graphs for all three phases of the design problem formulation process. In addition, we introduce the 
   concept of fixed vs. free connection edges to represent the
@@
   \label{ss:rules}
-  There are specific rules for the usage of these nodes and edges.
-  The driver node and the driven edge are only allowed to be present in PSGs. All 
-  other node and edge types can be present in any of the three graphs, subject to these restrictions: 
+  A rule set is provided for the usage of these nodes and edges.
+  The driver node and the driven edge are allowed to be present only in PSGs.
+  Other node and edge types can be present in any of the three graph types, subject to the following restrictions: 
   \begin{enumerate}
-  \item A model node can only have one edge directed to or from another model node.
-  \item A model node can only have fixed connection edges directed in or out.
+  \item A model node may have only one edge directed to or from another model node.
+  \item A model node may have only fixed connection edges directed in or out.
   \item A model node must have at least one edge directed in and at least one edge 
     directed out.
-  \item If a variable node has an outgoing edge to a model node then it may not have 
-    any other outgoing edges.
+  \item If a variable node has an outgoing edge to a model node, then it may not have other outgoing edges.
   \end{enumerate}
@@
   is the number of connection edges directed out and it is denoted as $\txt{deg}^+(v)$.
-  The degree of a given node is only a function of the connection edges 
-  attached to it. The number of driven edges is not relevant since any number 
-  of drivers could be involved, at different parts of an solution process. For 
-  example in a sequential optimization where a global optimizer is first applied
-  and a gradient based one is applied second, design variables would have incomming 
+  The degree of a given node is a function only of the connection edges 
+  attached to it. The number of driven edges is not relevant because any number 
+  of drivers could be involved in different parts of an solution process. For 
+  example, in a sequential optimization in which a global optimizer is applied first ((BG: first and second?))
+  and a gradient based optimizer is applied second, design variables would have incomming 
   driven edges from both optimizers but this would not affect the indegree of those
@@
 
-  We also define the \emph{upper indegree limit}
+  Specifially for variable nodes we also define the \emph{upper indegree limit}
 
@@
   node $v$ with $\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, $v$ 
-  must have exactly one incomming explicit edge or the graph is invalid. 
+  must have exactly one incomming explicit edge or the graph is deemed invalid. 
   Two possiblities exist for voilating these limit conditions:
@@
       \begin{equation} \txt{deg}^-(v) < \txt{deg}_l^-(v) \label{e:hole} \end{equation}
-      This represents a lack of information being supplied to the node.
+      A hole represents a lack of information being supplied to the node.
     \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
@@
 
-  An algorithm for detecting and managing these violations if proposed in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}.
+  The presence of holes and collisions in a graph represent issues that will give
+  rise to an invalid problem formulation. An algorithm for detecting and 
+  managing these violations if proposed in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}.
 

### 2013-04-19 intro.tex
Dr. German's edits

@@
     The number of analysis tools required in multidisciplinary design optimization (MDO) studies is growing
-    in parallel with the increasing scope of typical MDO problems. An example of this growth in scope can
+    in parallel with the increasing scope of typical problems. An example of this growth in scope can
     be observed in the historical evolution of the disciplines involved in MDO problems in aircraft design.
@@
     problems for aircraft design also often include tools for aircraft noise and emissions. In sum, it is now
-    commonplace for 5-10 analysis tools to be employed for typical aircraft design optimization studies [[REF]].
+    commonplace for 5--10 analysis tools to be employed for typical aircraft design optimization studies [[REF]].
 
@@
 
-    As the size and complexity of engineering systems grow, the time and expense for setting up 
-    analysis models grow with them. Multidisciplinary Design Analysis and Optimization (MDAO)
-    frameworks such as OpenMDAO\cite{Gray2012} and ModelCenter have enabled a new level of analysis tool integration 
-    and paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. That 
-    new capability has created a new challenge. Even for models with 10's of analyses there could be hundreds or thousands
-    of variables that need to be linked together. Commonly different analyses will provide competing values for the 
-    same physical quantity, and these conflicts need to be resolved somehow. To make matters worse, the competeing values 
+    The time and expense for setting up 
+    analysis models is growing in conjunction with the increasing scope of MDO problems. 
+    Multidisciplinary Design Analysis and Optimization (MDAO)
+    frameworks such as OpenMDAO\cite{Gray2012} and ModelCenter ((BG: TM?)) have enabled a new level of analysis tool integration 
+    and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. That 
+    new capability has created a new challenge. Even models with tens of analyses could include hundreds or thousands
+    of variables that are interdependent and must be linked in the framework. 
+    A common situation is that different analyses provide differing values for the 
+    same physical quantity, and these conflicts need to be resolved. These differing values 
     may not even be in the same format. These occurrences are particularly acute when analysis tools have differing fidelities. For
@@
 
-    With all the added complexity from larger models it is not hard to imagine that the task of combining all the analyses into a 
+    With the added complexity from larger models, it is plausible that the task of combining all the analyses into a 
     consistent system model capable of solving a relevant engineering design 
-    problem could easily become much more costly than createing the discipline 
-    analyses themselves. Just as for a large system the couplings between the 
-    disciplines begin to dominate the design space, the couplings between the 
-    analyses begin to dominate the job of setting up the model. It is this problem of 
+    problem could approach the cost and time requirements of creating any of the discipline 
+    analyses themselves. For these large scale MDO problems, the couplings between the 
+    analyses begin to dominate the effort required in setting up the model. It is this problem of 
     determining sets of analysis tools and their interconnectivities to form realistic
     multidisciplinary problems that is the subject of this paper. We are motivated by the following notional
-    but real problem of organizing an MDO study for a complex system:
+    but realistic problem of organizing an MDO study for a complex system:
 
+    \begin{quote}
     A new system is being designed for which there is little or no historical precedent. The system
-    is complex, as measured by the number of disciplines and/or components involved in the
-    analysis. A general optimization problem statement has been formulated based on system-
-    level objectives and constraints; however, it is unclear which engineering analysis tools should
-    be interconnected in order to solve the optimization problem. A team of disciplinary and/
-    or component design engineers has been formed in which each engineer has expertise in a
+    is complex, as measured by the number of coupled disciplines and/or components involved in the
+    analysis. A general optimization problem statement has been formulated based on system--level objectives and constraints; however, it is unclear which engineering analysis tools should
+    be interconnected in order to solve the optimization problem. A team of disciplinary and/or component design engineers has been formed in which each engineer has expertise in a
     particular analysis tool or component model. The engineers meet to discuss the approach to
-    interconnecting their tools to achieve the required system-level MDO model.
+    interconnecting their tools to achieve the required system--level MDO model.
+    \end{quote}
 
@@
     interconnections between the analysis tools proposed by the engineers. Graph operations are then
-    conducted reduce the MCG down to a \emph{fundamental problem graphs (FPG)} that describes the minimum set of analysis
-    tools needed to solve the specified system-level design problem. The FPG does not predispose
+    conducted to reduce the MCG down to a \emph{fundamental problem graph (FPG)} that describes the minimum set of analysis
+    tools needed to solve the specified system--level design problem. ((DJP: The minimum set will be an FPG but not all FPGs are the minimum set)) The FPG does not predispose
     any particular solution procedure for the problem; any relevant MDO solution architecture, e.g. MDF,
-    IDF, CO, BLISS, could be selected \emph{post facto} to implement the system-level optimization by
+    IDF, CO, BLISS, could be selected \emph{post facto} to implement the system--level optimization by
     constructing a \emph{problem solution graph (PSG)} and the corresponding problem solution approach.
@@
     The paper is organized as follows. First, we describe the differences between a fundamental problem
-    formulation, which is based only on the system-level optimization problem statement that the
+    formulation, which is based only on the system--level optimization problem statement that the
     engineers desire to solve and on the available analysis tools, and a specific problem formulation, which
-    additionally presumes a specific solution to the problem. Next, we survey the literature in applying
+    additionally presumes a specific MDO solution approach to the problem. Next, we survey the literature in applying
     graph theoretic and formal language approaches to multidisciplinary design problem formulation. 
@@
 	\label{s:specific vs fundamental}
-    The Fundamental Problem Formulation (FPF) for any given problem will be constant regardless 
-    of which MDAO framework, optimization algorithm, iterative solver, or execution sequence
-    is used to solve the problem. However, it is very easy to formulate a problem without 
+    The Fundamental Problem Formulation (FPF) is our terminology for a statement of the overall system--level optimization problem that does not depend on the choice of solution--specifc elements such as the MDAO framework, optimization algorithm, iterative solver, or execution sequence
+    used to solve the problem. However, it is very easy to formulate a problem without 
     any reference to a specific solution strategy, but still impose a a specific execution sequence.
     For example, consider the Sellar test problem \cite{AIAA:sellar}:
+
+((can we either make the $t$ a subscript or wrap parenthesis around it, i.e. $y^{(t)}$?))
 
@@
     and then $D_1$. Since the formulations in Eqs. \ref{eqn:simple} and 
-    \ref{eqn:simple2} both describe the same problem then neither can be the
-    FPF. They are both specific versions of a more fundamental description of 
-    the problem that is common between them. We present the fundamental 
-    formulation as follows:
+    \ref{eqn:simple2} describe the same problem then neither can be the
+    FPF. They are specific alterative expressions of a more fundamental problem description 
+    that we call the FPF:
 
@@
 	\label{s:existing syntax}
-    As shown above, the mathematical language for specifying problem formulations is very general and can be used both for 
-    fundamental and specific problem formulations. Tedford and Martins used the above mathematical syntax to specify the 
+    As indicated in the examples in Sec.~\ref{s:specific vs fundamental}, the mathematical language for specifying problem formulations is very general and can be used both for 
+    fundamental and specific problem formulations. Tedford and Martins used this mathematical syntax to specify the 
     FPF for a set of test problems and also to describe specific formulations for solving them with a 
-    number of optimization architectures\cite{Tedford2009}. Their work demonstrates clearly how multiple specific 
-    problem formulations can all relate back to a common FPG. The challenge with using this 
+    number of optimization architectures\cite{Tedford2009}. Their work demonstrates how multiple specific 
+    problem formulations can relate to a common FPG. The challenge with this 
     traditional mathematical syntax is that it is not easily manipulated or analyzed.
-    A number of graph-based methods have been used successfully to translate the 
+
+    A number of graph--based methods have been used successfully to translate the 
     mathematical syntax into a more useful computational form. 
-
-    Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools where off 
-    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Since a DSM describes a square adjacency matrix, 
+    Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools.
+    Off--diagonal elements of the matrix indicate coupling\cite{Steward1981}. Since a DSM describes a square adjacency matrix, 
     it can be represented in an equivalent directed graph where nodes represent analysis tools and 
     edges represent information dependence between those tools. The ordering of elements in a DSM can be used to indicate 
-    execution order.  For more complex problems, choosing the proper order to run analysis tools is a non-trival task. 
-    Rogers et. al developed DeMAID to manipulate a DSM to find an ordering for analysis tools that 
+    execution order.  For more complex problems, choosing the proper order to run analysis tools is a non--trival task. 
+    Rogers et al.~developed DeMAID to manipulate a DSM to find an ordering for analysis tools that 
     reduces the cost of solving highly coupled systems\cite{rogers1996,rogers1996demaid}. This re-ordering is done through 
@@
 
-    Despite it's attractive properties, a DSM by itself is insufficient to describe complete problem formulations. 
-    Traditional DSM only captures information about data dependency between analyses. 
-    Objective and constraint information is missing from the description of the problem. 
-    An alternative matrix-based syntax, called a Functional Dependency Table (FDT), was proposed by Wagner and Papalambros. 
+    A DSM by itself is insufficient to describe complete problem formulations because it captures only information about data dependency between analyses; objective and constraint information is missing from the DSM description of the problem.
+    An alternative matrix--based syntax, called a Functional Dependency Table (FDT), was proposed by Wagner and Papalambros. 
     FDT represents the relationship between functions, including objectives and constraints, and specific variables that affect 
     them\cite{Wagner1993}. Similar to DSM, FDT also describes an adjacency matrix of a graph. Unlike the DSM graph, 
-    however, the graph is undirected and nodes can represent analysis tools, objectives, 
+    however, the FDT graph is undirected and nodes can represent analysis tools, objectives, 
     or constraints. Edges between nodes represent a dependence on the same 
@@
     more efficient optimization problem decompositions\cite{Michelena1997}. While FDT succeeds at capturing the 
-    information about objectives and constraints, it can not capture the coupled data dependency that DSM captures. For instance, 
-    we know from the FPF in Eq. \ref{eqn:simple_fpf} that the objective, $f$, is dependent on the 
-    output, $m$, of analysis $A$. The dependence of $m$ on $A$ is now captured 
-    in FDT in Fig. \ref{fig:FDT_simple} alone. This missing information means 
+    information about objectives and constraints, its lack of directed edges implies that it cannot capture the coupled data dependency that ((DJP: \emph{a}?)) DSM captures. For example, 
+    the objective, $f$, from the FPF in Eq. \ref{eqn:simple_fpf} is dependent on 
+    output $m$ of analysis $A$. The dependence of $m$ on $A$ is now captured 
+    in FDT in Fig. \ref{fig:FDT_simple} alone. This missing information implies 
     that, while FDT is very useful for partitioning problems, it is also not 
@@
     FDT and DSM\cite{alexandrov2004}. REMS retains the square adjacency 
-    matrix form from DSM, but by adding in the new elements it partially 
-    combined a traditional DSM with an FDT. This allows REMS to represent data 
+    matrix form from DSM, but by adding the objectives and constraints, it partially 
+    combines a traditional DSM with an FDT. This formulation allows REMS to represent data 
     dependency between multiple analysis tools as well as between analysis tools and
@@
     transition between multiple solution strategies while keeping a constant 
-    graph representation of the fundamental problem formulation. REMS only provides
-    syntax for variables and functions, but does not provide for inclusion 
+    graph representation of the fundamental problem formulation. REMS provides
+    syntax for variables and functions but does not facilitate inclusion 
     of solvers or optimizers in the graph.
@@
     additional information included in an XDSM, Lu and Martins applied both 
-    ordering and partitioning algorithms on an MDAO test problem named the 
+    ordering and partitioning algorithms on an MDAO test problem termed the 
     Scalable Problem \cite{Lu2012}.
@@
         \includegraphics[height=.25\textheight]{XDSM/simple}
-        \caption{XDSM for Eq. \ref{eqn:simple}, with a gauss-siedel iteration and MDF solution architecture. \label{fig:XDSM_simple}}
+        \caption{XDSM for Eq. \ref{eqn:simple}, with a Gauss-Seidel iteration and MDF solution architecture. \label{fig:XDSM_simple}}
         \end{center}
@@
 
-    Although XDSM captures part of the functional aspects of FDT, it 
+    Although XDSM captures many of the functional aspects of FDT, it 
     requires the use of solver and optimizer blocks to represent 
     the relationship between design variables and objectives/constraints. 
-    By introducing solver or optimizer blocks XDSM automatically provides
-    some kind of solution strategy. The XDSM for Eq. \ref{eqn:simple} is 
+    By introducing solver or optimizer blocks, XDSM automatically implies a strategy. 
+The XDSM for Eq. \ref{eqn:simple} is 
     given in Figure \ref{fig:XDSM_simple}. This diagram is shown with an 
-    assumed gauss-siedel iteration scheme and a MDF solution architecture. 
-    Hence XDSM is too specific for use with a fundamental problem formulation.
+    assumed Gauss-Seidel iteration scheme and an MDF solution architecture. 
+    Hence, the XDSM syntax description is too specific for use with a fundamental problem formulation.
 
-    Of all the methods, REMS comes the closest to fully describing a fundamental 
-    problem formulation, but lacks the ability to represent optimziers and 
+    Of these methods, REMS can describe most aspects of a fundamental 
+    problem formulation, but lacks the ability to represent optimizers and 
     solvers in more specific problem formulations. While XDSM does include 
-    optimizers and constraints, it relies on their presence and can't be used 
+    optimizers and constraints, it relies on their presence is not formulated  
     for a more fundamental problem statement. We propose a new graph syntax that 

### 2013-04-19 graph_syntax.tex
fixing a few last things

@@
   contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
-  For a \emph{directed graph*}((DJP: why the asterisk?)) (or \emph{digraph}) we construct $E$ as a set of ordered pairs instead 
+  For a \emph{directed graph} (or \emph{digraph}) we construct $E$ as a set of ordered pairs instead 
   of a set of sets. Each ordered pair represents an edge starting at the node 
@@
   of drivers could be involved in different parts of an solution process. For 
-  example, in a sequential optimization in which a global optimizer is applied first ((BG: first and second?))
-  and a gradient based optimizer is applied second, design variables would have incomming 
-  driven edges from both optimizers but this would not affect the indegree of those
-  variable nodes.
+  example in a sequential optimization in which a global optimizer is used first 
+  followed by a gadient based optimizer second to refine global optimizer 
+  result, design variables would have incomming driven edges from both optimizers. 
+  These driven edges would not affect the indegree of those variable nodes.
 

### 2013-04-19 intro.tex
fixing a few last things

@@
     Multidisciplinary Design Analysis and Optimization (MDAO)
-    frameworks such as OpenMDAO\cite{Gray2012} and ModelCenter ((BG: TM?)) have enabled a new level of analysis tool integration 
+    frameworks such as OpenMDAO\cite{Gray2012} and \textregistered{ModelCenter} have enabled a new level of analysis tool integration 
     and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. That 
@@
     more efficient optimization problem decompositions\cite{Michelena1997}. While FDT succeeds at capturing the 
-    information about objectives and constraints, its lack of directed edges implies that it cannot capture the coupled data dependency that ((DJP: \emph{a}?)) DSM captures. For example, 
+    information about objectives and constraints, its lack of directed edges 
+    implies that it cannot capture the coupled data dependency that a DSM captures. For example, 
     the objective, $f$, from the FPF in Eq. \ref{eqn:simple_fpf} is dependent on 

### 2013-04-22 graph_syntax.tex
adding in brians updates to section 6

@@
 \label{s:graph representation}
-In this section we will make use of the Sellar problem, described in 
-Eqn. \ref{eqn:simple_fpf}. The graph representation of the Sellar problem, 
-using the new stynax, is presented in Fig. \ref{f:sellar_graph_full}. Throughout
-the rest of this section, different sub-graphs or slight modifications of 
-the full graph will be used to demonstrate the important MDAO constructs within
-this graph syntax.
+In this section we present our graph syntax applied to the description of MDAO 
+Problems. We make use of the Sellar problem, described in 
+Eqn. \ref{eqn:simple_fpf} as an example to illustrate the implementation of the syntax. 
+The graph representation of the Sellar problem, using the new stynax, is 
+presented in Fig. \ref{f:sellar_graph_full}. In this section sub-graphs or 
+slight modifications of the full graph will be used to demonstrate the 
+important MDAO constructs described by this graph syntax.
 
@@
 \label{ss:analysis blocks and connections}
-Analysis tools take in a number input variables and then perform some work to calculate 
-the values for their respective outputs. In an MDAO graph, this process is 
-represented by a group of nodes and edges called an \emph{analysis block}, 
-shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
+Analysis tools take in a number input variables and then calculate 
+the values for their respective outputs. We represent this process
+by a group of nodes and edges called an \emph{analysis block}, 
+shown in Fig. \ref{f:analysis block}. Within an analysis block, each variable 
 node represents a single input or output and is connected 
 to a single model node via a fixed edge. Note that in Fig.~\ref{f:analysis block}, 
-the analysis block contains two model nodes, with a single edge connecting them. 
+the analysis block includes two model nodes, with a single edge connecting them. 
 This edge represents the necessary calculations to map given inputs 
-to proper output values. If constructing a weighted graph, the weight calculation edge 
-provides the opportunity to computational cost. However, it is also allowed to 
-skip this edge and have both incomming and outgoing edges to variables through a 
-single model node for a given analysis block. Although analysis blocks are 
-comprised of multiple nodes and edges, since all the edges within them are 
-fixed, they become fixed sub-graphs within a graph. In this way, they cannot be 
-altered but they can be removed as a group.
+to the outputs of the anlysis. This calculation edge provides the opportunity 
+to encode computational cost or other characteristics of the analysis code as 
+a weight in a weighted graph formulation. If associating a weight is 
+unnecessary, the calculation edge may be omitted and all input and output 
+variables can be connected to a single model node for a given analysis block. 
+Since all the edges within analysis blocks are fixed, the blocks themselves are 
+fixed sub-graphs within the overall MDAO problem graph. The connectivity of 
+nodes and edges in an analysis block cannot be altered for the purposes of MDAO 
+problem formulation; however, analysis block sub-graphs can be added or removed 
+from the MDAO problem graph as needed.
 
@@
 an output by the manner in which they are connected. As shown in Fig.~\ref{f:analysis block}, 
-inputs are any variable nodes that have an outgoing edge into a model 
-node. Conversely, outputs are any variable nodes that have an incoming edge from a model node.
+inputs are represented as variable nodes that have an outgoing edge into a model 
+node. Conversely, outputs represented as any variable nodes that have an incoming 
+edge from a model node.
 
 MDAO problems require that information be passed between sets of analyses. When 
-information from the output of one analysis block is passed, or connected, to the 
-input of another analysis block, a new connection edge is added connecting the two 
-variable nodes involved in the exchange. These connection edges are free, unlike the edges 
-within an analysis block, can be added or removed depending on the needs
+information from one output of an analysis block is passed, or connected, to an 
+input of another analysis block, a new connection edge is added between the two 
+variable nodes involved in the exchange. These connection edges are free, and unlike the edges 
+within an analysis block, the edges can be added or removed depending on the needs
 of a given design problem. In other words, free connection edges are created by 
-engineers linking the output of one tool to the input of another one. Note that 
-it is allowed for a single output to have outgoing connection edges to multiple 
+engineers linking an output of one tool to an input of another one. It is 
+allowable for a single output to have outgoing connection edges to multiple 
 downstream inputs.
@@
 \subsection{Design Variables}
-For the Sellar problem, there are three design variables, $x_1^*$, $z_1^*$, and $z_2^*$. In Fig. 
-\ref{f:sellar_graph_full}, these are the only three holes (nodes without out 
-incomming edges) in the graph.
+Design variables are the input variables that are free to be 
+changed by the designer or by an optimizer in an MDAO study. In a graph design
+variables will be a set of variable nodes, $V_{dv}$, identified as follows:
+
+\begin{equation}
+  V_{dv} = \{V \st \txt{deg}_l^-(v)=0\}
+\end{equation}
+
+For the Sellar problem, there are three design variables: $x_1^*$, $z_1^*$, and 
+$z_2^*$. In Fig. \ref{f:designvars} all variable from the Sellar problem are 
+annotated with their lower indegree limit and the three design variables are identified 
+in bold.
 
@@
   \begin{center}
-  [NOTE: Re-work this one to match $D_2$ from Sellar]
+    [Graph Here....]
     \includegraphics[width=.6\textwidth]{images/design_vars_graph}
   \end{center}
-  \caption{Notional graph with two potential design variables \label{f:designvars}}
+  \caption{Sellar problem graph with lower indegree limits for all variables \label{f:designvars}}
 \end{figure}
 
-More formally stated, given a MCG with the $\txt{deg}_l^-(v)=1$ for all nodes, 
-there will likely be a set of holes preventing a valid FPG from being obtained. 
-Each hole represents a node that could potentially become a design variable. 
-It is up to the designer to examine each and decide if it is appropriate to 
-consider it a design variable, in which case the designer shall set the lower 
-indegree limit to be zero. However, some variable holes may not actually be 
-suitable as a design variable. For instance, if an aircraft mission analysis 
-code requires a drag input which ends up being a hole then it's likely best to 
-leave$\txt{deg}_l^-(v)=1$ and find a new analysis code to fill the hole in graph.
+In Fig. \ref{f:designvars}, none of the design variable nodes have any incomming 
+edges. In general, in a MCG or a FPG, design variables will 
+show up as nodes without any incomming edges. For a valid PSG, design variables 
+could have one or more incomming driven edges representing a value determined by 
+a optimizer.
 
-So a design variable is defined as any variable node with $\txt{deg}_l^-(v)=0$. 
-This, by definition, means that design variables are not holes in the graph, since 
-they have not violated their lower indegree limit. Regardless, 
-when working with a MCG or FPG, only connection edges are allowed, so a design 
-variable in either graph will be a variable with no incomming edges edges. For a valid PSG 
-any design variable node would have at least one, but possibly multiple, incomming 
-driven edges from one or more driver nodes.
+Note the although there are no incomming edges, design variables are not holes 
+in the graph as defined by Eq. \ref{e:hole}. Since they have $\txt{deg}_l^-(v)=0$ 
+having no incomming connection edges does not violate their lower indegree limit. 
+However, the concept of holes can be useful for identifying potential design variables
+given an arbitrary MCG. If you first set $\txt{deg}_l^-(v)=1$ for all variable nodes, 
+and then find the resulint holes, you will have a list of variable nodes that 
+could potentially be design variables. Some variables will be true design variables, 
+while others might represent missing information in the problem formulation. For 
+instance, if there is a hole for a drag input of an aircraft mission analysis code 
+then this is not likely a good choice for a design variable. So it is up to the designer to 
+examine each hole and decide if it's an appropriate design variable. If 
+the variable should become a design variable then set $\txt{deg}_l^-(v)=0$. 
+If a variable should not be a design variable, then leave it as a hole which 
+needs to be addressed in a manner discussed in section \ref{ss:obtaining FPG}.
 
-\subsection{Objectives, and Constraints}
+\subsection{Objectives and Constraints}
 \label{ss:objectives and constraints}
-Like design variables, objectives and constraints are also constructs that need 
-to be identified when setting up a given design problem. In the case of 
-objective functions single output values could be selected, but commonly multiple 
-values are assembeled together via simple algebraic expressions to form a composite 
-objective function. (e.g. The objective function from Eqn. \ref{eqn:simple_fpf} is
-$x_1^2+z_2+y_1+e^{-y_2}$). Constraints are always given in the 
-form of either an inequality or an equality. (e.g. a constraint from Eqn. 
-\ref{eqn:simple_fpf} is $\frac{y_1}{3.16}-1\geq0$.) By convention, 
-constraints are usually given such that some expression will be less than or equal 
-to zero, so any positive value would violate the constraint. For both objective 
-and constraints, these simple expressions take some inputs and map them to an
-output value of significance to the overall design problem.
+Objectives and constraints are constructs that need to be identified when 
+describing and MDAO problem. In the case of objective functions a single output 
+value generated by an analysis block could be selected, but commonly multiple 
+values are assembeled together to form a composite objective function. 
+Generally both objective and constraint functions accept a set of inputs and map 
+them to an output value of significance to the overall design problem.
 
-These operations, though very simple, are effectively the same kind of task 
-performed by an analysis block. A constraint or objective function can therfore 
-be represented as another analysis block on the graph, with it's own input and 
+The operations of implementing composite objective and constraint functions, 
+although typically simple, are effectively the identical to the task performed by an 
+analysis block. A composit constraint or objective function can therfore 
+be represented as an additional analysis block within the graph, with its own input and 
 output variable nodes. Although fundamentally no different than an analysis block, 
-for clarity and convience it is useful to distinguish between regular analysis 
-blocks and those that arise from the addition of objectives or constraints to 
-the graph. Therefore, we define an \emph{expression block}, as the collection of variable and model 
-nodes related to a given objective or constraint function. Fig. \ref{f:obj-cons}
-highlights the objective and constraint blocks from the Sellar problem.
+for clarity and convience, it is useful to distinguish between analysis 
+blocks corresponding to analysis codes and those that arise from the addition of 
+objectives or constraints to the graph. Therefore, we define an 
+\emph{expression block}, as the collection of variable and model nodes related 
+to a given objective or constraint function. Fig. \ref{f:obj-cons} highlights 
+the objective and constraint blocks from the Sellar problem. If the relevant 
+objective and constraint values are direct results from other analysis blocks 
+and composite functions are not needed, then expression blocks may be omitted 
+from the MDAO problem graph.
 
@@
 
-  A simple definition of a global node is any node that involves information 
-  from more than one analysis block. Likewise, a local node is any node 
-  that involves information from only a single analysis block. As stated before, variable 
-  nodes with fixed edges directed to model nodes (as part of an analysis block) are inherently local 
-  and similarly model nodes are also local. Global variables are a part of many 
-  MDAO problems and must be representable in the graph. Follwoing our simple definition
-  any variable node that had multiple incomming or outgoing edges, would be global. 
-  In Fig. \ref{f:sellar_graph_full} the variable nodes for $x_1^*$, $z_1^*$, and $z_2^*$ 
-  have multiple outgoing edges and would be considered global.
+  The set of local nodes, $V_l$, is the collection of 
+  nodes with exactly one outgoing connection edge.
+
+  \begin{equation}
+    V_{l} = \{V \st \txt{deg}^+(v)=1\}
+    \label{e:local_definition}
+  \end{equation}
+
+  In section \ref{s:specific vs fundamental} variable nodes that are part of 
+  an analysis block are inherently local and similarly model nodes 
+  are also local. Following the definition in Eq. \ref{e:local_definition}
+  we can now see why. Variable nodes connected to model nodes are only allowed to 
+  have one outgoing edge, and hence will always by definition be local variables.
+
+  The set of global nodes, $V_g$, is the collection of all nodes with more than 
+  one outgoing connection edge.
+
+  \begin{equation}
+    V_{g} = \{V \st \txt{deg}^+(v)>1\}
+    \label{e:global_definition}
+  \end{equation}
+
+  Follwoing our definition any variable node that had multiple 
+  outgoing edges, is global. In Fig. \ref{f:sellar_graph_full} 
+  the variable nodes for $x_1^*$, $z_1^*$, and $z_2^*$ have multiple outgoing 
+  edges and would be considered global.
 
@@
   arises from the explicit treatment of objectives and constraints as separate 
-  expression blocks. Normally in the Sellar problem $x_1$ is considered a local 
-  variable because it only directly affects analysis block $D_1$. By forcing
-  the expansion of $F$ into an expression block with its own inputs, then $F$ 
-  gets it's own copy of the $x_1$ variable. This necessitates the creation 
-  of the global $x_1^*$ node in the graph with two outgoing edges to link the 
-  two $x_1$ inputs nodes in the different blocks. Interestingly, using this 
-  definition, where $x_1^*$ is a global variable fits nicely within 
-  the structure of Braun's Colaborative Optimization (CO) archtiecture \cite{braun1996thesis}. 
-  The rules of setting up a problem with CO require that each discipline be 
-  given an independent local copy of all global variables. All local 
-  variables are retained uniquely within their respective disciplines, except 
-  when a local design variable shows up explicitely in the objective 
-  or global constraint. In that case a global variable is created, and the discipline again 
-  gets an independent local copy of it. Given our definition of global vs local
-  variable nodes, no such exception for local variables in global expressions is 
-  necessary. When you include a given variable in an expression block, it becomes 
-  global by definition and should be treated as such in CO.
+  expression blocks. Normally in the Sellar problem, $x_1$ is considered a local 
+  variable because it directly affects only analysis block $D_1$. By expanding
+  $F$ into an expression block with its own inputs, then $F$ is assigned its own local 
+  copy of the $x_1$ variable. This formulation necessitates the creation of the global $x_1^*$ 
+  node in the graph with two outgoing edges to link the two $x_1$ inputs nodes 
+  in the different blocks. Interestingly, the classification of $x_1^*$ 
+  as a global variable fits nicely within the structure of the Colaborative 
+  Optimization (CO) archtiecture \cite{braun1996thesis}. The rules for setting up 
+  a problem with CO require that each discipline be given an independent local 
+  copy of all global variables. All local variables are retained uniquely 
+  within their respective disciplines, except when a local design variable appears 
+  explicitely in the objective or global constraint. In that case a global 
+  variable is created, and the analysis block again is again assigned an 
+  independent local copy of it. Given our definition of global vs local variable 
+  nodes, no such exception for local variables in global expressions is 
+  necessary. Whenever a given variable is included in an expression block, 
+  it forces the creation of a global variable node in the graph which will 
+  result in the correct structure for CO automatically.
 
-  Analysis and expression blocks can also have a locality defined for the set of 
-  nodes and edges within the block. Their locality is determined by the 
-  incomming and outgoing edges from the block as a whole. Fig. 
+  Analysis and expression blocks can themselves be local or global, but the 
+  criteria for determining this is slightly different than for variable nodes. 
+  In this case, you consider the set of connection edges for the block
+  (e.g. from all the variable nodes within the block) as a set. Additionally, 
+  you must consider both incoming and outgoing edges, where for variables 
+  only outgoing connection edges were relevant.
+
+  A local analysis block is defined as a block with incomming edges from 
+  exactly one other analysis block and outgoing edges from exactly one other analysis 
+  block.
+
+  A global analysis block is defined as
+
+  Fig. 
   \ref{f:sellar_graph_full} has expression blocks for each constraint, 

### 2013-04-22 graph_syntax.tex
stuff

@@
 Problems. We make use of the Sellar problem, described in 
-Eqn. \ref{eqn:simple_fpf} as an example to illustrate the implementation of the syntax. 
+Eq. \ref{eqn:simple_fpf} as an example to illustrate the implementation of the syntax. 
 The graph representation of the Sellar problem, using the new stynax, is 
@@
 
-  Note that the variable $x_1$ from the Sellar problem is usually refered to as 
+  The variable $x_1$ from the Sellar problem is usually refered to as 
   a local variable, which is contradicted by the given graph. This discrepancy 
@@
 
-  A local analysis block is defined as a block with incomming edges from 
-  exactly one other analysis block and outgoing edges from exactly one other analysis 
-  block.
+  A local analysis block is 
+  defined as a block with incomming edges from exactly one other analysis 
+  block and outgoing edges from exactly one other analysis block. A global 
+  analysis block is defined as a block with incomming edges or outgoing edges 
+  to more than one analysis block.
 
-  A global analysis block is defined as
+  [[MATH HERE]]
 
-  Fig. 
-  \ref{f:sellar_graph_full} has expression blocks for each constraint, 
+  Looking at the Sellar problem in 
+  Fig. \ref{f:sellar_graph_full} we see the expression blocks for each constraint, 
   $G_1$ and $G_2$. Both constraints are local to their respective analysis 
@@
 
-  Simply stating that any node or block with incomming or outgoing edges
-  from more than one other node or block is global works glosses over a more 
-  subtle aspect of some MDAO problems. The locality of any node in a graph is 
+  Catagorizing any node or analysis block block with incomming or outgoing edges
+  from more than one source as global neglects a more 
+  subtle aspect of some MDAO problems. Namley the locality of any node in a graph is 
   a relative property. For instance, a single node might have outgoing edges 
-  to two separate analysis blocks, $A$ and $B$, but none to a third, $C$. Then 
-  you could say that this node was global with respect to $A$ and $B$ separately, but local 
-  to the group $A,B$. This situation produces a natural 
-  hierarchy in a graph, where you could form increasingly smaller groups as you 
-  segment problem into finer localities.
+  to two separate analysis blocks, $A$ and $B$, but none to a third, $C$. This 
+  node was global with respect to $A$ and $B$ separately, but local 
+  to the group $A,B$. This situation produces a natural hierarchy in a graph, 
+  from which one could form ever smaller groups as the problem is segmented into 
+  finer localities.
 
-  When solving a problem using a heirarchical decomposition approach, 
-  there are various techniques for partitioning that are effective in different 
-  scenarios \cite{krishnamachari1997optimal,michelena1997hypergraph,sobieszczanski1997,Perez2004,allison2009optimal}. 
-  Tosserams et al proposed a language based syntax for describing the partitioning of problems, named $\Psi$
-  \cite{tosserams2010specification}. $\Psi$ provides an opposing perspective to 
-  this work, where they compose hierarchies from the bottom up, into larger and 
-  larger groups. However, $\Psi$ provides a compiler that can produce a standard 
-  problem representation of the assembled problem, and a number of example converters
-  from that standard representation to application specific formats.
-
-  [[Could probably build this and include it in appendix]]
+  When solving a problem using a hierarchical decomposition approach, 
+  there are various techniques for partitioning a problem into hierarchical 
+  localities that are effective in different scenarios \cite{krishnamachari1997optimal,michelena1997hypergraph,sobieszczanski1997,Perez2004,allison2009optimal}. 
+  Tosserams \textit{et al.} proposed a language based-syntax for describing the 
+  partitioning of problems, named $\Psi$\cite{tosserams2010specification}. 
+  $\Psi$ provides an opposing perspective to this work, where they compose 
+  hierarchies from the bottom up, into larger and larger groups. However, 
+  $\Psi$ provides a compiler that can produce a standard problem representation 
+  of the assembled problem and a number of example converters from that 
+  standard representation to application specific formats.
 
@@
   Coupling exists in a design problem when a set of two or more analyses each depend on the
-  others outputs. In the Sellar problem from Eqn \ref{eqn:simple_fpf} the two 
-  coupling constraints, $H_1$ and $H_2$ provide a reciprical dependence between 
-  $D_1$ and $D_2$. In a graph, these constraints show up as connection edges between 
+  other's outputs. In the Sellar problem from Eqn \ref{eqn:simple_fpf} the two 
+  coupling constraints, $H_1$ and $H_2$, provide a reciprical dependence between 
+  $D_1$ and $D_2$. In a graph, these constraints appear as connection edges between 
   the outputs of each analysis block. The relevant connection edges are highlighted in 
-  Fig. \ref{f:coupling}, where they form a cycle. Cycles of connection edges are the characteristic 
-  structure that represents coupling.
+  Fig. \ref{f:coupling}, where they form a closed path between two analysis blocks,
+  called a cycle. Cycles of connection edges are the characteristic structure that 
+  represents coupling.
 
@@
 
-  It is important to note that the coupling cycles do not contain 
-  any driver nodes. Solver, optimizer, or other iterative process is 
-  invovled in the coupling definition--though one will be necessary for a valid 
-  PSG. Additionally, a coupling cycle has no inherant start or end. It would be
-  valid to pick any node in the cycle as a starting point, and proceed around the
-  loop untill you get back to your starting point. For the Sellar Problem, picking 
-  $D_1$ as the starting point would yeild a problem as given in 
-  Eqn.~\ref{eqn:simple}, while picking $D_2$ would yield the problem as given in 
-  Eqn.~\ref{eqn:simple2}.
+  Coupling cycles do not contain driver nodes in the MCG of FPG representation. 
+  Solvers, optimizers, and other iterative processes are not invovled in the coupling 
+  definition--though one will be necessary to build a valid PSG from a given FPG. 
+  Additionally, a coupling cycle has no inherent start or end. It would be acceptable to select
+  node in the cycle as a starting point and proceed around the
+  loop until you get back to the starting point. For the Sellar Problem, selecting 
+  $D_1$ as the starting point would yield a problem as given in 
+  Eq.~\ref{eqn:simple}, whereas selecting $D_2$ would yield the problem as given in 
+  Eq.~\ref{eqn:simple2}.
 
   Larger problems can contain more complex cycles in their FPG, indicating more 
-  complex coupling between analyses. A cycle can contain more than 
-  just two analysis blocks. Multiple independent cycles can exist, indicating 
-  independent coupling relationships. Cycles can also overlap, where the same analysis 
+  complex coupling between analyses. For example A cycle can involve more than 
+  just two analysis blocks. Multiple independent cycles could also exist, indicating 
+  independent coupling relationships. Cycles can also overlap, meaning that the same analysis 
   blocks are involved in multiple different coupling cycles. All of these situations
@@
   become difficult. In the present work, Sec.~\ref{s:example problem}.\ref{ss:obtaining an FPG} 
-  will demonstrate how building an FPG from an MCG provides an opportunity to 
+  demonstrates how building an FPG from an MCG provides an opportunity to 
   identify and potentailly reduce the number of cycles in a graph.
 
-  If large amounts of coupling are present it can help convergence 
-  rates to look for an effective ordering for the execution of analyses.
+  If many couplings are present, convergence rates can be improved by 
+  searching for an effective ordering for the execution of analyses.
   Rogers' DEMAID tool uses a genetic algorithm to find an ordering that minimizes 
   the overall coupling of the system by separating independent cycles in the 
-  graph \cite{rogers1996,rogers1996demaid}. Rodgers work focused on the matrix 
-  form of the DSM for ordering optimization. Lu and Martins more recent work 
-  worked with a weighted form of the DSM and used an iterative clustering approach 
-  to perform a similar task to DEMAID \cite{Lu2012}.
-
-  However, when using a graph representation, the ordering problem can be greatly 
-  simplified. For any given coupling cycle, a Gauss-Seidel iteration pattern 
-  is automatically given simply folowing the cycle around the graph. If only 
-  once cycle is present, then the task reduces down to selecting a starting point
-  in the cycle. When sub-cycles are present each sub-cycle ordering is also 
-  given automatically by following the graph, but now a there is freedom at which 
-  level to manage the iteration over the cycles. Ordering of many cycles and 
-  sub-cycles is still a challenge, but the total search space is reduced by 
-  always following the ordering from the graph from a selected starting node.
+  graph \cite{rogers1996,rogers1996demaid}. Rogers work focused on the matrix 
+  form of the DSM for ordering optimization. Lu and Martins more recently leveraged 
+  a weighted form of the DSM and used an iterative clustering approach to perform a 
+  similar task to DEMAID \cite{Lu2012}.
 
@@
   A multi--fidelity problem can be characterized by two or more different analyses 
-  each calculating the same data. Mapping this type of multi--fidelity situation 
-  onto a graph will yield two analysis blocks, each with their own output 
-  variable nodes, having outgoing edges to a single variable in a third analysis 
-  block that needs the data as input. Figure \ref{f:collision_example} shows 
-  a modified version of the Sellar problem with a new analysis, $D_0$, representing a 
-  low fidelity version of $D_1$.
+  each calculating the same data. Representing this type of multi--fidelity 
+  construct in a graph will yield two analysis blocks, each with its own output 
+  variable nodes, having outgoing edges that connect to a single variable in a 
+  third analysis block that needs the data as input. So the set of all multi-fidelity
+  variables, $V_{mf}$, is given as:
+
+  \begin{equation}
+    V_{mf} = \{ V \st \txt{deg}^-(v)>1 \} \cup \{ V \st \txt{deg}^-_u(v)>=\txt{deg}^-(v) \}
+  \end{equation}
+
+  In a multi-fidelity problem, a given variable node may have multiple incomming 
+  edges without causing a collision as defined by by Eq.~\ref{e:collision}. 
+  Fig.~\ref{f:collision_example} shows a modified version of the Sellar problem 
+  with a new analysis, $D_0$, representing a low fidelity version of $D_1$. 
+  In this modified graph, the annotations indicate the upper indegree limit. For 
+  $D_2$, $\txt{deg}^-_u(D_2.y_1) = 2$ which avoids a collsion. A  
+  $\txt{deg}^-_u(v) > 1$ for any variable in a graph represents a decision to 
+  allow multiple fidelities to interact at that part of the graph in order to alieviate
+  a conflict. In section \ref{ss:obtaining FPG} we present an algorithm for finding 
+  conflicts within a given MCG so that a designer can make the necessary decisisons about each 
+  one in turn.
 
@@
     [INSERT GRAPH HERE]
-    \caption{Graph of a notional multi-fidelity problem \label{f:collision_example}}
+    \caption{Graph of the modified Sellar, multi--fidelity problem with $\txt{deg}^-_u(v)$ annotations on variable nodes\label{f:collision_example}}
   \end{center}
@@
 
-  When an input variable node has more than one incoming edge, designers must make
-  some kind of decision about how to manage the flow of information in their problem. 
-  In Eqn.~\ref{e:collision} we defined a collision as the situation where
-  there are more incoming connection edges than the upper indegree limit. So desginers can
-  indicate how to handle conflicting edges in a problem graph by setting the upper indegree limit
-  for a given node. In more concrete terms, setting this limit equates to defining 
-  the part of the graph in question as single or multi--fidelity.
-
-  When $\txt{deg}_u^-(y)=1$, then any conflicting edges in the graph will cause a collision
-  and a choice between the colliding analysis blocks must be made. This results in a 
-  single fidelity design problem. On the other hand if $\txt{deg}_u^-(y)=2$, then two 
-  different analysis blocks can co-exist with a given graph without causing a 
-  collision. The design problem is then a multi-fidelity problem.
-
-  Multi-fidelity problems are always charaterized by graphs with variable 
-  nodes that have an upper-degree-limit greater than one. These problems require 
+  Multi--fidelity problems are always charaterized by graphs with variable 
+  nodes that have an $\txt{deg}^-_u(v) > 1$. These problems require 
   special techniques for resolving the confliciting edges by introducing some mechanism
@@
   run\cite{march2012provably,alexandrov2001approximation,Huang_Allen_Notz_Miller_2006}.
+  The specifics of this mechanism are not given in any form withing an MCG or 
+  FPG. Instead the multi-fidelity mechanism specifics would be represented as a 
+  driver in a PSG derived from any given multi-fidelity FPG.

### 2013-04-22 graph_syntax.tex
finishing the last of the eqn additions

@@
 
-The new syntax is intended to represent three phases of the design problem formulation process. 
-In the initial problem definition
-  phase, the specific analysis tools and design goals are identified. Next, 
-  a single formal problem formulation is identified that specifies design variables, constraints, objectives, 
-  analysis tools, and all other elements required to represent the overall MDAO problem. 
-Lastly, a specific procedure for solving the problem is selected, e.g. selecting an MDAO optimization architecture. Using the proposed graph syntax, these phases 
-  can be represented with the following three graphs:
+  The new syntax is intended to represent three phases of the design problem 
+  formulation process. In the initial problem definition phase, the specific 
+  analysis tools and design goals are identified. Next, a single formal problem 
+  formulation is identified that specifies design variables, constraints, 
+  objectives, analysis tools, and all other elements required to represent 
+  the overall MDAO problem. Lastly, a specific procedure for solving the problem 
+  is selected, e.g. selecting an MDAO optimization architecture. Using 
+  the proposed graph syntax, these phases can be represented with the following 
+  three graphs:
+
   \begin{itemize}
@@
 
-  A \emph{path} $P=(V,E)$ from $x_0$ to $x_k$ in graph $G$ is a subgraph of $G$ with $V = \{x_0,x_1,\ldots,x_k\}$ and $E = \{(x_0,x_1),(x_1,x_2),\ldots,(x_{k-1},x_k)\}$.
-  A \emph{reverse path} $P_R$ in graph $G$ is a path on $R$, where $R$ is the reverse graph of $G$ obtained by switching the orientation of every edge.
+  A \emph{path} $P=(V,E)$ from $x_0$ to $x_k$ in graph $G$ is a subgraph of $G$ 
+  with $V = \{x_0,x_1,\ldots,x_k\}$ and $E = \{(x_0,x_1),(x_1,x_2),\ldots,(x_{k-1},x_k)\}$.
+  A \emph{reverse path} $P_R$ in graph $G$ is a path on $R$, where $R$ is the 
+  reverse graph of $G$ obtained by switching the orientation of every edge.
 
@@
     \item[variable:] represents scalar or array data
+    \begin{equation} 
+      V_{\txt{var}} = \{v \st t_{\txt{node}}(v) = \txt{`variable'}\}
+    \end{equation}
+
     \item[model:] responsible for mapping inputs to outputs
+    \begin{equation} 
+      V_{\txt{model}} = \{v \st t_{\txt{node}}(v) = \txt{`model'}\}
+    \end{equation}
+
     \item[driver:] control logic capable of managing iteration
+    \begin{equation} 
+      V_{\txt{driver}} = \{v \st t_{\txt{node}}(v) = \txt{`driver'}\}
+    \end{equation}
+
   \end{description}
@@
   can either be fixed (cannot be removed from graph) or free (can be removed). 
+    \begin{equation} 
+      E_{\txt{conn}} = \{e \st t_{\txt{edge}}(e) = \txt{`connection'}\}
+    \end{equation}
   \item [driven edge:] Passing of information from a driver node to a 
@@
     free and can be added or removed from the graph.
+     \begin{equation} 
+       E_{\txt{driven}} = \{e \st t_{\txt{edge}}(e) = \txt{`driven'}\}
+     \end{equation} 
   \end{description}
@@
 
-  In Alexandrov and Lewis's REMS syntax only two node types (variable and function) and one edge type are included \cite{alexandrov2004}. 
- We have chosen to adopt the terminology ``model'' node instead of ``function'' node to be consistent with modern MDAO framework terminonlogy. The present work 
-  adds one additional node types and one additional edge type to the syntax to allow descriptive
+  In Alexandrov and Lewis's REMS syntax only two node types (variable and function) 
+  and one edge type are included \cite{alexandrov2004}. We have chosen to adopt 
+  the terminology ``model'' node instead of ``function'' node to be consistent 
+  with modern MDAO framework terminonlogy. The present work adds one additional 
+  node types and one additional edge type to the syntax to allow descriptive
   graphs for all three phases of the design problem formulation process. In addition, we introduce the 
@@
   The driver node and the driven edge are allowed to be present only in PSGs.
-  Other node and edge types can be present in any of the three graph types, subject to the following restrictions: 
+  Other node and edge types can be present in any of the three graph types, 
+  subject to the following restrictions: 
   \begin{enumerate}
@@
 by a group of nodes and edges called an \emph{analysis block}, 
-shown in Fig. \ref{f:analysis block}. Within an analysis block, each variable 
-node represents a single input or output and is connected 
-to a single model node via a fixed edge. Note that in Fig.~\ref{f:analysis block}, 
-the analysis block includes two model nodes, with a single edge connecting them. 
-This edge represents the necessary calculations to map given inputs 
-to the outputs of the anlysis. This calculation edge provides the opportunity 
+shown in Fig. \ref{f:analysis block}. Each analysis block is 
+denoted by:
+
+\begin{equation}
+  A_i=(V_{A_i},E_{A_i}),\ for\ i\in I, 
+\end{equation}
+
+Where $m$ is the number of analysis tools represented in the graph 
+and $I=\{1,2,\ldots,m\}$.
+
+Within a given analysis block, variable node represents a single input or 
+output and is connected to a single model node via a fixed connection edge. Note that 
+in Fig.~\ref{f:analysis block}, the analysis block includes two model nodes, 
+with a single edge connecting them. This edge represents the necessary 
+calculations to map given inputs to the outputs of the anlysis. This calculation edge provides the opportunity 
 to encode computational cost or other characteristics of the analysis code as 
 a weight in a weighted graph formulation. If associating a weight is 
-unnecessary, the calculation edge may be omitted and all input and output 
+unnecessary, the calculation edge and second model node may be omitted and all input and output 
 variables can be connected to a single model node for a given analysis block. 
@@
 \begin{equation}
-  V_{dv} = \{V \st \txt{deg}_l^-(v)=0\}
+  V_{\txt{dv}} = \{V_{\txt{var}} \st \txt{deg}_l^-(v)=0\}
 \end{equation}
@@
 blocks corresponding to analysis codes and those that arise from the addition of 
-objectives or constraints to the graph. Therefore, we define an 
-\emph{expression block}, as the collection of variable and model nodes related 
-to a given objective or constraint function. Fig. \ref{f:obj-cons} highlights 
-the objective and constraint blocks from the Sellar problem. If the relevant 
+objectives or constraints to the graph. Therefore, we denote an \emph{expression block}, as:
+
+\begin{equation}
+  F_j=(V_{F_j},E_{F_j}),\ for\ j\in J, 
+\end{equation}
+
+Where $m$ is the number of expressions in the graph
+and $J=\{1,2,\ldots,m\}$. For the sake of consistency of naming conventions
+$F_j$, $G_j$, and $H_j$ are used to deonte expression blocks for objectives, inequality
+constraints, and equality constraints. If the relevant 
 objective and constraint values are direct results from other analysis blocks 
 and composite functions are not needed, then expression blocks may be omitted 
-from the MDAO problem graph.
+from the MDAO problem graph. Fig. \ref{f:obj-cons} highlights the 
+objective and constraint blocks from the Sellar problem.
 
@@
 
-  The set of local nodes, $V_l$, is the collection of 
+  The set of local nodes, $V_\txt{l}$, is the collection of 
   nodes with exactly one outgoing connection edge.
@@
   \begin{equation}
-    V_{l} = \{V \st \txt{deg}^+(v)=1\}
+    V_{\txt{l}} = \{V_{\txt{var}} \st \txt{deg}^+(v)=1\}
     \label{e:local_definition}
@@
 
-  The set of global nodes, $V_g$, is the collection of all nodes with more than 
+  The set of global nodes, $V_\txt{g}$, is the collection of all nodes with more than 
   one outgoing connection edge.
@@
   \begin{equation}
-    V_{g} = \{V \st \txt{deg}^+(v)>1\}
+    V_{\txt{g}} = \{V_{\txt{var}} \st \txt{deg}^+(v)>1\}
     \label{e:global_definition}
@@
   expression blocks. Normally in the Sellar problem, $x_1$ is considered a local 
-  variable because it directly affects only analysis block $D_1$. By expanding
+  variable because it directly affects only analysis block $A_1$. By expanding
   $F$ into an expression block with its own inputs, then $F$ is assigned its own local 
@@
   $G_1$ and $G_2$. Both constraints are local to their respective analysis 
-  blocks. But the expression block for $F$ has incomming edges from $D_1$ and 
-  $D_2$, so it would considered global.
+  blocks. But the expression block for $F$ has incomming edges from $A_1$ and 
+  $A_2$, so it would considered global.
 
@@
   there are various techniques for partitioning a problem into hierarchical 
-  localities that are effective in different scenarios \cite{krishnamachari1997optimal,michelena1997hypergraph,sobieszczanski1997,Perez2004,allison2009optimal}. 
+  localities that are effective in different scenarios 
+  \cite{krishnamachari1997optimal,michelena1997hypergraph,sobieszczanski1997,Perez2004,allison2009optimal}. 
   Tosserams \textit{et al.} proposed a language based-syntax for describing the 
@@
 \subsection{Coupling Between Analyses}
-
-  Coupling exists in a design problem when a set of two or more analyses each depend on the
+  Coupling is described by the presence of a a closed path between two or more
+  analysis blocks, called a cycle. A path, given by $P=(V_{path},E_{path})$, 
+  contains a cycle if at least two edges in the path refrence the same node, 
+  $E = \{(x_0,x_1),(x_1,x_2),\ldots,(x_{k-1},x_k),(x_k,x_0)\}$. These cycles 
+  decribe the reciprical dependence of a set of two or more analyses on each 
   other's outputs. In the Sellar problem from Eqn \ref{eqn:simple_fpf} the two 
   coupling constraints, $H_1$ and $H_2$, provide a reciprical dependence between 
-  $D_1$ and $D_2$. In a graph, these constraints appear as connection edges between 
+  $A_1$ and $A_2$. In a graph, these constraints appear as connection edges between 
   the outputs of each analysis block. The relevant connection edges are highlighted in 
-  Fig. \ref{f:coupling}, where they form a closed path between two analysis blocks,
-  called a cycle. Cycles of connection edges are the characteristic structure that 
-  represents coupling.
+  Fig. \ref{f:coupling}, where they form
 
@@
   loop until you get back to the starting point. For the Sellar Problem, selecting 
-  $D_1$ as the starting point would yield a problem as given in 
-  Eq.~\ref{eqn:simple}, whereas selecting $D_2$ would yield the problem as given in 
+  $A_1$ as the starting point would yield a problem as given in 
+  Eq.~\ref{eqn:simple}, whereas selecting $A_2$ would yield the problem as given in 
   Eq.~\ref{eqn:simple2}.
@@
   \begin{equation}
-    V_{mf} = \{ V \st \txt{deg}^-(v)>1 \} \cup \{ V \st \txt{deg}^-_u(v)>=\txt{deg}^-(v) \}
+    V_{\txt{mf}} = \{ V_{\txt{var}} \st 1 < \txt{deg}^-(v) \leq \txt{deg}^-_u(v) \}
   \end{equation}
@@
   Fig.~\ref{f:collision_example} shows a modified version of the Sellar problem 
-  with a new analysis, $D_0$, representing a low fidelity version of $D_1$. 
+  with a new analysis, $D_0$, representing a low fidelity version of $A_1$. 
   In this modified graph, the annotations indicate the upper indegree limit. For 
-  $D_2$, $\txt{deg}^-_u(D_2.y_1) = 2$ which avoids a collsion. A  
+  $A_2$, $\txt{deg}^-_u(A_2.y_1) = 2$ which avoids a collsion. A  
   $\txt{deg}^-_u(v) > 1$ for any variable in a graph represents a decision to 

### 2013-04-22 graph_syntax.tex
define t_{edge} and t_{node} functions

@@
   \label{ss:types}
-  We now define different types of nodes and edges with specific properties to compose graphs used to represent an MDAO problem. 
-  There are three node types:  
+  To distinguish between different kinds of nodes, we define the function 
+  $t_{\txt{node}}(v)$ which returns one of three different types of nodes:
+
   \begin{description}
@@
 
-  In addition to the three node types, there are two edge types:
+  In addition we define the function $t_{\txt{edge}}(e)$ which returns one of 
+  two different types of edges:
 

### 2013-05-08 example_problem.tex
Major update

@@
 	This section presents an example problem to demonstrate the process of obtaining the FPG and the advantages of doing so. 
-	For this example the algorithm from Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG} was implemented in the NetworkX package of the Python programming language. 
-	While any language with the ability to manipulate a graph should be applicable, NetworkX was chosen due to its built--in features, such as cycle detection and shortest path algorithms.
+	For this example, the algorithm from Sec.~\ref{ss:obtaining FPG} was implemented in the NetworkX package of the Python programming language. 
+	While any language with the ability to manipulate graphs is applicable, NetworkX was chosen due to its built-in features, such as cycle detection and shortest path algorithms.
 
-	The example task is the conceptual sizing of a single--aisle subsonic transport using an MDAO framework with objectives for performance and total weight for a required mission. 
-	The full set of analysis codes available for use is given in Table \ref{t:analysis codes}.
+	The example task is to create an FPG for the conceptual sizing of a single-aisle subsonic transport using a set of analysis tools with objectives including performance and gross weight for a required mission.
+	The full set of analysis codes available for use is provided in Table \ref{t:analysis codes}.
 	\begin{table}[htbp]
@@
 	\end{table}%
-	Each of these analysis codes contribute individual disciplinary analysis but they are not mutually exclusive. 
+	Each analysis code contributes individual disciplinary analysis capability, but the outputs of the codes are not mutually exclusive. 
 	For example, VORLAX and PMARC are both aerodynamics codes that predict inviscid drag but with different levels of fidelity. 
-	The Flight Optimization System (FLOPS) is included twice to represent two different uses. 
-	FLOPSa denotes FLOPS being executed for both mission performance and engine analysis, while FLOPSb indicates that FLOPS is executed for only mission performance analysis. 
-	This representation is useful for ``supercodes'' that are capable of being used in different ways and with different sets of inputs and outputs.
+	The Flight Optimization System (FLOPS) is included twice to represent two different configurations corresponding to different uses. 
+	FLOPSa denotes FLOPS implemented to execute both mission performance and engine analysis, while FLOPSb indicates  FLOPS implemented to execute for only mission performance analysis. 
+	This representation is useful for ``supercodes'' that are capable of being implemented in different ways and with different sets of inputs and outputs.
+The enumeration of these analysis tools and objectives concludes step (A) in Sec.~\ref{ss:process}.
 
-	In this example, the MCG is formed using a consistant variable naming convention and then connecting every variable with the same name, though this is not the only way to form an MCG. 
-	Table \ref{t:ins and outs} presents the full list of variables in the leftmost column and then indicates whether the variable is an input or an output of each analysis code. 
-	Some of the variables, such as geometry and performance, represent many different variables which have been bundled together due to there similarity. 
-	This bundling is not necessary and does not limit the generality of this example; it is done to simplify the presentation.
+Step (B) is the production of the MCG.
+	In this example, the MCG is formed using a consistant variable naming convention and then connecting all variables with the same name with connection edges. 
+	Table \ref{t:ins and outs} presents the full list of variables in the leftmost column and indicates whether the variable is an input or an output of each analysis code. 
+	Some variables, such as geometry and performance, represent groups of variables that are passed as vectors or other structures due to their similarity. 
+	This bundling of variables is not fundamental and does not limit the generality of this example; rather, it simplifies the presentation.
 	\begin{table}[htb!]
@@
 
-	The MCG $M$ is formed in four steps:
+	In this example, the MCG $M$ is formed in four steps:
 	\begin{enumerate}
 	\item An analysis block is created for each analysis code using the information in each column of Table \ref{t:ins and outs}. 
-	Each analysis block is formed by creating variable nodes for each input and directing them into a single model node, which is directed into another model node, which is then directed into variable nodes corresponding to each output.
-	A sample analysis block is shown for analysis code FLOPSb in Fig.~\ref{f:FLOPSb analysis block}.
+	Each analysis block is formed by first creating variable nodes for each input and adding directed edges into a single model node.
+An edge is directed from this model node into a second model node, which is then directed into variable nodes corresponding to each output.
+	A sample analysis block is shown for analysis code FLOPSa in Fig.~\ref{f:FLOPSb analysis block}.
 	\begin{figure}[htb!]
 	  \begin{center}
-		\includegraphics[width=6in]{images/MCG_edit_holes}
+		\includegraphics[width=2.75in]{images/FLOPSa_analysis_block}
 	  \end{center}
-	  \caption{Sample analysis block for analysis code FLOPSb using Table \ref{t:ins and outs}}
+	  \caption{Sample analysis block for analysis code FLOPSa using Table \ref{t:ins and outs}}
 	\label{f:FLOPSb analysis block}
@@
 
-	\item Variable nodes are created to represent the objectives for performance and total weight.
+	\item Expression blocks are created to represent the objectives for performance and total weight.
 
-	\item Variable nodes are created to represent the geometry variable and mission variable as global inputs.
+	\item A variable nodes is created to represent the geometry variable as a global input. Any global inputs incorrectly omitted will be identified in step (D).
 
-	\item Connection edges are created from each variable node to every other variable node representing variables with the same name. 
-	The direction is determined by whether the variable node has an edge directed into or out of a model node (local input or local output).
+	\item Connection edges are created that connect each variable node to every other variable node representing variables with the same name. 
+	The direction is determined by whether the variable node has an edge directed into or out of a model node, i.e. whether it is a local input or a local output.
 	\end{enumerate}
-	The resulting MCG is shown in Fig.~\ref{f:MCG}. 
-	Emerging cycles are shown in yellow and indicate potential cycles that may exist when an FPG is formed.
+	The resulting MCG is shown in Fig.~\ref{f:MCG holes}. 
+	Cycles are shown as dashed lines. FPGs formed from this MCG may or may not retain these cycles.
 	\begin{figure}[htb!]
 	  \begin{center}
-		\includegraphics[width=6in]{images/MCG_edit}
+		\includegraphics[width=4.5in]{images/MCG_edit_holes}
 	  \end{center}
 	  \caption{Maximal connectivity graph for the subsonic transport example problem}
+	\label{f:MCG holes}
+	\end{figure}
+
+To implement step (C) in Sec.~\ref{ss:process}, the indegree of every variable node is set to unity to reflect a single fidelity analysis for which the design variables have not yet been selected.
+
+%	As described in Sec.~\ref{ss:process}, the process of obtaining an FPG from the given MCG is expected to be an iterative process in which the user applies the algorithm to determine whether or not an FPG can be obtained,
+%	and then modifies the indegree limits or the MCG to obtain a satisfactory FPG.
+%	For example, after determining the location of a hole, the user can change the indegree limit to make the node a design variable, or the user could change the MCG by adding a new analysis code such that the hole is resolved.
+%
+%	As presented in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}, the first step to obtaining an FPG is to detect holes. 
+%	To begin, we set the lower indegree limit for every node to be unity.
+
+Step (D) now procededs by executing the FPG algorithm.
+An FPG cannot be obtained because the variable nodes representing the number of engines are holes for NPSS, FLOPSa, and FLOPSb, and the variable nodes representing the mission definition are holes for FLOPSa and FLOPSb.
+These holes propagated upstream and created holes in the objectives, thereby preventing a valid FPG.
+
+Step (D)(a) begins to resolve this conflict by creating a global input for the number of engines.
+In this example, the number of engines was not originally included as a global input to demonstrate how the process responds to this error.
+Step (D)(b) then reduces the lower and upper indegree limits for the variable node representing the mission definition to zero, implying that these nodes must be specified as design variables (see Table \ref{t:variable node classification}).
+The new MCG is shown in Fig.~\ref{f:MCG}.
+	\begin{figure}[htb!]
+	  \begin{center}
+		\includegraphics[width=4.5in]{images/MCG_edit}
+	  \end{center}
+	  \caption{The updated MCG with a new global input.}
 	\label{f:MCG}
@@
 
-\subsection{Obtaining an FPG}
-	\label{ss:obtaining an FPG}
-	The process of obtaining an FPG from the given MCG is expected to be an iterative process in which the user applies the algorithm to determine whether or not an FPG can be obtained, 
-	and then modifies the indegree limits or the MCG to obtain a satisfactory FPG. 
-	For example, after determining the location of a hole, the user can change the indegree limit to make the node a design variable, or the user could change the MCG by adding a new analysis code such that the hole is resolved.
+	Although an FPG may now be obtained, this process has not provided a method to decide which edges to retain when resolving a collision, and these decisions will likely determine which analysis tools are included in the FPG.
+These decisions are left to the discretion of the designer in the context of the specific implementation.
+	However, the graph-based approach presented in this paper enables standard graph algorithms to be employed to automate these decisions based on considerations of metrics related to the graph or other data related to the analysis.
+	An example is a metric to obtain an FPG with the fewest possible number of cycles. 
+In this example, the FPG with the fewest number cycles is shown in Fig.~\ref{f:FPG fewest cycles}, in which the edges belonging to a cycle are inticated by dashed lines. 
+For this example problem, the cycles were detected using the implementation of Johnson's algorithm \cite{Johnson1975} in the Python package NetworkX. 
+There are two cycles which arise from each local output from PDCYL being directed into FLOPSa and then back into PDCYL. 
+%((DP:discuss cycle equivalency))
+%((mention edge weight for center edge))
 
-	As presented in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}, the first step to obtaining an FPG is to detect holes. 
-	To start, we set the lower indegree limit for every node to be unity. 
-	In this case, it turns out that the variable nodes representing the number of engines is a hole for NPSS, FLOPSa, and FLOPSb. 
-	This implies that an FPG cannot be obtained from the given MCG, as shown in Fig.~\ref{f:MCG nengines}. 
-	In this figure, the green nodes indicate analysis blocks with a hole, and the cyan nodes indicate analysis blocks that became unused after the invalid analyses were removed ((connections were removed, color edges too)).
+	Similar alternatives would be to minimize the number of analysis blocks involved in cycles, counting multiplicity, or to minimize the length of the longest cycle (called the \emph{circumference} of the graph).
+Since these graphs will always be relatively small (on the order of hundreds of nodes) the
+	computational expense for examining all possible FPGs is negligble compared to the expected expense of using the problem formulation to obtain a solution.
 	\begin{figure}[htb!]
 	  \begin{center}
-		\includegraphics[width=.6\textwidth]{images/MCG_nengines}
-	  \end{center}
-	  \caption{The MCG with one global variable missing. The resulting holes are shown in green and unused analyses shown in cyan.}
-	\label{f:MCG nengines}
-	\end{figure}
-	Therefore, we shall now create another variable node to represent the number of engines as a global input and also create connection edges to supply this new input.
-
-	While Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG} indicated the process for obtaining an FPG, it did not specify how to decide which edges to keep when resolving a collision, that freedom is left to the specific implementation.
-	The benefit of this approach is that standard graph theory algorithms can be used to make these decisions. 
-	The first example of this is to obtain an FPG with the fewest possible number of cycles, which are found using the implementation of Johnson's algorithm \cite{Johnson1975} in the Python package NetworkX.
-	The FPG with the fewest cycles is shown in Fig.~\ref{f:FPG fewest cycles}, which indicates one cycle shown in yellow. 
-	A similar alternative would be to minimize the number of analysis blocks involved in cycles, counting multiplicity, or to minimize the length of the longest cycle.
-Since these graphs will always be relatively small (on the order of hundres of nodes) the
-	computational expense for examining all possible FPGs is negligble.
-	\begin{figure}[htb!]
-	  \begin{center}
-		\includegraphics[width=.6\textwidth]{images/FPG_fewest_cycles}
+		\includegraphics[width=4.5in]{images/FPG_edit_fewest_cycles}
 	  \end{center}
@@
 
-	Alternatively, it may be desirable to resolve collisions by choosing the higher fidelity codes, such as NPSS in this case.
+	Alternatively, it may be desirable to resolve collisions by choosing the higher fidelity analysis tools.
 	One method to ensure that the edges directed from certain analysis blocks are those chosen to resolve conflicts is to use a ranking system.
-	For this case, consider the rankings for each analysis code given in Table \ref{t:rankings}.
+Each analysis block is assigned a value by the designer describing how desireable it is for this block to be present in the FPG.
+The connection edges directed out of each analysis block are then given a weight equal to the value assigned to the analysis block.
+Finally, each collision is resolved by selected the edges with the highest weights.
+For the current example problem, consider the rankings for each analysis code given in Table \ref{t:rankings}.
 	\begin{table}[htbp]
@@
 	\end{table}%
-	Step two in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG} (the resolution of collisions) is now conducted by selecting the edges directed from the analysis blocks with the highest rank for each collision.
-	The resulting FPG is shown in Fig.~\ref{f:FPG highest rank}, and this graph has four cycles, which are enumerated in Table \ref{t:cycles}.
+	The resulting FPG has four cycles and is shown in Fig.~\ref{f:FPG highest rank}. 
+%This graph has four cycles, which are enumerated in Table \ref{t:cycles}.
 	\begin{figure}[htb!]
 	  \begin{center}
-		\includegraphics[width=.6\textwidth]{images/FPG_highest_rank}
+		\includegraphics[width=4.5in]{images/FPG_edit_ranking}
 	  \end{center}
@@
 	\end{figure}
-	\begin{table}[htbp]
-	  \centering
-	  \caption{Cycles for the FPG obtained using ranking}
-		\begin{tabular}{ccccccc}
-		\toprule
-		output/input & analysis code & output/input & analysis code & output/input & analysis code & output/input \\
-		\midrule
-		engine weight & FLOPSb & drag  & NPSS  & engine performance & WATE  & engine weight \\
-		engine performance & FLOPSb & drag  & NPSS  & engine performance &       &  \\
-		fuselage weight & FLOPSb & total weight & PDCYL & fuselage weight &       &  \\
-		wing weight & FLOBSb & total weight & PDCYL & wing weight &       &  \\
-		\bottomrule
-		\end{tabular}%
-	  \label{t:cycles}
-	\end{table}%
+%	\begin{table}[htbp]
+%	  \centering
+%	  \caption{Cycles for the FPG obtained using ranking}
+%		\begin{tabular}{ccccccc}
+%		\toprule
+%		output/input & analysis code & output/input & analysis code & output/input & analysis code & output/input \\
+%		\midrule
+%		engine weight & FLOPSb & drag  & NPSS  & engine performance & WATE  & engine weight \\
+%		engine performance & FLOPSb & drag  & NPSS  & engine performance &       &  \\
+%		fuselage weight & FLOPSb & total weight & PDCYL & fuselage weight &       &  \\
+%		wing weight & FLOBSb & total weight & PDCYL & wing weight &       &  \\
+%		\bottomrule
+%		\end{tabular}%
+%	  \label{t:cycles}
+%	\end{table}%
 
-	Finally, now consider the case where it is desired to have the inviscid drag input into FLOPSb be a multi--fidelity input, meaning that it is sought to use multiple analysis codes to calculate the same variable.
-	This is done simply by setting the upper indegree limit for this node as two and then repreating the (automated) process. 
+	Finally, consider a case in which 
+the inviscid drag input into FLOPSb is a multi-fidelity input, meaning that 
+ multiple analysis codes calculate the same variable. 
+	This multi-fidelity formulation is implemented by setting the upper indegree limit for this node as two and then repreating the (automated) process. 
 	In this case, both VORLAX and PMARC are retained, resulting in an FPG that omits only FLOPSa.
-%\clearpage
-%\newtheorem{name}{Printed output}
-%\begin{name}
-%asdf
-%\end{name}
-%\begin{proof}[asdf]
-%asdf
-%\end{proof}
-
-%\begin{table}[h!]
-% \begin{center}
-%  \caption{Example problem analysis codes}
-%  \label{t:analysis codes}
-%  \begin{tabular}{ccc} \hline 
-%Analysis code & input variables & output variables \\ \hline
-%VSP & geometry & wetted area \\
-%PDCYL & geometry & fuselage weight \\
-%	& total weight & wing weight \\
-%NPSS & number of engines & engine performance \\
-%	& drag &	\\
-%VORLAX & geometry & inviscid drag \\
-%PMARC & geometry & inviscid drag \\
-%WATE & geometry & engine weight \\
-%	& engine performance &	\\ 
-%FLOPSa & number of engines & engine weight \\
-%	& mission & performance \\
-%	& geometry & total weight \\
-%	& wetted area & engine performance \\
-%	& fuselage weight & drag \\
-%FLOPSb & number of engines & drag \\
-%	& mission & performance \\
-%	& geometry & total weight \\
-%	& wetted area & 	 \\
-%	& engine weight & 	\\
-%	& wing weight & 	\\
-%	& inviscid drag & 	\\
-%	& engine performance & 	\\ \hline
-%  \end{tabular}
-% \end{center}
-% \vspace{-15pt}
-%\end{table}

### 2013-05-08 graph_syntax.tex
Major update

--- 2013-04-22 graph_syntax.tex
@@
 
-\section{Requirements for New Graph Syntax}
-  \label{s:requirements}
-  The goal of the new graph--based syntax presented here is to enable the general structure of an MDAO problem to be described independently of any solution information, 
-  while still being able to accomodate the more specific case when a solution 
-  strategy is applied. In order to achieve this goal, 
-  the graph syntax needs to accomodate a number of MDAO problem constructs: 
-  \begin{itemize}
-    \item Analysis tools and their interconnections
-    \item Design variables, objectives, and constraints
-    \item Local and global information
-    \item Coupling between analyses
-    \item Multi-fidelity analyses
-  \end{itemize}
-
-  The new syntax is intended to represent three phases of the design problem 
-  formulation process. In the initial problem definition phase, the specific 
-  analysis tools and design goals are identified. Next, a single formal problem 
-  formulation is identified that specifies design variables, constraints, 
-  objectives, analysis tools, and all other elements required to represent 
-  the overall MDAO problem. Lastly, a specific procedure for solving the problem 
-  is selected, e.g. selecting an MDAO optimization architecture. Using 
-  the proposed graph syntax, these phases can be represented with the following 
-  three graphs:
-
-  \begin{itemize}
-    \item Maximal Connectivity Graph (MCG)
-    \item Fundamental Problem Graph (FPG)
-    \item Problem Solution Graph (PSG)
-  \end{itemize}
-
-  The \emph{maximal connectivity graph} represents the first phase of the problem formulation with all 
-  analysis codes being considered and all possible connections between them also present. The second graph 
-  is the \emph{fundamental problem graph}, which is the smallest possible graph 
-  that still fully defines a given problem formulation that could in principle be solved by an MDAO technique. Finally, a \emph{specific problem formulation} 
-  may be represented by including additional edges and nodes to represent the 
-  solution strategy being employed to solve the problem.
-
-  The relationship between these three graphs is depicted in Figs.~\ref{f:tree} and \ref{f:hourglass}. 
-  The tree diagram demonstrates the fact that it is generally possible to obtain 
-  multiple FPGs from a single maximal connectivity graph. This  may correspond to 
-  different down--selections of analysis codes, different connections between them, 
-  or both. Each down--selection shrinks the number of possible FPGs that could be reached 
-  until only one is remaining. Then, from a single FPG, different PSGs may be obtained by implementing 
-  different solution strategies. If you consider the size of a graph to be the sum of all of its
-  edges and nodes then the hourglass shape in Fig. \ref{f:hourglass} illustrates how
-  the FPG is obtained from the MCG by removing nodes and edges, 
-  and the PSG is obtained from the FPG by adding nodes and edges.
-
-  \begin{figure}[htb!]
-    \centering
-    \subfigure[number of possible graphs]{
-    \includegraphics[width=2.0in]{images/tree}
-    \label{f:tree}
-    }
-    \subfigure[graph size]{
-    \includegraphics[width=2.0in]{images/hourglass}
-    \label{f:hourglass}
-    }
-  \caption{The relationship between the MCG, FPG, and PSG.}
-  \end{figure}
-
-\section{Syntax Definition}
+\section{Graph-Based Syntax Definition and Formalisms}
   \label{s:syntax definition}
-  In this section, we present the necessary graph theoretic fundamentals to 
+  In this section we present our graph-based syntax, beginning with the necessary graph theoretic fundamentals to 
   construct the graphs discussed in Sec.~\ref{s:requirements}. 
-  The notation used in this work is adapted from Diestel \cite{Diestel2010}. 
+  The notation used in this paper is adapted from Diestel \cite{Diestel2010}. 
   A \emph{graph} is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, 
-  which means that the elements of $E$ are 2--element subsets of $V$. The set $V$ 
+  which means that the elements of $E$ are 2-element subsets of $V$. The set $V$ 
   contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
@@
   indicated by the first entry and directed to the node indicated by the second 
-  entry. Edge $e$ = $(x,y)$ may be referred to simply as $xy$. For node $v \in V$ 
-  the edges directed out from $v$ are denoted by $E(v)$ and the edges directed into $v$ are given 
+  entry. An edge $e$ = $(x,y)$ may be referred to simply as $xy$. The edges directed out from node $v$ are denoted by $E(v)$ and the edges directed into $v$ are given 
   by $E^{-1}(v)$. 
-  If $E$ is not a one--to--one mapping, $E(v)$ may be the empty set, a single node, or a set of nodes.
+  If $E$ is not a one-to-one mapping, $E(v)$ may be the empty set, a single node, or a set of nodes.
   \begin{figure}[htb!]
@@
   with $V = \{x_0,x_1,\ldots,x_k\}$ and $E = \{(x_0,x_1),(x_1,x_2),\ldots,(x_{k-1},x_k)\}$. 
+Path $P$ is a \emph{cycle} if $x_0 = x_k$.
   A \emph{reverse path} $P_R$ in graph $G$ is a path on $R$, where $R$ is the 
@@
   \label{ss:types}
-  We now define different types of nodes and edges with specific properties to compose graphs used to represent an MDAO problem. 
-  There are three node types:  
+Along with these fundamentals of graph theory, we categorize nodes and edges into distinct types in order to provide an intuitive approach to MDAO problem formulation. The three node types are
   \begin{description}
-    \item[variable:] represents scalar or array data
-    \begin{equation} 
-      V_{\txt{var}} = \{v \st t_{\txt{node}}(v) = \txt{`variable'}\}
-    \end{equation}
+    \item[\bf{variable:}] represents scalar or array data, inputs and outputs,
+    \item[\bf{model:}] represents the computation peformed by analysis tools,
+    \item[\bf{driver:}] control logic capable of managing iterations (PSG),
+  \end{description}
+and the three edge types are
+%$V_\txt{var}$, $V_\txt{model}$, $V_\txt{driver}$
+  \begin{description}
+  \item[\bf{connection edge:}] exchange of information external to analysis tools,
+  \item[\bf{model edge:}] exchange of information internal to analysis tools,
+  \item[\bf{driven edge:}] passing of information from a driver node to a variable node (PSG).
+  \end{description}
+Figure \ref{f:sellar types} demonstrates the usage of these node and edge types via the Sellar problem.
+Model nodes are indicated by squares, variable nodes are indicated by circles, connection edges are indicated by dotted lines, and model edges are indicated by solid lines.
+\begin{figure}[htb!]
+  \begin{center}
+    \includegraphics[width=4.0in]{images/sellar_types}
+  \end{center}
+  \caption{Sellar problem represented as a graph.}
+\label{f:sellar types}
+\end{figure}
 
-    \item[model:] responsible for mapping inputs to outputs
-    \begin{equation} 
-      V_{\txt{model}} = \{v \st t_{\txt{node}}(v) = \txt{`model'}\}
-    \end{equation}
+% ((JG: A single variable node can have many incomming driven edges. These edges are 
+%    free and can be added or removed from the graph.))
+%
+%  [[Insert legend for node and edge types]] ((DP: I'm not sure what the legend is supposed to be))
 
-    \item[driver:] control logic capable of managing iteration
-    \begin{equation} 
-      V_{\txt{driver}} = \{v \st t_{\txt{node}}(v) = \txt{`driver'}\}
-    \end{equation}
-
-  \end{description}
-
-  In addition to the three node types, there are two edge types:
-
-  \begin{description}
-  \item[connection edge:] Exchange of information between two nodes. These edges 
-  can either be fixed (cannot be removed from graph) or free (can be removed). 
-    \begin{equation} 
-      E_{\txt{conn}} = \{e \st t_{\txt{edge}}(e) = \txt{`connection'}\}
-    \end{equation}
-  \item [driven edge:] Passing of information from a driver node to a 
-    variable node. A single variable node can have many incomming driven edges. These edges are 
-    free and can be added or removed from the graph.
-     \begin{equation} 
-       E_{\txt{driven}} = \{e \st t_{\txt{edge}}(e) = \txt{`driven'}\}
-     \end{equation} 
-  \end{description}
-
-  [[Insert legend for node and edge types]]
+  A rule set is provided for the usage of these node and edge types to provide a structure to the graph-based representation.
+  The driver node and the driven edge are allowed to be present only in PSGs, whereas the other node and edge types can be present in any of the three graph types, 
+  subject to the following restrictions: 
+  \begin{enumerate}
+  \item A model node may have only one edge directed to or from another model node.
+  \item A model node may have only model edges directed in or out.
+  \item A model node must have at least one edge directed in and at least one edge 
+    directed out.
+  \item If a variable node has an outgoing model edge, then it may not have other outgoing edges.
+  \item If a variable node has an incoming model edge, then it may not have any additional incoming edges.
+  \end{enumerate}
 
@@
   the terminology ``model'' node instead of ``function'' node to be consistent 
-  with modern MDAO framework terminonlogy. The present work adds one additional 
-  node types and one additional edge type to the syntax to allow descriptive
-  graphs for all three phases of the design problem formulation process. In addition, we introduce the 
+  with modern MDAO framework terminology. The present work adds one additional 
+  node type and two additional edge types to the syntax to allow descriptive
+  graphs for all three phases of the design problem formulation process indicated in Figs.~\ref{f:tree} and \ref{f:hourglass}.
+\begin{comment}
+In addition, we introduce the 
   concept of fixed vs. free connection edges to represent the
   static relationship between variable and model nodes from a single analysis, 
-  vs the flexible relationship between variables from different analyses.
+  as opposed to the flexible relationship between variables from different analyses. 
+\end{comment}
 
-\subsection{Rules for Nodes and Edges}
-  \label{ss:rules}
-  A rule set is provided for the usage of these nodes and edges.
-  The driver node and the driven edge are allowed to be present only in PSGs.
-  Other node and edge types can be present in any of the three graph types, 
-  subject to the following restrictions: 
-  \begin{enumerate}
-  \item A model node may have only one edge directed to or from another model node.
-  \item A model node may have only fixed connection edges directed in or out.
-  \item A model node must have at least one edge directed in and at least one edge 
-    directed out.
-  \item If a variable node has an outgoing edge to a model node, then it may not have other outgoing edges.
-  \end{enumerate}
+\subsection{Analysis Blocks}
+\label{ss:analysis blocks}
+Analysis tools take in a set of input variables and then calculate 
+the values for their respective outputs. We represent this process
+by a digraph called an \emph{analysis block}; a notional analysis block is shown in Fig.~\ref{f:analysis block}. 
+%Each analysis block is a directed graph denoted as $ A=(V_{A},E_{A})$
+As indicated in this figure, analysis blocks comprise three sets of nodes representing the distinct local inputs, the analysis (computation), and the distinct local outputs.
+The local input and local outputs nodes are variable nodes, while the nodes representing the analysis are model nodes.
+All of the edges within the analysis block are model edges, and are considered fixed to the analysis block.
+This graph structure demonstrated by Fig.~\ref{f:analysis block} satisfies the rules listed in Sec.~\ref{ss:types}.
+Reciprocally, the rules ensure that analysis tools are represented via this analysis block structure.
+\begin{figure}[htb]
+    \begin{center}
+    \includegraphics[width=3.0in]{images/analysis_block}
+    \end{center}
+    \vspace{-10pt}
+\caption{Example analysis block with node types labeled.}
+\label{f:analysis block}
+\end{figure}
 
-\subsection{indegree and outdegree}
+The model edge connecting the two model nodes represents the necessary calculations to map the inputs to the outputs of the analysis. 
+This edge also provides the opportunity to encode computational cost or other characteristics of the analysis code as a weight in a weighted graph formulation. 
+Since all the edges within analysis blocks are fixed, the blocks themselves are fixed sub-graphs within the overall MDAO problem graph. 
+The connectivity of nodes and edges in an analysis block cannot be altered for the purposes of MDAO problem formulation; 
+however, analysis block sub-graphs can be added or removed from the MDAO problem graph as needed.
+Variable nodes in an analysis block can be distinguished as either an input or an output by the manner in which they are connected. 
+As shown in Fig.~\ref{f:analysis block}, inputs are represented as variable nodes that have an outgoing edge into a model node, and outputs are represented as variable nodes that have an incoming edge from a model node.
+
+%MDAO problems require that information be passed between sets of analyses ((DP: needs better transition)).
+%For this purpose, connection edges are placed to link analysis blocks
+% 
+%When information from one output of an analysis block is passed to an 
+%input of another analysis block, a new connection edge is added between the two 
+%variable nodes involved in the exchange. 
+%These connection edges are free, and unlike the edges within an analysis block, the edges can be added or removed depending on the needs of a given design problem. 
+%In other words, free connection edges are created by engineers linking an output of one tool to an input of another one. 
+%It is allowable for a single output to have outgoing connection edges to multiple downstream inputs.
+
+\subsection{Objectives, Constraints, and Global Inputs}
+\label{ss:objectives and constraints}
+Along with analysis tools, objectives, constraints, and global inputs also need to be represented.
+In the case of objective functions, a single output value generated by an analysis block could be selected, but commonly multiple values are assembeled together to form a composite objective function.
+Generally, both objective and constraint functions accept a set of inputs and map them to an output value of significance to the overall design problem.
+
+The operations of implementing composite objective and constraint functions, although typically simple, are effectively identical to the task performed by an analysis block. 
+A composite constraint or objective function can therefore be represented as an analysis block within the graph with its own input and output variable nodes. 
+Although fundamentally no different than an analysis block, for clarity and convience, it is useful to distinguish between analysis blocks corresponding to analysis codes and those that arise from the addition of objectives or constraints to the graph.
+Therefore, we use an \emph{expression block} to represent an objective or constraint. 
+These expression block graphs follow the same structure as analysis blocks presented in Sec.~\ref{ss:analysis blocks}.
+
+Inputs to multiple analysis tools are considered global inputs.
+These are represented in the graph syntax by single variable nodes refered to as global inputs---the definition of global with respect to a node will be defined subsequently.
+%For the sake of consistency of naming conventions $F_j$, $G_j$, and $H_j$ are used to deonte expression blocks for objectives, inequality constraints, and equality constraints. 
+%If the relevant objective and constraint values are direct results from other analysis blocks and composite functions are not needed, then expression blocks may be omitted from the MDAO problem graph.
+Figure \ref{f:obj-cons} demonstrates the use of global inputs (black circles), analysis blocks (dotted boxes), and expression blocks (dashed boxes), for the Sellar problem graph.
+\begin{figure}[htb!]
+  \begin{center}
+    \includegraphics[width=4.0in]{images/sellar_obj_and_cons}
+  \end{center}
+  \caption{Sellar problem graph indicating global inputs (black circles), analysis blocks (dotted boxes), and expression blocks (dashed boxes).}
+\label{f:obj-cons}
+\end{figure}
+
+\subsection{Indegree and Outdegree Limits}
   \label{s:indegree-outdegree}
-  The \emph{indegree} of a node is the number of connection edges directed in and 
+	To address the MDAO notion of design variables, we first introduce the concept of the degree of a node.
+  In a digraph, \emph{indegree} of a node is the number of edges directed in and 
   is denoted as $\txt{deg}^-(v)$, and the \emph{outdegree} 
-  is the number of connection edges directed out and it is denoted as $\txt{deg}^+(v)$.
-  The degree of a given node is a function only of the connection edges 
-  attached to it. The number of driven edges is not relevant because any number 
-  of drivers could be involved in different parts of an solution process. For 
-  example in a sequential optimization in which a global optimizer is used first 
-  followed by a gadient based optimizer second to refine global optimizer 
-  result, design variables would have incomming driven edges from both optimizers. 
-  These driven edges would not affect the indegree of those variable nodes.
+  is the number of edges directed out and it is denoted as $\txt{deg}^+(v)$ \cite{Diestel2010}.
+  In this paper, the indegree and outdegree of a given node refer only to the number of the connection edges 
+  directed into or out of the node, respectively. 
+%The number of driven edges is not considered relevant because any number 
+%  of drivers could be involved in different parts of a solution process. For 
+%  example, in a sequential optimization in which a global optimizer is used first 
+%  followed by a gradient-based optimizer second to refine global optimizer 
+%  result, design variables would have incomming driven edges from both optimizers. 
+%  These driven edges would not affect the indegree of those variable nodes.
 
-  Specifially for variable nodes we also define the \emph{upper indegree limit}
-
+  We may now define the \emph{upper indegree limit} 
   \begin{equation}
@@
   These user-specified limits govern the number of connection edges that may 
-  be directed into a variable node for a valid graph. Consider a variable 
+  be directed into a node for a valid graph; of course, $\txt{deg}_l^-(v) \leq \txt{deg}_u^-(v)$ must be satisfied. 
+For example, consider a variable 
   node $v$ with $\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, $v$ 
-  must have exactly one incomming explicit edge or the graph is deemed invalid. 
-  Two possiblities exist for voilating these limit conditions:
+  must have exactly one incoming connection edge or the graph is deemed an invalid problem formulation.
 
+\section{MDAO Constructs Derived from the Graph-Based Syntax}
+\label{s:graph representation}
+In Sec.~\ref{s:syntax definition} we presented the structure of the graph-based representation.
+We now discuss how these formalisms provide the remaining MDAO problem constructs identified in Sec.~\ref{s:requirements}.
+
+%To begin, we make use of the Sellar problem, described in 
+%Eq. \ref{eqn:simple_fpf} as an example to illustrate the implementation of the syntax. 
+%The graph representation of the Sellar problem, using the new stynax, is 
+%presented in Fig. \ref{f:sellar_graph_full}. In this section sub-graphs or 
+%slight modifications of the full graph will be used to demonstrate the 
+%important MDAO constructs described by this graph syntax.
+%\begin{figure}[htb!]
+%    \begin{center}
+%    \includegraphics[width=4in]{images/sellar_graph_full}
+%    \end{center}
+%    \vspace{-10pt}
+%\caption{Graph of the Sellar problem formulation}
+%\label{f:sellar_graph_full}
+%\end{figure}
+
+\subsection{Design Variables, Holes, and Collisions}
+Design variables are input variables whose values are free to be 
+changed by the designer or by an optimizer in an MDAO study. In the graph-based syntax, a node corresponds to a design variable when the indegree limit of the node is prescribed as zero.
+
+If the indegree limits are violated by the number of incoming connection edges (the indegree of the node), the graph is regarded as an invalid problem formulation.
+There are two classifications for these violations:
   \begin{description}
-    \item[hole: ] The number of incomming edges is less than the lower indegree limit:
-      \begin{equation} \txt{deg}^-(v) < \txt{deg}_l^-(v) \label{e:hole} \end{equation}
-      A hole represents a lack of information being supplied to the node.
-    \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
-      \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \label{e:collision}\end{equation}
+    \item[\bf{hole:}] The number of incoming edges is less than the lower indegree limit:
+      \begin{equation} 
+\txt{deg}^-(v) < \txt{deg}_l^-(v). 
+\label{e:hole} 
+\end{equation}
+      A hole represents a lack of information being supplied to the variable node.
+This implies that the analysis tool being represented by the analysis block would not be capable of producing outputs.
+    \item[\bf{collision:}] The number of incoming edges is greater than the upper indegree limit. 
+      \begin{equation} 
+\txt{deg}^-(v) > \txt{deg}_u^-(v). 
+\label{e:collision}
+\end{equation}A collision represents redundant information being supplied to the variable node from two or more sources.
+This conflict implies an ambiguity as to which information is to be used as an input for the analysis tool.
   \end{description}
 
-  The presence of holes and collisions in a graph represent issues that will give
-  rise to an invalid problem formulation. An algorithm for detecting and 
-  managing these violations if proposed in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}.
+%For the Sellar problem, there are three design variables: $x_1^*$, $z_1^*$, and 
+%$z_2^*$. In Fig. \ref{f:designvars} each variable is 
+%annotated with the lower indegree limit, and the three design variables are identified 
+%in bold((DP:check)). 
+%
+%\begin{figure}[htb!]
+%  \begin{center}
+%    [Graph Here....]
+%    \includegraphics[width=.6\textwidth]{images/design_vars_graph}
+%  \end{center}
+%  \caption{Sellar problem graph with lower indegree limits for all variables \label{f:designvars}}
+%\end{figure}
+%
+%In Fig. \ref{f:designvars}, the design variable are void of incoming edges. 
+%For a valid PSG, design variables could have one or more incoming driven edges representing a value determined by an optimizer.
 
-\section{Graph Representation of MDAO Constructs}
-\label{s:graph representation}
-In this section we present our graph syntax applied to the description of MDAO 
-Problems. We make use of the Sellar problem, described in 
-Eq. \ref{eqn:simple_fpf} as an example to illustrate the implementation of the syntax. 
-The graph representation of the Sellar problem, using the new stynax, is 
-presented in Fig. \ref{f:sellar_graph_full}. In this section sub-graphs or 
-slight modifications of the full graph will be used to demonstrate the 
-important MDAO constructs described by this graph syntax.
+  The presence of holes and collisions in a graph represent conflicts that will give
+  rise to an invalid problem formulation. It is expected that a maximal connectivity graph will posess these conflicts, whereas a fundamental problem graph will not.
 
-\begin{figure}[htb!]
-    \begin{center}
-    \includegraphics[width=4in]{images/sellar_graph_full}
-    \end{center}
-    \vspace{-10pt}
-\caption{Graph of the Sellar problem formulation}
-\label{f:sellar_graph_full}
-\end{figure}
+%However, if the graph remains unchanged while the user changes the lower indegree limit of the node to now be zero, then the node becomes a design variable.
 
-\subsection{Analysis Blocks and Connections}
-\label{ss:analysis blocks and connections}
-Analysis tools take in a number input variables and then calculate 
-the values for their respective outputs. We represent this process
-by a group of nodes and edges called an \emph{analysis block}, 
-shown in Fig. \ref{f:analysis block}. Each analysis block is 
-denoted by:
+%In some cases, the distinction between a hole and a design variable is rather subtle. Both have no incoming connection edges, but 
+%
+%
+%Note that although there are no incoming edges, design variables are not holes 
+%in the graph as defined by Eq. \ref{e:hole}. Since they have $\txt{deg}_l^-(v)=0$ 
+%having no incoming connection edges does not violate their lower indegree limit. 
+%However, the concept of holes can be useful for identifying potential design variables
+%given an arbitrary MCG. If you first set $\txt{deg}_l^-(v)=1$ for all variable nodes, 
+%and then find the resulint holes, you will have a list of variable nodes that 
+%could potentially be design variables. Some variables will be true design variables, 
+%while others might represent missing information in the problem formulation. For 
+%instance, if there is a hole for a drag input of an aircraft mission analysis code 
+%then this is not likely a good choice for a design variable. So it is up to the designer to 
+%examine each hole and decide if it's an appropriate design variable. If 
+%the variable should become a design variable then set $\txt{deg}_l^-(v)=0$. 
+%If a variable should not be a design variable, then leave it as a hole which 
+%needs to be addressed in a manner discussed in section \ref{ss:obtaining FPG}.
 
-\begin{equation}
-  A_i=(V_{A_i},E_{A_i}),\ for\ i\in I, 
-\end{equation}
+\subsection{Local vs Global Nodes ((recommend omission))}
 
-Where $m$ is the number of analysis tools represented in the graph 
-and $I=\{1,2,\ldots,m\}$.
+  In this graph-based syntax, a variable node is considered \emph{local} if it has exactly one outgoing connection edge or \emph{global} if it has more than one outgoing connection edge.
+Therefore, a variable node may be used to represent a global input if it has at least two outgoing connection edges.
+Otherwise, the input is formulated as a design variable that is local to a specific analysis block.
+%  The set of local nodes, $V_\txt{l}$, is the collection of 
+%  nodes with exactly one outgoing connection edge. 
+%
+%  \begin{equation}
+%    V_{\txt{l}} = \{V_{\txt{var}} \st \txt{deg}^+(v)=1\}
+%    \label{e:local_definition}
+%  \end{equation}
+%  In section \ref{s:specific vs fundamental} variable nodes that are part of 
+%  an analysis block are inherently local and similarly model nodes 
+%  are also local. Following the definition in Eq. \ref{e:local_definition}
+%  we can now see why. Variable nodes connected to model nodes are only allowed to 
+%  have one outgoing edge, and hence will always by definition be local variables.
 
-Within a given analysis block, variable node represents a single input or 
-output and is connected to a single model node via a fixed connection edge. Note that 
-in Fig.~\ref{f:analysis block}, the analysis block includes two model nodes, 
-with a single edge connecting them. This edge represents the necessary 
-calculations to map given inputs to the outputs of the anlysis. This calculation edge provides the opportunity 
-to encode computational cost or other characteristics of the analysis code as 
-a weight in a weighted graph formulation. If associating a weight is 
-unnecessary, the calculation edge and second model node may be omitted and all input and output 
-variables can be connected to a single model node for a given analysis block. 
-Since all the edges within analysis blocks are fixed, the blocks themselves are 
-fixed sub-graphs within the overall MDAO problem graph. The connectivity of 
-nodes and edges in an analysis block cannot be altered for the purposes of MDAO 
-problem formulation; however, analysis block sub-graphs can be added or removed 
-from the MDAO problem graph as needed.
+%  The set of global nodes, $V_\txt{g}$, is the collection of all nodes with more than 
+%  one outgoing connection edge. (())
+%
+%  \begin{equation}
+%    V_{\txt{g}} = \{V_{\txt{var}} \st \txt{deg}^+(v)>1\}
+%    \label{e:global_definition}
+%  \end{equation}
 
-\begin{figure}[htb!]
-    \begin{center}
-    \includegraphics[width=4in]{images/analysis_block}
-    \end{center}
-    \vspace{-10pt}
-\caption{Example analysis block. The each node type and edge type is labeled in italics and annotated parenthetically.}
-\label{f:analysis block}
-\end{figure}
+%  Following our definition any variable node that had multiple 
+%  outgoing edges, is global. In Fig. \ref{f:sellar_graph_full} 
+%  the variable nodes for $x_1^*$, $z_1^*$, and $z_2^*$ have multiple outgoing 
+%  edges and would be considered global. 
+In the Sellar problem (Fig.~\ref{f:sellar graph}) the variable nodes $x_1^*$, $z_1^*$, and $z_2^*$ have multiple outgoing edges and are therefore considered global inputs. 
+  The variable $x_1$ is usually refered to as 
+  a local variable, which is contradicted by the given graph ((DP: can we cite someone calling this local?)). This discrepancy 
+  arises from the explicit treatment of objectives and constraints as separate 
+  expression blocks. This variable is often considered a local 
+  variable because it directly affects only analysis block $A_1$. However, by expanding
+  $F$ into an expression block with its own inputs, $F$ maintains its own local 
+  copy of the $x_1$ variable. This formulation necessitates the creation of the global 
+  node $x_1^*$ in the graph with two outgoing edges to link the two $x_1$ inputs nodes 
+  in the respective blocks.
 
-Variable nodes in an analysis block can be distinguished as either an input or 
-an output by the manner in which they are connected. As shown in Fig.~\ref{f:analysis block}, 
-inputs are represented as variable nodes that have an outgoing edge into a model 
-node. Conversely, outputs represented as any variable nodes that have an incoming 
-edge from a model node.
-
-MDAO problems require that information be passed between sets of analyses. When 
-information from one output of an analysis block is passed, or connected, to an 
-input of another analysis block, a new connection edge is added between the two 
-variable nodes involved in the exchange. These connection edges are free, and unlike the edges 
-within an analysis block, the edges can be added or removed depending on the needs
-of a given design problem. In other words, free connection edges are created by 
-engineers linking an output of one tool to an input of another one. It is 
-allowable for a single output to have outgoing connection edges to multiple 
-downstream inputs.
-
-\subsection{Design Variables}
-Design variables are the input variables that are free to be 
-changed by the designer or by an optimizer in an MDAO study. In a graph design
-variables will be a set of variable nodes, $V_{dv}$, identified as follows:
-
-\begin{equation}
-  V_{\txt{dv}} = \{V_{\txt{var}} \st \txt{deg}_l^-(v)=0\}
-\end{equation}
-
-For the Sellar problem, there are three design variables: $x_1^*$, $z_1^*$, and 
-$z_2^*$. In Fig. \ref{f:designvars} all variable from the Sellar problem are 
-annotated with their lower indegree limit and the three design variables are identified 
-in bold.
-
-\begin{figure}[htb!]
-  \begin{center}
-    [Graph Here....]
-    \includegraphics[width=.6\textwidth]{images/design_vars_graph}
-  \end{center}
-  \caption{Sellar problem graph with lower indegree limits for all variables \label{f:designvars}}
-\end{figure}
-
-In Fig. \ref{f:designvars}, none of the design variable nodes have any incomming 
-edges. In general, in a MCG or a FPG, design variables will 
-show up as nodes without any incomming edges. For a valid PSG, design variables 
-could have one or more incomming driven edges representing a value determined by 
-a optimizer.
-
-Note the although there are no incomming edges, design variables are not holes 
-in the graph as defined by Eq. \ref{e:hole}. Since they have $\txt{deg}_l^-(v)=0$ 
-having no incomming connection edges does not violate their lower indegree limit. 
-However, the concept of holes can be useful for identifying potential design variables
-given an arbitrary MCG. If you first set $\txt{deg}_l^-(v)=1$ for all variable nodes, 
-and then find the resulint holes, you will have a list of variable nodes that 
-could potentially be design variables. Some variables will be true design variables, 
-while others might represent missing information in the problem formulation. For 
-instance, if there is a hole for a drag input of an aircraft mission analysis code 
-then this is not likely a good choice for a design variable. So it is up to the designer to 
-examine each hole and decide if it's an appropriate design variable. If 
-the variable should become a design variable then set $\txt{deg}_l^-(v)=0$. 
-If a variable should not be a design variable, then leave it as a hole which 
-needs to be addressed in a manner discussed in section \ref{ss:obtaining FPG}.
-
-\subsection{Objectives and Constraints}
-\label{ss:objectives and constraints}
-Objectives and constraints are constructs that need to be identified when 
-describing and MDAO problem. In the case of objective functions a single output 
-value generated by an analysis block could be selected, but commonly multiple 
-values are assembeled together to form a composite objective function. 
-Generally both objective and constraint functions accept a set of inputs and map 
-them to an output value of significance to the overall design problem.
-
-The operations of implementing composite objective and constraint functions, 
-although typically simple, are effectively the identical to the task performed by an 
-analysis block. A composit constraint or objective function can therfore 
-be represented as an additional analysis block within the graph, with its own input and 
-output variable nodes. Although fundamentally no different than an analysis block, 
-for clarity and convience, it is useful to distinguish between analysis 
-blocks corresponding to analysis codes and those that arise from the addition of 
-objectives or constraints to the graph. Therefore, we denote an \emph{expression block}, as:
-
-\begin{equation}
-  F_j=(V_{F_j},E_{F_j}),\ for\ j\in J, 
-\end{equation}
-
-Where $m$ is the number of expressions in the graph
-and $J=\{1,2,\ldots,m\}$. For the sake of consistency of naming conventions
-$F_j$, $G_j$, and $H_j$ are used to deonte expression blocks for objectives, inequality
-constraints, and equality constraints. If the relevant 
-objective and constraint values are direct results from other analysis blocks 
-and composite functions are not needed, then expression blocks may be omitted 
-from the MDAO problem graph. Fig. \ref{f:obj-cons} highlights the 
-objective and constraint blocks from the Sellar problem.
-
-\begin{figure}[htb!]
-  \begin{center}
-    \includegraphics[width=.6\textwidth]{images/obj_const_graph}
-  \end{center}
-  \caption{Notional graph with and objective and constraint nodes \label{f:obj-cons}}
-\end{figure}
-
-\subsection{Local vs Global Nodes}
-
-  The set of local nodes, $V_\txt{l}$, is the collection of 
-  nodes with exactly one outgoing connection edge.
-
-  \begin{equation}
-    V_{\txt{l}} = \{V_{\txt{var}} \st \txt{deg}^+(v)=1\}
-    \label{e:local_definition}
-  \end{equation}
-
-  In section \ref{s:specific vs fundamental} variable nodes that are part of 
-  an analysis block are inherently local and similarly model nodes 
-  are also local. Following the definition in Eq. \ref{e:local_definition}
-  we can now see why. Variable nodes connected to model nodes are only allowed to 
-  have one outgoing edge, and hence will always by definition be local variables.
-
-  The set of global nodes, $V_\txt{g}$, is the collection of all nodes with more than 
-  one outgoing connection edge.
-
-  \begin{equation}
-    V_{\txt{g}} = \{V_{\txt{var}} \st \txt{deg}^+(v)>1\}
-    \label{e:global_definition}
-  \end{equation}
-
-  Follwoing our definition any variable node that had multiple 
-  outgoing edges, is global. In Fig. \ref{f:sellar_graph_full} 
-  the variable nodes for $x_1^*$, $z_1^*$, and $z_2^*$ have multiple outgoing 
-  edges and would be considered global.
-
-  The variable $x_1$ from the Sellar problem is usually refered to as 
-  a local variable, which is contradicted by the given graph. This discrepancy 
-  arises from the explicit treatment of objectives and constraints as separate 
-  expression blocks. Normally in the Sellar problem, $x_1$ is considered a local 
-  variable because it directly affects only analysis block $A_1$. By expanding
-  $F$ into an expression block with its own inputs, then $F$ is assigned its own local 
-  copy of the $x_1$ variable. This formulation necessitates the creation of the global $x_1^*$ 
-  node in the graph with two outgoing edges to link the two $x_1$ inputs nodes 
-  in the different blocks. Interestingly, the classification of $x_1^*$ 
-  as a global variable fits nicely within the structure of the Colaborative 
-  Optimization (CO) archtiecture \cite{braun1996thesis}. The rules for setting up 
+This classification of $x_1^*$ in the Sellar problem
+  as a global variable agrees with the structure of the Collaborative 
+  Optimization (CO) archtiecture \cite{braun1996thesis}. The rules for constructing 
   a problem with CO require that each discipline be given an independent local 
@@
   within their respective disciplines, except when a local design variable appears 
-  explicitely in the objective or global constraint. In that case a global 
-  variable is created, and the analysis block again is again assigned an 
-  independent local copy of it. Given our definition of global vs local variable 
+  explicitely in the objective or global constraint. In this case, a global 
+  variable is created and the analysis block is again assigned an 
+  independent local copy of it. Following our definition of global and local variable 
   nodes, no such exception for local variables in global expressions is 
-  necessary. Whenever a given variable is included in an expression block, 
+  necessary. Whenever a variable is included in an expression block, 
   it forces the creation of a global variable node in the graph which will 
-  result in the correct structure for CO automatically.
+automatically result in the correct CO structure.
 
-  Analysis and expression blocks can themselves be local or global, but the 
-  criteria for determining this is slightly different than for variable nodes. 
-  In this case, you consider the set of connection edges for the block
-  (e.g. from all the variable nodes within the block) as a set. Additionally, 
-  you must consider both incoming and outgoing edges, where for variables 
-  only outgoing connection edges were relevant.
+  Analysis and expression blocks are also regarded local or global but the 
+  criteria for determining this is slightly different than for variable nodes---both incoming and outgoing edges must be considered.
+%  In this case, you consider the set of connection edges for the block
+%  (e.g. from all the variable nodes within the block) as a set. Additionally, 
+%, whereas for variables
+%  only outgoing connection edges are relevant. 
+  An analysis block is local if all incoming edges are from exactly one analysis 
+  block and all outgoing edges are from exactly one other analysis block. A global 
+  analysis block is defined as a block with incoming edges or outgoing edges 
+  to more than one analysis block.   In the Sellar problem, the expression blocks for each constraint, 
+  $G_1$ and $G_2$, are local to their respective analysis 
+  blocks, whereas the expression block for $F$ has incoming edges from $A_1$ and 
+  $A_2$, therefore it is considered global.
 
-  A local analysis block is 
-  defined as a block with incomming edges from exactly one other analysis 
-  block and outgoing edges from exactly one other analysis block. A global 
-  analysis block is defined as a block with incomming edges or outgoing edges 
-  to more than one analysis block.
-
-  [[MATH HERE]]
-
-  Looking at the Sellar problem in 
-  Fig. \ref{f:sellar_graph_full} we see the expression blocks for each constraint, 
-  $G_1$ and $G_2$. Both constraints are local to their respective analysis 
-  blocks. But the expression block for $F$ has incomming edges from $A_1$ and 
-  $A_2$, so it would considered global.
-
-  Catagorizing any node or analysis block block with incomming or outgoing edges
+  Catagorizing any node or analysis block with incoming or outgoing edges
   from more than one source as global neglects a more 
@@
 \subsection{Coupling Between Analyses}
-  Coupling is described by the presence of a a closed path between two or more
-  analysis blocks, called a cycle. A path, given by $P=(V_{path},E_{path})$, 
-  contains a cycle if at least two edges in the path refrence the same node, 
-  $E = \{(x_0,x_1),(x_1,x_2),\ldots,(x_{k-1},x_k),(x_k,x_0)\}$. These cycles 
-  decribe the reciprical dependence of a set of two or more analyses on each 
-  other's outputs. In the Sellar problem from Eqn \ref{eqn:simple_fpf} the two 
+In MDAO, coupling is the mutual dependence the outputs of two or more analysis codes. In this graph-based syntax, coupling is described by the presence of a cycle between two or more analysis blocks. 
+%A path, given by $P=(V_\txt{path},E_\txt{path})$, contains a cycle if at least two edges in the path refrence the same node,  $E = \{(x_0,x_1),(x_1,x_2),\ldots,(x_{k-1},x_k),(x_k,x_0)\}$. 
+%Cycles decribe the reciprical dependence of a set of two or more analyses on each other's outputs. 
+In the Sellar problem from Eqn \ref{eqn:simple_fpf} the two 
   coupling constraints, $H_1$ and $H_2$, provide a reciprical dependence between 
   $A_1$ and $A_2$. In a graph, these constraints appear as connection edges between 
-  the outputs of each analysis block. The relevant connection edges are highlighted in 
-  Fig. \ref{f:coupling}, where they form
+  the outputs of each analysis block. The edges belonging to a cycle in the Sellar problem are highlighted in 
+  Fig. \ref{f:sellar cycles}, where they form ((sentence is unfinished))
 
-  \begin{figure}
+ \begin{figure}[htb!]
   \begin{center}
-      %\includegraphics[height=.25\textheight]{}
-      [INSERT GRAPH HERE]
-      \caption{Graph of a notional problem with simple coupling \label{f:coupling}}
+    \includegraphics[width=4.0in]{images/sellar_cycles}
   \end{center}
+  \caption{Sellar problem graph with dashed edges indicating participation in a cycle.}
+\label{f:sellar cycles}
 \end{figure}
 
-  Coupling cycles do not contain driver nodes in the MCG of FPG representation. 
+  Coupling cycles do not contain driver nodes in the MCG or FPG representation. 
   Solvers, optimizers, and other iterative processes are not invovled in the coupling 
-  definition--though one will be necessary to build a valid PSG from a given FPG. 
-  Additionally, a coupling cycle has no inherent start or end. It would be acceptable to select
+  definition; however, one or more solvers will be necessary to build a valid PSG from a given FPG. 
+  Additionally, a coupling cycle has no inherent start or end point. It would be acceptable to select
   node in the cycle as a starting point and proceed around the
-  loop until you get back to the starting point. For the Sellar Problem, selecting 
+  loop until the starting point is reached again. For the Sellar Problem, selecting 
   $A_1$ as the starting point would yield a problem as given in 
@@
   Larger problems can contain more complex cycles in their FPG, indicating more 
-  complex coupling between analyses. For example A cycle can involve more than 
-  just two analysis blocks. Multiple independent cycles could also exist, indicating 
+  complex coupling between analyses. For example, a cycle can involve more than 
+   two analysis blocks. Multiple independent cycles could also exist, indicating 
   independent coupling relationships. Cycles can also overlap, meaning that the same analysis 
@@
   arise naturally as the size of problems grows, and managing this coupling may
-  become difficult. In the present work, Sec.~\ref{s:example problem}.\ref{ss:obtaining an FPG} 
+  become difficult. In the present work, Sec.~\ref{s:example problem}
   demonstrates how building an FPG from an MCG provides an opportunity to 
@@
 
-\subsection{Multi--fidelity Problems}
+\subsection{Multi-fidelity Problems}
   \label{ss:multi-fideliy problems}
-  A multi--fidelity problem can be characterized by two or more different analyses 
-  each calculating the same data. Representing this type of multi--fidelity 
-  construct in a graph will yield two analysis blocks, each with its own output 
-  variable nodes, having outgoing edges that connect to a single variable in a 
-  third analysis block that needs the data as input. So the set of all multi-fidelity
-  variables, $V_{mf}$, is given as:
+  A multi-fidelity MDAO problem can be characterized by a formulation in which two or more different analyses 
+  each calculating the same data.
+Multi-fidelity is represented in a graph by a variable node having an indegree greater than one, which means that multiple connection edges are directed into it.
+When the upper indegree limit of a variable node is set above one, then the node is implied to allow multiple fidelity instances of the variable.
+When the lower indegree limit of a variable node is set above one, then multiple fidelity instances are required..
+%  variables, $V_{mf}$, is given as: 
+%
+%  \begin{equation}
+%    V_{\txt{mf}} = \{ V_{\txt{var}} \st 1 < \txt{deg}^-(v) \leq \txt{deg}^-_u(v) \}
+%  \end{equation}
 
-  \begin{equation}
-    V_{\txt{mf}} = \{ V_{\txt{var}} \st 1 < \txt{deg}^-(v) \leq \txt{deg}^-_u(v) \}
-  \end{equation}
-
-  In a multi-fidelity problem, a given variable node may have multiple incomming 
-  edges without causing a collision as defined by by Eq.~\ref{e:collision}. 
+  In a multi-fidelity problem, a given variable node may have multiple incoming 
+  edges without implying a collision as defined by by Eq.~\ref{e:collision}. 
   Fig.~\ref{f:collision_example} shows a modified version of the Sellar problem 
@@
   $\txt{deg}^-_u(v) > 1$ for any variable in a graph represents a decision to 
-  allow multiple fidelities to interact at that part of the graph in order to alieviate
+  allow multiple fidelities to interact at that part of the graph in order to allieviate
   a conflict. In section \ref{ss:obtaining FPG} we present an algorithm for finding 
@@
   one in turn. 
-
   \begin{figure}
@@
     [INSERT GRAPH HERE]
-    \caption{Graph of the modified Sellar, multi--fidelity problem with $\txt{deg}^-_u(v)$ annotations on variable nodes\label{f:collision_example}}
+    \caption{Graph of the modified Sellar, multi-fidelity problem with $\txt{deg}^-_u(v)$ annotations on variable nodes\label{f:collision_example}}
   \end{center}
@@
 
-  Multi--fidelity problems are always charaterized by graphs with variable 
+  Multi-fidelity problems are always charaterized by graphs with variable 
   nodes that have an $\txt{deg}^-_u(v) > 1$. These problems require 
@@
   FPG. Instead the multi-fidelity mechanism specifics would be represented as a 
-  driver in a PSG derived from any given multi-fidelity FPG.
+  driver in a PSG derived from a multi-fidelity FPG.
+
+Table \ref{t:variable node classification} provides examples to summarize the classification of a variable node as a hole, design variable, single valid input (nominal), collision, or multi-fidelity.
+
+\begin{table}[htbp]
+  \centering
+  \caption{Variable Node Classification}
+    \begin{tabular}{ccccc}
+    \toprule
+    $\txt{deg}^-_l(v)$ & $\txt{deg}^-_u(v)$ & $\txt{deg}^-(v)$ & valid & classification \\
+    \midrule
+    0     & 0     & 0     & yes   & design variable \\
+    0     & 0     & 1     & yes   & collision \\
+    0     & 1     & 0     & yes   & design variable \\
+    0     & 1     & 1     & yes   & single input \\
+    1     & 1     & 0     & no    & hole \\
+    1     & 1     & 1     & yes   & single input \\
+    1     & 1     & $>1$ & no    & collision \\
+    1     & 2     & 1     & yes   & supplied input \\
+    1     & 2     & 2     & yes   & multi-fidelity \\
+    2     & 2     & $<2$ & no    & hole \\
+    2     & 2     & 2     & yes   & multi-fidelity \\
+    2     & 2     & $>2$     & no   & collision \\
+    \bottomrule
+    \end{tabular}%
+  \label{t:variable node classification}%
+\end{table}%

### 2013-05-08 mcg_fpg.tex
Major update

@@
-\section{Building MCG and FPG}
+\section{The MCG and FPG}
     \label{s:building graphs}
+We now discuss how the graph-based syntax defined in Sec.~\ref{s:syntax definition} is used to represent an MDAO problem in its entirety. This discussion is based on the MCG and the FPG.
+
 \subsection{Maximal Connectivity Graph}
     \label{ss:MCG}
-    To construct the maximal connectivity graph, we assume that a set of codes, global inputs, and objectives and constraints (collectively called global outputs) are provided. 
-    The codes are represented by analysis block graphs ((ABGs?)) $A_i=(V_{A_i},E_{A_i}), \ i\in I$, where $m$ is the number of codes and $I=\{1,2,\ldots,m\}$. 
-    The global inputs are represented as a set of variable nodes $V_\txt{in}$, and the global outputs are represented by a set of variable nodes $V_\txt{out}$. 
-    We assume that $V_\txt{in}$, $V_\txt{out}$, and each $A_i$ are given, and that any potential connection between variables is given in the form of connection edges in the set $C_M$. 
-    Then we may construct the maximal connectivity graph $M=(V_M,E_M)$ as
+For a specified MDAO problem, the maximal connectivity graph represents every analysis tool, objective, constraint, and global input being considered, as well as every possible interconnection among them.
+The definition of the MCG is given via construction.
+    To construct the maximal connectivity graph, we presume that a set of analysis tools, global inputs, objectives, and constraints are provided. 
+    Each of the $m$ analysis codes correspond to an index $i\in I_A$, $I_A=\{1,2,\ldots,m\}$, and are represented by an analysis block graph $G_{A_i}=(V_{A_i},E_{A_i})$.
+    Each of the $n$ expression blocks correspond to an index $i\in I_E$, $I_E=\{1,2,\ldots,n\}$, and are represented by an expression block graph $G_{E_i}=(V_{E_i},E_{E_i})$.
+    Finally, the global inputs are represented as a set of variable nodes $V_\txt{in}$.
+    We presume that $V_\txt{in}$, each $G_{A,i}$, and each $G_{E,i}$ are given, and that any potential connection between variables is specified in the form of connection edges in the set $E_{M,\txt{con}}$.
+One method for defining these connections is to use a consistant variable naming convention.
+    We may then construct the maximal connectivity graph $M=(V_M,E_M)$ as
     \begin{IEEEeqnarray*}{rCl}
-    V_M & = & V_\txt{in} \cup V_\txt{out} \cup \left( \bigcup_{i \in I} V_{A_i} \right), \\
-    E_M & = & C_M \cup \left( \bigcup_{i \in I} E_{A_i} \right),
+    V_M & = & V_\txt{in} \cup \left( \bigcup_{i \in I_A} V_{A_i} \right) \cup \left( \bigcup_{i \in I_E} V_{E_i} \right), \\
+    E_M & = & E_{M,\txt{con}} \cup \left( \bigcup_{i \in I_A} E_{A_i} \right)  \cup \left( \bigcup_{i \in I_E} E_{E_i} \right).
     \end{IEEEeqnarray*}
-    The MCG $M$ is uniquely determined by the given set of analysis blocks, the required outputs, and the given global inputs. 
+    The MCG $M$ is uniquely determined by the given set of analysis blocks, global inputs, and expression blocks. 
     In the cases where the set of global inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.
+
+The nodes and edges in $M$ can be partioned in sets according to their type:
+\begin{equation}
+V_M = V_{M,\txt{var}} \cup V_{M,\txt{mod}}, \ V_{M,\txt{var}} \cap V_{M,\txt{mod}} = \emptyset,
+\end{equation}
+where $V_{M,\txt{var}}$ and $V_{M,\txt{mod}}$ are the sets of variable nodes and model nodes, respecively;
+\begin{equation}
+E_M = E_{M,\txt{con}} \cup E_{M,\txt{mod}}, \ E_{M,\txt{con}} \cap E_{M,\txt{mod}} = \emptyset,
+\end{equation}
+where $ E_{M,\txt{con}}$ and $E_{M,\txt{mod}}$ are the sets of connection edges and model edges, respectively. These sets will be referenced in the process for obtaining an FPG.
 
@@
     \label{ss:FPG}
-    We now define the fundamental problem graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions, which are explained subsequently,
+    We now define the fundamental problem graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions:
     \begin{enumerate}
-    \item[(1)] $F \subset M$ and $V_\txt{out} \subset V_F$
-    \item[(2)] $\displaystyle{\forall i \in I, \txt{ if } F \cap A_i \neq \emptyset \txt{ then } A_i \subset F}$
-    \item[(3)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
-    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_\txt{out}}$
+    \item[(1)] $F \subset M$ and $G_{E_i} \subset F  \ \forall i \in I_E$
+%    \item[(1)] $F \subset M$ and $V_\txt{out} \subset V_F$
+    \item[(2)] $\displaystyle{\forall i \in I_A, \txt{ if } F \cap G_{A_i} \neq \emptyset \txt{ then } G_{A_i} \subset F}$
+%    \item[(2)] $\displaystyle{\forall i \in I, \txt{ if } F \cap A_i \neq \emptyset \txt{ then } A_i \subset F}$
+    \item[(3)] $\displaystyle{\forall v \in V_F \txt{ with }v \in V_{M,\txt{var}} \txt{ and }  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
+    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_{E_i}}$
+%    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_\txt{out}}$
     \end{enumerate}
@@
     %Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
-    Requirement (1) asserts that only the nodes and edges provided by the maximal connectivity graph can be used in the fundamental problem graph and that every global output must be included. 
-    Requirement (2) requires that each analysis block must be entirely present or entirely absent. 
-    Requirement (3) stipulates that the number of edges directed into each variable node must be within the lower and upper in-degree limits; 
-    if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is a \emph{collision}.
-    Lastly, requirement (4) ensures that only the nodes that are being used are included in the FPG by requiring that for every node a reverse path exists from at least one global output to that node.
+    Condition (1) asserts that only the nodes and edges provided by the maximal connectivity graph can be used in the fundamental problem graph and that every expression block must be included. 
+    Condition (2) requires that analysis blocks must be included or excluded in their entirety. 
+    Condition (3) stipulates that the number of edges directed into each variable node must be within the lower and upper in-degree limits; 
+    if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is the location of a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is the location of a \emph{collision}.
+    Lastly, condition (4) ensures that only the nodes that are being used are included in the FPG by requiring that for every node a reverse path exists from at least one expression block to that node.
     The reverse graph $R_F$ is obtained from $F$ by simply switching the orientation of every edge in $E_F$.
@@
 
-\subsection{Obtaining the Fundamental Problem Graph}
+\subsection{Algorithm for obtaining the Fundamental Problem Graph from the MCG}
     \label{ss:obtaining FPG}
-    In general, there may be multiple different graphs that satisfy the FPG conditions in Sec.~\ref{ss:FPG}, though there may be none at all. 
-    Here, we describe a process for obtaining an FPG by starting with the MCG and disconnecting connection edges until the FPG conditions are met. 
-    Then the problem is reduced to deciding which connection edges to remove.
-
-((describe pruning as a subroutine enforcing step 4 and then step 2))
+    In general, there may be zero, one, or many graphs that satisfy the FPG conditions in Sec.~\ref{ss:FPG}.
+    Here, we describe an algorithm for obtaining an FPG by starting with the MCG and disconnecting connection edges until the FPG conditions are met. 
+    With this approach, the problem is reduced to deciding which connection edges to remove.
+First, we describe a pruning subroutine used within the algorithm, and next we describe the algorithm itself.
 
     \begin{description}
-    \item[Subroutine: Pruning] 
-        Given a digraph $G_1 = (V_1,E_1)$, let the subroutine denoted as $P_\txt{prune}$ operate on $G_1$ to produce a digraph $G_2 = (V_2,E_2)$ that satisfies requirements (2) 
-        and (4), and write $G_2 = P_\txt{prune} ( G_1 )$. 
-        This is accomplished by first creating the reverse graph of $G_1$, $R = (V_1,E_R)$.
-Then adding a new node $x$ to $V_R$ and adding edges directed from this node to each of the global outputs, i.e.,
+    \item[\bf{Subroutine: Pruning}] 
+        Given a digraph $G_1 = (V_1,E_1)$, let the subroutine denoted as $P_\txt{prune}$ operate on $G_1$ to produce a new digraph $G_2$ as
 	\begin{equation}
-        \forall y \in V_\txt{out}, (x,y) \in E_R
+	G_2 = P_\txt{prune}(G_1) = (V_2,E_2),
 	\end{equation} 
-        The set of nodes that may be reached from the global outputs is constructed as
+	where $G_2$ now satisfies requirements (2) and (4) in Sec.~\ref{ss:FPG}.
+        Pruning is accomplished by first creating the reverse graph of $G_1$, $R = (V_1,E_R)$, where
 	\begin{equation}
-        U = \{ n \in V_1 \st \exists P \txt{ a path from $x$ to $n$}, \ P \subset R \}.
+	E_R = \{ (x,y) \st (y,x) \in E_1\}.
 	\end{equation}
-        The list of analysis blocks with at least one node that may be reach is constructed as
+	Next, a new node $b$ is added to $V_1$ and edges directed from this node to each of the expression blocks are added:
     \begin{equation}
-        I_1 = \{ i \in I \st V_{A_i} \cap U \neq \emptyset \}
+        \forall i \in I_E,\forall v \in V_{E_i}, \txt{ if } \txt{deg}^+(v)=0 \txt{ then } (b,v) \in E_R
     \end{equation}
-        The set of nodes to remove is
+        The set of nodes that may be reached from $b$ is constructed as
     \begin{equation}
-        N = \{ v \in V_1 \st v \in V_{A_i} \txt{ for } i \in I_1, \txt{ or } v \in V_\txt{in}\setminus U \},
+        U = \{ v \in V_1 \st \exists P \txt{ a path from $b$ to $v$}, \ P \subset R \}.
         \end{equation}
-        which means that $N$ is composed of unused analysis block nodes and unused global input nodes. Then the new set of nodes is created as
+	Because $b$ is directed into only the expression blocks, any path from $b$ necessarily provides a path from at least one expression block, which means the node is being used in the problem formulation.
+
+        The list of analysis blocks with at least one node in $U$ is constructed as
         \begin{equation}
-        V_2 = V_1 \setminus N,
+        I_2 = \{ i \in I_A \st V_{A_i} \cap U \neq \emptyset \}
         \end{equation}
-        and edges involving the removed nodes are also deleted
+        Any analysis block not in $I_2$ should be removed from the graph because none of its outputs contribute to the problem.
+Global inputs that are not being used are also removed.
+The set of nodes to remove is then
         \begin{equation}
-        E_2 = E_1 \setminus \{ (a,b) \st a \in N \txt{ or } b \in N  \}.
+        N = (V_\txt{in}\setminus U )  \cup \left( \bigcup_{i \notin I_2} V_{A_i} \right),
         \end{equation}
-        The set of connection edges can be obtained as
+and the new set of nodes is created as
         \begin{equation}
-        C_2 = \{(x,y) \in V_{2} \st \sim(x \in V_{A_i} \txt{ and } y \in V_{A_i} \txt{ for some } i \in I_1)\},
+        V_2 = V_1 \setminus N.
         \end{equation}
-        which means $C_2 \cap V_{A_i} = \emptyset  \ \forall i \in I_1$, and $E_2 = C_2 \cup \left( \bigcup_{i \in I_1} E_{A_i} \right)$.
+        Edges involving the removed nodes are also deleted as
+        \begin{equation}
+        E_2 = E_1 \setminus \{ (x,y) \st x \in N \txt{ or } y \in N  \}.
+        \end{equation}
+        The set of connection edges can be extracted by considering only edges whos endpoints are not in the same analysis block:
+        \begin{equation}
+        E_{2,\txt{con}} = \{(x,y) \in V_{2} \st \sim(x \in V_{A_i} \txt{ and } y \in V_{A_i} \txt{ for some } i \in I_2)\}.
+        \end{equation}
+        %which means $C_2 \cap V_{A_i} = \emptyset  \ \forall i \in I_1$, and $E_2 = C_2 \cup \left( \bigcup_{i \in I_1} E_{A_i} \right)$. 
+\end{description}
 
-    \item[Step 1: Holes] 
-        The first step is to detect holes and disconnect the connection edges following them. 
-        These connection edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs.
+\noindent The algorithm for obtaining an FPG can now be described as follows:
+\begin{description}
+    \item[\bf{Step 1: Holes}] 
+        The first step is to detect holes and disconnect the first set of connection edges downstream of them, as indicated in Fig.~\ref{f:hole}. 
+        These connection edges are removed because they represent output variables which cannot be determined because the analysis function does not have adequate inputs.
 
-        To start, the an initial graph is created as a pruned copy of the MCG
+        To begin the process, an initial graph is created as a pruned copy of the MCG
         \begin{equation}
@@
         \end{equation}
-        where $F_0 = (V_{F,0},E_{F,0})$, and $C_{F,0}$ is also obtained as described. 
-        The set of variable nodes which are holes is created as
+        where $F_0 = (V_{F_0},E_{F_0})$, and $C_{F_0,\txt{con}}$ is also obtained as described. 
+        The set of variable nodes which are holes is identified as
         \begin{equation}
-        H = \{v \in V_{F,0} \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
+        H = \{v \in V_{F_0} \st v \in V_{M,\txt{var}} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
         \end{equation}
-        which is the set of variable nodes with fewer incoming edges than are allowed by the lower in-degree limit.
-        Then the updated set of edges is created by removing the edges preceding or succeeding the analysis block:
+        which is the set of variable nodes with fewer incoming edges than are allowed by the lower indegree limit.
+        The updated set of connection edges is then created by removing the edges preceding or succeeding the analysis block:
         \begin{equation}
-        C_{F,1} = C_{F,0} \setminus \{(x,y) \in C_{F,0} \st x \in V_{A_i} \txt{ or } y \in V_{A_i}, \txt{ and } H\cap V_{A_i} \neq \emptyset \},
+        E_{F_1,\txt{con}} = E_{F_0,\txt{con}} \setminus \{(x,y) \in E_{F_0,\txt{con}} \st x \in V_{A_i} \txt{ or } y \in V_{A_i}, \txt{ and } H\cap V_{A_i} \neq \emptyset \}.
     \end{equation}
-    %deg-v is now recaculated for the new C. 
-    and this step is demonstrated by Fig.~\ref{f:hole}. Because removing these edges can create new holes, this step must be repeated until no more holes are found.
-        If this step every finds a global output to be a hole, meaning $V_\txt{out} \cap H \neq \emptyset$, then the algorithm stops because an FPG cannot be obtained. 
-        Otherwise, it is guaranteed that an FPG can be obtained.
+Because removing these edges can create new holes, this step must be repeated until no additional holes are found.
+        If the hole identification step identifies a variable node in an expression block as hole, meaning $V_{E_i} \cap H \neq \emptyset$ for some $i \in I_E$, then the algorithm terminates because an FPG cannot be obtained.
+
+If this step finishes without finding a hole in an expression block, it is guaranteed that an FPG can be obtained because $F_1$ now satisfies all four conditions from Sec.~\ref{ss:FPG} expect for (3), which requires there to be no holes or collisions in the graph.
+Since no more holes remain, and collisions can always be resolved without producing a hole, (see Table \ref{t:variable node classification}), then it is guaranteed than an FPG can be obtained.
+
     \begin{figure}[htb!]
@@
 
-    \item[Step 2: Collisions] 
-        The second step is to detect collisions and to then disconnect enough connection edges so that the collisions are resolved. 
-        The set of variable nodes which contain collisions is created as
+    \item[\bf{Step 2: Collisions}] 
+        The second step is to detect collisions and to disconnect precisely the number of connection edges required such that all collisions are resolved without introducting holes. 
+        The set of variable nodes at which collisions occur is 
     \begin{equation}
-    S_\txt{nodes} = \{v \in V_M \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
+    S_\txt{nodes} = \{v \in V_{M,\txt{var}} \st \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
     \end{equation}
-    For each collision node we can construct a set contained the edges directed in. The set containing all of these sets is constructed as
+    For each collision node we can construct a set containing the edges directed into the node. The set containing all of these sets is constructed as
     \begin{equation}
@@
     Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$. 
-    An example collision is shown in Fig.~\ref{f:collision} to indicate the definition of $S_{\txt{edges},j}$. 
+    An example collision is shown in Fig.~\ref{f:collision} to exemplify the definition of $S_{\txt{edges},j}$. 
     We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$ (which are sets). 
-    Then we may construct sets of edges as
+    We may then construct sets of edges as
     \begin{equation}
-    B_j = \big \{e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \txt{ with } K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
+    B_j = \big \{e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \txt{ with } \txt{deg}_u^-(v_j) \leq K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
     \end{equation}
-        which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper in--degree limit of $v_j$. 
-        The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include. Then let
+        which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper and lower indegree limits of $v_j$.  
+The construction of each $B_j$ corresponds to making a decision about which edges to include and which edges not to include. 
+%An upper bound on the number of such decisions that need to be made is 
+%\begin{equation}
+%\prod_{j \in J} 	
+%	\begin{pmatrix}
+%		|S_{\txt{edges},j}| \\
+%		\txt{deg}_u^-(v_j)
+%	\end{pmatrix}
+%\end{equation}
+%where $(:)$ refers to standard $n$-choose-$k$ notation.
+Let the new set of connection edges be
     \begin{equation}
-    C_{F,2} = \{ e \in C_{F,1} \st e \in B_j \txt{ for some } j \in J\},
+    E_{F_2,\txt{con}} = \{ e \in E_{F_1,\txt{con}} \st e \in B_j \txt{ for some } j \in J\}.
     \end{equation}
+The set of all edges is created by removing the connection edges not in $E_{F_2,\txt{con}}$:
     \begin{equation}
-    E_{F,2} = E_M \setminus (C_M \setminus C_{F,2})
+    E_{F,2} = E_{F_0} \setminus (E_{M,\txt{con}} \setminus E_{F_2,\txt{con}}),
     \end{equation}
-    and
+    which gives
     \begin{equation}
-    F_2 = (V_M,E_{F,2})
+    F_2 = (V_{F_0},E_{F,2}).
     \end{equation}
@@
 
-    \item[Step 3: Finalize] 
-        The third and final step is to prune the graph one last time
+    \item[\bf{Step 3: Finalize}] 
+        The third and final step is to prune the graph a final time to exclude any analysis blocks that became unneeded after the collisions were resolved:
     \begin{equation}
@@
     \end{equation}
-        and $F = (V_F,E_F)$.
-        The indices of the used analysis blocks can then be found as
+        with $F = (V_F,E_F)$.
+        The indices of the analysis blocks implemented in the FPG can then be found as
     \begin{equation}
-        I_F = \{ i \in I \st A_i \cap F \neq \emptyset \}
+        I_F = \{ i \in I_A \st G_{A_i} \subset F \}.
     \end{equation}
@@
 
-This process will always provide an FPG if one exists. If an FPG does not exist, this process will reveal the limiting factors preventing a valid formulation.
+This algorithm will always provide an FPG if one exists. If an FPG does not exist, the limiting factors preventing a valid formulation are revealed.
+
+\subsection{Suggested Process for Using the FPG Algorithm}
+\label{ss:process}
+Section \ref{ss:obtaining FPG} provided an algorithm for obtaining an FPG from the given MCG.
+However, if an FPG cannot be obtained, this algorithm can be applied as part of a larger process in which the designer changes the problem or the supplied analyses to attempt to obtain an FPG.
+The following steps detail the suggested procedure for obtaining an FPG:
+\begin{enumerate}
+\item[\bf{(A)}] Begin with a set of global inputs, analysis codes, objectives, and constraints.
+\item[\bf{(B)}] Build the MCG as described in Sec.~\ref{ss:MCG}.
+\item[\bf{(C)}] Set indegree limits for variable nodes representing local inputs as described in Sec.~\ref{s:indegree-outdegree} and Table \ref{t:variable node classification}
+\item[\bf{(D)}] Run the FPG algorithm described in Sec.~\ref{ss:obtaining FPG}. If a valid FPG is unattainable:
+		\begin{enumerate}
+\item Change the MCG by adding analysis blocks and/or global inputs, which means identifying new analyses to include in a potential MDAO workflow.
+\item Change indegree limits (see Table \ref{t:variable node classification}).
+\end{enumerate}
+\item[\bf{(E)}] Repeat from step (A) until an FPG is obtained.
+\end{enumerate}

### 2013-05-08 problem_formulation.tex
updated per brian's suggestions
Major update

@@
-\documentclass[referee]{svjour3} % insert '[draft]' option to show overfull boxes
+%\documentclass[referee]{svjour3} % insert '[draft]' option to show overfull boxes
+\documentclass[onecolumn]{svjour3} % insert '[draft]' option to show overfull boxes
 
@@
 \author{
-  David Pate, %
-  Justin Gray,%
-  Dr. Brian German 
+  David J. Pate, %
+  Justin Gray, %
+  Brian J. German 
  }
 
- \institute { Justin Gray \at NASA Glenn Research Center, Mail Stop 5-11, 21000 Brookpark Rd Clevland OH 44135 %
-  \and David Pate \at Gatech \and Dr. Brian German \at Gatech}
+ \institute { Justin Gray 
+	\at NASA Glenn Research Center, Mail Stop 5-11, 21000 
+		Brookpark Rd Clevland OH 44135 %
+	\and David J. Pate 
+	\at Graduate Research Assistant, Georgia Institute of Technology, 270 Ferst Drive, Atlanta, GA, 30332, U.S.A.
+	\and Brian German 
+	\at Assistant Professor, Georgia Institute of Technology, 270 Ferst Drive, Atlanta, GA, 30332, U.S.A.}
 
@@
 \begin{abstract}
-   blah blah blah
+The formulation of multidisciplinary design, analysis, and optimization (MDAO) problems has become increasingly complex as the number of analysis tools and design variables included in typical studies has grown.
+This growth in the scale and scope of MDAO problems has been motivated by the need to incorporate additional design disciplines and to expand the parametric design space to enable the exploration of unconventional design concepts.
+In this context, given a large set of disciplinary analysis tools, the problem of determining a feasible data flow between tools to produce a specified set of system-level outputs is combinatorically challenging.
+The difficulty is compounded in multi-fidelity problems, which are of increasing interest to the MDAO community.
+In this paper, we propose an approach for addressing this problem based on the formalism of graph theory.
+The approach begins by constructing the maximal connectivity graph (MCG) describing all possible interconnections between a set of analysis tools.
+Graph operations are then conducted to reduce the MCG to a fundamental problem graph (FPG) that describes the connectivity of analysis tools needed to solve a specified system-level design problem.
+The FPG does not predispose a particular solution procedure; any relevant MDO solution architecture could be selected to implement the optimization.
+The approach is applied to an example problem to formulate an FPG for a commercial aircraft MDAO study.
 \end{abstract}
@@
 \input{example_problem}
-
-\section{Applications}
-
-\section{Conclusions}
+\input{conclusions}
 

### 2013-05-08 conclusions.tex

\section*{Conclusions}
In this paper, we have presented an approach for describing MDAO problems with constructs and syntax from graph theory.  
Our graph description shares similarities to other approaches such as REMS, $\Psi$, FDT, DSM, and XDSM, but it provides new constructs tailored to applications in early phases of MDAO problem formulation before specific problem solution approaches are envisioned.  
In particular, we introduce the concepts of the Maximally Connected Graph (MCG) and the Fundamental Problem Graph (FPG).

The MCG addresses the question, ``Given a set of analysis tools, what are all of the variable interconnections between them that could be established?''  
The MCG provides a structured formalism to identify \textit{holes} and \textit{conflicts} in interconnections between input and output variables implied by the set of analysis tools.  
A hole corresponds to a variable that is not calculated in the MCG data flow but is needed by other tools.   
To create a valid problem formulation, all holes must be ``filled'' either by assigning the corresponding variable as a design variable to be set by the user or an optimizer, or by introducing additional tools to compute the variable.

A conflict corresponds to a situation in which two or more analysis tools compute separate instances of the same variable.  
To create a valid problem formulation, all conflicts must be resolved either by removing redundant analysis tools or by introducing logic or solvers into the data flow to choose between or aggregate multiple instances during execution.  
The identification of conflicts in an MCG is especially important for problem formulations involving multiples analysis tools of different fidelities; in this context, a conflict represents an opportunity to formulate a multi-fidelity problem that can take best advantage of the efficiency/accuracy tradeoff between different analysis tools.  
Resolving conflicts and filling holes in an MCG requires choices by the user to specify the overall MDAO problem that he/she desires to solve, i.e. to produce the required set of system-level outputs given a specified set of design variables.

The FPG is a graph that describes a data connection structure corresponding to an MDAO problem formulation free of conflicts and holes.  
An FPG is the result of user choices to fill holes and resolve conflicts in the initial MCG.  
One benefit of using graph theory is the standard algorithms, such as cycle detection, which can be used to inform the user's decisions.
Typically, many different FPGs could be attained from a given MCG, depending on the user choices.  
The number of possible FPGs that can be attained by selecting different design variables or introducing additional analysis tools to fill holes and/or resolve conflicts can be viewed as a measure of the freedom available to the user in implementing the available analysis tools to formulate valid MDAO problems.  
In Section 5, we present an algorithm for identifying holes and conflicts in an MCG that can be applied by the user in a process to arrive at valid FPGs.  
In Section 6, we provide an example application of formulating an FPG from an MCG for a commercial aircraft design problem.

Once a desired FPG is identified, the graph syntax described in the paper can be leveraged to formulate a Problem Solution Graph (PSG) describes particular MDAO solution architectures, e.g. MDF, IDF, BLISS, ATC, that can be used to solve the problem defined by the FPG.  
Many possible PSGs could be formulated to solve a problem corresponding to a specified FPG.   
The application of the syntax to PSGs has not been illustrated in the paper.

For simple problems with few analysis tools and variables, the formulation of a valid problem description is straightforward.  
However, MDAO problems continue to grow in scale and complexity.  
As the numbers of analysis tools and variables have increased, it has become increasingly challenging and time consuming for engineering teams to determine how the multiple analysis tools can be interconnected to produce valid problem formulations, to know when other tools must be introduced, and to determine the number of free variables in the problem that can or should be varied by the designer or an optimizer.  
The graph formalism presented in this paper is intended to offer value in this context of increasing problem complexity by providing a formal approach to identify and resolve missing and redundant data in order to create valid MDAO problem formulations.

### 2013-05-08 intro.tex

@@
     The number of analysis tools required in multidisciplinary design optimization (MDO) studies is growing
-    in parallel with the increasing scope of typical problems. An example of this growth in scope can
+    in parallel with the increasing scope of typical problems. An example can
     be observed in the historical evolution of the disciplines involved in MDO problems in aircraft design.
@@
     deformations, through the linking of aerodynamic vortex lattice or panel methods with structural finite
-    element models [[REF]]. Subsequently, flight performance and life cycle economics tools were integrated
-    into MDO analysis workflows for conceptual and preliminary design studies [[REF]]. Currently, MDO
+    element models \cite{Cramer1994}. Subsequently, flight performance and life cycle economics tools were integrated
+    into MDO analysis workflows for conceptual and preliminary design studies \cite{Sobieski1998}. Currently, MDO
     problems for aircraft design also often include tools for aircraft noise and emissions. In sum, it is now
-    commonplace for 5--10 analysis tools to be employed for typical aircraft design optimization studies [[REF]].
+    commonplace for 5--10 analysis tools to be employed for typical aircraft design optimization studies.
 
@@
     part, by consideration of additional disciplines. Current trends on the horizon for aircraft MDO studies
-    include incorporation of manufacturing analyses [[REF]], subsystem performance [[REF]], and fleet-level
-    aggregate models of emissions, noise, or economics [[REF]].
+    include incorporation of manufacturing analyses \cite{deWeck2007}, subsystem performance \cite{Dean2009,Gavel2006}, and models of emissions, noise, or economics \cite{Antoine2004,Rallabhandi2007,Kirby2008}.
 
@@
     Multidisciplinary Design Analysis and Optimization (MDAO)
-    frameworks such as OpenMDAO\cite{Gray2012} and \textregistered{ModelCenter} have enabled a new level of analysis tool integration 
-    and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. That 
+    frameworks such as OpenMDAO\cite{Gray2012} and {ModelCenter}\textregistered have enabled a new level of analysis tool integration 
+    and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. This 
     new capability has created a new challenge. Even models with tens of analyses could include hundreds or thousands
@@
     is complex, as measured by the number of coupled disciplines and/or components involved in the
-    analysis. A general optimization problem statement has been formulated based on system--level objectives and constraints; however, it is unclear which engineering analysis tools should
+    analysis. A general optimization problem statement has been formulated based on system-level objectives and constraints; however, it is unclear which engineering analysis tools should
     be interconnected in order to solve the optimization problem. A team of disciplinary and/or component design engineers has been formed in which each engineer has expertise in a
     particular analysis tool or component model. The engineers meet to discuss the approach to
-    interconnecting their tools to achieve the required system--level MDO model.
+    interconnecting their tools to achieve the required system-level MDO model.
     \end{quote}
@@
     of the concepts proposed in REMS with the goal of allowing the graphs to more closely interact with 
-    existing design processes and MDAO frameworks. The approach begins by 
+    existing design processes and MDAO frameworks.
+
+The approach begins by 
     constructing the \emph{maximal connectivity graph (MCG)} describing all possible
     interconnections between the analysis tools proposed by the engineers. Graph operations are then
-    conducted to reduce the MCG down to a \emph{fundamental problem graph (FPG)} that describes the minimum set of analysis
-    tools needed to solve the specified system--level design problem. ((DJP: The minimum set will be an FPG but not all FPGs are the minimum set)) The FPG does not predispose
+    conducted to reduce the MCG down to a \emph{fundamental problem graph (FPG)} that describes the set of analysis
+    tools needed to solve the specified system-level design problem.
     any particular solution procedure for the problem; any relevant MDO solution architecture, e.g. MDF,
-    IDF, CO, BLISS, could be selected \emph{post facto} to implement the system--level optimization by
+    IDF, CO, BLISS, could be selected \emph{post facto} to implement the system-level optimization by
     constructing a \emph{problem solution graph (PSG)} and the corresponding problem solution approach.
@@
     The paper is organized as follows. First, we describe the differences between a fundamental problem
-    formulation, which is based only on the system--level optimization problem statement that the
+    formulation, which is based only on the system-level optimization problem statement that the
     engineers desire to solve and on the available analysis tools, and a specific problem formulation, which
@@
 
-\section{Specific vs Fundamental Problem Formulation }
+\section{Motivation}
+\subsection{Specific vs Fundamental Problem Formulation }
 	\label{s:specific vs fundamental}
-    The Fundamental Problem Formulation (FPF) is our terminology for a statement of the overall system--level optimization problem that does not depend on the choice of solution--specifc elements such as the MDAO framework, optimization algorithm, iterative solver, or execution sequence
-    used to solve the problem. However, it is very easy to formulate a problem without 
-    any reference to a specific solution strategy, but still impose a a specific execution sequence.
+    The Fundamental Problem Formulation (FPF) is our terminology for a statement of the overall system-level optimization problem that does not depend on the choice of solution-specifc elements such as the MDAO framework, optimization algorithm, iterative solver, or execution sequence
+    used to solve the problem. The FPF contrasts with problem statements that are written without 
+    reference to a specific solution strategy, but that still imply a specific execution sequence.
     For example, consider the Sellar test problem \cite{AIAA:sellar}: 
-
-((can we either make the $t$ a subscript or wrap parenthesis around it, i.e. $y^{(t)}$?))
-
     \begin{align}
-        given & \ \ y_1 = D_1(x_1,\hat{y_2},z_1,z_2) \notag
+        \txt{given} & \ \ y_1 = D_1(x_1,\hat{y_2},z_1,z_2) \notag
         \\      & \ \ y_2 = D_2(y_1,z_1,z_2) \notag
-        \\min. &\ \ F(x_1,y_1,y_2,z_2) \notag
-        \\w.r.t. & \ \ x_1,y_1,\hat{y_2},z_1,z_2 \notag
-        \\s.t. & \ \ H(y_2,\hat{y_2}) = 0 \notag 
+        \\\txt{min.} &\ \ F(x_1,y_1,y_2,z_2) \notag
+        \\\txt{w.r.t.} & \ \ x_1,y_1,\hat{y_2},z_1,z_2 \notag
+        \\\txt{s.t.} & \ \ H(y_2,\hat{y_2}) = 0 \notag 
         \\     & \ \ G_1(y_1) \geq 0 \notag
@@
 
-    $D_1$ and $D_2$ represent analysis tools, $F$, $G_1$, and $G_2$ are the objective 
+    Here, $D_1$ and $D_2$ represent analysis tools, $F$, $G_1$, and $G_2$ are the objective 
     and constraint functions respectively, and $H$ is the coupling constraint. 
-    Equation \ref{eqn:simple} makes an inherent assumption about the execution 
+    Equation \ref{eqn:simple} implies the execution 
     order for the problem. Analysis $D_1$ outputs $y_1$, which is an input to $D_2$. 
-    Hence, $D_1$ should be run before $D_2$ with with $y_2$ being iterated on 
-    to convergence with $\hat{y_2}$. However, a slightly different formulation 
-    is equally valid and still represents the exact same problem:
+    Hence, $D_1$ should be run before $D_2$ with with $y_2$ iterating
+    to convergence with $\hat{y_2}$. A slightly different formulation 
+    is equally valid and still represents the same problem:
 
     \begin{align}
-        given & \ \ y_2 = D_2(\hat{y_1},z_1,z_2) \notag
+        \txt{given} & \ \ y_2 = D_2(\hat{y_1},z_1,z_2) \notag
         \\      & \ \ y_1 = D_1(x_1,y_2,z_1,z_2) \notag
-        \\min. &\ \ F(x_1,y_1,y_2,z_2) \notag
-        \\w.r.t. & \ \ x_1,\hat{y_1},y_2,z_1,z_2 \notag
-        \\s.t. & \ \ H(y_1,\hat{y_1}) = 0 \notag 
+        \\\txt{min.} &\ \ F(x_1,y_1,y_2,z_2) \notag
+        \\\txt{w.r.t.} & \ \ x_1,\hat{y_1},y_2,z_1,z_2 \notag
+        \\\txt{s.t.} & \ \ H(y_1,\hat{y_1}) = 0 \notag 
         \\     & \ \ G_1(y_1) \geq 0 \notag
@@
     Notice that the formulation from Eq. \ref{eqn:simple} includes 
-    $\hat{y_2}$, while Eq. \ref{eqn:simple2} includes $\hat{y_2}$. The result is that
+    $\hat{y_2}$, while Eq. \ref{eqn:simple2} includes $\hat{y_1}$. The result is that
     $D_1$ is now dependent on the output of $D_2$, and $y_1$ will be iterated on to 
     convergence with $\hat{y_1}$. Now the problem has to be solved by running $D_2$ first 
-    and then $D_1$. Since the formulations in Eqs. \ref{eqn:simple} and 
-    \ref{eqn:simple2} describe the same problem then neither can be the
-    FPF. They are specific alterative expressions of a more fundamental problem description 
+    and then $D_1$.
+
+Equations \ref{eqn:simple} and 
+    \ref{eqn:simple2}  are specific alterative expressions of a more fundamental problem description 
     that we call the FPF:
@@
     \begin{align}
-        given & \ \ y_1 = D_1(x_1,\hat{y_2},z_1,z_2) \notag
+        \txt{given} & \ \ y_1 = D_1(x_1,\hat{y_2},z_1,z_2) \notag
         \\      & \ \ y_2 = D_2(\hat{y_1},z_1,z_2) \notag
-        \\min. &\ \ F(x_1,y_1,y_2,z_2) \notag
-        \\w.r.t. & \ \ x_1,\hat{y_1},\hat{y_2},z_1,z_2 \notag
-        \\s.t. & \ \ H_1(y_1,\hat{y_1}) = 0 \notag
+        \\\txt{min.} &\ \ F(x_1,y_1,y_2,z_2) \notag
+        \\\txt{w.r.t.} & \ \ x_1,\hat{y_1},\hat{y_2},z_1,z_2 \notag
+        \\\txt{s.t.} & \ \ H_1(y_1,\hat{y_1}) = 0 \notag
         \\     & \ \ H_2(y_2,\hat{y_2}) = 0 \notag 
@@
 
-    The formulation in Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} 
-    and Eq. \ref{eqn:simple2} because it has two coupling constraints which both must be 
-    met. The result is that Eq. \ref{eqn:simple_fpf} required the presence of both 
+    The formulation of Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} 
+    and Eq. \ref{eqn:simple2} because it includes two coupling constraints. 
+The result is that Eq. \ref{eqn:simple_fpf} requires the presence of both 
     $\hat{y_1}$ and $\hat{y_2}$, which means that $D_1$ and $D_2$ each now operate with 
-    their own local copies of the coupling variables. This fully decouples the problem so that 
+    their own local copies of the coupling variables. This formulation fully decouples the problem so that 
     either $D_1$ or $D_2$ could be run first or both could be run simultaneously. 
+\begin{comment}
     Alexandrov and Lewis, in their work with REMS, assert that fully decoupling 
@@
     from the basic chain rule.
+\end{comment}
 
-\section{Existing Graph-Based Syntax}
+Figure \ref{f:sellar graph} depicts the Sellar problem given by Eq.~\ref{eqn:simple_fpf} in the graph-based syntax which will be presented in this paper.
+This example will serve to highlight the features of the syntax as they are introduced.
+\begin{figure}[htb!]
+  \begin{center}
+    \includegraphics[width=4.0in]{images/sellar}
+  \end{center}
+  \caption{Sellar problem represented as a graph.}
+\label{f:sellar graph}
+\end{figure}
+
+\subsection{Existing Graph-Based Syntax}
 	\label{s:existing syntax}
@@
 
-    A number of graph--based methods have been used successfully to translate the 
+    A number of graph-based methods have been used successfully to translate the 
     mathematical syntax into a more useful computational form. 
     Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools.
-    Off--diagonal elements of the matrix indicate coupling\cite{Steward1981}. Since a DSM describes a square adjacency matrix, 
-    it can be represented in an equivalent directed graph where nodes represent analysis tools and 
+    Off-diagonal elements of the matrix indicate coupling\cite{Steward1981}. Since a DSM describes a square adjacency matrix, 
+    it can be represented in an equivalent directed graph in which nodes represent analysis tools and 
     edges represent information dependence between those tools. The ordering of elements in a DSM can be used to indicate 
-    execution order.  For more complex problems, choosing the proper order to run analysis tools is a non--trival task. 
+    execution order.  For more complex problems, choosing the proper order to run analysis tools is a challenging task. 
     Rogers et al.~developed DeMAID to manipulate a DSM to find an ordering for analysis tools that 
@@
     A DSM by itself is insufficient to describe complete problem formulations because it captures only information about data dependency between analyses; objective and constraint information is missing from the DSM description of the problem.
-    An alternative matrix--based syntax, called a Functional Dependency Table (FDT), was proposed by Wagner and Papalambros. 
+    An alternative matrix-based syntax, called a Functional Dependency Table (FDT), was proposed by Wagner and Papalambros. 
     FDT represents the relationship between functions, including objectives and constraints, and specific variables that affect 
@@
     information about objectives and constraints, its lack of directed edges 
-    implies that it cannot capture the coupled data dependency that a DSM captures. For example, 
-    the objective, $f$, from the FPF in Eq. \ref{eqn:simple_fpf} is dependent on 
-    output $m$ of analysis $A$. The dependence of $m$ on $A$ is now captured 
+    implies that it cannot capture the coupled data dependency that a DSM captures. 
+For example, the objective, $f$, from the FPF in Eq. \ref{eqn:simple_fpf} is dependent on 
+    output $m$ of analysis $A$ ((BG: I don't quite get this)). The dependence of $m$ on $A$ is now captured 
     in FDT in Fig. \ref{fig:FDT_simple} alone. This missing information implies 
     that, while FDT is very useful for partitioning problems, it is also not 
-    sufficient to contain a complete problem formulation.
-
-    \begin{figure}
+    sufficient to completely describe a problem formulation. 
+    \begin{figure}[htb!]
         \begin{center}
@@
     definition of MDAO architectures. Martins uses XDSM to describe 13 different 
-    optimziation architectures in a survey paper that provides a novel 
+    optimization architectures in a survey paper that provides a novel 
     classification of the different techniques \cite{Lambe2011}. With the 
@@
     assumed Gauss-Seidel iteration scheme and an MDF solution architecture.
-    Hence, the XDSM syntax description is too specific for use with a fundamental problem formulation.
 
@@
     problem formulation, but lacks the ability to represent optimizers and 
-    solvers in more specific problem formulations. While XDSM does include 
-    optimizers and constraints, it relies on their presence is not formulated  
+    solvers in more specific problem formulations. XDSM relies on the provision of  
+    optimizers and constraints and is not formulated  
     for a more fundamental problem statement. We propose a new graph syntax that 
-    combines the key features of REMS and XDSM while retaining the flexiblity to 
+    combines the relevant features of REMS and XDSM while retaining the flexiblity to 
     represent a fundamental problem formulation free of solution specific information.
+
+\subsection{Requirements for a New Graph Syntax}
+  \label{s:requirements}
+  The goal of the new graph-based syntax presented here is to enable the general structure of an MDAO problem to be described independently of any solution information, 
+  while still being able to accomodate the more specific case when a solution 
+  strategy is applied. In order to achieve this goal, 
+  the graph syntax needs to accomodate a number of MDAO problem constructs: 
+  \begin{itemize}
+    \item Analysis tools and their interconnections
+    \item Design variables, objectives, and constraints
+    \item Local and global information
+    \item Coupling between analyses
+    \item Multi-fidelity analyses
+  \end{itemize}
+
+  The new syntax is intended to represent three phases of the design problem 
+  formulation process. In the initial problem definition phase, the specific 
+  analysis tools and design goals are identified. Next, a single formal problem 
+  formulation is identified that specifies design variables, constraints, 
+  objectives, analysis tools, and all other elements required to represent 
+  the overall MDAO problem. Lastly, a specific procedure for solving the problem 
+  is selected, e.g. selecting an MDAO optimization architecture. Using 
+  the proposed graph syntax, these phases can be represented with the following 
+  three graphs:
+  \begin{itemize}
+    \item Maximal Connectivity Graph (MCG)
+    \item Fundamental Problem Graph (FPG)
+    \item Problem Solution Graph (PSG)
+  \end{itemize}
+
+  The \emph{maximal connectivity graph} represents the first phase of the problem formulation with all 
+  analysis tools being considered and all possible connections between them also present. The second graph 
+  is the \emph{fundamental problem graph}, which is comprises only the analyses participating to solve the problem. Finally, a \emph{problem solution graph} 
+  may be represented by including additional edges and nodes to represent the 
+  solution strategy being employed to solve the problem. This paper focuses on the MCG and FPG and does not describe the PSG in detail.
+
+  The relationship between these three graphs is depicted in Figs.~\ref{f:tree} and \ref{f:hourglass}. 
+  The tree diagram demonstrates the fact that it is generally possible to obtain 
+  multiple FPGs from a single maximal connectivity graph. The multiple FPGs may correspond to 
+  different down-selections of analysis tools, different connections between the tools, 
+  or both. Each down-selection shrinks the number of possible FPGs that could be reached 
+  until only one remainins. For a given FPG, different PSGs may be obtained by implementing 
+  different solution strategies. Considering the size of a graph to be the sum of all of its
+  edges and nodes, the hourglass shape in Fig. \ref{f:hourglass} qualitatively illustrates how
+  the FPG is obtained from the MCG by \emph{removing} nodes and edges, 
+  and the PSG is obtained from the FPG by \emph{adding} nodes and edges.
+  \begin{figure}[htb!]
+    \centering
+    \subfigure[number of possible graphs]{
+    \includegraphics[width=1.75in]{images/tree}
+    \label{f:tree}
+    }
+    \subfigure[graph size]{
+    \includegraphics[width=1.75in]{images/hourglass}
+    \label{f:hourglass}
+    }
+  \caption{The relationship between the MCG, FPG, and PSG.}
+  \end{figure}

### 2013-05-09 graph_syntax.tex
another round of edits

@@
 
-\subsection{Local vs Global Nodes ((recommend omission))}
+\subsection{Local vs Global Nodes }
 
-  In this graph-based syntax, a variable node is considered \emph{local} if it has exactly one outgoing connection edge or \emph{global} if it has more than one outgoing connection edge.
-Therefore, a variable node may be used to represent a global input if it has at least two outgoing connection edges.
-Otherwise, the input is formulated as a design variable that is local to a specific analysis block.
-%  The set of local nodes, $V_\txt{l}$, is the collection of 
-%  nodes with exactly one outgoing connection edge. 
-%
-%  \begin{equation}
-%    V_{\txt{l}} = \{V_{\txt{var}} \st \txt{deg}^+(v)=1\}
-%    \label{e:local_definition}
-%  \end{equation}
-%  In section \ref{s:specific vs fundamental} variable nodes that are part of 
-%  an analysis block are inherently local and similarly model nodes 
-%  are also local. Following the definition in Eq. \ref{e:local_definition}
-%  we can now see why. Variable nodes connected to model nodes are only allowed to 
-%  have one outgoing edge, and hence will always by definition be local variables.
+  All global variable nodes have more than one outgoing edge. Conversely all local 
+  variable nodes have only a single outgoing edge. Fig.~\ref{f:sellar graph} contains 
+  three global variables that are not present in the problem formulation given by 
+  Eq.~\ref{eqn:simple_fpf}: $x_1^*$, $z_1^*$, and $z_2^*$. The presence of the 
+  new nodes is required in order to represent the equality of the given variables that
+  was implied by their shared names in Eq.~\ref{eqn:simple_fpf}.
 
-%  The set of global nodes, $V_\txt{g}$, is the collection of all nodes with more than 
-%  one outgoing connection edge. (())
-%
-%  \begin{equation}
-%    V_{\txt{g}} = \{V_{\txt{var}} \st \txt{deg}^+(v)>1\}
-%    \label{e:global_definition}
-%  \end{equation}
-
-%  Following our definition any variable node that had multiple 
-%  outgoing edges, is global. In Fig. \ref{f:sellar_graph_full} 
-%  the variable nodes for $x_1^*$, $z_1^*$, and $z_2^*$ have multiple outgoing 
-%  edges and would be considered global. 
-In the Sellar problem (Fig.~\ref{f:sellar graph}) the variable nodes $x_1^*$, $z_1^*$, and $z_2^*$ have multiple outgoing edges and are therefore considered global inputs. 
-  The variable $x_1$ is usually refered to as 
-  a local variable, which is contradicted by the given graph ((DP: can we cite someone calling this local?)). This discrepancy 
+  The variable $x_1$ is usually refered to as a local variable, which is 
+  contradicted by the given graph \cite{sellar1996}. This discrepancy 
   arises from the explicit treatment of objectives and constraints as separate 
-  expression blocks. This variable is often considered a local 
+  expression blocks. This variable is could be considered a local 
   variable because it directly affects only analysis block $A_1$. However, by expanding
@@
   This classification of $x_1^*$ in the Sellar problem
-  as a global variable agrees with the structure of the Collaborative 
+  as a global variable works well within the structure of the Collaborative 
   Optimization (CO) archtiecture \cite{braun1996thesis}. The rules for constructing 
@@
 
-  Analysis and expression blocks are also regarded local or global but the 
-  criteria for determining this is slightly different than for variable nodes---both incoming and outgoing edges must be considered.
-%  In this case, you consider the set of connection edges for the block
-%  (e.g. from all the variable nodes within the block) as a set. Additionally, 
-%, whereas for variables
-%  only outgoing connection edges are relevant. 
-  An analysis block is local if all incoming edges are from exactly one analysis 
-  block and all outgoing edges are from exactly one other analysis block. A global 
-  analysis block is defined as a block with incoming edges or outgoing edges 
-  to more than one analysis block.   In the Sellar problem, the expression blocks for each constraint, 
-  $G_1$ and $G_2$, are local to their respective analysis 
-  blocks, whereas the expression block for $F$ has incoming edges from $A_1$ and 
-  $A_2$, therefore it is considered global.
-
-  Catagorizing any node or analysis block with incoming or outgoing edges
-  from more than one source as global neglects a more 
-  subtle aspect of some MDAO problems. Namley the locality of any node in a graph is 
-  a relative property. For instance, a single node might have outgoing edges 
-  to two separate analysis blocks, $A$ and $B$, but none to a third, $C$. This 
-  node was global with respect to $A$ and $B$ separately, but local 
-  to the group $A,B$. This situation produces a natural hierarchy in a graph, 
-  from which one could form ever smaller groups as the problem is segmented into 
-  finer localities.
-
-  When solving a problem using a hierarchical decomposition approach, 
-  there are various techniques for partitioning a problem into hierarchical 
-  localities that are effective in different scenarios 
-  \cite{krishnamachari1997optimal,michelena1997hypergraph,sobieszczanski1997,Perez2004,allison2009optimal}. 
-  Tosserams \textit{et al.} proposed a language based-syntax for describing the 
-  partitioning of problems, named $\Psi$\cite{tosserams2010specification}. 
-  $\Psi$ provides an opposing perspective to this work, where they compose 
-  hierarchies from the bottom up, into larger and larger groups. However, 
-  $\Psi$ provides a compiler that can produce a standard problem representation 
-  of the assembled problem and a number of example converters from that 
-  standard representation to application specific formats.
-
 \subsection{Coupling Between Analyses}
-In MDAO, coupling is the mutual dependence the outputs of two or more analysis codes. In this graph-based syntax, coupling is described by the presence of a cycle between two or more analysis blocks. 
-%A path, given by $P=(V_\txt{path},E_\txt{path})$, contains a cycle if at least two edges in the path refrence the same node,  $E = \{(x_0,x_1),(x_1,x_2),\ldots,(x_{k-1},x_k),(x_k,x_0)\}$. 
-%Cycles decribe the reciprical dependence of a set of two or more analyses on each other's outputs. 
-In the Sellar problem from Eqn \ref{eqn:simple_fpf} the two 
-  coupling constraints, $H_1$ and $H_2$, provide a reciprical dependence between 
-  $A_1$ and $A_2$. In a graph, these constraints appear as connection edges between 
-  the outputs of each analysis block. The edges belonging to a cycle in the Sellar problem are highlighted in 
-  Fig. \ref{f:sellar cycles}, where they form ((sentence is unfinished))
+  In MDAO, coupling is the mutual dependence the outputs of two or more analysis 
+  codes. In this graph-based syntax, coupling is described by the presence of a 
+  cycle between two or more analysis blocks. In the Sellar problem from Eqn \ref{eqn:simple_fpf} 
+  a reciprical dependence between $D_1$ and $D_2$ exists through the variabled $y_1$ and $y_2$. 
+  The edges belonging to a cycle in the Sellar problem are highlighted in 
+  Fig. \ref{f:sellar cycles} where they form a closed path between analysis blocks $D_1$ and $D_2$.
 
@@
     \end{center}
-  \caption{Sellar problem graph with dashed edges indicating participation in a cycle.}
+    \caption{Sellar problem graph with dashed edges indicating participation in a cycle.
     \label{f:sellar cycles}
+    }
   \end{figure}
@@
   \label{ss:multi-fideliy problems}
-  A multi-fidelity MDAO problem can be characterized by a formulation in which two or more different analyses 
-  each calculating the same data.
-Multi-fidelity is represented in a graph by a variable node having an indegree greater than one, which means that multiple connection edges are directed into it.
-When the upper indegree limit of a variable node is set above one, then the node is implied to allow multiple fidelity instances of the variable.
-When the lower indegree limit of a variable node is set above one, then multiple fidelity instances are required..
-%  variables, $V_{mf}$, is given as: 
-%
-%  \begin{equation}
-%    V_{\txt{mf}} = \{ V_{\txt{var}} \st 1 < \txt{deg}^-(v) \leq \txt{deg}^-_u(v) \}
-%  \end{equation}
+  A multi-fidelity MDAO problem can be characterized by a formulation in which 
+  two or more different analyses each calculating the same data. Multi-fidelity 
+  is represented in a graph by a variable node having an indegree greater than 
+  one, which means that multiple connection edges are directed into it. When the 
+  upper indegree limit of a variable node is set above one, then the node is 
+  implied to allow multiple fidelity instances of the variable. When the lower 
+  indegree limit of a variable node is set above one, then multiple fidelity instances are required.
 

### 2013-05-09 intro.tex
another round of edits

@@
     tools needed to solve the specified system-level design problem.
-    any particular solution procedure for the problem; any relevant MDO solution architecture, e.g. MDF,
+    Any particular solution procedure for the problem; any relevant MDO solution architecture, e.g. MDF,
     IDF, CO, BLISS, could be selected \emph{post facto} to implement the system-level optimization by
@@
 	\label{s:specific vs fundamental}
-    The Fundamental Problem Formulation (FPF) is our terminology for a statement of the overall system-level optimization problem that does not depend on the choice of solution-specifc elements such as the MDAO framework, optimization algorithm, iterative solver, or execution sequence
+    The Fundamental Problem Formulation (FPF) is our terminology for a 
+    statement of the overall system-level optimization problem that does not 
+    depend on the choice of solution-specifc elements such as the MDAO 
+    framework, optimization algorithm, iterative solver, or execution sequence
     used to solve the problem. The FPF contrasts with problem statements that are written without 
     reference to a specific solution strategy, but that still imply a specific execution sequence.
-    For example, consider the Sellar test problem \cite{AIAA:sellar}: 
+    For example, consider the Sellar test problem \cite{AIAA:sellar}, with an FPF given as follows:
+
+    \begin{align}
+        \txt{given} & \ \ y_1 = D_1(x_1,y_2,z_1,z_2) \notag
+        \\      & \ \ y_2 = D_2(y_1,z_1,z_2) \notag
+        \\\txt{min.} &\ \ F(x_1,y_1,y_2,z_2) \notag
+        \\\txt{w.r.t.} & \ \ x_1,y_1,y_2,z_1,z_2 \notag
+        \\\txt{s.t.} & \ \ G_1(y_1) \geq 0 \notag
+        \\     & \ \ G_2(y_2) \geq 0
+        \label{eqn:simple_fpf}
+    \end{align}
+
+    In Eq. \ref{eqn:simple_fpf}, $y_1$ is an output of $D_1$ as well as an input 
+    to $D_2$. Similarly, $y_2$ is an output of $D_2$ as well as an input 
+    to $D_1$. This implies that $D_1$ and $D_2$ are coupled together by means 
+    of their reciprical dependence. However no information is given as to weather
+    $D_1$ or $D_2$ should be run before the other. Given Eq \ref{eqn:simple_fpf}, 
+    it would be equally valid to run $D_1$ first, $D_2$ first, or both simultaneously.
+
+    Eq. \ref{eqn:simple} gives a modified problem formulation where $D_1$ is run 
+    first. A new variable is introduced, $\hat{y_2}$, to break the dependence of
+    $D_1$ on $D_2$. Also a new coupling constraint, $H$, to enforce the consistency
+    that would otherwise be lost due to $\hat{y_2}$. Equation \ref{eqn:simple} is 
+    a specific alternative expression of the more fundamental problem description 
+    given in Eq. \ref{eqn:simple_fpf}.
+
     \begin{align}
@@
 
-    Here, $D_1$ and $D_2$ represent analysis tools, $F$, $G_1$, and $G_2$ are the objective 
-    and constraint functions respectively, and $H$ is the coupling constraint. 
-    Equation \ref{eqn:simple} implies the execution 
-    order for the problem. Analysis $D_1$ outputs $y_1$, which is an input to $D_2$. 
-    Hence, $D_1$ should be run before $D_2$ with with $y_2$ iterating
-    to convergence with $\hat{y_2}$. A slightly different formulation 
-    is equally valid and still represents the same problem:
-
-    \begin{align}
-        \txt{given} & \ \ y_2 = D_2(\hat{y_1},z_1,z_2) \notag
-        \\      & \ \ y_1 = D_1(x_1,y_2,z_1,z_2) \notag
-        \\\txt{min.} &\ \ F(x_1,y_1,y_2,z_2) \notag
-        \\\txt{w.r.t.} & \ \ x_1,\hat{y_1},y_2,z_1,z_2 \notag
-        \\\txt{s.t.} & \ \ H(y_1,\hat{y_1}) = 0 \notag 
-        \\     & \ \ G_1(y_1) \geq 0 \notag
-        \\     & \ \ G_2(y_2) \geq 0
-        \label{eqn:simple2}
-    \end{align}
-
-    Equation \ref{eqn:simple2} differs only slightly from Eq. \ref{eqn:simple}. 
-    Notice that the formulation from Eq. \ref{eqn:simple} includes 
-    $\hat{y_2}$, while Eq. \ref{eqn:simple2} includes $\hat{y_1}$. The result is that
-    $D_1$ is now dependent on the output of $D_2$, and $y_1$ will be iterated on to 
-    convergence with $\hat{y_1}$. Now the problem has to be solved by running $D_2$ first 
-    and then $D_1$.
-
-Equations \ref{eqn:simple} and 
-    \ref{eqn:simple2}  are specific alterative expressions of a more fundamental problem description 
-    that we call the FPF:
-
-    \begin{align}
-        \txt{given} & \ \ y_1 = D_1(x_1,\hat{y_2},z_1,z_2) \notag
-        \\      & \ \ y_2 = D_2(\hat{y_1},z_1,z_2) \notag
-        \\\txt{min.} &\ \ F(x_1,y_1,y_2,z_2) \notag
-        \\\txt{w.r.t.} & \ \ x_1,\hat{y_1},\hat{y_2},z_1,z_2 \notag
-        \\\txt{s.t.} & \ \ H_1(y_1,\hat{y_1}) = 0 \notag
-        \\     & \ \ H_2(y_2,\hat{y_2}) = 0 \notag 
-        \\     & \ \ G_1(y_1) \geq 0 \notag
-        \\     & \ \ G_2(y_2) \geq 0
-        \label{eqn:simple_fpf}
-    \end{align}
-
-    The formulation of Eq. \ref{eqn:simple_fpf} differs from both Eq. \ref{eqn:simple} 
-    and Eq. \ref{eqn:simple2} because it includes two coupling constraints. 
-The result is that Eq. \ref{eqn:simple_fpf} requires the presence of both 
-    $\hat{y_1}$ and $\hat{y_2}$, which means that $D_1$ and $D_2$ each now operate with 
-    their own local copies of the coupling variables. This formulation fully decouples the problem so that 
-    either $D_1$ or $D_2$ could be run first or both could be run simultaneously. 
-\begin{comment}
-    Alexandrov and Lewis, in their work with REMS, assert that fully decoupling 
-    the analyses is key to achieving reconfigurability in the problem 
-    formulation\cite{alexandrov2004}. The key feature they identify to allow this decoupling 
-    is the creation of independent variables for each analysis. Tosserams et. al. 
-    in their development of the $\psi$ language for heirarchical problem composition
-    make the same requirement that all analyses get their own local copies of 
-    variables in order to use them as the fundamental building blocks for 
-    problems\cite{tosserams2010specification}. Martins and Hwang derive the same requirement,
-    from a different perspective, in their work on computing analytic derivatives 
-    for MDAO problems\cite{Martins2012}. They determined that local copies of all variables 
-    allows for derivation of all relevant methods for calculating analytic derivatives 
-    from the basic chain rule.
-\end{comment}
-
-Figure \ref{f:sellar graph} depicts the Sellar problem given by Eq.~\ref{eqn:simple_fpf} in the graph-based syntax which will be presented in this paper.
-This example will serve to highlight the features of the syntax as they are introduced.
+    Figure \ref{f:sellar graph} depicts the Sellar problem given by 
+    Eq.~\ref{eqn:simple_fpf} in the graph-based syntax which will be presented 
+    in this paper. This example will serve to highlight the features of the 
+    syntax as they are introduced.
     \begin{figure}[htb!]
@@
 	\label{s:existing syntax}
-    As indicated in the examples in Sec.~\ref{s:specific vs fundamental}, the mathematical language for specifying problem formulations is very general and can be used both for 
+    As indicated in the examples in Sec.~\ref{s:specific vs fundamental}, the mathematical 
+    language for specifying problem formulations is very general and can be used both for 
     fundamental and specific problem formulations. Tedford and Martins used this mathematical syntax to specify the 
@@
 
-    A DSM by itself is insufficient to describe complete problem formulations because it captures only information about data dependency between analyses; objective and constraint information is missing from the DSM description of the problem.
+    A DSM by itself is insufficient to describe complete problem formulations because it 
+    captures only information about data dependency between analyses; objective and 
+    constraint information is missing from the DSM description of the problem. 
     An alternative matrix-based syntax, called a Functional Dependency Table (FDT), was proposed by Wagner and Papalambros. 
@@
     implies that it cannot capture the coupled data dependency that a DSM captures. 
-For example, the objective, $f$, from the FPF in Eq. \ref{eqn:simple_fpf} is dependent on 
-    output $m$ of analysis $A$ ((BG: I don't quite get this)). The dependence of $m$ on $A$ is now captured 
-    in FDT in Fig. \ref{fig:FDT_simple} alone. This missing information implies 
+    The FDT in Fig. \ref{fig:FDT_simple} shows that $D_1$ is dependent on $y_2$ and 
+    that $D_2$ is dependent on $y_1$, but you can not infer their coupled dependence from 
+    that information alone. This missing information implies 
     that, while FDT is very useful for partitioning problems, it is also not 
@@
             \hline
-                 & $x$ & $y$ & $y^t$ & $z$ & $z^t$ & m \\ \hline
-            $A$  & 1  & 1    &       &     &       &   \\ \hline
-            $B$  & 1  &      &       & 1   &       &   \\ \hline
-            $f$  &    &      &       &     &       & 1 \\ \hline
-            $g1$ &    &      &       & 1   & 1     &   \\ \hline
-            $g2$ &    & 1    & 1     &     &       &   \\
+                   & $x_1$ & $y_1$ & $y_2$ & $z_1$ & $z_2$ \\ \hline
+            $D_1$  & 1     &       & 1     & 1     & 1     \\ \hline
+            $D_2$  &       & 1     &       & 1     & 1     \\ \hline
+            $F$    & 1     & 1     & 1     &       & 1     \\ \hline
+            $G_1$  &       & 1     &       &       &       \\ \hline
+            $G_2$  &       &       & 1     &       &       \\
             \hline
@@
         \includegraphics[height=.25\textheight]{XDSM/simple}
-        \caption{XDSM for Eq. \ref{eqn:simple}, with a Gauss-Seidel iteration and MDF solution architecture. \label{fig:XDSM_simple}}
+        \caption{XDSM for Eq. \ref{eqn:simple}, with a Gauss-Seidel iteration 
+          and MDF solution architecture. \label{fig:XDSM_simple}}
         \end{center}

### 2013-05-09 graph_syntax.tex
removing old Fig 1.

@@
   graphs for all three phases of the design problem formulation process indicated in Figs.~\ref{f:tree} and \ref{f:hourglass}.
-\begin{comment}
-In addition, we introduce the 
-  concept of fixed vs. free connection edges to represent the
-  static relationship between variable and model nodes from a single analysis, 
-  as opposed to the flexible relationship between variables from different analyses. 
-\end{comment}
 
@@
   All global variable nodes have more than one outgoing edge. Conversely all local 
-  variable nodes have only a single outgoing edge. Fig.~\ref{f:sellar graph} contains 
+  variable nodes have only a single outgoing edge. Fig.~\ref{f:sellar types} contains 
   three global variables that are not present in the problem formulation given by 
@@
   The variable $x_1$ is usually refered to as a local variable, which is 
-  contradicted by the given graph \cite{sellar1996}. This discrepancy 
+  contradicted by the given graph \cite{Sellar1996}. This discrepancy 
   arises from the explicit treatment of objectives and constraints as separate 
@@
   loop until the starting point is reached again. For the Sellar Problem, selecting 
-  $A_1$ as the starting point would yield a problem as given in 
-  Eq.~\ref{eqn:simple}, whereas selecting $A_2$ would yield the problem as given in 
-  Eq.~\ref{eqn:simple2}.
-
+  $D_1$ as the starting point would yield a problem as given in 
+  Eq.~\ref{eqn:simple}.
   Larger problems can contain more complex cycles in their FPG, indicating more 

### 2013-05-09 graph_syntax.tex
fixing A vs D

@@
 \end{figure}
-
-% ((JG: A single variable node can have many incomming driven edges. These edges are 
-%    free and can be added or removed from the graph.))
-%
-%  [[Insert legend for node and edge types]] ((DP: I'm not sure what the legend is supposed to be))
 
@@
   expression blocks. This variable is could be considered a local 
-  variable because it directly affects only analysis block $A_1$. However, by expanding
+  variable because it directly affects only analysis block $D_1$. However, by expanding
   $F$ into an expression block with its own inputs, $F$ maintains its own local 
@@
   Fig.~\ref{f:collision_example} shows a modified version of the Sellar problem 
-  with a new analysis, $D_0$, representing a low fidelity version of $A_1$. 
+  with a new analysis, $D_0$, representing a low fidelity version of $D_1$. 
   In this modified graph, the annotations indicate the upper indegree limit. For 
-  $A_2$, $\txt{deg}^-_u(A_2.y_1) = 2$ which avoids a collsion. A  
+  $D_2$, $\txt{deg}^-_u(D_2.y_1) = 2$ which avoids a collsion. A  
   $\txt{deg}^-_u(v) > 1$ for any variable in a graph represents a decision to 

### 2013-05-09 graph_syntax.tex
update

@@
   \begin{description}
-    \item[\bf{variable:}] represents scalar or array data, inputs and outputs,
-    \item[\bf{model:}] represents the computation peformed by analysis tools,
-    \item[\bf{driver:}] control logic capable of managing iterations (PSG),
+    \item[\bf{variable node:}] represents scalar or array data, inputs and outputs,
+    \item[\bf{function node:}] represents the computation peformed by analysis tools,
+    \item[\bf{driver node:}] control logic capable of managing iterations (PSG),
   \end{description}
 and the three edge types are
-%$V_\txt{var}$, $V_\txt{model}$, $V_\txt{driver}$
+
   \begin{description}
   \item[\bf{connection edge:}] exchange of information external to analysis tools,
-  \item[\bf{model edge:}] exchange of information internal to analysis tools,
+  \item[\bf{function edge:}] exchange of information internal to analysis tools,
   \item[\bf{driven edge:}] passing of information from a driver node to a variable node (PSG).
@@
 Figure \ref{f:sellar types} demonstrates the usage of these node and edge types via the Sellar problem.
-Model nodes are indicated by squares, variable nodes are indicated by circles, connection edges are indicated by dotted lines, and model edges are indicated by solid lines.
+function nodes are indicated by squares, variable nodes are indicated by circles, connection edges are indicated by dotted lines, and function edges are indicated by solid lines.
 \begin{figure}[htb!]
@@
   \begin{enumerate}
-  \item A model node may have only one edge directed to or from another model node.
-  \item A model node may have only model edges directed in or out.
-  \item A model node must have at least one edge directed in and at least one edge 
+  \item A function node may have only one edge directed to or from another function node.
+  \item A function node may have only function edges directed in or out.
+  \item A function node must have at least one edge directed in and at least one edge 
     directed out.
-  \item If a variable node has an outgoing model edge, then it may not have other outgoing edges.
-  \item If a variable node has an incoming model edge, then it may not have any additional incoming edges.
+  \item If a variable node has an outgoing function edge, then it may not have other outgoing edges.
+  \item If a variable node has an incoming function edge, then it may not have any additional incoming edges.
   \end{enumerate}
@@
   In Alexandrov and Lewis's REMS syntax only two node types (variable and function) 
-  and one edge type are included \cite{alexandrov2004}. We have chosen to adopt 
-  the terminology ``model'' node instead of ``function'' node to be consistent 
-  with modern MDAO framework terminology. The present work adds one additional 
+  and one edge type are included \cite{alexandrov2004}. The present work adds one additional 
   node type and two additional edge types to the syntax to allow descriptive
@@
 As indicated in this figure, analysis blocks comprise three sets of nodes representing the distinct local inputs, the analysis (computation), and the distinct local outputs.
-The local input and local outputs nodes are variable nodes, while the nodes representing the analysis are model nodes.
-All of the edges within the analysis block are model edges, and are considered fixed to the analysis block.
+The local input and local outputs nodes are variable nodes, while the nodes representing the analysis are function nodes.
+All of the edges within the analysis block are function edges, and are considered fixed to the analysis block.
 This graph structure demonstrated by Fig.~\ref{f:analysis block} satisfies the rules listed in Sec.~\ref{ss:types}.
@@
 
-The model edge connecting the two model nodes represents the necessary calculations to map the inputs to the outputs of the analysis. 
+The function edge connecting the two function nodes represents the necessary calculations to map the inputs to the outputs of the analysis. 
 This edge also provides the opportunity to encode computational cost or other characteristics of the analysis code as a weight in a weighted graph formulation. 
@@
 Variable nodes in an analysis block can be distinguished as either an input or an output by the manner in which they are connected. 
-As shown in Fig.~\ref{f:analysis block}, inputs are represented as variable nodes that have an outgoing edge into a model node, and outputs are represented as variable nodes that have an incoming edge from a model node.
+As shown in Fig.~\ref{f:analysis block}, inputs are represented as variable nodes that have an outgoing edge into a function node, and outputs are represented as variable nodes that have an incoming edge from a function node.
 
-%MDAO problems require that information be passed between sets of analyses ((DP: needs better transition)).
-%For this purpose, connection edges are placed to link analysis blocks
-% 
-%When information from one output of an analysis block is passed to an 
-%input of another analysis block, a new connection edge is added between the two 
-%variable nodes involved in the exchange. 
-%These connection edges are free, and unlike the edges within an analysis block, the edges can be added or removed depending on the needs of a given design problem. 
-%In other words, free connection edges are created by engineers linking an output of one tool to an input of another one. 
-%It is allowable for a single output to have outgoing connection edges to multiple downstream inputs.
-
-\subsection{Objectives, Constraints, and Global Inputs}
+\subsection{Objectives, Constraints, and Inputs}
 \label{ss:objectives and constraints}
-Along with analysis tools, objectives, constraints, and global inputs also need to be represented.
+Along with analysis tools, objectives, constraints, and inputs also need to be represented.
 In the case of objective functions, a single output value generated by an analysis block could be selected, but commonly multiple values are assembeled together to form a composite objective function.
@@
 
-Inputs to multiple analysis tools are considered global inputs.
-These are represented in the graph syntax by single variable nodes refered to as global inputs---the definition of global with respect to a node will be defined subsequently.
-%For the sake of consistency of naming conventions $F_j$, $G_j$, and $H_j$ are used to deonte expression blocks for objectives, inequality constraints, and equality constraints. 
-%If the relevant objective and constraint values are direct results from other analysis blocks and composite functions are not needed, then expression blocks may be omitted from the MDAO problem graph.
-Figure \ref{f:obj-cons} demonstrates the use of global inputs (black circles), analysis blocks (dotted boxes), and expression blocks (dashed boxes), for the Sellar problem graph.
+Finally, inputs to the problem formulation are represented in the graph by individual variable nodes with connection edges directed out to corresponding local inputs of analysis blocks, and they are referred to simply as \emph{inputs}.
+These nodes serve to indicate that the variable nodes they are connected to represent the same variable in the problem formulation.
+These nodes serve to unify the variable nodes they are connected to. That is, they indicate that these nodes refer to the same input variable.
+
+Figure \ref{f:obj-cons} demonstrates the use of inputs (black circles), analysis blocks (dotted boxes), and expression blocks (dashed boxes), for the Sellar problem graph.
 \begin{figure}[htb!]
@@
   \end{center}
-  \caption{Sellar problem graph indicating global inputs (black circles), analysis blocks (dotted boxes), and expression blocks (dashed boxes).}
+  \caption{Sellar problem graph indicating inputs (black circles), analysis blocks (dotted boxes), and expression blocks (dashed boxes).}
 \label{f:obj-cons}
@@
   directed into or out of the node, respectively.
-%The number of driven edges is not considered relevant because any number 
-%  of drivers could be involved in different parts of a solution process. For 
-%  example, in a sequential optimization in which a global optimizer is used first 
-%  followed by a gradient-based optimizer second to refine global optimizer 
-%  result, design variables would have incomming driven edges from both optimizers. 
-%  These driven edges would not affect the indegree of those variable nodes.
 
@@
 
-%For the Sellar problem, there are three design variables: $x_1^*$, $z_1^*$, and 
-%$z_2^*$. In Fig. \ref{f:designvars} each variable is 
-%annotated with the lower indegree limit, and the three design variables are identified 
-%in bold((DP:check)). 
-%
-%\begin{figure}[htb!]
-%  \begin{center}
-%    [Graph Here....]
-%    \includegraphics[width=.6\textwidth]{images/design_vars_graph}
-%  \end{center}
-%  \caption{Sellar problem graph with lower indegree limits for all variables \label{f:designvars}}
-%\end{figure}
-%
-%In Fig. \ref{f:designvars}, the design variable are void of incoming edges. 
-%For a valid PSG, design variables could have one or more incoming driven edges representing a value determined by an optimizer.
-
   The presence of holes and collisions in a graph represent conflicts that will give
-  rise to an invalid problem formulation. It is expected that a maximal connectivity graph will posess these conflicts, whereas a fundamental problem graph will not.
-
-%However, if the graph remains unchanged while the user changes the lower indegree limit of the node to now be zero, then the node becomes a design variable.
-
-%In some cases, the distinction between a hole and a design variable is rather subtle. Both have no incoming connection edges, but 
-%
-%
-%Note that although there are no incoming edges, design variables are not holes 
-%in the graph as defined by Eq. \ref{e:hole}. Since they have $\txt{deg}_l^-(v)=0$ 
-%having no incoming connection edges does not violate their lower indegree limit. 
-%However, the concept of holes can be useful for identifying potential design variables
-%given an arbitrary MCG. If you first set $\txt{deg}_l^-(v)=1$ for all variable nodes, 
-%and then find the resulint holes, you will have a list of variable nodes that 
-%could potentially be design variables. Some variables will be true design variables, 
-%while others might represent missing information in the problem formulation. For 
-%instance, if there is a hole for a drag input of an aircraft mission analysis code 
-%then this is not likely a good choice for a design variable. So it is up to the designer to 
-%examine each hole and decide if it's an appropriate design variable. If 
-%the variable should become a design variable then set $\txt{deg}_l^-(v)=0$. 
-%If a variable should not be a design variable, then leave it as a hole which 
-%needs to be addressed in a manner discussed in section \ref{ss:obtaining FPG}.
-
-\subsection{Local vs Global Nodes }
-
-  All global variable nodes have more than one outgoing edge. Conversely all local 
-  variable nodes have only a single outgoing edge. Fig.~\ref{f:sellar types} contains 
-  three global variables that are not present in the problem formulation given by 
-  Eq.~\ref{eqn:simple_fpf}: $x_1^*$, $z_1^*$, and $z_2^*$. The presence of the 
-  new nodes is required in order to represent the equality of the given variables that
-  was implied by their shared names in Eq.~\ref{eqn:simple_fpf}.
-
-  The variable $x_1$ is usually refered to as a local variable, which is 
-  contradicted by the given graph \cite{Sellar1996}. This discrepancy 
-  arises from the explicit treatment of objectives and constraints as separate 
-  expression blocks. This variable is could be considered a local 
-  variable because it directly affects only analysis block $D_1$. However, by expanding
-  $F$ into an expression block with its own inputs, $F$ maintains its own local 
-  copy of the $x_1$ variable. This formulation necessitates the creation of the global 
-  node $x_1^*$ in the graph with two outgoing edges to link the two $x_1$ inputs nodes 
-  in the respective blocks.
-
-  This classification of $x_1^*$ in the Sellar problem
-  as a global variable works well within the structure of the Collaborative 
-  Optimization (CO) archtiecture \cite{braun1996thesis}. The rules for constructing 
-  a problem with CO require that each discipline be given an independent local 
-  copy of all global variables. All local variables are retained uniquely 
-  within their respective disciplines, except when a local design variable appears 
-  explicitely in the objective or global constraint. In this case, a global 
-  variable is created and the analysis block is again assigned an 
-  independent local copy of it. Following our definition of global and local variable 
-  nodes, no such exception for local variables in global expressions is 
-  necessary. Whenever a variable is included in an expression block, 
-  it forces the creation of a global variable node in the graph which will 
-  automatically result in the correct CO structure.
+  rise to an invalid problem formulation. It is expected that a maximal connectivity graph could indicate these conflicts, whereas a fundamental problem graph will not.
 
@@
   edges without implying a collision as defined by by Eq.~\ref{e:collision}. 
-  Fig.~\ref{f:collision_example} shows a modified version of the Sellar problem 
+An  upper indegree limit above unity for any variable in a graph represents a decision to 
+  allow multiple fidelities to interact at that part of the graph in order to allieviate
+  a conflict. 
+  Figure \ref{f:collision_example} shows a modified version of the Sellar problem 
   with a new analysis, $D_0$, representing a low fidelity version of $D_1$. 
-  In this modified graph, the annotations indicate the upper indegree limit. For 
-  $D_2$, $\txt{deg}^-_u(D_2.y_1) = 2$ which avoids a collsion. A  
-  $\txt{deg}^-_u(v) > 1$ for any variable in a graph represents a decision to 
-  allow multiple fidelities to interact at that part of the graph in order to allieviate
-  a conflict. In section \ref{ss:obtaining FPG} we present an algorithm for finding 
+  In this modified version, the variable nodes with an upper indegree limit of two are highlighted in gray, and changing these upper degree limits for the variable nodes representing $y_1$ in $D_2$, $F$, and $G_1$ avoids a collision.
+In section \ref{ss:obtaining FPG} we present an algorithm for finding 
   conflicts within a given MCG so that a designer can make the necessary decisisons about each 
@@
     \begin{center}
-      %\includegraphics[height=.25\textheight]{}
-    [INSERT GRAPH HERE]
-    \caption{Graph of the modified Sellar, multi-fidelity problem with $\txt{deg}^-_u(v)$ annotations on variable nodes\label{f:collision_example}}
+      \includegraphics[width=4in]{images/sellar_mulfi}
+    \caption{Graph of the modified Sellar problem with multi-fidelity nodes highlighed in gray.\label{f:collision_example}}
   \end{center}

### 2013-05-09 mcg_fpg.tex
update

@@
     \label{ss:MCG}
-For a specified MDAO problem, the maximal connectivity graph represents every analysis tool, objective, constraint, and global input being considered, as well as every possible interconnection among them.
+For a specified MDAO problem, the maximal connectivity graph represents every analysis tool, objective, constraint, and input being considered, as well as every possible interconnection among them.
 The definition of the MCG is given via construction.
-    To construct the maximal connectivity graph, we presume that a set of analysis tools, global inputs, objectives, and constraints are provided. 
+    To construct the maximal connectivity graph, we presume that a set of analysis tools, inputs, objectives, and constraints are provided. 
     Each of the $m$ analysis codes correspond to an index $i\in I_A$, $I_A=\{1,2,\ldots,m\}$, and are represented by an analysis block graph $G_{A_i}=(V_{A_i},E_{A_i})$.
     Each of the $n$ expression blocks correspond to an index $i\in I_E$, $I_E=\{1,2,\ldots,n\}$, and are represented by an expression block graph $G_{E_i}=(V_{E_i},E_{E_i})$.
-    Finally, the global inputs are represented as a set of variable nodes $V_\txt{in}$.
+    Finally, the inputs are represented as a set of variable nodes $V_\txt{in}$.
     We presume that $V_\txt{in}$, each $G_{A,i}$, and each $G_{E,i}$ are given, and that any potential connection between variables is specified in the form of connection edges in the set $E_{M,\txt{con}}$.
@@
     \end{IEEEeqnarray*}
-    The MCG $M$ is uniquely determined by the given set of analysis blocks, global inputs, and expression blocks. 
-    In the cases where the set of global inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.
+    The MCG $M$ is uniquely determined by the given set of analysis blocks, inputs, and expression blocks. 
+    In the cases where the set of inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.
 
@@
 \end{equation}
-where $V_{M,\txt{var}}$ and $V_{M,\txt{mod}}$ are the sets of variable nodes and model nodes, respecively;
+where $V_{M,\txt{var}}$ and $V_{M,\txt{mod}}$ are the sets of variable nodes and function nodes, respecively;
 \begin{equation}
@@
 \end{equation}
-where $ E_{M,\txt{con}}$ and $E_{M,\txt{mod}}$ are the sets of connection edges and model edges, respectively. These sets will be referenced in the process for obtaining an FPG.
+where $ E_{M,\txt{con}}$ and $E_{M,\txt{mod}}$ are the sets of connection edges and function edges, respectively. These sets will be referenced in the process for obtaining an FPG.
 
@@
     %Requirement (2) stipulates that each analysis blocks in $F$ must have at least one local output that is being used. 
-    %Requirement (3) stipulates that only the global inputs that are being used should be included. 
+    %Requirement (3) stipulates that only the inputs that are being used should be included. 
     %Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
@@
     With this approach, the problem is reduced to deciding which connection edges to remove.
-First, we describe a pruning subroutine used within the algorithm, and next we describe the algorithm itself.
 
-    \begin{description}
-    \item[\bf{Subroutine: Pruning}] 
-        Given a digraph $G_1 = (V_1,E_1)$, let the subroutine denoted as $P_\txt{prune}$ operate on $G_1$ to produce a new digraph $G_2$ as
-	\begin{equation}
-	G_2 = P_\txt{prune}(G_1) = (V_2,E_2),
-	\end{equation} 
-	where $G_2$ now satisfies requirements (2) and (4) in Sec.~\ref{ss:FPG}.
-        Pruning is accomplished by first creating the reverse graph of $G_1$, $R = (V_1,E_R)$, where
-	\begin{equation}
-	E_R = \{ (x,y) \st (y,x) \in E_1\}.
-	\end{equation}
-	Next, a new node $b$ is added to $V_1$ and edges directed from this node to each of the expression blocks are added:
-    \begin{equation}
-        \forall i \in I_E,\forall v \in V_{E_i}, \txt{ if } \txt{deg}^+(v)=0 \txt{ then } (b,v) \in E_R
-    \end{equation}
-        The set of nodes that may be reached from $b$ is constructed as
-    \begin{equation}
-        U = \{ v \in V_1 \st \exists P \txt{ a path from $b$ to $v$}, \ P \subset R \}.
-        \end{equation}
-	Because $b$ is directed into only the expression blocks, any path from $b$ necessarily provides a path from at least one expression block, which means the node is being used in the problem formulation.
-
-        The list of analysis blocks with at least one node in $U$ is constructed as
-        \begin{equation}
-        I_2 = \{ i \in I_A \st V_{A_i} \cap U \neq \emptyset \}
-        \end{equation}
-        Any analysis block not in $I_2$ should be removed from the graph because none of its outputs contribute to the problem.
-Global inputs that are not being used are also removed.
-The set of nodes to remove is then
-        \begin{equation}
-        N = (V_\txt{in}\setminus U )  \cup \left( \bigcup_{i \notin I_2} V_{A_i} \right),
-        \end{equation}
-and the new set of nodes is created as
-        \begin{equation}
-        V_2 = V_1 \setminus N.
-        \end{equation}
-        Edges involving the removed nodes are also deleted as
-        \begin{equation}
-        E_2 = E_1 \setminus \{ (x,y) \st x \in N \txt{ or } y \in N  \}.
-        \end{equation}
-        The set of connection edges can be extracted by considering only edges whos endpoints are not in the same analysis block:
-        \begin{equation}
-        E_{2,\txt{con}} = \{(x,y) \in V_{2} \st \sim(x \in V_{A_i} \txt{ and } y \in V_{A_i} \txt{ for some } i \in I_2)\}.
-        \end{equation}
-        %which means $C_2 \cap V_{A_i} = \emptyset  \ \forall i \in I_1$, and $E_2 = C_2 \cup \left( \bigcup_{i \in I_1} E_{A_i} \right)$. 
-\end{description}
-
-\noindent The algorithm for obtaining an FPG can now be described as follows:
 \begin{description}
@@
 
-        To begin the process, an initial graph is created as a pruned copy of the MCG
+        To begin the process, an initial graph is created as a copy of the MCG
         \begin{equation}
-        F_0 = P_\txt{prune}(M),
+        F_0 = M,
         \end{equation}
-        where $F_0 = (V_{F_0},E_{F_0})$, and $C_{F_0,\txt{con}}$ is also obtained as described. 
-        The set of variable nodes which are holes is identified as
+        with $F_0 = (V_{F_0},E_{F_0})$, and $C_{F_0,\txt{con}}$ is the set of connection edges. 
+        The set of variable nodes that are holes is identified as
         \begin{equation}
@@
 The construction of each $B_j$ corresponds to making a decision about which edges to include and which edges not to include. 
-%An upper bound on the number of such decisions that need to be made is 
-%\begin{equation}
-%\prod_{j \in J} 	
-%	\begin{pmatrix}
-%		|S_{\txt{edges},j}| \\
-%		\txt{deg}_u^-(v_j)
-%	\end{pmatrix}
-%\end{equation}
-%where $(:)$ refers to standard $n$-choose-$k$ notation.
 Let the new set of connection edges be
@@
     \begin{equation}
-    F_2 = (V_{F_0},E_{F,2}).
+    F_2 = (V_{F_0},E_{F_2}).
     \end{equation}
@@
     \item[\bf{Step 3: Finalize}] 
-        The third and final step is to prune the graph a final time to exclude any analysis blocks that became unneeded after the collisions were resolved:
+        The third and final step is to prune the graph to exclude any analysis blocks that became unneeded after the collisions were resolved.
+        This is accomplished by first creating the reverse graph of $F_2$, $R = (V_R,E_R)$, where
 	\begin{equation}
-        F = P_\txt{prune}(F_2),
+	E_R = \{ (x,y) \st (y,x) \in E_{F_2}\}, \ V_R = V_{F_2}
 	\end{equation}
-        with $F = (V_F,E_F)$.
-        The indices of the analysis blocks implemented in the FPG can then be found as
+	Next, a new node $b$ is added to $V_R$ and edges directed from this node to each of the expression blocks are added:
+    \begin{equation}
+        \forall i \in I_E,\forall v \in V_{E_i}, \txt{ if } \txt{deg}^+(v)=0 \txt{ then } (b,v) \in E_R
+    \end{equation}
+        The set of nodes that may be reached from $b$ is then constructed as
+    \begin{equation}
+        U = \{ v \in V_R \st \exists P \txt{ a path from $b$ to $v$}, \ P \subset R \}.
+        \end{equation}
+	Because $b$ is directed into only the expression blocks, any path from $b$ necessarily provides a path from at least one expression block, which means the node is being used in the problem formulation.
+
+        The list of analysis blocks with at least one node in $U$ is constructed as
+        \begin{equation}
+        I_F = \{ i \in I_A \st V_{A_i} \cap U \neq \emptyset \}
+        \end{equation}
+        Any analysis block not in $I_F$ should be removed from the graph because none of its outputs contribute to the problem.
+Inputs that are not being used are also removed.
+The set of nodes to remove is then
+        \begin{equation}
+        N = (V_\txt{in}\setminus U )  \cup \left( \bigcup_{i \notin I_F} V_{A_i} \right),
+        \end{equation}
+and the new set of nodes is created as
+        \begin{equation}
+        V_F = V_{F_2} \setminus N.
+        \end{equation}
+        Edges involving the removed nodes are also deleted as
+        \begin{equation}
+        E_F = E_{F_2} \setminus \{ (x,y) \st x \in N \txt{ or } y \in N  \}.
+        \end{equation}
+
+The fundamental problem graph is then
+\begin{equation}
+F = (V_F,E_F).
+\end{equation}
+        If desired, the indices of the analysis blocks implemented in the FPG can then be found as
     \begin{equation}
@@
     \end{equation}
-%       \begin{
-%       \begin{equation}
-%       I_F =  \{i \in I \ | \ \exists v \in V_{A_i} \txt{ such that }  t_\txt{node}(E_{F,2}^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) > 0 \},
-%       \end{equation}
-%       where the degree is calculated with respect to $F_2$. 
-%       This set excludes both the analysis blocks that were unused in the original MCG and those that became unused in step 1 or step 2.
-%
-%       The set of used global inputs is constructed as
-%       \begin{equation}
-%       V_{\txt{in},F} = \big\{v \in V_\txt{in} \ \big| \ |\{(x,y) \in E_{F,2}(v) \st y \in V_{A_i} \txt{ for } i \in I_F  \}| > 1 \big\},
-%       \end{equation}
-%       which requires that at least two edges directed out of the nodes in $V_{\txt{in},F}$ must be to the analysis blocks indexed by $I_F$.
-%
-%       Finally, the set of nodes and edges describing a fundamental problem graph is given by
-%       \begin{equation}
-%       V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right),
-%       \end{equation}
-%       \begin{equation}
-%       C_{F} = \{ (x,y) \in C_{F,2} \st x,y \in V_F\},
-%       \end{equation}
-%       \begin{equation}
-%       E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right),
-%       \end{equation}
-%       \begin{equation}
-%       F = (V_F,E_F).
-%       \end{equation}
-
+        Also, the set of connection edges can be extracted by considering only edges whos endpoints are not in the same analysis block:
+        \begin{equation}
+        E_{F,\txt{con}} = \{(x,y) \in E_{F} \st \sim(x \in V_{A_i} \txt{ and } y \in V_{A_i} \txt{ for some } i \in I_F)\}.
+        \end{equation}
 \end{description}
@@
 \begin{enumerate}
-\item[\bf{(A)}] Begin with a set of global inputs, analysis codes, objectives, and constraints.
+\item[\bf{(A)}] Begin with a set of inputs, analysis codes, objectives, and constraints.
 \item[\bf{(B)}] Build the MCG as described in Sec.~\ref{ss:MCG}.
@@
 		\begin{enumerate}
-\item Change the MCG by adding analysis blocks and/or global inputs, which means identifying new analyses to include in a potential MDAO workflow.
+\item Change the MCG by adding analysis blocks and/or inputs, which means identifying new analyses to include in a potential MDAO workflow.
 \item Change indegree limits (see Table \ref{t:variable node classification}).
@@
 \end{enumerate}
+This process illustrates how the algorithm for obtaining a valid FPG is to be applied.
+However, this process also serves to illustrate how a designer may revist the MCG after an FPG was already produced.
+For example, if a designer wants to change an input to be multi-fidelity, new analysis blocks can be added to the original MCG and indegree limits can be modified accordingly, which corresponds to steps (A)--(C), and then the algorithm can be applied again.

### 2013-05-09 problem_formulation.tex
update

@@
 
- \title{A Graph Theory Based Approach to MDAO Problem Formulation}
+ \title{A Graph Theoretic Approach to Problem Formulation for Multidisciplinary Design Optimization}
 
@@
     AAO      & All-At- \\
-    MDAO     & Multidisciplinary Design Analysis and Optimization \\
+    MDAO     & Multidisciplinary Design Analysis and \\ & Optimization \\
     FPF      & Fundamental Problem Formulation \\

### 2013-05-09 conclusions.tex
started editing conclusion

@@
 \section*{Conclusions}
-In this paper, we have presented an approach for describing MDAO problems with constructs and syntax from graph theory.  
-Our graph description shares similarities to other approaches such as REMS, $\Psi$, FDT, DSM, and XDSM, but it provides new constructs tailored to applications in early phases of MDAO problem formulation before specific problem solution approaches are envisioned.
-In particular, we introduce the concepts of the Maximally Connected Graph (MCG) and the Fundamental Problem Graph (FPG).
+In this paper, we have presented an approach for describing  MDAO problems with 
+with a graph based syntax.  Our graph description shares 
+similarities to other approaches such as REMS, FDT, DSM, and XDSM, but 
+it provides new constructs tailored to applications in early phases of the design 
+process where a problem formulation has not yet been fully formed.   
+In particular, we introduce the concepts of the Maximal Connectivity Graph (MCG) 
+and the Fundamental Problem Graph (FPG).
 
-The MCG addresses the question, ``Given a set of analysis tools, what are all of the variable interconnections between them that could be established?''  
-The MCG provides a structured formalism to identify \textit{holes} and \textit{conflicts} in interconnections between input and output variables implied by the set of analysis tools.  
-A hole corresponds to a variable that is not calculated in the MCG data flow but is needed by other tools.   
-To create a valid problem formulation, all holes must be ``filled'' either by assigning the corresponding variable as a design variable to be set by the user or an optimizer, or by introducing additional tools to compute the variable.
+The MCG addresses the question, ``Given a set of analysis tools and design objectives, what are all 
+of the potential problem formulations that exist to solve a design problem?''  
+The MCG provides a structured formalism to reduce the (potentially very large) 
+set of valid combinations of analysis tools to a single subset and to identify
+conflicts that could prevent a valid problem formulation.
+Conflicts can originate from two sources: holes and collisisons. A hole 
+corresponds to a variable that is not calculated by any analysis block in the 
+MCG but is needed by other blocks. To create a valid problem formulation, all 
+holes must be ``filled'' either by assigning the corresponding variable as a 
+design variable to be set by the user or an optimizer, or by introducing 
+additional tools to compute the variable.  A conflict corresponds to a 
+situation in which two or more analysis tools compute values for the same variable.  
+To create a valid problem formulation, all conflicts must be resolved either by 
+removing redundant analysis tools or by allowing the redundencies to remain 
+in the form of a multifidelity problem. For the latter case, a special solution 
+method will need to be employed to manage the multifidelity situation and decide 
+when to execute the different analysis tools. Resolving collisions and filling 
+holes in an MCG both require choices by the designer to shape the overall
+MDAO problem that he/she desires to solve.
 
-A conflict corresponds to a situation in which two or more analysis tools compute separate instances of the same variable.  
-To create a valid problem formulation, all conflicts must be resolved either by removing redundant analysis tools or by introducing logic or solvers into the data flow to choose between or aggregate multiple instances during execution.
-The identification of conflicts in an MCG is especially important for problem formulations involving multiples analysis tools of different fidelities; in this context, a conflict represents an opportunity to formulate a multi-fidelity problem that can take best advantage of the efficiency/accuracy tradeoff between different analysis tools.
-Resolving conflicts and filling holes in an MCG requires choices by the user to specify the overall MDAO problem that he/she desires to solve, i.e.
-to produce the required set of system-level outputs given a specified set of design variables.
+The FPG represents the problem as an engineer would describe it: ``Given a specific set of tools, 
+with a set of free design vairables, find design that best meets a given objective
+within the given constraints.'' Although an FPG uniquely defines a single MDAO problem, 
+any given MCG can potentially contain a number of different FPG's. In Section 
+\ref{s:building graphs} we provide an algorithm 
+that reduces an MCG, via the resolution of conflicts in the graph, down to an FPG. 
+How one goes about resolving the conflicts will determine which particular FPG is reached. 
+One benefit of using graph theory is the standard tools, such as cycle 
+detection, which can be used to inform the user's decisions and make the process of conflict 
+resolution much easier.
 
-The FPG is a graph that describes a data connection structure corresponding to an MDAO problem formulation free of conflicts and holes.  
-An FPG is the result of user choices to fill holes and resolve conflicts in the initial MCG.  
-One benefit of using graph theory is the standard algorithms, such as cycle detection, which can be used to inform the user's decisions.
-Typically, many different FPGs could be attained from a given MCG, depending on the user choices.  
-The number of possible FPGs that can be attained by selecting different design variables or introducing additional analysis tools to fill holes and/or resolve conflicts can be viewed as a measure of the freedom available to the user in implementing the available analysis tools to formulate valid MDAO problems.
-In Section 5, we present an algorithm for identifying holes and conflicts in an MCG that can be applied by the user in a process to arrive at valid FPGs.  
-In Section 6, we provide an example application of formulating an FPG from an MCG for a commercial aircraft design problem.
+In Section \ref{s:example problem}, we provide an example application of formulating different FPGs from an MCG 
+for a commercial aircraft design problem. This problem includes 8 
+analysis tools and approximately 20 variables. Despite this relatively modest size, 
+the MCG is rich enough to provide for a number of different valid FPGs. This demonstrates
+the inherant value of the MCG to the design process by providing a mechanism for selecting 
+an FPG based on important factors such as available design cycle time, consideration of 
+coupling effects, or desired fidelity. Even after downselecting to a single FPG, the 
+MCG itself still remains a usefull tool. No real design problem is static, and will always
+evolve throughout the design process. New tools may be introduced to account for previously 
+unexpected effects and new constraints or objectives can be introduced. When this occurs, 
+the MCG provides the opportunity to account for these changes by starting over with the 
+algorithm from Section \ref{s:building graphs} to downselect to the most effective new FPG. 
+By repeating this process you allow for any new interdisciplinary couplings and  
+information conflicts to play a role in selecting the most appropriate FPG.
 

### 2013-05-09 conclusions.tex

@@
 The MCG addresses the question, ``Given a set of analysis tools and design objectives, what are all 
-of the potential problem formulations that exist to solve a design problem?''  
-The MCG provides a structured formalism to reduce the (potentially very large) 
+of the potential problem formulations that could exist to solve a design problem?''  
+The MCG provides a structured formalism to reduce the, potentially very large, 
 set of valid combinations of analysis tools to a single subset and to identify
@@
 MCG but is needed by other blocks. To create a valid problem formulation, all 
-holes must be ``filled'' either by assigning the corresponding variable as a 
-design variable to be set by the user or an optimizer, or by introducing 
-additional tools to compute the variable.  A conflict corresponds to a 
-situation in which two or more analysis tools compute values for the same variable.  
-To create a valid problem formulation, all conflicts must be resolved either by 
-removing redundant analysis tools or by allowing the redundencies to remain 
+holes must be ``filled'' either by selecting the corresponding variable as a 
+design variable, or by introducing additional analysis tools to compute the variable.  
+A collision happens when two or more analysis tools compute values for the same variable.  
+To create a valid problem formulation, all collisions must be resolved either by 
+removing redundant analysis tools or by explicitly allowing the redundencies to remain 
 in the form of a multifidelity problem. For the latter case, a special solution 
@@
 holes in an MCG both require choices by the designer to shape the overall
-MDAO problem that he/she desires to solve.
+MDAO problem to meet the needs of the problem he/she is trying to solve.
 
 The FPG represents the problem as an engineer would describe it: ``Given a specific set of tools, 
-with a set of free design vairables, find design that best meets a given objective
+with a set of design vairables, find design that best meets a given objective
 within the given constraints.'' Although an FPG uniquely defines a single MDAO problem, 
-any given MCG can potentially contain a number of different FPG's. In Section 
+there is not necessarily a unique FPG for any given MCG. In Section 
 \ref{s:building graphs} we provide an algorithm 
@@
 How one goes about resolving the conflicts will determine which particular FPG is reached. 
-One benefit of using graph theory is the standard tools, such as cycle 
+One benefit of using graph theory is the standard techniques, such as cycle 
 detection, which can be used to inform the user's decisions and make the process of conflict 
@@
 
-In Section \ref{s:example problem}, we provide an example application of formulating different FPGs from an MCG 
-for a commercial aircraft design problem. This problem includes 8 
+In Section \ref{s:example problem}, we provide an example of 
+formulating different FPGs from an MCG for a commercial aircraft design problem. This problem includes 8 
 analysis tools and approximately 20 variables. Despite this relatively modest size, 
@@
 coupling effects, or desired fidelity. Even after downselecting to a single FPG, the 
-MCG itself still remains a usefull tool. No real design problem is static, and will always
+MCG itself still remains a useful tool. No real design problem is static, and will always
 evolve throughout the design process. New tools may be introduced to account for previously 
@@
 algorithm from Section \ref{s:building graphs} to downselect to the most effective new FPG. 
-By repeating this process you allow for any new interdisciplinary couplings and  
+While it is possible that this new FPG will be very similar to the previous one, 
+it is also very possible that it will be different in significant ways. 
+By repeating downselection algorithm from the MCG you allow for any new interdisciplinary couplings and  
 information conflicts to play a role in selecting the most appropriate FPG.
 
-Once a desired FPG is identified, the graph syntax described in the paper can be leveraged to formulate a Problem Solution Graph (PSG) describes particular MDAO solution architectures, e.g.
-MDF, IDF, BLISS, ATC, that can be used to solve the problem defined by the FPG.
-Many possible PSGs could be formulated to solve a problem corresponding to a specified FPG.   
-The application of the syntax to PSGs has not been illustrated in the paper.
+Once a desired FPG is identified, the graph syntax described in the paper can be 
+leveraged to formulate a Problem Solution Graph (PSG). This graph includes all of the 
+same information from the FPG while additionally describing the particular MDAO 
+solution architecture, e.g. MDF, IDF, BLISS, ATC, that will be used to solve the problem.  
+Many possible PSGs could be formulated to solve a problem corresponding to a 
+specified FPG. The application of the syntax to PSGs has not been illustrated in the paper, 
+but the basic concept of applying multiple solution strategies to a given problem 
+formulation has been well established by a number of MDAO architecture 
+benchmarking efforts\cite{Tedford2009,gray2013}. Furthermore the validity of applying 
+graph syntax to the description of specific problem solutions has been demonstrated 
+by XDSM.
 

### 2013-05-09 example_problem.tex
update

@@
 	This section presents an example problem to demonstrate the process of obtaining the FPG and the advantages of doing so. 
-	For this example, the algorithm from Sec.~\ref{ss:obtaining FPG} was implemented in the NetworkX package of the Python programming language. 
-	While any language with the ability to manipulate graphs is applicable, NetworkX was chosen due to its built-in features, such as cycle detection and shortest path algorithms.
-
 The example task is to create an FPG for the conceptual sizing of a single-aisle subsonic transport using a set of analysis tools with objectives including performance and gross weight for a required mission.
@@
 	Table \ref{t:ins and outs} presents the full list of variables in the leftmost column and indicates whether the variable is an input or an output of each analysis code. 
-	Some variables, such as geometry and performance, represent groups of variables that are passed as vectors or other structures due to their similarity. 
+	Some variables, such as geometry and performance, represent groups of variables that are passed as arrays or other data structures due to their similarity. 
 	This bundling of variables is not fundamental and does not limit the generality of this example; rather, it simplifies the presentation.
@@
 	\item An analysis block is created for each analysis code using the information in each column of Table \ref{t:ins and outs}. 
-	Each analysis block is formed by first creating variable nodes for each input and adding directed edges into a single model node.
-An edge is directed from this model node into a second model node, which is then directed into variable nodes corresponding to each output.
+	Each analysis block is formed by first creating variable nodes for each input and adding directed edges into a single function node.
+An edge is directed from this function node into a second function node, which is then directed into variable nodes corresponding to each output.
 	A sample analysis block is shown for analysis code FLOPSa in Fig.~\ref{f:FLOPSb analysis block}.
@@
 
-	\item A variable nodes is created to represent the geometry variable as a global input. Any global inputs incorrectly omitted will be identified in step (D).
+	\item A variable nodes is created to represent the geometry variable as a input. Any inputs incorrectly omitted will be identified in step (D).
 
 	\item Connection edges are created that connect each variable node to every other variable node representing variables with the same name. 
-	The direction is determined by whether the variable node has an edge directed into or out of a model node, i.e. whether it is a local input or a local output.
+	The direction is determined by whether the variable node has an edge directed into or out of a function node, i.e. whether it is a local input or a local output.
 	\end{enumerate}
@@
 	  \begin{center}
-		\includegraphics[width=4.5in]{images/MCG_edit_holes}
+		\includegraphics[width=6in]{images/MCG_edit_holes}
 	  \end{center}
@@
 
-Step (D)(a) begins to resolve this conflict by creating a global input for the number of engines.
-In this example, the number of engines was not originally included as a global input to demonstrate how the process responds to this error.
+Step (D)(a) begins to resolve this conflict by creating a input for the number of engines.
+In this example, the number of engines was not originally included as a input to demonstrate how the process responds to this error.
 Step (D)(b) then reduces the lower and upper indegree limits for the variable node representing the mission definition to zero, implying that these nodes must be specified as design variables (see Table \ref{t:variable node classification}).
@@
 	  \begin{center}
-		\includegraphics[width=4.5in]{images/MCG_edit}
+		\includegraphics[width=6in]{images/MCG_edit}
 	  \end{center}
-	  \caption{The updated MCG with a new global input.}
+	  \caption{The updated MCG with a new input.}
 	\label{f:MCG}
@@
 There are two cycles which arise from each local output from PDCYL being directed into FLOPSa and then back into PDCYL.
-%((DP:discuss cycle equivalency))
-%((mention edge weight for center edge))
 
 	Similar alternatives would be to minimize the number of analysis blocks involved in cycles, counting multiplicity, or to minimize the length of the longest cycle (called the \emph{circumference} of the graph).
-Since these graphs will always be relatively small (on the order of hundreds of nodes) the
-	computational expense for examining all possible FPGs is negligble compared to the expected expense of using the problem formulation to obtain a solution.
+It is beyond the scope of this work to explore efficient methods for finding the FPGs meeting these criteria.
+However, as discussed by Gabow\cite{Gabow1985}, graph algorithms typically scale well with the size of the graph.
+This suggests that the computation cost of finding an FPG should be small.
 	\begin{figure}[htb!]
@@
 	  \begin{center}
-		\includegraphics[width=4.5in]{images/FPG_edit_ranking}
+		\includegraphics[width=6in]{images/FPG_edit_ranking}
 	  \end{center}

### 2013-05-09 intro.tex
removing old Fig 1.
update

@@
     Our goal is to develop formalism for expressing analysis interconnectivity and for determining feasible
-    data flows to assist an engineering team conducting this task. Because the problem deals with
-    interconnectivity, we base our approach on the representations and techniques of graph theory. The 
-    proposed graph syntax is closely related to the REMS system proposed by 
-    Alexandrov and Lewis\cite{alexandrov2004}. The approach taken here should be viewed as an extention 
-    of the concepts proposed in REMS with the goal of allowing the graphs to more closely interact with 
-    existing design processes and MDAO frameworks.
-
-The approach begins by 
-    constructing the \emph{maximal connectivity graph (MCG)} describing all possible
+    analysis tool sets to assist an engineering team conducting this task. Because the problem deals with
+    interconnectivity, we base our approach on the representations and techniques of graph theory.
+    The approach begins by constructing the \emph{maximal connectivity graph (MCG)} describing all possible
     interconnections between the analysis tools proposed by the engineers. Graph operations are then
@@
     The Fundamental Problem Formulation (FPF) is our terminology for a 
-    statement of the overall system-level optimization problem that does not 
-    depend on the choice of solution-specifc elements such as the MDAO 
-    framework, optimization algorithm, iterative solver, or execution sequence
-    used to solve the problem. The FPF contrasts with problem statements that are written without 
-    reference to a specific solution strategy, but that still imply a specific execution sequence.
+    statement of the overall system-level optimization problem that contains only 
+    information about analysis tools, design variables, objectives, and constraints without any 
+    information about how to solve the problem. It does not depend on the choice 
+    of solution-specific elements such as the MDAO framework, optimization 
+    algorithm, iterative solver, or execution sequence. This description represents 
+    the problem from the point of view of a engineer specifying the details of a design problem 
+    without constrainig how it should be solved. The lack of solution information 
+    in the FPF contrasts with problem statements that are written with reference 
+    to a specific solution strategy or with implication of a specific execution sequence.
     For example, consider the Sellar test problem \cite{AIAA:sellar}, with an FPF given as follows:
@@
 
-    In Eq. \ref{eqn:simple_fpf}, $y_1$ is an output of $D_1$ as well as an input 
+    In Eq.~\ref{eqn:simple_fpf}, $y_1$ is an output of $D_1$ as well as an input 
     to $D_2$. Similarly, $y_2$ is an output of $D_2$ as well as an input 
@@
 
-    Eq. \ref{eqn:simple} gives a modified problem formulation where $D_1$ is run 
+    The Sellar problem is very simple. There is only a single coupling interaction 
+    between the two disciplines and there are a very limited set of variables. 
+    That makes it relatively easy to describe the problem given in 
+    Eq.~\ref{eqn:simple_fpf}. However, for larger more complex problems, 
+    it is much harder to identify the fundamental problem formulation. Interdisciplinary 
+    couplings are not always obvious and the set of analsyis tools and variables are 
+    much larger. To further compount the problem engineers may already have some 
+    solution strategy or execution order they are used to applying to problems. 
+    Hence, it is common to see problem formulations given in a less general form 
+    as in Eq. \ref{eqn:simple}. Here $D_1$ must be run 
     first. A new variable is introduced, $\hat{y_2}$, to break the dependence of
-    $D_1$ on $D_2$. Also a new coupling constraint, $H$, to enforce the consistency
-    that would otherwise be lost due to $\hat{y_2}$. Equation \ref{eqn:simple} is 
-    a specific alternative expression of the more fundamental problem description 
-    given in Eq. \ref{eqn:simple_fpf}.
+    $D_1$ on $D_2$. Also a new coupling constraint, $H$, is added to enforce the 
+    consistency that would otherwise be lost due to $\hat{y_2}$. Equation \ref{eqn:simple} is 
+    a more specific, but equally valid, expression of the Sellar problem that could easily be 
+    generated through an incremental approach to constructing the problem.
 
@@
     \end{align}
-
-    Figure \ref{f:sellar graph} depicts the Sellar problem given by 
-    Eq.~\ref{eqn:simple_fpf} in the graph-based syntax which will be presented 
-    in this paper. This example will serve to highlight the features of the 
-    syntax as they are introduced.
-    \begin{figure}[htb!]
-      \begin{center}
-        \includegraphics[width=4.0in]{images/sellar}
-      \end{center}
-      \caption{Sellar problem represented as a graph.}
-    \label{f:sellar graph}
-    \end{figure}
 
@@
     variable. Michelena and Papalambros made use of the FDT to solve a graph partitioning problem that yielded 
-    more efficient optimization problem decompositions\cite{Michelena1997}. While FDT succeeds at capturing the 
+    more efficient optimization problem decompositions\cite{Michelena1997}. Allison 
+    extended this work to couple adjacency matrix information with the FDT to 
+    account for system coupling in an automated partitioning scheme\cite{Allison2008} While FDT succeeds at capturing the 
     information about objectives and constraints, its lack of directed edges 
@@
     sufficient to completely describe a problem formulation.
+
     \begin{figure}[htb!]
@@
     syntax for variables and functions but does not facilitate inclusion 
-    of solvers or optimizers in the graph.
+    of solvers or optimizers in the graph.  The 
+    proposed graph syntax is closely related to REMS. The syntax proposed here should 
+    be viewed as an extention to REMS with the goal of allowing the graphs to more 
+    closely interact with existing design processes and MDAO frameworks.
 
@@
     By introducing solver or optimizer blocks, XDSM automatically implies a strategy. 
-    The XDSM for Eq. \ref{eqn:simple} is 
-    given in Figure \ref{fig:XDSM_simple}. This diagram is shown with an 
+    The XDSM for Eq. \ref{eqn:simple} is given in Figure \ref{fig:XDSM_simple}. 
+    This diagram is shown with an 
     assumed Gauss-Seidel iteration scheme and an MDF solution architecture.
@@
   \label{s:requirements}
-  The goal of the new graph-based syntax presented here is to enable the general structure of an MDAO problem to be described independently of any solution information, 
+  The goal of the new graph-based syntax presented here is to enable the general 
+  structure of an MDAO problem to be described independently of any solution information, 
   while still being able to accomodate the more specific case when a solution 
@@
     \item Design variables, objectives, and constraints
-    \item Local and global information
     \item Coupling between analyses
@@
 
-  The \emph{maximal connectivity graph} represents the first phase of the problem formulation with all 
-  analysis tools being considered and all possible connections between them also present. The second graph 
-  is the \emph{fundamental problem graph}, which is comprises only the analyses participating to solve the problem. Finally, a \emph{problem solution graph} 
-  may be represented by including additional edges and nodes to represent the 
-  solution strategy being employed to solve the problem. This paper focuses on the MCG and FPG and does not describe the PSG in detail.
+  The \emph{maximal connectivity graph} represents the first phase of the problem 
+  formulation with all analysis tools being considered and all possible connections 
+  between them also present. The second graph is the \emph{fundamental problem graph}, 
+  which is comprises only the analyses participating to solve the problem. 
+  Finally, a \emph{problem solution graph} may be represented by including additional 
+  edges and nodes to represent the solution strategy being employed to solve the 
+  problem. This paper focuses on the MCG and FPG and does not describe the PSG in detail.
 

### 2013-05-10 conclusions.tex
working on conclusion.

@@
 
-Once a desired FPG is identified, the graph syntax described in the paper can be 
+Once a desired FPG is identified, the driver node and driven edge described in the paper can be 
 leveraged to formulate a Problem Solution Graph (PSG). This graph includes all of the 
@@
 formulation has been well established by a number of MDAO architecture 
-benchmarking efforts\cite{Tedford2009,gray2013}. Furthermore the validity of applying 
+benchmarking efforts. Furthermore the validity of applying 
 graph syntax to the description of specific problem solutions has been demonstrated 
@@
 
-For simple problems with few analysis tools and variables, the formulation of a valid problem description is straightforward.  
-However, MDAO problems continue to grow in scale and complexity.  
-As the numbers of analysis tools and variables have increased, it has become increasingly challenging and time consuming for engineering teams to determine how the multiple analysis tools can be interconnected to produce valid problem formulations, to know when other tools must be introduced, and to determine the number of free variables in the problem that can or should be varied by the designer or an optimizer.
-The graph formalism presented in this paper is intended to offer value in this context of increasing problem complexity by providing a formal approach to identify and resolve missing and redundant data in order to create valid MDAO problem formulations.
+For simple problems with few analysis tools and variables, the formulation of a 
+valid problem description is straightforward.  However, MDAO problems continue to 
+grow in scale and complexity. As the numbers of analysis tools and variables have 
+increased, it has become increasingly challenging and time consuming for engineering 
+teams to determine how the multiple analysis tools can be interconnected to produce 
+valid problem formulations. Large complex problems also present signficant challenges
+as the design process evolves requiring the integration of additional and higher-fidelity 
+analysis tools into an existing model. The graph syntax presented in this paper 
+is intended to offer value in this context of increasing problem complexity by 
+providing a formal approach to identify and resolve data conflicts in a set of interconnected
+analysis tools and to help arrange those tools into an effective and efficient system model
+where MDAO tools and techniques can be quickly and easily applied.

### 2013-05-10 intro.tex
working on conclusion.

@@
     FPF for a set of test problems and also to describe specific formulations for solving them with a 
-    number of optimization architectures\cite{Tedford2009}. Their work demonstrates how multiple specific 
-    problem formulations can relate to a common FPG. The challenge with this 
+    number of optimization architectures\cite{Tedford2009}. This enabled them to produce a set of 
+    results benchmarking the performance of different optimziation architectures against a fixed set of 
+    problems. Gray et al. also benchmarked the performance of a set of MDAO architectures\cite{Gray2013}. 
+    This work looked at a larger set of architectures and proposed the use of OpenMDAO as a platform to begin 
+    building a larger community developed set of test problems and architectures. Both benchmarking efforts 
+    demonstrate how multiple specific problem formulations can relate to a common FPG and prove the 
+    value of a common problem descrition. The challenge with this 
     traditional mathematical syntax is that it is not easily manipulated or analyzed.

### 2013-05-10 conclusions.tex
last tweaks

@@
 valid problem formulations. Large complex problems also present signficant challenges
-as the design process evolves requiring the integration of additional and higher-fidelity 
+as the design process evolves requiring the integration high-fidelity 
 analysis tools into an existing model. The graph syntax presented in this paper 

### 2013-08-07 conclusions.tex
files for submitted version and reviewer 3's comments

--- 2013-05-09 conclusions.tex
@@
 \section*{Conclusions}
-In this paper, we have presented an approach for describing  MDAO problems with 
-with a graph based syntax.  Our graph description shares 
-similarities to other approaches such as REMS, FDT, DSM, and XDSM, but 
-it provides new constructs tailored to applications in early phases of the design 
-process where a problem formulation has not yet been fully formed.   
-In particular, we introduce the concepts of the Maximal Connectivity Graph (MCG) 
-and the Fundamental Problem Graph (FPG).
+In this paper, we have presented an approach for describing MDAO problems with constructs and syntax from graph theory. 
+Our graph description shares similarities to other approaches such as REMS, $\Psi$, FDT, DSM, and XDSM, but it provides new constructs tailored to applications in early phases of MDAO problem formulation before specific problem solution approaches are envisioned.
+In particular, we introduce the concepts of the Maximally Connected Graph (MCG) and the Fundamental Problem Graph (FPG).
 
-The MCG addresses the question, ``Given a set of analysis tools and design objectives, what are all 
-of the potential problem formulations that could exist to solve a design problem?''  
-The MCG provides a structured formalism to reduce the, potentially very large, 
-set of valid combinations of analysis tools to a single subset and to identify
-conflicts that could prevent a valid problem formulation.
-Conflicts can originate from two sources: holes and collisisons. A hole 
-corresponds to a variable that is not calculated by any analysis block in the 
-MCG but is needed by other blocks. To create a valid problem formulation, all 
-holes must be ``filled'' either by selecting the corresponding variable as a 
-design variable, or by introducing additional analysis tools to compute the variable.  
-A collision happens when two or more analysis tools compute values for the same variable.  
-To create a valid problem formulation, all collisions must be resolved either by 
-removing redundant analysis tools or by explicitly allowing the redundencies to remain 
-in the form of a multifidelity problem. For the latter case, a special solution 
-method will need to be employed to manage the multifidelity situation and decide 
-when to execute the different analysis tools. Resolving collisions and filling 
-holes in an MCG both require choices by the designer to shape the overall
-MDAO problem to meet the needs of the problem he/she is trying to solve.
+The MCG addresses the question, ``Given a set of analysis tools, what are all of the variable interconnections between them that could be established?''  
+The MCG provides a structured formalism to identify \textit{holes} and \textit{collisions} in interconnections between input and output variables implied by the set of analysis tools. 
+A hole corresponds to a variable that is not calculated in the MCG data flow but is needed by other tools. 
+To create a valid problem formulation, all holes must be ``filled'' either by assigning the corresponding variable as a design variable to be set by the user or an optimizer, or by introducing additional tools to compute the variable.
 
-The FPG represents the problem as an engineer would describe it: ``Given a specific set of tools, 
-with a set of design vairables, find design that best meets a given objective
-within the given constraints.'' Although an FPG uniquely defines a single MDAO problem, 
-there is not necessarily a unique FPG for any given MCG. In Section 
-\ref{s:building graphs} we provide an algorithm 
-that reduces an MCG, via the resolution of conflicts in the graph, down to an FPG. 
-How one goes about resolving the conflicts will determine which particular FPG is reached. 
-One benefit of using graph theory is the standard techniques, such as cycle 
-detection, which can be used to inform the user's decisions and make the process of conflict 
-resolution much easier.
+A collision corresponds to a situation in which two or more analysis tools compute separate instances of the same variable. 
+To create a valid problem formulation, all collisions must be resolved either by removing redundant analysis tools or by introducing logic or solvers into the data flow to choose between or aggregate multiple instances during execution.
+The identification of collisions in an MCG is especially important for problem formulations involving multiples analysis tools of different fidelities; in this context, a collision represents an opportunity to formulate a multi-fidelity problem that can take best advantage of the efficiency/accuracy tradeoff between different analysis tools.
+Resolving collisions and filling holes in an MCG requires choices by the user to specify the overall MDAO problem that he/she desires to solve, i.e.
+to produce the required set of system-level outputs given a specified set of design variables.
+One benefit of a graph theoretic approach is the standard algorithms which can be used to inform the user's decisions, such as cycle detection, minimum spanning trees, and shortest path algorithms.
 
-In Section \ref{s:example problem}, we provide an example of 
-formulating different FPGs from an MCG for a commercial aircraft design problem. This problem includes 8 
-analysis tools and approximately 20 variables. Despite this relatively modest size, 
-the MCG is rich enough to provide for a number of different valid FPGs. This demonstrates
-the inherant value of the MCG to the design process by providing a mechanism for selecting 
-an FPG based on important factors such as available design cycle time, consideration of 
-coupling effects, or desired fidelity. Even after downselecting to a single FPG, the 
-MCG itself still remains a useful tool. No real design problem is static, and will always
-evolve throughout the design process. New tools may be introduced to account for previously 
-unexpected effects and new constraints or objectives can be introduced. When this occurs, 
-the MCG provides the opportunity to account for these changes by starting over with the 
-algorithm from Section \ref{s:building graphs} to downselect to the most effective new FPG. 
-While it is possible that this new FPG will be very similar to the previous one, 
-it is also very possible that it will be different in significant ways. 
-By repeating downselection algorithm from the MCG you allow for any new interdisciplinary couplings and  
-information conflicts to play a role in selecting the most appropriate FPG.
+The FPG is a graph that describes a data connection structure corresponding to an MDAO problem formulation free of collisions and holes. 
+An FPG is the result of user choices to fill holes and resolve collisions in the initial MCG. 
+Typically, many different FPGs could be attained from a given MCG, depending on the user choices. 
+The number of possible FPGs that can be attained by selecting different design variables or introducing additional analysis tools to fill holes and/or resolve collisions can be viewed as a measure of the freedom available to the user in implementing the available analysis tools to formulate valid MDAO problems.
+In Sec.~\ref{s:building graphs}, we present an algorithm for identifying holes and conflicts in an MCG that can be applied by the user in a process to arrive at valid FPGs. 
+In Sec.~\ref{s:example problem}, we provide an example application of formulating an FPG from an MCG for a commercial aircraft design problem.
 
-Once a desired FPG is identified, the graph syntax described in the paper can be 
-leveraged to formulate a Problem Solution Graph (PSG). This graph includes all of the 
-same information from the FPG while additionally describing the particular MDAO 
-solution architecture, e.g. MDF, IDF, BLISS, ATC, that will be used to solve the problem.  
-Many possible PSGs could be formulated to solve a problem corresponding to a 
-specified FPG. The application of the syntax to PSGs has not been illustrated in the paper, 
-but the basic concept of applying multiple solution strategies to a given problem 
-formulation has been well established by a number of MDAO architecture 
-benchmarking efforts\cite{Tedford2009,gray2013}. Furthermore the validity of applying 
-graph syntax to the description of specific problem solutions has been demonstrated 
-by XDSM.
+Once a desired FPG is identified, the graph syntax described in the paper can be leveraged to formulate a Problem Solution Graph (PSG).
+The PSG describes particular MDAO solution architectures, e.g.
+MDF, IDF, BLISS, ATC, that can be used to solve the problem defined by the FPG.
+Many possible PSGs could be formulated to solve a problem corresponding to a specified FPG.  
+The application of the syntax to PSGs has not been illustrated in the paper.
 

### 2013-08-07 example_problem.tex
files for submitted version and reviewer 3's comments

@@
 	This section presents an example problem to demonstrate the process of obtaining the FPG and the advantages of doing so. 
-The example task is to create an FPG for the conceptual sizing of a single-aisle subsonic transport using a set of analysis tools with objectives including performance and gross weight for a required mission.
+The example task is to create an FPG for the conceptual sizing of a single-aisle subsonic transport aircraft using a set of analysis tools with objectives including performance and gross weight for a required mission.
 	The full set of analysis codes available for use is provided in Table \ref{t:analysis codes}.
@@
 	  \centering
-	  \caption{Analysis codes for subsonic transport sizing}
+	  \caption{Analysis tools for subsonic transport sizing}
 		\begin{tabular}{cc}
 		\toprule
-		analysis code & description \\
+		analysis tool & description \\
 		\midrule
@@
 		NPSS  & engine sizing and performance \\
-		VORLAX & aerodynamics using the vortex lattice method \\
-		PMARC & aerodynamics using a low-order panel method \\
+		VORLAX & vortex lattice method  aerodynamics \\
+		PMARC & low-order panel method  aerodynamics \\
 		WATE  & engine weight estimation \\
-		FLOPSa & mission performance, \\
-		  & engine sizing, and weight estimation \\
+		FLOPSa & mission performance \\
+		  & engine sizing and weight estimation \\
 		FLOPSb & mission performance only \\
@@
 	\end{table}%
-	Each analysis code contributes individual disciplinary analysis capability, but the outputs of the codes are not mutually exclusive. 
-	For example, VORLAX and PMARC are both aerodynamics codes that predict inviscid drag but with different levels of fidelity. 
+	Each analysis tool contributes individual disciplinary analysis capability, but the outputs of the tools are not mutually exclusive. 
+	For example, VORLAX and PMARC are both aerodynamics codes that predict induced drag but with different levels of fidelity. 
 	The Flight Optimization System (FLOPS) is included twice to represent two different configurations corresponding to different uses. 
-	FLOPSa denotes FLOPS implemented to execute both mission performance and engine analysis, while FLOPSb indicates  FLOPS implemented to execute for only mission performance analysis. 
+	FLOPSa denotes FLOPS implemented to execute both mission performance and engine analysis, while FLOPSb indicates  FLOPS implemented to execute only mission performance analysis. 
 	This representation is useful for ``supercodes'' that are capable of being implemented in different ways and with different sets of inputs and outputs.
@@
 Step (B) is the production of the MCG.
-	In this example, the MCG is formed using a consistant variable naming convention and then connecting all variables with the same name with connection edges. 
-	Table \ref{t:ins and outs} presents the full list of variables in the leftmost column and indicates whether the variable is an input or an output of each analysis code. 
+	In this example, the MCG is formed by using a consistant variable naming convention and then connecting all variables with the same name with connection edges. 
+	Table \ref{t:ins and outs} presents the full list of variables in the leftmost column and indicates whether the variable is an input or an output of each analysis tool. 
 	Some variables, such as geometry and performance, represent groups of variables that are passed as arrays or other data structures due to their similarity. 
 	This bundling of variables is not fundamental and does not limit the generality of this example; rather, it simplifies the presentation.
-	\begin{table}[htb!]
+	\begin{table*}[htb!]
 	  \centering
@@
 	  \label{t:ins and outs}
-	\end{table}
+	\end{table*}
 
@@
 	Each analysis block is formed by first creating variable nodes for each input and adding directed edges into a single function node. 
-An edge is directed from this function node into a second function node, which is then directed into variable nodes corresponding to each output.
+A function edge is created from this function node into a second function node, which is then directed into variable nodes corresponding to each output. 
 	A sample analysis block is shown for analysis code FLOPSa in Fig.~\ref{f:FLOPSb analysis block}.
@@
 	  \begin{center}
-		\includegraphics[width=2.75in]{images/FLOPSa_analysis_block}
+		\includegraphics[width=3.25in]{images/FLOPSa_analysis_block}
 	  \end{center}
@@
 
-	\item A variable nodes is created to represent the geometry variable as a input. Any inputs incorrectly omitted will be identified in step (D).
+	\item A variable node is created to represent the geometry variable as a input. Any inputs incorrectly omitted will be identified in step (D).
 
@@
 	The resulting MCG is shown in Fig.~\ref{f:MCG holes}. 
-	Cycles are shown as dashed lines. FPGs formed from this MCG may or may not retain these cycles.
-	\begin{figure}[htb!]
+	Cycles are shown as dashed lines. FPGs formed from this MCG may or may not retain these cycles,  depending on user choices.
+	\begin{figure*}[htb!]
 	  \begin{center}
@@
 	\label{f:MCG holes}
-	\end{figure}
+	\end{figure*}
 
@@
 
-%	As described in Sec.~\ref{ss:process}, the process of obtaining an FPG from the given MCG is expected to be an iterative process in which the user applies the algorithm to determine whether or not an FPG can be obtained,
-%	and then modifies the indegree limits or the MCG to obtain a satisfactory FPG.
-%	For example, after determining the location of a hole, the user can change the indegree limit to make the node a design variable, or the user could change the MCG by adding a new analysis code such that the hole is resolved.
-%
-%	As presented in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}, the first step to obtaining an FPG is to detect holes. 
-%	To begin, we set the lower indegree limit for every node to be unity.
-
 Step (D) now procededs by executing the FPG algorithm. 
 An FPG cannot be obtained because the variable nodes representing the number of engines are holes for NPSS, FLOPSa, and FLOPSb, and the variable nodes representing the mission definition are holes for FLOPSa and FLOPSb.
-These holes propagated upstream and created holes in the objectives, thereby preventing a valid FPG.
+These holes are propagated upstream by the FPG algorithm and create holes in the objectives, thereby preventing a valid FPG.
 
-Step (D)(a) begins to resolve this conflict by creating a input for the number of engines.
-In this example, the number of engines was not originally included as a input to demonstrate how the process responds to this error.
+Step (D)(a) begins to resolve this conflict by creating an input for the number of engines. 
+In this example, the number of engines was not originally included as an input to demonstrate how the process can reveal the need to select a variable as an input.
 Step (D)(b) then reduces the lower and upper indegree limits for the variable node representing the mission definition to zero, implying that these nodes must be specified as design variables (see Table \ref{t:variable node classification}).
 The new MCG is shown in Fig.~\ref{f:MCG}.
-	\begin{figure}[htb!]
+	\begin{figure*}[htb!]
 	  \begin{center}
@@
 	\label{f:MCG}
-	\end{figure}
+	\end{figure*}
 
@@
 	An example is a metric to obtain an FPG with the fewest possible number of cycles. 
-In this example, the FPG with the fewest number cycles is shown in Fig.~\ref{f:FPG fewest cycles}, in which the edges belonging to a cycle are inticated by dashed lines. 
+In this example, the FPG with the fewest cycles is shown in Fig.~\ref{f:FPG fewest cycles}, in which the edges belonging to a cycle are inticated by dashed lines. 
 For this example problem, the cycles were detected using the implementation of Johnson's algorithm \cite{Johnson1975} in the Python package NetworkX. 
@@
 It is beyond the scope of this work to explore efficient methods for finding the FPGs meeting these criteria.
-However, as discussed by Gabow\cite{Gabow1985}, graph algorithms typically scale well with the size of the graph.
-This suggests that the computation cost of finding an FPG should be small.
-	\begin{figure}[htb!]
+However, as discussed by Gabow \cite{Gabow1985}, graph algorithms typically scale well with the size of the graph.
+This suggests that the computation cost of finding an FPG should be small even for large problems.
+	\begin{figure*}[htb!]
 	  \begin{center}
@@
 	\label{f:FPG fewest cycles}
-	\end{figure}
+	\end{figure*}
 
-	Alternatively, it may be desirable to resolve collisions by choosing the higher fidelity analysis tools.
-	One method to ensure that the edges directed from certain analysis blocks are those chosen to resolve conflicts is to use a ranking system.
+	Alternatively, it may be desirable to resolve collisions by preferentially choosing analysis tools with favorable properties, such as higher accuracy or minimal run time, whenever possible.
+	One method to implement an FPG in this paradigm is to use a ranking system.
 Each analysis block is assigned a value by the designer describing how desireable it is for this block to be present in the FPG.
-The connection edges directed out of each analysis block are then given a weight equal to the value assigned to the analysis block.
-Finally, each collision is resolved by selected the edges with the highest weights.
+The connection edges directed out of each analysis block are then assigned a weight equal to the value assigned to the analysis block.
+Finally, each collision is resolved by selecting the edges with the highest weights.
 For the current example problem, consider the rankings for each analysis code given in Table \ref{t:rankings}.
@@
 	The resulting FPG has four cycles and is shown in Fig.~\ref{f:FPG highest rank}. 
-%This graph has four cycles, which are enumerated in Table \ref{t:cycles}.
-	\begin{figure}[htb!]
+	\begin{figure*}[htb!]
 	  \begin{center}
@@
 	  \end{center}
-	  \caption{FPG obtained using the ranking system.}
+	  \caption{FPG obtained using a ranking system.}
 	\label{f:FPG highest rank}
-	\end{figure}
-%	\begin{table}[htbp]
-%	  \centering
-%	  \caption{Cycles for the FPG obtained using ranking}
-%		\begin{tabular}{ccccccc}
-%		\toprule
-%		output/input & analysis code & output/input & analysis code & output/input & analysis code & output/input \\
-%		\midrule
-%		engine weight & FLOPSb & drag  & NPSS  & engine performance & WATE  & engine weight \\
-%		engine performance & FLOPSb & drag  & NPSS  & engine performance &       &  \\
-%		fuselage weight & FLOPSb & total weight & PDCYL & fuselage weight &       &  \\
-%		wing weight & FLOBSb & total weight & PDCYL & wing weight &       &  \\
-%		\bottomrule
-%		\end{tabular}%
-%	  \label{t:cycles}
-%	\end{table}%
+	\end{figure*}
 
 	Finally, consider a case in which 
-the inviscid drag input into FLOPSb is a multi-fidelity input, meaning that 
+the induced drag input into FLOPSb is a multi-fidelity input, meaning that 
  multiple analysis codes calculate the same variable. 

### 2013-08-07 graph_syntax.tex
files for submitted version and reviewer 3's comments

@@
   indicated by the first entry and directed to the node indicated by the second 
-  entry. An edge $e$ = $(x,y)$ may be referred to simply as $xy$. The edges directed out from node $v$ are denoted by $E(v)$ and the edges directed into $v$ are given 
-  by $E^{-1}(v)$. 
-  If $E$ is not a one-to-one mapping, $E(v)$ may be the empty set, a single node, or a set of nodes.
+  entry. An edge $e$ = $(x,y)$ may simply be referred to as $xy$. The edges 
+  directed out from node $v$ are denoted by $E(v)$ and the edges directed into $v$ are denoted 
+  by $E^{-1}(v)$. $E(v)$ may be the empty set, a single edge, or a set of edges, and likewise for $E^{-1}(v)$.
   \begin{figure}[htb!]
@@
   The set of sets $\mathcal A = \{A_i \st i \in I\}$ is called an indexed family of sets with index $i$ and 
-  indexing set $I$\cite{smith2006}. 
+  indexing set $I$ \cite{smith2006}. 
   The union over this family of sets can be described in a few different ways:
@@
   \label{ss:types}
-Along with these fundamentals of graph theory, we categorize nodes and edges into distinct types in order to provide an intuitive approach to MDAO problem formulation. The three node types are
+We categorize nodes and edges 
+  into distinct types in order to provide an intuitive approach to MDAO problem formulation. The three node types are
     \begin{description}
@@
       \item[\bf{function node:}] represents the computation peformed by analysis tools,
-    \item[\bf{driver node:}] control logic capable of managing iterations (PSG),
+      \item[\bf{driver node:}] represents control logic capable of managing iterations (present only in a PSG),
     \end{description}
@@
   \begin{description}
-  \item[\bf{connection edge:}] exchange of information external to analysis tools,
-  \item[\bf{function edge:}] exchange of information internal to analysis tools,
-  \item[\bf{driven edge:}] passing of information from a driver node to a variable node (PSG).
+  \item[\bf{connection edge:}] represents exchange of information external to analysis tools,
+  \item[\bf{function edge:}] represents exchange of information internal to analysis tools,
+  \item[\bf{driven edge:}] represents passing of information from a driver node to a variable node (present only in a PSG).
   \end{description}
-Figure \ref{f:sellar types} demonstrates the usage of these node and edge types via the Sellar problem.
-function nodes are indicated by squares, variable nodes are indicated by circles, connection edges are indicated by dotted lines, and function edges are indicated by solid lines.
+  Figure \ref{f:sellar types} demonstrates the usage of these node and edge 
+  types via the Sellar problem. 
+In this figure, function nodes are indicated by squares, 
+  variable nodes are indicated by circles, connection edges are indicated by 
+  dotted lines, and function edges are indicated by solid lines.
   \begin{figure}[htb!]
     \begin{center}
-    \includegraphics[width=4.0in]{images/sellar_types}
+      \includegraphics[width=3.25in]{images/sellar_types}
     \end{center}
@@
   A rule set is provided for the usage of these node and edge types to provide a structure to the graph-based representation.
-  The driver node and the driven edge are allowed to be present only in PSGs, whereas the other node and edge types can be present in any of the three graph types, 
+  The driver node and the driven edge are allowed only in PSGs, whereas the other node and edge types can be present in any of the three graph types, 
   subject to the following restrictions: 
@@
 
-  In Alexandrov and Lewis's REMS syntax only two node types (variable and function) 
-  and one edge type are included \cite{alexandrov2004}. The present work adds one additional 
+ Alexandrov and Lewis's REMS syntax includes only two node types (variable and function) 
+  and one edge type   \cite{alexandrov2004}. The present work adds one additional 
   node type and two additional edge types to the syntax to allow descriptive
-  graphs for all three phases of the design problem formulation process indicated in Figs.~\ref{f:tree} and \ref{f:hourglass}.
+  graphs for all three phases of the design problem formulation process discussed in Sec. \ref{s:requirements}.
 
@@
 \label{ss:analysis blocks}
-Analysis tools take in a set of input variables and then calculate 
+  Analysis tools take in a set of input variables and calculate 
   the values for their respective outputs. We represent this process
@@
   %Each analysis block is a directed graph denoted as $ A=(V_{A},E_{A})$
-As indicated in this figure, analysis blocks comprise three sets of nodes representing the distinct local inputs, the analysis (computation), and the distinct local outputs.
-The local input and local outputs nodes are variable nodes, while the nodes representing the analysis are function nodes.
-All of the edges within the analysis block are function edges, and are considered fixed to the analysis block.
-This graph structure demonstrated by Fig.~\ref{f:analysis block} satisfies the rules listed in Sec.~\ref{ss:types}.
-Reciprocally, the rules ensure that analysis tools are represented via this analysis block structure.
+  As indicated in this figure, analysis blocks comprise three sets of nodes 
+  representing the distinct local inputs, the analysis (computation), and the 
+  distinct local outputs. The local input and local output nodes are variable 
+  nodes, while the nodes representing the analysis are function nodes. All of 
+  the edges within the analysis block are function edges, and are considered 
+  fixed to the analysis block. This graph structure demonstrated by 
+  Fig.~\ref{f:analysis block} satisfies the rules listed in Sec.~\ref{ss:types}. 
+  Conversely, the rules ensure that analysis tools are represented via this analysis block structure.
+
   \begin{figure}[htb]
@@
 
-The function edge connecting the two function nodes represents the necessary calculations to map the inputs to the outputs of the analysis. 
-This edge also provides the opportunity to encode computational cost or other characteristics of the analysis code as a weight in a weighted graph formulation. 
-Since all the edges within analysis blocks are fixed, the blocks themselves are fixed sub-graphs within the overall MDAO problem graph. 
+  The function edge connecting the two function nodes represents the necessary 
+  calculations to map the inputs to the outputs of the analysis. 
+  This edge also provides the opportunity to encode computational cost or other 
+  characteristics of the analysis code as a weight in a weighted graph formulation. 
+  Because all the edges within analysis blocks are fixed, the blocks themselves are fixed sub-graphs within the overall MDAO problem graph. 
   The connectivity of nodes and edges in an analysis block cannot be altered for the purposes of MDAO problem formulation; 
@@
   Along with analysis tools, objectives, constraints, and inputs also need to be represented.
-In the case of objective functions, a single output value generated by an analysis block could be selected, but commonly multiple values are assembeled together to form a composite objective function.
+  In the case of objective functions, a single output value generated by an analysis block could be selected, but commonly, multiple output values from different analysis tools are aggregated together to form a composite objective function.
   Generally, both objective and constraint functions accept a set of inputs and map them to an output value of significance to the overall design problem.
@@
   A composite constraint or objective function can therefore be represented as an analysis block within the graph with its own input and output variable nodes. 
-Although fundamentally no different than an analysis block, for clarity and convience, it is useful to distinguish between analysis blocks corresponding to analysis codes and those that arise from the addition of objectives or constraints to the graph.
+  Although fundamentally no different than an analysis block, for clarity and convenience, it is useful to distinguish between analysis blocks corresponding to analysis codes and those that arise from the addition of objectives or constraints to the graph.
   Therefore, we use an \emph{expression block} to represent an objective or constraint. 
@@
   Finally, inputs to the problem formulation are represented in the graph by individual variable nodes with connection edges directed out to corresponding local inputs of analysis blocks, and they are referred to simply as \emph{inputs}.
-These nodes serve to indicate that the variable nodes they are connected to represent the same variable in the problem formulation.
-These nodes serve to unify the variable nodes they are connected to. That is, they indicate that these nodes refer to the same input variable.
+  These nodes serve to indicate that the variable nodes within analysis or expression blocks to which they are connected represent the same variable in the problem formulation.
 
@@
     \begin{center}
-    \includegraphics[width=4.0in]{images/sellar_obj_and_cons}
+      \includegraphics[width=3.25in]{images/sellar_obj_and_cons}
     \end{center}
@@
   is the number of edges directed out and it is denoted as $\txt{deg}^+(v)$ \cite{Diestel2010}.
-  In this paper, the indegree and outdegree of a given node refer only to the number of the connection edges 
-  directed into or out of the node, respectively.
 
@@
   These user-specified limits govern the number of connection edges that may 
-  be directed into a node for a valid graph; of course, $\txt{deg}_l^-(v) \leq \txt{deg}_u^-(v)$ must be satisfied. 
-For example, consider a variable 
-  node $v$ with $\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, $v$ 
+  be directed into a node for a valid graph; of course, the specification must satisfy $\txt{deg}_l^-(v) \leq \txt{deg}_u^-(v)$.
+  For example, consider a variable node $v$ with $\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, $v$ 
   must have exactly one incoming connection edge or the graph is deemed an invalid problem formulation.
@@
   \label{s:graph representation}
-In Sec.~\ref{s:syntax definition} we presented the structure of the graph-based representation.
+  In Sec.~\ref{s:syntax definition}, we presented the structure of the graph-based representation.
 We now discuss how these formalisms provide the remaining MDAO problem constructs identified in Sec.~\ref{s:requirements}.
-
-%To begin, we make use of the Sellar problem, described in 
-%Eq. \ref{eqn:simple_fpf} as an example to illustrate the implementation of the syntax. 
-%The graph representation of the Sellar problem, using the new stynax, is 
-%presented in Fig. \ref{f:sellar_graph_full}. In this section sub-graphs or 
-%slight modifications of the full graph will be used to demonstrate the 
-%important MDAO constructs described by this graph syntax.
-%\begin{figure}[htb!]
-%    \begin{center}
-%    \includegraphics[width=4in]{images/sellar_graph_full}
-%    \end{center}
-%    \vspace{-10pt}
-%\caption{Graph of the Sellar problem formulation}
-%\label{f:sellar_graph_full}
-%\end{figure}
 
@@
   Design variables are input variables whose values are free to be 
-changed by the designer or by an optimizer in an MDAO study. In the graph-based syntax, a node corresponds to a design variable when the indegree limit of the node is prescribed as zero.
+  changed by the designer or by an optimizer in an MDAO study. In the graph-based 
+  syntax, a node corresponds to a design variable if the indegree limits of the node are prescribed as zero.
 
-If the indegree limits are violated by the number of incoming connection edges (the indegree of the node), the graph is regarded as an invalid problem formulation.
-There are two classifications for these violations:
+  If the indegree limits are violated by the number of incoming connection edges 
+  (the indegree of the node), then the graph is regarded as an invalid problem formulation. There are two classifications for these violations:
     \begin{description}
@@
             A hole represents a lack of information being supplied to the variable node. 
-This implies that the analysis tool being represented by the analysis block would not be capable of producing outputs.
-    \item[\bf{collision:}] The number of incoming edges is greater than the upper indegree limit. 
+            This implies that the analysis tool being represented by the analysis block 
+            would not be capable of producing outputs.
+          \item[\bf{collision:}] The number of incoming edges is greater than the upper indegree limit:
             \begin{equation} 
@@
       \label{e:collision}
-\end{equation}A collision represents redundant information being supplied to the variable node from two or more sources.
-This conflict implies an ambiguity as to which information is to be used as an input for the analysis tool.
+      \end{equation}A collision represents redundant information being supplied to 
+      the variable node from two or more sources. This conflict implies an 
+      ambiguity as to which information is to be used as an input for the analysis tool.
+	In multifidelity problems (discussed subsequently), the upper indegree limit could be specified as $deg_u^- > 2$, which implies that a collision is not noted in the case of multiple inputs.
   \end{description}
@@
   The presence of holes and collisions in a graph represent conflicts that will give
-  rise to an invalid problem formulation. It is expected that a maximal connectivity graph could indicate these conflicts, whereas a fundamental problem graph will not.
+  rise to an invalid problem formulation. 
+The process of obtaining an FPG from an MCG will reveal these conflicts.
 
 \subsection{Coupling Between Analyses}
-  In MDAO, coupling is the mutual dependence the outputs of two or more analysis 
-  codes. In this graph-based syntax, coupling is described by the presence of a 
-  cycle between two or more analysis blocks. In the Sellar problem from Eqn \ref{eqn:simple_fpf} 
-  a reciprical dependence between $D_1$ and $D_2$ exists through the variabled $y_1$ and $y_2$. 
+  In MDAO, coupling is the mutual dependence between two or more analysis 
+  tools and their respective outputs. In this graph-based syntax, coupling is described by the presence of a 
+  cycle between two or more analysis blocks. In the Sellar problem from Eq.~\ref{eqn:simple_fpf} 
+  a mutual dependence between $D_1$ and $D_2$ exists through the variables $y_1$ and $y_2$. 
   The edges belonging to a cycle in the Sellar problem are highlighted in 
-  Fig. \ref{f:sellar cycles} where they form a closed path between analysis blocks $D_1$ and $D_2$.
+  Fig. \ref{f:sellar cycles}.  These edges collectively form a closed path between analysis blocks $D_1$ and $D_2$.
 
@@
   \begin{center}
-    \includegraphics[width=4.0in]{images/sellar_cycles}
+    \includegraphics[width=3.25in]{images/sellar_cycles}
   \end{center}
@@
   Solvers, optimizers, and other iterative processes are not invovled in the coupling 
-  definition; however, one or more solvers will be necessary to build a valid PSG from a given FPG. 
-  Additionally, a coupling cycle has no inherent start or end point. It would be acceptable to select
+  definition; however, one or more solvers will be required to build a valid PSG from a given FPG  that includes cycles. 
+  Additionally, a coupling cycle has no inherent start or end point. It would be acceptable to select a
   node in the cycle as a starting point and proceed around the
@@
   Eq.~\ref{eqn:simple}.
-  Larger problems can contain more complex cycles in their FPG, indicating more 
+  Larger problems can involve more complex cycles in their FPG, indicating more 
   complex coupling between analyses. For example, a cycle can involve more than 
@@
   searching for an effective ordering for the execution of analyses.
-  Rogers' DEMAID tool uses a genetic algorithm to find an ordering that minimizes 
+  Rogers' DeMAID tool uses a genetic algorithm to find an ordering that minimizes 
   the overall coupling of the system by separating independent cycles in the 
@@
   a weighted form of the DSM and used an iterative clustering approach to perform a 
-  similar task to DEMAID \cite{Lu2012}.
+  similar task to DeMAID \cite{Lu2012}.
 
@@
   A multi-fidelity MDAO problem can be characterized by a formulation in which 
-  two or more different analyses each calculating the same data. Multi-fidelity 
+  two or more different analyses each calculate the same data. Multi-fidelity 
   is represented in a graph by a variable node having an indegree greater than 
@@
   In a multi-fidelity problem, a given variable node may have multiple incoming 
-  edges without implying a collision as defined by by Eq.~\ref{e:collision}. 
-An  upper indegree limit above unity for any variable in a graph represents a decision to 
-  allow multiple fidelities to interact at that part of the graph in order to allieviate
-  a conflict. 
+  edges without implying a collision as defined by Eq.~\ref{e:collision}. 
   Figure \ref{f:collision_example} shows a modified version of the Sellar problem 
   with a new analysis, $D_0$, representing a low fidelity version of $D_1$. 
-  In this modified version, the variable nodes with an upper indegree limit of two are highlighted in gray, and changing these upper degree limits for the variable nodes representing $y_1$ in $D_2$, $F$, and $G_1$ avoids a collision.
-In section \ref{ss:obtaining FPG} we present an algorithm for finding 
-  conflicts within a given MCG so that a designer can make the necessary decisisons about each 
+  In this modified version, the variable nodes with an upper indegree limit of two are highlighted in gray, and changing the upper indegree limits for the variable nodes representing $y_1$ in $D_2$, $F$, and $G_1$ avoids a collision.
+In Sec.~\ref{ss:obtaining FPG} we present an algorithm for locating 
+  conflicts within a given MCG such that a designer can make the necessary decisisons about each 
   one in turn. 
@@
     \begin{center}
-      \includegraphics[width=4in]{images/sellar_mulfi}
+      \includegraphics[width=3.25in]{images/sellar_mulfi}
     \caption{Graph of the modified Sellar problem with multi-fidelity nodes highlighed in gray.\label{f:collision_example}}
@@
 
-  Multi-fidelity problems are always charaterized by graphs with variable 
+  Multi-fidelity problems are always charaterized by graphs in which at least some of the variable 
   nodes that have an $\txt{deg}^-_u(v) > 1$. These problems require 
@@
   to manage when each of the different fidelity analyses are 
-  run\cite{march2012provably,alexandrov2001approximation,Huang_Allen_Notz_Miller_2006}.
-  The specifics of this mechanism are not given in any form withing an MCG or 
-  FPG. Instead the multi-fidelity mechanism specifics would be represented as a 
+  run \cite{march2012provably,alexandrov2001approximation,Huang_Allen_Notz_Miller_2006}.
+  The specifics of this mechanism are not represented within an MCG or 
+  FPG. Instead, the multi-fidelity mechanism specifics would be represented as a 
   driver in a PSG derived from a multi-fidelity FPG.
 
-Table \ref{t:variable node classification} provides examples to summarize the classification of a variable node as a hole, design variable, single valid input (nominal), collision, or multi-fidelity.
+Table \ref{t:variable node classification} summarizes the classification of a variable node as a hole, design variable, single valid input (nominal), collision, or multi-fidelity node.
 

### 2013-08-07 intro.tex
files for submitted version and reviewer 3's comments

--- 2013-05-09 intro.tex
@@
     Multidisciplinary Design Analysis and Optimization (MDAO)
-    frameworks such as OpenMDAO\cite{Gray2012} and {ModelCenter}\textregistered have enabled a new level of analysis tool integration 
+    frameworks such as OpenMDAO \cite{Gray2012} and {ModelCenter}\textregistered have enabled a new level of analysis tool integration 
     and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. This 
@@
     A common situation is that different analyses provide differing values for the 
-    same physical quantity, and these conflicts need to be resolved. These differing values 
-    may not even be in the same format. These occurrences are particularly acute when analysis tools have differing fidelities. For
+    same physical quantity, often in different data formats, and these conflicts need to be resolved. 
+   These occurrences are particularly acute in situations in which analysis tools have differing fidelities. For
     example, an abstracted aerodynamic analysis such as an empirical drag buildup model may return only
@@
     Our goal is to develop formalism for expressing analysis interconnectivity and for determining feasible
-    data flows to assist an engineering team conducting this task. Because the problem deals with
-    interconnectivity, we base our approach on the representations and techniques of graph theory. The 
-    proposed graph syntax is closely related to the REMS system proposed by 
-    Alexandrov and Lewis\cite{alexandrov2004}. The approach taken here should be viewed as an extention 
-    of the concepts proposed in REMS with the goal of allowing the graphs to more closely interact with 
-    existing design processes and MDAO frameworks.
-
-The approach begins by 
-    constructing the \emph{maximal connectivity graph (MCG)} describing all possible
+    analysis tool sets to assist an engineering team conducting this task. Because the problem deals with
+    interconnectivity, we base our approach on the representations and techniques of graph theory.
+    The approach begins by constructing the \emph{maximal connectivity graph (MCG)} describing all possible
     interconnections between the analysis tools proposed by the engineers. Graph operations are then
-    conducted to reduce the MCG down to a \emph{fundamental problem graph (FPG)} that describes the set of analysis
+    conducted to reduce the MCG to a \emph{fundamental problem graph (FPG)} that describes the set of analysis
     tools needed to solve the specified system-level design problem.
-    Any particular solution procedure for the problem; any relevant MDO solution architecture, e.g. MDF,
-    IDF, CO, BLISS, could be selected \emph{post facto} to implement the system-level optimization by
-    constructing a \emph{problem solution graph (PSG)} and the corresponding problem solution approach.
-    The concept of the FPG and the identification of feasible FPGs are the main contributions of the paper.
+    Any particular solution procedure for the problem, such as the choice of an MDO architecture, e.g. MDF,
+    IDF, CO, BLISS, could then be described by
+    constructing a \emph{problem solution graph (PSG)}.
+    The concept of the FPG and the identification of feasible FPGs from an MCG are the main contributions of the paper.
 
@@
     engineers desire to solve and on the available analysis tools, and a specific problem formulation, which
-    additionally presumes a specific MDO solution approach to the problem. Next, we survey the literature in applying
+    additionally presumes a specific MDO solution approach to the problem. Next, we survey the literature related to
     graph theoretic and formal language approaches to multidisciplinary design problem formulation. 
@@
     determining the MCG and FPG. Finally, we present an example problem based on an MDO analysis of a 
-    commercial aircraft and discuss additional applications enabled by our approach.
+    commercial aircraft.
 
@@
 	\label{s:specific vs fundamental}
-    The Fundamental Problem Formulation (FPF) is our terminology for a 
-    statement of the overall system-level optimization problem that does not 
-    depend on the choice of solution-specifc elements such as the MDAO 
-    framework, optimization algorithm, iterative solver, or execution sequence
-    used to solve the problem. The FPF contrasts with problem statements that are written without 
-    reference to a specific solution strategy, but that still imply a specific execution sequence.
+    The Fundamental Problem Formulation (FPF) is our terminology for a statement of the overall system-level optimization problem that contains only  information about analysis tools, design variables, objectives, and constraints without reference to a solution approach.
+In particular, the FPF does not depend on the choice of solution-specific elements such as the MDAO framework, optimization algorithm, iterative solver, or execution sequence. 
+This description represents  the problem from the point of view of an engineer specifying the design problem that he/she desires to solve without specifying how the problem should be solved. 
+The lack of solution information in the FPF contrasts with optimization problem statements that are written with reference to a specific solution strategy or with implication of a specific execution sequence.
 For example, consider the Sellar test problem \cite{AIAA:sellar}, with an FPF given as follows:
@@
 
-    In Eq. \ref{eqn:simple_fpf}, $y_1$ is an output of $D_1$ as well as an input 
+    In Eq.~\ref{eqn:simple_fpf}, $y_1$ is an output of $D_1$ as well as an input 
     to $D_2$. Similarly, $y_2$ is an output of $D_2$ as well as an input 
     to $D_1$. This implies that $D_1$ and $D_2$ are coupled together by means 
-    of their reciprical dependence. However no information is given as to weather
-    $D_1$ or $D_2$ should be run before the other. Given Eq \ref{eqn:simple_fpf}, 
+    of their mutual dependence. However no information is given as to weather
+    $D_1$ or $D_2$ should be run before the other. Given Eq.~\ref{eqn:simple_fpf}, 
     it would be equally valid to run $D_1$ first, $D_2$ first, or both simultaneously. 
+Any MDAO root finding method or compatibility constraint formulation could be implemented to achieve consistency between $D_1$ and $D_2$.
 
-    Eq. \ref{eqn:simple} gives a modified problem formulation where $D_1$ is run 
-    first. A new variable is introduced, $\hat{y_2}$, to break the dependence of
-    $D_1$ on $D_2$. Also a new coupling constraint, $H$, to enforce the consistency
-    that would otherwise be lost due to $\hat{y_2}$. Equation \ref{eqn:simple} is 
-    a specific alternative expression of the more fundamental problem description 
-    given in Eq. \ref{eqn:simple_fpf}.
+    The Sellar problem is simple, with only a single coupling interaction 
+    between the two disciplines and a very limited set of variables. 
+ However, for larger and more complex problems, 
+    it is much more difficult to identify the fundamental problem formulation. Interdisciplinary 
+    couplings are not always apparent and the set of analysis tools and variables are 
+    much larger.
+
+An example of a more specific problem statement for the Sellar problem is presented in Eq. 2.  
+In this formulation, the implication is that  $D_1$ must be run 
+    first. A new variable, $\hat{y_2}$, is introduced to break the direct dependence of
+    $D_1$ on $D_2$, and a new coupling constraint, $H$, is added to enforce 
+    consistency. Equation \ref{eqn:simple} is 
+    an equally valid expression of the Sellar problem that could be 
+    generated  based on a preferred solution approach.
 
@@
 
-    Figure \ref{f:sellar graph} depicts the Sellar problem given by 
-    Eq.~\ref{eqn:simple_fpf} in the graph-based syntax which will be presented 
-    in this paper. This example will serve to highlight the features of the 
-    syntax as they are introduced.
-    \begin{figure}[htb!]
-      \begin{center}
-        \includegraphics[width=4.0in]{images/sellar}
-      \end{center}
-      \caption{Sellar problem represented as a graph.}
-    \label{f:sellar graph}
-    \end{figure}
-
-\subsection{Existing Graph-Based Syntax}
+\subsection{Background in Graph-Based Descriptions of MDAO Problems}
 	\label{s:existing syntax}
     As indicated in the examples in Sec.~\ref{s:specific vs fundamental}, the mathematical 
-    language for specifying problem formulations is very general and can be used both for 
-    fundamental and specific problem formulations. Tedford and Martins used this mathematical syntax to specify the 
+    language for specifying optimization problem formulations is very general and can be used both for 
+    fundamental and specific problem formulations. Tedford and Martins used this syntax to specify the 
     FPF for a set of test problems and also to describe specific formulations for solving them with a 
-    number of optimization architectures\cite{Tedford2009}. Their work demonstrates how multiple specific 
-    problem formulations can relate to a common FPG. The challenge with this 
-    traditional mathematical syntax is that it is not easily manipulated or analyzed.
+    number of optimization architectures \cite{Tedford2009}.
+By solving different specific problem statements corresponding to the same FPF, they were able to  benchmark the performance of different optimization architectures against a fixed set of
+    problems. Gray et al. also benchmarked the performance of a set of MDAO architectures  with a similar approach \cite{Gray2013}. 
+    This work examined a larger set of architectures and proposed the use of OpenMDAO as a platform to build a larger community-developed set of test problems and architectures.
+Both of these benchmarking efforts
+    demonstrate how multiple specific problem formulations can relate to a common FPF and indicate the 
+    value of a common problem descrition. The challenge with this 
+    traditional syntax is that it is not easily manipulated or analyzed with automatic procedures to explore alternate problem formulations.
 
@@
     Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools.
-    Off-diagonal elements of the matrix indicate coupling\cite{Steward1981}. Since a DSM describes a square adjacency matrix, 
+    Off-diagonal elements of the matrix indicate coupling \cite{Steward1981}. Since a DSM describes a square adjacency matrix, 
     it can be represented in an equivalent directed graph in which nodes represent analysis tools and 
@@
     Rogers et al.~developed DeMAID to manipulate a DSM to find an ordering for analysis tools that 
-    reduces the cost of solving highly coupled systems\cite{rogers1996,rogers1996demaid}. This re-ordering is done through 
-    row operations on the DSM matrix and yields multiple specific problem 
+    reduces the cost of solving highly coupled systems \cite{rogers1996,rogers1996demaid}. This re-ordering is done through 
+    operations on the DSM matrix and yields multiple specific problem 
     formulations which all solve the same FPF.
 
-    A DSM by itself is insufficient to describe complete problem formulations because it 
+    A DSM itself is insufficient to describe complete optimization problem formulations because it 
     captures only information about data dependency between analyses; objective and 
@@
     An alternative matrix-based syntax, called a Functional Dependency Table (FDT), was proposed by Wagner and Papalambros. 
-    FDT represents the relationship between functions, including objectives and constraints, and specific variables that affect 
-    them\cite{Wagner1993}. Similar to DSM, FDT also describes an adjacency matrix of a graph. Unlike the DSM graph, 
+    The FDT represents the relationship between functions, including objectives and constraints, and specific variables that affect 
+    them \cite{Wagner1993}. Similar to DSM, FDT also describes an adjacency matrix of a graph. Unlike the DSM graph, 
     however, the FDT graph is undirected and nodes can represent analysis tools, objectives, 
@@
     variable. Michelena and Papalambros made use of the FDT to solve a graph partitioning problem that yielded 
-    more efficient optimization problem decompositions\cite{Michelena1997}. While FDT succeeds at capturing the 
+    more efficient optimization problem decompositions \cite{Michelena1997}. Allison 
+    extended this work to incorporate adjacency matrix information with the FDT to 
+    account for system coupling in an automated partitioning scheme \cite{Allison2008}.
+
+While FDT succeeds at capturing the 
     information about objectives and constraints, its lack of directed edges 
-    implies that it cannot capture the coupled data dependency that a DSM captures. 
-    The FDT in Fig. \ref{fig:FDT_simple} shows that $D_1$ is dependent on $y_2$ and 
-    that $D_2$ is dependent on $y_1$, but you can not infer their coupled dependence from 
+    implies that it cannot describe the coupled data dependency that a DSM captures. 
+Figure \ref{fig:FDT_simple}  shows the FDT for the Sellar problem described by Eq.~\ref{eqn:simple_fpf}.
+    The FDT shows that $D_1$ is dependent on $y_2$ and 
+    that $D_2$ is dependent on $y_1$, but the coupled dependence cannot be inferred from 
     that information alone. This missing information implies 
-    that, while FDT is very useful for partitioning problems, it is also not 
-    sufficient to completely describe a problem formulation. 
+    that, although FDT is very useful for partitioning problems, it is not 
+    sufficient to completely describe the data flows in a problem formulation.
+
     \begin{figure}[htb!]
@@
     incorporates objectives and constraints into a graph, effectively combining 
-    FDT and DSM\cite{alexandrov2004}. REMS retains the square adjacency 
+    FDT and DSM \cite{alexandrov2004}. REMS retains the square adjacency 
     matrix form from DSM, but by adding the objectives and constraints, it partially 
@@
     dependency between multiple analysis tools as well as between analysis tools and
-    objective/constraint functions. Additionaly, REMS address the need to be able to 
-    transition between multiple solution strategies while keeping a constant 
+    objective/constraint functions. Additionaly, REMS address the need to
+    transition between multiple solution strategies while maintaining a single consistent  
     graph representation of the fundamental problem formulation. REMS provides
     syntax for variables and functions but does not facilitate inclusion 
-    of solvers or optimizers in the graph.
+    of solvers or optimizers in the graph.  The 
+     graph syntax proposed in this paper could  
+    be viewed as an extention to REMS with the goal of allowing the graphs to  describe analysis couplings and more 
+    closely interact with existing design processes and MDAO frameworks.
 
@@
     in an Extended DSM (XDSM) in order to capture a more complete description 
-    of solution strategies for MDAO problems\cite{Lambe2012}. Unlike REMS, 
-    XDSM also includes nodes for solvers and optimziers to enable complete 
+    of solution strategies for MDAO problems \cite{Lambe2012}. Unlike REMS, 
+    XDSM also includes nodes for solvers and optimizers to enable complete 
     definition of MDAO architectures. Martins uses XDSM to describe 13 different 
@@
     additional information included in an XDSM, Lu and Martins applied both 
-    ordering and partitioning algorithms on an MDAO test problem termed the 
+    ordering and partitioning algorithms to an MDAO test problem termed the 
     Scalable Problem \cite{Lu2012}.
@@
     the relationship between design variables and objectives/constraints. 
-    By introducing solver or optimizer blocks, XDSM automatically implies a strategy. 
-    The XDSM for Eq. \ref{eqn:simple} is 
-    given in Figure \ref{fig:XDSM_simple}. This diagram is shown with an 
+    By introducing solver or optimizer blocks, XDSM automatically implies a solution strategy. 
+    The XDSM for Eq. \ref{eqn:simple} is given in Figure \ref{fig:XDSM_simple}. 
+    This diagram is shown with an 
     assumed Gauss-Seidel iteration scheme and an MDF solution architecture.
 
-    Of these methods, REMS can describe most aspects of a fundamental 
-    problem formulation, but lacks the ability to represent optimizers and 
-    solvers in more specific problem formulations. XDSM relies on the provision of  
-    optimizers and constraints and is not formulated  
-    for a more fundamental problem statement. We propose a new graph syntax that 
-    combines the relevant features of REMS and XDSM while retaining the flexiblity to 
+In this paper, we propose a new graph syntax that 
+    combines certain features of REMS and XDSM. 
+The approach allows interactions and couplings of analysis tools and objective and constraint functions to be described while retaining the flexiblity to 
     represent a fundamental problem formulation free of solution specific information.
@@
   \label{s:requirements}
-  The goal of the new graph-based syntax presented here is to enable the general structure of an MDAO problem to be described independently of any solution information, 
+  The goal of the graph-based syntax presented here is to enable the general 
+  structure of an MDAO problem to be described independently of any solution information, 
   while still being able to accomodate the more specific case when a solution 
@@
 
-  The new syntax is intended to represent three phases of the design problem 
+  The syntax is intended to represent three phases of the design problem 
   formulation process. In the initial problem definition phase, the specific 
@@
   is selected, e.g. selecting an MDAO optimization architecture. Using 
-  the proposed graph syntax, these phases can be represented with the following 
-  three graphs:
+  the proposed graph syntax, the outcome of these phases can be represented with the following graphs:
   \begin{itemize}
@@
 
-  The \emph{maximal connectivity graph} represents the first phase of the problem formulation with all 
-  analysis tools being considered and all possible connections between them also present. The second graph 
-  is the \emph{fundamental problem graph}, which is comprises only the analyses participating to solve the problem. Finally, a \emph{problem solution graph} 
-  may be represented by including additional edges and nodes to represent the 
-  solution strategy being employed to solve the problem. This paper focuses on the MCG and FPG and does not describe the PSG in detail.
+  The \emph{maximal connectivity graph} represents the first phase of the problem 
+  formulation with all analysis tools being considered and all possible connections 
+  between them also present. The second phase of problem formulation results in the \emph{fundamental problem graph}, 
+  which comprises only the analysis tools, objectives, constraints, and variables to solve the problem. 
+  The final phase results in a \emph{problem solution graph} which includes additional 
+  edges and nodes to represent the solution strategy being employed to solve the 
+  problem. This paper focuses on the MCG and FPG and does not describe the PSG in detail.
 
-  The relationship between these three graphs is depicted in Figs.~\ref{f:tree} and \ref{f:hourglass}. 
-  The tree diagram demonstrates the fact that it is generally possible to obtain 
+  Comparisons of the number and size of each of these graphs are depicted in Figs.~\ref{f:tree} and \ref{f:hourglass}. 
+  The tree diagram illustrates that it is generally possible to obtain 
   multiple FPGs from a single maximal connectivity graph. The multiple FPGs may correspond to 
   different down-selections of analysis tools, different connections between the tools, 
-  or both. Each down-selection shrinks the number of possible FPGs that could be reached 
-  until only one remainins. For a given FPG, different PSGs may be obtained by implementing 
+  or both. Each down-selection reduces the number of possible FPGs that could be reached 
+  until only one remains. For a given FPG, however, different PSGs may be obtained by implementing 
   different solution strategies. Considering the size of a graph to be the sum of all of its
@@
   and the PSG is obtained from the FPG by \emph{adding} nodes and edges.
+These additions correspond to optimizers, solvers, and compatibility constraints required in particular MDAO solution architectures.
 \begin{figure}[htb!]

### 2013-08-07 mcg_fpg.tex
files for submitted version and reviewer 3's comments

@@
     \label{s:building graphs}
-We now discuss how the graph-based syntax defined in Sec.~\ref{s:syntax definition} is used to represent an MDAO problem in its entirety. This discussion is based on the MCG and the FPG.
+We now discuss how the graph-based syntax defined in Secs.~\ref{s:syntax definition} and \ref{s:graph representation}  is used to represent an MDAO problem in its entirety.
+This discussion is based on the MCG and the FPG.
 
@@
     Each of the $m$ analysis codes correspond to an index $i\in I_A$, $I_A=\{1,2,\ldots,m\}$, and are represented by an analysis block graph $G_{A_i}=(V_{A_i},E_{A_i})$.
-    Each of the $n$ expression blocks correspond to an index $i\in I_E$, $I_E=\{1,2,\ldots,n\}$, and are represented by an expression block graph $G_{E_i}=(V_{E_i},E_{E_i})$.
+    Each of the $n$ expression blocks correspond to an index $i\in I_E$, $I_E=\{1,2,\ldots,n\}$, and is represented by expression block graph $G_{E_i}=(V_{E_i},E_{E_i})$.
     Finally, the inputs are represented as a set of variable nodes $V_\txt{in}$.
-    We presume that $V_\txt{in}$, each $G_{A,i}$, and each $G_{E,i}$ are given, and that any potential connection between variables is specified in the form of connection edges in the set $E_{M,\txt{con}}$.
+    We presume that $V_\txt{in}$, each $G_{A_i}$, and each $G_{E_i}$ are given, and that any potential connection between variables is specified in the form of connection edges in the set $E_{M,\txt{con}}$.
 One method for defining these connections is to use a consistant variable naming convention.
+
     We may then construct the maximal connectivity graph $M=(V_M,E_M)$ as
@@
 \begin{equation}
-V_M = V_{M,\txt{var}} \cup V_{M,\txt{mod}}, \ V_{M,\txt{var}} \cap V_{M,\txt{mod}} = \emptyset,
+V_M = V_{M,\txt{var}} \cup V_{M,\txt{fun}}, \ V_{M,\txt{var}} \cap V_{M,\txt{fun}} = \emptyset,
 \end{equation}
-where $V_{M,\txt{var}}$ and $V_{M,\txt{mod}}$ are the sets of variable nodes and function nodes, respecively;
+where $V_{M,\txt{var}}$ and $V_{M,\txt{mod}}$ are the sets of variable nodes and function nodes, respectively;
 \begin{equation}
-E_M = E_{M,\txt{con}} \cup E_{M,\txt{mod}}, \ E_{M,\txt{con}} \cap E_{M,\txt{mod}} = \emptyset,
+E_M = E_{M,\txt{con}} \cup E_{M,\txt{fun}}, \ E_{M,\txt{con}} \cap E_{M,\txt{fun}} = \emptyset,
 \end{equation}
-where $ E_{M,\txt{con}}$ and $E_{M,\txt{mod}}$ are the sets of connection edges and function edges, respectively. These sets will be referenced in the process for obtaining an FPG.
+where $ E_{M,\txt{con}}$ and $E_{M,\txt{fun}}$ are the sets of connection edges and function edges, respectively. These sets will be referenced in the process for obtaining an FPG.
 
-\subsection{Fundamental Problem Graph}
+\subsection{Fundamental Problem Graph (FPG)}
     \label{ss:FPG}
@@
 %    \item[(2)] $\displaystyle{\forall i \in I, \txt{ if } F \cap A_i \neq \emptyset \txt{ then } A_i \subset F}$
-    \item[(3)] $\displaystyle{\forall v \in V_F \txt{ with }v \in V_{M,\txt{var}} \txt{ and }  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
-    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_{E_i}}$
+    \item[(3)] $\displaystyle{\forall v \in V_F \txt{ with }v \in V_{M,\txt{var}}}$ \txt{ and } \\
+			 $\displaystyle{ \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
+    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F }$ \\
+			$\displaystyle{\txt{ from } x \txt{ to } v \txt{ with } x\in V_{E_i}}$
 %    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_\txt{out}}$
@@
     if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is the location of a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is the location of a \emph{collision}.
-    Lastly, condition (4) ensures that only the nodes that are being used are included in the FPG by requiring that for every node a reverse path exists from at least one expression block to that node.
+    Lastly, condition (4) ensures that only the nodes that are used in the MDAO problem formulation are included in the FPG by requiring that for every node a reverse path exists from at least one expression block to that node.
     The reverse graph $R_F$ is obtained from $F$ by simply switching the orientation of every edge in $E_F$.
@@
 
-\subsection{Algorithm for obtaining the Fundamental Problem Graph from the MCG}
+\subsection{Algorithm for Obtaining the Fundamental Problem Graph from the MCG}
     \label{ss:obtaining FPG}
     In general, there may be zero, one, or many graphs that satisfy the FPG conditions in Sec.~\ref{ss:FPG}.
-    Here, we describe an algorithm for obtaining an FPG by starting with the MCG and disconnecting connection edges until the FPG conditions are met. 
+    Here, we describe an algorithm for obtaining an FPG from a given MCG by starting with the MCG and disconnecting connection edges until the FPG conditions are met. 
     With this approach, the problem is reduced to deciding which connection edges to remove.
@@
         \end{equation}
-        with $F_0 = (V_{F_0},E_{F_0})$, and $C_{F_0,\txt{con}}$ is the set of connection edges. 
+        where $F_0 = (V_{F_0},E_{F_0})$, and $C_{F_0,\txt{con}}$ is the set of connection edges. 
         The set of variable nodes that are holes is identified as
@@
         which is the set of variable nodes with fewer incoming edges than are allowed by the lower indegree limit.
-        The updated set of connection edges is then created by removing the edges preceding or succeeding the analysis block:
+        An updated set of edges is then created by removing the connection edges preceding or succeeding the analysis block:
+\begin{IEEEeqnarray*}{rCl}
+        E_{F_1} = E_{F_0} \setminus &\{& (x,y) \in E_{F_0,\txt{con}} \st x \in V_{A_i} \\
+& &  \txt{ or } y \in V_{A_i}, \txt{ and } H\cap V_{A_i} \neq \emptyset \},
+\end{IEEEeqnarray*}
+and the graph is updated as
 \begin{equation}
-        E_{F_1,\txt{con}} = E_{F_0,\txt{con}} \setminus \{(x,y) \in E_{F_0,\txt{con}} \st x \in V_{A_i} \txt{ or } y \in V_{A_i}, \txt{ and } H\cap V_{A_i} \neq \emptyset \}.
+F_1 = (V_{F_0},E_{F_1}).
 \end{equation}
+
 Because removing these edges can create new holes, this step must be repeated until no additional holes are found.
@@
 
-If this step finishes without finding a hole in an expression block, it is guaranteed that an FPG can be obtained because $F_1$ now satisfies all four conditions from Sec.~\ref{ss:FPG} expect for (3), which requires there to be no holes or collisions in the graph.
-Since no more holes remain, and collisions can always be resolved without producing a hole, (see Table \ref{t:variable node classification}), then it is guaranteed than an FPG can be obtained.
+If this step terminates without identifying a hole in an expression block, it is guaranteed that an FPG can be obtained because $F_1$ now satisfies all conditions from Sec.~\ref{ss:FPG} except for (3), which requires that no holes or collisions be present in the graph.
+Since the graph is free of holes, and collisions can always be resolved without producing a hole, (see Table \ref{t:variable node classification}), then it is guaranteed than an FPG can be obtained.
 
@@
     \begin{equation}
-    S_\txt{nodes} = \{v \in V_{M,\txt{var}} \st \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
+    S_\txt{nodes} = \{v \in V_{M,\txt{var}} \st \txt{deg}^-(v) > \txt{deg}_u^-(v) \},
     \end{equation}
+where the degree is now calculated with respect to $F_1$.
     For each collision node we can construct a set containing the edges directed into the node. The set containing all of these sets is constructed as
@@
     An example collision is shown in Fig.~\ref{f:collision} to exemplify the definition of $S_{\txt{edges},j}$. 
-    We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$ (which are sets). 
+    $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$ (which are sets). 
     We may then construct sets of edges as
-    \begin{equation}
-    B_j = \big \{e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \txt{ with } \txt{deg}_u^-(v_j) \leq K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
-    \end{equation}
+\begin{IEEEeqnarray*}{rCl}
+    B_j =  &\big\{& e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \\
+& & \txt{ with } \txt{deg}_u^-(v_j) \leq K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
+\end{IEEEeqnarray*}
+
         which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper and lower indegree limits of $v_j$.  
 The construction of each $B_j$ corresponds to making a decision about which edges to include and which edges not to include. 
-Let the new set of connection edges be
+Let the new set of connection edges be denoted
     \begin{equation}
@@
     \begin{equation}
-        \forall i \in I_E,\forall v \in V_{E_i}, \txt{ if } \txt{deg}^+(v)=0 \txt{ then } (b,v) \in E_R
+        \forall i \in I_E,\forall v \in V_{E_i}, \txt{ if } \txt{deg}^-(v)=0 \txt{ then } (b,v) \in E_R.
     \end{equation}
@@
 \end{equation}
-        If desired, the indices of the analysis blocks implemented in the FPG can then be found as
-    \begin{equation}
-        I_F = \{ i \in I_A \st G_{A_i} \subset F \}.
-    \end{equation}
-        Also, the set of connection edges can be extracted by considering only edges whos endpoints are not in the same analysis block:
-        \begin{equation}
-        E_{F,\txt{con}} = \{(x,y) \in E_{F} \st \sim(x \in V_{A_i} \txt{ and } y \in V_{A_i} \txt{ for some } i \in I_F)\}.
-        \end{equation}
+        If desired, the set of connection edges can be extracted by considering only edges whose endpoints are not in the same analysis block:
+\begin{IEEEeqnarray*}{rCl}
+E_{F,\txt{con}} = &\{&(x,y) \in E_{F} \st \sim(x \in V_{A_i} \txt{ and } y \in V_{A_i}  \\
+& & \txt{for some } i \in I_F)\}.
+\end{IEEEeqnarray*}
 \end{description}
 
-This algorithm will always provide an FPG if one exists. If an FPG does not exist, the limiting factors preventing a valid formulation are revealed.
+This algorithm will always provide an FPG if one exists. If an FPG does not exist, the implementation reveals the limiting factors that prevent a valid FPG from being achieved.
 
-\subsection{Suggested Process for Using the FPG Algorithm}
+\subsection{Suggested Process for Appyling the FPG Algorithm}
 \label{ss:process}
 Section \ref{ss:obtaining FPG} provided an algorithm for obtaining an FPG from the given MCG. 
-However, if an FPG cannot be obtained, this algorithm can be applied as part of a larger process in which the designer changes the problem or the supplied analyses to attempt to obtain an FPG.
+The algorithm can be applied as part of a broader process in which the designer changes the problem formulation or the supplied analysis tools to attempt to obtain an FPG.
 The following steps detail the suggested procedure for obtaining an FPG:
 \begin{enumerate}
-\item[\bf{(A)}] Begin with a set of inputs, analysis codes, objectives, and constraints.
+\item[\bf{(A)}] Begin with a set of inputs, analysis tools, objectives, and constraints.
 \item[\bf{(B)}] Build the MCG as described in Sec.~\ref{ss:MCG}.
-\item[\bf{(C)}] Set indegree limits for variable nodes representing local inputs as described in Sec.~\ref{s:indegree-outdegree} and Table \ref{t:variable node classification}
+\item[\bf{(C)}] Set indegree limits for variable nodes representing local inputs as described in Sec.~\ref{s:indegree-outdegree} and Table \ref{t:variable node classification}.
 \item[\bf{(D)}] Run the FPG algorithm described in Sec.~\ref{ss:obtaining FPG}. If a valid FPG is unattainable:
 		\begin{enumerate}
-\item Change the MCG by adding analysis blocks and/or inputs, which means identifying new analyses to include in a potential MDAO workflow.
+\item Change the MCG by adding analysis blocks and/or inputs, which means identifying new analysis tools to include in a potential MDAO workflow.
 \item Change indegree limits (see Table \ref{t:variable node classification}).
@@
 \end{enumerate}
-This process illustrates how the algorithm for obtaining a valid FPG is to be applied.
-However, this process also serves to illustrate how a designer may revist the MCG after an FPG was already produced.
-For example, if a designer wants to change an input to be multi-fidelity, new analysis blocks can be added to the original MCG and indegree limits can be modified accordingly, which corresponds to steps (A)--(C), and then the algorithm can be applied again.
+This process illustrates how the algorithm for obtaining a valid FPG can be applied.
+However, the process also serves to illustrate how a designer may revist the MCG after an FPG has already been produced.
+For example, if a designer desires to change an input to be multi-fidelity, new analysis blocks can be added to the original MCG, the indegree limits can be modified accordingly, and the algorithm can be applied again.

### 2013-08-07 problem_formulation.tex
files for submitted version and reviewer 3's comments

--- 2013-05-08 problem_formulation.tex
@@
 %\documentclass[referee]{svjour3} % insert '[draft]' option to show overfull boxes
-\documentclass[onecolumn]{svjour3} % insert '[draft]' option to show overfull boxes
+\documentclass[twocolumn]{svjour3} % insert '[draft]' option to show overfull boxes
 
- \title{A Graph Theory Based Approach to MDAO Problem Formulation}
+ \title{A Graph Theoretic Approach to Problem Formulation for Multidisciplinary Design Analysis and Optimization}
 
 \author{
-  David J. Pate, %
-  Justin Gray, %
+  David J. Pate,\\ %
+  Justin Gray, and\\%
   Brian J. German 
@@
 
- \institute { Justin Gray 
+ \institute { David J. Pate 
+	\at Graduate Research Assistant, Georgia Institute of Technology, 270 Ferst Drive, Atlanta, GA, 30332, U.S.A.
+	\and Justin Gray 
 	\at NASA Glenn Research Center, Mail Stop 5-11, 21000 
 		Brookpark Rd Clevland OH 44135 %
-	\and David J. Pate 
-	\at Graduate Research Assistant, Georgia Institute of Technology, 270 Ferst Drive, Atlanta, GA, 30332, U.S.A.
 	\and Brian German 
@@
 \usepackage{lscape}
-\usepackage{hyperref}
 \usepackage{appendix}
-\usepackage[section]{placeins}
 \usepackage[retainorgcmds]{IEEEtrantools}
@@
 \usepackage{booktabs}
+\usepackage{multicol}
 
@@
 
-\section*{Nomenclature}
-
-\begin{tabular}{l l} 
-    AAO      & All-At- \\
-    MDAO     & Multidisciplinary Design Analysis and Optimization \\
-    FPF      & Fundamental Problem Formulation \\
-\end{tabular}
+%\section*{Nomenclature}
+%
+%\begin{tabular}{l l} 
+%    DSM	   & Design structure matrix \\
+%    FDT      & Functional dependency table \\
+%    FPG      & Fundamental problem graph \\
+%    FPF      & Fundamental problem formulation \\
+%    MCG      & Maximal connectivity graph \\
+%    MDAO     & Multidisciplinary Design Analysis and \\ & Optimization \\
+%    MDO      & Multidisciplinary Design Optimization \\
+%    REMS     & Reconfigurable multidisciplinary synthesis
+%\end{tabular}
 

### 2013-08-08 reviewer_response.tex
edited for first round of reviews

\documentclass{aiaa-tc}

\title{reviewer responses for: \\ A Graph Theoretic Approach to Problem Formulation for 
Multidisciplinary Design Analysis and Optimization}

\usepackage{color}

\newenvironment{rev}{\vspace{2em}\itshape \textcolor{red}}{}

\begin{document}
\maketitle

\section{Reviewer1}

\rev{As the authors point out the design of a major product, such as a civil airliner, 
is complex and the level of complexity increases as design requirements change i.e. the 
need to reduce carbon emissions, the need to reduce maintenance downtime etc. and the 
number of design teams involved in creating a new product increases in number and 
distribution. Their paper is, I think, attempting to address the need to support 
design teams in confronting this situation but have not clearly defined where their work 
fits into the design process \\\\
Broadly speaking there are three major tasks or steps that have to be worked 
through when a globally distributed team is setting up a design system that allows it to 
effectively deploy an MDO methodology to support the design of a product such as a new 
civil airliner: 
\\\\1. The team must decide which design tools are required at each stage of the design 
process as the design proceeds down the time-line this includes the data flow required 
between the various design centers and the design history preservation requirements,
\\\\2. This combination of requirements must be converted into a set of interacting tools 
together with data flow and control programs that constitute a template for a 
Computational Design Engine (CDE),
\\\\3. Finally this template has to be instantiated into a usable CDE software system 
which might take advantage of commercially available software such as that found in Simula etc.
\\\\The paper is clearly focused on step 2 above and it advances a sound argument for the use 
of graph theory to establish an appropriate set of outcomes from such a step and, whilst 
it implies the existence of the other two steps, it does not state this early enough in 
the paper and with sufficient detail to allow a practicing engineer to understand the 
role being played by the authors contribution. I would recommend that the authors clearly 
define the role that their graph theoretic approach plays in this triple set of steps 
leading to the creation of effective CDE.}

The reviewer has correctly identified the major motivation for this work. The goal is indeed to 
assist engineers and teams of engineers in the process of building a model out of a set of interconnected 
tools.

We believe that the distrinction between phases 1 and 2 from the comments are covered adequately 
in the bulk of the introduction and especially in the notional problem described there in.

However, to clarify the disctinction between phases 2 and 3 we hav added the following near 
line 52, page 3 of the original submission:

\emph{%
Our goal is to develop formalism for expressing analysis interconnectivity and for determining feasible
    analysis tool sets to assist an engineering team conducting this task. Because the problem deals with
    interconnectivity, we base our approach on the representations and techniques of graph theory.
    The approach begins by constructing the \emph{maximal connectivity graph (MCG)} describing all possible
    interconnections between the analysis tools proposed by the engineers. Graph operations are then
    conducted to reduce the MCG to a \emph{fundamental problem graph (FPG)} that describes the set of analysis
    tools needed to solve the specified system-level design problem. The concept of the FPG and the identification of feasible FPGs from an MCG are the main contributions of the paper. \\\\
    This information in an FPG represents the engineering design problem that needs to be solved, but it is not 
    itself sufficient to actually run a design optimization. It does not provide infomration 
    about how to solve the problem. Even after the problem is defined an integration platform or framework needs 
    to be selected and appropriate optimization methods identified. This last step essentially 
    involves taking the FPG and turning it into a usable model. We breifly consider methods to further 
    manipulate the FPG into a \emph{problem solution graph (PSG)} which could be usefull in this 
    task, but that is not the focus of this paper. The work is primarily concerned with applying graph 
    theory to the creation of an FPG and the benefits to the design process by doing so.}

\end{document}

### 2013-08-08 intro.tex
edited for first round of reviews

@@
     conducted to reduce the MCG to a \emph{fundamental problem graph (FPG)} that describes the set of analysis
-    tools needed to solve the specified system-level design problem.
-    Any particular solution procedure for the problem, such as the choice of an MDO architecture, e.g. MDF,
-    IDF, CO, BLISS, could then be described by
-    constructing a \emph{problem solution graph (PSG)}.
-    The concept of the FPG and the identification of feasible FPGs from an MCG are the main contributions of the paper.
+    tools needed to solve the specified system-level design problem. The concept of the FPG and the identification of feasible FPGs from an MCG are the main contributions of the paper.
+
+    This information in an FPG represents the engineering design problem that needs to be solved. 
+    However, it does not provide details about how the team should go about 
+    solving it. Even after the problem is defined an integration platform or framework needs 
+    to be selected and appropriate optimization methods identified. This last step essentially 
+    involves taking the FPG and turning it into a usable model. We breifly consider methods to further 
+    manipulate the FPG into a \emph{problem solution graph (PSG)} which could be usefull in this 
+    task, but that is not the focus of this paper.
 

### 2013-09-03 reviewer_response.tex
added comments from other reviews into respons doc

@@
 
-\section{Reviewer1}
+\section{Reviewer 1}
+
+\rev{This idea of PSG was mentioned but was not covered in 
+the paper. For completeness and value proposition sake, I would recommend to 
+add at least half section to show how it will help and look like. You can 
+take the simplest formulation of your choice and show how MCG and FPG will 
+translate and look as a PSG. A brief discussion of advantages vs 
+disadvantages will be useful. }
+
+\rev{ The idea of “holes” and “collision” is explained a few times in the paper. 
+It can be reduced. }
+
+\rev{The review of other graph based techniques is good and highlights their 
+limitations, but I would like to see a discussion of what challenges a 
+practitioner may face while trying to use the approach suggested in this 
+paper. The MCG and FPG will cover lot of information like objectives and 
+constraints, but at what expense. Comparing XDSM with MCG or FPG, one might 
+find XDSM as more easy to read and comprehend. For a similar size problem, 
+FPG looks far more complex and bigger.}
+
+\rev{If PSG is covered even to a small 
+extent (like explained in first comment) I would suggest or like to see the 
+PSG equivalent of problem formulation shown in Figure 2 with XDSM.}
+
+\rev{Is it really adding a lot of value to have graphs like MCG or FPG which 
+are very explicit in nature and can become difficult to read}
+
+\rev{Can the graph be compressed in a logical way to reduce the number of nodes 
+and their interconnections while still conveying important and critical 
+information for that stage (MCG stage or FPG stage) ? Reducing cycles is 
+discussed but the graph may still look complex.}
+
+\rev{Looking at these graphs (for example Fig. 8), a node for example Z2 
+appears several times. It may be useful to figure out holes and collision, 
+but once it is resolved, can those nodes be consolidated to provide simpler 
+visualization ?}
+
+\rev{As the problem size increases, how will it affect the size and complexity 
+of such graphs.}
+
+\rev{
+I will assume that creation of such a graph and interconnections will be 
+expected to be done by some kind of software tool or wizard. A brief 
+discussion about why this MCG, FPG and PSG idea is better to lead to such a 
+system can be useful. Although this itself can be a separate research and 
+paper, but why this idea can provide a better foundation can be valuable.
+}
+
+\rev{Minor comments:
+Section 1 Line 46: I think other commercial tools such as iSight, 
+modeFRONTIER should also be mentioned since their usage is more widespread 
+compared to OpenMDAO.}
+
+\rev{Page 13 line 33 column 2: Typo : “Step (D) now PROCEDEDS … “ Page 16 line 38 
+column 2: Typo : “… formulations involving MULTIPLES analysis ….”}
+
+\section{Reviewer 2}
+
+\rev{Page 1, Line 46: 2nd column...check spacing after ``ModelCenter''.}
+
+\rev{Page 2, Line 38: 2nd column...as your motivation was really done in the 
+Introduction section, consider changing the heading on section 2 to 
+``Background'' or something similar.  You don't really provide motivation in 
+this section anyhow.}
+
+\rev{Page 3, Line 19, 1st column: change ``weather'' to ``whether''}
+
+\rev{Page 3, Line 21, 2nd column: change ``descrition'' to ``description''}
+
+\rev{Page 4, Line 37, 1st column: what does ``REMS'' stand for?}
+
+\rev{Page 4, Line 45, 1st column: change ``REMS address'' to ``REMS addresses''}
+
+\rev{Page 4, Lines 35 to 55, 1st column:  This paragraph is a bit awkward in 
+structure.  In the second to last sentence you state, "REMS provides syntax 
+for variables and functions but does not facilitate inclusion of solvers or 
+optimizers in the graph."  The reader then expects your method to address 
+these shortcomings in the next sentence but instead you state, "...the goal 
+of allowing graphs to describe analysis couplings and more closely interact 
+with existing design processes and MDAO frameworks," which you never really 
+called out as a shortcoming of REMS in the first place.  I think you could 
+improve the flow/wording here.}
+
+\rev{Line 10, 1st column, Page 10:  This equation (missing a number) has a LOT of 
+unions, variables, etc...in it.  It would help tremendously to provide a 
+verbal description of at least one of these equations to help with 
+translation into the real world.  This would be similar to what you did 
+following the steps (1) to (4) in the next subsection.}
+
+\rev{Lines 5 - 12, 2nd column, Page 12:  Just a general question spurred by the 
+text here.  What criteria would be applied to determine which FPG is better 
+than another?}
+
+\rev{Line 57, 1st column, Page 13: change ``as a input'' to ``as an input''}
+
+\rev{Now for the request of more significant importance.  This is coming from 
+someone with significant experience in MDO and who uses a lot of the 
+analysis tools you specified in your case study on a daily basis.  I really 
+liked this article but felt it was lacking some punch in the results 
+section.  This was namely in the form of quantitative results.  You vaguely 
+give some in the form of the different resulting connectivity graphs and 
+this was good but to really drive home the value of your research to a 
+potential user I think you could do a bit better with very minimal work. 
+The potential user of this methodology will be concerned with one of two 
+things: 1) minimizing computational time or 2) increasing the accuracy of 
+the results.  I think you've given examples of these two scenarios already 
+in figures 15 and 16, so why not provide a table of the CPU times (1) and/or 
+deviation from higher order analyses (2) for each of these graphs.  Doesn't 
+have to necessarily be these two metrics but something to really tie the 
+output of your methodology to something tangible for comparison purposes. 
+You could even include the metrics for a randomly obtained FPG as a 
+baseline.  Then the person reading the article can say, "Wow, given a set of 
+analysis tools and a general problem, I can employ this method to obtain an 
+FPG that significantly reduces my overall computational time." or "Wow, 
+given a set of analysis tools and a general problem, I can employ this 
+method to obtain an FPG that significantly improves my overall computational 
+accuracy while keeping my computional efficiency as high as possible."}}
+
+\rev{If you are worried about length, I think the conclusions section may be cut 
+in half and still accomplish its intended purpose.}
+
+\section{Reviewer 3}
 
@@
     conducted to reduce the MCG to a \emph{fundamental problem graph (FPG)} that describes the set of analysis
-    tools needed to solve the specified system-level design problem. The concept of the FPG and the identification of feasible FPGs from an MCG are the main contributions of the paper. \\\\
+    tools needed to solve the specified system-level design problem. The concept of the FPG and 
+    the identification of feasible FPGs from an MCG are the main contributions of the paper. \\\\
     This information in an FPG represents the engineering design problem that needs to be solved, but it is not 

### 2013-09-03 reviewer_response.tex
extra curly fixed

@@
 method to obtain an FPG that significantly improves my overall computational 
-accuracy while keeping my computional efficiency as high as possible."}}
+accuracy while keeping my computional efficiency as high as possible."}
 

### 2013-09-03 intro.tex
extra curly fixed

@@
 
-    This information in an FPG represents the engineering design problem that needs to be solved. 
-    However, it does not provide details about how the team should go about 
-    solving it. Even after the problem is defined an integration platform or framework needs 
+    This information in an FPG represents the engineering design problem that needs to be solved, but it is not 
+    itself sufficient to actually run a design optimization. It does not provide infomration 
+    about how to solve the problem. Even after the problem is defined an integration platform or framework needs 
     to be selected and appropriate optimization methods identified. This last step essentially 
@@
     manipulate the FPG into a \emph{problem solution graph (PSG)} which could be usefull in this 
-    task, but that is not the focus of this paper.
+    task, but that is not the focus of this paper. The work is primarily concerned with applying graph 
+    theory to the creation of an FPG and the benefits to the design process by doing so.
 

### 2013-09-04 reviewer_response.tex
adding in first PSG details

@@
 disadvantages will be useful. }
+
+We agree that this was too quickly addressed in the paper. While we feel that 
+discussion of the PSG in any significant depth is too big of a topic to fit into 
+this paper, it does warrant a more complete description than was given in the 
+original draft.
+
+We have added sections 3.5, 4.4 in the appropriate 
+parts of the paper to address the the driver nodes and driven edges which 
+together distinguish a PSG from and FPG.  We have also added added section 5.5
+which gives an example of a simple PSG and compares a PSG with its close relative
+and XDSM diagram.
 

### 2013-09-04 graph_syntax.tex
adding in first PSG details

@@
   \label{s:syntax definition}
-  In this section we present our graph-based syntax, beginning with the necessary graph theoretic fundamentals to 
-  construct the graphs discussed in Sec.~\ref{s:requirements}. 
+  In this section we present our graph-based syntax, beginning with the necessary graph 
+  theoretic fundamentals to construct the graphs discussed in Sec.~\ref{s:requirements}. 
   The notation used in this paper is adapted from Diestel \cite{Diestel2010}. 
@@
 We categorize nodes and edges 
-  into distinct types in order to provide an intuitive approach to MDAO problem formulation. The three node types are
+  into distinct types in order to provide an intuitive approach to MDAO problem formulation. 
+  The three node types are
     \begin{description}
       \item[\bf{variable node:}] represents scalar or array data, inputs and outputs,
-      \item[\bf{function node:}] represents the computation peformed by analysis tools,
-      \item[\bf{driver node:}] represents control logic capable of managing iterations (present only in a PSG),
+      \item[\bf{function node:}] represents the computation performed by analysis tools,
+      \item[\bf{driver node:}] represents control logic capable of managing iterations 
+      (present only in a PSG),
     \end{description}
@@
   \item[\bf{function edge:}] represents exchange of information internal to analysis tools,
-  \item[\bf{driven edge:}] represents passing of information from a driver node to a variable node (present only in a PSG).
+  \item[\bf{driven edge:}] represents passing of information from a driver node to a 
+  variable node (present only in a PSG).
   \end{description}
@@
 
-  A rule set is provided for the usage of these node and edge types to provide a structure to the graph-based representation.
-  The driver node and the driven edge are allowed only in PSGs, whereas the other node and edge types can be present in any of the three graph types, 
+  A rule set is provided for the usage of these node and edge types to provide a structure 
+  to the graph-based representation. The driver node and the driven edge are allowed only in 
+  PSGs, whereas the other node and edge types can be present in any of the three graph types, 
   subject to the following restrictions: 
@@
     directed out.
-  \item If a variable node has an outgoing function edge, then it may not have other outgoing edges.
-  \item If a variable node has an incoming function edge, then it may not have any additional incoming edges.
+  \item If a variable node has an outgoing function edge, then it may not have other 
+  outgoing edges.
+  \item If a variable node has an incoming function edge, then it may not have any 
+  additional incoming edges.
   \end{enumerate}
@@
   node type and two additional edge types to the syntax to allow descriptive
-  graphs for all three phases of the design problem formulation process discussed in Sec. \ref{s:requirements}.
+  graphs for all three phases of the design problem formulation process discussed 
+  in Sec. \ref{s:requirements}.
 
@@
   the values for their respective outputs. We represent this process
-  by a digraph called an \emph{analysis block}; a notional analysis block is shown in Fig.~\ref{f:analysis block}. 
+  by a digraph called an \emph{analysis block}; a notional analysis block is shown 
+  in Fig.~\ref{f:analysis block}. 
   %Each analysis block is a directed graph denoted as $ A=(V_{A},E_{A})$
@@
   Fig.~\ref{f:analysis block} satisfies the rules listed in Sec.~\ref{ss:types}. 
-  Conversely, the rules ensure that analysis tools are represented via this analysis block structure.
+  Conversely, the rules ensure that analysis tools are represented via this analysis 
+  block structure.
 
@@
   characteristics of the analysis code as a weight in a weighted graph formulation. 
-  Because all the edges within analysis blocks are fixed, the blocks themselves are fixed sub-graphs within the overall MDAO problem graph. 
-  The connectivity of nodes and edges in an analysis block cannot be altered for the purposes of MDAO problem formulation; 
-  however, analysis block sub-graphs can be added or removed from the MDAO problem graph as needed.
-  Variable nodes in an analysis block can be distinguished as either an input or an output by the manner in which they are connected. 
-  As shown in Fig.~\ref{f:analysis block}, inputs are represented as variable nodes that have an outgoing edge into a function node, and outputs are represented as variable nodes that have an incoming edge from a function node.
+  Because all the edges within analysis blocks are fixed, the blocks themselves are fixed 
+  sub-graphs within the overall MDAO problem graph. 
+  The connectivity of nodes and edges in an analysis block cannot be altered for the 
+  purposes of MDAO problem formulation; 
+  however, analysis block sub-graphs can be added or removed from the MDAO problem 
+  graph as needed. Variable nodes in an analysis block can be distinguished as either an 
+  input or an output by the manner in which they are connected. As shown in 
+  Fig.~\ref{f:analysis block}, inputs are represented as variable nodes that have an 
+  outgoing edge into a function node, and outputs are represented as variable nodes that 
+  have an incoming edge from a function node.
 
@@
   Along with analysis tools, objectives, constraints, and inputs also need to be represented.
-  In the case of objective functions, a single output value generated by an analysis block could be selected, but commonly, multiple output values from different analysis tools are aggregated together to form a composite objective function.
-  Generally, both objective and constraint functions accept a set of inputs and map them to an output value of significance to the overall design problem.
+  In the case of objective functions, a single output value generated by an analysis 
+  block could be selected, but commonly, multiple output values from different analysis 
+  tools are aggregated together to form a composite objective function. 
+  Generally, both objective and constraint functions accept a set of inputs and map them 
+  to an output value of significance to the overall design problem.
 
-  The operations of implementing composite objective and constraint functions, although typically simple, are effectively identical to the task performed by an analysis block. 
-  A composite constraint or objective function can therefore be represented as an analysis block within the graph with its own input and output variable nodes. 
-  Although fundamentally no different than an analysis block, for clarity and convenience, it is useful to distinguish between analysis blocks corresponding to analysis codes and those that arise from the addition of objectives or constraints to the graph.
+  The operations of implementing composite objective and constraint functions, 
+  although typically simple, are effectively identical to the task performed by an analysis block. 
+  A composite constraint or objective function can therefore be represented as an 
+  analysis block within the graph with its own input and output variable nodes. 
+  Although fundamentally no different than an analysis block, for clarity and 
+  convenience, it is useful to distinguish between analysis blocks corresponding to 
+  analysis codes and those that arise from the addition of objectives or constraints to the graph. 
   Therefore, we use an \emph{expression block} to represent an objective or constraint. 
-  These expression block graphs follow the same structure as analysis blocks presented in Sec.~\ref{ss:analysis blocks}.
+  These expression block graphs follow the same structure as analysis blocks 
+  presented in Sec.~\ref{ss:analysis blocks}.
 
-  Finally, inputs to the problem formulation are represented in the graph by individual variable nodes with connection edges directed out to corresponding local inputs of analysis blocks, and they are referred to simply as \emph{inputs}.
-  These nodes serve to indicate that the variable nodes within analysis or expression blocks to which they are connected represent the same variable in the problem formulation.
+  Finally, inputs to the problem formulation are represented in the graph by individual 
+  variable nodes with connection edges directed out to corresponding local inputs of 
+  analysis blocks, and they are referred to simply as \emph{inputs}. 
+  These nodes serve to indicate that the variable nodes within analysis or 
+  expression blocks to which they are connected represent the same variable in the 
+  problem formulation.
 
-  Figure \ref{f:obj-cons} demonstrates the use of inputs (black circles), analysis blocks (dotted boxes), and expression blocks (dashed boxes), for the Sellar problem graph.
+  Figure \ref{f:obj-cons} demonstrates the use of inputs (black circles), analysis 
+  blocks (dotted boxes), and expression blocks (dashed boxes), for the Sellar problem 
+  graph.
   \begin{figure}[htb!]
@@
     \end{center}
-    \caption{Sellar problem graph indicating inputs (black circles), analysis blocks (dotted boxes), and expression blocks (dashed boxes).}
+    \caption{Sellar problem graph indicating inputs (black circles), analysis blocks 
+    (dotted boxes), and expression blocks (dashed boxes).}
   \label{f:obj-cons}
@@
   \label{s:indegree-outdegree}
-	To address the MDAO notion of design variables, we first introduce the concept of the degree of a node.
-  In a digraph, \emph{indegree} of a node is the number of edges directed in and 
-  is denoted as $\txt{deg}^-(v)$, and the \emph{outdegree} 
+	To address the MDAO notion of design variables, we first introduce the concept of the 
+  degree of a node. In a digraph, \emph{indegree} of a node is the number of edges 
+  directed in and is denoted as $\txt{deg}^-(v)$, and the \emph{outdegree} 
   is the number of edges directed out and it is denoted as $\txt{deg}^+(v)$ \cite{Diestel2010}.
@@
   These user-specified limits govern the number of connection edges that may 
-  be directed into a node for a valid graph; of course, the specification must satisfy $\txt{deg}_l^-(v) \leq \txt{deg}_u^-(v)$.
-  For example, consider a variable node $v$ with $\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. In this case, $v$ 
-  must have exactly one incoming connection edge or the graph is deemed an invalid problem formulation.
+  be directed into a node for a valid graph; of course, the specification must 
+  satisfy $\txt{deg}_l^-(v) \leq \txt{deg}_u^-(v)$.
+  For example, consider a variable node $v$ with $\txt{deg}_u^-(v) = \txt{deg}_l^-(v) = 1$. 
+  In this case, $v$ must have exactly one incoming connection edge or the graph is 
+  deemed an invalid problem formulation.
+
+  \subsection{Driver Nodes and Driven Edges}
+  Within this syntax, all iterative processes are represented with Driver nodes. 
+  The driver node shares some basic qualities with the model node. It can have incoming 
+  and outgoing function edges to and from variable nodes. For example, many optimizers take 
+  convergence tolerance as input and output iteration counts and final objective value. 
+  With respect to variable nodes and function edges, drivers behave exactly the same as
+  model nodes and are subject to the same rules governing their use in the graph. In this
+  context, driver nodes can be part of driver blocks which behave just like analysis 
+  blocks.
+
+  Driven edges are the distinguishing characteristic associated with driver nodes and
+  driver blocks. Driven edges don't follow the same rules as other edges. Driven edges  
+  will always connect a driver node to a variable node (in either direction), but they 
+  are not considered when counting the indegree or outdegree of the variable node.
+  Hence a single variable node could have multiple incoming driven edges from different 
+  driver blocks. Figure \ref{f:driver block} shows a notional example where a driver block 
+  represents and optimizer solving an unconstrained minimization problem.
+
+  \begin{figure}[htb]
+    \begin{center}
+    \includegraphics[width=3.0in]{images/driver_block}
+    \end{center}
+    \vspace{-10pt}
+  \caption{Example driver block with node types labeled.}
+  \label{f:driver block}
+  \end{figure}
 
@@
   \label{s:graph representation}
-  In Sec.~\ref{s:syntax definition}, we presented the structure of the graph-based representation.
-We now discuss how these formalisms provide the remaining MDAO problem constructs identified in Sec.~\ref{s:requirements}.
+  In Sec.~\ref{s:syntax definition}, we presented the structure of the graph-based 
+  representation. We now discuss how these formalisms provide the remaining MDAO problem 
+  constructs identified in Sec.~\ref{s:requirements}.
 
@@
   changed by the designer or by an optimizer in an MDAO study. In the graph-based 
-  syntax, a node corresponds to a design variable if the indegree limits of the node are prescribed as zero.
+  syntax, a node corresponds to a design variable if the indegree limits of the node are 
+  prescribed as zero.
 
   If the indegree limits are violated by the number of incoming connection edges 
-  (the indegree of the node), then the graph is regarded as an invalid problem formulation. There are two classifications for these violations:
+  (the indegree of the node), then the graph is regarded as an invalid problem formulation. 
+  There are two classifications for these violations:
     \begin{description}
@@
         would not be capable of producing outputs.
-          \item[\bf{collision:}] The number of incoming edges is greater than the upper indegree limit:
+      \item[\bf{collision:}] The number of incoming edges is greater than the upper 
+      indegree limit:
         \begin{equation} 
@@
         ambiguity as to which information is to be used as an input for the analysis tool.
-	In multifidelity problems (discussed subsequently), the upper indegree limit could be specified as $deg_u^- > 2$, which implies that a collision is not noted in the case of multiple inputs.
+	In multifidelity problems (discussed subsequently), the upper indegree limit could be 
+  specified as $deg_u^- > 2$, which implies that a collision is not noted in the case of 
+  multiple inputs.
   \end{description}
@@
   The presence of holes and collisions in a graph represent conflicts that will give
-  rise to an invalid problem formulation. 
-The process of obtaining an FPG from an MCG will reveal these conflicts.
+  rise to an invalid problem formulation. The process of obtaining an FPG from an 
+  MCG will reveal these conflicts.
 
@@
   In MDAO, coupling is the mutual dependence between two or more analysis 
-  tools and their respective outputs. In this graph-based syntax, coupling is described by the presence of a 
-  cycle between two or more analysis blocks. In the Sellar problem from Eq.~\ref{eqn:simple_fpf} 
+  tools and their respective outputs. In this graph-based syntax, coupling is 
+  described by the presence of a cycle between two or more analysis blocks. In the Sellar 
+  problem from Eq.~\ref{eqn:simple_fpf} 
   a mutual dependence between $D_1$ and $D_2$ exists through the variables $y_1$ and $y_2$. 
   The edges belonging to a cycle in the Sellar problem are highlighted in 
-  Fig. \ref{f:sellar cycles}.  These edges collectively form a closed path between analysis blocks $D_1$ and $D_2$.
+  Fig. \ref{f:sellar cycles}.  These edges collectively form a closed path between 
+  analysis blocks $D_1$ and $D_2$.
 
@@
     \end{center}
-    \caption{Sellar problem graph with dashed edges indicating participation in a cycle.
-\label{f:sellar cycles}
-    }
+    \caption{Sellar problem graph with dashed edges indicating %
+    participation in a cycle.\label{f:sellar cycles}}
   \end{figure}
@@
   Solvers, optimizers, and other iterative processes are not invovled in the coupling 
-  definition; however, one or more solvers will be required to build a valid PSG from a given FPG  that includes cycles. 
-  Additionally, a coupling cycle has no inherent start or end point. It would be acceptable to select a
+  definition; however, one or more solvers will be required to build a valid PSG from a 
+  given FPG  that includes cycles. Additionally, a coupling cycle has no inherent 
+  start or end point. It would be acceptable to select a
   node in the cycle as a starting point and proceed around the
@@
   demonstrates how building an FPG from an MCG provides an opportunity to 
-  identify and potentailly reduce the number of cycles in a graph.
+  identify and potentially reduce the number of cycles in a graph.
 
@@
   with a new analysis, $D_0$, representing a low fidelity version of $D_1$. 
-  In this modified version, the variable nodes with an upper indegree limit of two are highlighted in gray, and changing the upper indegree limits for the variable nodes representing $y_1$ in $D_2$, $F$, and $G_1$ avoids a collision.
-In Sec.~\ref{ss:obtaining FPG} we present an algorithm for locating 
-  conflicts within a given MCG such that a designer can make the necessary decisisons about each 
-  one in turn. 
+  In this modified version, the variable nodes with an upper indegree limit of two 
+  are highlighted in gray, and changing the upper indegree limits for the variable nodes 
+  representing $y_1$ in $D_2$, $F$, and $G_1$ avoids a collision. In Sec.~\ref{ss:obtaining FPG} 
+  we present an algorithm for locating conflicts within a given MCG such that a designer 
+  can make the necessary decisions about each one in turn. 
   \begin{figure}
@@
       \includegraphics[width=3.25in]{images/sellar_mulfi}
-    \caption{Graph of the modified Sellar problem with multi-fidelity nodes highlighed in gray.\label{f:collision_example}}
+    \caption{Graph of the modified Sellar problem with multi-fidelity nodes highlighted in gray.\label{f:collision_example}}
   \end{center}
@@
 
-  Multi-fidelity problems are always charaterized by graphs in which at least some of the variable 
+  Multi-fidelity problems are always characterized by graphs in which at least some of the variable 
   nodes that have an $\txt{deg}^-_u(v) > 1$. These problems require 
-  special techniques for resolving the confliciting edges by introducing some mechanism
+  special techniques for resolving the conflicting edges by introducing some mechanism
   to manage when each of the different fidelity analyses are 
@@
 
-Table \ref{t:variable node classification} summarizes the classification of a variable node as a hole, design variable, single valid input (nominal), collision, or multi-fidelity node.
+Table \ref{t:variable node classification} summarizes the classification of a variable 
+node as a hole, design variable, single valid input (nominal), collision, or 
+multi-fidelity node.
 
@@
 \end{table}%
+
+\subsection{Iterative Control: Drivers}
+  There are many iterative processes that are used to help solve MDAO problems such as 
+  solvers, optimizers, and Design of Experiment (DOE). Abstractly, even a human picking 
+  values guided by experience and trend could be viewed as an iterative process. 
+  These processes represent the control structures that govern how a design problem 
+  marches toward a solution but they do not fundamentally impact the definition of the 
+  problem to be solved. e.g. which optimizer you choose does not change the problem you 
+  are solving but will change the path you follow to solve it. Hence, as a rule, driver 
+  nodes and driven edges are excluded from both the MCG and FPG which deal exclusively 
+  with problem formulation. They only come into play with in the PSG where the issue of 
+  how to solve a specific problem is addressed. In fact, at least one driver block is 
+  required to be present in any PSG in order to select values for the design variables.

### 2013-09-04 mcg_fpg.tex
adding in first PSG details

@@
-\section{The MCG and FPG}
+\section{The MCG, FPG, and PSG}
     \label{s:building graphs}
@@
 For example, if a designer desires to change an input to be multi-fidelity, new analysis blocks can be added to the original MCG, the indegree limits can be modified accordingly, and the algorithm can be applied again.
+
+\subsection{Problem Solution Graph}
+Once a suitable FPG has been obtained the issue of how to structure a solution strategy 
+can be addressed. Similar to the way an MCG can be manipulated into a FPG, an FPG can be
+further manipulated into a PSG. To get from an MCG to an FPG you eliminated analysis blocks and
+connection edges to reduce the problem to something solvable, but to get from an FPG to a PSG you 
+add in driver blocks, driven edges to describe the structure of the solution strategy being applied 
+to the problem.

### 2013-09-04 reviewer_response.tex
more revisions

@@
 
+It is our conclusion that the PSG represents a sub-set of the information in the XDSM
+diagram, but that for data connectivity part of the graphs XDSM and PSG represent 
+equivalent graphs. It would be interesting future work to consider incorporating a new
+process edge type into the PSG to make it functionally equivalent to a complete XDSM, 
+however that would be a significant amount of extra work and worth of it's own paper. 
+To do that topic justice, the original XDSM work would need to be converted for every 
+case considered into the proposed syntax to ensure that they were completely equivalent.
+
 \rev{ The idea of “holes” and “collision” is explained a few times in the paper. 
 It can be reduced. }
+
+Searching the paper, the concepts is discussed often, but only formally defined once 
+in section 4.1. In 5.3 a mathematical description of the set of nodes that meet
+the criteria for holes and edges is given, but this is different than the definition.
+
+We have checked both sections and removed the gross reducency, but feel that keeping 
+both equations makes for a clearer message in the paper. The concept is a bit subtle 
+since a node can be either a design variable or a hole, but not both. Similarlly, 
+a collision can occur when a variable has two incomming edges or not if the problem 
+is multi-fidelity. So a little extra explanation helps to keep this clear.
 
@@
 FPG looks far more complex and bigger.}
+
+We agree that XDSM is more visually informative and consise. However this was not the 
+goal of the new syntax. Instead, the goal was to make a graph structure that 
+was easier to operate on algorithmically. As a consequence the graph grows larger and 
+less human readable. This trade is acceptable for two reasons:
+
+\begin{enumerate}
+    \item It should be possible to convert between the XDSM format and the proposed syntax 
+    in a lossless manner. This is true for the gray lines of XDSM but not the black ones.
+    Only data connections are shown in current syntax. 
+    \item Algorithmic methods are a useful way of analyzing problem structure and it makes 
+    sense to have a graph structure that works well for this use, even at the expense 
+    of human readability. 
+\end{enumerate}
+
+See the revised end of the introduction (section 1) where this is more clearly stated.
 

### 2013-09-04 graph_syntax.tex
more revisions

@@
   Hence a single variable node could have multiple incoming driven edges from different 
-  driver blocks. Figure \ref{f:driver block} shows a notional example where a driver block 
-  represents and optimizer solving an unconstrained minimization problem.
+  driver blocks, e.g. in a sequential optimization strategy, both drivers would at 
+  different times, pick values for the same variable. Figure \ref{f:driver block} shows 
+  a notional example where a driver block represents and optimizer solving an 
+  unconstrained minimization problem.
 

### 2013-09-04 intro.tex
more revisions

@@
     analyses begin to dominate the effort required in setting up the model. It is this problem of 
-    determining sets of analysis tools and their interconnectivities to form realistic
+    determining sets of analysis tools and their inter-connectivities to form realistic
     multidisciplinary problems that is the subject of this paper. We are motivated by the following notional
@@
     This information in an FPG represents the engineering design problem that needs to be solved, but it is not 
-    itself sufficient to actually run a design optimization. It does not provide infomration 
+    itself sufficient to actually run a design optimization. It does not provide information 
     about how to solve the problem. Even after the problem is defined an integration platform or framework needs 
     to be selected and appropriate optimization methods identified. This last step essentially 
-    involves taking the FPG and turning it into a usable model. We breifly consider methods to further 
-    manipulate the FPG into a \emph{problem solution graph (PSG)} which could be usefull in this 
+    involves taking the FPG and turning it into a usable model. We briefly consider methods to further 
+    manipulate the FPG into a \emph{problem solution graph (PSG)} which could be useful in this 
     task, but that is not the focus of this paper. The work is primarily concerned with applying graph 
@@
     demonstrate how multiple specific problem formulations can relate to a common FPF and indicate the 
-    value of a common problem descrition. The challenge with this 
+    value of a common problem description. The challenge with this 
     traditional syntax is that it is not easily manipulated or analyzed with automatic procedures to explore alternate problem formulations.
@@
     dependency between multiple analysis tools as well as between analysis tools and
-    objective/constraint functions. Additionaly, REMS address the need to
+    objective/constraint functions. Additionally, REMS address the need to
     transition between multiple solution strategies while maintaining a single consistent  
@@
     syntax for variables and functions but does not facilitate inclusion 
-    of solvers or optimizers in the graph.  The 
-     graph syntax proposed in this paper could  
-    be viewed as an extention to REMS with the goal of allowing the graphs to  describe analysis couplings and more 
+    of solvers or optimizers in the graph.  The graph syntax proposed in this paper could  
+    be viewed as an extension to REMS with the goal of allowing the graphs to  describe analysis couplings and more 
     closely interact with existing design processes and MDAO frameworks.
@@
     The XDSM for Eq. \ref{eqn:simple} is given in Figure \ref{fig:XDSM_simple}. 
-    This diagram is shown with an 
-    assumed Gauss-Seidel iteration scheme and an MDF solution architecture.
+    This diagram is shown with an assumed Gauss-Seidel iteration scheme and an MDF solution architecture.
 
-In this paper, we propose a new graph syntax that 
-    combines certain features of REMS and XDSM. 
-The approach allows interactions and couplings of analysis tools and objective and constraint functions to be described while retaining the flexiblity to 
-    represent a fundamental problem formulation free of solution specific information.
+    In this paper, we propose a new graph syntax that combines certain features of REMS and XDSM. 
+    Despite sharing some features with both of these other graph based approaches, the syntax
+    proposed here has a fundamentally different goal which is complimentary to both of them. 
+    REMS and XDSM provide excellent human readable formats for an MDAO problem description. This 
+    syntax is designed to provide a more effective machine readable format for MDAO problem definition. 
+    In the interest of making it easier to algorithmically operate on the graphs, the specified 
+    format is both more verbose and less visually informative. The benefit is a more consistent 
+    and easily utilized graph from a algorithm perspective.
 
@@
   structure of an MDAO problem to be described independently of any solution information, 
-  while still being able to accomodate the more specific case when a solution 
+  while still being able to accommodate the more specific case when a solution 
   strategy is applied. In order to achieve this goal, 
-  the graph syntax needs to accomodate a number of MDAO problem constructs: 
+  the graph syntax needs to accommodate a number of MDAO problem constructs: 
   \begin{itemize}

### 2013-09-04 mcg_fpg.tex
more revisions

@@
     \item[\bf{Step 1: Holes}] 
-        The first step is to detect holes and disconnect the first set of connection edges downstream of them, as indicated in Fig.~\ref{f:hole}. 
-        These connection edges are removed because they represent output variables which cannot be determined because the analysis function does not have adequate inputs.
+    The first step is to detect holes and disconnect the first set of connection edges 
+    downstream of them, as indicated in Fig.~\ref{f:hole}.
 
@@
     \end{equation}
-        which is the set of variable nodes with fewer incoming edges than are allowed by the lower indegree limit.
-        An updated set of edges is then created by removing the connection edges preceding or succeeding the analysis block:
+    which is the set of variable nodes with fewer incoming edges than are allowed by 
+    the lower indegree limit. An updated set of edges is then created by removing the 
+    connection edges preceding or succeeding the analysis block:
     \begin{IEEEeqnarray*}{rCl}
@@
 
-Because removing these edges can create new holes, this step must be repeated until no additional holes are found.
-        If the hole identification step identifies a variable node in an expression block as hole, meaning $V_{E_i} \cap H \neq \emptyset$ for some $i \in I_E$, then the algorithm terminates because an FPG cannot be obtained.
+    Because removing these edges can create new holes, this step must be repeated 
+    until no additional holes are found. If the hole identification step identifies a 
+    variable node in an expression block as hole, meaning 
+    $V_{E_i} \cap H \neq \emptyset$ for some $i \in I_E$, then the algorithm terminates 
+    because an FPG cannot be obtained.
 
-If this step terminates without identifying a hole in an expression block, it is guaranteed that an FPG can be obtained because $F_1$ now satisfies all conditions from Sec.~\ref{ss:FPG} except for (3), which requires that no holes or collisions be present in the graph.
-Since the graph is free of holes, and collisions can always be resolved without producing a hole, (see Table \ref{t:variable node classification}), then it is guaranteed than an FPG can be obtained.
+    If this step terminates without identifying a hole in an expression block, it is 
+    guaranteed that an FPG can be obtained because $F_1$ now satisfies all conditions 
+    from Sec.~\ref{ss:FPG} except for (3), which requires that no holes or collisions 
+    be present in the graph. Since the graph is free of holes, and collisions can 
+    always be resolved without producing a hole, (see Table 
+    \ref{t:variable node classification}), then it is guaranteed than an FPG can be 
+    obtained.
 
@@
 add in driver blocks, driven edges to describe the structure of the solution strategy being applied 
-to the problem.
+to the problem. For example, Figure \ref{f:selllar_psg} shows the PSG that solves the Sellar problem 
+with an MDF solution strategy. There are two driver blocks in this PSG, one for the solver that 
+ensures compatibility and another for the optimizer that picks the design variables.
+
+\begin{figure}[htb]
+    \begin{center}
+    \includegraphics[width=3.0in]{images/sellar_psg}
+    \end{center}
+    \vspace{-10pt}
+    \caption{Example PSG for solving the Sellar problem with MDF strategy}
+    \label{f:sellar_psg}
+\end{figure}
+
+The PSG in figure \ref{f:selllar_psg} can be compared to the XDSM for the same situation 
+in Fig. \ref{fig:XDSM_simple}. While the two diagrams contain much of the same information 
+there is a significant difference. The PSG describes the data passing that happens between 
+the drivers and the rest of the problem, but it does not describe when that data passing 
+will occur (the execution order of the blocks). In the proposed graph syntax edges describe 
+the passing of data between one node an another, but they do not define when that 
+communication happens.
+
+XDSM diagrams do contain information about data passing as well as when that data should be passed. 
+XDSM essentially represents two related, but independent graphs simultaneously. 
+This is accomplished by including two distinct sets of edges, one for data 
+connections and a second indicating execution sequence. As seen in Fig. \ref{fig:XDSM_simple}
+data connections are bold-gray edges and execution connections are numbered-black edges. The
+execution edges, the back edges, in XDSM represent and additional graph that
+follows a completely different set of rules and behaviors from the data connections graph. 
+On the other hand the data described by the gray edges, 
+the data passing information, is equivalent to the information in the PSG. XDSM presents this 
+information in a more compact and human readable manner by aggregating groups of related variables
+into combined nodes. The more verbose form presented in a PSG is less human readable, but more easily 
+machine readable as a consequence of a more consistent structure.

### 2013-09-04 reviewer_response.tex
finished reviewer 1 comments

@@
 
+See new Fig. 8 and Fig. 13
+
 \rev{Is it really adding a lot of value to have graphs like MCG or FPG which 
 are very explicit in nature and can become difficult to read}
+
+Their value is not derived from their ability to represent the data in a human readable 
+form, but in it's power as a machine readable one. Section 5.3 introduces an 
+algorithm for testing if graphs are in fact reduced down to fundamental problems and 
+to reduce them to one if they are not. As problems grow larger and more complex, 
+this kind of algorithm becomes ever more valuable.
+
+It could be implemented into a framework or stand alone tool to help teams develop the
+specifics of a design problem before they started working to code the modules and such. 
+Section 6 shows that for even simple problems, the graph (as viewed by an algorithm) is 
+non-trivial. It also shows that there are multiple ways to solve problems. As 
+problems grow even a small amount from the size given in the example it's clear that 
+algorithmic analysis useful. The graph syntax derives its usefulness from 
+the effectiveness of describing the graph to a computer.
 
@@
 information for that stage (MCG stage or FPG stage) ? Reducing cycles is 
-discussed but the graph may still look complex.}
+discussed but the graph may still look complex.
 
-\rev{Looking at these graphs (for example Fig. 8), a node for example Z2 
+Looking at these graphs (for example Fig. 8), a node for example Z2 
 appears several times. It may be useful to figure out holes and collision, 
@@
 
+There are a number of ways the graph could be simplified down to make it more 
+human readable, but this was not the primary goal of the work. You could combine 
+variables or remove them all together and just represent component to component 
+connections like a pure DSM.
+
+It should be possible to convert the entire PSG graph into the that used by the 
+data-connectivity part of an XDSM. You would have all the information in a more 
+compact and visually insightful form. You could always convert that back into the 
+more verbose syntax here as well. We would argue that the XDSM is more human readable, 
+but this syntax is a closer representation of how a framework would view the problem.
+
+Given our goals, the verbose nature of the graphs is a necessary side effect.
+
 \rev{As the problem size increases, how will it affect the size and complexity 
 of such graphs.}
+
+The answer to this question is not trivial, as it depends on a number of factors 
+such as the number of analyses and the number of variables in those analyses. 
+However, speaking roughly, you could measure the complexity of a graph by the total 
+number of nodes and edges.
+
+So the number of nodes will grow roughly linearly with the number of analysis tools 
+and the number of variables.
+The number of fixed edges will grow linearly with the number of variables as well.
+The number of connection edges depends entirely on the specifics of the problem, but since any variable can only have a finite number of incoming edges (usually 1, but maybe 2 or 3 in multi-fidelity case) you can avoid an unbounded
+growth there. So very roughly speaking, the graph will grow O(N+M) where N is the number 
+of variables and M is the number of analyses.
+
+If you assume large problems have hundreds of components and millions of variables, 
+you're talking about graphs with a few million nodes and edges. These are very large, but 
+not outside the relm of what is manageable by todays graph libraries.
 
@@
 
+See the new section 2.4 for a short section describing our envisioned usage 
+of the graph language.
+
+[TODO: Brian and David can flush this out better than me...]
+
 \rev{Minor comments:
@@
 
-\rev{Page 13 line 33 column 2: Typo : “Step (D) now PROCEDEDS … “ Page 16 line 38 
+Agreed. Done and incorporated into Section 1.
+
+\rev{Page 13 line 33 column 2: Typo : “Step (D) now PROCEDEDS … “}
+
+fixed.
+
+\rev{Page 16 line 38 
 column 2: Typo : “… formulations involving MULTIPLES analysis ….”}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 method to obtain an FPG that significantly improves my overall computational 
-accuracy while keeping my computional efficiency as high as possible."}
+accuracy while keeping my computational efficiency as high as possible."}
 

### 2013-09-04 conclusions.tex
finished reviewer 1 comments

@@
 To create a valid problem formulation, all collisions must be resolved either by removing redundant analysis tools or by introducing logic or solvers into the data flow to choose between or aggregate multiple instances during execution.
-The identification of collisions in an MCG is especially important for problem formulations involving multiples analysis tools of different fidelities; in this context, a collision represents an opportunity to formulate a multi-fidelity problem that can take best advantage of the efficiency/accuracy tradeoff between different analysis tools.
+The identification of collisions in an MCG is especially important for problem formulations involving multiple analysis tools of different fidelities; in this context, a collision represents an opportunity to formulate a multi-fidelity problem that can take best advantage of the efficiency/accuracy tradeoff between different analysis tools.
 Resolving collisions and filling holes in an MCG requires choices by the user to specify the overall MDAO problem that he/she desires to solve, i.e.

### 2013-09-04 example_problem.tex
finished reviewer 1 comments

@@
 Step (B) is the production of the MCG.
-	In this example, the MCG is formed by using a consistant variable naming convention and then connecting all variables with the same name with connection edges. 
+	In this example, the MCG is formed by using a consistent variable naming convention and then connecting all variables with the same name with connection edges. 
 	Table \ref{t:ins and outs} presents the full list of variables in the leftmost column and indicates whether the variable is an input or an output of each analysis tool. 
@@
 
-Step (D) now procededs by executing the FPG algorithm. 
+Step (D) now proceeds by executing the FPG algorithm. 
 An FPG cannot be obtained because the variable nodes representing the number of engines are holes for NPSS, FLOPSa, and FLOPSb, and the variable nodes representing the mission definition are holes for FLOPSa and FLOPSb.
@@
 	One method to implement an FPG in this paradigm is to use a ranking system.
-Each analysis block is assigned a value by the designer describing how desireable it is for this block to be present in the FPG.
+Each analysis block is assigned a value by the designer describing how desirable it is for this block to be present in the FPG.
 The connection edges directed out of each analysis block are then assigned a weight equal to the value assigned to the analysis block.
@@
  multiple analysis codes calculate the same variable. 
-	This multi-fidelity formulation is implemented by setting the upper indegree limit for this node as two and then repreating the (automated) process. 
+	This multi-fidelity formulation is implemented by setting the upper indegree limit for this node as two and then repeating the (automated) process. 
 	In this case, both VORLAX and PMARC are retained, resulting in an FPG that omits only FLOPSa.

### 2013-09-04 intro.tex
finished reviewer 1 comments

@@
     Multidisciplinary Design Analysis and Optimization (MDAO)
-    frameworks such as OpenMDAO \cite{Gray2012} and {ModelCenter}\textregistered have enabled a new level of analysis tool integration 
+    frameworks such as {ModelCenter}\textregistered, iSigt\textregistered, ModeFrontier\textregistered, and OpenMDAO \cite{Gray2012}and have enabled a new level of analysis tool integration 
     and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. This 
@@
   \end{figure}
+
+\subsection{Potential Applications}
+
+The syntax proposed here for defining graphs is partially motivated by the unique needs
+of working with very large problems that have thousands of design variables. Such problems are 
+large enough that many traditional design methods stumble simply because the problems 
+are so large that one person can not maintain a complete view of the problem. Hence as problems 
+grow to that size, software tools become needed to manage the complexity of the problem itself.
+
+Even for small problems, algorithmic exploration of problem structure can yield deeper understanding 
+of problems and more effective solution methods, as demonstrated by the work with the FDT. The syntax 
+described here along with the rules for applying that syntax to the three phases of the design 
+process provides a strong foundation to begin developing new algorithms for analyzing and modifying 
+problem structure. This foundation makes it much easier to incorporate a standard view of 
+problem definition into MDAO frameworks and by extension to implement any new algorithms inside 
+those frameworks.

### 2013-09-04 mcg_fpg.tex
finished reviewer 1 comments

@@
 add in driver blocks, driven edges to describe the structure of the solution strategy being applied 
-to the problem. For example, Figure \ref{f:selllar_psg} shows the PSG that solves the Sellar problem 
+to the problem. For example, Figure \ref{f:sellar_psg} shows the PSG that solves the Sellar problem 
 with an MDF solution strategy. There are two driver blocks in this PSG, one for the solver that 
@@
     \begin{center}
-    \includegraphics[width=3.0in]{images/sellar_psg}
+    \includegraphics[width=3.0in]{images/sellar_mdf_psg}
     \end{center}
@@
 
-The PSG in figure \ref{f:selllar_psg} can be compared to the XDSM for the same situation 
+The PSG in figure \ref{f:sellar_psg} can be compared to the XDSM for the same situation 
 in Fig. \ref{fig:XDSM_simple}. While the two diagrams contain much of the same information 

### 2013-09-05 reviewer_response.tex
addresses second reviewers comments

@@
 \rev{Page 1, Line 46: 2nd column...check spacing after ``ModelCenter''.}
+Fixed
 
@@
 this section anyhow.}
+Section title changed
 
 \rev{Page 3, Line 19, 1st column: change ``weather'' to ``whether''}
+fixed
 
 \rev{Page 3, Line 21, 2nd column: change ``descrition'' to ``description''}
+fixed
 
 \rev{Page 4, Line 37, 1st column: what does ``REMS'' stand for?}
+"Reconfigurability in MDO Problem Synthesis". Full name added to the paper.
 
 \rev{Page 4, Line 45, 1st column: change ``REMS address'' to ``REMS addresses''}
+fixed
 
@@
 
+This paragraph was poorly worded. We have re-worded it to attempt 
+to make the issue with REMS more clear and to address more bluntly 
+what it fails to allow.
+
 \rev{Line 10, 1st column, Page 10:  This equation (missing a number) has a LOT of 
@@
 
+equation numbers were added, see new equations 8 and 9. Also, verbal descriptions
+were added to accompany the set notations.
+
 \rev{Lines 5 - 12, 2nd column, Page 12:  Just a general question spurred by the 
@@
 
+We chose not to address the issue of seleting one FPG over another. The answer 
+to this quetion depends heavily on specific situation. If you're short on time, 
+a smaller FPG with less cycles and less variables may be preferable. If you're trying
+to resolve some small issue with nacelle-wing interference then you may choose to go with a
+more high fidelity and computationally expensive FPG.
+
+Our main point was that the MCG provides a good foundation for deriving the FPG, and that 
+it can be revisited later in the design process as things change. You are not stuck modifying the 
+FPG you initally select.
+
 \rev{Line 57, 1st column, Page 13: change ``as a input'' to ``as an input''}
+fixed
 
@@
 
+We agree that this sort of quantitative data would be very valuable. However, we 
+contend that this data needs to be larger in scope than just that which could be obtained 
+by executing the example problem given here. Instead we think that to do that topic 
+justice, you would need to address a wide range of problems with different graph structural 
+characteristics. The challenge is that any quantitative measures given are inherantly 
+dependent on particular details of a problem. So to address this question fully you 
+would need a range of problems to consider. This work, while very valuable, would 
+fill an entire paper in it's own right and it's something we will consider in future 
+work.
+
 \rev{If you are worried about length, I think the conclusions section may be cut 
 in half and still accomplish its intended purpose.}
+
+Althought the length of the paper was not excessive, we agree that the conclusion 
+could be more concise. We have shortened the conclusion and re-woreded it to make 
+it more clear.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2013-09-05 conclusions.tex
addresses second reviewers comments

@@
 In this paper, we have presented an approach for describing MDAO problems with constructs and syntax from graph theory. 
-Our graph description shares similarities to other approaches such as REMS, $\Psi$, FDT, DSM, and XDSM, but it provides new constructs tailored to applications in early phases of MDAO problem formulation before specific problem solution approaches are envisioned.
-In particular, we introduce the concepts of the Maximally Connected Graph (MCG) and the Fundamental Problem Graph (FPG).
+Our graph description shares similarities to other approaches such as REMS, $\Psi$, FDT, 
+DSM, and XDSM, but it provides new constructs tailored to algorithmic analysis and 
+manipulation. The proposed syntax has applications in early phases of MDAO problem 
+formulation, especially for problems with large numbers of variables and analyses. 
+In particular, we introduce the concepts of the Maximally Connected Graph (MCG) and 
+the Fundamental Problem Graph (FPG).
 
-The MCG addresses the question, ``Given a set of analysis tools, what are all of the variable interconnections between them that could be established?''  
-The MCG provides a structured formalism to identify \textit{holes} and \textit{collisions} in interconnections between input and output variables implied by the set of analysis tools. 
-A hole corresponds to a variable that is not calculated in the MCG data flow but is needed by other tools. 
-To create a valid problem formulation, all holes must be ``filled'' either by assigning the corresponding variable as a design variable to be set by the user or an optimizer, or by introducing additional tools to compute the variable.
+The MCG addresses the question, ``Given a set of analysis tools, 
+what are all of the variable interconnections between them that could be established?''  
+The MCG provides a structured formalism to identify \textit{holes} and 
+\textit{collisions} in interconnections between the set of analysis tools. 
+In order to achieve a valid problem formulation, all of the holes and collisions must 
+be resolved. We present an algorithm to perform this resolution which gurantees that you 
+will find an FPG if one exists. In some cases this algorithm still relies on some input from the 
+designers to make decisions. One benefit of a graph theoretic approach is the standard 
+algorithms which can be used to inform the user's decisions, such as cycle detection, 
+minimum spanning trees, and shortest path algorithms.
 
-A collision corresponds to a situation in which two or more analysis tools compute separate instances of the same variable. 
-To create a valid problem formulation, all collisions must be resolved either by removing redundant analysis tools or by introducing logic or solvers into the data flow to choose between or aggregate multiple instances during execution.
-The identification of collisions in an MCG is especially important for problem formulations involving multiple analysis tools of different fidelities; in this context, a collision represents an opportunity to formulate a multi-fidelity problem that can take best advantage of the efficiency/accuracy tradeoff between different analysis tools.
-Resolving collisions and filling holes in an MCG requires choices by the user to specify the overall MDAO problem that he/she desires to solve, i.e.
-to produce the required set of system-level outputs given a specified set of design variables.
-One benefit of a graph theoretic approach is the standard algorithms which can be used to inform the user's decisions, such as cycle detection, minimum spanning trees, and shortest path algorithms.
+The FPG is a graph that describes a data connection structure corresponding 
+to an MDAO problem formulation free of collisions and holes. 
+An FPG is the result of user choices to fill holes and resolve collisions in 
+the initial MCG. Typically, many different FPGs could be attained from a given MCG, 
+depending on the user choices. The number of possible FPGs that can be attained by 
+selecting different design variables or introducing additional analysis tools to 
+fill holes and/or resolve collisions can be viewed as a measure of the freedom 
+available to the user in implementing the available analysis tools to 
+formulate valid MDAO problems. In Sec.~\ref{s:example problem}, 
+we provide an example application of formulating multiple FPGs from an MCG 
+for a commercial aircraft design problem.
 
-The FPG is a graph that describes a data connection structure corresponding to an MDAO problem formulation free of collisions and holes. 
-An FPG is the result of user choices to fill holes and resolve collisions in the initial MCG. 
-Typically, many different FPGs could be attained from a given MCG, depending on the user choices. 
-The number of possible FPGs that can be attained by selecting different design variables or introducing additional analysis tools to fill holes and/or resolve collisions can be viewed as a measure of the freedom available to the user in implementing the available analysis tools to formulate valid MDAO problems.
-In Sec.~\ref{s:building graphs}, we present an algorithm for identifying holes and conflicts in an MCG that can be applied by the user in a process to arrive at valid FPGs. 
-In Sec.~\ref{s:example problem}, we provide an example application of formulating an FPG from an MCG for a commercial aircraft design problem.
-
-Once a desired FPG is identified, the graph syntax described in the paper can be leveraged to formulate a Problem Solution Graph (PSG).
-The PSG describes particular MDAO solution architectures, e.g.
-MDF, IDF, BLISS, ATC, that can be used to solve the problem defined by the FPG.
-Many possible PSGs could be formulated to solve a problem corresponding to a specified FPG.  
-The application of the syntax to PSGs has not been illustrated in the paper.
-
-For simple problems with few analysis tools and variables, the formulation of a valid problem description is straightforward. 
-However, MDAO problems continue to grow in scale and complexity. 
-As the numbers of analysis tools and variables have increased, it has become increasingly challenging and time consuming for engineering teams to determine how the multiple analysis tools can be interconnected to produce valid problem formulations, to know when other tools must be introduced, and to determine the number of free variables in the problem that can or should be varied by the designer or an optimizer.
-The graph formalism presented in this paper is intended to offer value in this context of increasing problem complexity by providing a formal approach to identify and resolve missing and redundant data in order to create valid MDAO problem formulations.
+For simple problems with few analysis tools and variables, the formulation of a 
+valid problem description is straightforward. However, MDAO problems continue to 
+grow in scale and complexity. As the numbers of analysis tools and variables have 
+increased, it has become increasingly challenging and time consuming for engineering 
+teams to determine how the multiple analysis tools can be interconnected to produce 
+valid problem formulations, to know when other tools must be introduced, and to 
+determine the number of free variables in the problem that can or should be 
+varied by the designer or an optimizer. The graph formalism presented in this 
+paper is intended to offer value in this context of increasing problem complexity by 
+providing a formal approach to programatically identify and resolve missing and 
+redundant data in order to create valid MDAO problem formulations.

### 2013-09-05 intro.tex
addresses second reviewers comments

@@
     Multidisciplinary Design Analysis and Optimization (MDAO)
-    frameworks such as {ModelCenter}\textregistered, iSigt\textregistered, ModeFrontier\textregistered, and OpenMDAO \cite{Gray2012}and have enabled a new level of analysis tool integration 
+    frameworks such as iSight\textsuperscript{\textregistered},\\ ModelCenter\textsuperscript{\textregistered}, ModeFrontier\textsuperscript{\textregistered}, and OpenMDAO\cite{Gray2012} and have enabled a new level of analysis tool integration
     and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. This 
@@
 
-\section{Motivation}
+\section{Background}
 \subsection{Specific vs Fundamental Problem Formulation }
@@
     to $D_1$. This implies that $D_1$ and $D_2$ are coupled together by means 
-    of their mutual dependence. However no information is given as to weather
+    of their mutual dependence. However no information is given as to whether
     $D_1$ or $D_2$ should be run before the other. Given Eq.~\ref{eqn:simple_fpf}, 
@@
 
-    Alexandrov and Lewis introduced a graph based syntax called REMS which 
-    incorporates objectives and constraints into a graph, effectively combining 
-    FDT and DSM \cite{alexandrov2004}. REMS retains the square adjacency 
+    Alexandrov and Lewis introduced a graph based syntax called Reconfigurability in 
+    MDO Problem Synthesis (REMS) which incorporates objectives and constraints 
+    into a graph, effectively combining FDT and DSM \cite{alexandrov2004}. REMS retains the square adjacency 
     matrix form from DSM, but by adding the objectives and constraints, it partially 
@@
     dependency between multiple analysis tools as well as between analysis tools and
-    objective/constraint functions. Additionally, REMS address the need to
+    objective/constraint functions. Additionally, REMS addresses the need to
     transition between multiple solution strategies while maintaining a single consistent  
-    graph representation of the fundamental problem formulation. REMS provides
-    syntax for variables and functions but does not facilitate inclusion 
-    of solvers or optimizers in the graph.  The graph syntax proposed in this paper could  
-    be viewed as an extension to REMS with the goal of allowing the graphs to  describe analysis couplings and more 
-    closely interact with existing design processes and MDAO frameworks.
+    graph representation of the fundamental problem formulation. However REMS does not provide 
+    a mechanism for inclusion of solvers or optimizers in the graph. This fundamentally limits REMS 
+    from describing the specifics of a different solution strategies as applied to a specific problem.
 

### 2013-09-05 mcg_fpg.tex
addresses second reviewers comments

@@
 
-    We may then construct the maximal connectivity graph $M=(V_M,E_M)$ as
-    \begin{IEEEeqnarray*}{rCl}
-    V_M & = & V_\txt{in} \cup \left( \bigcup_{i \in I_A} V_{A_i} \right) \cup \left( \bigcup_{i \in I_E} V_{E_i} \right), \\
-    E_M & = & E_{M,\txt{con}} \cup \left( \bigcup_{i \in I_A} E_{A_i} \right)  \cup \left( \bigcup_{i \in I_E} E_{E_i} \right).
-    \end{IEEEeqnarray*}
+    We may then construct the maximal connectivity graph $M=(V_M,E_M)$. $V_M$ is 
+    the union of the sets of variable nodes from the inputs, analysis blocks, and expression blocks. 
+    \begin{equation}
+    V_M = V_\txt{in} \cup \left( \bigcup_{i \in I_A} V_{A_i} \right) \cup \left( \bigcup_{i \in I_E} V_{E_i} \right)
+    \end{equation}
+    $E_M$ is the union of all connection and fixed edges from the all the analysis blocks and expression blocks. 
+    \begin{equation}
+    E_M = E_{M,\txt{con}} \cup \left( \bigcup_{i \in I_A} E_{A_i} \right)  \cup \left( \bigcup_{i \in I_E} E_{E_i} \right)
+    \end{equation}
     The MCG $M$ is uniquely determined by the given set of analysis blocks, inputs, and expression blocks. 

### 2013-09-11 reviewer_response.tex
Update with BG and DP edits to paper and reviewer response

--- 2013-09-04 reviewer_response.tex
@@
 
-\title{reviewer responses for: \\ A Graph Theoretic Approach to Problem Formulation for 
-Multidisciplinary Design Analysis and Optimization}
+\title{A Graph Theoretic Approach to Problem Formulation for 
+Multidisciplinary Design Analysis and Optimization \\  (Reviewer comments appear below in red text and the authors' responses appear in black text)}
 
@@
 
-We agree that this was too quickly addressed in the paper. While we feel that 
-discussion of the PSG in any significant depth is too big of a topic to fit into 
-this paper, it does warrant a more complete description than was given in the 
+We agree that the PSG was too quickly addressed in the paper. Although we feel that 
+discussion of the PSG in significant depth would be too lengthy for the present paper, we agree that the topic does warrant a more complete discussion than that provided in the 
 original draft.
 
-We have added sections 3.5, 4.4 in the appropriate 
-parts of the paper to address the the driver nodes and driven edges which 
-together distinguish a PSG from and FPG.  We have also added added section 5.5
-which gives an example of a simple PSG and compares a PSG with its close relative
-and XDSM diagram.
-
-It is our conclusion that the PSG represents a sub-set of the information in the XDSM
-diagram, but that for data connectivity part of the graphs XDSM and PSG represent 
-equivalent graphs. It would be interesting future work to consider incorporating a new
-process edge type into the PSG to make it functionally equivalent to a complete XDSM, 
-however that would be a significant amount of extra work and worth of it's own paper. 
-To do that topic justice, the original XDSM work would need to be converted for every 
-case considered into the proposed syntax to ensure that they were completely equivalent.
+We have added subsections 3.5 and 4.4  to address the the driver nodes and driven edges which 
+together distinguish a PSG from an FPG.  We have also added subsection 5.5,
+which provides an example of a simple PSG and compares and contrasts it with an XDSM diagram.
 
@@
 
-Searching the paper, the concepts is discussed often, but only formally defined once 
-in section 4.1. In 5.3 a mathematical description of the set of nodes that meet
-the criteria for holes and edges is given, but this is different than the definition.
+In reviewing the paper, the authors agree that the concepts of holes and collisions are discussed perhaps too often.  The terms are formally defined only once 
+in section 4.1. Section 5.3 provides a mathematical description of the set of nodes that meet
+the criteria for holes and edges  in the context of the FPG algorithm, but it is our opinion that this discussion should stand separately from the definitions.
 
-We have checked both sections and removed the gross reducency, but feel that keeping 
-both equations makes for a clearer message in the paper. The concept is a bit subtle 
-since a node can be either a design variable or a hole, but not both. Similarlly, 
-a collision can occur when a variable has two incomming edges or not if the problem 
-is multi-fidelity. So a little extra explanation helps to keep this clear.
+We have checked sections 4.1 and 5.3 and removed many instances of redundancy, including informal definitions repeated several times in the text.
+Nonetheless, we feel that preserving the separate mathematical discussions in sections 4.1 and 5.3 makes for a clearer message in the paper.
+The concept is somewhat subtle
+since, in our definitions, a node can be either a design variable or a hole, but not both. Similarly, 
+a collision can occur when a variable has two incomming edges if the problem is single fidelity but not if the problem 
+is multi-fidelity. The additional discussion is intended to clarify theses subtleties.
 
@@
 
-We agree that XDSM is more visually informative and consise. However this was not the 
-goal of the new syntax. Instead, the goal was to make a graph structure that 
-was easier to operate on algorithmically. As a consequence the graph grows larger and 
-less human readable. This trade is acceptable for two reasons:
+We agree that XDSM is more visually informative and concise. However, an easily human-readable format was not the 
+goal of the new syntax. Instead, the goal is to make a graph structure that 
+was easier to operate on algorithmically to achieve feasible MDO data flows. As a consequence of this goal, the graph grows larger and becomes
+less human readable. We believe that this tradeoff is acceptable for two reasons:
 
 \begin{enumerate}
-    \item It should be possible to convert between the XDSM format and the proposed syntax 
-    in a lossless manner. This is true for the gray lines of XDSM but not the black ones.
-    Only data connections are shown in current syntax. 
-    \item Algorithmic methods are a useful way of analyzing problem structure and it makes 
-    sense to have a graph structure that works well for this use, even at the expense 
+    \item It should be possible to convert between the gray lines of the XDSM format and the proposed syntax 
+    in a lossless manner. 
+    It is our conclusion that the PSG in our syntax represents a subset of the information in the XDSM diagram; namely, the data connectivity elements of XDSM and PSG are equivalent graphs.
+The implication is that it would be possible to write an algorithm and associated code to "abstract" the FPG into a partial XDSM diagram that could then be elaborated by the user to add the solution-specific elements.
+It may then be possible to write a PSG from an XDSM, if the additional information about all of the individual variable connections from the FPG graph were preserved in the background.
+    (This is true for the gray lines of XDSM but not the black ones.
+    Only data connections are shown in new syntax.) 
+    \item Algorithmic methods are a useful way for analyzing MDO problem structure, so it seems appropriate to have a graph structure that works well for this use, even at the expense 
     of human readability. 
@@
 
-See the revised end of the introduction (section 1) where this is more clearly stated.
+Please see the revised end of the introduction (Section 1) in which the algorithmic purpose of the new graph syntax is now stated more clearly.
 
@@
 
+We have added  Fig. 8 and Fig. 13 and associated discussion to address this request.
+
 \rev{Is it really adding a lot of value to have graphs like MCG or FPG which 
 are very explicit in nature and can become difficult to read}
+
+As discussed above, we believe that the value of the MCG and FPG is not derived from their ability to represent the data in a human readable 
+form, but in their straightforward translation to machine readable formats. Section 5.3 introduces an 
+algorithm for testing whether MDO data connectivity graphs are FPGs and for reducing graphs to FPGs. As problems grow larger and more complex, 
+this type of algorithm is expected to become ever more valuable.
+
+Such an algorithm could be implemented into a framework or stand-alone tool to help teams to develop the
+specifics of a design problem before they begin detailed development work. 
+Section 6 shows that for even simple problems, the graph (as viewed by an algorithm) is 
+non-trivial. It also shows that there are multiple ways to solve problems. As 
+problems grow even a small amount from the size of the example becomes invaluable. In sum, it is our belief that the graph syntax derives its usefulness from 
+its effectiveness in describing MDO problem graphs in computational syntax.
 
@@
 information for that stage (MCG stage or FPG stage) ? Reducing cycles is 
-discussed but the graph may still look complex.}
-
-\rev{Looking at these graphs (for example Fig. 8), a node for example Z2 
+discussed but the graph may still look complex.
+\\ \indent
+Looking at these graphs (for example Fig. 8), a node for example Z2 
 appears several times. It may be useful to figure out holes and collision, 
@@
 
+There are a number of ways the graph could be simplified to make it more human readable.
+Variables could be combined, or they could be removed altogether such that the graph represents only component to component connections like a traditional DSM.
+
+As discussed in the response above, it should be possible to convert the entire PSG graph into the data-connectivity part of an XDSM.
+This would present all of the information in a more compact and visually insightful form, e.g.
+an XDSM.
+By preserving the variable connectivity information, one could always convert back into the more verbose PSG syntax  if needed.
+We would argue that the XDSM is more human readable, but the new syntax is a representation of how a framework tool, e.g.
+OpenMDAO or ModelCenter, could view the problem.
+
+Given our goals, the verbose nature of the graphs is a necessary side effect.
+
 \rev{As the problem size increases, how will it affect the size and complexity 
 of such graphs.}
+
+The answer to this question is not trivial, as it depends on a number of factors such as the number of analyses and the number of variables in those analyses.
+However, roughly speaking, one could measure the complexity of a graph by the total number of nodes and edges.
+The number of nodes grows roughly linearly with the number of analysis tools and the number of variables.
+The number of fixed edges grows linearly with the number of variables as well.
+The number of connection edges depends entirely on the specifics of the problem, but, since any variable will have only a small finite number of incoming edges (usually 1, but maybe 2 or 3 in multi-fidelity case) one can avoid an unbounded growth.
+Therefore, very roughly speaking, the graph will grow as O(N+M) where N is the number of variables and M is the number of analyses.
+
+If one assumes that large problems have hundreds of components and millions of variables, then graphs with a few million nodes and edges are possible.
+These numbers are very large, but are not outside of the realm of what is readily manageable and rapidly evaluated by today's graph libraries in common programming languages.
 
@@
 
+See the new section 2.4 for a short section describing our envisioned usage 
+of the graph language. 
+We have also mentioned in Section 5.3 that the algorithm can be directly implemented in any programming language capable of managing lists, such as Python.
+
 \rev{Minor comments:
@@
 
-\rev{Page 13 line 33 column 2: Typo : “Step (D) now PROCEDEDS … “ Page 16 line 38 
+Agreed. We have now mentioned this tools alongside OpenMDAO in  Section 1.
+
+\rev{Page 13 line 33 column 2: Typo : “Step (D) now PROCEDEDS … “}
+
+Fixed.
+
+\rev{Page 16 line 38 
 column 2: Typo : “… formulations involving MULTIPLES analysis ….”}
+
+((let's discuss))
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \rev{Page 1, Line 46: 2nd column...check spacing after ``ModelCenter''.}
+
+Fixed
 
@@
 
+Section title changed to "Background", as suggested.
+
 \rev{Page 3, Line 19, 1st column: change ``weather'' to ``whether''}
+
+Fixed
 
@@
 
+Fixed
+
 \rev{Page 4, Line 37, 1st column: what does ``REMS'' stand for?}
 
+``Reconfigurability in MDO Problem Synthesis''. The full name has been added to the paper.
+
 \rev{Page 4, Line 45, 1st column: change ``REMS address'' to ``REMS addresses''}
+
+Fixed
 
@@
 
+The authors agree that this paragraph was poorly worded. We have re-worded it to attempt 
+to make the issue with REMS more clear.
+The text now states (last two sentences, first column of p.
+4): ``...
+REMS does not provide a mechanism for inclusion of solvers or optimizers in the graph.
+This fundamentally limits REMS from describing the specifics of different solution strategies as applied to a particular problem.''  We have removed the oddly juxtaposed statement about the goal of our graph syntax that was noted by the reviewer from the paragraph.
+However, based on the comments by Reviewer 1 noted above, we have now added an additional description of how our method can capture optimizers and solvers in the context of the problem solution graph (PSG).
+
 \rev{Line 10, 1st column, Page 10:  This equation (missing a number) has a LOT of 
@@
 
+The authors agree.  We have added equation numbers (now Eqns. 8 and 9). We have also provided verbal descriptions to accompany the set notations.
+
 \rev{Lines 5 - 12, 2nd column, Page 12:  Just a general question spurred by the 
@@
 
+After considerable deliberation, we chose not to address the issue of how one should select one FPG in preference to another  in the context of the present paper. The answer 
+to this question depends heavily on the specific situation. If one is limited by run-time of the MDO workflow, 
+a smaller FPG with fewer cycles and variables may be preferable.
+If one is trying to resolve a detailed physical question, e.g.
+subtle design changes associated with nacelle-wing interference, then it may be preferable  to select a
+more high fidelity and computationally expensive FPG. These notions are introduced in the example problem, but further development was beyond the scope of the paper.
+
+The MCG provides a good foundation for finding the only FPG or indicate multiple possible FPGs.
+The user can then select the best FPG for the present stage of the design process, but this decision can  be revisited later in the process as the design goals change.
+
 \rev{Line 57, 1st column, Page 13: change ``as a input'' to ``as an input''}
+
+Fixed
 
@@
 method to obtain an FPG that significantly improves my overall computational 
-accuracy while keeping my computional efficiency as high as possible."}
+accuracy while keeping my computational efficiency as high as possible."}
+
+We agree that this sort of quantitative comparison would be very valuable. However, we 
+believe that, in order not to be misleading, this comparison needs to be larger in scope than just that which could be obtained 
+by executing the example problem given currently presented in the paper. We believe that one would need to address a wide range of problems with different graph structural 
+characteristics and different inherent nonlinear features in the analyses in order to make statements that are more generalizable. The challenge is that quantitative measures are inherently 
+dependent on particular details of a problem. 
+For example, for coupled analyses, total CPU time is highly dependent on the number of iterations required to converge the system of nonlinear functions to a consistent solution within a specified tolerance.
+This convergence is tied very closely to the specifics of the coupling and the nonlinearity.
+We had previously considered the idea of selecting ``optimal FPGs'' in terms of aggregate metrics that could be specified \textit{a priori} as meta-data for a particular analysis, but defining these metrics without considering problem coupling and nonlinearity proved especially challenging.
+To address this issue well would therefore necessitate the evaluation of many example problems with different characteristics.  This work, while very valuable, would 
+likely require an entire paper in its own right. We thank the reviewer for this suggestion, and we consider it to be a great topic for future work in this area.
 
@@
 in half and still accomplish its intended purpose.}
+
+We agree that the conclusion section
+could be more concise. We have shortened it by two paragraphs  and improved the wording.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 tools. 
+We believe that the distinction between phases 1 and 2 in the reviewer's  comments are covered adequately 
+in the the introduction and especially in the notional problem described in the indented text on p. 2.
+However, to clarify the distinction between phases 2 and 3 indicated by the reviewer, we modified the text following the indented text at the bottom of p. 2 to read as follows:
 
-We believe that the distrinction between phases 1 and 2 from the comments are covered adequately 
-in the bulk of the introduction and especially in the notional problem described there in.
-
-However, to clarify the disctinction between phases 2 and 3 we hav added the following near 
-line 52, page 3 of the original submission:
-
-\emph{%
-Our goal is to develop formalism for expressing analysis interconnectivity and for determining feasible
-    analysis tool sets to assist an engineering team conducting this task. Because the problem deals with
-    interconnectivity, we base our approach on the representations and techniques of graph theory.
-    The approach begins by constructing the \emph{maximal connectivity graph (MCG)} describing all possible
-    interconnections between the analysis tools proposed by the engineers. Graph operations are then
-    conducted to reduce the MCG to a \emph{fundamental problem graph (FPG)} that describes the set of analysis
-    tools needed to solve the specified system-level design problem. The concept of the FPG and 
-    the identification of feasible FPGs from an MCG are the main contributions of the paper. \\\\
-    This information in an FPG represents the engineering design problem that needs to be solved, but it is not 
-    itself sufficient to actually run a design optimization. It does not provide infomration 
-    about how to solve the problem. Even after the problem is defined an integration platform or framework needs 
-    to be selected and appropriate optimization methods identified. This last step essentially 
-    involves taking the FPG and turning it into a usable model. We breifly consider methods to further 
-    manipulate the FPG into a \emph{problem solution graph (PSG)} which could be usefull in this 
-    task, but that is not the focus of this paper. The work is primarily concerned with applying graph 
-    theory to the creation of an FPG and the benefits to the design process by doing so.}
+\emph{ ((copy and paste the last two paragraphs of Section 2.3))}
 

### 2013-09-11 conclusions.tex
Update with BG and DP edits to paper and reviewer response

@@
 In order to achieve a valid problem formulation, all of the holes and collisions must 
-be resolved. We present an algorithm to perform this resolution which gurantees that you 
-will find an FPG if one exists. In some cases this algorithm still relies on some input from the 
+be resolved. We present an algorithm to perform this resolution which guarantees that an FPG will be found if one exists. In some cases, this algorithm relies on additional input from the 
 designers to make decisions. One benefit of a graph theoretic approach is the standard 
-algorithms which can be used to inform the user's decisions, such as cycle detection, 
+algorithms that can be used to inform the user's decisions, such as cycle detection, 
 minimum spanning trees, and shortest path algorithms.
@@
 paper is intended to offer value in this context of increasing problem complexity by 
-providing a formal approach to programatically identify and resolve missing and 
+providing a formal approach to systematically identify and resolve missing and 
 redundant data in order to create valid MDAO problem formulations.

### 2013-09-11 example_problem.tex
addresses second reviewers comments
Update with BG and DP edits to paper and reviewer response

@@
 
-	\item A variable node is created to represent the geometry variable as a input. Any inputs incorrectly omitted will be identified in step (D).
+	\item A variable node is created to represent the geometry variable as an input. Any inputs incorrectly omitted will be identified in step (D).
 
@@
 These decisions are left to the discretion of the designer in the context of the specific implementation.
-	However, the graph-based approach presented in this paper enables standard graph algorithms to be employed to automate these decisions based on considerations of metrics related to the graph or other data related to the analysis.
+	However, the graph-based approach presented in this paper enables standard graph algorithms to be employed to automate these decisions based on considerations of metrics related to the graph or other data related to the analyses.
 	An example is a metric to obtain an FPG with the fewest possible number of cycles. 
-In this example, the FPG with the fewest cycles is shown in Fig.~\ref{f:FPG fewest cycles}, in which the edges belonging to a cycle are inticated by dashed lines. 
+In this example, the FPG with the fewest cycles is shown in Fig.~\ref{f:FPG fewest cycles}, in which the edges belonging to a cycle are indicated by dashed lines. 
 For this example problem, the cycles were detected using the implementation of Johnson's algorithm \cite{Johnson1975} in the Python package NetworkX. 

### 2013-09-11 graph_syntax.tex
addresses second reviewers comments
Update with BG and DP edits to paper and reviewer response

--- 2013-09-04 graph_syntax.tex
@@
   indicated by the first entry and directed to the node indicated by the second 
-  entry. An edge $e$ = $(x,y)$ may simply be referred to as $xy$. The edges 
+  entry. The edges 
   directed out from node $v$ are denoted by $E(v)$ and the edges directed into $v$ are denoted 
@@
   \end{figure}
-  As an example, for the directed graph shown in Fig.~\ref{f:example directed graph} we have
+  As an example, for the directed graph shown in Fig.~\ref{f:example directed graph}, we have
   \begin{IEEEeqnarray*}{rCl}
@@
 
-  Lastly, the cardinality of a set $B$ is the number of elements in $B$ and is denoted as $|B|$.
+  Lastly, the cardinality of a set $B$ is the number of elements in $B$ and is denoted as $|B|$, and the set difference is denoted with a backslash, i.e. $A \setminus B = \{x \in A \st x \notin B\}$.
 
@@
   \item[\bf{function edge:}] represents exchange of information internal to analysis tools,
-  \item[\bf{driven edge:}] represents passing of information from a driver node to a 
+  \item[\bf{driven edge:}] represents passing of information between a driver node and a 
   variable node (present only in a PSG).
@@
 \label{ss:analysis blocks}
-  Analysis tools take in a set of input variables and calculate 
+  Analysis tools accept a set of input variables and calculate 
   the values for their respective outputs. We represent this process
@@
   the edges within the analysis block are function edges, and are considered 
-  fixed to the analysis block. This graph structure demonstrated by 
+  fixed to the analysis block. The graph structure demonstrated by 
   Fig.~\ref{f:analysis block} satisfies the rules listed in Sec.~\ref{ss:types}. 
@@
   \subsection{Driver Nodes and Driven Edges}
-  Within this syntax, all iterative processes are represented with Driver nodes. 
+  Within this syntax, all iterative processes are represented with driver nodes. 
   The driver node shares some basic qualities with the model node. It can have incoming 
   and outgoing function edges to and from variable nodes. For example, many optimizers take 
-  convergence tolerance as input and output iteration counts and final objective value. 
-  With respect to variable nodes and function edges, drivers behave exactly the same as
+  convergence tolerance as input, and they output iteration counts and the final objective value. 
+  With respect to variable nodes and function edges, drivers behave similarly to
   model nodes and are subject to the same rules governing their use in the graph. In this
-  context, driver nodes can be part of driver blocks which behave just like analysis 
+  context, driver nodes can be part of driver blocks, which behave identically to analysis 
   blocks.
@@
   Driven edges are the distinguishing characteristic associated with driver nodes and
-  driver blocks. Driven edges don't follow the same rules as other edges. Driven edges  
-  will always connect a driver node to a variable node (in either direction), but they 
+  driver blocks. Driven edges do not follow the same rules as other edges: they  
+   always connect a driver node to a variable node (in either direction), but they 
   are not considered when counting the indegree or outdegree of the variable node.
-  Hence a single variable node could have multiple incoming driven edges from different 
-  driver blocks. Figure \ref{f:driver block} shows a notional example where a driver block 
-  represents and optimizer solving an unconstrained minimization problem.
+  Hence, a single variable node could have multiple incoming driven edges from different 
+  driver blocks. This accommodates a sequential optimization strategy in which both drivers, at 
+  different times, select values for the same variable. Figure \ref{f:driver block} shows 
+  a notional example in which a driver block represents an optimizer solving an 
+  unconstrained minimization problem.
 
@@
     \vspace{-10pt}
-  \caption{Example driver block with node types labeled.}
+  \caption{Example driver block implementation. The driver, analysis, and expression blocks are dash-dotted, dotted, and dashed, respectively. The driver nodes are represented by rotated squares.}
   \label{f:driver block}
@@
   Design variables are input variables whose values are free to be 
-  changed by the designer or by an optimizer in an MDAO study. In the graph-based 
+  changed by the designer or an optimizer in an MDAO study. In the graph-based 
   syntax, a node corresponds to a design variable if the indegree limits of the node are 
@@
 	In multifidelity problems (discussed subsequently), the upper indegree limit could be 
-  specified as $deg_u^- > 2$, which implies that a collision is not noted in the case of 
+  specified as $\txt{deg}_u^- > 2$, which implies that a collision is not noted in the case of 
   multiple inputs.
@@
   In MDAO, coupling is the mutual dependence between two or more analysis 
-  tools and their respective outputs. In this graph-based syntax, coupling is 
+  tools and their respective outputs. In our graph-based syntax, coupling is 
   described by the presence of a cycle between two or more analysis blocks. In the Sellar 
-  problem from Eq.~\ref{eqn:simple_fpf} 
+  problem from Eq.~\ref{eqn:simple_fpf}, 
   a mutual dependence between $D_1$ and $D_2$ exists through the variables $y_1$ and $y_2$. 
-  The edges belonging to a cycle in the Sellar problem are highlighted in 
+  The edges belonging to this cycle in the Sellar problem are highlighted in 
   Fig. \ref{f:sellar cycles}.  These edges collectively form a closed path between 
@@
 
-  Coupling cycles do not contain driver nodes in the MCG or FPG representation. 
-  Solvers, optimizers, and other iterative processes are not invovled in the coupling 
+  Coupling cycles do not contain driver nodes or driven edges since these 
+  are not allowed in the MCG or FPG representations. 
+  Solvers, optimizers, and other iterative processes are not involved in the coupling 
   definition; however, one or more solvers will be required to build a valid PSG from a 
-  given FPG  that includes cycles. Additionally, a coupling cycle has no inherent 
+  given FPG  that includes cycles.
+
+  Additionally, a coupling cycle has no inherent 
   start or end point. It would be acceptable to select a
   node in the cycle as a starting point and proceed around the
-  loop until the starting point is reached again. For the Sellar Problem, selecting 
-  $D_1$ as the starting point would yield a problem as given in 
+  loop until the starting point is reached again. For the Sellar problem, selecting 
+  $D_1$ as the starting point would yield the form of the stated in 
   Eq.~\ref{eqn:simple}.
-  Larger problems can involve more complex cycles in their FPG, indicating more 
-  complex coupling between analyses. For example, a cycle can involve more than 
+
+  Larger problems can involve more complex cycles in their corresponding FPG, indicating more 
+  complex coupling between analyses. For example, a cycle could incorporate more than 
    two analysis blocks. Multiple independent cycles could also exist, indicating 
-  independent coupling relationships. Cycles can also overlap, meaning that the same analysis 
-  blocks are involved in multiple different coupling cycles. All of these situations
-  arise naturally as the size of problems grows, and managing this coupling may
-  become difficult. In the present work, Sec.~\ref{s:example problem}
+  independent coupling relationships. Cycles could also overlap, meaning that the same analysis 
+  blocks belong to multiple different coupling cycles. All of these situations
+  arise naturally as problems grow in size, and managing these increasingly complex  couplings is particularly challenging. 
+  In the present work, Sec.~\ref{s:example problem}
   demonstrates how building an FPG from an MCG provides an opportunity to 
@@
   \label{ss:multi-fideliy problems}
-  A multi-fidelity MDAO problem can be characterized by a formulation in which 
+  A multi-fidelity MDAO problem can be characterized as a formulation in which 
   two or more different analyses each calculate the same data. Multi-fidelity 
@@
   Multi-fidelity problems are always characterized by graphs in which at least some of the variable 
-  nodes that have an $\txt{deg}^-_u(v) > 1$. These problems require 
+  nodes have an $\txt{deg}^-_u(v) > 1$. These problems require 
   special techniques for resolving the conflicting edges by introducing some mechanism
@@
 node as a hole, design variable, single valid input (nominal), collision, or 
-multi-fidelity node.
+multi-fidelity node in relation to the indegree limits.
 
@@
     0     & 0     & 0     & yes   & design variable \\
-    0     & 0     & 1     & yes   & collision \\
+    0     & 0     & 1     & no   & collision \\
     0     & 1     & 0     & yes   & design variable \\
@@
 \subsection{Iterative Control: Drivers}
-  There are many iterative processes that are used to help solve MDAO problems such as 
-  solvers, optimizers, and Design of Experiment (DOE). Abstractly, even a human picking 
-  values guided by experience and trend could be viewed as an iterative process. 
-  These processes represent the control structures that govern how a design problem 
-  marches toward a solution but they do not fundamentally impact the definition of the 
-  problem to be solved. e.g. which optimizer you choose does not change the problem you 
-  are solving but will change the path you follow to solve it. Hence, as a rule, driver 
-  nodes and driven edges are excluded from both the MCG and FPG which deal exclusively 
-  with problem formulation. They only come into play with in the PSG where the issue of 
-  how to solve a specific problem is addressed. In fact, at least one driver block is 
-  required to be present in any PSG in order to select values for the design variables.
+  There are many iterative processes  such as 
+  solvers, optimizers, and designs of experiments that are used to solve MDAO problems. 
+  These processes represent the control structures that govern the path toward a solution, but they do not fundamentally impact the definition of the 
+  problem to be solved. Hence, driver 
+  nodes and driven edges are excluded from both the MCG and FPG, which deal exclusively 
+  with problem formulation. These nodes and edges are used only in the PSG, in which the issue of 
+  how to solve a specific problem is addressed. Indeed, at least one driver block is 
+  required to be present in any PSG in order to select values for the design variables during the optimization process.

### 2013-09-11 intro.tex
Update with BG and DP edits to paper and reviewer response

--- 2013-09-04 intro.tex
@@
 
-    The time and expense for setting up 
-    analysis models is growing in conjunction with the increasing scope of MDO problems. 
     Multidisciplinary Design Analysis and Optimization (MDAO)
-    frameworks such as OpenMDAO \cite{Gray2012} and {ModelCenter}\textregistered have enabled a new level of analysis tool integration 
+    frameworks such as iSight\textsuperscript{\textregistered},\\ ModelCenter\textsuperscript{\textregistered}, ModeFrontier\textsuperscript{\textregistered}, and OpenMDAO\cite{Gray2012} have enabled a new level of analysis tool integration
     and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. This 
-    new capability has created a new challenge. Even models with tens of analyses could include hundreds or thousands
+    new capability for tool integration has created a new challenge. Even models with tens of analyses could include hundreds or thousands
     of variables that are interdependent and must be linked in the framework.
-    A common situation is that different analyses provide differing values for the 
+
+    A common situation is that different analyses provide differing predictions for the 
     same physical quantity, often in different data formats, and these conflicts need to be resolved. 
@@
     interconnections between the analysis tools proposed by the engineers. Graph operations are then
-    conducted to reduce the MCG to a \emph{fundamental problem graph (FPG)} that describes the set of analysis
-    tools needed to solve the specified system-level design problem. The concept of the FPG and the identification of feasible FPGs from an MCG are the main contributions of the paper.
+    conducted to reduce the MCG to a \emph{fundamental problem graph (FPG)} that describes a set of analysis
+    tools that collectively specify a system-level design problem. The concept of the FPG and the identification of feasible FPGs from an MCG are the main contributions of the paper.
 
-    This information in an FPG represents the engineering design problem that needs to be solved, but it is not 
-    itself sufficient to actually run a design optimization. It does not provide information 
-    about how to solve the problem. Even after the problem is defined an integration platform or framework needs 
+    The information in an FPG represents the engineering design problem that needs to be solved, but the FPG does not 
+    itself provide sufficient information to run a design optimization. In particular, it does not provide information 
+    about \textit{how} to solve the problem. Even after the problem is defined, an integration platform or framework needs 
     to be selected and appropriate optimization methods identified. This last step essentially 
-    involves taking the FPG and turning it into a usable model. We briefly consider methods to further 
+    involves converting the FPG into a usable model. We briefly consider methods to further 
     manipulate the FPG into a \emph{problem solution graph (PSG)} which could be useful in this 
-    task, but that is not the focus of this paper. The work is primarily concerned with applying graph 
-    theory to the creation of an FPG and the benefits to the design process by doing so.
+    task, but the PSG is not the primary focus of this paper. Rather, the work is primarily concerned with applying graph 
+    theory to the creation of an FPG and the benefits to the design process obtained by doing so.
 
@@
 The lack of solution information in the FPF contrasts with optimization problem statements that are written with reference to a specific solution strategy or with implication of a specific execution sequence.
-For example, consider the Sellar test problem \cite{AIAA:sellar}, with an FPF given as follows:
-
+For example, consider the Sellar test problem \cite{AIAA:sellar} with an FPF given as follows:
     \begin{align}
@@
     \end{align}
-
     In Eq.~\ref{eqn:simple_fpf}, $y_1$ is an output of $D_1$ as well as an input 
     to $D_2$. Similarly, $y_2$ is an output of $D_2$ as well as an input 
-    to $D_1$. This implies that $D_1$ and $D_2$ are coupled together by means 
-    of their mutual dependence. However no information is given as to weather
-    $D_1$ or $D_2$ should be run before the other. Given Eq.~\ref{eqn:simple_fpf}, 
-    it would be equally valid to run $D_1$ first, $D_2$ first, or both simultaneously. 
-Any MDAO root finding method or compatibility constraint formulation could be implemented to achieve consistency between $D_1$ and $D_2$.
+    to $D_1$. This structure implies that $D_1$ and $D_2$ are coupled together by means 
+    of their mutual dependence. However, no information is given as to whether
+    $D_1$ or $D_2$ should be executed before the other. Given Eq.~\ref{eqn:simple_fpf}, 
+    it would be equally valid to execute $D_1$ first, $D_2$ first, or both simultaneously. 
+An MDAO root finding method or compatibility constraint formulation could be implemented to achieve consistency between $D_1$ and $D_2$.
 
@@
     generated  based on a preferred solution approach.
-
     \begin{align}
@@
     \end{align}
-
 \subsection{Background in Graph-Based Descriptions of MDAO Problems}
@@
     mathematical syntax into a more useful computational form. 
-    Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools.
-    Off-diagonal elements of the matrix indicate coupling \cite{Steward1981}. Since a DSM describes a square adjacency matrix, 
-    it can be represented in an equivalent directed graph in which nodes represent analysis tools and 
+    Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools \cite{Steward1981}.
+    Off-diagonal elements of the matrix indicate coupling. Since a DSM describes a square adjacency matrix, 
+    it can be represented as an equivalent directed graph in which nodes represent analysis tools and 
     edges represent information dependence between those tools. The ordering of elements in a DSM can be used to indicate 
@@
     constraint information is missing from the DSM description of the problem. 
-    An alternative matrix-based syntax, called a Functional Dependency Table (FDT), was proposed by Wagner and Papalambros. 
+    An alternative matrix-based syntax, called a Functional Dependency Table (FDT), was proposed by Wagner and Papalambros \cite{Wagner1993}. 
     The FDT represents the relationship between functions, including objectives and constraints, and specific variables that affect 
-    them \cite{Wagner1993}. Similar to DSM, FDT also describes an adjacency matrix of a graph. Unlike the DSM graph, 
+    them. Similar to DSM, FDT also describes an adjacency matrix of a graph. Unlike the DSM graph, 
     however, the FDT graph is undirected and nodes can represent analysis tools, objectives, 
@@
 
-    While FDT succeeds at capturing the 
+    Although FDT succeeds at capturing the 
     information about objectives and constraints, its lack of directed edges 
     implies that it cannot describe the coupled data dependency that a DSM captures. 
-    Figure \ref{fig:FDT_simple}  shows the FDT for the Sellar problem described by Eq.~\ref{eqn:simple_fpf}.
-    The FDT shows that $D_1$ is dependent on $y_2$ and 
+    Table \ref{t:FDT_simple}  shows the FDT for the Sellar problem described by Eq.~\ref{eqn:simple_fpf}.
+    The FDT indicates that $D_1$ is dependent on $y_2$ and 
     that $D_2$ is dependent on $y_1$, but the coupled dependence cannot be inferred from 
@@
 
-    \begin{figure}[htb!]
-        \begin{center}
+\begin{table}[htb!]
+  \centering
+        \caption{Sellar Problem FDT}
         \begin{tabular}{|c|c|c|c|c|c|c|}
@@
         \end{tabular}
-        \caption{Functional Dependency Table (FDT) for Eq. \ref{eqn:simple_fpf} \label{fig:FDT_simple}}
-        \end{center}
-    \end{figure}
+ \label{t:FDT_simple}
+\end{table}%
 
-    Alexandrov and Lewis introduced a graph based syntax called REMS which 
-    incorporates objectives and constraints into a graph, effectively combining 
-    FDT and DSM \cite{alexandrov2004}. REMS retains the square adjacency 
-    matrix form from DSM, but by adding the objectives and constraints, it partially 
+    Alexandrov and Lewis introduced a graph based syntax called Reconfigurability in 
+    MDO Problem Synthesis (REMS) which incorporates objectives and constraints 
+    into a graph, effectively combining FDT and DSM \cite{alexandrov2004}. REMS retains the square adjacency 
+    matrix from DSM, but by adding the objectives and constraints, it partially 
     combines a traditional DSM with an FDT. This formulation allows REMS to represent data 
     dependency between multiple analysis tools as well as between analysis tools and
-    objective/constraint functions. Additionally, REMS address the need to
+    objective/constraint functions. Additionally, REMS addresses the need to
     transition between multiple solution strategies while maintaining a single consistent  
-    graph representation of the fundamental problem formulation. REMS provides
-    syntax for variables and functions but does not facilitate inclusion 
-    of solvers or optimizers in the graph.  The graph syntax proposed in this paper could  
-    be viewed as an extension to REMS with the goal of allowing the graphs to  describe analysis couplings and more 
-    closely interact with existing design processes and MDAO frameworks.
+    graph representation of the fundamental problem formulation. However, REMS does not provide 
+    a mechanism for inclusion of solvers or optimizers in the graph. This fundamentally limits REMS 
+    from describing the specifics of different solution strategies as applied to a specific problem.
 
@@
     XDSM also includes nodes for solvers and optimizers to enable complete 
-    definition of MDAO architectures. Martins uses XDSM to describe 13 different 
+    definition of MDAO architectures. Martins uses XDSM to describe thirteen different 
     optimization architectures in a survey paper that provides a novel 
@@
     Despite sharing some features with both of these other graph based approaches, the syntax
-    proposed here has a fundamentally different goal which is complimentary to both of them. 
-    REMS and XDSM provide excellent human readable formats for an MDAO problem description. This 
-    syntax is designed to provide a more effective machine readable format for MDAO problem definition. 
-    In the interest of making it easier to algorithmically operate on the graphs, the specified 
-    format is both more verbose and less visually informative. The benefit is a more consistent 
-    and easily utilized graph from a algorithm perspective.
+    proposed here has a fundamentally different goal which is complementary to both of them. 
+    REMS and XDSM provide excellent human readable formats for an MDAO problem description, whereas the  
+    syntax described in this paper is designed to provide an effective machine readable format for MDAO problem definition. 
+    In the interest of simplifying algorithmic operations on the graphs, the specified 
+    format is both more verbose and less visually informative. The benefit is a consistent 
+    and easily utilized graph from a algorithmic perspective.
 
@@
   formulation process. In the initial problem definition phase, the specific 
-  analysis tools and design goals are identified. Next, a single formal problem 
+  analysis tools and design goals are identified. Next, a single problem 
   formulation is identified that specifies design variables, constraints, 
@@
   the overall MDAO problem. Lastly, a specific procedure for solving the problem 
-  is selected, e.g. selecting an MDAO optimization architecture. Using 
+  is selected, e.g. an MDAO optimization architecture. Using 
   the proposed graph syntax, the outcome of these phases can be represented with the following graphs:
@@
   edges and nodes to represent the solution strategy being employed to solve the 
-  problem. This paper focuses on the MCG and FPG and does not describe the PSG in detail.
+  problem.
 
@@
     \centering
-    \subfigure[number of possible graphs]{
+    \subfigure[Number of possible graphs]{
     \includegraphics[width=1.75in]{images/tree}
@@
     }
-    \subfigure[graph size]{
+    \subfigure[Graph size]{
     \includegraphics[width=1.75in]{images/hourglass}
@@
   \end{figure}
+
+\subsection{Potential Applications}
+
+The syntax proposed here for defining graphs is partially motivated by the unique needs
+in formulating very large problems that have thousands of design variables. Such problems are 
+large enough that many traditional design methods stumble simply because one person cannot maintain a complete view of the problem. Hence, as problems 
+grow, software tools become needed to manage the complexity of the problem itself.
+
+Even for small problems, algorithmic exploration of problem structure can yield deeper understanding 
+of problems and more effective solution methods, as demonstrated by the work by Michelena and Papalambros \cite{Michelena1997} and Allison \cite{Allison2008} in applying FDT to study problem partitioning.
+The syntax and rule set described
+in this paper
+provide a foundation for the development of new algorithms for analyzing and modifying 
+problem structure. This foundation is intended to facilitate a standard view of 
+problem definition and automatic methods for investigating different  problem formulations within MDAO frameworks.

### 2013-09-11 mcg_fpg.tex
Update with BG and DP edits to paper and reviewer response

@@
 The definition of the MCG is given via construction.
+
     To construct the maximal connectivity graph, we presume that a set of analysis tools, inputs, objectives, and constraints are provided. 
@@
     We may then construct the maximal connectivity graph $M=(V_M,E_M)$. $V_M$ is 
-    the union of the sets of variable nodes from the inputs, analysis blocks, and expression blocks. 
+    the union of the sets of variable nodes from the inputs, analysis blocks, and expression blocks: 
     \begin{equation}
-    V_M = V_\txt{in} \cup \left( \bigcup_{i \in I_A} V_{A_i} \right) \cup \left( \bigcup_{i \in I_E} V_{E_i} \right)
+    V_M = V_\txt{in} \cup \left( \bigcup_{i \in I_A} V_{A_i} \right) \cup \left( \bigcup_{i \in I_E} V_{E_i} \right),
     \end{equation}
-    $E_M$ is the union of all connection and fixed edges from the all the analysis blocks and expression blocks. 
+    and
+    $E_M$ is the union of all connection edges and fixed edges from all of the analysis blocks and expression blocks. 
     \begin{equation}
-    E_M = E_{M,\txt{con}} \cup \left( \bigcup_{i \in I_A} E_{A_i} \right)  \cup \left( \bigcup_{i \in I_E} E_{E_i} \right)
+    E_M = E_{M,\txt{con}} \cup \left( \bigcup_{i \in I_A} E_{A_i} \right)  \cup \left( \bigcup_{i \in I_E} E_{E_i} \right).
     \end{equation}
     The MCG $M$ is uniquely determined by the given set of analysis blocks, inputs, and expression blocks. 
-    In the cases where the set of inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.
+    When the set of inputs is not known a priori, the process of obtaining the FPG will determine the required inputs, as discussed subsequently.
 
-The nodes and edges in $M$ can be partioned in sets according to their type:
+The nodes and edges in $M$ can be partitioned into sets according to their type:
 \begin{equation}
@@
 %    \item[(2)] $\displaystyle{\forall i \in I, \txt{ if } F \cap A_i \neq \emptyset \txt{ then } A_i \subset F}$
-    \item[(3)] $\displaystyle{\forall v \in V_F \txt{ with }v \in V_{M,\txt{var}}}$ \txt{ and } \\
+    \item[(3)] $\displaystyle{\forall v \in V_F \txt{ with }v \in V_{M,\txt{var}}}$ \\
 			 $\displaystyle{ \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
     \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F }$ \\
-			$\displaystyle{\txt{ from } x \txt{ to } v \txt{ with } x\in V_{E_i}}$
+			$\displaystyle{\txt{ from some } x \txt{ to } v \txt{ with } x\in V_{E_i}}$
 %    \item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_\txt{out}}$
@@
     %Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
-    Condition (1) asserts that only the nodes and edges provided by the maximal connectivity graph can be used in the fundamental problem graph and that every expression block must be included. 
+    Condition (1) asserts that only the nodes and edges provided by the maximal connectivity graph can be included in the fundamental problem graph and that every expression block must be present. 
     Condition (2) requires that analysis blocks must be included or excluded in their entirety. 
-    Condition (3) stipulates that the number of edges directed into each variable node must be within the lower and upper in-degree limits; 
-    if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is the location of a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is the location of a \emph{collision}.
-    Lastly, condition (4) ensures that only the nodes that are used in the MDAO problem formulation are included in the FPG by requiring that for every node a reverse path exists from at least one expression block to that node.
+    Condition (3) stipulates that the number of edges directed into each variable node must be within the lower and upper in-degree limits.
+    Lastly, condition (4) ensures that only the nodes that are used in the MDAO problem formulation are included in the FPG by requiring that, for every node, a reverse path exists from at least one expression block to that node.
     The reverse graph $R_F$ is obtained from $F$ by simply switching the orientation of every edge in $E_F$.
@@
     With this approach, the problem is reduced to deciding which connection edges to remove.
+Because this algorithm uses only basic set operations, it can be directly implemented in any programming language capable of managing lists, e.g.
+the Python programming language.
 
@@
     \end{equation}
-where the degree is now calculated with respect to $F_1$.
+where the indegree is now calculated with respect to $F_1$.
     For each collision node we can construct a set containing the edges directed into the node. The set containing all of these sets is constructed as
@@
 The construction of each $B_j$ corresponds to making a decision about which edges to include and which edges not to include.
-Let the new set of connection edges be denoted
-    \begin{equation}
-    E_{F_2,\txt{con}} = \{ e \in E_{F_1,\txt{con}} \st e \in B_j \txt{ for some } j \in J\}.
-    \end{equation}
-The set of all edges is created by removing the connection edges not in $E_{F_2,\txt{con}}$:
-    \begin{equation}
-    E_{F,2} = E_{F_0} \setminus (E_{M,\txt{con}} \setminus E_{F_2,\txt{con}}),
-    \end{equation}
-    which gives
+
+The set of all edges is created by removing the discarded edges:
+    \begin{IEEEeqnarray}{rCl}
+    \nonumber E_{F,2} = E_{F_1} \setminus & \{ & e \in  E_{M,\txt{con}} \st e \in S_{\txt{edges},j} \txt{ and } \\ & & e \notin B_j \txt{ for some } j \in J\},
+    \end{IEEEeqnarray}
+    which then gives
     \begin{equation}
@@
         \end{equation}
-	Because $b$ is directed into only the expression blocks, any path from $b$ necessarily provides a path from at least one expression block, which means the node is being used in the problem formulation.
+	Because $b$ is directed into only the expression blocks, any path from $b$ necessarily provides a path from at least one expression block, which means the node is included in the problem formulation.
 
@@
         \begin{equation}
-        V_F = V_{F_2} \setminus N.
+        V_F = V_{F_0} \setminus N.
         \end{equation}
@@
 \subsection{Problem Solution Graph}
-Once a suitable FPG has been obtained the issue of how to structure a solution strategy 
-can be addressed. Similar to the way an MCG can be manipulated into a FPG, an FPG can be
-further manipulated into a PSG. To get from an MCG to an FPG you eliminated analysis blocks and
-connection edges to reduce the problem to something solvable, but to get from an FPG to a PSG you 
-add in driver blocks, driven edges to describe the structure of the solution strategy being applied 
-to the problem. For example, Figure \ref{f:sellar_psg} shows the PSG that solves the Sellar problem 
-with an MDF solution strategy. There are two driver blocks in this PSG, one for the solver that 
-ensures compatibility and another for the optimizer that picks the design variables.
-
+Once a suitable FPG has been obtained, the issue of how to structure a solution strategy 
+can be addressed. Similar to the way an MCG can be manipulated to form an FPG, an FPG can be
+further manipulated to form a PSG. To obtain an FPG from an MCG, analysis blocks and
+connection edges are eliminated to achieve a solvable problem. To obtain a PSG from an FPG, driver blocks and driven edges are added to describe the structure of the solution strategy being applied 
+to the problem. For example, Fig. \ref{f:sellar_psg} shows the PSG that solves the Sellar problem 
+with an MDF solution strategy. The PSG contains two driver blocks, one for the solver that 
+ensures compatibility and another for the optimizer that selects the design variables. 
 \begin{figure}[htb]
@@
     \vspace{-10pt}
-    \caption{Example PSG for solving the Sellar problem with MDF strategy}
+    \caption{Example PSG for solving the Sellar problem with MDF strategy. Driver blocks are indicated by dashed boxes and driven edges are indicated by dashed lines.}
     \label{f:sellar_psg}
@@
 
-The PSG in figure \ref{f:sellar_psg} can be compared to the XDSM for the same situation 
-in Fig. \ref{fig:XDSM_simple}. While the two diagrams contain much of the same information 
-there is a significant difference. The PSG describes the data passing that happens between 
-the drivers and the rest of the problem, but it does not describe when that data passing 
-will occur (the execution order of the blocks). In the proposed graph syntax edges describe 
-the passing of data between one node an another, but they do not define when that 
-communication happens.
+The PSG in Fig. \ref{f:sellar_psg} can be compared to the XDSM for the same problem, shown 
+in Fig. \ref{fig:XDSM_simple}, although the two diagrams contain much of the same information, 
+there is one significant difference. The PSG describes the data interchange between 
+the drivers and the remaining nodes, but it does not describe when that data interchange 
+will occur (the execution order of the blocks). 
+%In the proposed graph syntax edges describe the passing of data between one node an another, but they do not define when that communication happens.
 
-XDSM diagrams do contain information about data passing as well as when that data should be passed. 
-XDSM essentially represents two related, but independent graphs simultaneously. 
-This is accomplished by including two distinct sets of edges, one for data 
-connections and a second indicating execution sequence. As seen in Fig. \ref{fig:XDSM_simple}
-data connections are bold-gray edges and execution connections are numbered-black edges. The
-execution edges, the back edges, in XDSM represent and additional graph that
-follows a completely different set of rules and behaviors from the data connections graph. 
-On the other hand the data described by the gray edges, 
-the data passing information, is equivalent to the information in the PSG. XDSM presents this 
+An XDSM is equivalent to two related but independent graphs with two distinct sets of edges, one for data 
+connections and a second indicating execution sequence. In Fig. \ref{fig:XDSM_simple}
+data connections are indicated in bold-gray edges and execution connections are shown as numbered black edges. The
+execution edges in XDSM correspond to an additional graph that
+follows different rules than the data connections graph. 
+On the other hand, the information described by the gray edges, 
+the data interchange information, is equivalent to the information in the PSG. XDSM presents this 
 information in a more compact and human readable manner by aggregating groups of related variables
-into combined nodes. The more verbose form presented in a PSG is less human readable, but more easily 
-machine readable as a consequence of a more consistent structure.
+into combined nodes. The more verbose form presented in a PSG is less human readable but more easily 
+machine readable as a consequence of the formal graph structure that can be implemented and interpreted by commonly available software for graph manipulations.

### 2013-09-11 problem_formulation.tex
Update with BG and DP edits to paper and reviewer response

@@
 The formulation of multidisciplinary design, analysis, and optimization (MDAO) problems has become increasingly complex as the number of analysis tools and design variables included in typical studies has grown.
-This growth in the scale and scope of MDAO problems has been motivated by the need to incorporate additional design disciplines and to expand the parametric design space to enable the exploration of unconventional design concepts.
-In this context, given a large set of disciplinary analysis tools, the problem of determining a feasible data flow between tools to produce a specified set of system-level outputs is combinatorically challenging.
+This growth in the scale and scope of MDAO problems has been motivated by the need to incorporate additional disciplines and to expand the parametric design space to enable the exploration of unconventional design concepts.
+In this context, given a large set of disciplinary analysis tools, the problem of determining a feasible data flow between tools to produce a specified set of system-level outputs is combinatorially challenging.
 The difficulty is compounded in multi-fidelity problems, which are of increasing interest to the MDAO community.
@@
 The FPG does not predispose a particular solution procedure; any relevant MDO solution architecture could be selected to implement the optimization.
+The solution architecture can be represented in a problem solution graph (PSG).
 The approach is applied to an example problem to formulate an FPG for a commercial aircraft MDAO study.

### 2013-09-27 reviewer_response.tex
Submitted version of revisions 1

@@
-\documentclass{aiaa-tc}
+\documentclass{article} %{aiaa-tc}
 
-\title{A Graph Theoretic Approach to Problem Formulation for 
-Multidisciplinary Design Analysis and Optimization \\  (Reviewer comments appear below in red text and the authors' responses appear in black text)}
+	\addtolength{\oddsidemargin}{-.875in}
+	\addtolength{\evensidemargin}{-.875in}
+	\addtolength{\textwidth}{1.75in}
+
+	\addtolength{\topmargin}{-.875in}
+	\addtolength{\textheight}{1.75in}
+
+\date{}
+\setcounter{secnumdepth}{0}
+
+\title{\large Authors' Responses to the Reviewers for the Manuscript,\\
+\large ``A Graph Theoretic Approach to Problem Formulation for
+Multidisciplinary Design Analysis and Optimization'' \\  \vspace{0.25in} \normalsize (Reviewer comments are set in red text and the authors' responses are set in black text)}
 
@@
 
-\newenvironment{rev}{\vspace{2em}\itshape \textcolor{red}}{}
+%\newenvironment{rev}{\vspace{2em}\itshape \textcolor{red}}{}
+\newenvironment{rev}{\vspace{2em} \textcolor{red}}{}
 
@@
 
-We agree that the PSG was too quickly addressed in the paper. Although we feel that 
-discussion of the PSG in significant depth would be too lengthy for the present paper, we agree that the topic does warrant a more complete discussion than that provided in the 
-original draft.
-
-We have added subsections 3.5 and 4.4  to address the the driver nodes and driven edges which 
+Although we feel that discussion of the PSG in significant depth would be too lengthy for the present paper, we agree that the topic does warrant a more complete discussion than that provided in the
+original draft.  We have added subsections 3.5 and 4.4 to address the the driver nodes and driven edges which
 together distinguish a PSG from an FPG.  We have also added subsection 5.5,
@@
 
-In reviewing the paper, the authors agree that the concepts of holes and collisions are discussed perhaps too often.  The terms are formally defined only once 
-in section 4.1. Section 5.3 provides a mathematical description of the set of nodes that meet
-the criteria for holes and edges  in the context of the FPG algorithm, but it is our opinion that this discussion should stand separately from the definitions.
+The authors agree that the concepts of holes and collisions were discussed too often.
+The terms are formally defined in section 4.1.
+Section 5.3 provides a mathematical description of the set of nodes that meet
+the criteria for holes and collisions in the context of the FPG algorithm, and it is our opinion that this discussion should stand separately from the definitions.
 
-We have checked sections 4.1 and 5.3 and removed many instances of redundancy, including informal definitions repeated several times in the text.
+We have checked sections 4.1 and 5.3 and have removed most instances of redundancy in other parts of the paper, including informal definitions repeated several times in the text.
 Nonetheless, we feel that preserving the separate mathematical discussions in sections 4.1 and 5.3 makes for a clearer message in the paper.
-The concept is somewhat subtle
-since, in our definitions, a node can be either a design variable or a hole, but not both. Similarly, 
-a collision can occur when a variable has two incomming edges if the problem is single fidelity but not if the problem 
-is multi-fidelity. The additional discussion is intended to clarify theses subtleties.
+The concept is somewhat subtle since, in our definitions, a node can be either a design variable or a hole, but not both.
+Similarly, a collision occurs when a variable has two incoming edges if the problem is single fidelity but does not occur in this situation if the problem is multi-fidelity.
+The additional discussion is intended to clarify theses subtleties.
 
@@
 We agree that XDSM is more visually informative and concise. However, an easily human-readable format was not the
-goal of the new syntax. Instead, the goal is to make a graph structure that 
-was easier to operate on algorithmically to achieve feasible MDO data flows. As a consequence of this goal, the graph grows larger and becomes
-less human readable. We believe that this tradeoff is acceptable for two reasons:
+goal of the new syntax. Instead, the goal is to create a graph structure that
+can be manipulated algorithmically to achieve feasible MDO data flows.
+As a consequence of this goal, the graph grows larger and becomes less human readable.
+We believe that this tradeoff is acceptable for two reasons:
 
@@
     It is our conclusion that the PSG in our syntax represents a subset of the information in the XDSM diagram; namely, the data connectivity elements of XDSM and PSG are equivalent graphs.
-The implication is that it would be possible to write an algorithm and associated code to "abstract" the FPG into a partial XDSM diagram that could then be elaborated by the user to add the solution-specific elements.
+The implication is that it would be possible to write an algorithm and associated code to ``abstract'' the FPG into a partial XDSM diagram that could then be elaborated by the user to add the solution-specific elements.
 It may then be possible to write a PSG from an XDSM, if the additional information about all of the individual variable connections from the FPG graph were preserved in the background.
-    (This is true for the gray lines of XDSM but not the black ones.
-    Only data connections are shown in new syntax.) 
+    (This is true for the gray lines of XDSM but not the black ones. Only data connections are described in our new syntax.)
     \item Algorithmic methods are a useful way for analyzing MDO problem structure, so it seems appropriate to have a graph structure that works well for this use, even at the expense
@@
 Please see the revised end of the introduction (Section 1) in which the algorithmic purpose of the new graph syntax is now stated more clearly.
+We have also added the following text in the last paragraph of Section 2.2 to clarify the issue:
+
+\begin{quotation} 
+\textit{REMS and XDSM provide excellent
+human readable formats for an MDAO problem description, whereas the syntax described in this paper is designed to provide an effective machine readable format
+for MDAO problem definition.
+In the interest of simplifying algorithmic operations on the graphs, the specified format is both more verbose and less visually informative.
+The benefit is a consistent and easily utilized
+graph from a algorithmic perspective.}
+\end{quotation}
 
@@
 specifics of a design problem before they begin detailed development work.
-Section 6 shows that for even simple problems, the graph (as viewed by an algorithm) is 
-non-trivial. It also shows that there are multiple ways to solve problems. As 
-problems grow even a small amount from the size of the example becomes invaluable. In sum, it is our belief that the graph syntax derives its usefulness from 
+Section 6 shows that for even simple problems, the graph is
+non-trivial.
+As problems grow even a small amount from the size of the example, an algorithmic approach for studying problem structure becomes invaluable.
+In sum, it is our belief that the graph syntax derives its usefulness from
 its effectiveness in describing MDO problem graphs in computational syntax.
@@
 and their interconnections while still conveying important and critical
-information for that stage (MCG stage or FPG stage) ? Reducing cycles is 
+information for that stage (MCG stage or FPG stage)? Reducing cycles is
 discussed but the graph may still look complex.
@@
 OpenMDAO or ModelCenter, could view the problem.
-
 Given our goals, the verbose nature of the graphs is a necessary side effect.
@@
 
-See the new section 2.4 for a short section describing our envisioned usage 
-of the graph language. 
-We have also mentioned in Section 5.3 that the algorithm can be directly implemented in any programming language capable of managing lists, such as Python.
+Please see the new section 2.4 for a brief discussion of our envisioned usage
+of the graph language in a software tool.
+We have also mentioned in Section 5.3 that the algorithm can be straightforwardly implemented in any programming language capable of managing lists, such as Python.
 
@@
 
-Agreed. We have now mentioned this tools alongside OpenMDAO in  Section 1.
+Agreed. We have now mentioned these tools alongside OpenMDAO in  Section 1.
 
@@
 
-((let's discuss))
+Fixed.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Fixed
+Fixed.
 
@@
 
-Section title changed to "Background", as suggested.
+Section title changed to ``Background'' as suggested.
 
@@
 
-``Reconfigurability in MDO Problem Synthesis''. The full name has been added to the paper.
+``Reconfigurability in MDO Problem Synthesis''. The full name has been added to the paper to define the acronym.
 
@@
 The authors agree that this paragraph was poorly worded. We have re-worded it to attempt
-to make the issue with REMS more clear.
+to make the comparison with REMS more clear.
 The text now states (last two sentences, first column of p.
@@
 
-The MCG provides a good foundation for finding the only FPG or indicate multiple possible FPGs.
-The user can then select the best FPG for the present stage of the design process, but this decision can  be revisited later in the process as the design goals change.
+The MCG provides a good foundation for identifying multiple possible FPGs.
+The user can then select the best FPG for the present stage of the design process based on his/her own criteria.
+This decision can be revisited later in the process as the design goals and associated criteria change.
 
@@
 believe that, in order not to be misleading, this comparison needs to be larger in scope than just that which could be obtained
-by executing the example problem given currently presented in the paper. We believe that one would need to address a wide range of problems with different graph structural 
+by executing the example problem currently presented in the paper. We believe that one would need to address a wide range of problems with different graph structural
 characteristics and different inherent nonlinear features in the analyses in order to make statements that are more generalizable. The challenge is that quantitative measures are inherently
@@
 We agree that the conclusion section
-could be more concise. We have shortened it by two paragraphs  and improved the wording.
+could be more concise. We have shortened it by two paragraphs and improved the wording.  In particular, we have removed the unnecessary recap of collisions and holes that was noted by Reviewer 1.
 
@@
 
-\emph{ ((copy and paste the last two paragraphs of Section 2.3))}
+\begin{quotation}
+    \textit{Our goal is to develop formalism for expressing analysis interconnectivity and for determining feasible
+    analysis tool sets to assist an engineering team conducting this task. Because the problem deals with
+    interconnectivity, we base our approach on the representations and techniques of graph theory.
+    The approach begins by constructing the \emph{maximal connectivity graph (MCG)} describing all possible
+    interconnections between the analysis tools proposed by the engineers. Graph operations are then
+    conducted to reduce the MCG to a \emph{fundamental problem graph (FPG)} that describes a set of analysis
+    tools that collectively specify a system-level design problem. The concept of the FPG and the identification of feasible FPGs from an MCG are the main contributions of the paper.}
+
+   \textit{ The information in an FPG represents the engineering design problem that needs to be solved, but the FPG does not
+    itself provide sufficient information to run a design optimization. In particular, it does not provide information
+    about \textit{how} to solve the problem. Even after the problem is defined, an integration platform or framework needs
+    to be selected and appropriate optimization methods identified. This last step essentially
+    involves converting the FPG into a usable model. We briefly consider methods to further
+    manipulate the FPG into a \emph{problem solution graph (PSG)} which could be useful in this
+    task, but the PSG is not the primary focus of this paper. Rather, the work is primarily concerned with applying graph
+    theory to the creation of an FPG and the benefits to the design process obtained by doing so.}
+\end{quotation}
 

### 2013-09-27 conclusions.tex
Submitted version of revisions 1

@@
 formulation, especially for problems with large numbers of variables and analyses.
+
 In particular, we introduce the concepts of the Maximally Connected Graph (MCG) and
 the Fundamental Problem Graph (FPG).
-
 The MCG addresses the question, ``Given a set of analysis tools,
@@
 
+%Techniques such as REMS and $\Psi$ focus on developing partitioned subproblems corresponding a given fundamental problem formulation.
+
+Our graph-based approach is intended primarily to infer possible fundamental problem formulations, via their corresponding FPGs.
+However, the syntax also facilitates the development of a Problem Solution Graph (PSG) that describes the MDAO partitioning and coordination strategy based on a specified FPG.
+The formulation of a PSG could take advantage of other methods such as $\Psi$, REMS, and XDSM to assist in the specification of subproblem partitions and the addition of coordination-specific elements such as drivers and compatibility constraints.
+
 For simple problems with few analysis tools and variables, the formulation of a

### 2013-09-27 intro.tex
Submitted version of revisions 1

@@
     and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. This
-    new capability for tool integration has created a new challenge. Even models with tens of analyses could include hundreds or thousands
-    of variables that are interdependent and must be linked in the framework.
+    new capability for tool integration has created a new challenge: Even models with tens of analyses could include hundreds or thousands of variables that are interdependent and must be linked in the framework.
 
@@
 
-    Lamb and Martins also included objectives and constraint functions as nodes 
+Tosserams et al.
+developed a technique for problem partitioning named $\Psi$ that, like REMS, is based on a formal language specification \cite{Tosserams2010}.
+The $\Psi$ method allows the user to specify how an MDO problem is to be decomposed into multiple optimization subproblems.
+These subproblems may involve one or several analyses and multiple design variables.
+Unlike REMS which builds subproblems automatically, $\Psi$ is intended to enable a manual process of subproblem construction based on the designers’ goals.
+After subproblems are built, the language includes constructs to link them together to form into a system-level optimization problem statement.
+$\Psi$ does not predispose the coordination strategy in the MDO architecture; any applicable approach can be used to achieve the required consistency among subproblems during the system level optimization.
+Compiler and generator functions have been developed for $\Psi$ to facilitate a straightforward specification of coordination approaches for MDO frameworks.
+
+    Lamb and Martins included objectives and constraint functions as nodes
     in an Extended DSM (XDSM) in order to capture a more complete description

### 2013-09-27 problem_formulation.tex
Submitted version of revisions 1

@@
 The FPG does not predispose a particular solution procedure; any relevant MDO solution architecture could be selected to implement the optimization.
-The solution architecture can be represented in a problem solution graph (PSG).
-The approach is applied to an example problem to formulate an FPG for a commercial aircraft MDAO study.
+Finally, the solution architecture can be represented in a problem solution graph (PSG). The graph approach is applied to an example problem based on a commercial aircraft MDAO study.
 \end{abstract}
